### 2021-05-03 latex/ro-crate.tex
Moved to latex/

% PLEASE USE THIS FILE AS A TEMPLATE
% Check file iosart2x.tex for more examples

% add. options: [seceqn,secthm,crcready]
\documentclass[ds]{iosart2x}
\usepackage[smartEllipses]{markdown}

%\usepackage{dcolumn}

%%%%%%%%%%% Put your definitions here

%%%%%%%%%%% End of definitions

\pubyear{0000}
\volume{0}
\firstpage{1}
\lastpage{1}

\begin{document}

\begin{frontmatter}

%\pretitle{}
\title{Article title}
\runtitle{Running head title}
%\subtitle{}

% For one author:
%\author{\inits{N.}\fnms{Name1} \snm{Surname1}\ead[label=e1]{first@somewhere.com}\ead[label=o1,orcid]{0000-0000-0000-0001}}
%\address{Department first, \orgname{University or Company name},
%Abbreviate US states, \cny{Country}\printead[presep={\\}]{e1,o1}}

% Two or more authors:
\author[A]{\inits{N.}\fnms{Name1} \snm{Surname1}\ead[label=e1]{first@somewhere.com}\ead[label=o1,orcid]{0000-0000-0000-0001}%
\thanks{Corresponding author. \printead{e1}.}},
\author[B]{\inits{N.N.}\fnms{Name2 Name2} \snm{Surname2}\ead[label=e2]{second@somewhere.com}\ead[label=o2,orcid]{0000-0000-0000-0002}}
and
\author[C]{\inits{N.-N.}\fnms{Name3-Name3} \snm{Surname3}\ead[label=e3]{third@somewhere.com}\ead[label=o3,orcid]{0000-0000-0000-0003}}
\address[A]{Department first, \orgname{University or Company name},
Abbreviate US states, \cny{Country}\printead[presep={\\}]{e1,o1}}
\address[B]{Department first, \orgname{University or Company name},
Abbreviate US states, \cny{Country}\printead[presep={\\}]{e2,o2}}
\address[C]{Department first, \orgname{University or Company name},
Abbreviate US states, \cny{Country}\printead[presep={\\}]{e3,o3}}

%\begin{review}{editor}
%\reviewer{\fnms{First} \snm{Editor}\ead[label=oe1,orcid]{}\printead[presep={\ }]{oe1}}
%\reviewer{\fnms{Second} \snm{Editor}\ead[label=oe2,orcid]{}\printead[presep={\ }]{oe2}}
%\end{review}
%\begin{review}{solicited}
%\reviewer{\fnms{First} \snm{Solicited reviewer}\ead[label=os1,orcid]{}\printead[presep={\ }]{os1}}
%\end{review}

\begin{abstract}
Abstract text.
\end{abstract}

\begin{keyword}
\kwd{Keyword one}
\kwd{keyword two}
\end{keyword}

\end{frontmatter}

%%%%%%%%%%% The article body starts:

%\section{}\label{s1}

%\subsection{}\label{s1.1}

\markdownInput{../content/10.background.md}
\markdownInput{../content/20.rocrate.md}
\markdownInput{../content/30.formal-definition.md}

%\begin{figure}[t]
%\includegraphics{}
%\caption{Figure caption.}\label{f1}
%\end{figure}

%\begin{table*}
%\caption{} \label{t1}
%\begin{tabular}{lll}
%\hline
%&&\\
%&&\\
%\hline
%\end{tabular}
%\end{table*}

%%%%%%%%%%% The bibliography starts:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  ios1.bst will be used to                               %%
%%  create a .BBL file for submission.                     %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\nocite{*} 
% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{ios1}           % Style BST file.
\bibliography{bibliography}        % Bibliography file (usually '*.bib')

% or include bibliography directly:
%\begin{thebibliography}{0}
%\bibitem{r1} F. Author, Information about cited object.
%
%\bibitem{r2} S. Author and T. Author, Information about cited object.
%\end{thebibliography}

\end{document}

### 2021-05-03 latex/ro-crate.tex
markdown as well in ../output
include all markdown files
authors

@@
-% PLEASE USE THIS FILE AS A TEMPLATE
-% Check file iosart2x.tex for more examples
+%% ro-crate paper
+%% (c) 2020-2021 The University of Manchester
+%% 
+%% SPDX-License-Identifier: CC-BY-4.0
 
@@
 \usepackage[smartEllipses]{markdown}
+\def\markdownOptionOutputDir{../output}
+
+%\markdownSetup{
+%    defaultOptions.cacheDir = 
+%}
 
@@
 %\pretitle{}
-\title{Article title}
-\runtitle{Running head title}
+\title{Packaging research artifacts with RO-Crate}
+\runtitle{Packaging research artifacts with RO-Crate}
 %\subtitle{}
@@
 % Two or more authors:
-\author[A]{\inits{N.}\fnms{Name1} \snm{Surname1}\ead[label=e1]{first@somewhere.com}\ead[label=o1,orcid]{0000-0000-0000-0001}%
+\author[A]{\inits{S.S-R.}\fnms{Stian} \snm{Soiland-Reyes}\ead[label=e1]{soiland-reyes@manchester.ac.uk}\ead[label=o1,orcid]{0000-0001-9842-9718}%
 \thanks{Corresponding author. \printead{e1}.}},
-\author[B]{\inits{N.N.}\fnms{Name2 Name2} \snm{Surname2}\ead[label=e2]{second@somewhere.com}\ead[label=o2,orcid]{0000-0000-0000-0002}}
+\author[B]{\inits{P.S.}\fnms{Peter} \snm{Sefton}\ead[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}}
 and
-\author[C]{\inits{N.-N.}\fnms{Name3-Name3} \snm{Surname3}\ead[label=e3]{third@somewhere.com}\ead[label=o3,orcid]{0000-0000-0000-0003}}
-\address[A]{Department first, \orgname{University or Company name},
-Abbreviate US states, \cny{Country}\printead[presep={\\}]{e1,o1}}
-\address[B]{Department first, \orgname{University or Company name},
-Abbreviate US states, \cny{Country}\printead[presep={\\}]{e2,o2}}
-\address[C]{Department first, \orgname{University or Company name},
-Abbreviate US states, \cny{Country}\printead[presep={\\}]{e3,o3}}
+\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}}
+\author[C]{\inits{P.G.}\fnms{Paul} \snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}}
+\address[A]{Computer Science, \orgname{The University of Manchester},
+\cny{UK}\printead[presep={\\}]{e1,o1}}
+\address[B]{Faculty of Science, \orgname{University Technology Sydney},
+\cny{Australia}\printead[presep={\\}]{e2,o2}}
+\address[C]{Informatics Institute, \orgname{University of Amsterdam},
+\cny{The Netherlands}\printead[presep={\\}]{e3,o3}}
 
@@
 \begin{keyword}
-\kwd{Keyword one}
-\kwd{keyword two}
+\kwd{Data publishing}
+\kwd{Data packaging}
+\kwd{FAIR}
+\kwd{Linked Data}
+\kwd{Metadata}
+\kwd{Reproducibility}
+\kwd{Research Object}
 \end{keyword}
@@
 \markdownInput{../content/20.rocrate.md}
-\markdownInput{../content/30.formal-definition.md}
+\markdownInput{../content/21.conceptual.md}
+\markdownInput{../content/22.implementation.md}
+\markdownInput{../content/23.community.md}
+\markdownInput{../content/30.tooling.md}
+\markdownInput{../content/40.inuse.md}
+\markdownInput{../content/50.relatedwork.md}
+\markdownInput{../content/60.conclusion.md}
+%\markdownInput{../content/80.formal-definition.md}
 

### 2021-05-03 latex/ro-crate.tex
author mess

@@
 \documentclass[ds]{iosart2x}
+\usepackage[utf8]{inputenc}
 \usepackage[smartEllipses]{markdown}
@@
 % Two or more authors:
-\author[A]{\inits{S.S-R.}\fnms{Stian} \snm{Soiland-Reyes}\ead[label=e1]{soiland-reyes@manchester.ac.uk}\ead[label=o1,orcid]{0000-0001-9842-9718}%
+\author[A,B]{\inits{S.S-R.}\fnms{Stian} \snm{Soiland-Reyes}\ead[label=e1]{soiland-reyes@manchester.ac.uk}\ead[label=o1,orcid]{0000-0001-9842-9718}%
 \thanks{Corresponding author. \printead{e1}.}},
-\author[B]{\inits{P.S.}\fnms{Peter} \snm{Sefton}\ead[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}}
-and
-\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}}
-\author[C]{\inits{P.G.}\fnms{Paul} \snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}}
+\author[C]{\inits{P.S.}\fnms{Peter} \snm{Sefton}\ead[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}},
+\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}},
+\author[B]{\inits{P.G.}\fnms{Paul} \snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}}
 \address[A]{Computer Science, \orgname{The University of Manchester},
 \cny{UK}\printead[presep={\\}]{e1,o1}}
-\address[B]{Faculty of Science, \orgname{University Technology Sydney},
-\cny{Australia}\printead[presep={\\}]{e2,o2}}
-\address[C]{Informatics Institute, \orgname{University of Amsterdam},
-\cny{The Netherlands}\printead[presep={\\}]{e3,o3}}
+\address[B]{Informatics Institute, \orgname{University of Amsterdam},
+\cny{The Netherlands}\printead[presep={\\}]{e2,o2}}
+\address[C]{Faculty of Science, \orgname{University Technology Sydney},
+\cny{Australia}\printead[presep={\\}]{e3,o3}}
 

### 2021-05-03 latex/ro-crate.tex
markdown settings

@@
 \usepackage[utf8]{inputenc}
-\usepackage[smartEllipses]{markdown}
+\usepackage[smartEllipses,tableCaptions,pipeTAbles,fencedCode,footnotes,underscores]{markdown}
 \def\markdownOptionOutputDir{../output}
@@
 
-\pubyear{0000}
+\pubyear{2021}
 \volume{0}
@@
 % Two or more authors:
-\author[A,B]{\inits{S.S-R.}\fnms{Stian} \snm{Soiland-Reyes}\ead[label=e1]{soiland-reyes@manchester.ac.uk}\ead[label=o1,orcid]{0000-0001-9842-9718}%
+\author[A,B]{\fnms{Stian} \snm{Soiland-Reyes}\ead[label=e1]{soiland-reyes@manchester.ac.uk}\ead[label=o1,orcid]{0000-0001-9842-9718}%
 \thanks{Corresponding author. \printead{e1}.}},

### 2021-05-03 build/build.sh
symlink hack

@@ -11,6 +11,12 @@ export TZ=Etc/UTC
 # Default Python to read/write text files using UTF-8 encoding
 export LC_ALL=en_US.UTF-8
 
+# Overleaf workaround, where symlinks are not permitted in git
+ln -fs v/latest/manuscript.pdf webpage/
+ln -fs v/latest/images webpage/
+ln -fs v/latest/index.html webpage/
+
+
 # Generate reference information
 echo >&2 "Retrieving and processing reference metadata"
 manubot process \

### 2021-05-03 latex/ro-crate.tex
More latex settings..

@@
 \documentclass[ds]{iosart2x}
-\usepackage[utf8]{inputenc}
-\usepackage[smartEllipses,tableCaptions,pipeTAbles,fencedCode,footnotes,underscores]{markdown}
+%\usepackage[utf8]{inputenc}
+
+% From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
+\usepackage{amsmath}
+\usepackage{hyperref}
+\usepackage{fontspec}
+\usepackage{unicode-math}
+\usepackage{graphicx}
+
+%%
+
+\DeclareGraphicsExtensions{.png,.pdf}
+\graphicspath{../content/images/}
+\usepackage[smartEllipses,tableCaptions,pipeTables,fencedCode,footnotes,underscores,hybrid]{markdown}
 \def\markdownOptionOutputDir{../output}
@@
 \begin{document}
+
+\setmathfont{STIX2Math}[
+Extension={.otf},
+Path=./STIX2fonts/,
+Scale=1]
+\setmainfont{STIX2Text}[
+Extension={.otf},
+Path=./STIX2fonts/,
+UprightFont={*-Regular},
+BoldFont={*-Bold},
+ItalicFont={*-Italic},
+BoldItalicFont={*-BoldItalic}]
 
@@
 \markdownInput{../content/60.conclusion.md}
-%\markdownInput{../content/80.formal-definition.md}
+\markdownInput{../content/80.formal-definition.md}
 

### 2021-05-04 content/80.formal-definition.tex
fixing more DOIs and markdown

# Appendix

## Formalizing RO-Crate in First Order Logic

Below is an attempt to formalize the concept of RO-Crate as a set of relations using First Order Logic:

### Language

    𝕃ro-crate = { Property(p), Class(c), Literal(x), ℝ, 𝕊 }
    𝔻 =  𝕀𝕣𝕚
    𝕀𝕣𝕚 ≡  { IRIs as defined in <https://tools.ietf.org/html/rfc3987> }
    ℝ ≡  { real or integer numbers }
    𝕊 ≡  { literal strings }

The domain of discourse is the set of 𝕀𝕣𝕚 identifiers (notation `<http://example.com/>`), with additional descriptions using numbers ℝ (notation `13.37`) and literal strings 𝕊 (notation `“Hello”`).

From this formalized language `𝕃ro-crate` a RO-Crate can be interpreted in any representation that can gather these descriptions, their properties, classes, and literal attributes.

### Minimal RO-Crate

Below is using `𝕃ro-crate` to define a minimal RO-Crate:

    RO-Crate(R) ⊨  Root(R) ∧ Mentions(R, R) ∧
        hasPart(R, d) ∧ Mentions(R, d) ∧ DataEntity(d) ∧
        Mentions(R, c) ∧ ContextualEntity(c)
    ∀r Root(r) →  Dataset(r) ∧ name(r, n)∧ description(r, d)
                            ∧ published(r, date) ∧ license(e, l)
    ∀e∀n name(e, n) →  Literal(n)
    ∀e∀d description(e, d) →  Literal(d)
    ∀e∀date datePublished(e, date) →  Literal(date)
    ∀e∀l license(e, l) →  ContextualEntity(l)
    DataEntity(e) ≡  File(e) ⊕ Dataset(e)
    Entity(e) ≡  DataEntity(e) ∨ ContextualEntity(e)
    ∀e Entity(e) → Class(e)
    Mentions(R, s) ⊨  Relation(s, p, e)  ⊕  Attribute(s, p, l)
    Relation(s, p, o) ⊨ Entity(s) ∧ Property(p) ∧ Entity(o)
    Attribute(s, p, x) ⊨ Entity(s)  ∧ Property(p) ∧ Literal(x)
    Literal(x) ≡  x ∈ ℝ  ⊕   x ∈ 𝕊

An `RO-Crate(R)` is defined as a self-described _Root Data Entity_, which describes and contains parts (_data entities_), which are further described in _contextual entities_.  These terms align with their use in the [RO-Crate 1.1 terminology](https://www.researchobject.org/ro-crate/1.1/terminology).

The `Root(r)` is a type of Dataset(r), and must have the metadata to literal attributes to provide a `name`, `description` and `datePublished`, as well as a contextual entity identifying its license. These predicates correspond to the RO-Crate 1.1 [requirements for the root data entity](https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#direct-properties-of-the-root-data-entity).

The concept of an `Entity(e)` is introduced as being either a DataEntity(e), a `ContextualEntity(e)`, or [both](https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities); and must be typed with at least one `Class(e)`.

For simplicity in this formalization (and to assist production rules below) `R` is a constant representing a single RO-Crate, typically written to independent RO-Crate Metadata files. `R` is used by `Mentions(R, e)` to indicate that `e` is an Entity described by the RO-Crate and therefore its metadata (a set of Relation and Attribute predicates) form part of the RO-Crate serialization. `Relation(s, p, o)` and `Attribute(s, p, x)` are defined as a _subject-predicate-object_ triple pattern from an `Entity(s)` using a `Property(p)` to either another `Entity(o)` or a `Literal(x)` value.

### Example of formalized RO-Crate

The below is an example RO-Crate represented using the above formalization, assuming a base URI of `http://example.com/ro/123/`:

    RO-Crate(<http://example.com/ro/123/>)
    name(<http://example.com/ro/123/, 
        “Data files associated with the manuscript:Effects of …”)
    description(<http://example.com/ro/123/, 
        “Palliative care planning for nursing home residents …")
    datePublished(<http://example.com/ro/123/>, “2017")
    license(<http://example.com/ro/123/>, 
        <https://creativecommons.org/licenses/by-nc-sa/3.0/au/>
    ContextualEntity(<https://creativecommons.org/licenses/by-nc-sa/3.0/au/>)
    name(<https://creativecommons.org/licenses/by-nc-sa/3.0/au/, 
        “Attribution-NonCommercial-ShareAlike 3.0 Australia (CC BY-NC-SA 3.0 AU)”)

    hasPart(<http://example.com/ro/123/>, <http://example.com/ro/123/file.txt>)
    File(<http://example.com/ro/123/survey.csv>)
    name(<http://example.com/ro/123/survey.csv>, “Survey of care providers”)
    hasPart(<http://example.com/ro/123/>, <http://www.example.om/ro/123/folder/>)
    Dataset(<http://example.com/ro/123/interviews/>)
    name(<http://example.com/ro/123/interviews/>, 
        “Audio recordings of care provider interviews”)

In reality many additional attributes from schema.org types like <http://schema.org/Dataset> and <http://schema.org/CreativeWork> would be used to further describe the RO-Crate and its entities, but as these are optional they do not form part of this formalization.

### Mapping to RDF with schema.org

A formalized RO-Crate can be mapped to different serializations. Below follows a mapping to RDF using schema.org.

    Dataset(d) →  type(d, <http://schema.org/Dataset>)
    File(f) →  type(f, <http://schema.org/MediaObject>)
    Property(p) →  type(p, <http://www.w3.org/2000/01/rdf-schema#Property>)
    Class(c) →  type(c, <http://www.w3.org/2000/01/rdf-schema#Class>)
    CreativeWork(e) →  ContextualEntity(e) ∧ type(e, <http://schema.org/CreativeWork>)
    hasPart(e, t) →  Relation(e, <http://schema.org/hasPart>, t)
    type(e, t) →  Relation(e, <http://www.w3.org/1999/02/22-rdf-syntax-ns#type>, t) ∧ Class(t)
    name(e, n) →  Attribute(e, <http://schema.org/name>, n)
    description(e, d) →  Attribute(e, <http://schema.org/description>, d)
    datePublished(e, date) →  Attribute(e, <http://schema.org/datePublished>, date)
    license(e, l) →  Relation(e, <http://schema.org/license>, l) ∧ CreativeWork(l)

Note that in the JSON-LD serialization of RO-Crate the expression of `Class` and `Property `is typically indirect, as the JSON-LD `@context` maps to schema.org IRIs, which when resolved as Linked Data embeds their formal definition as RDFa.

### RO-Crate 1.1 Metadata File Descriptor

An important RO-Crate principle is that of being **self-describing**. Therefore the serialization of the RO-Crate into a file should also describe itself in a [Metadata File Descriptor](https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#ro-crate-metadata-file-descriptor), indicating it is about (describing) the RO-Crate root data entity, and that it conformsTo a particular version of the RO-Crate specification:

    about(s,o) →  Relation(s, <http://schema.org/about>, o)
    conformsTo(s,o) →  Relation(s, <http://purl.org/dc/terms/conformsTo>, R)
    MetadataFileDescriptor(m) →  ( CreativeWork(m) ∧ about(m,R) ∧ RO-Crate(R) ∧ 
        conformsTo(m, <https://w3id.org/ro/crate/1.1>) )

Note that although the metadata file necessarily is an _information resource_ written to disk or served over the network (e.g. as JSON-LD), it is not considered to be a contained _part_ of the RO-Crate in the form of a _data entity_, rather it is described only as a _contextual entity_.

While in the conceptual model the _RO-Crate Metadata File_ can be seen as the top-level node that describes the _RO-Crate Root_, in the formal model (and the JSON-LD format) the metadata file descriptor is an additional contextual entity and not affecting the depth-limit of the RO-Crate.

### Forward-chained Production Rules for JSON-LD

Combining the above predicates and schema.org mapping with rudimentary JSON templates, these forward-chaining production rules can output JSON-LD according to the RO-Crate 1.1 specification[^2]:

    Mentions(R, s) ∧ Relation(s, p, o) →  Mentions(R, o)
    i ∈ 𝕀𝕣𝕚 → "i"
    r ∈ ℝ →  r
    s ∈ 𝕊 → "s"
    ∀s∀p∀o Relation(s,p,o) →  { "@id": s,
                                p: { "@id": o }
                              }     
    ∀s∀p∀v Attribute(s,p,v) →  { "@id": s,
                                p: v 
                               }
    ∀r∀c  RO-Crate(r) →  { "@graph": [ Mentions(r, c)* ] }
    R ⊨  <./>
    MetadataFileDescriptor(<ro-crate-metadata.json>)

This exposes the first order logic domain of discourse of IRIs, with rational numbers and strings as their corresponding JSON-LD representation. These production rules first grow the graph of `R` by adding a transitive rule that anything described in `R` which is related to `o` means that `o` is also mentioned by the RO-Crate `R`. For simplicity this rule is one-way; in practice the JSON-LD graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities.

[^2]:
    Limitations: The full list of types, relations and attribute properties from the RO-Crate specification are not included. Examples shown include `datePublished`, `CreativeWork` and `name`. Contextual entities not related from the RO-Crate (e.g. using inverse relations to a data entity) would not be covered by the single direction `Mentions(R, s)` production rule; see [issue #122](https://github.com/ResearchObject/ro-crate/issues/122). The `datePublished(e, date)` rule do not include syntax checks for the ISO 8601 datetime format. Compared with RO-Crate examples, this generated JSON-LD does not use a `@context` as the IRIs are produced unshortened; a post-step could do JSON-LD Flattening with a versioned RO-Crate context.

### 2021-05-04 latex/ro-crate.tex
use ro-crate.bib
markdown settings
fixing more DOIs and markdown
using iosevka light

@@
 %\usepackage[utf8]{inputenc}
+
+%\usepackage[backend=biber]{biblatex}
+%\addbibresource{ro-crate.bib}
 
@@
 \graphicspath{../content/images/}
-\usepackage[smartEllipses,tableCaptions,pipeTables,fencedCode,footnotes,underscores,hybrid]{markdown}
+\usepackage[citations,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
 \def\markdownOptionOutputDir{../output}
@@
 
+% Welcome to the 2000s with UNICODE fonts!
+% Adapted from
+% https://www.overleaf.com/learn/latex/Articles%2FOpenType-based_math_typesetting:_An_introduction_to_the_STIX2_OpenType_fonts
 \setmathfont{STIX2Math}[
@@
 BoldItalicFont={*-BoldItalic}]
+% https://typeof.net/Iosevka/ mono font has 
+% all; the logic symbols we need
+\setmonofont{iosevka-slab}[
+Extension={.otf},
+Path=./ttf-iosevka-slab-6.1.3/,
+Scale=MatchLowercase,
+UprightFont={*-light},
+ItalicFont={*-lightitalic}
+]
 
@@
 
-%\section{}\label{s1}
-
-%\subsection{}\label{s1.1}
-
+% Every *.md file in content in order, except 00, 90
 \markdownInput{../content/10.background.md}
@@
 \markdownInput{../content/80.formal-definition.md}
+% TODO: Recode formal 
+%\input{../content/80.formal-definition.tex}
 
@@
 \bibliographystyle{ios1}           % Style BST file.
-\bibliography{bibliography}        % Bibliography file (usually '*.bib')
+\bibliography{ro-crate}        % Bibliography file (usually '*.bib')
 

### 2021-05-04 content/80.formal-definition.tex
Cite RFC and use `backticks` font

@@
     𝔻 =  𝕀𝕣𝕚
-    𝕀𝕣𝕚 ≡  { IRIs as defined in <https://tools.ietf.org/html/rfc3987> }
+    𝕀𝕣𝕚 ≡  { IRIs as defined in RFC3987 }
     ℝ ≡  { real or integer numbers }
@@
 
-The domain of discourse is the set of 𝕀𝕣𝕚 identifiers (notation `<http://example.com/>`), with additional descriptions using numbers ℝ (notation `13.37`) and literal strings 𝕊 (notation `“Hello”`).
+The domain of discourse is the set of `𝕀𝕣𝕚` identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`), with additional descriptions using numbers ℝ (notation `13.37`) and literal strings `𝕊` (notation `“Hello”`).
 
-From this formalized language `𝕃ro-crate` a RO-Crate can be interpreted in any representation that can gather these descriptions, their properties, classes, and literal attributes.
+From this formalized language `𝕃ro-crate` an RO-Crate can be interpreted in any representation that can gather these descriptions, their properties, classes, and literal attributes.
 
@@
 
-The `Root(r)` is a type of Dataset(r), and must have the metadata to literal attributes to provide a `name`, `description` and `datePublished`, as well as a contextual entity identifying its license.
+The `Root(r)` is a type of `Dataset(r)`, and must have the metadata to literal attributes to provide a `name`, `description` and `datePublished`, as well as a contextual entity identifying its license.
 These predicates correspond to the RO-Crate 1.1 [requirements for the root data entity](https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#direct-properties-of-the-root-data-entity).

### 2021-05-04 latex/ro-crate.tex

@@
     Path=./STIX2fonts/,
-Scale=1]
+    Scale=1
+]
 \setmainfont{STIX2Text}[
@@
     ItalicFont={*-Italic},
-BoldItalicFont={*-BoldItalic}]
+    BoldItalicFont={*-BoldItalic}
+]
 % https://typeof.net/Iosevka/ mono font has 

### 2021-05-04 build/build.sh
include bibtex refs

@@ -15,6 +15,7 @@ export LC_ALL=en_US.UTF-8
 ln -fs v/latest/manuscript.pdf webpage/
 ln -fs v/latest/images webpage/
 ln -fs v/latest/index.html webpage/
+ln -s ../latex/ro-crate.bib content/manual-references-ro-crate.bib
 
 
 # Generate reference information

### 2021-05-04 latex/ro-crate.tex
tidy up latex comments
Updates from Overleaf

@@
 \documentclass[ds]{iosart2x}
-%\usepackage[utf8]{inputenc}
 
-%\usepackage[backend=biber]{biblatex}
-%\addbibresource{ro-crate.bib}
+%%%%%%%%%%% Put your definitions here
 
@@
 \usepackage{amsmath}
-\usepackage{hyperref}
+% \usepackage{hyperref}
 \usepackage{fontspec}
-\usepackage{unicode-math}
-\usepackage{graphicx}
-
-%%
+% \usepackage{unicode-math}
+% \usepackage{graphicx}
 
@@
 \def\markdownOptionOutputDir{../output}
-
-%\markdownSetup{
-%    defaultOptions.cacheDir = 
-%}
-
-%\usepackage{dcolumn}
-
-%%%%%%%%%%% Put your definitions here
 
@@
     Scale=MatchLowercase,
-    UprightFont={*-light},
+    UprightFont={*-thin},
     ItalicFont={*-lightitalic}
@@
 
-% Every *.md file in content in order, except 00, 90
+% Every *.md file in content/ in order, except 00, 90
 \markdownInput{../content/10.background.md}
@@
 \markdownInput{../content/80.formal-definition.md}
-% TODO: Recode formal 
+% TODO: Recode formal definition to LaTeX math mode?
 %\input{../content/80.formal-definition.tex}
 
-%\begin{figure}[t]
-%\includegraphics{}
-%\caption{Figure caption.}\label{f1}
-%\end{figure}
-
-%\begin{table*}
-%\caption{} \label{t1}
-%\begin{tabular}{lll}
-%\hline
-%&&\\
-%&&\\
-%\hline
-%\end{tabular}
-%\end{table*}
-
 %%%%%%%%%%% The bibliography starts:
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  ios1.bst will be used to                               %%
-%%  create a .BBL file for submission.                     %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-% or include bibliography directly:
-%\begin{thebibliography}{0}
-%\bibitem{r1} F. Author, Information about cited object.
-%
-%\bibitem{r2} S. Author and T. Author, Information about cited object.
-%\end{thebibliography}
-
 \end{document}

### 2021-05-04 latex/ro-crate.tex
More figures from powerpoint

@@
 \usepackage{fontspec}
-% \usepackage{unicode-math}
+\usepackage{unicode-math}
 % \usepackage{graphicx}
@@
 \setmonofont{iosevka-slab}[
-    Extension={.otf},
+    Extension={.ttf},
     Path=./ttf-iosevka-slab-6.1.3/,
     Scale=MatchLowercase,
-    UprightFont={*-thin},
+    UprightFont={*-light},
     ItalicFont={*-lightitalic}

### 2021-05-05 build/build.sh
attempt to get figures working again with Manubot

@@ -17,6 +17,11 @@ ln -fs v/latest/images webpage/
 ln -fs v/latest/index.html webpage/
 ln -s ../latex/ro-crate.bib content/manual-references-ro-crate.bib
 
+# Replace silly paths for LaTeX
+sed -i 's,../content/images/,images/,g' content/*md
+# .. and use SVG instead of PDF
+sed -i 's,\.pdf,.svg/,g' content/*md
+
 
 # Generate reference information
 echo >&2 "Retrieving and processing reference metadata"

### 2021-05-05 build/build.sh
correct file ending

@@ -20,7 +20,7 @@ ln -s ../latex/ro-crate.bib content/manual-references-ro-crate.bib
 # Replace silly paths for LaTeX
 sed -i 's,../content/images/,images/,g' content/*md
 # .. and use SVG instead of PDF
-sed -i 's,\.pdf,.svg/,g' content/*md
+sed -i 's,\.pdf,.svg,g' content/*md
 
 
 # Generate reference information

### 2021-05-05 build/build.sh
aaaah

@@ -22,6 +22,14 @@ sed -i 's,../content/images/,images/,g' content/*md
 # .. and use SVG instead of PDF
 sed -i 's,\.pdf,.svg,g' content/*md
 
+## Horrible latex seds for the figure captions
+sed -i -E 's/^!\[(.*)\]\((.*) "([^"]*)"\)/![\3](\2 "\1")/' content/*md
+sed -i -E 's/\\textbf{([^}]*)}/**\1**/g' content/*md
+sed -i -E 's/\\emph{([^}]*)}/*\1*/g' content/*md
+sed -i -E 's/\\texttt{([^}]*)}/`\1`/g' content/*md
+sed -i -E 's/\\cite{([^}]*)}/[@\1]/g' content/*md
+sed -i -E 's/\\citet{([^}]*)}/@\1/g' content/*md
+
 
 # Generate reference information
 echo >&2 "Retrieving and processing reference metadata"

### 2021-05-05 build/build.sh
sed sed SED

@@ -23,12 +23,12 @@ sed -i 's,../content/images/,images/,g' content/*md
 sed -i 's,\.pdf,.svg,g' content/*md
 
 ## Horrible latex seds for the figure captions
-sed -i -E 's/^!\[(.*)\]\((.*) "([^"]*)"\)/![\3](\2 "\1")/' content/*md
-sed -i -E 's/\\textbf{([^}]*)}/**\1**/g' content/*md
-sed -i -E 's/\\emph{([^}]*)}/*\1*/g' content/*md
-sed -i -E 's/\\texttt{([^}]*)}/`\1`/g' content/*md
-sed -i -E 's/\\cite{([^}]*)}/[@\1]/g' content/*md
-sed -i -E 's/\\citet{([^}]*)}/@\1/g' content/*md
+sed -E -i 's/^!\[(.*)\]\((.*) "([^"]*)"\)/![\3](\2 "\1")/g' content/*md
+sed -E -i 's/\\textbf{([^}]*)}/**\1**/g' content/*md
+sed -E -i 's/\\emph{([^}]*)}/*\1*/g' content/*md
+sed -E -i 's/\\texttt{([^}]*)}/`\1`/g' content/*md
+sed -E -i 's/\\cite{([^}]*)}/[@\1]/g' content/*md
+sed -E -i 's/\\citet{([^}]*)}/@\1/g' content/*md
 
 
 # Generate reference information

### 2021-05-05 build/build.sh
more sed is good for you

@@ -24,11 +24,11 @@ sed -i 's,\.pdf,.svg,g' content/*md
 
 ## Horrible latex seds for the figure captions
 sed -E -i 's/^!\[(.*)\]\((.*) "([^"]*)"\)/![\3](\2 "\1")/g' content/*md
-sed -E -i 's/\\textbf{([^}]*)}/**\1**/g' content/*md
-sed -E -i 's/\\emph{([^}]*)}/*\1*/g' content/*md
-sed -E -i 's/\\texttt{([^}]*)}/`\1`/g' content/*md
-sed -E -i 's/\\cite{([^}]*)}/[@\1]/g' content/*md
-sed -E -i 's/\\citet{([^}]*)}/@\1/g' content/*md
+sed -E -i 's/\\textbf\{([^}]*)\}/**\1**/g' content/*md
+sed -E -i 's/\\emph\{([^}]*)\}/*\1*/g' content/*md
+sed -E -i 's/\\texttt\{([^}]*)\}/`\1`/g' content/*md
+sed -E -i 's/\\cite\{([^}]*)\}/[@\1]/g' content/*md
+sed -E -i 's/\\citet\{([^}]*)\}/@\1/g' content/*md
 
 
 # Generate reference information

### 2021-05-05 build/build.sh
avoid latex comments in manubot

@@ -29,6 +29,8 @@ sed -E -i 's/\\emph\{([^}]*)\}/*\1*/g' content/*md
 sed -E -i 's/\\texttt\{([^}]*)\}/`\1`/g' content/*md
 sed -E -i 's/\\cite\{([^}]*)\}/[@\1]/g' content/*md
 sed -E -i 's/\\citet\{([^}]*)\}/@\1/g' content/*md
+# remove latex comments
+sed -E -i 's/^%.*//g' content/*md
 
 
 # Generate reference information

### 2021-05-06 latex/ro-crate.tex
formal definitions are in appendix
appendix moved down
Updates from Overleaf

@@
 
-\DeclareGraphicsExtensions{.png,.pdf}
+\DeclareGraphicsExtensions{.pdf,.png,.svg}
 \graphicspath{../content/images/}
@@
 ]
+
 % https://typeof.net/Iosevka/ mono font has 
@@
 \author[C]{\inits{P.S.}\fnms{Peter} \snm{Sefton}\ead[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}},
+\author[D]{RO-Crate Community},
 \author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}},
-\author[B]{\inits{P.G.}\fnms{Paul} \snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}}
+\author[B]{\inits{P.G.}\fnms{Paul} 
+\snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}}
 \address[A]{Computer Science, \orgname{The University of Manchester},
@@
 \cny{Australia}\printead[presep={\\}]{e3,o3}}
+\address[D]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
 
@@
 \markdownInput{../content/40.inuse.md}
+\markdownInput{../content/41.workflows.md}
+\markdownInput{../content/42.regulatory.md}
+\markdownInput{../content/43.culturalheritage.md}
 \markdownInput{../content/50.relatedwork.md}
 \markdownInput{../content/60.conclusion.md}
+\markdownInput{../content/70.acknowledgements.md}
+
+%%%%%%%%%%% The bibliography starts:
+\nocite{*} 
+\bibliographystyle{ios1}
+\bibliography{ro-crate}
+
+%%%% Appendix
+
+\appendix
 \markdownInput{../content/80.formal-definition.md}
@@
 %\input{../content/80.formal-definition.tex}
-
-%%%%%%%%%%% The bibliography starts:
-
-\nocite{*} 
-% if your bibliography is in bibtex format, use those commands:
-\bibliographystyle{ios1}           % Style BST file.
-\bibliography{ro-crate}        % Bibliography file (usually '*.bib')
+\markdownInput{../content/89.community.md}
 

### 2021-05-09 content/31.tooling-table.tex
Sunday fun with tables in LaTeX

\begin{table}[ht]
	\centering
	\begin{tabular}{llll}
		\toprule
		\textbf{Tool name} & Targets & Language / Platform & Status \\
		\multicolumn{4}{l}{\it Brief Description} \\
		\midrule

		\textbf{Describo} \citep{describo} & Research Data Managers & NodeJS (Desktop) & RC \\
		\multicolumn{4}{l}{\it Interactive desktop application to create, update and export RO-Crates for different profiles} \\

		\textbf{Describo Online} \citep{describo-online} & Platform developers & NodeJS (Web) & Alpha \\
		\multicolumn{4}{l}{\it Web-based application to create RO-Crates using cloud storage} \\

		\textbf{ro-crate-excel} \citep{ro-crate-excel} & Data managers & JavaScript \\
		\multicolumn{4}{l}{\it Command-line tool to help create RO-Crates and HTML-readable rendering} \\

		\textbf{ro-crate-html-js} \citep{ro-crate-html-js} & Developers & JavaScript & Beta \\
		\multicolumn{4}{l}{\it HTML rendering of RO-Crate}  \\

		\textbf{ro-crate-js} \citep{ro-crate-js} & Research Data Managers & JavaScript & Alpha \\
		\multicolumn{4}{l}{\it Library for creating/manipulating crates; basic validation code} \\

		\textbf{ro-crate-ruby} \citep{ro-crate-ruby} & Developers & Ruby & Beta \\
		\multicolumn{4}{l}{\it Ruby library for reading/writing RO-Crate, with workflow support} \\

		\textbf{ro-crate-py} \citep{ro-crate-py} & Developers & Python & Beta \\
		\multicolumn{4}{l}{\it Object-oriented Python library for reading/writing RO-Crate} \\

		\textbf{WorkflowHub} \citep{about-workflowhub} & Workflow users & Ruby & Beta \\
		\multicolumn{4}{l}{\it Workflow repository; imports and exports Workflow RO-Crate} \\

		\textbf{Modern PARADISEC} \citep{modpdsc} & Repository managers & Platform & Beta \\
		\multicolumn{4}{l}{\it Cultural Heritage portal based on OCFL and RO-Crate} \\

		\textbf{ONI express} \citep{arkisto-data-portal} & Repository managers & Platform & Beta \\
		\multicolumn{4}{l}{\it Platform for publishing data and documents stored in an OCFL repository via a web interface} \\

		\textbf{ocfl-tools} \citep{ocfl-tools} & Developers & JavaScript (CLI) & Beta \\
		\multicolumn{4}{l}{\it Tools for managing RO-Crates in an OCFL repository}\\

		\textbf{RO Composer} \citep{ro-composer} & Repository developers & Java & Alpha \\ 
		\multicolumn{4}{l}{\it REST API for gradually building ROs for given profile} \\

		\textbf{galaxy2cwl} \citep{galaxy2cwl} & Workflow developers & Python & Alpha \\
		\multicolumn{4}{l}{\it Wraps Galaxy workflow as Workflow RO-Crate} \\
		\bottomrule
	\end{tabular}
	\caption{Applications and libraries implementing RO-Crate, targeting different types of users across multiple programming languages. Status is indicative as assessed by this work.}
	\label{tab:tools}
\end{table}

### 2021-05-10 content/80.formal-definition.tex
More LaTeX IsN'tiT

@@
-# Appendix
+%% TODO: Convert to LaTex math mode. See GitHub rendering for original Unicode symbols
+% https://github.com/stain/ro-crate-paper/blob/master/content/80.formal-definition.md
 
+% Text remains for now as Markdown
+
+\begin{markdown}
 ## Formalizing RO-Crate in First Order Logic
@@
 
-    𝕃ro-crate = { Property(p), Class(c), Literal(x), ℝ, 𝕊 }
-    𝔻 =  𝕀𝕣𝕚
-    𝕀𝕣𝕚 ≡  { IRIs as defined in RFC3987 }
+\end{markdown}
+
+    $\mathbb{L}$ro-crate = { Property(p), Class(c), Literal(x), ℝ, $\mathbb{S}$ }
+    $\mathbb{D}$ =  $\mathbb{Iri}$
+    $\mathbb{Iri}$ ≡  { IRIs as defined in RFC3987 }
     ℝ ≡  { real or integer numbers }
-    𝕊 ≡  { literal strings }
+    $\mathbb{S}$ ≡  { literal strings }
 
-The domain of discourse is the set of `𝕀𝕣𝕚` identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`), with additional descriptions using numbers ℝ (notation `13.37`) and literal strings `𝕊` (notation `“Hello”`).
+\begin{markdown}
+The domain of discourse is the set of $\mathbb{Iri}$ identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`), with additional descriptions using numbers ℝ (notation `13.37`) and literal strings $\mathbb{S}$ (notation `“Hello”`).
 
-From this formalized language `𝕃ro-crate` an RO-Crate can be interpreted in any representation that can gather these descriptions, their properties, classes, and literal attributes.
+From this formalized language `$\mathbb{L}$ro-crate` an RO-Crate can be interpreted in any representation that can gather these descriptions, their properties, classes, and literal attributes.  
+\end{markdown}
 
+\begin{markdown}
 ### Minimal RO-Crate
 
-Below is using `𝕃ro-crate` to define a minimal RO-Crate:
+Below is using `$\mathbb{L}$ro-crate` to define a minimal RO-Crate:
+\end{markdown}
 
@@
     Attribute(s, p, x) ⊨ Entity(s)  ∧ Property(p) ∧ Literal(x)
-    Literal(x) ≡  x ∈ ℝ  ⊕   x ∈ 𝕊
+    Literal(x) ≡  x ∈ ℝ  ⊕   x ∈  $\mathbb{S}$
 
+\begin{markdown}
 An `RO-Crate(R)` is defined as a self-described _Root Data Entity_, which describes and contains parts (_data entities_), which are further described in _contextual entities_.
@@
 `Relation(s, p, o)` and `Attribute(s, p, x)` are defined as a _subject-predicate-object_ triple pattern from an `Entity(s)` using a `Property(p)` to either another `Entity(o)` or a `Literal(x)` value.
+\end{markdown}
 
+\begin{markdown}
 ### Example of formalized RO-Crate
@@
 The below is an example RO-Crate represented using the above formalization, assuming a base URI of `http://example.com/ro/123/`:
+\end{markdown}
 
@@
 
+\begin{markdown}
 In reality many additional attributes from schema.org types like <http://schema.org/Dataset> and <http://schema.org/CreativeWork> would be used to further describe the RO-Crate and its entities, but as these are optional they do not form part of this formalization.
+\end{markdown}
 
@@
     Mentions(R, s) ∧ Relation(s, p, o) →  Mentions(R, o)
-    i ∈ 𝕀𝕣𝕚 → "i"
+    i ∈ $\mathbb{Iri}$ → "i"
     r ∈ ℝ →  r
-    s ∈ 𝕊 → "s"
+    s ∈ $\mathbb{S}$ → "s"
     ∀s∀p∀o Relation(s,p,o) →  { "@id": s,

### 2021-05-10 content/31.tooling-table.tex
Fixed table
Added SCHEMA and LifeMonitor and RO-Crate DMP to table

@@
-\begin{table}[ht]
+\begin{table}[htbp]
 	\centering
@@
 
+		\textbf{Life Monitor} \citep{about-lifemonitor} & Workflow developers & Python & Beta \\
+		\multicolumn{4}{l}{\it Workflow testing and monitoring service; Workflow Testing profile of RO-Crate} \\
+
+        \textbf{SCHeMa} & Workflow users & PHP & Alpha \\
+        \multicolumn{4}{l}{\it Workflow execution using RO-Crate as exchange mechanism } \\
+
+		\textbf{galaxy2cwl} \citep{galaxy2cwl} & Workflow developers & Python & Alpha \\
+		\multicolumn{4}{l}{\it Wraps Galaxy workflow as Workflow RO-Crate} \\
+
 		\textbf{Modern PARADISEC} \citep{modpdsc} & Repository managers & Platform & Beta \\
@@
 
-		\textbf{galaxy2cwl} \citep{galaxy2cwl} & Workflow developers & Python & Alpha \\
-		\multicolumn{4}{l}{\it Wraps Galaxy workflow as Workflow RO-Crate} \\
+        \textbf{Ro-Crate\_2\_ma-DMP} \citep{doi:10.5281/zenodo.3903463} & Data Management Plan users & Python & Beta \\
+        \multicolumn{4}{l}{\it Convert between machine-actionable data management plans (maDMP) and RO-Crate \citep{doi:10.4126/frl01-006423291} }
+
 		\bottomrule

### 2021-05-10 content/31.tooling-table.tex
citation for SCHeMa
for some reason there are two maDMP mappings

@@
 
-        \textbf{SCHeMa} & Workflow users & PHP & Alpha \\
-        \multicolumn{4}{l}{\it Workflow execution using RO-Crate as exchange mechanism } \\
+        \textbf{SCHeMa} \citep{athenarc-schema} & Workflow users & PHP & Alpha \\
+        \multicolumn{4}{l}{\it Workflow execution using RO-Crate as exchange mechanism \citep{doi:10.5281/zenodo.4671709} } \\
 
@@
 
+		\textbf{RDA maDMP Mapper} [@doi:10.5281/zenodo.3922136] & Data Management Plan users & Python & Beta \\
+		\multicolumn{4}{l}{\it Mapping between machine-actionable data management plans (maDMP) and RO-Crate\citep{doi:10.4126/frl01-006423291} }
+
         \textbf{Ro-Crate\_2\_ma-DMP} \citep{doi:10.5281/zenodo.3903463} & Data Management Plan users & Python & Beta \\
-        \multicolumn{4}{l}{\it Convert between machine-actionable data management plans (maDMP) and RO-Crate \citep{doi:10.4126/frl01-006423291} }
+        \multicolumn{4}{l}{\it Convert between machine-actionable data management plans (maDMP) and RO-Crate }
 

### 2021-05-11 content/31.tooling-table.tex
Updates from Overleaf

--- 2021-05-10 content/31.tooling-table.tex
@@
 
-        \textbf{SCHeMa} & Workflow users & PHP & Alpha \\
-        \multicolumn{4}{l}{\it Workflow execution using RO-Crate as exchange mechanism } \\
+        \textbf{SCHeMa} \citep{arxiv:2103.13138v1} & Workflow users & PHP & Alpha \\
+        \multicolumn{4}{l}{\it Workflow execution using RO-Crate as exchange mechanism} \\
 
@@
 
+		\textbf{RDA maDMP Mapper} \citep{doi:10.5281/zenodo.3922136} & Data Management Plan users & Python & Beta \\
+		\multicolumn{4}{l}{\it Mapping between machine-actionable data management plans (maDMP) and RO-Crate\citep{doi:10.4126/frl01-006423291} } \\
+
         \textbf{Ro-Crate\_2\_ma-DMP} \citep{doi:10.5281/zenodo.3903463} & Data Management Plan users & Python & Beta \\
-        \multicolumn{4}{l}{\it Convert between machine-actionable data management plans (maDMP) and RO-Crate \citep{doi:10.4126/frl01-006423291} }
+        \multicolumn{4}{l}{\it Convert between machine-actionable data management plans (maDMP) and RO-Crate } \\
 

### 2021-05-14 content/80.formal-definition.tex
Changes from Nick and others

@@
 
-From this formalized language `$\mathbb{L}$ro-crate` an RO-Crate can be interpreted in any representation that can gather these descriptions, their properties, classes, and literal attributes.  
+From this formalised language `$\mathbb{L}$ro-crate` an RO-Crate can be interpreted in any representation that can gather these descriptions, their properties, classes, and literal attributes.  
 \end{markdown}

### 2021-05-14 content/80.formal-definition.tex

@@
-%% TODO: Convert to LaTex math mode. See GitHub rendering for original Unicode symbols
+%% TODO: Being onverted to LaTex math mode. See GitHub rendering for original Unicode symbols
 % https://github.com/stain/ro-crate-paper/blob/master/content/80.formal-definition.md
 
-% Text remains for now as Markdown
+% Text remains for now as Markdown blocks
 
@@
 
-    $\mathbb{L}$ro-crate = { Property(p), Class(c), Literal(x), ℝ, $\mathbb{S}$ }
-    $\mathbb{D}$ =  $\mathbb{Iri}$
-    $\mathbb{Iri}$ ≡  { IRIs as defined in RFC3987 }
-    ℝ ≡  { real or integer numbers }
-    $\mathbb{S}$ ≡  { literal strings }
+Definition of language $\mathcal{L}_{rocrate}$:
+
+\begin{eqnarray*}
+    \mathcal{L}_{rocrate}   & \equiv & { Property(p), Class(c), 
+                            Literal(x), \mathbb{R}, \mathbb{S} } \\
+    \mathbb{D}              & \equiv & \mathbb{IRI} \\
+    \mathbb{IRI}            & \equiv & { \text{IRIs as defined in RFC3987} } \\
+    \mathbb{R}              & \equiv & { \text{real or integer numbers} } \\
+    \mathbb{S}              & \equiv & { \text{literal strings} }
+\end{eqnarray*}
 
 \begin{markdown}
-The domain of discourse is the set of $\mathbb{Iri}$ identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`), with additional descriptions using numbers ℝ (notation `13.37`) and literal strings $\mathbb{S}$ (notation `“Hello”`).
+The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`), with additional descriptions using numbers $\mathbb{R}$ (notation `13.37`) and literal strings $\mathbb{S}$ (notation `"Hello"`).
 
-From this formalised language `$\mathbb{L}$ro-crate` an RO-Crate can be interpreted in any representation that can gather these descriptions, their properties, classes, and literal attributes.  
+From this formalised language $\mathcal{L}_{rocrate}$ we can interpret an RO-Crate in any representation that can gather these descriptions, their properties, classes, and literal attributes.  
 \end{markdown}
@@
 
-Below is using `$\mathbb{L}$ro-crate` to define a minimal RO-Crate:
+Below we use $\mathcal{L}_{rocrate}$ to define a minimal RO-Crate:
 \end{markdown}
 
-    RO-Crate(R) ⊨  Root(R) ∧ Mentions(R, R) ∧
-        hasPart(R, d) ∧ Mentions(R, d) ∧ DataEntity(d) ∧
-        Mentions(R, c) ∧ ContextualEntity(c)
-    ∀r Root(r) →  Dataset(r) ∧ name(r, n)∧ description(r, d)
-                            ∧ published(r, date) ∧ license(e, l)
-    ∀e∀n name(e, n) →  Literal(n)
-    ∀e∀d description(e, d) →  Literal(d)
-    ∀e∀date datePublished(e, date) →  Literal(date)
-    ∀e∀l license(e, l) →  ContextualEntity(l)
-    DataEntity(e) ≡  File(e) ⊕ Dataset(e)
-    Entity(e) ≡  DataEntity(e) ∨ ContextualEntity(e)
-    ∀e Entity(e) → Class(e)
-    Mentions(R, s) ⊨  Relation(s, p, e)  ⊕  Attribute(s, p, l)
-    Relation(s, p, o) ⊨ Entity(s) ∧ Property(p) ∧ Entity(o)
-    Attribute(s, p, x) ⊨ Entity(s)  ∧ Property(p) ∧ Literal(x)
-    Literal(x) ≡  x ∈ ℝ  ⊕   x ∈  $\mathbb{S}$
+\begin{eqnarray*}
+ROCrate(R)                                 & \models & Root(R) \land Mentions(R, R) \land hasPart(R, d) \land \\
+                                            & & Mentions(R, d) \land DataEntity(d) \land \\
+                                            & & Mentions(R, c) \land ContextualEntity(c) \\
+\forall r Root(r)                           & \Rightarrow & Dataset(r) \land name(r, n) \land description(r, d) \land \\
+                                            & &             published(r, date) \land license(e, l) \\
+\forall e \forall n \ name(e, n)            & \Rightarrow & Literal(n) \\
+\forall e \forall s \ description(e, s)     & \Rightarrow & Literal(s) \\
+\forall e \forall d \ datePublished(e, d)   & \Rightarrow & Literal(d) \\
+\forall e \forall l \ license(e, l)         & \Rightarrow & ContextualEntity(l) \\
+DataEntity(e)                               & \equiv &      File(e) \oplus Dataset(e) \\
+Entity(e)                                   & \equiv &      DataEntity(e) \lor ContextualEntity(e) \\
+\forall e \ Entity(e)                       & \Rightarrow & Class(e) \\
+Mentions(R, s)                              & \models &     Relation(s, p, e) \oplus Attribute(s,  p, l) \\
+Relation(s, p, o)                           & \models &     Entity(s) \land Property(p) \land  Entity(o) \\
+Attribute(s, p, x)                          & \models &     Entity(s) \land Property(p) \land Literal(x) \\
+Literal(x)                                  & \equiv &      x \in \mathbb{R} \oplus x \in \mathbb{S}
+\end{eqnarray*}
 
 \begin{markdown}
-An `RO-Crate(R)` is defined as a self-described _Root Data Entity_, which describes and contains parts (_data entities_), which are further described in _contextual entities_.
+An $ROCrate(R)$ is defined as a self-described _Root Data Entity_, which describes and contains parts (_data entities_), which are further described in _contextual entities_.
 These terms align with their use in the [RO-Crate 1.1 terminology](https://www.researchobject.org/ro-crate/1.1/terminology).
 
-The `Root(r)` is a type of `Dataset(r)`, and must have the metadata to literal attributes to provide a `name`, `description` and `datePublished`, as well as a contextual entity identifying its license.
+The $Root(r)$ is a type of $Dataset(r)$, and must have the metadata to literal attributes to provide a $name$, $description$ and $datePublished$, as well as a contextual entity identifying its license.
 These predicates correspond to the RO-Crate 1.1 [requirements for the root data entity](https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#direct-properties-of-the-root-data-entity).
 
-The concept of an `Entity(e)` is introduced as being either a DataEntity(e), a `ContextualEntity(e)`, or [both](https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities); and must be typed with at least one `Class(e)`.
+The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or [both](https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities); and must be typed with at least one $Class(e)$.
 
-For simplicity in this formalization (and to assist production rules below) `R` is a constant representing a single RO-Crate, typically written to independent RO-Crate Metadata files.
-`R` is used by `Mentions(R, e)` to indicate that `e` is an Entity described by the RO-Crate and therefore its metadata (a set of Relation and Attribute predicates) form part of the RO-Crate serialization.
-`Relation(s, p, o)` and `Attribute(s, p, x)` are defined as a _subject-predicate-object_ triple pattern from an `Entity(s)` using a `Property(p)` to either another `Entity(o)` or a `Literal(x)` value.
+For simplicity in this formalization (and to assist production rules below) $R$ is a constant representing a single RO-Crate, typically written to independent RO-Crate Metadata files.
+$R$ is used by $Mentions(R, e)$ to indicate that $e$ is an Entity described by the RO-Crate and therefore its metadata (a set of $Relation$ and $Attribute$ predicates) form part of the RO-Crate serialization.
+$Relation(s, p, o)$ and $Attribute(s, p, x)$ are defined as a _subject-predicate-object_ triple pattern from an $Entity(s)$ using a $Property(p)$ to either another $Entity(o)$ or a $Literal(x)$ value.
 \end{markdown}
@@
 
-The below is an example RO-Crate represented using the above formalization, assuming a base URI of `http://example.com/ro/123/`:
-\end{markdown}
+The below is an example RO-Crate represented using the above formalization, assuming a base URI of `<http://example.com/ro/123/>`:
 
-    RO-Crate(<http://example.com/ro/123/>)
+    ROCrate(<http://example.com/ro/123/>)
     name(<http://example.com/ro/123/, 
@@
         “Audio recordings of care provider interviews”)
+\end{markdown}
 
@@
 
+\begin{markdown}
 ### Mapping to RDF with schema.org
@@
 
-Note that in the JSON-LD serialization of RO-Crate the expression of `Class` and `Property `is typically indirect, as the JSON-LD `@context` maps to schema.org IRIs, which when resolved as Linked Data embeds their formal definition as RDFa.
+Note that in the JSON-LD serialization of RO-Crate the expression of $Class$ and $Property $is typically indirect, as the JSON-LD $@context$ maps to schema.org IRIs, which when resolved as Linked Data embeds their formal definition as RDFa.
+\end{markdown}
+
+\begin{markdown}
 
@@
 While in the conceptual model the _RO-Crate Metadata File_ can be seen as the top-level node that describes the _RO-Crate Root_, in the formal model (and the JSON-LD format) the metadata file descriptor is an additional contextual entity and not affecting the depth-limit of the RO-Crate.
+\end{markdown}
 
+\begin{markdown}
 ### Forward-chained Production Rules for JSON-LD
@@
 This exposes the first order logic domain of discourse of IRIs, with rational numbers and strings as their corresponding JSON-LD representation.
-These production rules first grow the graph of `R` by adding a transitive rule that anything described in `R` which is related to `o` means that `o` is also mentioned by the RO-Crate `R`.
+These production rules first grow the graph of $R$ by adding a transitive rule – anything described in $R$ which is related to $o$, means that $o$ is also mentioned by the $ROCrate(R)$.
 For simplicity this rule is one-way; in practice the JSON-LD graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities.
@@
     Limitations: The full list of types, relations and attribute properties from the RO-Crate specification are not included.
-Examples shown include `datePublished`, `CreativeWork` and `name`.
+Examples shown include $datePublished$, $CreativeWork$ and $name$.
 Contextual entities not related from the RO-Crate (e.g.
-using inverse relations to a data entity) would not be covered by the single direction `Mentions(R, s)` production rule; see [issue #122](https://github.com/ResearchObject/ro-crate/issues/122).
-The `datePublished(e, date)` rule do not include syntax checks for the ISO 8601 datetime format.
-Compared with RO-Crate examples, this generated JSON-LD does not use a `@context` as the IRIs are produced unshortened; a post-step could do JSON-LD Flattening with a versioned RO-Crate context.
+using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see [issue 122](https://github.com/ResearchObject/ro-crate/issues/122).
+The $datePublished(e, date)$ rule do not include syntax checks for the ISO 8601 datetime format.
+Compared with RO-Crate examples, this generated JSON-LD does not use a $@context$ as the IRIs are produced unshortened; a post-step could do JSON-LD Flattening with a versioned RO-Crate context.
+\end{markdown}

### 2021-05-14 content/80.formal-definition.tex
definition lists..

@@
-%% TODO: Being onverted to LaTex math mode. See GitHub rendering for original Unicode symbols
+%% TODO: Convert to LaTex math mode. See GitHub rendering for original Unicode symbols
 % https://github.com/stain/ro-crate-paper/blob/master/content/80.formal-definition.md
 
-% Text remains for now as Markdown blocks
+% Text remains for now as Markdown
 
@@
 \begin{markdown}
-The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`), with additional descriptions using numbers $\mathbb{R}$ (notation `13.37`) and literal strings $\mathbb{S}$ (notation `"Hello"`).
+The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`), with additional descriptions using numbers $\mathbb{R}$ (notation $13.37$) and literal strings $\mathbb{S}$ (notation $\text{“Hello”}$).
 
@@
 The below is an example RO-Crate represented using the above formalization, assuming a base URI of `<http://example.com/ro/123/>`:
+\end{markdown}
 
-    ROCrate(<http://example.com/ro/123/>)
-    name(<http://example.com/ro/123/, 
-        “Data files associated with the manuscript:Effects of …”)
-    description(<http://example.com/ro/123/, 
-        “Palliative care planning for nursing home residents …")
-    datePublished(<http://example.com/ro/123/>, “2017")
-    license(<http://example.com/ro/123/>, 
-        <https://creativecommons.org/licenses/by-nc-sa/3.0/au/>
-    ContextualEntity(<https://creativecommons.org/licenses/by-nc-sa/3.0/au/>)
-    name(<https://creativecommons.org/licenses/by-nc-sa/3.0/au/, 
-        “Attribution-NonCommercial-ShareAlike 3.0 Australia (CC BY-NC-SA 3.0 AU)”)
-
-    hasPart(<http://example.com/ro/123/>, <http://example.com/ro/123/file.txt>)
-    File(<http://example.com/ro/123/survey.csv>)
-    name(<http://example.com/ro/123/survey.csv>, “Survey of care providers”)
-    hasPart(<http://example.com/ro/123/>, <http://www.example.om/ro/123/folder/>)
-    Dataset(<http://example.com/ro/123/interviews/>)
-    name(<http://example.com/ro/123/interviews/>, 
-        “Audio recordings of care provider interviews”)
-\end{markdown}
+\allowdisplaybreaks
+\begin{eqnarray*}
+&& ROCrate(\texttt{<http://example.com/ro/123/>}) \\
+&& name(\texttt{<http://example.com/ro/123/>}, \\
+&& \ \ \ \ \ \text{“Data files associated with the manuscript:Effects of …”}) \\
+&& description(\texttt{<http://example.com/ro/123/}, \\
+&& \ \ \ \ \ \text{“Palliative care planning for nursing home residents …”}) \\
+&& license(\texttt{<http://example.com/ro/123/>}, \\ 
+&& \ \ \ \ \ \texttt{<https://creativecommons.org/licenses/by-nc-sa/3.0/au/>}) \\ 
+&& datePublished(\texttt{<http://example.com/ro/123/>}, \text{“2017-02-23”}) \\ 
+&& hasPart(\texttt{<http://example.com/ro/123/>}, 
+        \texttt{<http://example.com/ro/123/file.txt>}) \\
+&& hasPart(\texttt{<http://example.com/ro/123/>}, 
+        \texttt{<http://www.example.om/ro/123/folder/>}) \\
+\\
+&& ContextualEntity(\texttt{<https://creativecommons.org/licenses/by-nc-sa/3.0/au/>}) \\
+&& name(\texttt{<https://creativecommons.org/licenses/by-nc-sa/3.0/au/>},  \\
+&& \ \ \ \ \  \text{“Attribution-NonCommercial-ShareAlike 3.0 Australia (CC BY-NC-SA 3.0 AU)”}) \\
+\\
+&& File(\texttt{<http://example.com/ro/123/survey.csv>}) \\
+&& name(\texttt{<http://example.com/ro/123/survey.csv>}, 
+        \text{“Survey of care providers”}) \\
+\\
+&& Dataset(\texttt{<http://example.com/ro/123/interviews/>}) \\
+&& name(\texttt{<http://example.com/ro/123/interviews/>},  \\
+&& \ \ \ \ \  \text{“Audio recordings of care provider interviews”}) 
+\end{eqnarray*}
 

### 2021-05-14 latex/ro-crate.tex
Added Simone as co-author in latex
Sunday fun with tables in LaTeX
Fixed table
More LaTeX IsN'tiT
some blurb about DMPs
Reviews, credits, formal definitions in tex
definition lists..

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 % \usepackage{hyperref}
@@
 \graphicspath{../content/images/}
-\usepackage[citations,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
+\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
 \def\markdownOptionOutputDir{../output}
+
+% for table resizing
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 \author[C]{\inits{P.S.}\fnms{Peter} \snm{Sefton}\ead[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}},
+\author[E]{\inits{S.L.}\fnms{Simone} \snm{Leo}\ead[label=e5]{simone.leo@crs4.it}\ead[label=o5,orcid]{0000-0001-8271-5429}},
 \author[D]{RO-Crate Community},
-\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}},
 \author[B]{\inits{P.G.}\fnms{Paul} 
 \snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}}
+\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}},
 \address[A]{Computer Science, \orgname{The University of Manchester},
@@
 \cny{Australia}\printead[presep={\\}]{e3,o3}}
+\address[E]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA), \cny{Italy}\printead[presep={\\}]{e5,o5}}
 \address[D]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
@@
 \begin{abstract}
-Abstract text.
+\markdownInput{../content/01.abstract.md}
 \end{abstract}
@@
 \markdownInput{../content/10.background.md}
+
 \markdownInput{../content/20.rocrate.md}
@@
 \markdownInput{../content/23.community.md}
+
+\input{../content/31.tooling-table.tex}
+%\markdownInput{../content/31.tooling-table.md}
 \markdownInput{../content/30.tooling.md}
+
 \markdownInput{../content/40.inuse.md}
@@
 \markdownInput{../content/43.culturalheritage.md}
+\markdownInput{../content/44.dmp.md}
+
 \markdownInput{../content/50.relatedwork.md}
@@
 % TODO: Recode formal definition to LaTeX math mode?
-%\input{../content/80.formal-definition.tex}
+%\include{../content/80.formal-definition}
+
 \markdownInput{../content/89.community.md}

### 2021-05-17 latex/ro-crate.tex
More edits and authors

@@
 \author[C]{\inits{P.S.}\fnms{Peter} \snm{Sefton}\ead[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}},
+\author[I]{\inits{M.C.}\fnms{Mercè} \snm{Crosas}\ead[label=e9]{mcrosas@g.harvard.edu}\ead[label=o9,orcid]{0000-0003-1304-1939}},
+\author[G]{\inits{M.LR.}\fnms{Marco} \snm{La Rosa}\ead[label=e7]{m@lr.id.au}\ead[label=o7,orcid]{0000-}},
 \author[E]{\inits{S.L.}\fnms{Simone} \snm{Leo}\ead[label=e5]{simone.leo@crs4.it}\ead[label=o5,orcid]{0000-0001-8271-5429}},
+\author[F]{\inits{M.P.}\fnms{Marc} \snm{Portier}\ead[label=e6]{marc.portier@vliz.be}\ead[label=o6,orcid]{0000-0002-9648-6484}},
+\author[H]{\inits{D.G.}\fnms{Daniel} \snm{Garijo}\ead[label=e8]{dgarijo@fi.upm.es}\ead[label=o8,orcid]{0000-0003-0454-7145}},
+\author[I]{\inits{A.T.}\fnms{Ana} \snm{Trisovic}\ead[label=e9]{ana.trisovic@gmail.com}\ead[label=o9,orcid]{0000-0003-1991-0533}},
+
 \author[D]{RO-Crate Community},
@@
 \address[D]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
+
+\address[H]{Ontology Engineering Group, \orgname{Universidad Politécnica de Madrid},
+Madrid, \cny{Spain}\printead[presep={\\}]{e8,o8}}
+
+\address[I]{Institute for Quantitative Social Science, \orgname{Harvard University} Cambridge, MA,  \cny{USA}\printead[presep={\\}]{e9,o9}}
 

### 2021-05-17 latex/ro-crate.tex
Sync

@@
 \author[I]{\inits{M.C.}\fnms{Mercè} \snm{Crosas}\ead[label=e9]{mcrosas@g.harvard.edu}\ead[label=o9,orcid]{0000-0003-1304-1939}},
+\author[H]{\inits{D.G.}\fnms{Daniel} \snm{Garijo}\ead[label=e8]{dgarijo@fi.upm.es}\ead[label=o8,orcid]{0000-0003-0454-7145}},
 \author[G]{\inits{M.LR.}\fnms{Marco} \snm{La Rosa}\ead[label=e7]{m@lr.id.au}\ead[label=o7,orcid]{0000-}},
@@
 \author[F]{\inits{M.P.}\fnms{Marc} \snm{Portier}\ead[label=e6]{marc.portier@vliz.be}\ead[label=o6,orcid]{0000-0002-9648-6484}},
-\author[H]{\inits{D.G.}\fnms{Daniel} \snm{Garijo}\ead[label=e8]{dgarijo@fi.upm.es}\ead[label=o8,orcid]{0000-0003-0454-7145}},
 \author[I]{\inits{A.T.}\fnms{Ana} \snm{Trisovic}\ead[label=e9]{ana.trisovic@gmail.com}\ead[label=o9,orcid]{0000-0003-1991-0533}},
@@
 \author[B]{\inits{P.G.}\fnms{Paul} 
-\snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}}
-\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}},
+\snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}},
+\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}}
+
 \address[A]{Computer Science, \orgname{The University of Manchester},
@@
 \address[D]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
-
 \address[H]{Ontology Engineering Group, \orgname{Universidad Politécnica de Madrid},
 Madrid, \cny{Spain}\printead[presep={\\}]{e8,o8}}
+\address[I]{Institute for Quantitative Social Science, \orgname{Harvard University} Cambridge, MA,  \cny{USA}\printead[presep={\\}]{e9,o9}}
 
-\address[I]{Institute for Quantitative Social Science, \orgname{Harvard University} Cambridge, MA,  \cny{USA}\printead[presep={\\}]{e9,o9}}
+%% TODO: Fix this mess - it seems we need one \address per author
+%% as this address also print their individual email and orcid,
+%% meaning some organization will repeat.
 

### 2021-05-17 latex/ro-crate.tex
author list and affiliations

--- 2021-05-14 latex/ro-crate.tex
@@
 % Two or more authors:
-\author[A,B]{\fnms{Stian} \snm{Soiland-Reyes}\ead[label=e1]{soiland-reyes@manchester.ac.uk}\ead[label=o1,orcid]{0000-0001-9842-9718}%
-\thanks{Corresponding author. \printead{e1}.}},
-\author[C]{\inits{P.S.}\fnms{Peter} \snm{Sefton}\ead[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}},
-\author[E]{\inits{S.L.}\fnms{Simone} \snm{Leo}\ead[label=e5]{simone.leo@crs4.it}\ead[label=o5,orcid]{0000-0001-8271-5429}},
-\author[D]{RO-Crate Community},
-\author[B]{\inits{P.G.}\fnms{Paul} 
-\snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}}
-\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}},
-\address[A]{Computer Science, \orgname{The University of Manchester},
-\cny{UK}\printead[presep={\\}]{e1,o1}}
-\address[B]{Informatics Institute, \orgname{University of Amsterdam},
-\cny{The Netherlands}\printead[presep={\\}]{e2,o2}}
-\address[C]{Faculty of Science, \orgname{University Technology Sydney},
-\cny{Australia}\printead[presep={\\}]{e3,o3}}
-\address[E]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA), \cny{Italy}\printead[presep={\\}]{e5,o5}}
-\address[D]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
+\author[SSR1,SSR2]{\inits{S.SR.}\fnms{Stian}
+    \snm{Soiland-Reyes}\ead[label=eSSR]{soiland-reyes@manchester.ac.uk}\ead[label=oSSR,orcid]{0000-0001-9842-9718}%
+    \thanks{Corresponding author. \printead{eSSR}.}}, 
+\author[PS]{\inits{P.S.}\fnms{Peter}
+    \snm{Sefton}\ead[label=ePS]{Peter.Sefton@uts.edu.au}\ead[label=oPS,orcid]{0000-0002-3545-944X}},
+\author[MC]{\inits{M.C.}\fnms{Mercè}
+    \snm{Crosas}\ead[label=eMC]{mcrosas@g.harvard.edu}\ead[label=oMC,orcid]{0000-0003-1304-1939}},
+\author[JMF]{\inits{JM.F.}\fnms{José María}
+    \snm{Fernández}\ead[label=eJMF]{jose.m.fernandez@bsc.es}\ead[label=oJMF,orcid]{0000-0002-4806-5140}},
+\author[DG]{\inits{D.G.}\fnms{Daniel}
+    \snm{Garijo}\ead[label=eDG]{dgarijo@fi.upm.es}\ead[label=oDG,orcid]{0000-0003-0454-7145}},
+\author[MLR]{\inits{M.LR.}\fnms{Marco}
+    \snm{La Rosa}\ead[label=eMLR]{m@lr.id.au}},
+\author[SL]{\inits{S.L.}\fnms{Simone}
+    \snm{Leo}\ead[label=eSL]{simone.leo@crs4.it}\ead[label=oSL,orcid]{0000-0001-8271-5429}},
+\author[MP]{\inits{M.P.}\fnms{Marc}
+    \snm{Portier}\ead[label=eMP]{marc.portier@vliz.be}\ead[label=oMP,orcid]{0000-0002-9648-6484}},
+\author[AT]{\inits{A.T.}\fnms{Ana}
+    \snm{Trisovic}\ead[label=eMP]{ana.trisovic@gmail.com}\ead[label=oAT,orcid]{0000-0003-1991-0533}},
+\author[RO]{RO-Crate Community},
+\author[PG]{\inits{P.G.}\fnms{Paul}
+    \snm{Groth}\ead[label=ePG]{p.t.groth@uva.nl}\ead[label=oPG,orcid]{0000-0003-0183-6910}},
+\author[CG]{\inits{C.G.}\fnms{Carole}
+    \snm{Goble}\ead[label=eCG]{carole.goble@manchester.ac.uk}\ead[label=oCG,orcid]{0000-0003-1219-2137}}
+\address[SSR1]{Computer Science, 
+    \orgname{The University of Manchester}, \cny{UK}}
+\address[SSR2]{Informatics Institute, \orgname{University of Amsterdam},
+    \cny{The Netherlands}\printead[presep={\\}]{eSSR,oSSR}}
+\address[PS]{Faculty of Science, \orgname{University Technology Sydney},
+    \cny{Australia}\printead[presep={\\}]{ePS,oPS}}
+\address[MC]{Institute for Quantitative Social Science, 
+    \orgname{Harvard University}, Cambridge, MA,  \cny{USA}\printead[presep={\\}]{oMC}}
+\address[JMF]{\orgname{Barcelona Supercomputing Center}, 
+    Barcelona \cny{Spain}\printead[presep={\\}]{oJMF}}
+\address[DG]{Ontology Engineering Group, 
+    \orgname{Universidad Politécnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{oDG}}
+\address[MLR]{\printead{eMLR}}
+\address[SL]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA),
+    \cny{Italy}\printead[presep={\\}]{oSL}}
+\address[MP]{\orgname{Vlaams Instituut voor de Zee}, Oostende,
+    \cny{Belgium}\printead[presep={\\}]{oMP}}
+\address[AT]{Institute for Quantitative Social Science,
+    \orgname{Harvard University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{oAT}}
+\address[RO]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
+\address[PG]{Informatics Institute, \orgname{University of Amsterdam},
+    \cny{The Netherlands}\printead[presep={\\}]{ePG,oPG}}
+\address[CG]{Computer Science, \orgname{The University of Manchester},
+    \cny{UK}\printead[presep={\\}]{eCG,oCG}}
+
+%% TODO: Fix this LaTeX mess - it seems we need one \address per author
+%% as this address also print their individual email and orcid,
+%% meaning some organization will repeat.
 

### 2021-05-18 build/build.sh
copy ro-crate* in build

@@ -138,4 +138,8 @@ if [ "${SPELLCHECK:-}" = "true" ]; then
   rm output/expanded-spelling-errors.txt
 fi
 
+# Copy RO-Crate metadata
+# TODO: Add date, commit etc.
+cp ro-crate/ro-crate-metadata.json ro-crate/ro-crate-preview.html output/
+
 echo >&2 "Build complete"

### 2021-05-19 content/31.tooling-table.tex
Changes from Nick and others
Reviews, credits, formal definitions in tex
Updates from Overleaf

@@
+% Please update in 31.tooling-table.md first!
 \begin{table}[htbp]
@@
 
-		\textbf{Life Monitor} \citep{about-lifemonitor} & Workflow developers & Python & Beta \\
+		\textbf{Life Monitor} \citep{about-lifemonitor} & Workflow developers & Python & Alpha \\
 		\multicolumn{4}{l}{\it Workflow testing and monitoring service; Workflow Testing profile of RO-Crate} \\
@@
 
+        \textbf{CheckMyCrate} \citep{CheckMyCrate} & Developers & Python (CLI) & Alpha \\
+        \multicolumn{4}{l}{\it Validation according to Workflow RO-Crate profile} \\
+
 		\bottomrule

### 2021-05-19 latex/ro-crate.tex
Promoted contributors to authors

@@
 %\pretitle{}
-\title{Packaging research artifacts with RO-Crate}
-\runtitle{Packaging research artifacts with RO-Crate}
+\title{Packaging research artefacts with RO-Crate}
+\runtitle{Packaging research artefacts with RO-Crate}
 %\subtitle{}
@@
     \snm{Sefton}\ead[label=ePS]{Peter.Sefton@uts.edu.au}\ead[label=oPS,orcid]{0000-0002-3545-944X}},
+\author[EOC]{\inits{E.OC.}\fnms{Eoghan}
+    \snm{Ó Carragáin}\ead[label=eEOC]{eoghan.ocarragain@ucc.ie}\ead[label=oEOC,orcid]{0000-0001-8131-2150}},
 \author[MC]{\inits{M.C.}\fnms{Mercè}
     \snm{Crosas}\ead[label=eMC]{mcrosas@g.harvard.edu}\ead[label=oMC,orcid]{0000-0003-1304-1939}},
-\author[JMF]{\inits{JM.F.}\fnms{José María}
+\author[LJC]{\inits{LJ.C.}\fnms{Leyla Jael}
+    \snm{Castro}\ead[label=eLJC]{ljgarcia@zbmed.de}\ead[label=oLJC,orcid]{0000-0003-3986-0510}},
+\author[FC]{\inits{F.C.}\fnms{Frederik}
+    \snm{Coppens}\ead[label=eFC]{frcop@psb.vib-ugent.be}\ead[label=oJMF,orcid]{0000-0001-6565-5145}},
+\author[JMF]{\inits{JM.F.}\fnms{José M.}
     \snm{Fernández}\ead[label=eJMF]{jose.m.fernandez@bsc.es}\ead[label=oJMF,orcid]{0000-0002-4806-5140}},

### 2021-05-19 latex/ro-crate.tex
Figures
author affiliations

@@
 % Two or more authors:
-\author[SSR1,SSR2]{\inits{S.SR.}\fnms{Stian}
+\author[SSR1,SSR2]{\inits{S.}\fnms{Stian}
     \snm{Soiland-Reyes}\ead[label=eSSR]{soiland-reyes@manchester.ac.uk}\ead[label=oSSR,orcid]{0000-0001-9842-9718}%
     \thanks{Corresponding author. \printead{eSSR}.}}, 
-\author[PS]{\inits{P.S.}\fnms{Peter}
+\author[PS]{\inits{P.}\fnms{Peter}
     \snm{Sefton}\ead[label=ePS]{Peter.Sefton@uts.edu.au}\ead[label=oPS,orcid]{0000-0002-3545-944X}},
-\author[EOC]{\inits{E.OC.}\fnms{Eoghan}
+\author[EOC]{\inits{E.}\fnms{Eoghan}
     \snm{Ó Carragáin}\ead[label=eEOC]{eoghan.ocarragain@ucc.ie}\ead[label=oEOC,orcid]{0000-0001-8131-2150}},
-\author[MC]{\inits{M.C.}\fnms{Mercè}
+\author[MC]{\inits{M.}\fnms{Mercè}
     \snm{Crosas}\ead[label=eMC]{mcrosas@g.harvard.edu}\ead[label=oMC,orcid]{0000-0003-1304-1939}},
-\author[LJC]{\inits{LJ.C.}\fnms{Leyla Jael}
+\author[LJC]{\inits{LJ.}\fnms{Leyla Jael}
     \snm{Castro}\ead[label=eLJC]{ljgarcia@zbmed.de}\ead[label=oLJC,orcid]{0000-0003-3986-0510}},
-\author[FC]{\inits{F.C.}\fnms{Frederik}
+\author[FC]{\inits{F.}\fnms{Frederik}
     \snm{Coppens}\ead[label=eFC]{frcop@psb.vib-ugent.be}\ead[label=oJMF,orcid]{0000-0001-6565-5145}},
-\author[JMF]{\inits{JM.F.}\fnms{José M.}
+\author[JMF]{\inits{JM.}\fnms{José M.}
     \snm{Fernández}\ead[label=eJMF]{jose.m.fernandez@bsc.es}\ead[label=oJMF,orcid]{0000-0002-4806-5140}},
-\author[DG]{\inits{D.G.}\fnms{Daniel}
+\author[DG]{\inits{D.}\fnms{Daniel}
     \snm{Garijo}\ead[label=eDG]{dgarijo@fi.upm.es}\ead[label=oDG,orcid]{0000-0003-0454-7145}},
-\author[MLR]{\inits{M.LR.}\fnms{Marco}
+\author[MLR]{\inits{M.}\fnms{Marco}
     \snm{La Rosa}\ead[label=eMLR]{m@lr.id.au}},
-\author[SL]{\inits{S.L.}\fnms{Simone}
+\author[SL]{\inits{S.}\fnms{Simone}
     \snm{Leo}\ead[label=eSL]{simone.leo@crs4.it}\ead[label=oSL,orcid]{0000-0001-8271-5429}},
-\author[MP]{\inits{M.P.}\fnms{Marc}
+\author[MP]{\inits{M.}\fnms{Marc}
     \snm{Portier}\ead[label=eMP]{marc.portier@vliz.be}\ead[label=oMP,orcid]{0000-0002-9648-6484}},
-\author[AT]{\inits{A.T.}\fnms{Ana}
+\author[AT]{\inits{A.}\fnms{Ana}
     \snm{Trisovic}\ead[label=eMP]{ana.trisovic@gmail.com}\ead[label=oAT,orcid]{0000-0003-1991-0533}},
 \author[RO]{RO-Crate Community},
-\author[PG]{\inits{P.G.}\fnms{Paul}
+\author[PG]{\inits{P.}\fnms{Paul}
     \snm{Groth}\ead[label=ePG]{p.t.groth@uva.nl}\ead[label=oPG,orcid]{0000-0003-0183-6910}},
-\author[CG]{\inits{C.G.}\fnms{Carole}
+\author[CG]{\inits{C.}\fnms{Carole}
     \snm{Goble}\ead[label=eCG]{carole.goble@manchester.ac.uk}\ead[label=oCG,orcid]{0000-0003-1219-2137}}
@@
     \cny{Australia}\printead[presep={\\}]{ePS,oPS}}
+\address[EOC]{\orgname{University College Cork},
+    \cny{Ireland}\printead[presep={\\}]{oEOC}}
 \address[MC]{Institute for Quantitative Social Science, 
     \orgname{Harvard University}, Cambridge, MA,  \cny{USA}\printead[presep={\\}]{oMC}}
+\address[LJC]{\orgname{ZB MED Information Centre for Life Sciences}, 
+    Köln, \cny{Germany}\printead[presep={\\}]{oLJC}}
+\address[FC]{\orgname{VIB-UGent Center for Plant Systems Biology}, 
+    Gent, \cny{Belgium}\printead[presep={\\}]{oFC}}
 \address[JMF]{\orgname{Barcelona Supercomputing Center}, 
-    Barcelona \cny{Spain}\printead[presep={\\}]{oJMF}}
+    Barcelona, \cny{Spain}\printead[presep={\\}]{oJMF}}
 \address[DG]{Ontology Engineering Group, 
@@
     \cny{UK}\printead[presep={\\}]{eCG,oCG}}
-
-%% TODO: Fix this LaTeX mess - it seems we need one \address per author
-%% as this address also print their individual email and orcid,
-%% meaning some organization will repeat.
 

### 2021-05-20 latex/ro-crate.tex
edits

@@
 \author[FC]{\inits{F.}\fnms{Frederik}
-    \snm{Coppens}\ead[label=eFC]{frcop@psb.vib-ugent.be}\ead[label=oJMF,orcid]{0000-0001-6565-5145}},
+    \snm{Coppens}\ead[label=eFC]{frcop@psb.vib-ugent.be}\ead[label=oFC,orcid]{0000-0001-6565-5145}},
 \author[JMF]{\inits{JM.}\fnms{José M.}
@@
     \snm{Goble}\ead[label=eCG]{carole.goble@manchester.ac.uk}\ead[label=oCG,orcid]{0000-0003-1219-2137}}
+
 \address[SSR1]{Computer Science, 
@@
 \address[LJC]{\orgname{ZB MED Information Centre for Life Sciences}, 
-    Köln, \cny{Germany}\printead[presep={\\}]{oLJC}}
+    Cologne, \cny{Germany}\printead[presep={\\}]{oLJC}}
 \address[FC]{\orgname{VIB-UGent Center for Plant Systems Biology}, 
@@
 
+\newpage % Because the author affiliations are too long!
 \begin{abstract}

### 2021-05-21 content/80.formal-definition.tex
more LaTeX in 80.formal-definition.tex
it does not seem to work with markdown sections though

@@
-%% TODO: Convert to LaTex math mode. See GitHub rendering for original Unicode symbols
-% https://github.com/stain/ro-crate-paper/blob/master/content/80.formal-definition.md
+% Because this file is added with \include, do not use
+% Markdown setions within.
 
-% Text remains for now as Markdown
-
-\begin{markdown}
-## Formalizing RO-Crate in First Order Logic
+\section{Formalizing RO-Crate in First Order Logic}
 
@@
 
-### Language
-
-\end{markdown}
+\subsection{Language}
 
@@
 \begin{eqnarray*}
-    \mathcal{L}_{rocrate}   & \equiv & { Property(p), Class(c), 
-                            Literal(x), \mathbb{R}, \mathbb{S} } \\
+    \mathcal{L}_{rocrate}   & \equiv & \big\{ Property(p), Class(c), 
+                            Value(x), \mathbb{R}, \mathbb{S} \big\} \\
     \mathbb{D}              & \equiv & \mathbb{IRI} \\
@@
 \begin{markdown}
-The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`), with additional descriptions using numbers $\mathbb{R}$ (notation $13.37$) and literal strings $\mathbb{S}$ (notation $\text{“Hello”}$).
+The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`)[^3], with additional descriptions using numbers $\mathbb{R}$ (notation $13.37$) and literal strings $\mathbb{S}$ (notation $\text{“Hello”}$).
 
 From this formalised language $\mathcal{L}_{rocrate}$ we can interpret an RO-Crate in any representation that can gather these descriptions, their properties, classes, and literal attributes.
+
+[^3]:
+    For simplicity, blank nodes are not included in this formalisation, as RO-Crate
+    recommends the use of IRI identifiers: \url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#describing-entities-in-json-ld} 
 \end{markdown}
@@
 
-Below we use $\mathcal{L}_{rocrate}$ to define a minimal RO-Crate:
+Below we use $\mathcal{L}_{rocrate}$ to define a minimal[^4] RO-Crate:
+
+[^4]: The full list of types, relations and attribute properties from the RO-Crate specification are not included. Examples shown include $datePublished$, $CreativeWork$ and $name$.
 \end{markdown}
@@
                                             & & Mentions(R, c) \land ContextualEntity(c) \\
-\forall r Root(r)                           & \Rightarrow & Dataset(r) \land name(r, n) \land description(r, d) \land \\
+\forall r \ Root(r)                         & \Rightarrow & Dataset(r) \land name(r, n) \land description(r, d) \land \\
                                             & &             published(r, date) \land license(e, l) \\
-\forall e \forall n \ name(e, n)            & \Rightarrow & Literal(n) \\
-\forall e \forall s \ description(e, s)     & \Rightarrow & Literal(s) \\
-\forall e \forall d \ datePublished(e, d)   & \Rightarrow & Literal(d) \\
+\forall e \forall n \ name(e, n)            & \Rightarrow & Value(n) \\
+\forall e \forall s \ description(e, s)     & \Rightarrow & Value(s) \\
+\forall e \forall d \ datePublished(e, d)   & \Rightarrow & Value(d) \\
 \forall e \forall l \ license(e, l)         & \Rightarrow & ContextualEntity(l) \\
@@
 Relation(s, p, o)                           & \models &     Entity(s) \land Property(p) \land  Entity(o) \\
-Attribute(s, p, x)                          & \models &     Entity(s) \land Property(p) \land Literal(x) \\
-Literal(x)                                  & \equiv &      x \in \mathbb{R} \oplus x \in \mathbb{S}
+Attribute(s, p, v)                          & \models &     Entity(s) \land Property(p) \land Value(v) \\
+Value(v)                                    & \equiv &      v \in \mathbb{R} \oplus v \in \mathbb{S}
 \end{eqnarray*}
@@
 $R$ is used by $Mentions(R, e)$ to indicate that $e$ is an Entity described by the RO-Crate and therefore its metadata (a set of $Relation$ and $Attribute$ predicates) form part of the RO-Crate serialization.
-$Relation(s, p, o)$ and $Attribute(s, p, x)$ are defined as a _subject-predicate-object_ triple pattern from an $Entity(s)$ using a $Property(p)$ to either another $Entity(o)$ or a $Literal(x)$ value.
+$Relation(s, p, o)$ and $Attribute(s, p, x)$ are defined as a _subject-predicate-object_ triple pattern from an $Entity(s)$ using a $Property(p)$ to either another $Entity(o)$ or a $Value(x)$ value.
 \end{markdown}
@@
 In reality many additional attributes from schema.org types like <http://schema.org/Dataset> and <http://schema.org/CreativeWork> would be used to further describe the RO-Crate and its entities, but as these are optional they do not form part of this formalization.
+
+### Mapping to RDF with schema.org
+
+A formalized RO-Crate in 
+$\mathcal{L}_{rocrate}$
+can be mapped to different serializations.
+
+Assume a simplified[^4] language 
+$\mathcal{L}_{RDF}$ 
+based on the [RDF abstract syntax](http://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/):
+
+[^4]: This simplification does not cover the extensive list of literal datatypes built-in to RDF 1.1, only strings and decimal real numbers. Likewise, language of literals are not included.
+
 \end{markdown}
 
-\begin{markdown}
-### Mapping to RDF with schema.org
+\begin{eqnarray*}
+\mathcal{L}_{RDF}           & \equiv &      \big\{ Triple(s,p,o), IRI(i), BlankNode(b), Literal(s), 
+    \mathbb{IRI}, \mathbb{S}, \mathbb{R}    \big\} \\
+\mathbb{D}_{RDF}            & \equiv &      \mathbb{S} \\
+\forall i \ IRI(i)          & \Rightarrow & i \in \mathbb{IRI} \\
+\forall s \ BlankNode(s)    & \Rightarrow & s \in \mathbb{S} \\
+\forall s \forall p \forall o \ 
+    Triple(s,p,o)           & \Rightarrow & \Big( IRI(s) \lor BlankNode(s) \Big) \land  \\
+                            & &             IRI(p) \land  \\
+                            & &             \Big(IRI(o) \lor BlankNode(o) \lor Literal(o) \Big) \\
+Literal(v)                  & \models &     Value(v) \land Datatype(v,t) \land IRI(t) \\
+\forall v \ Value(v)        & \Rightarrow & v \in \mathbb{S} \\
+LanguageTag(v, l)           & \equiv &      Datatype\big(v, \\
+    && \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns#langString>}\big)
+\end{eqnarray*}
 
-A formalized RO-Crate can be mapped to different serializations. Below follows a mapping to RDF using schema.org.
+Below follows a mapping from $\mathcal{L}_{rocrate}$ to $\mathcal{L}_{RDF}$ using schema.org as vocabulary:
 
-    Dataset(d) →  type(d, <http://schema.org/Dataset>)
-    File(f) →  type(f, <http://schema.org/MediaObject>)
-    Property(p) →  type(p, <http://www.w3.org/2000/01/rdf-schema#Property>)
-    Class(c) →  type(c, <http://www.w3.org/2000/01/rdf-schema#Class>)
-    CreativeWork(e) →  ContextualEntity(e) ∧ type(e, <http://schema.org/CreativeWork>)
-    hasPart(e, t) →  Relation(e, <http://schema.org/hasPart>, t)
-    type(e, t) →  Relation(e, <http://www.w3.org/1999/02/22-rdf-syntax-ns#type>, t) ∧ Class(t)
-    name(e, n) →  Attribute(e, <http://schema.org/name>, n)
-    description(e, d) →  Attribute(e, <http://schema.org/description>, d)
-    datePublished(e, date) →  Attribute(e, <http://schema.org/datePublished>, date)
-    license(e, l) →  Relation(e, <http://schema.org/license>, l) ∧ CreativeWork(l)
-
-Note that in the JSON-LD serialization of RO-Crate the expression of $Class$ and $Property $is typically indirect, as the JSON-LD $@context$ maps to schema.org IRIs, which when resolved as Linked Data embeds their formal definition as RDFa.
-\end{markdown}
+\begin{eqnarray*}
+Property(p)         & \Rightarrow &     type(p, \texttt{<http://www.w3.org/2000/01/rdf-schema#Property>})   \\
+Class(c)            & \Rightarrow &     type(c, \texttt{<http://www.w3.org/2000/01/rdf-schema#Class>})  \\
+Dataset(d)          & \Rightarrow &     type(d, \texttt{<http://schema.org/Dataset>})   \\
+File(f)             & \Rightarrow &     type(f, \texttt{<http://schema.org/MediaObject>})   \\
+CreativeWork(e)     & \Rightarrow &     ContextualEntity(e) \land type(e, \texttt{<http://schema.org/CreativeWork>})  \\
+hasPart(e, t)       & \Rightarrow &     Relation(e, \texttt{<http://schema.org/hasPart>}, t)    \\
+name(e, n)          & \Rightarrow &     Attribute(e, \texttt{<http://schema.org/name>}, n)  \\
+description(e, s)   & \Rightarrow &     Attribute(e, \texttt{<http://schema.org/description>}, s)   \\
+datePublished(e, d) & \Rightarrow &     Attribute(e, \texttt{<http://schema.org/datePublished>}, d) \\
+license(e, l)       & \Rightarrow &     Relation(e, \texttt{<http://schema.org/license>}, l) \land CreativeWork(l) \\
+type(e, t)          & \Rightarrow &     Relation(e, \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns#type>}, t) \\
+                    & & \land Class(t)   \\
+String(s)           & \equiv &          Value(s) \land  s \in \mathbb{S} \\
+String(s)           & \Rightarrow &     Datatype(s, \texttt{<http://www.w3.org/2001/XMLSchema#string>}) \\
+Decimal(d)          & \equiv &          Value(d) \land  d \in \mathbb{R} \\
+Decimal(d)          & \Rightarrow &     Datatype(d, \texttt{<http://www.w3.org/2001/XMLSchema#decimal>}) \\
+Relation(s,p,o)     & \Rightarrow &     Triple(s,p,o) \land IRI(s) \land IRI(o) \\
+Attribute(s,p,o)    & \Rightarrow &     Triple(s,p,o) \land IRI(s) \land Literal(o) \\
+\end{eqnarray*}
 
 \begin{markdown}
+Note that in the JSON-LD serialization of RO-Crate, the expression of $Class$ and $Property$ is typically indirect: The JSON-LD `@context` maps to schema.org IRIs, which, when resolved as Linked Data, embed their formal definition as RDFa.
 
@@
 Therefore the serialization of the RO-Crate into a file should also describe itself in a [Metadata File Descriptor](https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#ro-crate-metadata-file-descriptor), indicating it is about (describing) the RO-Crate root data entity, and that it conformsTo a particular version of the RO-Crate specification:
+\end{markdown}
 
-    about(s,o) →  Relation(s, <http://schema.org/about>, o)
-    conformsTo(s,o) →  Relation(s, <http://purl.org/dc/terms/conformsTo>, R)
-    MetadataFileDescriptor(m) →  ( CreativeWork(m) ∧ about(m,R) ∧ RO-Crate(R) ∧ 
-        conformsTo(m, <https://w3id.org/ro/crate/1.1>) )
+\begin{eqnarray*}
+about(s,o)      & \Rightarrow & Relation(s, \texttt{<http://schema.org/about>}, o)   \\
+conformsTo(s,o) & \Rightarrow & Relation(s, \texttt{<http://purl.org/dc/terms/conformsTo>}, o)   \\
+MetadataFile(m) & \Rightarrow & CreativeWork(m) \land about(m,R) ∧ ROCrate(R) \land    \\
+                &             & conformsTo(m, \texttt{<https://w3id.org/ro/crate/1.1>})
+\end{eqnarray*}
 
+\begin{markdown}
 Note that although the metadata file necessarily is an _information resource_ written to disk or served over the network (e.g.
@@
 
-    Mentions(R, s) ∧ Relation(s, p, o) →  Mentions(R, o)
-    i ∈ $\mathbb{Iri}$ → "i"
-    r ∈ ℝ →  r
-    s ∈ $\mathbb{S}$ → "s"
-    ∀s∀p∀o Relation(s,p,o) →  { "@id": s,
-                                p: { "@id": o }
-                              }     
-    ∀s∀p∀v Attribute(s,p,v) →  { "@id": s,
-                                p: v 
-                               }
-    ∀r∀c  RO-Crate(r) →  { "@graph": [ Mentions(r, c)* ] }
-    R ⊨  <./>
-    MetadataFileDescriptor(<ro-crate-metadata.json>)
+[^2]:
+    **Limitations:** Contextual entities not related from the RO-Crate (e.g.
+using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see [issue 122](https://github.com/ResearchObject/ro-crate/issues/122).
+The $datePublished(e, date)$ rule do not include syntax checks for the ISO 8601 datetime format.
+Compared with RO-Crate examples, this generated JSON-LD does not use a $@context$ as the IRIs are produced unshortened, a post-step could do JSON-LD Flattening with a versioned RO-Crate context.
+The `@type` expansion is included for clarity, even though this is also implied by the $type(e, t)$ expansion to $Relation(e, \texttt{xsd:type})$.
+\end{markdown}
 
+\begin{eqnarray*}
+Mentions(R, s) \land Relation(s,p,o)  
+                        & \Rightarrow & Mentions(R, o) \\
+IRI(i)                  & \Rightarrow & \texttt{"} i \texttt{"} \\
+Decimal(d)              & \Rightarrow & d \\
+String(s)               & \Rightarrow & \texttt{"} s \texttt{"} \\
+\forall e \forall t
+\ type(e, t)            & \Rightarrow & \texttt{\{"@id":}  e \texttt{,} \\
+&&                               \ \  \texttt{"@type":} t \\
+&&                              \texttt{\}} \\
+\forall s \forall p \forall o 
+\ Relation(s,p,o)  
+                        & \Rightarrow &  \texttt{\{"@id":}  s \texttt{,} \\
+&&                               \ \  p \texttt{: \{ "@id":} o \texttt{\}} \\
+&&                              \texttt{\}} \\
+\forall s \forall p \forall v
+\ Attribute(s,p,v)    & \Rightarrow &  \texttt{\{"@id":} s \texttt{,} \\
+&&                               \ \ p \texttt{:} v  \\
+&&                               \texttt{\}} \\ 
+\forall r  \forall c 
+    ROCrate(r)      & \Rightarrow &  \texttt{\{ "@graph": [} \\
+&& \ \ \ \ Mentions(r, c)* \\
+&& \ \ \ \texttt{]} \\
+&& \texttt{\}} \\
+R   & \equiv & \texttt{<./>}  \\
+R   & \Rightarrow &  MetadataFile(\texttt{<ro-crate-metadata.json>}) \\
+\end{eqnarray*}
+
+\begin{markdown}
 This exposes the first order logic domain of discourse of IRIs, with rational numbers and strings as their corresponding JSON-LD representation.
 These production rules first grow the graph of $R$ by adding a transitive rule – anything described in $R$ which is related to $o$, means that $o$ is also mentioned by the $ROCrate(R)$.
-For simplicity this rule is one-way; in practice the JSON-LD graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities.
-
-[^2]:
-    Limitations: The full list of types, relations and attribute properties from the RO-Crate specification are not included.
-Examples shown include $datePublished$, $CreativeWork$ and $name$.
-Contextual entities not related from the RO-Crate (e.g.
-using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see [issue 122](https://github.com/ResearchObject/ro-crate/issues/122).
-The $datePublished(e, date)$ rule do not include syntax checks for the ISO 8601 datetime format.
-Compared with RO-Crate examples, this generated JSON-LD does not use a $@context$ as the IRIs are produced unshortened; a post-step could do JSON-LD Flattening with a versioned RO-Crate context.
+For simplicity this rule is one-way; in practice the graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities.
 \end{markdown}

### 2021-05-23 content/80.formal-definition.tex
where have all the DOIs gone..

@@
 \begin{eqnarray*}
-    \mathcal{L}_{rocrate}   & \equiv & \big\{ Property(p), Class(c), 
+    \mathcal{L}_{rocrate}   & = & \big\{ Property(p), Class(c), 
                             Value(x), \mathbb{R}, \mathbb{S} \big\} \\
-    \mathbb{D}              & \equiv & \mathbb{IRI} \\
+    \mathbb{D}              & = & \mathbb{IRI} \\
     \mathbb{IRI}            & \equiv & { \text{IRIs as defined in RFC3987} } \\
@@
 
-\begin{markdown}
-The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers [@doi:10.17487/rfc3987] (notation `<http://example.com/>`)[^3], with additional descriptions using numbers $\mathbb{R}$ (notation $13.37$) and literal strings $\mathbb{S}$ (notation $\text{“Hello”}$).
+The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers \cite{doi:10.17487/rfc3987} (notation $\texttt{<http://example.com/>}$)\footnote{
+    For simplicity, blank nodes are not included in this formalisation, as RO-Crate
+    recommends the use of IRI identifiers: \url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#describing-entities-in-json-ld} 
+}, with additional descriptions using numbers $\mathbb{R}$ (notation $13.37$) and literal strings $\mathbb{S}$ (notation $\text{“Hello”}$).
 
@@
 
-[^3]:
-    For simplicity, blank nodes are not included in this formalisation, as RO-Crate
-    recommends the use of IRI identifiers: \url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#describing-entities-in-json-ld} 
-\end{markdown}
+\subsection{Minimal RO-Crate}
 
-\begin{markdown}
-### Minimal RO-Crate
-
-Below we use $\mathcal{L}_{rocrate}$ to define a minimal[^4] RO-Crate:
-
-[^4]: The full list of types, relations and attribute properties from the RO-Crate specification are not included. Examples shown include $datePublished$, $CreativeWork$ and $name$.
-\end{markdown}
+Below we use $\mathcal{L}_{rocrate}$ to define a minimal\footnote{
+    The full list of types, relations and attribute properties from the RO-Crate specification are not included. Examples shown include $datePublished$, $CreativeWork$ and $name$.
+} RO-Crate:
 
@@
 
-\begin{markdown}
-An $ROCrate(R)$ is defined as a self-described _Root Data Entity_, which describes and contains parts (_data entities_), which are further described in _contextual entities_.
-These terms align with their use in the [RO-Crate 1.1 terminology](https://www.researchobject.org/ro-crate/1.1/terminology).
+An $ROCrate(R)$ is defined as a self-described \emph{Root Data Entity}, which describes and contains parts (\emph{data entities}), which are further described in \emph{contextual entities}.
+These terms align with their use in the \href{https://www.researchobject.org/ro-crate/1.1/terminology}{RO-Crate 1.1 terminology}.
 
 The $Root(r)$ is a type of $Dataset(r)$, and must have the metadata to literal attributes to provide a $name$, $description$ and $datePublished$, as well as a contextual entity identifying its license.
-These predicates correspond to the RO-Crate 1.1 [requirements for the root data entity](https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#direct-properties-of-the-root-data-entity).
+These predicates correspond to the RO-Crate 1.1 \href{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#direct-properties-of-the-root-data-entity}{requirements for the root data entity}.
 
-The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or [both](https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities); and must be typed with at least one $Class(e)$.
+The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or \href{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities}{both}; and must be typed with at least one $Class(e)$.
 
@@
 $R$ is used by $Mentions(R, e)$ to indicate that $e$ is an Entity described by the RO-Crate and therefore its metadata (a set of $Relation$ and $Attribute$ predicates) form part of the RO-Crate serialization.
-$Relation(s, p, o)$ and $Attribute(s, p, x)$ are defined as a _subject-predicate-object_ triple pattern from an $Entity(s)$ using a $Property(p)$ to either another $Entity(o)$ or a $Value(x)$ value.
-\end{markdown}
+$Relation(s, p, o)$ and $Attribute(s, p, x)$ are defined as a \emph{subject—predicate—object} triple pattern from an $Entity(s)$ using a $Property(p)$ to either another $Entity(o)$ or a $Value(x)$ value.
 
-\begin{markdown}
-### Example of formalized RO-Crate
+\subsection{Example of formalized RO-Crate}
 
-The below is an example RO-Crate represented using the above formalization, assuming a base URI of `<http://example.com/ro/123/>`:
-\end{markdown}
+The below is an example RO-Crate represented using the above formalisation, assuming a base URI of \texttt{<http://example.com/ro/123/>}:
 
@@
 
-\begin{markdown}
-In reality many additional attributes from schema.org types like <http://schema.org/Dataset> and <http://schema.org/CreativeWork> would be used to further describe the RO-Crate and its entities, but as these are optional they do not form part of this formalization.
+In reality many additional attributes from schema.org types like \url{http://schema.org/Dataset} and \url{http://schema.org/CreativeWork} would be used to further describe the RO-Crate and its entities, but as these are optional they do not form part of this formalisation.
 
-### Mapping to RDF with schema.org
+\subsection{Mapping to RDF with schema.org}
 
-A formalized RO-Crate in 
-$\mathcal{L}_{rocrate}$
-can be mapped to different serializations.
-
-Assume a simplified[^4] language 
-$\mathcal{L}_{RDF}$ 
-based on the [RDF abstract syntax](http://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/):
-
-[^4]: This simplification does not cover the extensive list of literal datatypes built-in to RDF 1.1, only strings and decimal real numbers. Likewise, language of literals are not included.
-
-\end{markdown}
+A formalized RO-Crate in $\mathcal{L}_{rocrate}$ can be mapped to different serializations.
+Assume a simplified\footnote{
+ This simplification does not cover the extensive list of literal datatypes built-in to RDF 1.1, only strings and decimal real numbers. Likewise, language of literals are not included.
+} language $\mathcal{L}_{RDF}$ 
+based on the \href{http://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/}{RDF abstract syntax}:
 
@@
 
-\begin{markdown}
-Note that in the JSON-LD serialization of RO-Crate, the expression of $Class$ and $Property$ is typically indirect: The JSON-LD `@context` maps to schema.org IRIs, which, when resolved as Linked Data, embed their formal definition as RDFa.
+Note that in the JSON-LD serialization of RO-Crate, the expression of $Class$ and $Property$ is typically indirect: The JSON-LD \texttt{@context} maps to schema.org IRIs, which, when resolved as Linked Data, embed their formal definition as RDFa.
 
-### RO-Crate 1.1 Metadata File Descriptor
+\subsection{RO-Crate 1.1 Metadata File Descriptor}
 
-An important RO-Crate principle is that of being **self-describing**.
-Therefore the serialization of the RO-Crate into a file should also describe itself in a [Metadata File Descriptor](https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#ro-crate-metadata-file-descriptor), indicating it is about (describing) the RO-Crate root data entity, and that it conformsTo a particular version of the RO-Crate specification:
-\end{markdown}
+An important RO-Crate principle is that of being \textbf{self-describing} Therefore the serialisation of the RO-Crate into a file should also describe itself in a \href{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#ro-crate-metadata-file-descriptor}{Metadata File Descriptor}, indicating it is $about$ (describing) the RO-Crate root data entity, and that it $conformsTo$ a particular version of the RO-Crate specification:
 
@@
 
-\begin{markdown}
-Note that although the metadata file necessarily is an _information resource_ written to disk or served over the network (e.g.
-as JSON-LD), it is not considered to be a contained _part_ of the RO-Crate in the form of a _data entity_, rather it is described only as a _contextual entity_.
+Note that although the metadata file necessarily is an \emph{information resource} written to disk or served over the network (e.g.
+as JSON-LD), it is not considered to be a contained \emph{part} of the RO-Crate in the form of a \emph{data entity}, rather it is described only as a \emph{contextual entity}.
 
-While in the conceptual model the _RO-Crate Metadata File_ can be seen as the top-level node that describes the _RO-Crate Root_, in the formal model (and the JSON-LD format) the metadata file descriptor is an additional contextual entity and not affecting the depth-limit of the RO-Crate.
-\end{markdown}
+While in the conceptual model the \emph{RO-Crate Metadata File} can be seen as the top-level node that describes the \emph{RO-Crate Root}, in the formal model (and the JSON-LD format) the metadata file descriptor is an additional contextual entity and not affecting the depth-limit of the RO-Crate.
 
-\begin{markdown}
-### Forward-chained Production Rules for JSON-LD
+\subsection{Forward-chained Production Rules for JSON-LD}
 
-Combining the above predicates and schema.org mapping with rudimentary JSON templates, these forward-chaining production rules can output JSON-LD according to the RO-Crate 1.1 specification[^2]:
-
-[^2]:
-    **Limitations:** Contextual entities not related from the RO-Crate (e.g.
-using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see [issue 122](https://github.com/ResearchObject/ro-crate/issues/122).
+Combining the above predicates and schema.org mapping with rudimentary JSON templates, these forward-chaining production rules can output JSON-LD according to the RO-Crate 1.1 specification\footnote{
+    \textbf{Limitations:} Contextual entities not related from the RO-Crate (e.g.
+using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see \href{https://github.com/ResearchObject/ro-crate/issues/122}{issue 122}.
 The $datePublished(e, date)$ rule do not include syntax checks for the ISO 8601 datetime format.
 Compared with RO-Crate examples, this generated JSON-LD does not use a $@context$ as the IRIs are produced unshortened, a post-step could do JSON-LD Flattening with a versioned RO-Crate context.
-The `@type` expansion is included for clarity, even though this is also implied by the $type(e, t)$ expansion to $Relation(e, \texttt{xsd:type})$.
-\end{markdown}
+The \texttt{@type} expansion is included for clarity, even though this is also implied by the $type(e, t)$ expansion to $Relation(e, \texttt{xsd:type})$.
+}:
 
@@
 
-\begin{markdown}
 This exposes the first order logic domain of discourse of IRIs, with rational numbers and strings as their corresponding JSON-LD representation.
@@
 For simplicity this rule is one-way; in practice the graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities.
-\end{markdown}

### 2021-05-24 latex/ro-crate.tex
Updates from Overleaf
more LaTeX in 80.formal-definition.tex
it does not seem to work with markdown sections though
where have all the DOIs gone..
No need for Iosevko font anymore
ORCID for Marco

@@
 \usepackage{amssymb}
-% \usepackage{hyperref}
+\usepackage{hyperref}
+\usepackage{xcolor}
+\definecolor{darkerblue}{cmyk}{1,1,0,0.45}
+\hypersetup{
+    colorlinks=true,
+    linkcolor=black,
+    citecolor=black,
+    urlcolor=darkerblue}
+
 \usepackage{fontspec}
@@
 % \usepackage{graphicx}
+
+% Hyperlinks on DOI permalinks – just 20 years after DOI invention!
+% 
+% We had to make our own copy of 
+% https://ctan.org/pkg/doi?lang=en 
+% where commands are renamde to \doix
+% because iosart2x.cls and ios1.bst have 
+% a different use of \doiurl \doi etc.
+% We also set the doixtext prefix empty
+% to avoid doi:doi:
+\usepackage{doix}
+\renewcommand{\doixtext}{}
+\renewcommand{\doiurl}[1]{\doix{#1}}
 
@@
 % all; the logic symbols we need
-\setmonofont{iosevka-slab}[
-    Extension={.ttf},
-    Path=./ttf-iosevka-slab-6.1.3/,
-    Scale=MatchLowercase,
-    UprightFont={*-light},
-    ItalicFont={*-lightitalic}
-]
+%\setmonofont{iosevka-slab}[
+%    Extension={.ttf},
+%    Path=./ttf-iosevka-slab-6.1.3/,
+%    Scale=MatchLowercase,
+%    UprightFont={*-light},
+%    ItalicFont={*-lightitalic}
+%]
 
@@
 \author[MLR]{\inits{M.}\fnms{Marco}
-    \snm{La Rosa}\ead[label=eMLR]{m@lr.id.au}},
+    \snm{La Rosa}\ead[label=eMLR]{m@lr.id.au}\ead[label=oMLR,orcid]{0000-0001-5383-6993}},
 \author[SL]{\inits{S.}\fnms{Simone}
@@
     \orgname{Universidad Politécnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{oDG}}
-\address[MLR]{\printead{eMLR}}
+\address[MLR]{\printead{oMLR}}
 \address[SL]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA),
@@
 \markdownInput{../content/44.dmp.md}
+\markdownInput{../content/45.institutional-repos.md}
 
@@
 \appendix
-\markdownInput{../content/80.formal-definition.md}
-% TODO: Recode formal definition to LaTeX math mode?
-%\include{../content/80.formal-definition}
+%\markdownInput{../content/80.formal-definition.md}
+\input{../content/80.formal-definition}
 

### 2021-05-31 content/80.formal-definition.tex
80.formal-definition.tex  → 80.formal-definition.tmd

@@
 
-Below is an attempt to formalize the concept of RO-Crate as a set of relations using First Order Logic:
+Below is a formalization of the concept of RO-Crate as a set of relations using First Order Logic:
 
@@
 && hasPart(\texttt{<http://example.com/ro/123/>}, 
-        \texttt{<http://www.example.om/ro/123/folder/>}) \\
+        \texttt{<http://example.com/ro/123/interviews/>}) \\
 \\
@@
 \forall i \ IRI(i)          & \Rightarrow & i \in \mathbb{IRI} \\
-\forall s \ BlankNode(s)    & \Rightarrow & s \in \mathbb{S} \\
 \forall s \forall p \forall o \ 
@@
 using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see \href{https://github.com/ResearchObject/ro-crate/issues/122}{issue 122}.
-The $datePublished(e, date)$ rule do not include syntax checks for the ISO 8601 datetime format.
+The $datePublished(e, d)$ rule do not include syntax checks for the ISO 8601 datetime format.
 Compared with RO-Crate examples, this generated JSON-LD does not use a $@context$ as the IRIs are produced unshortened, a post-step could do JSON-LD Flattening with a versioned RO-Crate context.

### 2021-06-02 build/build.sh
more svg?

@@ -19,7 +19,8 @@ ln -s ../latex/ro-crate.bib content/manual-references-ro-crate.bib
 
 # Replace silly paths for LaTeX
 sed -i 's,../content/images/,images/,g' content/*md
-# .. and use SVG instead of PDF
+# .. and use PNG/SVG instead of PDF
+sed -i 's,ro-crate-bco-sep-of-concerns.pdf,ro-crate-bco-sep-of-concerns.png' content/42*md 
 sed -i 's,\.pdf,.svg,g' content/*md
 
 ## Horrible latex seds for the figure captions

### 2021-08-08 output/_markdown_ro-crate/1c64500a3160a6e96b6c36e70d7987b2.md.tex
Updates from Overleaf

\markdownRendererHeadingTwo{Conceptual Definition}\markdownRendererInterblockSeparator
{}A key premise of RO-Crate is the existence of a wide variety of resources on the Web that can help describe research. As such, RO-Crate relies on concepts from the Web, in particular that of Linked Data \markdownRendererCite{1}+{}{}{doi:10.2200/S00334ED1V01Y201102WBE001}. Figure \ref{fig:conceptual} shows the main conceptual elements involved in an RO-Crate; an RO-Metadata File(top) describes the research object using structured metadata including external references, coupled with the contained artefacts (bottom) bundled and described by the RO-Crate.\markdownRendererInterblockSeparator
{}\markdownRendererImage{Conceptual RO-Crate Overview}{../content/images/ro-crate-overview.pdf}{../content/images/ro-crate-overview.pdf}{\textbf{Conceptual overview of RO-Crate}. A \emph{Persistent Identifier} (PID) [@doi:10.1371/journal.pbio.2001414] identifies a \emph{Research Object} (RO), which is archived using BagIt \cite{doi:10.17487/rfc8493}, OCFL \cite{ocfl_2020}, git or ZIP. The RO is described within a \textit{RO Metadata File}, providing identifiers for authors using ORCID, organisations using [ROR](https://ror.org/) and licences such as Creative Commons. The \emph{RO-Crate content} is further described with its own metadata. Data can be embedded files and directories, as well as links to external web resources, PIDs [@10.1371/journal.pbio.2001414] and nested RO-Crates.}{#fig:conceptual width="90%"}\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Linked Data as a core principle}\markdownRendererInterblockSeparator
{}Linked Data principles \markdownRendererCite{1}+{}{}{doi:10.4018/978-1-60960-593-3.ch008} (use of IRIs to identify resources (i.e. artefacts), resolvable via HTTP, enriched with metadata and linked to each other) are core to RO-Crate; therefore IRIs\markdownRendererFootnote{IRIs\markdownRendererCite{1}+{}{}{doi:10.17487/rfc3987} are a generalisation of URIs (which include well-known http/https URLs), permitting international Unicode characters without \markdownRendererCodeSpan{\markdownRendererPercentSign{}}-encoding, commonly used on the browser address bar and in HTML5.} are used to identify an RO-Crate, its constituent parts and metadata descriptions, and the properties and classes used in the metadata. \markdownRendererInterblockSeparator
{}Linked Data make it possible for consumers to follow links for more (ideally both human- and machine-readable) information; as the RO-Crate relies on these principles, it can be sufficiently \markdownRendererEmphasis{self-describing} as artefacts can be interrelated using global identifiers, without needing to recursively fully describe every referenced artefact.\markdownRendererInterblockSeparator
{}[PROPOSED REPHRASE ON PREVIOUS PAR] RO Crates are \markdownRendererEmphasis{self-described}; and follow the Linked Data principles to describe all of their resources in both human and machine readable manner. Hence, resources are identified using global identifiers; and relationships between two resources are defined with links.\markdownRendererInterblockSeparator
{}The foundation on Linked Data and shared vocabularies also means multiple RO-Crates and other Linked Data resources can be indexed, combined, queried, integrated or transformed using existing Semantic Web technologies such as \markdownRendererLink{SPARQL}{https://www.w3.org/TR/sparql11-overview/}{https://www.w3.org/TR/sparql11-overview/}{} and knowledge graph triple stores.\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{RO-Crate is a self-described container}\markdownRendererInterblockSeparator
{}An RO-Crate is defined as a self-described \markdownRendererStrongEmphasis{Root Data Entity} that describes and contains \markdownRendererEmphasis{data entities}, which are further described using \markdownRendererEmphasis{contextual entities}. A \markdownRendererStrongEmphasis{data entity} is either a \markdownRendererEmphasis{file} (i.e. a set of bytes stored on disk somewhere) or a \markdownRendererEmphasis{directory} (i.e. dataset of named files and other directories) — while a \markdownRendererStrongEmphasis{contextual entity} exists outside the information system (e.g. a Person, a workflow language) and it is defined by its metadata. The representation of a \markdownRendererStrongEmphasis{data entity} as a set of bytes makes it possible to store a variety of research artefacts including not only data but also, for instance, software and text.\markdownRendererInterblockSeparator
{}The Root Data Entity is a directory, the RO-Crate root, identified by the presence of the \markdownRendererStrongEmphasis{RO-Crate Metadata File} (\markdownRendererCodeSpan{ro-crate-metadata.json}). This is a JSON-LD file that describes the RO-Crate, its content and related metadata using Linked Data. JSON-LD is an RDF serialisation that has become popular as it is easy to read by humans while also offers some advantages for data exchange on the Internet (e.g. it removes some of the cross-domain restrictions that can be present with XML). JSON-LD is the preferred and widely supported format by RO-Crate tools and community.\markdownRendererInterblockSeparator
{}The minimal \markdownRendererLink{requirements for the root data entity metadata}{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#direct-properties-of-the-root-data-entity}{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#direct-properties-of-the-root-data-entity}{} are \markdownRendererCodeSpan{name}, \markdownRendererCodeSpan{description} and \markdownRendererCodeSpan{datePublished}, as well as a contextual entity identifying its \markdownRendererCodeSpan{license} — additional metadata are frequently added depending on the purpose of the particular RO-Crate.\markdownRendererInterblockSeparator
{}RO-Crate can be stored, transferred or published in multiple ways, such as BagIt \markdownRendererCite{1}+{}{}{doi:10.17487/rfc8493}, Oxford Common File Layout \markdownRendererCite{1}+{}{}{ocfl_2020} (OCFL), downloadable ZIP archives in Zenodo or through dedicated online repositories, as well as published directly on the Web, e.g. using GitHub Pages. Combined with Linked Data identifiers, this caters for a diverse set of storage and access requirements across different scientific domains, from metagenomics workflows producing hundreds of gigabytes of genome data to cultural heritage records with access restrictions for personally identifiable data.\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Data Entities are described using Contextual Entities}\markdownRendererInterblockSeparator
{}RO-Crate distinguishes between \markdownRendererLink{data and contextual entities}{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities}{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities}{} in a similar way to HTTP terminology's early attempt to separate \markdownRendererEmphasis{information} (data) and \markdownRendererEmphasis{non-information} (contextual) resources \markdownRendererCite{1}+{}{}{httprange14}. Data entities are usually files and directories located by relative IRI references within the RO-Crate Root, but they can also be Web or restricted resources identified with absolute IRIs.\markdownRendererInterblockSeparator
{}As both types of entities are identified by IRIs, their distinction is allowed to be blurry; data entities can be located anywhere and be complex, while contextual entities can have a Web presence beyond their description inside the RO-Crate. For instance \markdownRendererCodeSpan{https://orcid.org/0000-0002-1825-0097} is primarily an identifier for a person, but secondarily also a web page that describes that person and their academic work. \markdownRendererInterblockSeparator
{}Any particular IRI might appear as a contextual entity in one RO-Crate and as a data entity in another; their distinction lies in the fact that data entities can be considered to be \markdownRendererEmphasis{contained} or captured by the RO-Crate, while contextual entities mainly \markdownRendererEmphasis{explain} the RO-Crate and its entities. It follows that an RO-Crate should have one or more data entities, although this is not a formal requirement. \markdownRendererInterblockSeparator
{}Figure \ref{fig:uml} shows a UML view of RO-Crate, highlighting the different types of data entities and contextual entities that can be aggregated and related.\markdownRendererInterblockSeparator
{}\markdownRendererImage{RO-Crate UML}{../content/images/ro-crate-uml.pdf}{../content/images/ro-crate-uml.pdf}{\textbf{UML model view of RO-Crate.} The \emph{RO-Crate Metadata File} conforms to a version of the specification, which mainly describes the \emph{RO-Crate Root Data Entity} representing the Research Object as a dataset. The RO-Crate aggregates \emph{data entities} (\texttt{hasPart}) which are further described using \emph{contextual entities}. Multiple types and relations from Schema.org allow annotations to be more specific, including figures, nested datasets, computational workflows, people, organisations, instruments and places.}{#fig:uml width="90%"}\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Best Practice Guide rather than strict specification}\markdownRendererInterblockSeparator
{}RO-Crate builds on Linked Data standards (JSON-LD \markdownRendererCite{1}+{}{}{sporny_2014}) and community-evolved vocabularies (\markdownRendererLink{Schema.org}{https://schema.org/}{https://schema.org/}{} \markdownRendererCite{1}+{}{}{doi:10.1145/2857274.2857276}). RO-Crate aims to build a set of best practices on how to practically apply these existing standards in a common way to describe research outputs and their provenance, without having to learn each of the underlying technologies in detail.\markdownRendererInterblockSeparator
{}As such, the \markdownRendererLink{RO-Crate 1.1}{https://w3id.org/ro/crate/1.1}{https://w3id.org/ro/crate/1.1}{} specification \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.4541002} can be seen as an opinionated and example-driven guide to writing Schema.org metadata as JSON-LD, which leaves it open for implementers to include additional metadata using other Schema.org types and properties, or even additional Linked Data vocabularies/ontologies or their own ad-hoc terms.\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Checking Simplicity}\markdownRendererInterblockSeparator
{}As previously stated, one aim of RO-Crate is to be conceptually simple. This simplicity has been repeatedly checked and confirmed through a community review process. For instance, in the discussion on supporting \markdownRendererLink{ad-hoc vocabularies}{https://github.com/ResearchObject/ro-crate/issues/71}{https://github.com/ResearchObject/ro-crate/issues/71}{} in RO-Crate, the community explored potential Linked Data solutions. The conventional wisdom in \markdownRendererLink{RDF best practice}{https://www.w3.org/TR/swbp-vocab-pub/}{https://www.w3.org/TR/swbp-vocab-pub/}{} is to establish a vocabulary with a new URI namespace, formalised using \markdownRendererLink{RDF Schema}{http://www.w3.org/TR/2014/REC-rdf-schema-20140225/}{http://www.w3.org/TR/2014/REC-rdf-schema-20140225/}{} or \markdownRendererLink{OWL}{http://www.w3.org/TR/2012/REC-owl2-overview-20121211/}{http://www.w3.org/TR/2012/REC-owl2-overview-20121211/}{} ontologies, however, as this may seem excessive learning curve for the use case of clarifying that \markdownRendererCodeSpan{sentence} in an RO-Crate refers to prisoner sentencing rather than a linguistic structure, the RO-Crate community instead agreed on a dual lightweight approach: (ⅰ) \markdownRendererLink{Document}{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#adding-new-or-ad-hoc-vocabulary-terms}{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#adding-new-or-ad-hoc-vocabulary-terms}{} how projects with their own web-presence can make a pure HTML-based vocabulary, and (ⅱ) provide a community-wide PID namespace under \markdownRendererLink{https://w3id.org/ro/terms/}{https://w3id.org/ro/terms/}{https://w3id.org/ro/terms/}{} that redirect to simple CSV files \markdownRendererLink{maintained in GitHub}{https://github.com/ResearchObject/ro-terms}{https://github.com/ResearchObject/ro-terms}{}. \markdownRendererInterblockSeparator
{}To further verify this idea, we have formalised the RO-Crate definition (see Appendix A). An important result of this exercise is that the underlying data structure of RO-Crate, although conceptually a graph, is represented as a tree limited to depth 2. The formalisation also emphasises the \markdownRendererEmphasis{boundedness} of the structure; namely, the fact that elements are specifically identified as being either semantically \markdownRendererEmphasis{contained} by the RO-Crate (\markdownRendererCodeSpan{hasPart}) or mainly referenced (\markdownRendererCodeSpan{mentions}) and typed as \markdownRendererEmphasis{external} to the Research Object (Contextual Entities). \relax

### 2021-08-08 output/_markdown_ro-crate/941d50e1d9479dc3f4480e58c6f60f66.md.tex
Updates from Overleaf

\markdownRendererHeadingTwo{RO-Crate Community}\markdownRendererInterblockSeparator
{}The RO-Crate conceptual model, implementation and best practices are developed by a growing community of researchers, developers and publishers. The RO-Crate community is a key aspect of its effectiveness in making research artefacts FAIR. Fundamentally, the community provides the overall context of the implementation and model and ensures its interoperability. \markdownRendererInterblockSeparator
{}The RO-Crate community consists of:\markdownRendererInterblockSeparator
{}\markdownRendererOlBeginTight
\markdownRendererOlItemWithNumber{1}a diverse set of people representing a variety of stakeholders;\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{2}a set of collective norms;\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{3}an open platform that facilitates communication (GitHub, Google Docs, monthly teleconferences).\markdownRendererOlItemEnd 
\markdownRendererOlEndTight \markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{People}\markdownRendererInterblockSeparator
{}The initial concept of RO-Crate was formed at the first Workshop on Research Objects (\markdownRendererLink{RO2018}{https://www.researchobject.org/ro2018/}{https://www.researchobject.org/ro2018/}{}), held as part of the IEEE conference on eScience. This workshop followed up on considerations made at a \markdownRendererLink{Research Data Alliance (RDA) meeting on Research Data Packaging}{https://rd-alliance.org/approaches-research-data-packaging-rda-11th-plenary-bof-meeting}{https://rd-alliance.org/approaches-research-data-packaging-rda-11th-plenary-bof-meeting}{} that found similar goals across multiple data packaging efforts \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.3250687}: simplicity, structured metadata and the use of JSON-LD.\markdownRendererInterblockSeparator
{}An important outcome of discussions that took place at RO2018 was the conclusion that the original Wf4Ever Research Object ontologies \markdownRendererCite{1}+{}{}{doi:10.1016/j.websem.2015.01.003}, in principle sufficient for packaging research artefacts with rich descriptions, were, in practice, considered inaccessible for regular programmers (e.g. web developers) and in danger of being incomprehensible for domain scientists due to their reliance on Semantic Web technologies and other ontologies.\markdownRendererInterblockSeparator
{}DataCrate \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.1445817} was presented at RO2018 as a promising lightweight alternative approach, and an agreement was made by a group of volunteers to attempt building “RO Lite” as a combination of DataCrate's implementation and Research Object's principles.\markdownRendererInterblockSeparator
{}This group, originally made up of library and Semantic Web experts, has subsequently grown to include domain scientists, developers, publishers and more. This perspective of multiple views led to the specification being used in a variety of domains, from bioinformatics and regulatory submissions to humanities and cultural heritage preservation. \markdownRendererInterblockSeparator
{}The RO-Crate community is strongly engaged with the European-wide biology/bioinformatics collaborative e-Infrastructure, ELIXIR, \markdownRendererCite{1}+{}{}{doi:10.1016/j.tibtech.2012.02.002}, along with European Open Science Cloud (EOSC) projects including EOSC-Life and FAIRplus. RO-Crate has also established collaborations with Bioschemas, GA4GH, OpenAIRE and multiple H2020 projects.\markdownRendererInterblockSeparator
{}A key set of stakeholders are developers; the RO-Crate community has made a point of attracting developers who can implement the specifications but, importantly, keeps “developer user experience” in mind. This means that the specifications are straightforward to implement and thus do not require expertise in technologies that are not widely deployed. \markdownRendererInterblockSeparator
{}This notion of catering to “developer user experience” is an example of the set of norms that have developed and now define the community. \markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Norms}\markdownRendererInterblockSeparator
{}The RO-Crate community is driven by conventions or notions that are prevalent within it but not formalised. Here, we distil what we as authors believe are the critical set of norms that have facilitated the development of RO-Crate and contributed to the ability for RO-Crate research packages to be FAIR. This is not to say that there are no other norms within the community or that everyone in the community holds these uniformly. Instead, what we emphasise is that these norms are helpful and also shaped by community practices. \markdownRendererInterblockSeparator
{}\markdownRendererOlBeginTight
\markdownRendererOlItemWithNumber{1}Simplicity\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{2}Developer friendliness\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{3}Focus on examples and best practices rather than rigorous specification\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{4}Reuse “just enough” Web standards\markdownRendererOlItemEnd 
\markdownRendererOlEndTight \markdownRendererInterblockSeparator
{}A core norm of RO-Crate is that of \markdownRendererStrongEmphasis{simplicity}, which sets the scene for how we guide developers to structure metadata with RO-Crate. We focus mainly on documenting simple approaches to the most common use cases, such as authors having an affiliation. This norm also influences our take on \markdownRendererStrongEmphasis{developer friendliness}; for instance, we are using the Web-native JSON format, allowing only a few of JSON-LD's flexible Linked Data features. Moreover, the RO-Crate documentation is largely built up by \markdownRendererStrongEmphasis{examples} showcasing \markdownRendererStrongEmphasis{best practices}, rather than rigorous specifications. In a sense, we are allowed to do this by building on existing \markdownRendererStrongEmphasis{Web standards} that themselves are defined rigorously, which we utilise \markdownRendererEmphasis{“\markdownRendererStrongEmphasis{just enough}”} in order to benefit from the advantages of Linked Data (e.g. extensions by namespaced vocabularies), without imposing too many developer choices or uncertainties (e.g. having to choose between the many RDF syntaxes). \markdownRendererInterblockSeparator
{}While the above norms alone could easily lead to the creation of “yet another” JSON format, we keep the goal of \markdownRendererStrongEmphasis{FAIR interoperability} of the captured metadata, and therefore follow closely FAIR best practices and current developments such as data citations, PIDs, open repositories and recommendations for sharing research outputs and software.\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Open Platforms}\markdownRendererInterblockSeparator
{}The critical infrastructure that enables the community around RO-Crate is the use of open development platforms. This underpins the importance of open community access to supporting FAIR. Specifically, it is difficult to build and consume FAIR research artefacts without being able to access the specifications, understand how they are developed, know about any potential implementation issues, and discuss usage to evolve best practices. \markdownRendererInterblockSeparator
{}It was seen as more important to document real-life examples and best practices rather than developing a rigorous specification. At the same time, we agreed to be opinionated on the syntactic form to reduce the jungle of implementation choices; we wanted to keep the important aspects of Linked Data to adhere to the FAIR principles while retaining the option of combining and extending the structured metadata using the existing Semantic Web stack, not just build “yet another” standalone JSON format. \markdownRendererInterblockSeparator
{}Further work during 2019 started adapting the DataCrate documentation through a more collaborative and exploratory \markdownRendererEmphasis{RO-Lite} phase, initially using Google Docs for review and discussion, then moving to GitHub as a collaboration space for developing what is now the RO-Crate specification, \markdownRendererLink{maintained as Markdown}{https://github.com/researchobject/ro-crate/}{https://github.com/researchobject/ro-crate/}{} in GitHub Pages and published through Zenodo. \markdownRendererInterblockSeparator
{}In addition to the typical Open Source-style development with GitHub issues and pull requests, the RO-Crate Community now has two regular monthly calls, a Slack channel and a mailing list for coordinating the project, and many of its participants collaborate on RO-Crate at multiple conferences and coding events such as the ELIXIR BioHackathon. The community is jointly developing the RO-Crate specification and Open Source tools, as well as providing support and considering new use cases. The \markdownRendererLink{RO-Crate Community}{https://www.researchobject.org/ro-crate/community}{https://www.researchobject.org/ro-crate/community}{} is open for anyone to join, to equally participate under a code of conduct, and currently has more than 40 members. \relax

### 2021-08-08 output/_markdown_ro-crate/a4f048493539fb8c35711fa80e7550d8.md.tex
Updates from Overleaf

\markdownRendererHeadingTwo{Digital Humanities: Cultural Heritage}\markdownRendererInterblockSeparator
{}\markdownRendererLink{PARADISEC}{https://www.paradisec.org.au/}{https://www.paradisec.org.au/}{} (the Pacific And Regional Archive for Digital Sources in Endangered Cultures) maintains a repository of more than 500,000 files documenting endangered languages across more than 16,000 items, collected over many years by researchers interviewing and recording native speakers across the region. As a proposed update of the 18 year old infrastructure, the \markdownRendererLink{Modern PARADISEC demonstrator}{https://mod.paradisec.org.au/}{https://mod.paradisec.org.au/}{} has been \markdownRendererLink{developed}{https://arkisto-platform.github.io/case-studies/paradisec/}{https://arkisto-platform.github.io/case-studies/paradisec/}{} to also help digitally preserve these artefacts using the \markdownRendererLink{Oxford Common File Layout}{https://ocfl.io/1.0/spec/}{https://ocfl.io/1.0/spec/}{} (OCFL) for file consistency and RO-Crate for structuring and capturing the metadata of each item. The existing PARADISEC data collection has been ported and captured as RO-Crates. A web portal then exposes the repository and its entries by indexing the RO-Crate metadata files using Elasticsearch as a “NoSQL” object database, presenting a domain-specific view of the items - the RO-Crate is “hidden” and does not change the user interface.\markdownRendererInterblockSeparator
{}This use case takes advantage of several RO-Crate features and principles. Firstly, the transcribed metadata are now independent of the PARADISEC platform and can be archived, preserved and processed in its own right, using Schema.org vocabularies augmented with PARADISEC-specific terms. The lightweight infrastructure with RO-Crate as the holder of itemised metadata in regular files (organised using OCFL\markdownRendererCite{1}+{}{}{ocfl_2020}, with checksums for integrity checking and versioning) also gives flexibility for future developments and maintenance; for example, potentially using Linked Data software such as a graph database, queried using SPARQL triple patterns across RO-Crates, or a “last resort” fallback to the generic RO-Crate HTML preview \markdownRendererCite{1}+{}{}{ro-crate-html-js}, which can be hosted as static files by any web server, in line with the approach taken by the Endings Project\markdownRendererFootnote{The Endings Project \url{https://endings.uvic.ca/} is a five-year project funded by the Social Sciences and Humanities Research Council (SSHRC) that is creating tools, principles, policies and recommendations for digital scholarship practitioners to create accessible, stable, long-lasting resources in the humanities.}.\relax

### 2021-08-08 output/_markdown_ro-crate/18bbe0122ea97198385a02765d1d7825.md.tex

\markdownRendererHeadingOne{RO-Crate}\markdownRendererInterblockSeparator
{}RO-Crate provides a lightweight approach to packaging research artefacts with their metadata. To illustrate this, let us imagine a research paper reporting on the sequence analysis of proteins obtained from an experiment on mice. The sequence output files, sequence analysis code, resulting data and reports summarising statistical measures or outputs are all important and inter-related research outputs, and consequently would ideally all be co-located in a directory and accompanied with their corresponding metadata. In reality, some of the artefacts (e.g. data or software) will be recorded as external references, not necessarily captured in a FAIR way. This directory, along with the relationships between its constituent digital artefacts, is what the RO-Crate model aims to represent, linking together all the elements pertaining to an experiment and required for its reproducibility. \markdownRendererInterblockSeparator
{}The question then arises as to how the directory with all this material should be packaged in a manner that is accessible and usable by others. By usable we mean not just readable by humans but programmatically accessible. A de facto approach to sharing collections of resources is through compressed archives (e.g. a zip file). This solves the problem of “packaging”, but it does not guarantee downstream access to all artefacts in a programmatic fashion, or the role of each file in that particular research. This leads to the need for explicit metadata about the contents of the folder, describing each and linking them together.\markdownRendererInterblockSeparator
{}Examples of metadata descriptions across a \markdownRendererLink{wide range of domains}{https://rdamsc.bath.ac.uk/scheme-index}{https://rdamsc.bath.ac.uk/scheme-index}{} abound within the literature, both in research data management (?cite) and within library and information systems (?cite). However, many of these approaches require knowledge of metadata schemas, particular annotation systems, or the use of obscure or complex software stacks. Indeed, particularly within research, these requirements have led to a lack of adoption and growing frustration with current tooling and specifications \markdownRendererCite{1}+{}{}{neylon_blog_post_2017}.\markdownRendererInterblockSeparator
{}RO-Crate seeks to address this complexity by:\markdownRendererInterblockSeparator
{}\markdownRendererOlBeginTight
\markdownRendererOlItemWithNumber{1}being easy to understand and conceptually simple;\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{2}providing a strong and opinionated guide regarding current best practices;\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{3}adopting de-facto standards that are widely used on the Web.\markdownRendererOlItemEnd 
\markdownRendererOlEndTight \markdownRendererInterblockSeparator
{}In the following sections we show how the RO-Crate specification and ecosystem achieves these goals, which concur in forming our definition of “lightweight”. \relax

### 2021-08-08 output/_markdown_ro-crate/6e307892a9f349f2a14d0e76a65d0100.md.tex

\markdownRendererHeadingTwo{Technical implementation of the model}\markdownRendererInterblockSeparator
{}The RO-Crate conceptual model has been realised using JSON-LD and Schema.org in a prescriptive form as discussed in the best practice approach. This technical approach again caters for simplicity. \markdownRendererInterblockSeparator
{}\markdownRendererLink{JSON-LD}{https://json-ld.org/}{https://json-ld.org/}{} provides a way to express Linked Data as a JSON structure, where a \markdownRendererEmphasis{context} provides mapping to RDF properties and classes. While JSON-LD cannot map arbitrary JSON structures to RDF, we found it does lower the barrier regarding Linked Data serialisations as it follows a JSON structure, nowadays a very common and popular format for data exchange.\markdownRendererInterblockSeparator
{}However, JSON-LD alone, has too many degrees of freedom and hidden complexities for software developers to reliably produce and consume without specialised expertise or software libraries. A large part of the RO-Crate specification is therefore dedicated to describing JSON structures. \markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{RO-Crate JSON-LD}\markdownRendererInterblockSeparator
{}RO-Crate mandates the use of \markdownRendererLink{flattened, compacted JSON-LD}{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html}{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html}{} where a single \markdownRendererCodeSpan{@graph} array contains all the data and contextual entities in a flat list. An example can be seen in the JSON-LD snippet below, describing a simple RO-Crate containing two datasets (data1.txt and data2.txt):\markdownRendererInterblockSeparator
{}\markdownRendererInputFencedCode{../output/_markdown_ro-crate/634f7904aa7296f7ec26373a016d51ed.verbatim}{json}\markdownRendererInterblockSeparator
{}\markdownRendererEmphasis{Figure X}: \markdownRendererEmphasis{Simplified RO-Crate JSON-LD showing the flattened compacted \markdownRendererCodeSpan{@graph} array}\markdownRendererInterblockSeparator
{}It can be argued that this is a more graph-like approach than the tree structure JSON would otherwise invite, and which is normally emphasised as a feature of JSON-LD in order to “hide” its RDF nature. \markdownRendererInterblockSeparator
{}However, we found that the use of trees for, e.g. \markdownRendererEmphasis{a \markdownRendererEmphasis{Person} entity appearing as author of a \markdownRendererCodeSpan{File} which nests under a \markdownRendererCodeSpan{Dataset}, \markdownRendererCodeSpan{hasPart}}, counter-intuitively leads one to consider the JSON-LD as an RDF Graph, since an identified \markdownRendererCodeSpan{Person} entity can then appear at multiple and repeated points of the tree (e.g. author of multiple files), necessitating node merging or duplication. \markdownRendererInterblockSeparator
{}By comparison, a single flat \markdownRendererCodeSpan{@graph} array approach means that applications can process and edit each entity as pure JSON by a simple lookup based on \markdownRendererCodeSpan{@id}. At the same time, lifting all entities to the same level emphasises Research Object's principle that describing the context and provenance is just as important as describing the data.\markdownRendererInterblockSeparator
{}RO-Crate reuses Schema.org, but provides its own versioned JSON-LD context, which has a similar flat list with the mapping from JSON-LD keys to their URI equivalents (e.g. \markdownRendererCodeSpan{author} maps to \markdownRendererLink{http://schema.org/author}{http://schema.org/author}{http://schema.org/author}{}). The rationale behind this decision is to support JSON-based RO-Crate applications that are largely unaware of JSON-LD, and thus still may want to process the \markdownRendererCodeSpan{@context} to find Linked Data definitions of unknown properties and types. Not reusing the official Schema.org context means RO-Crate is also able to map in additional vocabularies where needed, namely the \markdownRendererEmphasis{Portland Common Data Model} (PCDM) \markdownRendererCite{1}+{}{}{pcdm} for repositories and Bioschemas \markdownRendererCite{1}+{}{}{bioschemas_2017} for describing computational workflows.\markdownRendererInterblockSeparator
{}Similarly, rather than relying on implications from \markdownRendererCodeSpan{"@type": "@id"} annotations in the context, RO-Crate JSON-LD distinguishes explicitly between references to other entities (\markdownRendererCodeSpan{\markdownRendererLeftBrace{}"@id": "\markdownRendererHash{}alice"\markdownRendererRightBrace{}}) and string values (\markdownRendererCodeSpan{"Alice"}) - meaning RO-Crate applications can find the corresponding entity without parsing the \markdownRendererCodeSpan{@context}.\relax

### 2021-08-08 output/_markdown_ro-crate/9f734e70d2ecda4384f6c7600c9b9396.md.tex

\markdownRendererHeadingOne{RO-Crate Tooling}\markdownRendererInterblockSeparator
{}The work of the community led to the development of a number of tools for creating and using RO-Crates. Table \ref{tab:tools} shows the current set of implementations. Reviewing this list, one can see that tools support commonly used programming languages, including Python, JavaScript, and Ruby. Additionally, these tools can be integrated into commonly used research environments; in particular, the command line (\markdownRendererEmphasis{ro-crate-html-js}). Furthermore, there are tools that cater to the end-user (\markdownRendererEmphasis{Describo}, \markdownRendererEmphasis{Workflow Hub}). For example, Describo was developed to help researchers of the Australian \markdownRendererLink{Criminal Characters project}{https://criminalcharacters.com/}{https://criminalcharacters.com/}{} annotate historical prisoner records to gain greater insight into the history of Australia \markdownRendererCite{1}+{}{}{doi:10.1080/14490854.2020.1796500}. \markdownRendererInterblockSeparator
{}While the development of these tools is promising, our analysis of their maturity status shows that the majority of them are in the Beta stage. This is partly due to the fact that the RO-Crate specification itself only recently reached 1.0 status, in November 2019 \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.3541888}. Now that there is a fixed point of reference, and RO-Crate 1.1 (October 2020) \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.4031327} has stabilised based on feedback from application development, we expect to see a further increase in the maturity of these tools, along with the creation of new ones.\markdownRendererInterblockSeparator
{}Given the stage of the specification, these tools have been primarily targeted to developers, essentially providing them with the core libraries for working with RO-Crates. Another target has been that of research data managers who need to manage and curate large amounts of data. \relax

### 2021-08-08 output/_markdown_ro-crate/a2088708cc15c494d29d44db742568d6.md.tex

\markdownRendererHeadingOne{Profiles of RO-Crate in use}\markdownRendererInterblockSeparator
{}RO-Crate is fundamentally an infrastructure to help build FAIR research artefacts. In other words, the key question is whether RO-Crate can be used to share and (re)use research artefacts. Here we look at three research domains where RO-Crate is being applied: Bioinformatics, Regulatory Science and Cultural Heritages. In addition, we note how RO-Crate may have an important role as part of machine-actionable data management plans and institutional repositories.\markdownRendererInterblockSeparator
{}From these varied uses of RO-Crate we observe a natural differences in their detail level and the type of entities described by the RO-Crate. For instance, on submission of an RO-Crate to a workflow repository, it is reasonable to expect the RO-Crate to contain at least one workflow, ideally with a declared licence and workflow language. Specific additional recommendations such as on identifiers is also needed to meet the emerging requirements of \markdownRendererLink{FAIR Digital Objects}{https://fairdo.org/}{https://fairdo.org/}{}. \markdownRendererLink{Work has now begun}{https://github.com/ResearchObject/ro-crate/issues/153}{https://github.com/ResearchObject/ro-crate/issues/153}{} to formalise these different \markdownRendererEmphasis{profiles} of RO-Crates, which may impose additional constraints based on the needs of a specific domain or use case. \relax

### 2021-08-08 output/_markdown_ro-crate/bd47484f290d506e583cac18f875a772.md.tex

\markdownRendererHeadingOne{Introduction}\markdownRendererInterblockSeparator
{}The move towards open science has increased the need and demand for the publication of artefacts of the research process \markdownRendererCite{1}+{}{}{sefton_blog_post_2021}. This is particularly apparent in domains that rely on computational experiments; for example, the publication of software, datasets and records of the dependencies that such experiments rely on \markdownRendererCite{1}+{}{}{doi:10.1126/science.aah6168}. \markdownRendererInterblockSeparator
{}It is often argued that the publication of these assets, and specifically software \markdownRendererCite{1}+{}{}{doi:10.3233/DS-190026}, workflows \markdownRendererCite{1}+{}{}{https://doi.org/10.1162/dint_a_00033} and data, should follow the FAIR principles \markdownRendererCite{1}+{}{}{doi:10.1038/sdata.2016.18}; namely, that they are Findable, Accessible, Interoperable and Reusable. These principles are agnostic to the \markdownRendererEmphasis{implementation} strategy needed to comply with them. Hence, there has been an increasing amount of work in the development of platforms and specifications that aim to fulfil these goals \markdownRendererCite{1}+{}{}{isbn:9781315351148}. Important examples include data publication with rich metadata (e.g. Zenodo \markdownRendererCite{1}+{}{}{doi:10.3897/biss.3.37080}), domain-specific data deposition (e.g., PDB \markdownRendererCite{1}+{}{}{doi:10.1093/nar/gkl971}) and following practices for reproducible research software \markdownRendererCite{1}+{}{}{doi:10.1371/journal.pcbi.1003285} (e.g. use of containers). \markdownRendererInterblockSeparator
{}These strategies are focused primarily on one \markdownRendererEmphasis{type} of artefact. To address this, \markdownRendererCite{1}+{}{}{doi:10.1016/j.future.2011.08.004} introduced the notion of \markdownRendererStrongEmphasis{research objects} – \markdownRendererEmphasis{semantically rich aggregations of (potentially distributed) resources that provide a layer of structure on top of information delivered in a machine-readable format}. A Research Object combines the ability to bundle multiple types of artefacts together, such as CSV files, code, examples, and figures. This provides a compelling vision as an approach for implementing FAIR. However, existing research object implementations require a large technology stack, are tailored to a particular platform and are also not easily usable by end-users. \markdownRendererInterblockSeparator
{}To address this gap, a new community came together \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.3250687} to develop \markdownRendererStrongEmphasis{RO-Crate} - an \markdownRendererEmphasis{approach to package and aggregate research artefacts with their metadata and relationships}. The aim of this paper is to introduce RO-Crate and assess it as a strategy for making multiple types of research artefacts FAIR. Specifically, the contributions of this paper are as follows:\markdownRendererInterblockSeparator
{}\markdownRendererOlBeginTight
\markdownRendererOlItemWithNumber{1}an introduction to RO-Crate, its purpose and context;\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{2}a guide to the RO-Crate community and tooling;\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{3}and an exemplar usage of RO-Crate for different artefacts in different communities as well as its use as a connective tissue for such artefacts.\markdownRendererOlItemEnd 
\markdownRendererOlEndTight \markdownRendererInterblockSeparator
{}The rest of this paper is organised as follows. We first describe RO-Crate, the assumptions underlying it, and define RO-Crate technically and formally. We then proceed to introduce the community and tooling. We move to analyse RO-Crate with respect to usage in a diverse set of domains. Finally, we present related work and conclude with some remarks including RO-Crate highlights and future work. \relax

### 2021-08-08 output/_markdown_ro-crate/dd30b07fff9d6ccb8e0bef502350b086.md.tex

\markdownRendererHeadingTwo{Machine-actionable Data Management Plans}\markdownRendererInterblockSeparator
{}Machine-actionable Data Management Plans (maDMPs) have been proposed as an improvement to automate FAIR data management tasks in research \markdownRendererCite{1}+{}{}{doi:10.1371/journal.pcbi.1006750}, e.g. by using PIDs and controlled vocabularies to describe what happens to data over the research life cycle \markdownRendererCite{1}+{}{}{doi:10.1007/978-3-030-45442-5_15}. The Research Data Alliance's \markdownRendererEmphasis{DMP Common Standard} for maDMPs \markdownRendererCite{1}+{}{}{doi:10.15497/rda00039} is one such formalisation for expressing maDMPs, which can be expressed as Linked Data using the DMP Common Standard Ontology \markdownRendererCite{1}+{}{}{doi:10.4126/frl01-006423289}, a specialisation of the W3C Data Catalog Vocabulary (DCAT) \markdownRendererCite{1}+{}{}{dcat2}. RDA maDMPs are usually expressed using regular JSON, conforming to the DMP JSON Schema.\markdownRendererInterblockSeparator
{}A mapping has been produced between Research Object Crates and Machine-actionable Data Management Plans \markdownRendererCite{1}+{}{}{doi:10.4126/frl01-006423291}, implemented by the RO-Crate {RDA maDMP Mapper \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.3922136}. A similar mapping has been implemented by \markdownRendererCodeSpan{RO-Crate\markdownRendererUnderscore{}2\markdownRendererUnderscore{}ma-DMP} \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.3903463}. In both cases, a maDMP can be converted to a RO-Crate, or vice versa. In \markdownRendererCite{1}+{}{}{doi:10.4126/frl01-006423291} this functionality caters for two use cases:\markdownRendererInterblockSeparator
{}\markdownRendererOlBeginTight
\markdownRendererOlItemWithNumber{1}Start a skeleton data management plan based on an existing RO-Crate dataset, e.g. from an RO-Crate from WorkflowHub.\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{2}Instantiate an RO-Crate based on a data management plan.\markdownRendererOlItemEnd 
\markdownRendererOlEndTight \markdownRendererInterblockSeparator
{}An important difference here is that data management plans are (ideally) written in advance of data production, while RO-Crates are typically created to describe data after it has been generated. This approach shows the importance of \markdownRendererEmphasis{templating} to make both tasks more automatable and achievable, and how RO-Crate can fit into earlier stages of the research life cycle.\relax

### 2021-08-08 _markdown_ro-crate/22fbdbb418a36300535ce119da833f99.md.tex
Merge overleaf-2021-08-08-0724 into master [ci skip]
This build is based on
https://github.com/stain/ro-crate-paper/commit/ddb80a8f0dac3a15d4db06d53b8e5463e720cc7b.

This commit was created by the following CI build and job:
https://github.com/stain/ro-crate-paper/commit/ddb80a8f0dac3a15d4db06d53b8e5463e720cc7b/checks
https://github.com/stain/ro-crate-paper/runs/1109599515

An increasing amount of researchers support reproducibility by including pointers to and descriptions of datasets, software and methods in their publications. However, scientific articles may be ambiguous, incomplete and difficult to process by automated systems. In this paper we introduce RO-Crate, an open, community-driven, and lightweight approach to packaging research artefacts along with their metadata in a machine readable manner. Based on Schema.org annotations in JSON-LD, RO-Crate aims to establish best practices to formally describe metadata in an accessible and practical way for their use in a wide variety of situations. \markdownRendererInterblockSeparator
{}An RO-Crate is a structured archive of all the items that contributed to a research outcome, including their identifiers, provenance, relations and annotations. As a general purpose packaging framework for data and their metadata, RO-Crate is used across multiple areas, including bioinformatics, digital humanities and regulatory sciences. By applying "just enough" Linked Data standards, RO-Crate simplifies the process of making research outputs FAIR while also contributes to enhance research reproducibility.\relax

### 2021-08-08 _markdown_ro-crate/3ffef2c3bc11bc83c92df89b6fc00f76.md.tex
Merge overleaf-2021-08-08-0724 into master [ci skip]
This build is based on
https://github.com/stain/ro-crate-paper/commit/ddb80a8f0dac3a15d4db06d53b8e5463e720cc7b.

This commit was created by the following CI build and job:
https://github.com/stain/ro-crate-paper/commit/ddb80a8f0dac3a15d4db06d53b8e5463e720cc7b/checks
https://github.com/stain/ro-crate-paper/runs/1109599515

\markdownRendererHeadingTwo{Regulatory Sciences}\markdownRendererInterblockSeparator
{}\markdownRendererLink{BioCompute Objects}{https://www.biocomputeobject.org/}{https://www.biocomputeobject.org/}{} (BCO) \markdownRendererCite{1}+{}{}{doi:10.1371/journal.pbio.3000099} is a community-led effort to standardise submissions of computational workflows to biomedical regulators. For instance, a genomics sequencing pipeline, as part of a personalised cancer treatment study, can be submitted to the US Food and Drugs Administration (FDA) for approval. BCOs are formalised in the standard IEEE 2791-2020 \markdownRendererCite{1}+{}{}{doi:10.1109/IEEESTD.2020.9094416} as a combination of \markdownRendererLink{JSON Schemas}{https://opensource.ieee.org/2791-object/ieee-2791-schema/}{https://opensource.ieee.org/2791-object/ieee-2791-schema/}{} that define the structure of JSON metadata files describing exemplar workflow runs in detail, covering aspects such as the usability and error domain of the workflow, its runtime requirements, the reference datasets used and representative output data produced.\markdownRendererInterblockSeparator
{}BCOs provide a structured view over a particular workflow, informing regulators about its workings independently of the underlying workflow definition language. However, BCOs have only limited support for additional metadata\markdownRendererFootnote{IEEE 2791-2020 do permit user extensions in the \markdownRendererEmphasis{extension domain} by referencing additional JSON Schemas.}. For instance, while the BCO itself can indicate authors and contributors, and in particular regulators and their review decisions, it cannot describe the provenance of individual data files or workflow definitions. \markdownRendererInterblockSeparator
{}As a custom JSON format, BCOs cannot be extended with Linked Data concepts, except by adding an additional top-level JSON object formalised in another JSON Schema. A BCO and workflow submitted by upload to a regulator will also frequently consist of multiple cross-related files. Crucially, there is no way to tell whether a given \markdownRendererCodeSpan{*.json} file is a BCO file, except by reading its content and check for its \markdownRendererCodeSpan{spec\markdownRendererUnderscore{}version}. \markdownRendererInterblockSeparator
{}We can then consider how a BCO and its referenced artefacts can be packaged and transferred following FAIR principles. \markdownRendererLink{\markdownRendererStrongEmphasis{BCO RO-Crate}}{https://biocompute-objects.github.io/bco-ro-crate/}{https://biocompute-objects.github.io/bco-ro-crate/}{}\markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.4633732}, part of the BioCompute Object user guides, defines a set of best practices for wrapping a BCO with a workflow, together with its exemplar outputs in an RO-Crate, which then provides typing and additional provenance metadata of the individual files, workflow definition, referenced data and the BCO metadata itself. \markdownRendererInterblockSeparator
{}Here the BCO is responsible for describing the \markdownRendererEmphasis{purpose} of a workflow and its run at an abstraction level suitable for a domain scientist, while the more open-ended RO-Crate describes the surroundings of the workflow, classifying and relating its resources and providing provenance of their existence beyond the BCO. This emerging \markdownRendererEmphasis{separation of concerns} highlight how RO-Crate is used side-by-side of existing standards, even where there are apparent partial overlaps.\markdownRendererInterblockSeparator
{}A similar separation of concerns can be found if considering the RO-Crate as a set of files, where the \markdownRendererEmphasis{transport-level} metadata, such as checksum of files, are \markdownRendererLink{delegated to BagIt}{https://www.researchobject.org/ro-crate/1.1/appendix/implementation-notes.html#adding-ro-crate-to-bagit}{https://www.researchobject.org/ro-crate/1.1/appendix/implementation-notes.html#adding-ro-crate-to-bagit}{} manifests, a standard focusing on the preservation challenges of digital libraries\markdownRendererCite{1}+{}{}{doi:10.17487/rfc8493}. As such, RO-Crates are not required to iterate all the files in their folder hierarchy, only those that benefit from being described.\markdownRendererInterblockSeparator
{}Specifically, a BCO alone is insufficient for reliable re-execution of a workflow, which would need a compatible workflow engine depending on the workflow definition language, so IEEE 2791 recommends using Common Workflow Language \markdownRendererCite{1}+{}{}{arxiv:2105.07028} for interoperable pipeline execution. CWL itself relies on tool packaging in software containers using \markdownRendererLink{Docker}{https://www.docker.com/}{https://www.docker.com/}{} or \markdownRendererLink{Conda}{https://docs.conda.io/}{https://docs.conda.io/}{}. Thus, we can consider BCO RO-Crate as a stack: transport-level manifests of files (BagIt), provenance, typing and context of those files (RO-Crate), workflow overview and purpose (BCO), interoperable workflow definition (CWL) and tool distribution (Docker).\markdownRendererInterblockSeparator
{}\markdownRendererImage{separationofconcerns}{../content/images/ro-crate-bco-sep-of-concerns.pdf}{../content/images/ro-crate-bco-sep-of-concerns.pdf}{\textbf{Separation of Concerns in BCO RO-Crate}. BioCompute Object (IEEE2791) is a JSON file that structurally explains the purpose and implementation of a computational workflow, for instance implemented in Nextflow, that installs the workflow’s software dependencies as Docker containers or BioConda packages. An example execution of the workflow shows the different kinds of result outputs, which may be external, using GitHub LFS to support larger data. RO-Crate gathers all these local and external resources, relating them and giving individual descriptions, for instance permanent DOI identifiers for reused datasets accessed from Zenodo, but also adding external identifiers to attribute authors using ORCID or to identify which licences apply to individual resources. The RO-Crate and its local files are captured in a BagIt whose checksum ensures completeness, combined with Big Data Bag \cite{doi:10.1109/BigData.2016.7840618} features to “complete” the bag with large external files such as the workflow outputs}\relax

### 2021-08-08 _markdown_ro-crate/6294f39211071c8e56e22ae6c39e1fbd.md.tex
Merge overleaf-2021-08-08-0724 into master [ci skip]
This build is based on
https://github.com/stain/ro-crate-paper/commit/ddb80a8f0dac3a15d4db06d53b8e5463e720cc7b.

This commit was created by the following CI build and job:
https://github.com/stain/ro-crate-paper/commit/ddb80a8f0dac3a15d4db06d53b8e5463e720cc7b/checks
https://github.com/stain/ro-crate-paper/runs/1109599515

\markdownRendererHeadingTwo{Bioinformatics workflows}\markdownRendererInterblockSeparator
{}\markdownRendererLink{WorkflowHub.eu}{https://workflowhub.eu/}{https://workflowhub.eu/}{} is a European cross-domain registry of computational workflows, supported by European Open Science Cloud projects, e.g. \markdownRendererLink{EOSC-Life}{https://www.eosc-life.eu/}{https://www.eosc-life.eu/}{}, and research infrastructures including the pan-European bioinformatics network \markdownRendererLink{ELIXIR}{https://elixir-europe.org/}{https://elixir-europe.org/}{} \markdownRendererCite{1}+{}{}{doi:10.1016/j.tibtech.2012.02.002}. As part of promoting workflows as reusable tools, WorkflowHub includes documentation and high-level rendering of the workflow structure independent of its native workflow definition format. The rationale is that a domain scientist can browse all relevant workflows for their domain, before narrowing down their workflow engine requirements. As such, the WorkflowHub is intended largely as a registry of workflows already deposited in repositories specific to particular workflow languages and domains, such as UseGalaxy.eu \markdownRendererCite{1}+{}{}{doi:10.1371/journal.ppat.1008643} and Nextflow nf-core \markdownRendererCite{1}+{}{}{doi:10.1038/s41587-020-0439-x}. \markdownRendererInterblockSeparator
{}We here describe three different RO-Crate profiles developed for use with WorkflowHub.\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Profile for describing workflows}\markdownRendererInterblockSeparator
{}Being cross-domain, WorkflowHub has to cater for many different workflow systems. Many of these, for instance Nextflow \markdownRendererCite{1}+{}{}{doi:10.1038/nbt.3820} and Snakemake \markdownRendererCite{1}+{}{}{doi:10.1093/bioinformatics/bts480}, by virtue of their script-like nature, reference multiple neighbouring files typically maintained in a GitHub repository. This calls for a data exchange method that allows keeping related files together. WorkflowHub has tackled this problem by adopting RO-Crate as the packaging mechanism \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.4705078}, typing and annotating the constituent files of a workflow and — crucially — marking up the workflow language, as many workflow engines use common file extensions like \markdownRendererCodeSpan{*.xml} and \markdownRendererCodeSpan{*.json}. Workflows are further described with authors, licence, diagram previews and a listing of their inputs and outputs. RO-Crates can thus be used for interoperable deposition of workflows to WorkflowHub, but are also used as an archive for downloading workflows, embedding metadata registered with the WorkflowHub entry and translated workflow files such as abstract Common Workflow Language (CWL) \markdownRendererCite{1}+{}{}{arxiv:2105.07028} definitions and diagrams \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.4605654}. \markdownRendererInterblockSeparator
{}RO-Crate acts therefore as an interoperability layer between registries, repositories and users in WorkflowHub. The iterative development between WorkflowHub developers and the RO-Crate community heavily informed the creation of the Bioschemas \markdownRendererCite{1}+{}{}{bioschemas_2017} profile for \markdownRendererLink{Computational Workflows}{https://bioschemas.org/profiles/ComputationalWorkflow/1.0-RELEASE/}{https://bioschemas.org/profiles/ComputationalWorkflow/1.0-RELEASE/}{}, which again informed the \markdownRendererLink{RO-Crate 1.1 specification on workflows}{https://www.researchobject.org/ro-crate/1.1/workflows.html}{https://www.researchobject.org/ro-crate/1.1/workflows.html}{} and led to the RO-Crate Python library \markdownRendererCite{1}+{}{}{ro-crate-py} and WorkflowHub’s \markdownRendererLink{\markdownRendererStrongEmphasis{Workflow RO-Crate profile}}{https://about.workflowhub.eu/Workflow-RO-Crate/}{https://about.workflowhub.eu/Workflow-RO-Crate/}{}, which, in a similar fashion to RO-Crate itself, recommends which workflow resources and descriptions are required. This co-development across project boundaries exemplifies the drive for simplicity and for establishing best practices.\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Profile for recording workflow runs}\markdownRendererInterblockSeparator
{}While RO-Crates in WorkflowHub so far have been focused on workflows that are ready to be run, development of WorkflowHub are now creating a \markdownRendererStrongEmphasis{Workflow Run RO-Crate profile} for the purposes of benchmarking, testing and executing workflows. As such, RO-Crate serves as a container of both a \markdownRendererEmphasis{workflow definition} that may be executed and of a particular \markdownRendererEmphasis{workflow execution with test results}. This profile is a continuation of our previous work with capturing workflow provenance in a Research Object in CWLProv \markdownRendererCite{1}+{}{}{doi:10.1093/gigascience/giz095} and TavernaPROV \markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.51314}. In both cases, we used the PROV Ontology \markdownRendererCite{1}+{}{}{PROVO}, including details of every task execution with all the intermediate data, which required significant workflow engine integration\markdownRendererFootnote{CWLProv and TavernaProv predate RO-Crate, but use RO-Bundle\markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.12586}, a similar Research Object packaging method with JSON-LD metadata. }. To simplify from that approach, for this Workflow Run RO-Crate profile we will use a higher level \markdownRendererLink{schema.org provenance}{https://www.researchobject.org/ro-crate/1.1/provenance.html#software-used-to-create-files}{https://www.researchobject.org/ro-crate/1.1/provenance.html#software-used-to-create-files}{} for the input/output boundary of the overall workflow execution. This \markdownRendererEmphasis{Level 1 workflow provenance} \markdownRendererCite{1}+{}{}{doi:10.1093/gigascience/giz095} can be expressed generally across workflow languages with minimal engine changes, with the option of more detailed provenance traces as separate PROV resources in the RO-Crate. \markdownRendererInterblockSeparator
{}WorkflowHub has recently enabled minting of Digital Object Identifiers (DOIs), a PID commonly used for scholarly artefacts, for registered workflows, e.g. \markdownRendererCodeSpan{10.48546/workflowhub.workflow.56.1} \markdownRendererCite{1}+{}{}{doi:10.48546/workflowhub.workflow.56.1}, lowering the barrier for citing workflows as computational methods along with their FAIR metadata – captured within an RO-Crate. While it is not an aim for WorkflowHub to be a repository of workflow runs and their data, RO-Crates of \markdownRendererEmphasis{exemplar workflow runs} serve as useful workflow documentation, as well as being an exchange mechanism that preserve FAIR metadata in a diverse workflow execution environment.\markdownRendererInterblockSeparator
{}\markdownRendererHeadingThree{Profile for testing workflows}\markdownRendererInterblockSeparator
{}The value of computational workflows, however, is potentially undermined by the "collapse" over time of the software and services they depend upon: for instance, software dependencies can change in a non-backwards-compatible manner, or active maintenance may cease; an external resource, such as a reference index or a database query service, could shift to a different URL or modify its access protocol; or the workflow itself may develop hard-to-find bugs as it is updated. This can take a big toll on the workflow's reusability and on the reproducibility of any processes it evokes.\markdownRendererInterblockSeparator
{}For this reason, WorkflowHub is complemented by a monitoring and testing service called LifeMonitor\markdownRendererCite{1}+{}{}{about-lifemonitor}, also supported by EOSC-Life. LifeMonitor's main goal is to assist in the creation, periodic execution and monitoring of workflow tests, enabling the early detection of software collapse in order to minimise its detrimental effects. The communication of metadata related to workflow testing is achieved through the adoption of a \markdownRendererLink{\markdownRendererStrongEmphasis{Workflow Testing RO-Crate profile}}{https://crs4.github.io/life_monitor/workflow_testing_ro_crate}{https://crs4.github.io/life_monitor/workflow_testing_ro_crate}{} stacked on top of the \markdownRendererEmphasis{Workflow RO-Crate} profile. This further specialisation of Workflow RO-Crate allows to specify additional testing-related entities (test suites, instances, services, etc.), leveraging \markdownRendererLink{RO-Crate's extension mechanism}{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#extending-ro-crate}{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#extending-ro-crate}{} through the addition of terms from custom namespaces.\markdownRendererInterblockSeparator
{}In addition to showcasing RO-Crate's extensibility, the testing profile is an example of the format's flexibility and adaptability to the different needs of the research community. Though ultimately related to a computational workflow, in fact, most of the testing-specific entities are more about describing a protocol for interacting with a monitoring service than a set of research outputs and its associated metadata. Indeed, one of LifeMonitor's main functionalities is monitoring and reporting on test suites running on existing Continuous Integration (CI) services, which is described in terms of service URLs and job identifiers in the testing profile. In principle, in this context, data could disappear altogether, leading to an RO-Crate consisting entirely of contextual entities. Such an RO-Crate acts more as an exchange format for communication between services (WorkflowHub and LifeMonitor) than as an aggregator for research data and metadata, providing a good example of the format's high versatility.\relax

### 2021-08-08 _markdown_ro-crate/f8f3375d5d023f4f211e1a224d813c5f.md.tex
Merge overleaf-2021-08-08-0724 into master [ci skip]
This build is based on
https://github.com/stain/ro-crate-paper/commit/ddb80a8f0dac3a15d4db06d53b8e5463e720cc7b.

This commit was created by the following CI build and job:
https://github.com/stain/ro-crate-paper/commit/ddb80a8f0dac3a15d4db06d53b8e5463e720cc7b/checks
https://github.com/stain/ro-crate-paper/runs/1109599515

\markdownRendererHeadingTwo{Institutional data repositories – Harvard Data Commons}\markdownRendererInterblockSeparator
{}The concept of a Data Commons for research collaboration was originally defined as "cyber-infrastructure that co-locates data, storage, and computing infrastructure with commonly used tools for analysing and sharing data to create an interoperable resource for the research community" \markdownRendererCite{1}+{}{}{doi:10.1109/MCSE.2016.92}. More recently, it was established to integrate active data-intensive research with data management and archival best practices. It facilitates research by providing computational infrastructure where researchers can use, share and store data, software, workflows and other digital artefacts used in their studies. Furthermore, the Commons feature tools and services, such as computation clusters and storage for scalability, data repositories for disseminating and preserving regular, but also large or sensitive datasets, and other research assets. Multiple initiatives were undertaken to create Data Commons on national, research, and institutional levels. For example, the \markdownRendererLink{Australian Research Data Commons (ARDC)}{https://ardc.edu.au}{https://ardc.edu.au}{} \markdownRendererCite{1}+{}{}{doi:10.5334/dsj-2019-044} is a national initiative that enables local researchers and industries to access computing infrastructure, training, and curated datasets for data-intensive research. NCI’s \markdownRendererLink{Genomic Data Commons}{https://gdc.cancer.gov/}{https://gdc.cancer.gov/}{} (GDC) \markdownRendererCite{1}+{}{}{doi:10.1182/blood-2017-03-735654} provides the cancer research community with access to a vast volume of genomic and clinical data. Initiatives such as \markdownRendererLink{Research Data Alliance (RDA) Global Open Research Commons}{https://www.rd-alliance.org/groups/global-open-research-commons-ig}{https://www.rd-alliance.org/groups/global-open-research-commons-ig}{} propose standards on the implementation of Data Commons to avoid them becoming "data silos" and enable interoperability from one Data Commons to another.\markdownRendererInterblockSeparator
{}\markdownRendererStrongEmphasis{Harvard Data Commons} \markdownRendererCite{1}+{}{}{doi:10.7557/5.5422} aims to address data access and reuse challenges of cross-disciplinary research within a research institution. It brings together multiple institutional schools, libraries, computing centres and the \markdownRendererLink{Harvard Dataverse data repository}{https://dataverse.harvard.edu/}{https://dataverse.harvard.edu/}{}. \markdownRendererLink{Dataverse}{https://dataverse.org/}{https://dataverse.org/}{} \markdownRendererCite{1}+{}{}{doi:10.1045/january2011-crosas} is a free and open-source software platform to archive, share and cite research data. The Harvard Dataverse repository is the largest of 70 installations worldwide, containing over 100K datasets with about 1M data files. Toward the goal of facilitating collaboration and data discoverability and management within the university, Harvard Data Commons has the following primary objectives:\markdownRendererInterblockSeparator
{}\markdownRendererOlBeginTight
\markdownRendererOlItemWithNumber{1}integrating Harvard Research Computing with Harvard Dataverse by leveraging Globus endpoints \markdownRendererCite{1}+{}{}{doi:10.1109/MCC.2014.52} that will allow an automatic transfer of large datasets to the repository. In some cases, only the metadata will be transferred while the data stays stored in remote storage;\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{2}supporting advanced research workflows and providing packaging options for assets such as code and workflows in the Harvard Dataverse repository to enable reproducibility and reuse, and \markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{3}integrating repositories supported by Harvard, which are \markdownRendererLink{DASH}{https://dash.harvard.edu}{https://dash.harvard.edu}{}, the open access institutional repository, the Digital Repository Services (DRS) for preserving digital asset collections, and the Harvard Dataverse.\markdownRendererOlItemEnd 
\markdownRendererOlEndTight \markdownRendererInterblockSeparator
{}Particularly relevant to this paper is the second objective of the Harvard Data Commons, which aims to support the deposit of research artefacts to Harvard Dataverse with sufficient information in the metadata to allow their future reuse (Figure~\ref{fig:hdc}). Considering the requirements of incorporating data, code, and other artefacts from various institutional infrastructures, Harvard Data Commons is currently working on RO-Crate adaptation. The RO-Crate metadata provides the necessary structure to make all research artefacts FAIR. The Dataverse software already has extensive support for metadata, including the Data Documentation Initiative (DDI), Dublin Core, DataCite, and Schema.org. Incorporating RO-Crate, which has the flexibility to describe a wide range of research resources, will facilitate their seamless transition from one infrastructure to the other within the Harvard Data Commons.\markdownRendererInterblockSeparator
{}\markdownRendererImage{Harvard Data Commons}{../content/images/data-commons-ro-crate-figure-2.pdf}{../content/images/data-commons-ro-crate-figure-2.pdf}{\textbf{One aspect of Harvard Data Commons}. Automatic encapsulation and deposit of artefacts from data management tools used during active research at the Harvard Dataverse repository.}{#fig:hdc}\markdownRendererInterblockSeparator
{}Even though the Harvard Data Commons is specific to Harvard University, the overall vision and the three objectives can be abstracted and applied to other universities or research organisations. The Commons will be designed and implemented using standards and commonly-used approaches to make it interoperable and reusable by others.\relax

### 2021-08-09 output/_markdown_ro-crate/42df5638f5060e54b0e41156879c39de.md.tex

\markdownRendererHeadingOne{Conclusion}\markdownRendererInterblockSeparator
{}RO-Crate provides a lightweight approach to packaging digital research artefacts with structured metadata, assisting developers and researchers to produce and consume FAIR data archives of their ROs, including PIDs, context and provenance of research artefacts. Aggregated data may be large and distributed, or located in regular folders on a file system. \markdownRendererInterblockSeparator
{}As a set of best practice recommendations, developed by an open and broad community, RO-Crate shows how to use "just enough" Linked Data standards in a consistent way, with structured metadata using a rich base vocabulary that can cover general-purpose contextual relations, whilst retaining extensibility to domain- and application-specific uses. \markdownRendererInterblockSeparator
{}The adoption of simple web technologies in the RO-Crate specification has helped a rapid development of a wide variety of supporting open source tools and libraries. RO-Crate fits into the larger landscape of open scholarly communication and FAIR Digital Object infrastructure, and can be integrated into data repository platforms. RO-Crate can be applied as a data/metadata exchange mechanism, assist in long-term archival preservation of metadata and data, or simply used at small-scale by individual researchers. Thanks to its strong community support, new and improved profiles and tools are continuously added to the RO-Crate tooling landscape, making it easier for adopters to find examples and support for their own use case.\relax

### 2021-08-09 output/_markdown_ro-crate/48ea55320d214e1130b767589109006d.md.tex

\markdownRendererHeadingOne{Acknowledgements}\markdownRendererInterblockSeparator
{}\small\markdownRendererInterblockSeparator
{}This work has received funding from the European Commission's Horizon 2020 research and innovation programme for projects \markdownRendererLink{BioExcel-2}{https://cordis.europa.eu/project/id/823830}{https://cordis.europa.eu/project/id/823830}{} (H2020-INFRAEDI-2018-1 823830), \markdownRendererLink{IBISBA}{https://cordis.europa.eu/project/id/730976}{https://cordis.europa.eu/project/id/730976}{} (H2020-INFRAIA-2017-1-two-stage 730976, H2020-INFRADEV-2019-2 871118), \markdownRendererLink{EOSC-Life}{https://cordis.europa.eu/project/id/824087}{https://cordis.europa.eu/project/id/824087}{} (H2020-INFRAEOSC-2018-2 824087), \markdownRendererLink{SyntheSys+}{https://cordis.europa.eu/project/id/823827}{https://cordis.europa.eu/project/id/823827}{} (H2020-INFRAIA-2018-1 823827)\markdownRendererInterblockSeparator
{}A. Trisovic is funded by the Alfred P. Sloan Foundation (grant number P-2020-13988). Harvard Data Commons is supported by an award from Harvard University Information Technology (HUIT).\markdownRendererInterblockSeparator
{}\markdownRendererHeadingTwo{Contributions}\markdownRendererInterblockSeparator
{}Author contributions to this article and the RO-Crate projet according to the Contributor Roles Taxonomy \markdownRendererLink{CASRAI CrEDiT}{https://casrai.org/credit/}{https://casrai.org/credit/}{} \markdownRendererCite{1}+{}{}{doi:10.1087/20150211}:\markdownRendererInterblockSeparator
{}\markdownRendererDlBegin
\markdownRendererDlItem{Stian Soiland-Reyes}\markdownRendererDlDefinitionBegin Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Software, Visualization, Writing – original draft, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Peter Sefton}\markdownRendererDlDefinitionBegin Conceptualization, Investigation, Methodology, Project administration, Resources, Software, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Mercè Crosas}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Leyla Jael Castro}\markdownRendererDlDefinitionBegin Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Frederik Coppens}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{José M. Fernández}\markdownRendererDlDefinitionBegin Methodology, Software, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Daniel Garijo}\markdownRendererDlDefinitionBegin Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Marco La Rosa}\markdownRendererDlDefinitionBegin Software, Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Simone Leo}\markdownRendererDlDefinitionBegin Software, Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Eoghan Ó Carragáin}\markdownRendererDlDefinitionBegin Investigation, Methodology, Project administration, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Marc Portier}\markdownRendererDlDefinitionBegin Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Ana Trisovic}\markdownRendererDlDefinitionBegin Software, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{RO-Crate Community}\markdownRendererDlDefinitionBegin Investigation, Software, Validation, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Paul Groth}\markdownRendererDlDefinitionBegin Methodology, Supervision, Writing – original draft, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Carole Goble}\markdownRendererDlDefinitionBegin Conceptualization, Funding acquisition, Methodology, Project administration, Supervision, Visualization, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd 
\markdownRendererDlEnd\markdownRendererInterblockSeparator
{}We would also like to thank for contributions from:\markdownRendererInterblockSeparator
{}\markdownRendererDlBegin
\markdownRendererDlItem{Finn Bacall}\markdownRendererDlDefinitionBegin Software, Methodology\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Herbert Van de Sompel}\markdownRendererDlDefinitionBegin Writing – review \& editing \markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Ignacio Eguinoa}\markdownRendererDlDefinitionBegin Software, Methodology\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Nick Juty}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Oscar Corcho}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Stuart Owen}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Laura Rodríguez-Navas}\markdownRendererDlDefinitionBegin Software, Visualization, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd 
\markdownRendererDlEnd\markdownRendererInterblockSeparator
{}% CrEDiT taxonomy: % % Conceptualization % Data curation % Formal Analysis % Funding acquisition % Investigation % Methodology % Project administration % Resources % Software % Supervision % Validation % Visualization % Writing – original draft % Writing – review \& editing\relax

### 2021-08-09 output/_markdown_ro-crate/fbbd20ed29c69f8d0d09a7128f17add8.md.tex

\markdownRendererHeadingOne{Related Work}\markdownRendererInterblockSeparator
{}With the increasing digitalisation of research processes, there has been a significant call for the wider adoption of interoperable sharing of data and its associated metadata. For a comprehensive overview and recommendations, in particular for data, we refer to \markdownRendererCite{1}+{}{}{doi:10.1016/j.patter.2020.100136}, which highlights the wide variety of metadata and documentation that the literature prescribes for enabling data reuse. \markdownRendererInterblockSeparator
{}Here we focus on approaches for bundling research artefacts along with their metadata. This notion has a long history behind it \markdownRendererCite{1}+{}{}{doi:10.1190/1.1822162}, but recent approaches have followed three main strands: 1) publishing to centralised repositories; 2) packaging approaches similar to RO-Crate; and 3) bundling the computational workflow around a scientific experiment. \markdownRendererInterblockSeparator
{}% ## Repositories % % * DataCite Metadata \markdownRendererCite{1}+{}{}{doi:10.14454/3w3z-sa82} % * Open Science Framework\markdownRendererInterblockSeparator
{}\markdownRendererHeadingTwo{Bundling and Packaging Digital Research Artefacts}\markdownRendererInterblockSeparator
{}The challenge of describing computational workflows was one of the main motivations for the early proposal of \markdownRendererEmphasis{Research Objects} (RO) \markdownRendererCite{1}+{}{}{doi:10.1016/j.future.2011.08.004} as first-class citizens for sharing and publishing. The RO approach involves bundling datasets, workflows, scripts and results along with traditional dissemination materials like journal articles and presentations, forming a single package. Crucially, these resources are not just gathered, but also individually typed, described and related to each other using semantic vocabularies. As pointed out in \markdownRendererCite{1}+{}{}{doi:10.1016/j.future.2011.08.004} an open-ended \markdownRendererEmphasis{Linked Data} approach is not sufficient for scholarly communication: a common data model is also needed in addition to common and best practices for managing and annotating lifecycle, ownership, versioning and attributions.\markdownRendererInterblockSeparator
{}Considering the FAIR principles, we can say with hindsight that the initial RO approaches strongly targeted \markdownRendererEmphasis{Interoperability}, with a particular focus on the reproducibility of \markdownRendererEmphasis{in-silico experiments} involving computational workflows and the reuse of existing RDF vocabularies. \markdownRendererInterblockSeparator
{}The first implementation of Research Objects for sharing workflows in myExperiment \markdownRendererCite{1}+{}{}{doi:10.1093/nar/gkq429} was based on RDF ontologies \markdownRendererCite{1}+{}{}{newman2009}, building on Dublin Core, FOAF, SIOC, Creative Commons and OAI-ORE to form myExperiment ontologies for describing social networking, attribution and credit, annotations, aggregation packs, experiments, view statistics, contributions, and workflow components \markdownRendererCite{1}+{}{}{myExperimentOntology2009}.\markdownRendererInterblockSeparator
{}% * RO-Bag-It % * BDBag % * Ontologies ( OAI-ORE, OAC/AO/OA)\markdownRendererInterblockSeparator
{}\markdownRendererHeadingTwo{FAIR Digital Objects}\markdownRendererInterblockSeparator
{}FAIR Digital Objects (FDO) \markdownRendererCite{1}+{}{}{doi:10.3390/publications8020021} have been proposed as a conceptual framework for making digital resources available in a Digital Objects (DO) architecture that encourages active use of the objects and their metadata. In particular, an FDO has five parts: (i) The FDO \markdownRendererEmphasis{content}, bit sequences stored in an accessible repository; (ii) a \markdownRendererEmphasis{Persistent Identifier} (PID) such as a DOI that identifies the FDO and can resolve these parts; (iii) Associated rich \markdownRendererEmphasis{metadata}, as separate FDOs; (iv) Type definitions, also separate FDOs; (v) Associated \markdownRendererEmphasis{operations} for the given types. A Digital Object typed as a Collection aggregates other DOs by reference.\markdownRendererInterblockSeparator
{}As an "\markdownRendererLink{abstract protocol}{https://www.dona.net/sites/default/files/2018-11/DOIPv2Spec_1.pdf}{https://www.dona.net/sites/default/files/2018-11/DOIPv2Spec_1.pdf}{}", DOs could be implemented in multiple ways. One suggested implementation is the \markdownRendererLink{FAIR Digital Object Framework}{https://fairdigitalobjectframework.org/}{https://fairdigitalobjectframework.org/}{}, based on HTTP and the Linked Data Principles. While there is agreement on using PIDs based on DOIs, consensus on how to represent common metadata, core types and collections as FDOs has not yet been reached. We argue that RO-Crate can play an important role for FDOs:\markdownRendererInterblockSeparator
{}\markdownRendererOlBeginTight
\markdownRendererOlItemWithNumber{1}By providing a predictable and extensible serialisation of structured metadata.\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{2}By formalising how to aggregate digital objects as collections (and adding their context).\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{3}By providing a natural Metadata FDO in the form of the RO-Crate Metadata File.\markdownRendererOlItemEnd 
\markdownRendererOlItemWithNumber{4}By being based on Linked Data and schema.org vocabulary, meaning that PIDs already exist for common types and properties.\markdownRendererOlItemEnd 
\markdownRendererOlEndTight \markdownRendererInterblockSeparator
{}At the same time, it is clear that the goal of FDO is broader than that of RO-Crate; namely, FDOs are active objects with distributed operations, and add further constraints such as PIDs for every element. These features improve FAIR features of digital objects and are also useful for RO-Crate, but they also severely restrict the infrastructure that needs to be implemented and maintained in order for FDOs to remain available. RO-Crate, on the other hand, is more flexible: it can minimally be used within any file system structure, or ideally exposed through a range of Web-based scenarios. A \markdownRendererEmphasis{FAIR profile of RO-Crate} (e.g. enforcing PID usage) will fit well within a FAIR Digital Object ecosystem.\markdownRendererInterblockSeparator
{}\markdownRendererHeadingTwo{Packaging Workflows}\markdownRendererInterblockSeparator
{}The use of computational workflows, typically combining a chain of open source tools in an analytical pipeline, has gained prominence, in particular in the life sciences. Workflows may have initially been used to improve computational scalability, but they also assist in making computed data results FAIR \markdownRendererCite{1}+{}{}{doi:10.1162/dint_a_00033}, for instance by improving reproducibility \markdownRendererCite{1}+{}{}{10.1016/j.future.2017.01.012}, but also because programmatic data usage help propagate their metadata and provenance \markdownRendererCite{1}+{}{}{doi:10.1002/cpe.1228}. At the same time, however, workflows raise additional FAIR challenges, since they can be considered important research artefacts themselves, posing the problem of capturing and explaining the computational methods behind the analysis they perform \markdownRendererCite{1}+{}{}{doi:10.3233/DS-190026}.\markdownRendererInterblockSeparator
{}Even when researchers follow current best practices for workflow reproducibility, \markdownRendererCite{1}+{}{}{doi:10.1016/j.cels.2018.03.014} \markdownRendererCite{1}+{}{}{doi:10.1016/j.future.2017.01.012} the communication of outcomes through traditional academic publishing routes relying on a textual representation adds barriers that hinder reproducibility and FAIR use of the knowledge previously captured in the workflow.\markdownRendererInterblockSeparator
{}As a real-life example, let us look at a metagenomics article \markdownRendererCite{1}+{}{}{doi:10.1038/s41586-019-0965-1} where the authors have gone to extraordinary efforts to document the individual tools that have been reused, including their citations, versions, settings, parameters and combinations. The \markdownRendererEmphasis{Methods} section is 2 pages in tight double-columns with 24 additional references, supported by the availability of data on an FTP server (60 GB) \markdownRendererCite{1}+{}{}{ebi_ftp_umgs2019} and of open source code in GitHub \markdownRendererLink{Finn-Lab/MGS-gut}{https://github.com/Finn-Lab/MGS-gut}{https://github.com/Finn-Lab/MGS-gut}{} \markdownRendererCite{1}+{}{}{finn-lab-mgsgut}, including the pipeline as shell scripts and associated analysis scripts in R and Python.\markdownRendererInterblockSeparator
{}This attention to reporting detail for computational workflows is unfortunately not yet the norm, and although bioinformatics journals have strong \markdownRendererEmphasis{data availability} requirements, they frequently do not require authors to include or cite \markdownRendererEmphasis{software, scripts and pipelines} used for analysing and producing results \markdownRendererCite{1}+{}{}{soilandreyes_tweet_2020} – rather, authors might be penalised for doing so [cite?] as it would detrimentally count against arbitrary limits on number of pages and references.\markdownRendererInterblockSeparator
{}However detailed this additional information might be, another researcher who wants to reuse a particular computational method may first want to assess if the described tool or workflow is Re-runnable (executable at all), Repeatable (same results for original inputs on same platform), Reproducible (same results for original inputs with different platform or newer tools) and ultimately Reusable (similar results for different input data), Repurposable (reusing parts of the method for making a new method) or Replicable (rewriting the workflow following the method description). \markdownRendererCite{1}+{}{}{doi:10.3389/fninf.2017.00069}\markdownRendererCite{1}+{}{}{goble_presentation_2016}\markdownRendererInterblockSeparator
{}Following the textual description alone, researchers would be forced to jump straight to evaluate "Replicable" by rewriting the pipeline from scratch. This can be expensive and error-prone. They would firstly need to install all the software dependencies and download reference datasets. This can be a daunting task in and of itself, which may have to be repeated multiple times as workflows typically are developed at small scale on desktop computers, scaled up to local clusters, and potentially put into production using cloud instances, each of which will have different requirements for software installations.\markdownRendererInterblockSeparator
{}In recent years the situation has been greatly improved by software packaging and container technologies like Docker and Conda, which have seen increased adoption in life sciences \markdownRendererCite{1}+{}{}{doi:10.1007/s41019-017-0050-4} thanks to collaborative efforts such as BioConda \markdownRendererCite{1}+{}{}{doi:10.1038/s41592-018-0046-7} and BioContainers \markdownRendererCite{1}+{}{}{doi:10.1093/bioinformatics/btx192}, and support by Linux distributions (e.g. Debian Med \markdownRendererCite{1}+{}{}{doi:10.1186/1471-2105-11-S12-S5}). As of May 2021, more than 7000 software packages are available \markdownRendererLink{in BioConda alone}{https://anaconda.org/bioconda/}{https://anaconda.org/bioconda/}{}, and 9000 containers \markdownRendererLink{in BioContainers}{https://biocontainers.pro/#/registry}{https://biocontainers.pro/#/registry}{}. Docker and Conda have gained integration in workflow systems such as Snakemake \markdownRendererCite{1}+{}{}{doi:10.1093/bioinformatics/bts480}, Galaxy \markdownRendererCite{1}+{}{}{doi:10.1093/nar/gky379} and Nextflow \markdownRendererCite{1}+{}{}{doi:10.1038/nbt.3820}, meaning a downloaded workflow definition can now be executed on a "blank" machine (except for the workflow engine) with the underlying analytical tools installed on demand – but even here there is a reproducibility challenge, for instance \markdownRendererLink{Docker Hub's retention policy will expire container images after 6 months}{https://www.docker.com/blog/docker-hub-image-retention-policy-delayed-and-subscription-updates/}{https://www.docker.com/blog/docker-hub-image-retention-policy-delayed-and-subscription-updates/}{}, or lack of recording versions of transitive dependencies of Conda packages could cause incompatibilities if the packages are subsequently updated. Except for brief metadata in their repositories, these containers and packages do not capture any semantic relationships of their content – rather their opaqueness and wrapping of arbitrary binary tools makes such relationships harder to find.\markdownRendererInterblockSeparator
{}From this we see that computational workflows are themselves complex digital objects that needs to be recorded not just as files, but in the context of their execution environment, dependencies and analytical purpose in research – as well as their FAIR metadata (e.g. version, license, attribution and identifiers).\relax

### 2021-08-11 latex/ro-crate.tex
Ana's text

@@
 % From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
+\usepackage{fontspec}
+
 \usepackage{amsmath}
@@
 
-\usepackage{fontspec}
 \usepackage{unicode-math}
@@
     \snm{Sefton}\ead[label=ePS]{Peter.Sefton@uts.edu.au}\ead[label=oPS,orcid]{0000-0002-3545-944X}},
-\author[EOC]{\inits{E.}\fnms{Eoghan}
-    \snm{Ó Carragáin}\ead[label=eEOC]{eoghan.ocarragain@ucc.ie}\ead[label=oEOC,orcid]{0000-0001-8131-2150}},
 \author[MC]{\inits{M.}\fnms{Mercè}
@@
     \snm{Garijo}\ead[label=eDG]{dgarijo@fi.upm.es}\ead[label=oDG,orcid]{0000-0003-0454-7145}},
+\author[BG]{\inits{B.}\fnms{Björn}
+    \snm{Grüning}\ead[label=eBG]{gruening@informatik.uni-freiburg.de}\ead[label=oBG,orcid]{0000-0002-3079-6586}},
 \author[MLR]{\inits{M.}\fnms{Marco}
@@
     \snm{Leo}\ead[label=eSL]{simone.leo@crs4.it}\ead[label=oSL,orcid]{0000-0001-8271-5429}},
+\author[EOC]{\inits{E.}\fnms{Eoghan}
+    \snm{Ó Carragáin}\ead[label=eEOC]{eoghan.ocarragain@ucc.ie}\ead[label=oEOC,orcid]{0000-0001-8131-2150}},
 \author[MP]{\inits{M.}\fnms{Marc}
@@
     \orgname{Universidad Politécnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{oDG}}
-\address[MLR]{\printead{oMLR}}
+\address[BG]{Bioinformatics Group, Department of Computer Science,
+    \orgname{Albert-Ludwigs-University Freiburg}, Freiburg, \cny{Germany}\printead[presep={\\}]{oBG}}
+\address[MLR]{\orgname{PARADISEC}, Melbourne, \cny{Australia}\printead[presep={\\}]{oMLR}}
 \address[SL]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA),

### 2021-08-12 content/80.formal-definition.tex
data commons figure as svg
Synchronized with Google Docs version 2021-08-12 to overleaf
.. and a couple more edits

https://docs.google.com/document/d/11toi7xIDiDDeOq4fA_pr1GUDf0julX0Q/edit?usp=sharing&ouid=108141746798163896498&rtpof=true&sd=true

@@
 % Markdown setions within.
+% 
+% Edit here before syncing to 80.formal-definition.md for HTML version
 
@@
 && license(\texttt{<http://example.com/ro/123/>}, \\ 
-&& \ \ \ \ \ \texttt{<https://creativecommons.org/licenses/by-nc-sa/3.0/au/>}) \\ 
+&& \ \ \ \ \ \texttt{<https://spdx.org/licenses/CC-BY-4.0>}) \\ 
 && datePublished(\texttt{<http://example.com/ro/123/>}, \text{“2017-02-23”}) \\ 
@@
 \\
-&& ContextualEntity(\texttt{<https://creativecommons.org/licenses/by-nc-sa/3.0/au/>}) \\
-&& name(\texttt{<https://creativecommons.org/licenses/by-nc-sa/3.0/au/>},  \\
-&& \ \ \ \ \  \text{“Attribution-NonCommercial-ShareAlike 3.0 Australia (CC BY-NC-SA 3.0 AU)”}) \\
+&& ContextualEntity(\texttt{<https://spdx.org/licenses/CC-BY-4.0>}) \\
+&& name(\texttt{<https://spdx.org/licenses/CC-BY-4.0>},  \\
+&& \ \ \ \ \  \text{Creative Commons Attribution 4.0”}) \\
+\\
+&& ContextualEntity(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>}) \\
+&& name(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>},  \\
+&& \ \ \ \ \  \text{Creative Commons Attribution Non Commercial 4.0”}) \\
 \\
@@
 && name(\texttt{<http://example.com/ro/123/interviews/>},  \\
-&& \ \ \ \ \  \text{“Audio recordings of care provider interviews”}) 
+&& \ \ \ \ \  \text{“Audio recordings of care provider interviews”}) \\
+&& license(\texttt{<http://example.com/ro/123/interviews/>}, \\ 
+&& \ \ \ \ \ \texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>})
 \end{eqnarray*}
 
-In reality many additional attributes from schema.org types like \url{http://schema.org/Dataset} and \url{http://schema.org/CreativeWork} would be used to further describe the RO-Crate and its entities, but as these are optional they do not form part of this formalisation.
+Notable from this triple-like formalization is that a RO-Crate $R$ is fully represented as a tree at depth 2 helped by the use of $𝕀𝕣𝕚$ nodes.
+For instance the aggregation from the root entity $hasPart(…interviews/>)$ is at same level as the data entity’s property $license(…CC-BY-NC-4.0>)$ and that contextual entity’s attribute name $(…Non Commercial 4.0”)$.
+As shown in section RO-Crate JSON-LD, the RO-Crate Metadata File serialization is an equivalent shallow tree, although at depth 3 to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.
+
+In reality many additional attributes and contextual types from schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (_SHOULD_ requirements) they do not form part of this formalization
 
@@
 Note that in the JSON-LD serialization of RO-Crate, the expression of $Class$ and $Property$ is typically indirect: The JSON-LD \texttt{@context} maps to schema.org IRIs, which, when resolved as Linked Data, embed their formal definition as RDFa.
+Extensions may however include such term definitions directly in the RO-Crate.
 
@@
 
-An important RO-Crate principle is that of being \textbf{self-describing} Therefore the serialisation of the RO-Crate into a file should also describe itself in a \href{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#ro-crate-metadata-file-descriptor}{Metadata File Descriptor}, indicating it is $about$ (describing) the RO-Crate root data entity, and that it $conformsTo$ a particular version of the RO-Crate specification:
+An important RO-Crate principle is that of being \textbf{self-described} Therefore the serialisation of the RO-Crate into a file should also describe itself in a \href{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#ro-crate-metadata-file-descriptor}{Metadata File Descriptor}, indicating it is $about$ (describing) the RO-Crate root data entity, and that it $conformsTo$ a particular version of the RO-Crate specification:
 
@@
 
-Note that although the metadata file necessarily is an \emph{information resource} written to disk or served over the network (e.g.
-as JSON-LD), it is not considered to be a contained \emph{part} of the RO-Crate in the form of a \emph{data entity}, rather it is described only as a \emph{contextual entity}.
+Note that although the metadata file necessarily is an \emph{information resource} written to disk or served over the network (as JSON-LD), it is not considered to be a contained \emph{part} of the RO-Crate in the form of a \emph{data entity}, rather it is described only as a \emph{contextual entity}.
 
-While in the conceptual model the \emph{RO-Crate Metadata File} can be seen as the top-level node that describes the \emph{RO-Crate Root}, in the formal model (and the JSON-LD format) the metadata file descriptor is an additional contextual entity and not affecting the depth-limit of the RO-Crate.
+In the conceptual model the \emph{RO-Crate Metadata File} can be seen as the top-level node that describes the \emph{RO-Crate Root}, however in the formal model (and the JSON-LD format) the metadata file descriptor is an additional contextual entity that is not affecting the depth-limit of the RO-Crate.
 
@@
 These production rules first grow the graph of $R$ by adding a transitive rule – anything described in $R$ which is related to $o$, means that $o$ is also mentioned by the $ROCrate(R)$.
-For simplicity this rule is one-way; in practice the graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities.
+For simplicity this rule is one-way; in theory the graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities, but these are proposed to be bound to the root data entity with schema.org relation \href{http://schema.org/mentions}.

### 2021-08-13 content/80.formal-definition.tex

@@
 
-The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or \href{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities}{both}; and must be typed with at least one $Class(e)$.
+The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or both\footnote\url{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities}}; and must be typed with at least one $Class(e)$.
 

### 2021-10-03 response-to-review.md
response to review document

# Response to peer review, Data Science submission 706-1686

Response to  open peer review at <https://datasciencehub.net/paper/packaging-research-artefacts-ro-crate>

## Reviewer 1

_By Angelo A. <https://orcid.org/0000-0002-4763-3943>_

>  I will leave the latter to its future users in defining how it can be improved, based on their needs.

> I encourage the authors to add section numbers. Without them, it becomes hard to even scroll the document and reach the right location.

> Typos:
> - In section RO-Crate. “abuot” -> “about”
> - Figure 1 caption. “The RO is described within a ,”. The object of the sentence is missing.
> - In Data Entities are described using Contextual Entities section, you wrote “person, but secondarily also a web page and their academic work.” I think the sentence is incomplete.
> - In Table 1, the tool SCHeMa has the squared parenthesis signalling a citation, but there are question marks.
> - In Machine-actionable Data Management Plans section, there is a left curly bracket ({)
> - In Institutional data repositories – Harvard Data Commons, the reference to figure 3 is missing
> - In Institutional data repositories – Harvard Data Commons (PDF version) figure 3 is missing. In the HTML version, the figure 3 is available.
> - In Packaging Workflows, there is `[cite?]` signalling a missing citation.

> While reading the paper, I was eager to get to the evaluation section. At some point, in the Ensuring simplicity section, you state “One aim of RO-Crate is to be conceptually simple. > This simplicity has been repeatedly checked and confirmed through a community review process.”
> I was indeed curious to see if you had run a survey on what users thought about RO-Crate. Perhaps, a survey gathering feedback on the user experience. Also, the concept of simplicity is bound with the skills of the practitioner. To what extent, they found it simple?

> Also, I trust that RO-Crate is far from being complete (actually, you are in the process of releasing v1.2). What do you think needs to be improved in RO-Crate? What other new features do you plan to include? How much of this is shaped by the end-users?

## Reviewer 2

_By Alessia Bardi <https://orcid.org/0000-0002-1112-1292>_

> The most important remark is that I had not a good experience with the RO-Crate of this publication.
> I clicked on supplementary files 1128.html and went to https://datasciencehub.net/system/files/ds-supplementary-706-1128.html#r...
> - conformsTo is missing
> - the link to download returns 404
> - the RO-Crate check says it is not a valid RO-Crate
> The same happens for the version deposited on Zenodo (https://doi.org/10.5281/zenodo.5146227) linked in the abstract.

Section "Conceptual definition":
- Change ';' with ':' in "Figure 1 shows the main conceptual elements involved in an RO-Crate;"
- Figure 1: check the caption, there are missing words at "The RO is described within a , "
- Conceptual definition does not clarify what's the difference between a generic Research Object (as introduced in the previous section) and a RO-Crate

Section "Linked data as a foundation"
- Move the footnote about IRIs the first time you use the acronym, if you think the footnote is needed

Section "Data Entities are described using Contextual Entities"
The example of orcid is not very clear to me. Why would I need to include a contextual entity that describes the person? For example, as shown in listing 1, we have
"author": {"@id": "https://orcid.org/0000-0002-1825-0097"}

I can understand the suggestion to add

```
{ "@id": "https://orcid.org/0000-0002-1825-0097",
"@type": "Person",
"name": "Josiah Carberry"
},
```

for discoverability (and convenience) purposes, but following the linked data practices I could get the jsonld of the person with
`curl --location --request GET 'https://orcid.org/0000-0002-1825-0097' --header 'Accept: application/ld+json'`

- Figure 2 does not clearly show how the data entities can be linked to each other. Neither does listing 1.

Section "People"
- add links to all mentioned projects.
- Change ';' with ':' in "A key set of stakeholders are developers;"
- Table 1: check missing link for SCHeMa

Section "Profiles of RO-Crate in use"
- Cultural Heritages: remove the final 's'

Section "Machine-actionable Data Management Plans"
Pending brace at " RO-Crate {RDA maDMP Mapper"

Section 'Institutional data repositories – Harvard Data Commons"
- missing figure number at "in the metadata to allow their future reuse (Figure~)"

Section "FAIR Digital Objects"
- the link behind "As an “abstract protocol”" returns 404

Section "packaging workflows"
- Missing citation at "authors might be penalised for doing so [cite?] "

Section "Minimal RO-Crate"
- Align terms in the definition of minimal RO-crate (published vs datePublished)
- Check footnote 8

Section "Mapping to RDF with schema.org"
- The mapping for ContextualEntity(e) is missing

A couple of final questions, that can be addressed in authors' future work, is about the profiles. Could they limit the re-usability of RO-Crate and the data entities in it in different domain? Would the proliferation of profiles hinder the simplicity and set barriers to the adoption of the community tools that have been developed so far?

## Reviewer 3

_By anonymous:_

> Background: Incomplete or inappropriate

> Although the idea behind the RO-Crate seems interesting, in its current shape that is designed for a limited community with limited purposes. 
> The impact of RO-Crate for large uptake and long term could be formulated better.

> In the related work section, a better explanation and connection of RO-Crate to the relevant and known platforms is expected.

> Though a resource paper, there is no **evaluation or statistics** for comparison of RO-Crate to any other similar service. Even a **user study** from the LOD community would be helpful.

> - Following the given example, how do the authors deal with the massively missing parts of the output files, sequence analysis code, resulting data and reports of the existing research results of the past or even future? not every code and related artefacts of research outcomes are always available. Providing a statistic from repositories will be helpful.
> - Although the primary purpose of the RO-Crate specification is to assist developers in leveraging LOD, how can individual researchers benefit from RO-Crate or is it just a community-wide service for LOD developers?
> - It is mentioned that ‘RO-Crate is used across multiple areas’, but how can other less-technical domains benefit from RO-Crate?
> - How RO-Crate relates and differs from ORKG (https://www.orkg.org/orkg/) and other similar platforms?
> - RO-Crate is presented as ‘a structured archive’ and not called a knowledge graph (KG) as it seems to be more than a KG, but looking into the details of the service, it looks like a multimodal scholarly KG - what are the authors’ viewpoints here?

## Meta-review by editor

I have to confess it has been a pleasure reading this paper since it presents, with a lot of details, an important resource for the scholarly community at large, in particular in the context of current practices (such as those concerning FAIR and EOSC). All the reviewers agreed on this aspect, even if they provided some insights and questions that should be appropriately addressed in the camera-ready and that I will carefully check before publication.

### 2021-10-11 content/80.formal-definition.tex
Daniel's edits ++
Fixed accidental markdown

@@
 \forall r \ Root(r)                         & \Rightarrow & Dataset(r) \land name(r, n) \land description(r, d) \land \\
-                                            & &             published(r, date) \land license(e, l) \\
+                                            & &             datePublished(r, date) \land license(e, l) \\
 \forall e \forall n \ name(e, n)            & \Rightarrow & Value(n) \\
@@
 
-In reality many additional attributes and contextual types from schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (_SHOULD_ requirements) they do not form part of this formalization
+In reality many additional attributes and contextual types from schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (\textit{SHOULD} requirements) they do not form part of this formalization
 

### 2021-10-11 content/80.formal-definition.tex
text mode in math mode in text mode.
Latex....

@@
 Notable from this triple-like formalization is that a RO-Crate $R$ is fully represented as a tree at depth 2 helped by the use of $𝕀𝕣𝕚$ nodes.
-For instance the aggregation from the root entity $hasPart(…interviews/>)$ is at same level as the data entity’s property $license(…CC-BY-NC-4.0>)$ and that contextual entity’s attribute name $(…Non Commercial 4.0”)$.
+For instance the aggregation from the root entity $hasPart(\texttt{…interviews/>})$ is at same level as the data entity’s property $license(\texttt{…CC-BY-NC-4.0>})$ and that contextual entity’s attribute name $(\text{…Non Commercial 4.0”})$.
 As shown in section RO-Crate JSON-LD, the RO-Crate Metadata File serialization is an equivalent shallow tree, although at depth 3 to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.

### 2021-10-11 content/80.formal-definition.tex
All do the latex dance

--- 2021-10-11 content/80.formal-definition.tex
@@
 Notable from this triple-like formalization is that a RO-Crate $R$ is fully represented as a tree at depth 2 helped by the use of $𝕀𝕣𝕚$ nodes.
-For instance the aggregation from the root entity $hasPart(…interviews/>)$ is at same level as the data entity’s property $license(…CC-BY-NC-4.0>)$ and that contextual entity’s attribute name $(…Non Commercial 4.0”)$.
+For instance the aggregation from the root entity $hasPart(\texttt{…interviews/>})$ is at same level as the data entity’s property $license(\texttt{…CC-BY-NC-4.0>})$ and that contextual entity’s attribute $ name(\text{…Non Commercial 4.0”})$.
 As shown in section RO-Crate JSON-LD, the RO-Crate Metadata File serialization is an equivalent shallow tree, although at depth 3 to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.
 
-In reality many additional attributes and contextual types from schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (\textit{SHOULD} requirements) they do not form part of this formalization
+In reality many additional attributes and contextual types from schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (\textit{SHOULD} requirements) they do not form part of this formalization.
 

### 2021-10-13 build/build.sh
Remove LaTeX \_ escapes in DOIs

@@ -31,6 +31,8 @@ sed -E -i 's/\\cite\{([^}]*)\}/[@\1]/g' content/*md
 sed -E -i 's/\\citet\{([^}]*)\}/@\1/g' content/*md
 # remove latex comments
 sed -E -i 's/^%.*//g' content/*md
+# Remove LaTeX \_ escapes in DOIs
+sed -E -i 's/\_/_/g' latex/ro-crate.bib
 
 
 # Generate reference information

### 2021-10-13 content/80.formal-definition.tex
Escape # inside math mode

@@
 LanguageTag(v, l)           & \equiv &      Datatype\big(v, \\
-    && \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns#langString>}\big)
+    && \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#langString>}\big)
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-Property(p)         & \Rightarrow &     type(p, \texttt{<http://www.w3.org/2000/01/rdf-schema#Property>})   \\
-Class(c)            & \Rightarrow &     type(c, \texttt{<http://www.w3.org/2000/01/rdf-schema#Class>})  \\
+Property(p)         & \Rightarrow &     type(p, \texttt{<http://www.w3.org/2000/01/rdf-schema\#Property>})   \\
+Class(c)            & \Rightarrow &     type(c, \texttt{<http://www.w3.org/2000/01/rdf-schema\#Class>})  \\
 Dataset(d)          & \Rightarrow &     type(d, \texttt{<http://schema.org/Dataset>})   \\
@@
 license(e, l)       & \Rightarrow &     Relation(e, \texttt{<http://schema.org/license>}, l) \land CreativeWork(l) \\
-type(e, t)          & \Rightarrow &     Relation(e, \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns#type>}, t) \\
+type(e, t)          & \Rightarrow &     Relation(e, \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#type>}, t) \\
                     & & \land Class(t)   \\
 String(s)           & \equiv &          Value(s) \land  s \in \mathbb{S} \\
-String(s)           & \Rightarrow &     Datatype(s, \texttt{<http://www.w3.org/2001/XMLSchema#string>}) \\
+String(s)           & \Rightarrow &     Datatype(s, \texttt{<http://www.w3.org/2001/XMLSchema\#string>}) \\
 Decimal(d)          & \equiv &          Value(d) \land  d \in \mathbb{R} \\
-Decimal(d)          & \Rightarrow &     Datatype(d, \texttt{<http://www.w3.org/2001/XMLSchema#decimal>}) \\
+Decimal(d)          & \Rightarrow &     Datatype(d, \texttt{<http://www.w3.org/2001/XMLSchema\#decimal>}) \\
 Relation(s,p,o)     & \Rightarrow &     Triple(s,p,o) \land IRI(s) \land IRI(o) \\

### 2021-10-13 build/build.sh
more sed!

@@ -23,6 +23,7 @@ sed -i 's,../content/images/,images/,g' content/*md
 sed -i 's,\.pdf,.svg,g' content/*md
 
 ## Horrible latex seds for the figure captions
+sed -E -i 's/\ref{fig:/{@fig:/g' content/*md
 sed -E -i 's/^!\[(.*)\]\((.*) "([^"]*)"\)/![\3](\2 "\1")/g' content/*md
 sed -E -i 's/\\textbf\{([^}]*)\}/**\1**/g' content/*md
 sed -E -i 's/\\emph\{([^}]*)\}/*\1*/g' content/*md

### 2021-10-13 build/build.sh
more sed tricks

@@ -23,7 +23,7 @@ sed -i 's,../content/images/,images/,g' content/*md
 sed -i 's,\.pdf,.svg,g' content/*md
 
 ## Horrible latex seds for the figure captions
-sed -E -i 's/\ref{fig:/{@fig:/g' content/*md
+sed -E -i 's/\ref\{fig:/\{@fig:/g' content/*md
 sed -E -i 's/^!\[(.*)\]\((.*) "([^"]*)"\)/![\3](\2 "\1")/g' content/*md
 sed -E -i 's/\\textbf\{([^}]*)\}/**\1**/g' content/*md
 sed -E -i 's/\\emph\{([^}]*)\}/*\1*/g' content/*md

### 2021-10-13 build/build.sh
more escape escapes

@@ -23,7 +23,7 @@ sed -i 's,../content/images/,images/,g' content/*md
 sed -i 's,\.pdf,.svg,g' content/*md
 
 ## Horrible latex seds for the figure captions
-sed -E -i 's/\ref\{fig:/\{@fig:/g' content/*md
+sed -E -i 's/\\ref\{fig:/\{@fig:/g' content/*md
 sed -E -i 's/^!\[(.*)\]\((.*) "([^"]*)"\)/![\3](\2 "\1")/g' content/*md
 sed -E -i 's/\\textbf\{([^}]*)\}/**\1**/g' content/*md
 sed -E -i 's/\\emph\{([^}]*)\}/*\1*/g' content/*md
@@ -33,7 +33,7 @@ sed -E -i 's/\\citet\{([^}]*)\}/@\1/g' content/*md
 # remove latex comments
 sed -E -i 's/^%.*//g' content/*md
 # Remove LaTeX \_ escapes in DOIs
-sed -E -i 's/\_/_/g' latex/ro-crate.bib
+sed -E -i 's/\\_/_/g' latex/ro-crate.bib
 
 
 # Generate reference information

### 2021-10-13 content/80.formal-definition.tex
avoid latex footnote

@@
 
-The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or both\footnote\url{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities}}; and must be typed with at least one $Class(e)$.
+The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or \href{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities}{both}; and must be typed with at least one $Class(e)$.
 

### 2021-10-13 content/21a.conceptual-img.tex

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.95\textwidth]{content/images/ro-crate-overview.pdf}
    \caption{\textbf{Conceptual overview of RO-Crate}. A \emph{Persistent Identifier} (PID) \cite{doi:10.1371/journal.pbio.2001414} points to a \emph{Research Object} (RO), which may be archived using different packaging approaches like BagIt \cite{doi:10.17487/rfc8493}, OCFL \cite{ocfl_2020}, git or ZIP. The RO is described within a \emph{RO-Crate Metadata File}, providing identifiers for \emph{authors} using ORCID, \emph{organisations} using [ROR](https://ror.org/) and licences such as Creative Commons using SPDX identifiers. The \emph{RO-Crate content} is further described with additional metadata. Data can be embedded files and directories, as well as links to external web resources, PIDs and nested RO-Crates.}
    \label{fig:conceptual}
\end{figure}

### 2021-10-13 content/21e.contextual-entities-img.tex

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.9\textwidth]{content/images/ro-crate-uml.pdf}
    \caption{\textbf{UML model view of RO-Crate.} The \emph{RO-Crate Metadata File} conforms to a version of the specification; and contains a JSON-LD graph that describes the entities that make up the RO-Crate. The \emph{RO-Crate Root Data Entity} represent the Research Object as a dataset. The RO-Crate aggregates \emph{data entities} (\texttt{hasPart}) which are further described using \emph{contextual entities} (which may include aggregated and non-aggregated data entities). Multiple types and relations from Schema.org allow annotations to be more specific, including figures, nested datasets, computational workflows, people, organisations, instruments and places. Contextual entities not otherwise cross-referenced from other entities' properties (\emph{describes}) can be grouped under the root entity (\texttt{mentions}).}
    \label{fig:uml}
\end{figure}

### 2021-10-13 content/80.formal-definition.tex
more \# magic

@@
     For simplicity, blank nodes are not included in this formalisation, as RO-Crate
-    recommends the use of IRI identifiers: \url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#describing-entities-in-json-ld}
+    recommends the use of IRI identifiers: \url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#describing-entities-in-json-ld}
 }, with additional descriptions using numbers $\mathbb{R}$ (notation $13.37$) and literal strings $\mathbb{S}$ (notation $\text{“Hello”}$).

### 2021-10-13 content/42a.regulatory-img.tex

--- 2021-08-08 _markdown_ro-crate/3ffef2c3bc11bc83c92df89b6fc00f76.md.tex
@@
-\markdownRendererHeadingTwo{Regulatory Sciences}\markdownRendererInterblockSeparator
-{}\markdownRendererLink{BioCompute Objects}{https://www.biocomputeobject.org/}{https://www.biocomputeobject.org/}{} (BCO) \markdownRendererCite{1}+{}{}{doi:10.1371/journal.pbio.3000099} is a community-led effort to standardise submissions of computational workflows to biomedical regulators.
-For instance, a genomics sequencing pipeline, as part of a personalised cancer treatment study, can be submitted to the US Food and Drugs Administration (FDA) for approval.
-BCOs are formalised in the standard IEEE 2791-2020 \markdownRendererCite{1}+{}{}{doi:10.1109/IEEESTD.2020.9094416} as a combination of \markdownRendererLink{JSON Schemas}{https://opensource.ieee.org/2791-object/ieee-2791-schema/}{https://opensource.ieee.org/2791-object/ieee-2791-schema/}{} that define the structure of JSON metadata files describing exemplar workflow runs in detail, covering aspects such as the usability and error domain of the workflow, its runtime requirements, the reference datasets used and representative output data produced.\markdownRendererInterblockSeparator
-{}BCOs provide a structured view over a particular workflow, informing regulators about its workings independently of the underlying workflow definition language.
-However, BCOs have only limited support for additional metadata\markdownRendererFootnote{IEEE 2791-2020 do permit user extensions in the \markdownRendererEmphasis{extension domain} by referencing additional JSON Schemas.}.
-For instance, while the BCO itself can indicate authors and contributors, and in particular regulators and their review decisions, it cannot describe the provenance of individual data files or workflow definitions.
-\markdownRendererInterblockSeparator
-{}As a custom JSON format, BCOs cannot be extended with Linked Data concepts, except by adding an additional top-level JSON object formalised in another JSON Schema.
-A BCO and workflow submitted by upload to a regulator will also frequently consist of multiple cross-related files.
-Crucially, there is no way to tell whether a given \markdownRendererCodeSpan{*.json} file is a BCO file, except by reading its content and check for its \markdownRendererCodeSpan{spec\markdownRendererUnderscore{}version}.
-\markdownRendererInterblockSeparator
-{}We can then consider how a BCO and its referenced artefacts can be packaged and transferred following FAIR principles.
-\markdownRendererLink{\markdownRendererStrongEmphasis{BCO RO-Crate}}{https://biocompute-objects.github.io/bco-ro-crate/}{https://biocompute-objects.github.io/bco-ro-crate/}{}\markdownRendererCite{1}+{}{}{doi:10.5281/zenodo.4633732}, part of the BioCompute Object user guides, defines a set of best practices for wrapping a BCO with a workflow, together with its exemplar outputs in an RO-Crate, which then provides typing and additional provenance metadata of the individual files, workflow definition, referenced data and the BCO metadata itself.
-\markdownRendererInterblockSeparator
-{}Here the BCO is responsible for describing the \markdownRendererEmphasis{purpose} of a workflow and its run at an abstraction level suitable for a domain scientist, while the more open-ended RO-Crate describes the surroundings of the workflow, classifying and relating its resources and providing provenance of their existence beyond the BCO.
-This emerging \markdownRendererEmphasis{separation of concerns} highlight how RO-Crate is used side-by-side of existing standards, even where there are apparent partial overlaps.\markdownRendererInterblockSeparator
-{}A similar separation of concerns can be found if considering the RO-Crate as a set of files, where the \markdownRendererEmphasis{transport-level} metadata, such as checksum of files, are \markdownRendererLink{delegated to BagIt}{https://www.researchobject.org/ro-crate/1.1/appendix/implementation-notes.html#adding-ro-crate-to-bagit}{https://www.researchobject.org/ro-crate/1.1/appendix/implementation-notes.html#adding-ro-crate-to-bagit}{} manifests, a standard focusing on the preservation challenges of digital libraries\markdownRendererCite{1}+{}{}{doi:10.17487/rfc8493}.
-As such, RO-Crates are not required to iterate all the files in their folder hierarchy, only those that benefit from being described.\markdownRendererInterblockSeparator
-{}Specifically, a BCO alone is insufficient for reliable re-execution of a workflow, which would need a compatible workflow engine depending on the workflow definition language, so IEEE 2791 recommends using Common Workflow Language \markdownRendererCite{1}+{}{}{arxiv:2105.07028} for interoperable pipeline execution.
-CWL itself relies on tool packaging in software containers using \markdownRendererLink{Docker}{https://www.docker.com/}{https://www.docker.com/}{} or \markdownRendererLink{Conda}{https://docs.conda.io/}{https://docs.conda.io/}{}.
-Thus, we can consider BCO RO-Crate as a stack: transport-level manifests of files (BagIt), provenance, typing and context of those files (RO-Crate), workflow overview and purpose (BCO), interoperable workflow definition (CWL) and tool distribution (Docker).\markdownRendererInterblockSeparator
-{}\markdownRendererImage{separationofconcerns}{../content/images/ro-crate-bco-sep-of-concerns.pdf}{../content/images/ro-crate-bco-sep-of-concerns.pdf}{\textbf{Separation of Concerns in BCO RO-Crate}.
+\begin{figure}[t]
+    \centering
+    \includegraphics{content/images/ro-crate-bco-sep-of-concerns.pdf}
+    \caption{\textbf{Separation of Concerns in BCO RO-Crate}.
 BioCompute Object (IEEE2791) is a JSON file that structurally explains the purpose and implementation of a computational workflow, for instance implemented in Nextflow, that installs the workflow’s software dependencies as Docker containers or BioConda packages.
@@
 RO-Crate gathers all these local and external resources, relating them and giving individual descriptions, for instance permanent DOI identifiers for reused datasets accessed from Zenodo, but also adding external identifiers to attribute authors using ORCID or to identify which licences apply to individual resources.
-The RO-Crate and its local files are captured in a BagIt whose checksum ensures completeness, combined with Big Data Bag \cite{doi:10.1109/BigData.2016.7840618} features to “complete” the bag with large external files such as the workflow outputs}\relax
+The RO-Crate and its local files are captured in a BagIt whose checksum ensures completeness, combined with Big Data Bag \cite{doi:10.1109/BigData.2016.7840618} features to “complete” the bag with large external files such as the workflow outputs}
+    \label{fig:sep_concerns}
+\end{figure}

### 2021-10-13 content/80.formal-definition.tex
More footnotes for lovely PDF world

@@
 An $ROCrate(R)$ is defined as a self-described \emph{Root Data Entity}, which describes and contains parts (\emph{data entities}), which are further described in \emph{contextual entities}.
-These terms align with their use in the \href{https://www.researchobject.org/ro-crate/1.1/terminology}{RO-Crate 1.1 terminology}.
+These terms align with their use in the RO-Crate 1.1 terminology\footnote{
+    \url{https://www.researchobject.org/ro-crate/1.1/terminology}}.
 
 The $Root(r)$ is a type of $Dataset(r)$, and must have the metadata to literal attributes to provide a $name$, $description$ and $datePublished$, as well as a contextual entity identifying its license.
-These predicates correspond to the RO-Crate 1.1 \href{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#direct-properties-of-the-root-data-entity}{requirements for the root data entity}.
+These predicates correspond to the RO-Crate 1.1 requirements for the root data entity\footnote{
+    \href{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}
+}.
 
-The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or \href{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html#contextual-vs-data-entities}{both}; and must be typed with at least one $Class(e)$.
+The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or both\footnote{
+    \href{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html\#contextual-vs-data-entities}
+}; and must be typed with at least one $Class(e)$.
 
@@
 
-An important RO-Crate principle is that of being \textbf{self-described} Therefore the serialisation of the RO-Crate into a file should also describe itself in a \href{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html#ro-crate-metadata-file-descriptor}{Metadata File Descriptor}, indicating it is $about$ (describing) the RO-Crate root data entity, and that it $conformsTo$ a particular version of the RO-Crate specification:
+An important RO-Crate principle is that of being \textbf{self-described} Therefore the serialisation of the RO-Crate into a file should also describe itself in a Metadata File Descriptor\footnote{
+    \url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#ro-crate-metadata-file-descriptor}
+}, indicating it is $about$ (describing) the RO-Crate root data entity, and that it $conformsTo$ a particular version of the RO-Crate specification:
 
@@
     \textbf{Limitations:} Contextual entities not related from the RO-Crate (e.g.
-using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see \href{https://github.com/ResearchObject/ro-crate/issues/122}{issue 122}.
+using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see \href{https://github.com/ResearchObject/ro-crate/issues/122}{GitHub issue ResearchObject/ro-crate\#122}.
 The $datePublished(e, d)$ rule do not include syntax checks for the ISO 8601 datetime format.

### 2021-10-13 build/build.sh
Render <url>s

@@ -30,6 +30,7 @@ sed -E -i 's/\\emph\{([^}]*)\}/*\1*/g' content/*md
 sed -E -i 's/\\texttt\{([^}]*)\}/`\1`/g' content/*md
 sed -E -i 's/\\cite\{([^}]*)\}/[@\1]/g' content/*md
 sed -E -i 's/\\citet\{([^}]*)\}/@\1/g' content/*md
+sed -E -i 's/\\url\{([^}]*)\}/<\1>/g' content/*md
 # remove latex comments
 sed -E -i 's/^%.*//g' content/*md
 # Remove LaTeX \_ escapes in DOIs

### 2021-10-13 content/80.formal-definition.tex
Footnote URLs. Cite RDF 1.1

@@
 These predicates correspond to the RO-Crate 1.1 requirements for the root data entity\footnote{
-    \href{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}
+    \url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}
 }.
@@
 The concept of an $Entity(e)$ is introduced as being either a $DataEntity(e)$, a $ContextualEntity(e)$, or both\footnote{
-    \href{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html\#contextual-vs-data-entities}
+    \url{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html\#contextual-vs-data-entities}
 }; and must be typed with at least one $Class(e)$.
@@
 } language $\mathcal{L}_{RDF}$
-based on the \href{http://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/}{RDF abstract syntax}:
+based on the RDF abstract syntax \cite{rdfworkinggroup_2014}:
 

### 2021-10-22 content/21a.conceptual-img.tex
Updates from Overleaf

@@
 A \emph{Persistent Identifier} (PID) \cite{doi:10.1371/journal.pbio.2001414} points to a \emph{Research Object} (RO), which may be archived using different packaging approaches like BagIt \cite{doi:10.17487/rfc8493}, OCFL \cite{ocfl_2020}, git or ZIP.
-The RO is described within a \emph{RO-Crate Metadata File}, providing identifiers for \emph{authors} using ORCID, \emph{organisations} using [ROR](https://ror.org/) and licences such as Creative Commons using SPDX identifiers.
-The \emph{RO-Crate content} is further described with additional metadata.
+The RO is described within a \emph{RO-Crate Metadata File}, providing identifiers for \emph{authors} using ORCID, \emph{organisations} using ROR \cite{doi:10.6087/kcse.192} and licences such as Creative Commons using SPDX identifiers.
+The \emph{RO-Crate content} is further described with additional metadata following a Linked Data approach.
 Data can be embedded files and directories, as well as links to external web resources, PIDs and nested RO-Crates.}

### 2021-10-26 content/31.tooling-table.tex
more refs
More edits following review
references

@@
 		\textbf{ro-crate-excel} \citep{ro-crate-excel} & Data managers & JavaScript \\
-		\multicolumn{4}{l}{\it Command-line tool to help create RO-Crates and HTML-readable rendering} \\
+		\multicolumn{4}{l}{\it Command-line tool to create/edit RO-Crates with spreadsheets} \\
 
@@
 		\textbf{ro-crate-py} \citep{ro-crate-py} & Developers & Python & Beta \\
-		\multicolumn{4}{l}{\it Object-oriented Python library for reading/writing RO-Crate} \\
+		\multicolumn{4}{l}{\it Object-oriented Python library for reading/writing RO-Crate and use by Jupyter Notebook} \\
 
@@
 
-        \textbf{SCHeMa} \citep{arxiv:2103.13138v1} & Workflow users & PHP & Alpha \\
+        \textbf{SCHeMa} \citep{vergoulis2021schema} & Workflow users & PHP & Alpha \\
         \multicolumn{4}{l}{\it Workflow execution using RO-Crate as exchange mechanism} \\
@@
 		\textbf{RDA maDMP Mapper} \citep{doi:10.5281/zenodo.3922136} & Data Management Plan users & Python & Beta \\
-		\multicolumn{4}{l}{\it Mapping between machine-actionable data management plans (maDMP) and RO-Crate\citep{doi:10.4126/frl01-006423291} } \\
+		\multicolumn{4}{l}{\it Mapping between machine-actionable data management plans (maDMP) and RO-Crate \citep{doi:10.4126/frl01-006423291} } \\
 
@@
 
+        \textbf{RO-Crates-and-Excel} \citep{doi:10.5281/zenodo.5068950} & Data Managers & Java (CLI) & Alpha \\
+        \multicolumn{4}{l}{\it Describe column/data details of spreadsheets as RO-Crate using DataCube vocabulary} \\
+
 		\bottomrule
 	\end{tabular}
-	\caption{Applications and libraries implementing RO-Crate, targeting different types of users across multiple programming languages. Status is indicative as assessed by this work.}
+	\caption{Applications and libraries implementing RO-Crate, targeting different types of users across multiple programming languages.
+Status is indicative as assessed by this work (Alpha < Beta < Release Candidate (RC) < Release).}
 	\label{tab:tools}

### 2021-10-27 content/21a.conceptual-img.tex
language improvements

@@
 A \emph{Persistent Identifier} (PID) \cite{doi:10.1371/journal.pbio.2001414} points to a \emph{Research Object} (RO), which may be archived using different packaging approaches like BagIt \cite{doi:10.17487/rfc8493}, OCFL \cite{ocfl_2020}, git or ZIP.
-The RO is described within a \emph{RO-Crate Metadata File}, providing identifiers for \emph{authors} using ORCID, \emph{organisations} using ROR \cite{doi:10.6087/kcse.192} and licences such as Creative Commons using SPDX identifiers.
+The RO is described within a \emph{RO-Crate Metadata File}, providing identifiers for \emph{authors} using ORCID, \emph{organisations} using Research Organization Registry (ROR) \cite{doi:10.6087/kcse.192} and licences such as Creative Commons using SPDX identifiers.
 The \emph{RO-Crate content} is further described with additional metadata following a Linked Data approach.

### 2021-11-01 content/21e.contextual-entities-img.tex
more section labels

@@
     \includegraphics[width=0.9\textwidth]{content/images/ro-crate-uml.pdf}
-    \caption{\textbf{UML model view of RO-Crate.} The \emph{RO-Crate Metadata File} conforms to a version of the specification; and contains a JSON-LD graph that describes the entities that make up the RO-Crate.
+    \caption{\textbf{Simplified UML class diagram of RO-Crate.} The \emph{RO-Crate Metadata File} conforms to a version of the specification; and contains a JSON-LD graph that describes the entities that make up the RO-Crate.
 The \emph{RO-Crate Root Data Entity} represent the Research Object as a dataset.

### 2021-11-01 build/build.sh
use sec: references in markdown

@@ -31,6 +31,8 @@ sed -E -i 's/\\texttt\{([^}]*)\}/`\1`/g' content/*md
 sed -E -i 's/\\cite\{([^}]*)\}/[@\1]/g' content/*md
 sed -E -i 's/\\citet\{([^}]*)\}/@\1/g' content/*md
 sed -E -i 's/\\url\{([^}]*)\}/<\1>/g' content/*md
+sed -E -i 's/\\ref\{(sec:[^}]*)\}/{@sec:\1}/g' content/*md
+
 # remove latex comments
 sed -E -i 's/^%.*//g' content/*md
 # Remove LaTeX \_ escapes in DOIs

### 2021-11-01 build/build.sh
avoid sec:sec

@@ -31,7 +31,7 @@ sed -E -i 's/\\texttt\{([^}]*)\}/`\1`/g' content/*md
 sed -E -i 's/\\cite\{([^}]*)\}/[@\1]/g' content/*md
 sed -E -i 's/\\citet\{([^}]*)\}/@\1/g' content/*md
 sed -E -i 's/\\url\{([^}]*)\}/<\1>/g' content/*md
-sed -E -i 's/\\ref\{(sec:[^}]*)\}/{@sec:\1}/g' content/*md
+sed -E -i 's/\\ref\{(sec:[^}]*)\}/{@\1}/g' content/*md
 
 # remove latex comments
 sed -E -i 's/^%.*//g' content/*md

### 2021-11-01 content/80.formal-definition.tex
section reference for appendix
section reference

@@
 \section{Formalizing RO-Crate in First Order Logic}
+
+\label{sec:formaldefinition}
 
@@
 For instance the aggregation from the root entity $hasPart(\texttt{…interviews/>})$ is at same level as the data entity’s property $license(\texttt{…CC-BY-NC-4.0>})$ and that contextual entity’s attribute $ name(\text{…Non Commercial 4.0”})$.
-As shown in section RO-Crate JSON-LD, the RO-Crate Metadata File serialization is an equivalent shallow tree, although at depth 3 to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.
+As shown in section \ref{sec:jsonld}, the RO-Crate Metadata File serialization is an equivalent shallow tree, although at depth 3 to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.
 

### 2021-11-02 content/21e.contextual-entities-img.tex
language improvements (Stian)

--- 2021-10-13 content/21e.contextual-entities-img.tex
@@
     \includegraphics[width=0.9\textwidth]{content/images/ro-crate-uml.pdf}
-    \caption{\textbf{UML model view of RO-Crate.} The \emph{RO-Crate Metadata File} conforms to a version of the specification; and contains a JSON-LD graph that describes the entities that make up the RO-Crate.
+    \caption{\textbf{Simplified UML class diagram of RO-Crate.} The \emph{RO-Crate Metadata File} conforms to a version of the specification; and contains a JSON-LD graph \cite{sporny_2014} that describes the entities that make up the RO-Crate.
 The \emph{RO-Crate Root Data Entity} represent the Research Object as a dataset.

### 2021-11-10 content/42a.regulatory-img.tex
example is CWL instead of Nextflow in figure

@@
     \caption{\textbf{Separation of Concerns in BCO RO-Crate}.
-BioCompute Object (IEEE2791) is a JSON file that structurally explains the purpose and implementation of a computational workflow, for instance implemented in Nextflow, that installs the workflow’s software dependencies as Docker containers or BioConda packages.
+BioCompute Object (IEEE2791) is a JSON file that structurally explains the purpose and implementation of a computational workflow, for instance implemented in Common Workflow Language (CWL), that installs the workflow’s software dependencies as Docker containers or BioConda packages.
 An example execution of the workflow shows the different kinds of result outputs, which may be external, using GitHub LFS to support larger data.

### 2021-11-10 content/42a.regulatory-img.tex
period

@@
 RO-Crate gathers all these local and external resources, relating them and giving individual descriptions, for instance permanent DOI identifiers for reused datasets accessed from Zenodo, but also adding external identifiers to attribute authors using ORCID or to identify which licences apply to individual resources.
-The RO-Crate and its local files are captured in a BagIt whose checksum ensures completeness, combined with Big Data Bag \cite{doi:10.1109/BigData.2016.7840618} features to “complete” the bag with large external files such as the workflow outputs}
+The RO-Crate and its local files are captured in a BagIt whose checksum ensures completeness, combined with Big Data Bag \cite{doi:10.1109/BigData.2016.7840618} features to “complete” the bag with large external files such as the workflow outputs.}
     \label{fig:sep_concerns}

### 2021-11-18 content/42a.regulatory-img.tex
Final part of Alan's edits
Signed-off-by: Stian Soiland-Reyes <stain@apache.org>

@@
 BioCompute Object (IEEE2791) is a JSON file that structurally explains the purpose and implementation of a computational workflow, for instance implemented in Common Workflow Language (CWL), that installs the workflow’s software dependencies as Docker containers or BioConda packages.
-An example execution of the workflow shows the different kinds of result outputs, which may be external, using GitHub LFS to support larger data.
+An example execution of the workflow shows the different kinds of result outputs, which may be external, using GitHub LFS \cite{github-lfs} to support larger data.
 RO-Crate gathers all these local and external resources, relating them and giving individual descriptions, for instance permanent DOI identifiers for reused datasets accessed from Zenodo, but also adding external identifiers to attribute authors using ORCID or to identify which licences apply to individual resources.

### 2021-11-19 build/build.sh
Don't break .pdf hyperlinks

@@ -20,7 +20,7 @@ ln -s ../latex/ro-crate.bib content/manual-references-ro-crate.bib
 # Replace silly paths for LaTeX
 sed -i 's,../content/images/,images/,g' content/*md
 # .. and use SVG instead of PDF
-sed -i 's,\.pdf,.svg,g' content/*md
+#sed -i 's,\.pdf,.svg,g' content/*md
 
 ## Horrible latex seds for the figure captions
 sed -E -i 's/\\ref\{fig:/\{@fig:/g' content/*md

### 2021-11-26 content/80.formal-definition.tex
aligned tex/md of formal definitions
.. to cover Reviewer 2 improvements

@@
 Entity(e)                                   & \equiv &      DataEntity(e) \lor ContextualEntity(e) \\
-\forall e \ Entity(e)                       & \Rightarrow & Class(e) \\
+\forall e \ Entity(e)                       & \Rightarrow & type(e, c) \land Class(e) \\
+\forall e \ ContextualEntity(e)             & \Rightarrow & name(e, n)  \\
 Mentions(R, s)                              & \models &     Relation(s, p, e) \oplus Attribute(s,  p, l) \\
@@
     \url{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html\#contextual-vs-data-entities}
-}; and must be typed with at least one $Class(e)$.
+}.
+Any $Entity(e)$ must be typed with at least one $Class(c)$, and every $ContextualEntity(e)$ must also have a $name(e,n)$; this corresponds to expectations for any \emph{referenced contextual entity} (section \ref{sec:contextualentities}).
 
@@
 File(f)             & \Rightarrow &     type(f, \texttt{<http://schema.org/MediaObject>})   \\
+ContextualEntity(e) & \Rightarrow &     type(f, \texttt{<http://schema.org/Thing>})   \\
 CreativeWork(e)     & \Rightarrow &     ContextualEntity(e) \land type(e, \texttt{<http://schema.org/CreativeWork>})  \\

### 2021-11-27 build/build.sh
Removed nested ro-crate template
Now maintained in
https://github.com/researchobject/2021-packaging-research-artefacts-with-ro-crate
to appear at https://w3id.org/ro/doi/10.5281/zenodo.5146227

@@ -144,8 +144,4 @@ if [ "${SPELLCHECK:-}" = "true" ]; then
   rm output/expanded-spelling-errors.txt
 fi
 
-# Copy RO-Crate metadata
-# TODO: Add date, commit etc.
-cp ro-crate/ro-crate-metadata.json ro-crate/ro-crate-preview.html output/
-
 echo >&2 "Build complete"

### 2021-12-28 latex/ro-crate.tex
Updates from Overleaf
references
moved future work to new section
include future work
future work as part of conclusion
corrected ordering of affiliations
Typeset in LaTeX2e by VTeX Ltd., Vilnius, Lithuania

--- 2021-05-20 latex/ro-crate.tex
@@
-%% ro-crate paper
-%% (c) 2020-2021 The University of Manchester
-%% 
-%% SPDX-License-Identifier: CC-BY-4.0
-
-% add. options: [seceqn,secthm,crcready]
-\documentclass[ds]{iosart2x}
-
-%%%%%%%%%%% Put your definitions here
-
-% From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
-\usepackage{amsmath}
-\usepackage{amssymb}
-% \usepackage{hyperref}
-\usepackage{fontspec}
-\usepackage{unicode-math}
-% \usepackage{graphicx}
-
-\DeclareGraphicsExtensions{.pdf,.png,.svg}
-\graphicspath{../content/images/}
-\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
-\def\markdownOptionOutputDir{../output}
-
-% for table resizing
-\usepackage{booktabs}
-\usepackage{multirow}
-
-%%%%%%%%%%% End of definitions
-
+%BeginFileInfo
+%%Publisher=IOS
+%%Project=DS
+%%Manuscript=DS210053
+%%MS position=
+%%Stage=201
+%%TID=imikolaityte
+%%Pages=0
+%%Format=2019
+%%Distribution=vtex
+%%Destination=DVI
+%%PDF type=
+%%DVI.Maker=luadvi
+%%PS.Maker=luaps
+%%History1=2021.12.28 09:11
+%EndFileInfo
+%
+% Journal: Data Science (DS), IOS Press
+% Typeset in LaTeX2e by VTeX Ltd., Vilnius, Lithuania
+%
+%Spelling_date
+%\RequirePackage{float} % when algorithm.sty used in manuscript
+\documentclass[ds,v1.1.2,openaccess]{iosart2x}%% ,seceqn
+\usepackage{mathbh}
+%
+%\usepackage{}
+\usepackage{tmultirow}
+\usepackage{listings}
+%
+\PROOF
+%\PROOF
+%\CRC
+
+\volume{-1}
+\issue{-1}
 \pubyear{2021}
-\volume{0}
 \firstpage{1}
-\lastpage{1}
-
+\lastpage{42}
+\articletype{research-article} %%
+%http://www.vtex.vmt/docs/wiki/index.php/IOS_Press_instrukcijos_%28Kaunas%29#Straipsnio_tipas_.28.5Carticletype.7B.7D.29
+\aid{DS210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
+\doi{10.3233/DS-210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
+\vtexed{Ingrida}
+
+\startlocaldefs
+\setpkgattr{license}{type}{CC@BY} %% CC@BY@NC
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% indentation after theorem-like environments - auto
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%\newtheorem*{}{} %% unnumbered
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%\theoremstyle{plain}
+%%\newtheorem{}{}%[]
+%%\theoremstyle{definition}
+%%\newtheorem{}{}%[]
+\endlocaldefs
+%
 \begin{document}
-
-% Welcome to the 2000s with UNICODE fonts!
-% Adapted from
-% https://www.overleaf.com/learn/latex/Articles%2FOpenType-based_math_typesetting:_An_introduction_to_the_STIX2_OpenType_fonts
-\setmathfont{STIX2Math}[
-    Extension={.otf},
-    Path=./STIX2fonts/,
-    Scale=1
-]
-\setmainfont{STIX2Text}[
-    Extension={.otf},
-    Path=./STIX2fonts/,
-    UprightFont={*-Regular},
-    BoldFont={*-Bold},
-    ItalicFont={*-Italic},
-    BoldItalicFont={*-BoldItalic}
-]
-
-% https://typeof.net/Iosevka/ mono font has 
-% all; the logic symbols we need
-\setmonofont{iosevka-slab}[
-    Extension={.ttf},
-    Path=./ttf-iosevka-slab-6.1.3/,
-    Scale=MatchLowercase,
-    UprightFont={*-light},
-    ItalicFont={*-lightitalic}
-]
-
+%
 \begin{frontmatter}
-
+%
+\pretitle[XML]{Research Article}
 %\pretitle{}
 \title{Packaging research artefacts with RO-Crate}
+%\title{\thanks{}}
 \runtitle{Packaging research artefacts with RO-Crate}
 %\subtitle{}
-
-% For one author:
-%\author{\inits{N.}\fnms{Name1} \snm{Surname1}\ead[label=e1]{first@somewhere.com}\ead[label=o1,orcid]{0000-0000-0000-0001}}
-%\address{Department first, \orgname{University or Company name},
-%Abbreviate US states, \cny{Country}\printead[presep={\\}]{e1,o1}}
-
-% Two or more authors:
-\author[SSR1,SSR2]{\inits{S.}\fnms{Stian}
-    \snm{Soiland-Reyes}\ead[label=eSSR]{soiland-reyes@manchester.ac.uk}\ead[label=oSSR,orcid]{0000-0001-9842-9718}%
-    \thanks{Corresponding author. \printead{eSSR}.}}, 
-\author[PS]{\inits{P.}\fnms{Peter}
-    \snm{Sefton}\ead[label=ePS]{Peter.Sefton@uts.edu.au}\ead[label=oPS,orcid]{0000-0002-3545-944X}},
-\author[EOC]{\inits{E.}\fnms{Eoghan}
-    \snm{Ó Carragáin}\ead[label=eEOC]{eoghan.ocarragain@ucc.ie}\ead[label=oEOC,orcid]{0000-0001-8131-2150}},
-\author[MC]{\inits{M.}\fnms{Mercè}
-    \snm{Crosas}\ead[label=eMC]{mcrosas@g.harvard.edu}\ead[label=oMC,orcid]{0000-0003-1304-1939}},
-\author[LJC]{\inits{LJ.}\fnms{Leyla Jael}
-    \snm{Castro}\ead[label=eLJC]{ljgarcia@zbmed.de}\ead[label=oLJC,orcid]{0000-0003-3986-0510}},
-\author[FC]{\inits{F.}\fnms{Frederik}
-    \snm{Coppens}\ead[label=eFC]{frcop@psb.vib-ugent.be}\ead[label=oFC,orcid]{0000-0001-6565-5145}},
-\author[JMF]{\inits{JM.}\fnms{José M.}
-    \snm{Fernández}\ead[label=eJMF]{jose.m.fernandez@bsc.es}\ead[label=oJMF,orcid]{0000-0002-4806-5140}},
-\author[DG]{\inits{D.}\fnms{Daniel}
-    \snm{Garijo}\ead[label=eDG]{dgarijo@fi.upm.es}\ead[label=oDG,orcid]{0000-0003-0454-7145}},
-\author[MLR]{\inits{M.}\fnms{Marco}
-    \snm{La Rosa}\ead[label=eMLR]{m@lr.id.au}},
-\author[SL]{\inits{S.}\fnms{Simone}
-    \snm{Leo}\ead[label=eSL]{simone.leo@crs4.it}\ead[label=oSL,orcid]{0000-0001-8271-5429}},
-\author[MP]{\inits{M.}\fnms{Marc}
-    \snm{Portier}\ead[label=eMP]{marc.portier@vliz.be}\ead[label=oMP,orcid]{0000-0002-9648-6484}},
-\author[AT]{\inits{A.}\fnms{Ana}
-    \snm{Trisovic}\ead[label=eMP]{ana.trisovic@gmail.com}\ead[label=oAT,orcid]{0000-0003-1991-0533}},
-\author[RO]{RO-Crate Community},
-\author[PG]{\inits{P.}\fnms{Paul}
-    \snm{Groth}\ead[label=ePG]{p.t.groth@uva.nl}\ead[label=oPG,orcid]{0000-0003-0183-6910}},
-\author[CG]{\inits{C.}\fnms{Carole}
-    \snm{Goble}\ead[label=eCG]{carole.goble@manchester.ac.uk}\ead[label=oCG,orcid]{0000-0003-1219-2137}}
-
-\address[SSR1]{Computer Science, 
-    \orgname{The University of Manchester}, \cny{UK}}
-\address[SSR2]{Informatics Institute, \orgname{University of Amsterdam},
-    \cny{The Netherlands}\printead[presep={\\}]{eSSR,oSSR}}
-\address[PS]{Faculty of Science, \orgname{University Technology Sydney},
-    \cny{Australia}\printead[presep={\\}]{ePS,oPS}}
-\address[EOC]{\orgname{University College Cork},
-    \cny{Ireland}\printead[presep={\\}]{oEOC}}
-\address[MC]{Institute for Quantitative Social Science, 
-    \orgname{Harvard University}, Cambridge, MA,  \cny{USA}\printead[presep={\\}]{oMC}}
-\address[LJC]{\orgname{ZB MED Information Centre for Life Sciences}, 
-    Cologne, \cny{Germany}\printead[presep={\\}]{oLJC}}
-\address[FC]{\orgname{VIB-UGent Center for Plant Systems Biology}, 
-    Gent, \cny{Belgium}\printead[presep={\\}]{oFC}}
-\address[JMF]{\orgname{Barcelona Supercomputing Center}, 
-    Barcelona, \cny{Spain}\printead[presep={\\}]{oJMF}}
-\address[DG]{Ontology Engineering Group, 
-    \orgname{Universidad Politécnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{oDG}}
-\address[MLR]{\printead{eMLR}}
-\address[SL]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA),
-    \cny{Italy}\printead[presep={\\}]{oSL}}
-\address[MP]{\orgname{Vlaams Instituut voor de Zee}, Oostende,
-    \cny{Belgium}\printead[presep={\\}]{oMP}}
-\address[AT]{Institute for Quantitative Social Science,
-    \orgname{Harvard University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{oAT}}
-\address[RO]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
-\address[PG]{Informatics Institute, \orgname{University of Amsterdam},
-    \cny{The Netherlands}\printead[presep={\\}]{ePG,oPG}}
-\address[CG]{Computer Science, \orgname{The University of Manchester},
-    \cny{UK}\printead[presep={\\}]{eCG,oCG}}
-
-%\begin{review}{editor}
-%\reviewer{\fnms{First} \snm{Editor}\ead[label=oe1,orcid]{}\printead[presep={\ }]{oe1}}
-%\reviewer{\fnms{Second} \snm{Editor}\ead[label=oe2,orcid]{}\printead[presep={\ }]{oe2}}
-%\end{review}
-%\begin{review}{solicited}
-%\reviewer{\fnms{First} \snm{Solicited reviewer}\ead[label=os1,orcid]{}\printead[presep={\ }]{os1}}
-%\end{review}
-
-\newpage % Because the author affiliations are too long!
+%
+%%% initials without spaces
+%
+% {\inits{}\fnms{} \snm{} \suffix{}\ead[label=e?]{}
+%\ead[label=p1,phone]{}\ead[label=f1,fax]{}\ead[label=o?,orcid]{}
+%\thanks{}}
+%
+\begin{aug}
+\author[A,B]{\inits{S.}\fnms{Stian} \snm{Soiland-Reyes}\ead
+[label=e1]{soiland-reyes@manchester.ac.uk}\ead
+[label=o1,orcid]{0000-0001-9842-9718}\thanks{Corresponding author.
+\printead{e1}.}}
+\author[C]{\inits{P.}\fnms{Peter} \snm{Sefton}\ead
+[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}}
+\author[D]{\inits{M.}\fnms{Merc\`{e}} \snm{Crosas}\ead
+[label=o3,orcid]{0000-0003-1304-1939}}
+\author[E]{\inits{L.J.}\fnms{Leyla Jael} \snm{Castro}\ead
+[label=o4,orcid]{0000-0003-3986-0510}}
+\author[F]{\inits{F.}\fnms{Frederik} \snm{Coppens}\ead
+[label=o5,orcid]{0000-0001-6565-5145}}
+\author[G]{\inits{J.M.}\fnms{Jos\'{e} M.} \snm{Fern\'{a}ndez}\ead
+[label=o6,orcid]{0000-0002-4806-5140}}
+\author[H]{\inits{D.}\fnms{Daniel} \snm{Garijo}\ead
+[label=o7,orcid]{0000-0003-0454-7145}}
+\author[I]{\inits{B.}\fnms{Bj\"{o}rn} \snm{Gr\"{u}ning}\ead
+[label=o8,orcid]{0000-0002-3079-6586}}
+\author[J]{\inits{M.}\fnms{Marco} \snm{La Rosa}\ead
+[label=o9,orcid]{0000-0001-5383-6993}}
+\author[K]{\inits{S.}\fnms{Simone} \snm{Leo}\ead
+[label=o10,orcid]{0000-0001-8271-5429}}
+\author[L]{\inits{E.}\fnms{Eoghan} \snm{\'{O} Carrag\'{a}in}\ead
+[label=o11,orcid]{0000-0001-8131-2150}}
+\author[M]{\inits{M.}\fnms{Marc} \snm{Portier}\ead
+[label=o12,orcid]{0000-0002-9648-6484}}
+\author[N]{\inits{A.}\fnms{Ana} \snm{Trisovic}\ead
+[label=o13,orcid]{0000-0003-1991-0533}}
+\author[O]{\snm{RO-Crate Community}\ead
+[label=u1,url]{https://www.researchobject.org/ro-crate/community}}
+\author[P]{\inits{P.}\fnms{Paul} \snm{Groth}\ead
+[label=e15]{p.t.groth@uva.nl}\ead[label=o15,orcid]{0000-0003-0183-6910}}
+\author[Q]{\inits{C.}\fnms{Carole} \snm{Goble}\ead
+[label=e16]{carole.goble@manchester.ac.uk}\ead
+[label=o16,orcid]{0000-0003-1219-2137}}
+%
+%%%%%% printead order: TEL FAX E-MAIL URL
+%\thanks{Corresponding author: name surname, address. \printead{}.}
+%\thanks{Corresponding author. \printead{}.}
+%\thanks{}%% Move support info to acks
+%
+\runauthor{S. Soiland-Reyes et al.} %% auto
+% {\orgname{} \cny{}\printead[presep={\\}]{}}
+\address[A]{Department of Computer Science, \orgname{The University of
+Manchester}, \cny{UK}\printead[presep={\\}]{e1,o1}}
+\address[B]{Informatics Institute, \orgname{University of Amsterdam},
+\cny{The Netherlands}}
+\address[C]{Faculty of Science, \orgname{University Technology Sydney},
+\cny{Australia}\printead[presep={\\}]{e2,o2}}
+\address[D]{Institute for Quantitative Social Science, \orgname{Harvard
+University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{o3}}
+\address[E]{\orgname{ZB MED Information Centre for Life Sciences},
+Cologne, \cny{Germany}\printead[presep={\\}]{o4}}
+\address[F]{\orgname{VIB-UGent Center for Plant Systems Biology}, Gent,
+\cny{Belgium}\printead[presep={\\}]{o5}}
+\address[G]{\orgname{Barcelona Supercomputing Center}, Barcelona, \cny
+{Spain}\printead[presep={\\}]{o6}}
+\address[H]{Ontology Engineering Group, \orgname{Universidad Polit\'
+{e}cnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{o7}}
+\address[I]{Bioinformatics Group, Department of Computer Science,
+\orgname{Albert-Ludwigs-University Freiburg}, Freiburg, \cny
+{Germany}\printead[presep={\\}]{o8}}
+\address[J]{\orgname{PARADISEC}, Melbourne, \cny{Australia}\printead
+[presep={\\}]{o9}}
+\address[K]{\orgname{Center for Advanced Studies, Research, and
+Development in Sardinia} (CRS4), Pula (CA), \cny{Italy}\printead
+[presep={\\}]{o10}}
+\address[L]{\orgname{University College Cork}, \cny{Ireland}\printead
+[presep={\\}]{o11}}
+\address[M]{\orgname{Vlaams Instituut voor de Zee}, Oostende, \cny
+{Belgium}\printead[presep={\\}]{o12}}
+\address[N]{Institute for Quantitative Social Science, \orgname{Harvard
+University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{o13}}
+\address[O]{\printead{u1}~(see Appendix B)}
+\address[P]{Informatics Institute, \orgname{University of Amsterdam},
+\cny{The Netherlands}\printead[presep={\\}]{e15,o15}}
+\address[Q]{Department of Computer Science, \orgname{The University of
+Manchester}, \cny{UK}\printead[presep={\\}]{e16,o16}}
+\end{aug}
+%
+%%%%%%
+%
+\begin{review}{editor}
+\reviewer{\fnms{Silvio} \snm{Peroni}\ead
+[label=oe1,orcid]{0000-0003-0530-4305}\printead[presep={\ }]{oe1}}
+%\reviewer{\fnms{} \snm{}\ead[label=oe2,orcid]{}\printead[presep={\
+%}]{oe2}}
+\end{review}
+
+\begin{review}{solicited}
+\reviewer{\fnms{Angelo A.} \snm{Salatino}\ead
+[label=os1,orcid]{0000-0002-4763-3943}\printead[presep={\ }]{os1}}
+\reviewer{\fnms{Alessia} \snm{Bardi}\ead
+[label=os2,orcid]{0000-0002-1112-1292}\printead[presep={\ }]{os2}}
+\reviewer{\snm{one anonymous reviewer}}
+\end{review}
+%
+%\dedicated{}
+%
+% HISTORY:
+\received{\sday{13} \smonth{8} \syear{2021}}
+\accepted{\sday{23} \smonth{9} \syear{2021}}
+%
 \begin{abstract}
-\markdownInput{../content/01.abstract.md}
+%\parttitle{}
+An increasing number of researchers support reproducibility by
+including pointers to and
+descriptions of datasets, software and methods in their publications.
+However, scientific
+articles may be ambiguous, incomplete and difficult to process by
+automated systems. In
+this paper we introduce RO-Crate, an open, community-driven, and
+lightweight approach
+to packaging research artefacts along with their metadata in a machine
+readable manner.
+RO-Crate is based on \surl{Schema.org} annotations in JSON-LD, aiming to
+establish best
+practices to formally describe metadata in an accessible and practical
+way for their
+use in a wide variety of situations.
+
+An RO-Crate is a structured archive of all the items that contributed
+to a research
+outcome, including their identifiers, provenance, relations and
+annotations. As a general
+purpose packaging approach for data and their metadata, RO-Crate is
+used across multiple
+areas, including bioinformatics, digital humanities and regulatory
+sciences. By applying
+``just enough'' Linked Data standards, RO-Crate simplifies the process of
+making research outputs FAIR while also enhancing research reproducibility.
+
+An RO-Crate for this article\footnote{\url
+{https://w3id.org/ro/doi/10.5281/zenodo.5146227}}
+is archived at \url{https://doi.org/10.5281/zenodo.5146227}.
 \end{abstract}
-
-\begin{keyword}
+%
+\begin{keyword} %% 1 u.c.
 \kwd{Data publishing}
-\kwd{Data packaging}
+\kwd{data packaging}
 \kwd{FAIR}
 \kwd{Linked Data}
-\kwd{Metadata}
-\kwd{Reproducibility}
-\kwd{Research Object}
+\kwd{metadata}
+\kwd{reproducibility}
+\kwd{research object}
 \end{keyword}
-
+%
 \end{frontmatter}
-
-%%%%%%%%%%% The article body starts:
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% Main text entry area:
 % Every *.md file in content/ in order, except 00, 90
-\markdownInput{../content/10.background.md}
-
-\markdownInput{../content/20.rocrate.md}
-\markdownInput{../content/21.conceptual.md}
-\markdownInput{../content/22.implementation.md}
-\markdownInput{../content/23.community.md}
-
-\input{../content/31.tooling-table.tex}
-%\markdownInput{../content/31.tooling-table.md}
-\markdownInput{../content/30.tooling.md}
-
-\markdownInput{../content/40.inuse.md}
-\markdownInput{../content/41.workflows.md}
-\markdownInput{../content/42.regulatory.md}
-\markdownInput{../content/43.culturalheritage.md}
-\markdownInput{../content/44.dmp.md}
-
-\markdownInput{../content/50.relatedwork.md}
-\markdownInput{../content/60.conclusion.md}
-\markdownInput{../content/70.acknowledgements.md}
-
-%%%%%%%%%%% The bibliography starts:
-\nocite{*} 
-\bibliographystyle{ios1}
-\bibliography{ro-crate}
-
-%%%% Appendix
-
-\appendix
-\markdownInput{../content/80.formal-definition.md}
-% TODO: Recode formal definition to LaTeX math mode?
-%\include{../content/80.formal-definition}
-
-\markdownInput{../content/89.community.md}
-
+\section{Introduction}
+
+\label{sec:introduction}
+
+The move towards Open Science has increased the need and demand for the
+publication of artefacts of the research process
+\cite{sefton_blog_post_2021}. This is particularly apparent in domains that
+rely on computational experiments; for example, the publication of
+software, datasets and records of the dependencies that such
+experiments rely on \cite{doi:10.1126/science.aah6168}.
+
+It is often argued that the publication of these assets, and
+specifically software
+\cite{doi:10.3233/DS-190026}, workflows \cite{doi:10.1162/dint_a_00033} and
+data, should follow the FAIR principles \cite{doi:10.1038/sdata.2016.18};
+namely, that they are Findable, Accessible, Interoperable and Reusable.
+These principles are agnostic to the \textit{implementation} strategy needed
+to comply with them. Hence, there has been an increasing amount of work
+in the development of platforms and specifications that aim to fulfil
+these goals \cite{isbn:9781315351148}.
+
+Important examples include data publication with rich metadata (e.g.
+Zenodo \cite{doi:10.3897/biss.3.37080}), domain-specific data deposition
+(e.g. PDB \cite{doi:10.1093/nar/gkl971}) and following practices for
+reproducible research software \cite{doi:10.1371/journal.pcbi.1003285}
+(e.g. use of containers). While these platforms are useful, experience
+has shown that it is important to put greater emphasis on the
+interconnection of the multiple artefacts that make up the research
+process \cite{doi:10.1016/j.ijhcs.2020.102562}.
+
+The notion of \textbf{Research Objects} \cite{doi:10.1016/j.future.2011.08.004}
+(RO) was introduced to address this connectivity, providing
+semantically rich \textit{aggregations} of (potentially distributed) resources
+with a layer of structure over a research study; this is then to be
+delivered in a \textit{machine-readable format}.
+
+A Research Object combines the ability to bundle multiple types of
+artefacts together, such as spreadsheets, code, examples, and figures.
+The RO is augmented with annotations and relationships that describe
+the artefacts' \textit{context} (e.g. a CSV being used by a script, a figure
+being a result of a workflow).
+
+This notion of ROs provides a compelling vision as an approach for
+implementing FAIR data. However, existing Research Object
+implementations require a large technology stack
+\cite{doi:10.1016/j.websem.2015.01.003}, are typically tailored to a
+particular platform and are also not easily usable by end-users.
+
+To address this gap, a new community came together
+\cite{doi:10.5281/zenodo.3250687} to develop \textbf{RO-Crate} -- an \textit{approach
+to package and aggregate research artefacts with their metadata and
+relationships}. The aim of this paper is to introduce RO-Crate and
+assess it as a strategy for making multiple types of research artefacts
+FAIR. Specifically, the contributions of this paper are as follows:
+%
+\begin{enumerate}
+\item[1.] An introduction to RO-Crate, its purpose and context;
+\item[2.] A guide to the RO-Crate community and tooling;
+\item[3.] Examples of RO-Crate usage, demonstrating its value as connective
+tissue for different artefacts from different communities.
+\end{enumerate}
+
+The rest of this paper is organised as follows. We first describe
+RO-Crate through its development methodology that formed the RO-Crate
+concept, showing its foundations in Linked Data and emerging
+principles. We then define RO-Crate technically, before we introduce
+the community and tooling. We move to analyse RO-Crate with respect to
+usage in a diverse set of domains. Finally, we present related work and
+conclude with some remarks including RO-Crate highlights and future
+work. The appendix adds a formal definition of RO-Crate using
+First-Order logic.
+
+\section{RO-Crate} %{#rocrate}
+
+\label{sec:rocrate}
+
+RO-Crate aims to provide an approach to packaging research artefacts
+with their metadata that can be easily adopted. To illustrate this, let
+us imagine a research paper reporting on the sequence analysis of
+proteins obtained from an experiment on mice. The sequence output
+files, sequence analysis code, resulting data and reports summarising
+statistical measures are all important and inter-related research
+artefacts, and consequently would ideally all be co-located in a
+directory and accompanied with their corresponding metadata. In
+reality, some of the artefacts (e.g. data or software) will be recorded
+as external reference to repositories that are not necessarily
+following the FAIR principles. This conceptual directory, along with
+the relationships between its constituent digital artefacts, is what
+the RO-Crate model aims to represent, linking together all the elements
+of an experiment that are required for the experiment's reproducibility
+and reusability.
+
+The question then arises as to how the directory with all this material
+should be packaged in a manner that is accessible and usable by others.
+This means programmatically and automatically accessible by machines
+and human readable. A de facto approach to sharing collections of
+resources is through compressed archives (e.g. a zip file). This solves
+the problem of ``packaging'', but it does not guarantee downstream
+access to all artefacts in a programmatic fashion, nor describe the
+role of each file in that particular research. Both features, the
+ability to automatically access and reason about an object, are crucial
+and lead to the need for explicit metadata about the contents of the
+folder, describing each and linking them together.
+
+Examples of metadata descriptions across a wide range of
+domains\footnote{\url{https://rdamsc.bath.ac.uk/scheme-index}} abound within the
+literature, both in research data management
+\cite{doi:10.1007/s10209-016-0475-y,farnel_2014,doi:10.2777/620649}
+and within library and information
+systems\footnote{\url{https://www.loc.gov/librarians/standards}} \cite{chan_1995,doi:10.1515/9783598441844}. However, many of these approaches require
+knowledge of metadata schemas, particular annotation systems, or the
+use of complex software stacks. Indeed, particularly within research,
+these requirements have led to a lack of adoption and growing
+frustration with current tooling and specifications
+\cite{neylon_blog_post_2017,doi:10.1007/s00267-014-0258-2,doi:10.1038/s41597-020-0524-5}.
+
+RO-Crate seeks to address this complexity by:
+%
+\begin{enumerate}
+\item[1.] being conceptually simple and easy to understand for developers;
+\item[2.] providing strong, easy tooling for integration into community projects;
+\item[3.] providing a strong and opinionated guide regarding current best practices;
+\item[4.] adopting de-facto standards that are widely used on the Web.
+\end{enumerate}
+
+In the following sections we demonstrate how the RO-Crate specification
+and ecosystem achieve these goals.
+
+%% sub-subsections below to cater for figures in
+%% markdown vs LaTex
+\subsection{Development methodology}
+
+\label{sec:methodology}
+
+It is a good question as to what base level we assume for `conceptually
+simple'. We take simplicity to apply at two levels: for the
+\textit{developers} who produce the platforms and for the \textit{data practitioners}
+and users of those platforms.
+
+For our development methodology we followed the mantra of working
+closely with a small group to really get a deep understanding of
+requirements and ensure rapid feedback loops. We created a pool of
+early adopter projects from a range of disciplines and groups,
+primarily addressing developers of platforms. Thus the base level for
+simplicity was \textbf{developer friendliness}.
+
+We assumed a developer familiar with making Web applications with JSON
+data (who would then learn how to make \textit{RO-Crate JSON-LD}), which
+informed core design choices for our JSON-level documentation approach
+and RO-Crate serialization (Section~\ref{sec:implementation}). Our
+group of early adopters, growing as the community evolved, drove the
+RO-Crate requirements and provided feedback through our multiple
+communication channels including bi-monthly meetings, which we describe
+in Section~\ref{sec:community} along with the established norms.
+
+Addressing the simplicity of understanding and engaging with RO-Crate
+by data practitioners is through the platforms, for example with
+interactive tools (Section~\ref{sec:tooling}) like
+Describo\footnote{\url{https://uts-eresearch.github.io/describo/}} \cite{describo} and
+Jupyter notebooks \cite{doi:10.3233/978-1-61499-649-1-87}, and by close
+discussions with domain scientists on how to appropriately capture what
+they determine to be relevant metadata. This ultimately requires a new
+type of awareness and learning material separate from developer
+specifications, focusing on the simplicity of extensibility to serve
+the user needs, along with user-driven development of new RO-Crate
+Profiles specific for their needs (Section~\ref{sec:inuse}).
+
+\subsection{Conceptual definition}% {#conceptual}
+
+\label{sec:conceptual}
+
+A key premise of RO-Crate is the existence of a wide variety of
+resources on the Web that can help describe research. As such, RO-Crate
+relies on the Linked Data principles
+\cite{doi:10.2200/S00334ED1V01Y201102WBE001}. Figure~\ref{fig:conceptual}
+shows the main conceptual elements involved in an RO-Crate: The
+RO-Crate Metadata File (top) describes the Research Object using
+structured metadata including external references, coupled with the
+contained artefacts (bottom) bundled and described by the RO-Crate.
+
+The conceptual notion of a \textit{Research Object}
+\cite{doi:10.1016/j.future.2011.08.004} is thus realized with the RO-Crate
+model and serialized using Linked Data constructs within the RO-Crate
+metadata file.
+
+\begin{figure}%[t!]
+\includegraphics{210053f01}
+\caption{Conceptual overview of RO-Crate. A \emph{Persistent
+Identifier} (PID) \cite{doi:10.1371/journal.pbio.2001414} points to a
+\emph{Research Object} (RO), which may be archived using different
+packaging approaches like BagIt \cite{doi:10.17487/rfc8493}, OCFL \cite
+{ocfl_2020}, git or ZIP. The RO is described within a \emph{RO-Crate
+Metadata File}, providing identifiers for \emph{authors} using ORCID,
+\emph{organisations} using Research Organization Registry (ROR) \cite
+{doi:10.6087/kcse.192} and licences such as Creative Commons using SPDX
+identifiers. The \emph{RO-Crate content} is further described with
+additional metadata following a Linked Data approach. Data can be
+embedded files and directories, as well as links to external Web
+resources, PIDs and nested RO-Crates.}
+\label{fig:conceptual}
+\end{figure}
+
+\subsubsection{Linked Data as a foundation}% {#linkeddata}
+
+\label{sec:linkeddata}
+
+The \textbf{Linked Data} principles \cite{doi:10.4018/978-1-60960-593-3.ch008}
+(use of IRIs\footnote{\textbf{IRI}s \cite{doi:10.17487/rfc3987} are a generalisation
+ of \textit{URI}s
+(which include well-known http/https URLs), permitting international
+Unicode characters without percent encoding, commonly used on the
+browser address bar and in HTML5.} to identify resources (i.e. artefacts), resolvable via
+HTTP, enriched with metadata and linked to each other) are core to
+RO-Crate; therefore IRIs are used to identify an RO-Crate, its
+constituent parts and metadata descriptions, and the properties and
+classes used in the metadata.
+
+RO-Crates are \textit{self-described} and follow the Linked Data principles to
+describe all of their resources in both human and machine readable
+manner. Hence, resources are identified using global identifiers
+(absolute IRIs) where possible; and relationships between two resources
+are defined with links.
+
+The foundation of Linked Data and shared vocabularies also means that
+multiple RO-Crates and other Linked Data resources can be indexed,
+combined, queried, validated or transformed using existing Semantic Web
+technologies such as SPARQL,\footnote{\url{https://www.w3.org/TR/sparql11-overview}}
+SHACL\footnote{\url{https://www.w3.org/TR/shacl/}} and well established \textit{knowledge
+graph} triple stores like Apache Jena\footnote{\url{https://jena.apache.org/}} and
+OntoText GraphDB.\footnote{\url{https://www.ontotext.com/products/graphdb/}}
+
+The possibilities of consuming\footnote{Some consideration is needed in processing of RO-Crates as
+knowledge graphs, e.g. establishing absolute URIs for files inside a
+ZIP archive, detailed in the RO-Crate specification: \url
+{https://www.researchobject.org/ro-crate/1.1/appendix/relative-uris.html}.} RO-Crate metadata with such
+powerful tools gives another strong reason for using Linked Data as a
+foundation. This use of mature Web\footnote{Note that an RO-Crate is not required to be published on the
+Web, see Section~\ref{sec:selfdescribed}.} technologies also means its
+developers and consumers are not restricted to the Research Object
+aspects that have already been specified by the RO-Crate community, but
+can extend and integrate RO-Crate in multiple standardized ways.
+
+\subsubsection{RO-Crate is a self-described container}% {#selfdescribed}
+
+\label{sec:selfdescribed}
+
+An RO-Crate is defined\footnote{\url{https://www.researchobject.org/ro-crate/1.1/structure.html\#ro-crate-metadata-file-ro-crate-metadatajson}}
+as a self-described \textbf{Root Data Entity} that describes and contains
+\textit{data entities}, which are further described by referencing {contextual
+entities}. A \textbf{data entity} is either a \textit{file} (i.e. a byte sequence
+stored on disk somewhere) or a \textit{directory} (i.e. set of named files and
+other directories). A~file does not need to be stored inside the
+RO-Crate root, it can be referenced via a PID/IRI. A \textbf{contextual
+entity} exists outside the information system (e.g. a Person, a
+workflow language) and is stored solely by its metadata. The
+representation of a \textbf{data entity} as a byte sequence makes it
+possible to store a variety of research artefacts including not only
+data but also, for instance, software and text.
+
+The Root Data Entity is a directory, the \textit{RO-Crate Root}, identified by
+the presence of the \textbf{RO-Crate Metadata File} \texttt{ro-crate-metadata.json}
+(top of Fig.~\ref{fig:conceptual}). This is a JSON-LD file that
+describes the RO-Crate, its content and related metadata using Linked
+Data. JSON-LD is a W3C standard RDF serialisation that has become
+popular as it is easy to read by humans while also offers some
+advantages for data exchange on the Internet. JSON-LD, a subset of the
+widely supported and well-known JSON format, has tooling available for
+many programming languages.\footnote{\url{https://json-ld.org/\#developers}}
+
+The minimal requirements for the root data entity
+metadata\footnote{\url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}}
+are \texttt{name}, \texttt{description} and \texttt{datePublished}, as well as a contextual
+entity identifying its \texttt{license} -- additional metadata are commonly
+added to entities depending on the purpose of the particular RO-Crate.
+
+RO-Crates can be stored, transferred or published in multiple ways,
+e.g. BagIt \cite{doi:10.17487/rfc8493}, Oxford Common File Layout
+\cite{ocfl_2020} (OCFL), downloadable ZIP archives in Zenodo or through
+dedicated online repositories, as well as published directly on the
+Web, e.g. using GitHub Pages.\footnote{\url{https://pages.github.com/}} Combined
+with Linked Data identifiers, this caters for a diverse set of storage
+and access requirements across different scientific domains, from
+metagenomics workflows producing hundreds of gigabytes of genome data
+to cultural heritage records with access restrictions for personally
+identifiable data. Specific \textit{RO-Crate profiles} (Section~\ref{sec:profiles}) may constrain serialization and publication
+expectations, and require additional contextual types and properties.
+
+\subsubsection{Data entities are described using contextual entities}% {#contextualentities}
+
+\label{sec:contextualentities}
+
+RO-Crate distinguishes between data and contextual
+entities\footnote{\url{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html\#contextual-vs-data-entities}}
+in a similar way to HTTP terminology's early attempt to separate
+\textit{information} (data) and \textit{non-information} (contextual) resources
+\cite{httprange14}. Data entities are usually files and directories located
+by relative IRI references within the RO-Crate Root, but they can also
+be Web resources or restricted data identified with absolute IRIs,
+including \textit{Persistent Identifiers} (PIDs) \cite{doi:10.1371/journal.pbio.2001414}.
+
+As both types of entities are identified by IRIs, their distinction is
+allowed to be blurry; data entities can be located anywhere and be
+complex, while contextual entities can have a Web presence beyond their
+description inside the RO-Crate. For instance
+\texttt{https://orcid.org/0000-0002-1825-0097} is primarily an identifier for
+a person, but secondarily it is also a Web page and a way to refer to
+their academic work.
+
+A particular IRI may appear as a contextual entity in one RO-Crate and
+as a data entity in another; the distinction lies in the fact that data
+entities can be considered to be \textit{contained} or captured by that
+RO-Crate (\textit{RO Content} in Fig.~\ref{fig:conceptual}), while
+contextual entities mainly \textit{explain} an RO-Crate or its content
+(although this distinction is not a formal requirement).
+
+In RO-Crate, a referenced contextual entity (e.g. a person identified
+by ORCID) should always be described within the RO-Crate Metadata File
+with at least a \textit{type} and \textit{name}, even where their PID might resolve
+to further Linked Data. This is so that clients are not required to
+follow every link for presentation purposes, for instance HTML
+rendering. Similarly any imported extension
+terms\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#extending-ro-crate}}
+would themselves also have a human-readable description in the case
+where their PID does not resolve to human-readable documentation.
+
+Figure~\ref{fig:uml} shows a simplified UML class diagram of RO-Crate,
+highlighting the different types of data entities and contextual
+entities that can be aggregated and related. While an RO-Crate would
+usually contain one or more data entities (\texttt{hasPart}), it may also be a
+pure aggregation of contextual entities (\texttt{mentions}).
+
+\begin{figure}%[t!]
+\includegraphics{210053f02}
+\caption{Simplified UML class diagram of RO-Crate. The \emph
+{RO-Crate Metadata File} conforms to a version of the specification;
+and contains a JSON-LD graph \cite{sporny_2014} that describes the
+entities that make up the RO-Crate. The \emph{RO-Crate Root Data
+Entity} represent the Research Object as a dataset. The RO-Crate
+aggregates \emph{data entities} (\texttt{hasPart}) which are further
+described using \emph{contextual entities} (which may include
+aggregated and non-aggregated data entities). Multiple types and
+relations from \surl{Schema.org} allow annotations to be more specific,
+including figures, nested datasets, computational workflows, people,
+organisations, instruments and places. Contextual entities not
+otherwise cross-referenced from other entities' properties (\emph
+{describes}) can be grouped under the root entity (\texttt{mentions}).}
+\label{fig:uml}
+\end{figure}
+
+\subsubsection{Guide through recommended practices}% {#recommendedpractices}
+
+\label{sec:recommendedpractices}
+
+RO-Crate as a specification aims to build a set of recommended
+practices on how to practically apply existing standards in a common
+way to describe research outputs and their provenance, without having
+to learn each of the underlying technologies in detail.
+
+As such, the RO-Crate 1.1\footnote{\url{https://w3id.org/ro/crate/1.1}}
+specification \cite{doi:10.5281/zenodo.4541002} can be seen as an
+opinionated and example-driven guide to writing
+\surl{Schema.org}\footnote{\url{https://schema.org/}} \cite{doi:10.1145/2857274.2857276}
+metadata as JSON-LD \cite{sporny_2014} (see Section~\ref{sec:implementation}), which leaves it open for implementers to include
+additional metadata using other \surl{Schema.org} types and properties, or
+even additional Linked Data vocabularies/ontologies or their own ad-hoc terms.
+
+However the primary purpose of the RO-Crate specification is to assist
+developers in leveraging Linked Data principles for the focused purpose
+of describing Research Objects in a structured language, while reducing
+the steep learning curve otherwise associated in Semantic Web
+adaptation, like development of ontologies, identifiers, namespaces,
+and RDF serialization choices.
+\subsubsection{Ensuring simplicity}% {#simplicity}
+
+\label{sec:implicitly}
+
+One aim of RO-Crate is to be conceptually simple. This simplicity has
+been repeatedly checked and confirmed through an informal community
+review process. For instance, in the discussion on supporting ad-hoc
+vocabularies\footnote{\url{https://github.com/ResearchObject/ro-crate/issues/71}} in
+RO-Crate, the community explored potential Linked Data solutions. The
+conventional wisdom in RDF best
+practices\footnote{\url{https://www.w3.org/TR/swbp-vocab-pub/}} is to establish a
+vocabulary with a new URI namespace, formalised using RDF
+Schema\footnote{\url{http://www.w3.org/TR/2014/REC-rdf-schema-20140225/}} or
+OWL\footnote{\url{http://www.w3.org/TR/2012/REC-owl2-overview-20121211/}}
+ontologies. However, this may seem an excessive learning curve for
+non-experts in semantic knowledge representation, and the RO-Crate
+community instead agreed on a dual lightweight approach: (i)
+Document\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#adding-new-or-ad-hoc-vocabulary-terms}}
+how projects with their own Web-presence can make a pure HTML-based
+vocabulary, and (ii) provide a community-wide PID namespace under
+\texttt{https://w3id.org/ro/terms} that redirect to simple CSV files
+maintained in GitHub.\footnote{\url{https://github.com/ResearchObject/ro-terms}}
+
+To further verify this idea of simplicity, we have formalised the
+RO-Crate definition (see \textit{Appendix~\ref{sec:formaldefinition})}). An
+important result of this exercise is that the underlying data structure
+of RO-Crate, although conceptually a graph, is represented as a
+depth-limited tree. This formalisation also emphasises the
+\textit{boundedness} of the structure; namely, the fact that elements are
+specifically identified as being either semantically \textit{contained} by the
+RO-Crate as \textit{Data Entities} (\texttt{hasPart}) or mainly referenced
+(\texttt{mentions}) and typed as \textit{external} to the Research Object as
+\textit{Contextual Entities}. It is worth pointing out that this semantic
+containment can extend beyond the physical containment of files
+residing within the RO-Crate Root directory on a given storage system,
+as the RO-Crate data entities may include any data resource globally
+identifiable using IRIs.
+\subsubsection{Extensibility and RO-Crate profiles}% {#profiles}
+
+\label{sec:profiles}
+
+The RO-Crate specification provides a core set of conventions to
+describe research outputs using types and properties applicable across
+scientific domains. However we have found that domain-specific use of
+RO-Crate will, implicitly or explicitly, form a specialized \textbf{profile}
+of RO-Crate; i.e., {a set of conventions, types and properties that are
+minimally required and one can expect to be present in that subset of
+RO-Crates}. For instance, RO-Crates used for exchange of workflows will
+have to contain a data entity of type \texttt{ComputationalWorkflow}, or
+cultural heritage records should have a \texttt{contentLocation}.
+
+Making such profiles explicit allow further reliable programmatic
+consumption and generation of RO-Crates beyond the core types defined
+in the RO-Crate specification. Following the RO-Crate mantra of
+\textit{guidance over strictness}, profiles are mainly \textit{duck-typing} rather
+than strict syntactic or semantic types, but may also have
+corresponding machine-readable schemas at multiple levels (file
+formats, JSON, RDF shapes, RDFS/OWL semantics).
+
+The next version of the RO-Crate specification 1.2 will define a
+formalization\footnote{\url{https://www.researchobject.org/ro-crate/1.2-DRAFT/profiles}}
+for publishing and declaring conformance to RO-Crate profiles. Such a
+profile is primarily a human-readable document of before-mentioned
+expectations and conventions, but may also define a machine-readable
+profile as a \textbf{Profile Crate}: Another RO-Crate that describe the
+profile and in addition can list schemas for validation, compatible
+software, applicable repositories, serialization/packaging formats,
+extension vocabularies, custom JSON-LD contexts and examples (see for
+example the Workflow RO-Crate
+profile\footnote{\url{https://w3id.org/workflowhub/workflow-ro-crate/}}).
+
+In addition, there are sometimes existing domain-specific metadata
+formats, but they are either not RDF-based (and thus time-consuming to
+construct terms for in JSON-LD) or are at a different granularity level
+that might become overwhelming if represented directly in the RO-Crate
+Metadata file (e.g. W3C PROV bundle detailing every step execution of a
+workflow run \cite{doi:10.1093/gigascience/giz095}). RO-Crate allows such
+\textit{alternative metadata files} to co-exist, and be described as data
+entities with references to the standards and vocabularies they conform
+to. This simplifies further programmatic consumption even where no
+filename or file extension conventions have emerged for those metadata formats.
+
+Section~\ref{sec:inuse} examines the observed specializations of
+RO-Crate use in several domains and their emerging profiles.
+
+\subsection{Technical implementation of the RO-Crate model}% {#implementation}
+
+\label{sec:implementation}
+
+The RO-Crate conceptual model has been realised using JSON-LD and
+\surl{Schema.org} in a prescriptive form as discussed in Section~\ref{sec:conceptual}. These technical choices were made to cater for
+simplicity from a developer perspective (as introduced in Section~\ref{sec:methodology}).
+
+JSON-LD\footnote{\url{https://json-ld.org/}} \cite{sporny_2014} provides a way to
+express Linked Data as a JSON structure, where a \textit{context} provides
+mapping to RDF properties and classes. While JSON-LD cannot map
+arbitrary JSON structures to RDF, we found that it does lower the
+barrier compared to other RDF syntaxes, as the JSON syntax nowadays is
+a common and popular format for data exchange on the Web.
+
+However, JSON-LD alone has too many degrees of freedom and hidden
+complexities for software developers to reliably produce and consume
+without specialised expertise or large RDF software frameworks. A large
+part of the RO-Crate specification is therefore dedicated to describing
+the acceptable subset of JSON structures.
+
+\subsubsection{RO-Crate JSON-LD}% {#jsonld}
+
+\label{sec:jsonld}
+
+RO-Crate
+mandates\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html}}
+the use of flattened, compacted JSON-LD in the RO-Crate Metadata file
+\texttt{ro-crate-metadata.json}\footnote{The avid reader may spot that the RO-Crate Metadata file use the
+extension \texttt{.json} instead of \texttt{.jsonld}, this is to emphasize the
+developer expectations as a JSON format, while the file's JSON-LD
+nature is secondary. See \url
+{https://github.com/ResearchObject/ro-crate/issues/82}.}
+where a single \texttt{@graph} array contains all
+the data and contextual entities in a flat list. An example can be seen
+in the JSON-LD snippet in Listing \ref{lis1} below, describing a simple RO-Crate
+containing data entities described using contextual entities.
+
+\begin{listing}
+\includegraphics{210053i02}
+\caption{%\qq{}\protect\footnotemark[31]
+Simplified\protect\footnote{Recommended properties for types shown in
+Listing \ref{lis1} also include
+\texttt{affiliation}, \texttt{citation}, \texttt{contactPoint},
+\texttt{description},
+\texttt{encodingFormat}, \texttt{funder}, \texttt{geo}, \texttt{identifier},
+\texttt{keywords},
+\texttt{publisher}; these properties and corresponding contextual
+entities are
+excluded here for brevity. See complete example
+\url
+{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.}
+RO-Crate metadata file showing the
+flattened compacted JSON-LD \texttt{@graph} array containing the data entities
+and contextual entities, cross-referenced using \texttt{@id}. The
+\texttt{ro-crate-metadata.json} entity self-declares conformance with the
+RO-Crate specification using a versioned persistent identifier, further
+RO-Crate descriptions are on the root data entity \texttt{./} or any of the
+referenced data or contextual entities, as exemplified by the data
+entity \texttt{ImageObject} referencing contextual entities for
+\texttt{contentLocation} and \texttt{author} that differs from that of the overall
+RO-Crate. In this example, \texttt{about} of the CSV data entity reference the
+\texttt{ImageObject}, which then take the roles of both a data entity and
+contextual entity. While \texttt{Person} entities ideally are identified with
+ORCID PIDs as for Josiah, \texttt{\#alice} is here in contrast an RO-Crate
+local identifier, highlighting the pragmatic ``just enough'' Linked
+Data approach.}\label{lis1}
+\end{listing}
+
+In\footnotetext{Recommended properties for types shown in
+Listing \ref{lis1} also include
+\texttt{affiliation}, \texttt{citation}, \texttt{contactPoint},
+\texttt{description},
+\texttt{encodingFormat}, \texttt{funder}, \texttt{geo}, \texttt{identifier},
+\texttt{keywords},
+\texttt{publisher}; these properties and corresponding contextual
+entities are
+excluded here for brevity. See complete example
+\url
+{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.} this flattened profile of JSON-LD, each \texttt{{entity}} are directly
+under \texttt{@graph} and represents the RDF triples with a common \textit{subject}
+(\texttt{@id}), mapped \textit{properties} like \texttt{hasPart}, and \textit{objects} -- as
+either literal \texttt{"string"} values, referenced \texttt{{objects}} (which
+properties are listed in its own entity), or a JSON \texttt{[list]} of these.
+If processed as JSON-LD, this forms an RDF graph by matching the \texttt{@id}
+IRIs and applying the \texttt{@context} mapping to \surl{schema.org} terms.
+\normalsize
+
+ \subsubsection{Flattened JSON-LD}
+
+When JSON-LD 1.0 \cite{sporny_2014} was proposed, one of the motivations
+was to seamlessly apply an RDF nature on top of regular JSON as
+frequently used by Web APIs. JSON objects in APIs are frequently nested
+with objects at multiple levels, and the perhaps most common form of
+JSON-LD is the compacted
+form\footnote{\url{https://json-ld.org/spec/REC/json-ld/20140116/\#compacted-document-form}}
+which follows this expectation (JSON-LD
+1.1\footnote{\url{https://www.w3.org/TR/2020/REC-json-ld11-20200716/}} further
+expands these capabilities, e.g. allowing nested \texttt{@context} definitions).
+
+While this feature of JSON-LD can be seen as a way to ``hide'' its RDF
+nature, we found that the use of nested trees (e.g. a \texttt{Person} entity
+appearing as \texttt{author} of a \texttt{File} which nests under a \texttt{Dataset} with
+\texttt{hasPart}) counter-intuitively forces consumers to consider the JSON-LD
+as an RDF Graph, since an identified \texttt{Person} entity can appear at
+multiple and repeated points of the tree (e.g. author of multiple
+files), necessitating node merging or duplication, which can become
+complicated as this approach also invites the use of \textit{blank nodes}
+(entities missing \texttt{@id}).
+
+By comparison, a single flat \texttt{@graph} array approach, as required by
+RO-Crate, means that applications can choose to process and edit each
+entity as pure JSON by a simple lookup based on \texttt{@id}. At the same
+time, lifting all entities to the same level reflects the Research
+Object principles \cite{doi:10.1016/j.future.2011.08.004} in that
+describing the context and provenance is just as important as
+describing the data, and the requirement of \texttt{@id} of every entity
+forces RO-Crate generators to consciously consider existing IRIs and
+identifiers.\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#describing-entities-in-json-ld}}
+
+\subsubsection{JSON-LD context}
+
+In JSON-LD, the \texttt{@context} is a reference to another JSON-LD document
+that provides mapping from JSON keys to Linked Data term IRIs, and can
+enable various JSON-LD directives to cater for customized JSON
+structures for translating to RDF.
+
+RO-Crate reuses vocabulary terms and IRIs from \surl{Schema.org}, but provides
+its own versioned JSON-LD
+context,\footnote{\url{https://w3id.org/ro/crate/1.1/context}} which has a flat list
+with the mapping from JSON-LD keys to their URI equivalents (e.g. key
+\texttt{"author"} maps to the \url{http://schema.org/author} property).
+
+The rationale behind this decision is to support JSON-based RO-Crate
+applications that are largely unaware of JSON-LD, that still may want
+to process the \texttt{@context} to find or add Linked Data definitions of
+otherwise unknown properties and types. Not reusing the official
+\surl{Schema.org} context means RO-Crate is also able to map in additional
+vocabularies where needed, namely the \textit{Portland Common Data Model}
+(PCDM) \cite{pcdm} for repositories and Bioschemas \cite{bioschemas_2017} for
+describing computational workflows. RO-Crate profiles may
+extend\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#extending-ro-crate}}
+the \texttt{@context} to re-use additional domain-specific ontologies.
+
+Similarly, while the \surl{Schema.org} context
+currently\footnote{\url{https://schema.org/version/13.0/schemaorg-current-http.jsonld}}
+have \texttt{"@type": "@id"} annotations for object properties, RO-Crate
+JSON-LD distinguishes explicitly between references to other entities
+(\texttt{{"@id": "\#alice"}})
+and string values (\texttt{"Alice"}) -- meaning
+RO-Crate applications can find references for corresponding entities
+and IRIs without parsing the \texttt{@context} to understand a particular
+property. Notably this is exploited by the \texttt{ro-crate-html-js}
+\cite{ro-crate-html-js} tool to provide reliable HTML rendering for
+otherwise unknown properties and types.
+
+\subsection{RO-Crate community}% {#community}
+
+\label{sec:community}
+
+The RO-Crate conceptual model, implementation and best practices are
+developed by a growing community of researchers, developers and
+publishers. RO-Crate's community is a key aspect of its effectiveness
+in making research artefacts FAIR. Fundamentally, the community
+provides the overall context of the implementation and model and
+ensures its interoperability.
+
+The RO-Crate community consists of:
+\begin{enumerate}
+\item[1.] a diverse set of people representing a variety of stakeholders;
+\item[2.] a set of collective norms;
+\item[3.] an open platform that facilitates communication (GitHub, Google
+Docs, monthly teleconferences).
+\end{enumerate}
+\subsubsection{People}
+
+The initial concept of RO-Crate was formed at the first Workshop on
+Research Objects (RO2018\footnote{\url{https://www.researchobject.org/ro2018/}}),
+held as part of the IEEE conference on eScience. This workshop followed
+up on considerations made at a Research Data Alliance (RDA) meeting on
+Research Data
+Packaging\footnote{\url{https://rd-alliance.org/approaches-research-data-packaging-rda-11th-plenary-bof-meeting}}
+that found similar goals across multiple data packaging efforts
+\cite{doi:10.5281/zenodo.3250687}: simplicity, structured metadata and the
+use of JSON-LD.
+
+An important outcome of discussions that took place at RO2018 was the
+conclusion that the original Wf4Ever Research Object ontologies
+\cite{doi:10.1016/j.websem.2015.01.003}, in principle sufficient for
+packaging research artefacts with rich descriptions, were, in practice,
+considered inaccessible for regular programmers (e.g., Web developers)
+and in danger of being incomprehensible for domain scientists due to
+their reliance on Semantic Web technologies and other ontologies.
+
+DataCrate \cite{doi:10.5281/zenodo.1445817} was presented at RO2018 as a
+promising lightweight alternative approach, and an agreement was made
+by a group of volunteers to attempt building what was initially called
+\textit{``RO Lite''} as a combination of DataCrate's implementation and
+Research Object's principles.
+
+This group, originally made up of library and Semantic Web experts, has
+subsequently grown to include domain scientists, developers, publishers
+and more. This perspective of multiple views led to the specification
+being used in a variety of domains, from bioinformatics and regulatory
+submissions to humanities and cultural heritage preservation.
+
+The RO-Crate community is strongly engaged with the European-wide
+biology/bioinformatics collaborative e-Infrastructure, ELIXIR
+\cite{doi:10.1016/j.tibtech.2012.02.002}, along with European Open Science
+Cloud (EOSC\footnote{\url{https://eosc.eu/}}) projects including
+EOSC-Life,\footnote{\url{https://www.eosc-life.eu/}}
+FAIRplus,\footnote{\url{https://fairplus-project.eu/}}
+CS3MESH4EOSC\footnote{\url{https://cs3mesh4eosc.eu/}} and
+BY-COVID.\footnote{\url{https://by-covid.eu/}} RO-Crate has also established
+collaborations with Bioschemas \cite{bioschemas_2017}, GA4GH
+\cite{doi:10.1016/j.xgen.2021.100029}, OpenAIRE \cite{rettberg_2015_openaire}
+and multiple H2020 projects.
+
+A key set of stakeholders are developers: the RO-Crate community has
+made a point of attracting developers who can implement the
+specifications but, importantly, keeps ``developer user experience'' in
+mind. This means that the specifications are straightforward to
+implement and thus do not require expertise in technologies that are
+not widely deployed.
+
+This notion of catering to ``developer user experience'' is an example
+of the set of norms that have developed and now define the community.
+
+\subsubsection{Norms}
+
+The RO-Crate community is driven by informal conventions and notions
+that are prevalent but not neccessarily written down. Here, we distil
+what we as authors believe are the critical set of norms that have
+facilitated the development of RO-Crate and contributed to the ability
+for RO-Crate research packages to be FAIR. This is not to say that
+there are no other norms within the community nor that everyone in the
+community holds these uniformly. Instead, what we emphasise is that
+these norms are helpful and also shaped by community practices.
+\begin{enumerate}
+\item[1.] Simplicity
+\item[2.] Developer friendliness
+\item[3.] Focus on examples and best practices rather than rigorous specification
+\item[4.] Reuse ``just enough'' Web standards
+\end{enumerate}
+
+A core norm of RO-Crate is that of \textbf{simplicity}, which sets the scene
+for how we guide developers to structure metadata with RO-Crate. We
+focus mainly on documenting simple approaches to the most common use
+cases, such as authors having an affiliation. This norm also influences
+our take on \textbf{developer friendliness}; for instance, we are using the
+Web-native JSON format, allowing only a few of JSON-LD's flexible
+Linked Data features. Moreover, the RO-Crate documentation is largely
+built up by \textbf{examples} showcasing \textbf{best practices}, rather than
+rigorous specifications. We build on existing \textbf{Web standards} that
+themselves are defined rigorously, which we utilise \textit{``\textbf{just
+enough}''} in order to benefit from the advantages of Linked Data
+(e.g., extensions by namespaced vocabularies), without imposing too
+many developer choices or uncertainties (e.g., having to choose between
+the many RDF syntaxes).
+
+While the above norms alone could easily lead to the creation of ``yet
+another'' JSON format, we keep the goal of \textbf{FAIR interoperability} of
+the captured metadata, and therefore follow closely FAIR best practices
+and current developments such as data citations, PIDs, open
+repositories and recommendations for sharing research outputs and software.
+
+\subsubsection{Open platforms}
+
+The critical infrastructure that enables the community around RO-Crate
+is the use of open development platforms. This underpins the importance
+of open community access to supporting FAIR. Specifically, it is
+difficult to build and consume FAIR research artefacts without being
+able to access the specifications, understand how they are developed,
+know about any potential implementation issues, and discuss usage to
+evolve best practices.
+
+The development of RO-Crate was driven by capturing documentation of
+real-life examples and best practices rather than creating a rigorous
+specification. At the same time, we agreed to be opinionated on the
+syntactic form to reduce the jungle of implementation choices; we
+wanted to keep the important aspects of Linked Data to adhere to the
+FAIR principles while retaining the option of combining and extending
+the structured metadata using the existing Semantic Web stack, not just
+build a standalone JSON format.
+
+Further work during 2019 started adapting the DataCrate documentation
+through a more collaborative and exploratory \textit{RO Lite} phase, initially
+using Google Docs for review and discussion, then moving to GitHub as a
+collaboration space for developing what is now the RO-Crate
+specification, maintained as
+Markdown\footnote{\url{https://github.com/researchobject/ro-crate/}} in GitHub Pages
+and published through Zenodo.
+
+In addition to the typical Open Source-style development with GitHub
+issues and pull requests, the RO-Crate Community have, at time of
+writing, two regular monthly calls, a Slack channel and a mailing list
+for coordinating the project; also many of its participants collaborate
+on RO-Crate at multiple conferences and coding events such as the
+ELIXIR BioHackathon.\footnote{\url{https://biohackathon-europe.org/}} The community
+is jointly developing the RO-Crate specification and Open Source tools,
+as well as providing support and considering new use cases. The
+RO-Crate Community\footnote{\url{https://www.researchobject.org/ro-crate/community}}
+is open for anyone to join, to equally participate under a code of
+conduct, and as of October 2021 has more than 50 members (see Appendix~\ref{sec:communitylist}).
+
+% Please update in 31.tooling-table.md first!
+%
+\begin{table}[t!]%[htbp]
+\tabcolsep=0pt
+\caption{Applications and libraries implementing RO-Crate, targeting
+different types of users across multiple programming languages. Status
+is indicative as assessed by this work (Alpha $<$ Beta $<$ Release
+Candidate (RC) $<$ Release)}
+\label{tab1}
+%
+\begin{tabular*}{\textwidth}{@{\extracolsep{4in minus 4in}}llcc@{}}
+\hline
+\textbf{Tool name} & \multicolumn{1}{c}{\tmultirow{2}{*}{Targets}}
+& \tmultirow{2}{*}{Language/Platform} & \tmultirow{2}{*}{Status} \\
+\textit{Brief Description}&&& \\
+\hline
+\textbf{Describo} \cite{describo}                                                                                                          & Research Data Managers     & NodeJS (Desktop) & RC    \\
+\multicolumn{4}{@{}l@{}}{\textit{Interactive desktop application to create, update and export RO-Crates for different profiles}}                                                                               \\
+\textbf{Describo Online} \cite{describo-online}                                                                                            & Platform developers        & NodeJS (Web)     & Alpha \\
+\multicolumn{4}{@{}l@{}}{\textit{Web-based application to create RO-Crates using cloud storage}}                                                                                                               \\
+\textbf{ro-crate-excel} \cite{ro-crate-excel}                                                                                              & Data managers              & JavaScript       &       \\
+\multicolumn{4}{@{}l@{}}{\textit{Command-line tool to create/edit RO-Crates with spreadsheets}}                                                                                                                \\
+\textbf{ro-crate-html-js} \cite{ro-crate-html-js}                                                                                          & Developers                 & JavaScript       & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{HTML rendering of RO-Crate}}                                                                                                                                                  \\
+\textbf{ro-crate-js} \cite{ro-crate-js}                                                                                                    & Research Data Managers     & JavaScript       & Alpha \\
+\multicolumn{4}{@{}l@{}}{\textit{Library for creating/manipulating crates; basic validation code}}                                                                                                             \\
+\textbf{ro-crate-ruby} \cite{ro-crate-ruby}                                                                                                & Developers                 & Ruby             & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{Ruby library for reading/writing RO-Crate, with workflow support}}                                                                                                            \\
+\textbf{ro-crate-py} \cite{ro-crate-py}                                                                                                    & Developers                 & Python           & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{Object-oriented Python library for reading/writing RO-Crate and use by Jupyter Notebook}}                                                                                     \\
+\textbf{WorkflowHub} \cite{about-workflowhub}                                                                                              & Workflow users             & Ruby             & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{Workflow repository; imports and exports Workflow RO-Crate}}                                                                                                                  \\
+\textbf{Life Monitor} \cite{about-lifemonitor}                                                                                             & Workflow developers        & Python           & Alpha \\
+\multicolumn{4}{@{}l@{}}{\textit{Workflow testing and monitoring service; Workflow Testing profile of RO-Crate}}                                                                                               \\
+\textbf{SCHeMa} \cite{vergoulis2021schema}                                                                                                 & Workflow users             & PHP              & Alpha \\
+\multicolumn{4}{@{}l@{}}{\textit{Workflow execution using RO-Crate as exchange mechanism}}                                                                                                                     \\
+\textbf{galaxy2cwl} \cite{galaxy2cwl}                                                                                                      & Workflow developers        & Python           & Alpha \\
+\multicolumn{4}{@{}l@{}}{\textit{Wraps Galaxy workflow as Workflow RO-Crate}}                                                                                                                                  \\
+\textbf{Modern PARADISEC} \cite{modpdsc}                                                                                                   & Repository managers        & Platform         & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{Cultural Heritage portal based on OCFL and RO-Crate}}                                                                                                                        \\
+\textbf{ONI express} \cite{arkisto-data-portal}                                                                                            & Repository managers        & Platform         & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{Platform for publishing data and documents stored in an OCFL repository via a Web interface}}                                                                                 \\
+\textbf{ocfl-tools} \cite{ocfl-tools}                                                                                                      & Developers                 & JavaScript (CLI) & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{Tools for managing RO-Crates in an OCFL repository}}                                                                                                                          \\
+\textbf{RO Composer} \cite{ro-composer}                                                                                                    & Repository developers      & Java             & Alpha \\
+\multicolumn{4}{@{}l@{}}{\textit{REST API for gradually building ROs for given profile}}                                                                                                                       \\
+\textbf{RDA maDMP Mapper} \cite{doi:10.5281/zenodo.3922136}                                                                                & Data Management Plan users & Python           & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{Mapping between machine-actionable data management plans (maDMP) and RO-Crate \cite{doi:10.4126/frl01-006423291}}}                                                         \\
+\textbf{Ro-Crate\_2\_ma-DMP} \cite{doi:10.5281/zenodo.3903463}                                                                             & Data Management Plan users & Python           & Beta  \\
+\multicolumn{4}{@{}l@{}}{\textit{Convert between machine-actionable data management plans (maDMP) and RO-Crate}}                                                                                              \\
+\textbf{CheckMyCrate} \cite{CheckMyCrate}                                                                                                  & Developers                 & Python (CLI)     & Alpha \\
+\multicolumn{4}{@{}l@{}}{\textit{Validation according to Workflow RO-Crate profile}}                                                                                                                           \\
+\textbf{RO-Crates-and-Excel} \cite{doi:10.5281/zenodo.5068950}                                                                             & Data Managers              & Java (CLI)       & Alpha \\
+\multicolumn{4}{@{}l@{}}{\textit{Describe column/data details of spreadsheets as RO-Crate using DataCube vocabulary}}                                                                                          \\
+\hline
+\end{tabular*}\vspace*{10pt}
+\end{table}
+
+\section{RO-Crate tooling}% {#tooling}
+
+\label{sec:tooling}
+
+The work of the community has led to the development of a number of
+tools for creating and using RO-Crates. Table~\ref{tab1} shows the current set
+of implementations. Reviewing this list, one can see that tools support
+commonly used programming languages, including Python, JavaScript, and
+Ruby. Additionally, these tools can be integrated into commonly used
+research environments, in particular, the command line tool
+\textit{ro-crate-html-js} \cite{ro-crate-html-js} for creating a human-readable
+preview of an RO-Crate as a sidecar HTML file. Furthermore, there are
+tools that cater to end-users (\textit{Describo} \cite{describo}, \textit{WorkflowHub}
+\cite{about-workflowhub}), in order to simplify creating and managing
+RO-Crate. For example, Describo was developed to help researchers of
+the Australian Criminal Characters
+project\footnote{\url{https://criminalcharacters.com/}} annotate historical prisoner
+records to gain greater insight into the history of Australia
+\cite{doi:10.1080/14490854.2020.1796500}.
+
+While the development of these tools is promising, our analysis of
+their maturity status shows that the majority of them are in the Beta
+stage. This is partly due to the fact that the RO-Crate specification
+itself only recently reached 1.0 status, in November 2019
+\cite{doi:10.5281/zenodo.3541888}. Now that there is a fixed point of
+reference: With version 1.1 (October 2020)
+\cite{doi:10.5281/zenodo.4031327} RO-Crate has stabilised based on feedback
+from application development, and now we are seeing a further increase
+in the maturity of these tools, along with the creation of new ones.
+
+Given the stage of the specification, these tools have been primarily
+targeting developers, essentially providing them with the core
+libraries for working with RO-Crate. Another target has been that of
+research data managers who need to manage and curate large amounts of data.
+
+\section{Profiles of RO-Crate in use}% {#inuse}
+
+\label{sec:inuse}
+
+RO-Crate fundamentally forms part of an infrastructure to help build
+FAIR research artefacts. In other words, the key question is whether
+RO-Crate can be used to share and (re)use research artefacts. Here we
+look at three research domains where RO-Crate is being applied:
+Bioinformatics, Regulatory Science and Cultural Heritage. In addition,
+we note how RO-Crate may have an important role as part of
+machine-actionable data management plans and institutional repositories.
+
+From these varied uses of RO-Crate we observe natural differences in
+their detail level and the type of entities described by the RO-Crate.
+For instance, on submission of an RO-Crate to a workflow repository, it
+is reasonable to expect the RO-Crate to contain at least one workflow,
+ideally with a declared licence and workflow language. Specific
+additional recommendations such as on identifiers is also needed to
+meet the emerging requirements of FAIR Digital
+Objects.\footnote{\url{https://fairdo.org/}} Work has now
+begun\footnote{\url{https://github.com/ResearchObject/ro-crate/issues/153}} to
+formalise these different \textit{profiles} of RO-Crates, which may impose
+additional constraints based on the needs of a specific domain or use case.
+
+\subsection{Bioinformatics workflows}% {#workflows}
+
+\label{sec:workflows}
+
+WorkflowHub.eu\footnote{\url{https://workflowhub.eu/}} is a European cross-domain
+registry of computational workflows, supported by European Open Science
+Cloud projects, e.g. EOSC-Life,\footnote{\url{https://www.eosc-life.eu/}} and
+research infrastructures including the pan-European bioinformatics
+network ELIXIR\footnote{\url{https://elixir-europe.org/}}
+\cite{doi:10.1016/j.tibtech.2012.02.002}. As part of promoting workflows as
+reusable tools, WorkflowHub includes documentation and high-level
+rendering of the workflow structure independent of its native workflow
+definition format. The rationale is that a domain scientist can browse
+all relevant workflows for their domain, before narrowing down their
+workflow engine requirements. As such, the WorkflowHub is intended
+largely as a registry of workflows already deposited in repositories
+specific to particular workflow languages and domains, such as
+UseGalaxy.eu \cite{doi:10.1371/journal.ppat.1008643} and Nextflow nf-core
+\cite{doi:10.1038/s41587-020-0439-x}.
+
+We here describe three different RO-Crate profiles developed for use
+with WorkflowHub.
+
+\subsubsection{Profile for describing workflows}
+
+Being cross-domain, WorkflowHub has to cater for many different
+workflow systems. Many of these, for instance Nextflow
+\cite{doi:10.1038/nbt.3820} and Snakemake
+\cite{doi:10.1093/bioinformatics/bts480}, by virtue of their script-like
+nature, reference multiple neighbouring files typically maintained in a
+GitHub repository. This calls for a data exchange method that allows
+keeping related files together. WorkflowHub has tackled this problem by
+adopting RO-Crate as the packaging mechanism
+\cite{doi:10.5281/zenodo.4705078}, typing and annotating the constituent
+files of a workflow and -- crucially -- marking up the workflow
+language, as many workflow engines use common file extensions like
+\texttt{*.xml} and \texttt{*.json}. Workflows are further described with authors,
+license, diagram previews and a listing of their inputs and outputs.
+RO-Crates can thus be used for interoperable deposition of workflows to
+WorkflowHub, but are also used as an archive for downloading workflows,
+embedding metadata registered with the WorkflowHub entry and translated
+workflow files such as abstract Common Workflow Language (CWL)
+\cite{doi:10.1145/3486897} definitions and diagrams \cite{doi:10.5281/zenodo.4605654}.
+
+RO-Crate acts therefore as an interoperability layer between
+registries, repositories and users in WorkflowHub. The iterative
+development between WorkflowHub developers and the RO-Crate community
+heavily informed the creation of the Bioschemas \cite{bioschemas_2017}
+profile for Computational
+Workflows,\footnote{\url{https://bioschemas.org/profiles/ComputationalWorkflow/1.0-RELEASE/}}
+which again informed the RO-Crate 1.1 specification on
+workflows\footnote{\url{https://www.researchobject.org/ro-crate/1.1/workflows.html}}
+and led to the RO-Crate Python library \cite{ro-crate-py} and WorkflowHub's
+\textbf{Workflow RO-Crate
+profile},\footnote{\url{https://w3id.org/workflowhub/workflow-ro-crate/1.0}} which,
+in a similar fashion to RO-Crate itself, recommends which workflow
+resources and descriptions are required. This co-development across
+project boundaries exemplifies the drive for simplicity and for
+establishing best practices.
+
+\subsubsection{Profile for recording workflow runs}
+
+RO-Crates in WorkflowHub have so far been focused on workflows that are
+ready to be run, and development of WorkflowHub is now creating a
+\textbf{Workflow Run RO-Crate profile} for the purposes of benchmarking,
+testing and executing workflows. As such, RO-Crate serves as a
+container of both a \textit{workflow definition} that may be executed and of a
+particular \textit{workflow execution with test results}.
+
+This workflow run profile is a continuation of our previous work with
+capturing workflow provenance in a Research Object in CWLProv
+\cite{doi:10.1093/gigascience/giz095} and TavernaPROV
+\cite{doi:10.5281/zenodo.51314}. In both cases, we used the PROV Ontology
+\cite{CheckMyCrate}, including details of every task execution with all the
+intermediate data, which required significant workflow engine
+integration.\footnote{CWLProv and TavernaProv predate RO-Crate, but use
+RO-Bundle\cite{doi:10.5281/zenodo.12586}, a similar Research Object
+packaging method with JSON-LD metadata.}
+
+Simplifying from the CWLProv approach, the planned Workflow Run
+RO-Crate profile will use a high level \surl{Schema.org}
+provenance\footnote{\url{https://www.researchobject.org/ro-crate/1.1/provenance.html\#software-used-to-create-files}}
+for the input/output boundary of the overall workflow execution. This
+\textit{Level 1 workflow provenance} \cite{doi:10.1093/gigascience/giz095} can be
+expressed generally across workflow languages with minimal workflow
+engine changes, with the option of more detailed provenance traces as
+separate PROV artefacts in the RO-Crate as data entities. In the
+current development of Specimen Data
+Refinery\footnote{\url{https://github.com/DiSSCo/SDR}} \cite{doi:10.3897/rio.6.e57602}
+these RO-Crates will document the text recognition workflow runs of
+digitised biological specimens, exposed as FAIR Digital Objects
+\cite{doi:10.3390/publications8020021}.
+
+WorkflowHub has recently enabled minting of Digital Object Identifiers
+(DOIs), a PID commonly used for scholarly artefacts, for registered
+workflows, e.g. \texttt{10.48546/workflowhub.workflow.56.1}
+\cite{doi:10.48546/workflowhub.workflow.56.1}, lowering the barrier for
+citing workflows as computational methods along with their FAIR
+metadata -- captured within an RO-Crate. While it is not an aim for
+WorkflowHub to be a repository of workflow runs and their data,
+RO-Crates of \textit{exemplar workflow runs} serve as useful workflow
+documentation, as well as being an exchange mechanism that preserves
+FAIR metadata in a diverse workflow execution environment.
+
+\subsubsection{Profile for testing workflows}
+
+The value of computational workflows, however, is potentially
+undermined by the ``collapse'' over time of the software and services
+they depend upon: for instance, software dependencies can change in a
+non-backwards-compatible manner, or active maintenance may cease; an
+external resource, such as a reference index or a database query
+service, could shift to a different URL or modify its access protocol;
+or the workflow itself may develop hard-to-find bugs as it is updated.
+This \textit{workflow decay} can take a big toll on the workflow's reusability
+and on the reproducibility of any processes it evokes
+\cite{doi:10.1109/eScience.2012.6404482}.
+
+For this reason, WorkflowHub is complemented by a monitoring and
+testing service called LifeMonitor \cite{about-lifemonitor}, also supported
+by EOSC-Life. LifeMonitor's main goal is to assist in the creation,
+periodic execution and monitoring of workflow tests, enabling the early
+detection of software collapse in order to minimise its detrimental
+effects. The communication of metadata related to workflow testing is
+achieved through the adoption of a \textbf{Workflow Testing RO-Crate
+profile}\footnote{\url{https://lifemonitor.eu/workflow_testing_ro_crate}} stacked on
+top of the \textit{Workflow RO-Crate} profile. This further specialisation of
+Workflow RO-Crate allows to specify additional testing-related entities
+(test suites, instances, services, etc.), leveraging RO-Crate's
+extension
+mechanism\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#extending-ro-crate}}
+through the addition of terms from custom namespaces.
+
+In addition to showcasing RO-Crate's extensibility, the testing profile
+is an example of the format's flexibility and adaptability to the
+different needs of the research community. Though ultimately related to
+a computational workflow, in fact, most of the testing-specific
+entities are more about describing a protocol for interacting with a
+monitoring service than a set of research outputs and its associated
+metadata. Indeed, one of LifeMonitor's main functionalities is
+monitoring and reporting on test suites running on existing Continuous
+Integration (CI) services, which is described in terms of service URLs
+and job identifiers in the testing profile. In principle, in this
+context, data could disappear altogether, leading to an RO-Crate
+consisting entirely of contextual entities. Such an RO-Crate acts more
+as an exchange format for communication between services (WorkflowHub
+and LifeMonitor) than as an aggregator for research data and metadata,
+providing a good example of the format's high versatility.
+
+\subsection{Regulatory sciences}% {#regulatorysciences}
+
+\label{sec:regulatorysciences}
+
+BioCompute Objects\footnote{\url{https://biocomputeobject.org/}} (BCO)
+\cite{doi:10.1371/journal.pbio.3000099} is a community-led effort to
+standardise submissions of computational workflows to biomedical
+regulators. For instance, a genomics sequencing pipeline, as part of a
+personalised cancer treatment study, can be submitted to the US Food
+and Drugs Administration (FDA) for approval. BCOs are formalised in the
+standard IEEE 2791-2020 \cite{doi:10.1109/IEEESTD.2020.9094416} as a
+combination of JSON
+Schemas\footnote{\url{https://opensource.ieee.org/2791-object/ieee-2791-schema/}}
+that define the structure of JSON metadata files describing exemplar
+workflow runs in detail, covering aspects such as the usability and
+error domain of the workflow, its runtime requirements, the reference
+datasets used and representative output data produced.
+
+BCOs provide a structured view over a particular workflow, informing
+regulators about its workings independently of the underlying workflow
+definition language. However, BCOs have only limited support for
+additional metadata.\footnote{IEEE 2791-2020 do permit user extensions in the \textit{extension
+domain} by referencing additional JSON Schemas.} For instance, while the BCO itself can
+indicate authors and contributors, and in particular regulators and
+their review decisions, it cannot describe the provenance of individual
+data files or workflow definitions.
+
+As a custom JSON format, BCOs cannot be extended with Linked Data
+concepts, except by adding an additional top-level JSON object
+formalised in another JSON Schema. A BCO and workflow submitted by
+upload to a regulator will also frequently consist of multiple
+cross-related files. Crucially, there is no way to tell whether a given
+\texttt{*.json} file is a BCO file, except by reading its content and check
+for its \texttt{spec\_version}.
+
+We can then consider how a BCO and its referenced artefacts can be
+packaged and transferred following FAIR principles. \textbf{BCO
+RO-Crate}\footnote{\url{https://biocompute-objects.github.io/bco-ro-crate/}} \cite{doi:10.5281/zenodo.4633732},
+part of the BioCompute Object user guides, defines a set of best
+practices for wrapping a BCO with a workflow, together with its
+exemplar outputs in an RO-Crate, which then provides typing and
+additional provenance metadata of the individual files, workflow
+definition, referenced data and the BCO metadata itself.
+
+Here the BCO is responsible for describing the \textit{purpose} of a workflow
+and its run at an abstraction level suitable for a domain scientist,
+while the more open-ended RO-Crate describes the surroundings of the
+workflow, classifying and relating its resources and providing
+provenance of their existence beyond the BCO. This emerging \textit{separation
+of concerns} is shown in Fig.~\ref{fig:sep_concerns}, and highlights
+how RO-Crate is used side-by-side of existing standards and tooling,
+even where there are apparent partial overlaps.
+
+A similar separation of concerns can be found if considering the
+RO-Crate as a set of files, where the \textit{transport-level} metadata, such
+as checksum of files, are delegated to separate
+BagIt\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/implementation-notes.html\#adding-ro-crate-to-bagit}}
+manifests, a standard focusing on the preservation challenges of
+digital libraries \cite{doi:10.17487/rfc8493}. As such, RO-Crate metadata
+files are not required to iterate all the files in their folder
+hierarchy, only those that benefit from being described.
+
+Specifically, a BCO description alone is insufficient for reliable
+re-execution of a workflow, which would need a compatible workflow
+engine depending on the original workflow definition language, so IEEE
+2791 recommends using Common Workflow Language (CWL)
+\cite{doi:10.1145/3486897} for interoperable pipeline execution. CWL itself
+relies on tool packaging in software containers using
+Docker\footnote{\url{https://www.docker.com/}} or Conda.\footnote{\url{https://docs.conda.io/}}
+Thus, we can consider BCO RO-Crate as a stack: transport-level
+manifests of files (BagIt), provenance, typing and context of those
+files (RO-Crate), workflow overview and purpose (BCO), interoperable
+workflow definition (CWL) and tool distribution (Docker).
+
+%
+\begin{figure}%[t]
+\includegraphics{210053f03}
+\caption{Separation of Concerns in BCO RO-Crate. BioCompute
+Object (IEEE2791) is a JSON file that structurally explains the purpose
+and implementation of a computational workflow, for instance
+implemented in Common Workflow Language (CWL), that installs the
+workflow's software dependencies as Docker containers or BioConda
+packages. An example execution of the workflow shows the different
+kinds of result outputs, which may be external, using GitHub LFS \cite
+{github-lfs} to support larger data. RO-Crate gathers all these local
+and external resources, relating them and giving individual
+descriptions, for instance permanent DOI identifiers for reused
+datasets accessed from Zenodo, but also adding external identifiers to
+attribute authors using ORCID or to identify which licences apply to
+individual resources. The RO-Crate and its local files are captured in
+a BagIt whose checksum ensures completeness, combined with Big Data Bag
+\cite{doi:10.1109/BigData.2016.7840618} features to ``complete'' the
+bag with large external files such as the workflow outputs.}
+\label{fig:sep_concerns}
+\end{figure}
+%
+\subsection{Digital humanities: Cultural heritage}% {#culturalheritage}
+
+\label{sec:culturalheritage}
+
+The Pacific And Regional Archive for Digital Sources in Endangered
+Cultures (PARADISEC\footnote{\url{https://www.paradisec.org.au/}}) \cite{doi:10125/4567}
+maintains a repository of more than 500,000 files documenting
+endangered languages across more than 16,000 items, collected and
+digitized over many years by researchers interviewing and recording
+native speakers across the region.
+
+The Modern PARADISEC demonstrator\footnote{\url{https://mod.paradisec.org.au/}} has
+been
+proposed\footnote{\url{https://arkisto-platform.github.io/case-studies/paradisec/}}
+as an update to the 18 year old infrastructure, to also help long-term
+preservation of these artefacts in their digital form. The demonstrator
+uses RO-Crate to describe the overall structure and to capture the
+metadata of each item. The existing PARADISEC data collection has been
+ported and captured as RO-Crates. A Web portal then exposes the
+repository and its entries by indexing the RO-Crate metadata files,
+presenting a domain-specific view of the items -- the RO-Crate is
+``hidden'' and does not change the user interface.
+
+The PARADISEC use case takes advantage of several RO-Crate features and
+principles. Firstly, the transcribed metadata are now independent of
+the PARADISEC platform and can be archived, preserved and processed in
+its own right, using \surl{Schema.org} as base vocabulary and extended with
+PARADISEC-specific terms.
+
+In this approach, RO-Crate is the holder of itemised metadata, stored
+in regular files that are organised using Oxford Common File
+Layout\footnote{\url{https://ocfl.io/1.0/spec/}} (OCFL) \cite{ocfl_2020}, which ensures
+file integrity and versioning on a regular shared file system. This
+lightweight infrastructure also gives flexibility for future
+developments and maintenance. For example a consumer can use Linked
+Data software such as a graph database and query the whole corpora
+using SPARQL triple patterns across multiple RO-Crates. For long term
+digital preservation, beyond the lifetime of PARADISEC portals, a
+``last resort'' fallback is storing the generic RO-Crate HTML preview
+\cite{ro-crate-html-js}. Such human-readable rendering of RO-Crates can be
+hosted as static files by any Web server, in line with the approach
+taken by the Endings Project.\footnote{The Endings Project \url{https://endings.uvic.ca/} is a five-year
+project funded by the Social Sciences and Humanities Research Council
+(SSHRC) that is creating tools, principles, policies and
+recommendations for digital scholarship practitioners to create
+accessible, stable, long-lasting resources in the humanities.}
+
+\subsection{Machine-actionable Data Management Plans}% {#dmp}
+
+\label{sec:dmp}
+
+Machine-actionable Data Management Plans (maDMPs) have been proposed as
+an improvement to automate FAIR data management tasks in research
+\cite{doi:10.1371/journal.pcbi.1006750}; maDMPs use PIDs and controlled
+vocabularies to describe what happens to data over the research life
+cycle \cite{doi:10.1007/978-3-030-45442-5_15}. The Research Data Alliance's
+\textit{DMP Common Standard} for maDMPs \cite{doi:10.15497/rda00039} is one such
+formalisation for expressing maDMPs, which can be expressed as Linked
+Data using the DMP Common Standard Ontology
+\cite{doi:10.4126/frl01-006423289}, a specialisation of the W3C Data
+Catalog Vocabulary (DCAT) \cite{dcat2}. RDA maDMPs are usually expressed
+using regular JSON, conforming to the DMP JSON Schema.
+
+A mapping has been produced between Research Object Crates and
+Machine-actionable Data Management Plans
+\cite{doi:10.4126/frl01-006423291}, implemented by the RO-Crate RDA maDMP
+Mapper \cite{doi:10.5281/zenodo.3922136}. A similar mapping has been
+implemented by \texttt{RO-Crate\_2\_ma-DMP} \cite{doi:10.5281/zenodo.3903463}. In
+both cases, a maDMP can be converted to a RO-Crate, or vice versa. In
+\cite{doi:10.4126/frl01-006423291} this functionality caters for two use cases:
+\begin{enumerate}
+\item[1.] Start a skeleton data management plan based on an existing RO-Crate
+dataset, e.g. from an RO-Crate from WorkflowHub.
+\item[2.] Instantiate an RO-Crate based on a data management plan.
+\end{enumerate}
+
+An important nuance here is that data management plans are (ideally)
+written in \textit{advance} of data production, while RO-Crates are typically
+created to describe data \textit{after} it has been generated. What is
+significant to note in this approach is the importance of
+\textbf{templating} in order to make both tasks automatable and achievable,
+and how RO-Crate can fit into earlier stages of the research life cycle.
+\subsection{Institutional data repositories -- Harvard Data Commons}% {#institutionalrepos}
+
+\label{sec:institutionalrepos}
+
+The concept of a \textbf{Data Commons} for research collaboration was
+originally defined as \textit{``cyber-infrastructure that co-locates data,
+storage, and computing infrastructure with commonly used tools for
+analysing and sharing data to create an interoperable resource for the
+research community''} \cite{doi:10.1109/MCSE.2016.92}. More recently, Data
+Commons has been established to mean integration of active
+data-intensive research with data management and archival best
+practices, along with a supporting computational infrastructure.
+Furthermore, the Commons features tools and services, such as
+computation clusters and storage for scalability, data repositories for
+disseminating and preserving regular, but also large or sensitive
+datasets, and other research assets. Multiple initiatives were
+undertaken to create Data Commons on national, research, and
+institutional levels. For example, the Australian Research Data
+Commons (ARDC)\footnote{\url{https://ardc.edu.au}} \cite{doi:10.5334/dsj-2019-044} is a
+national initiative that enables local researchers and industries to
+access computing infrastructure, training, and curated datasets for
+data-intensive research. NCI's Genomic Data
+Commons\footnote{\url{https://gdc.cancer.gov/}} (GDC)
+\cite{doi:10.1182/blood-2017-03-735654} provides the cancer research
+community with access to a vast volume of genomic and clinical data.
+Initiatives such as Research Data Alliance (RDA) Global Open Research
+Commons\footnote{\url{https://www.rd-alliance.org/groups/global-open-research-commons-ig}}
+propose standards for the implementation of Data Commons to prevent
+them becoming ``data silos'' and thus, enable interoperability from one
+Data Commons to another.
+
+\textbf{Harvard Data Commons} \cite{doi:10.7557/5.5422} aims to address the
+challenges of data access and cross-disciplinary research within a
+research institution. It brings together multiple institutional
+schools, libraries, computing centres and the Harvard
+Dataverse\footnote{\url{https://dataverse.harvard.edu/}} data repository.
+Dataverse\footnote{\url{https://dataverse.org/}} \cite{doi:10.1045/january2011-crosas}
+is a free and open-source software platform to archive, share and cite
+research data. The Harvard Dataverse repository is the largest of 70
+Dataverse installations worldwide, containing over 120K datasets with
+about 1.3M data files (as of 2021-11-16). Working toward the goal of
+facilitating collaboration and data discoverability and management
+within the university, Harvard Data Commons has the following primary
+objectives:\looseness=1
+\begin{enumerate}
+\item[1.] the integration of Harvard Research Computing with Harvard Dataverse
+by leveraging Globus endpoints \cite{doi:10.1109/MCC.2014.52}; this will
+allow an automatic transfer of large datasets to the repository. In
+some cases, only the metadata will be transferred while the data stays
+stored in remote storage;
+\item[2.] support for advanced research workflows and providing packaging
+options for assets such as code and workflows in the Harvard Dataverse
+repository to enable reproducibility and reuse, and
+\item[3.] interation of repositories supported by Harvard, which include
+DASH,\footnote{\url{https://dash.harvard.edu/}} the open access institutional
+repository, the Digital Repository Services (DRS) for preserving
+digital asset collections, and the Harvard Dataverse.
+\end{enumerate}
+
+Particularly relevant to this article is the second objective of the
+Harvard Data Commons, which aims to support the deposit of research
+artefacts to Harvard Dataverse with sufficient information in the
+metadata to allow their future reuse (Fig.~\ref{fig:hdc}). To support
+the incorporation of data, code, and other artefacts from various
+institutional infrastructures, Harvard Data Commons is currently
+working on RO-Crate adaptation. The RO-Crate metadata provides the
+necessary structure to make all research artefacts FAIR. The Dataverse
+software already has extensive support for metadata, including the Data
+Documentation Initiative (DDI), Dublin Core, DataCite, and \surl{Schema.org}.
+Incorporating RO-Crate, which has the flexibility to describe a wide
+range of research resources, will facilitate their seamless transition
+from one infrastructure to the other within the Harvard Data Commons.
+
+Even though the Harvard Data Commons is specific to Harvard University,
+the overall vision and the three objectives can be abstracted and
+applied to other universities or research organisations. The Commons
+will be designed and implemented using standards and commonly-used
+approaches to make it interoperable and reusable by others.
+
+\begin{figure}%[t!]
+\includegraphics{210053f04}
+\caption{One aspect of Harvard Data Commons. Automatic
+encapsulation and deposit of artefacts from data management tools used
+during active research at the Harvard Dataverse repository.}
+\label{fig:hdc}
+\end{figure}
+
+ \section{Related work}% {#relatedwork}
+
+\label{sec:relatedwork}
+
+With the increasing digitisation of research processes, there has been
+a significant call for the wider adoption of interoperable sharing of
+data and its associated metadata. We refer to
+\cite{doi:10.1016/j.patter.2020.100136} for a comprehensive overview and
+recommendations, in particular for data; notably that review highlights
+the wide variety of metadata and documentation that the literature
+prescribes for enabling data reuse. Likewise, we suggest
+\cite{doi:10.1016/j.patter.2021.100322} that covers the importance of
+metadata standards in reproducible computational research.
+
+Here we focus on approaches for bundling research artefacts along with
+their metadata. This notion of publishing compound objects for
+scholarly communication has a long history behind it
+\cite{doi:10.1190/1.1822162,vandesompel_2007}, but recent approaches
+have followed three main strands: (1) publishing to centralised
+repositories; (2) packaging approaches similar to RO-Crate; and (3)
+bundling the computational workflow around a scientific experiment.
+
+ \subsection{Bundling and packaging digital research artefacts}
+
+Early work making the case for publishing compound scholarly
+communication units \cite{vandesompel_2007} led to the development of the
+Object Re-Use and Exchange
+model\footnote{\url{http://www.openarchives.org/ore/1.0/primer}} (OAI-ORE), providing
+a structured \textbf{resource map} of the digital artefacts that together
+support a scholarly output.
+
+The challenge of describing computational workflows was one of the main
+motivations for the early proposal of \textit{Research Objects} (RO)
+\cite{doi:10.1016/j.future.2011.08.004} as first-class citizens for sharing
+and publishing. The RO approach involves bundling datasets, workflows,
+scripts and results along with traditional dissemination materials like
+journal articles and presentations, forming a single package.
+Crucially, these resources are not just gathered, but also individually
+typed, described and related to each other using semantic vocabularies.
+As pointed out in \cite{doi:10.1016/j.future.2011.08.004} an open-ended
+\textit{Linked Data} approach is not sufficient for scholarly communication: a
+common data model is also needed in addition to common and best
+practices for managing and annotating lifecycle, ownership, versioning
+and attributions.
+
+Considering the FAIR principles \cite{doi:10.1038/sdata.2016.18}, we can
+say with hindsight that the initial RO approaches strongly targeted
+\textit{Interoperability}, with a particular focus on the reproducibility of
+\textit{in-silico experiments} involving computational workflows and the reuse
+of existing RDF vocabularies.
+
+The first implementation of Research Objects for sharing workflows in
+myExperiment \cite{doi:10.1093/nar/gkq429} was based on RDF ontologies
+\cite{newman2009}, building on Dublin Core, FOAF, SIOC, Creative Commons
+and OAI-ORE to form myExperiment ontologies for describing social
+networking, attribution and credit, annotations, aggregation packs,
+experiments, view statistics, contributions, and workflow components
+\cite{myExperimentOntology2009}.
+
+This initially workflow-centric approach was further formalized as the
+Wf4Ever Research Object Model \cite{doi:10.1016/j.websem.2015.01.003},
+which is a general-purpose research artefact description framework.
+This model is based on existing ontologies (FOAF, Dublin Core Terms,
+OAI-ORE and AO/OAC precursors to the W3C Web Annotation Model
+\cite{Ciccarese:17:WAD}) and adds specializations for workflow models and
+executions using W3C PROV-O \cite{PROVO}. The Research Object statements
+are saved in a \textit{manifest} (the OAI-ORE \textit{resource map}), with additional
+annotation resources containing user-provided details such as title and
+description.
+
+We now claim that one barrier for wider adoption of the Wf4Eer Research
+Object model for general packaging digital research artefacts was
+exactly this re-use of multiple existing vocabularies (FAIR principle
+I2: \textit{Metadata use vocabularies that follow FAIR principles}), which in
+itself is recogized as a challenge \cite{doi:10.3233/978-1-61499-660-6-9}.
+Adapters of the Wf4Ever RO model would have to navigate documentation
+of multiple overlapping ontologies, in addition to facing the usual
+Semantic Web development choices for RDF serialization formats,
+identifier minting and publishing resources on the Web.
+
+Several developments for Research Objects improved on this situation,
+such as ROHub used by Earth Sciences
+\cite{doi:10.1016/j.future.2019.03.046}, which provides a interactive
+user-interface for making Research Objects, along with Research Object
+Bundle \cite{doi:10.5281/zenodo.12586} (RO Bundle), which is a ZIP-archive
+embedding data files and a JSON-LD serialization of the manifest with
+mappings for a limited set of terms. RO Bundle was also used for
+storing detailed workflow run provenance (TavernaPROV
+\cite{doi:10.5281/zenodo.51314}).
+
+RO-Bundle evolved to Research Object BagIt
+archives,\footnote{\url{https://w3id.org/ro/bagit}} a variant of RO Bundle as a BagIt
+archive \cite{doi:10.17487/rfc8493}, used by Big Data Bags
+\cite{doi:10.1109/BigData.2016.7840618}, CWLProv
+\cite{doi:10.1093/gigascience/giz095} and WholeTale
+\cite{doi:10.3233/APC200107,doi:10.1109/eScience.2019.00068}.
+
+\subsection{FAIR Digital Objects}
+
+FAIR Digital Objects (FDO) \cite{doi:10.3390/publications8020021} have been
+proposed as a conceptual framework for making digital resources
+available in a Digital Objects (DO) architecture which encourages
+active use of the objects and their metadata. In particular, an FDO has
+five parts: (i) The FDO \textit{content}, bit sequences stored in an
+accessible repository; (ii) a \textit{Persistent Identifier} (PID) such as a
+DOI that identifies the FDO and can resolve these same parts; (iii)
+Associated rich \textit{metadata}, as separate FDOs; (iv) Type definitions,
+also separate FDOs; (v) Associated \textit{operations} for the given types. A
+Digital Object typed as a Collection aggregates other DOs by reference.
+
+The Digital Object Interface Protocol \cite{doip2.0} can be considered an
+``abstract protocol'' of requirements, DOs could be implemented in
+multiple ways. One suggested implementation is the FAIR Digital Object
+Framework,\footnote{\url{https://fairdigitalobjectframework.org/}} based on HTTP and
+the Linked Data Principles. While there is agreement on using PIDs
+based on DOIs, consensus on how to represent common metadata, core
+types and collections as FDOs has not yet been reached. We argue that
+RO-Crate can play an important role for FDOs:
+\begin{enumerate}
+\item[1.] By providing a predictable and extensible serialisation of
+structured metadata.
+\item[2.] By formalising how to aggregate digital objects as collections (and
+adding their context).
+\item[3.] By providing a natural Metadata FDO in the form of the RO-Crate
+Metadata File.
+\item[4.] By being based on Linked Data and \surl{schema.org} vocabulary, meaning
+that PIDs already exist for common types and properties.
+\end{enumerate}
+
+At the same time, it is clear that the goal of FDO is broader than that
+of RO-Crate; namely, FDOs are active objects with distributed
+operations, and add further constraints such as PIDs for every element.
+These features improve FAIR features of digital objects and are also
+useful for RO-Crate, but they also severely restrict the infrastructure
+that needs to be implemented and maintained in order for FDOs to remain
+accessible. RO-Crate, on the other hand, is more flexible: it can
+minimally be used within any file system structure, or ideally exposed
+through a range of Web-based scenarios. A \textit{FAIR profile of RO-Crate}
+(e.g. enforcing PID usage) will fit well within a FAIR Digital Object ecosystem.
+
+ \subsection{Packaging workflows}
+
+The use of computational workflows, typically combining a chain of
+tools in an analytical pipeline, has gained prominence in particular in
+the life sciences. Workflows might be used primarily to improve
+computational scalability, as well as to also assist in making computed
+data results FAIR \cite{doi:10.1162/dint_a_00033}, for instance by
+improving reproducibility \cite{doi:10.1016/j.future.2017.01.012}, but also
+because programmatic data usage help propagate their metadata and
+provenance \cite{doi:10.1002/cpe.1228}. At the same time, workflows raise
+additional FAIR challenges, since they can be considered important
+research artefacts themselves. This viewpoint poses the problem of
+capturing and explaining the computational methods of a pipeline in
+sufficient machine-readable detail \cite{doi:10.3233/DS-190026}.
+
+Even when researchers follow current best practices for workflow
+reproducibility \cite{doi:10.1016/j.cels.2018.03.014,doi:10.1016/j.future.2017.01.012}, the communication of computational
+outcomes through traditional academic publishing routes effectively
+adds barriers as authors are forced to rely on a textual manuscript
+representations. This hinder reproducibility and FAIR use of the
+knowledge previously captured in the workflow.
+
+As a real-life example, let us look at a metagenomics article
+\cite{doi:10.1038/s41586-019-0965-1} that describes a computational
+pipeline. Here the authors have gone to extraordinary efforts to
+document the individual tools that have been reused, including their
+citations, versions, settings, parameters and combinations. The
+\textit{Methods} section is two pages in tight double-columns with twenty four
+additional references, supported by the availability of data on an FTP
+server (60 GB) \cite{ebi_ftp_umgs2019} and of open source code in GitHub
+Finn-Lab/MGS-gut\footnote{\url{https://github.com/Finn-Lab/MGS-gut}}
+\cite{finn-lab-mgsgut}, including the pipeline as shell scripts and
+associated analysis scripts in R and Python.
+
+This attention to reporting detail for computational workflows is
+unfortunately not yet the norm, and although bioinformatics journals
+have strong \textit{data availability} requirements, they frequently do not
+require authors to include or cite \textit{software, scripts and pipelines}
+used for analysing and producing results \cite{soilandreyes_tweet_2020}.
+Indeed, in the absence of a specific requirement and an editorial
+policy to back it up -- such as eliminating the reference limit --
+authors are effectively discouraged from properly and comprehensively
+citing software \cite{doi:10.1038/s41592-019-0350-x}.
+
+However detailed this additional information might be, another
+researcher who wants to reuse a particular computational method may
+first want to assess if the described tool or workflow is Re-runnable
+(executable at all), Repeatable (same results for original inputs on
+same platform), Reproducible (same results for original inputs with
+different platform or newer tools) and ultimately Reusable (similar
+results for different input data), Repurposable (reusing parts of the
+method for making a new method) or Replicable (rewriting the workflow
+following the method description)
+\cite{doi:10.3389/fninf.2017.00069,goble_presentation_2016}.
+
+Following the textual description alone, researchers would be forced to
+jump straight to evaluate ``Replicable'' by rewriting the pipeline from
+scratch. This can be expensive and error-prone. They would firstly need
+to install all the software dependencies and download reference
+datasets. This can be a daunting task, which may have to be repeated
+multiple times as workflows typically are developed at small scale on
+desktop computers, scaled up to local clusters, and potentially put
+into production using cloud instances, each of which will have
+different requirements for software installations.
+
+In recent years the situation has been greatly improved by software
+packaging and container technologies like Docker and Conda, these
+technologies have been increasingly adopted in life sciences
+\cite{doi:10.1007/s41019-017-0050-4} thanks to collaborative efforts such
+as BioConda \cite{doi:10.1038/s41592-018-0046-7} and BioContainers
+\cite{doi:10.1093/bioinformatics/btx192}, and support by Linux
+distributions (e.g. Debian Med \cite{doi:10.1186/1471-2105-11-S12-S5}). As
+of November 2021, more than 9,000 software packages are available in
+BioConda alone,\footnote{\url{https://anaconda.org/bioconda/}} and 10,000 containers
+in BioContainers.\footnote{\url{https://biocontainers.pro/\#/registry}}
+
+Docker and Conda have been integrated into workflow systems such as
+Snakemake \cite{doi:10.1093/bioinformatics/bts480}, Galaxy
+\cite{doi:10.1093/nar/gky379} and Nextflow \cite{doi:10.1038/nbt.3820}, meaning
+a downloaded workflow definition can now be executed on a ``blank''
+machine (except for the workflow engine) with the underlying analytical
+tools installed on demand. Even with using containers there is a
+reproducibility challenge, for instance Docker Hub's retention policy
+will expire container images after six
+months,\footnote{\url{https://www.docker.com/blog/docker-hub-image-retention-policy-delayed-and-subscription-updates/}}
+or a lack of recording versions of transitive dependencies of Conda
+packages could cause incompatibilities if the packages are subsequently updated.
+
+These container and package systems only capture small amounts of
+metadata.\footnote{Docker and Conda can use \textit{build recipes}, a set of commands
+that construct the container image through downloading and installing
+its requirements. However these recipes are effectively another piece
+of software code, which may itself decay and become difficult to rerun.} In particular, they do not capture any of the semantic
+relationships between their content. Understanding these relationships
+is made harder by the opaque wrapping of arbitrary tools with unclear
+functionality, licenses and attributions.
+
+From this we see that computational workflows are themselves complex
+digital objects that need to be recorded not just as files, but in the
+context of their execution environment, dependencies and analytical
+purpose in research -- as well as other metadata (e.g. version,
+license, attribution and identifiers).
+
+It is important to note that having all these computational details in
+order to represent them in an RO-Crate is an ideal scenario -- in
+practice there will always be gaps of knowledge, and exposing all
+provenance details automatically would require improvements to the data
+sources, workflow, workflow engine and its dependencies. RO-Crate can
+be seen as a flexible annotation mechanism for augmenting automatic
+workflow provenance. Additional metadata can be added manually, e.g.
+for sensitive clinical data that cannot be publicly exposed\footnote{FAIR principle A2: \textit{M etadata are accessible, even when the
+data are no longer available} \cite{doi:10.1038/sdata.2016.18}.}, or
+to cite software that lack persistent identifiers. This inline
+\textit{FAIRifying} allows researchers to achieve ``just enough FAIR'' to
+explain their computational experiments.
+
+\section{Conclusion}
+
+\label{sec:conclusion}
+
+RO-Crate has been established as an approach to packaging digital
+research artefacts with structured metadata. This approach assists
+developers and researchers to produce and consume FAIR archives of
+their research.
+
+RO-Crate is formed by a set of best practice recommendations, developed
+by an open and broad community. These guidelines show how to use ``just
+enough'' Linked Data standards in a consistent way. The use of
+structured metadata with a rich base vocabulary can cover
+general-purpose contextual relations, with a Linked Data foundation
+that ensures extensibility to domain- and application-specific uses. We
+can therefore consider an RO-Crate not just as a structured data
+archive, but as a multimodal scholarly knowledge graph that can help
+``FAIRify'' and combine metadata of existing resources.
+
+The adoption of simple Web technologies in the RO-Crate specification
+has helped a rapid development of a wide variety of supporting open
+source tools and libraries. RO-Crate fits into the larger landscape of
+open scholarly communication and FAIR Digital Object infrastructure,
+and can be integrated into data repository platforms. RO-Crate can be
+applied as a data/metadata exchange mechanism, assist in long-term
+archival preservation of metadata and data, or simply used at a small
+scale by individual researchers. Thanks to its strong community
+support, new and improved profiles and tools are being continuously
+added to the RO-Crate tooling landscape, making it easier for adopters
+to find examples and support for their own use case.
+
+\subsection{Strictness vs flexibility}
+
+There is always a tradeoff between flexibility and strictness
+\cite{doi:10.1007/s11042-009-0397-2} when deciding on semantics of metadata
+models. Strict requirements make it easier for users and code to
+consume and populate a model, by reducing choices and having mandated
+``slots'' to fill in. But such rigidity can also restrict richness and
+applicability of the model, as it in turn enforce the initial
+assumptions about what can be described.
+
+RO-Crate attempts to strike a balance between these tensions, and
+provides a common metadata framework that encourages extensions.
+However, just like the RO-Crate specification can be thought of as a
+\textit{core profile} of \surl{schema.org} in JSON-LD, we cannot stress the
+importance of also establishing domain-specific RO-Crate profiles and
+conventions, as explored in Sections~\ref{sec:profiles} and \ref
+{sec:inuse}. Specialization comes hand-in-hand with the principle of
+\textit{graceful degradation}; RO-Crate applications and users are free to
+choose the semantic detail level they participate at, as long as they
+follow the common syntactic requirements.
+
+ \section{Future work}% {#futurework}
+
+\label{sec:futurework}
+
+The direction of future RO-Crate work is determined by the community
+around it as a collaborative effort. We currently plan on further
+outreach, building training material (including a comprehensive
+entry-level tutorial) and maturing the reference implementation
+libraries. We will also collect and build examples of RO-Crate
+\textit{consumption}, e.g. Jupyter Notebooks that query multiple crates using
+knowledge graphs. In addition, we are exploring ways to support some
+entity types requested by users, e.g. detailed workflow runs or
+container provenance, which do not have a good match in \surl{Schema.org}.
+Such support could be added, for instance, by integrating other
+vocabularies or by having separated (but linked) metadata files.
+
+Furthermore, we want to better understand how the community uses
+RO-Crate in practice and how it contrasts with other related efforts;
+this will help us to improve our specification and tools. By
+discovering commonalities in emerging usage (e.g. additional \surl{schema.org}
+types), the community helps to reduce divergence that could otherwise
+occur with proliferation of further RO-Crate profiles. We plan to
+gather feedback via user studies, with the Linked Open Data community
+or as part of EOSC Bring-your-own-Data training events.
+
+We operate in an open community where future and potential users of
+RO-Crate are actively welcomed to participate and contribute feedback
+and requirements. In addition we are targeting a wider audience through
+extensive outreach
+activities\footnote{\url{https://www.researchobject.org/ro-crate/outreach.html}} and
+by initiating new connections. Recent contacts include American
+Geophysical Union (AGU) on Data Citation Reliquary
+\cite{doi:10.5281/zenodo.4916734}, National Institute of Standards and
+Technology (NIST) on material science and
+InvenioRDM\footnote{\url{https://inveniosoftware.org/products/rdm/}} used by the
+Zenodo data repository. New Horizon Europe projects adapting RO-Crate
+include BY-COVID,\footnote{\url{https://by-covid.org/}} which aims to improve FAIR
+access to data on COVID-19 and other infectious diseases.
+
+The main addition in the upcoming 1.2 release of the RO-Crate
+specifications will be the formalization of
+profiles\footnote{\url{https://www.researchobject.org/ro-crate/1.2-DRAFT/profiles}}
+for different categories of crates. Additional entity types have been
+requested by users, e.g. workflow runs, business workflows, containers
+and software packages, tabular data structures; these are not always
+matched well with existing \surl{schema.org} types but may benefit from other
+vocabularies or even separate metadata files, e.g. from Frictionless
+Data.\footnote{\url{https://frictionlessdata.io/}} We will be further aligning with
+and collaborating with related research artefact description efforts
+like CodeMeta\footnote{\url{https://codemeta.github.io/}} for software metadata,
+Science-on-schema.org\footnote{\url{https://science-on-schema.org/}}
+\cite{doi:10.5281/zenodo.4477164} for datasets, FAIR Digital
+Objects\footnote{\url{https://fairdo.org/}} \cite{doi:10.3390/publications8020021} and
+activities in EOSC task forces\footnote{\url{https://www.eosc.eu/task-force-faq}}
+including the EOSC Interoperability Framework \cite{doi:10.2777/620649}.
+
+% CrEDiT taxonomy:
+%
+% Conceptualization
+% Data curation
+% Formal Analysis
+% Funding acquisition
+% Investigation
+% Methodology
+% Project administration
+% Resources
+% Software
+% Supervision
+% Validation
+% Visualization
+% Writing - original draft
+% Writing - review \& editing
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%% Supplementary data
+%\begin{esm}%[title=]
+%\esmtitle{Supplementary data}
+%\esmdescription{} is available at: \url{http://dx.doi.org/???}.
+%\esmfiletype{pdf} %file type list:
+%%http://www.freeformatter.com/mime-types-list.html#mime-types-list
+%\esmfilename{MANUSCRIPTsupp.pdf}
+%\end{esm}
+
+%% Put support info here:
+%\begin{ack}
+%\end{ack}
+\begin{acks}
+This work has received funding from the European Commission's Horizon
+2020 research and innovation programme for projects
+BioExcel-2\footnote{\url{https://cordis.europa.eu/project/id/823830}}
+(H2020-INFRAEDI-2018-1 823830), IBISBA
+1.0\footnote{\url{https://cordis.europa.eu/project/id/730976}}
+(H2020-INFRAIA-2017-1-two-stage 730976),
+PREP-IBISBA\footnote{\url{https://cordis.europa.eu/project/id/871118}}
+(H2020-INFRADEV-2019-2 871118),
+EOSC-Life\footnote{\url{https://cordis.europa.eu/project/id/824087}}
+(H2020-INFRAEOSC-2018-2 824087),
+SyntheSys+\footnote{\url{https://cordis.europa.eu/project/id/823827}}
+(H2020-INFRAIA-2018-1 823827). From the Horizon Europe Framework
+Programme this work has received funding for
+BY-COVID\footnote{\url{https://cordis.europa.eu/project/id/101046203}}
+(HORIZON-INFRA-2021-EMERGENCY-01 101046203).
+
+Bj\"{o}rn Gr\"{u}ning is supported by DataPLANT (NFDI 7/1 --
+42077441\footnote{\url{https://gepris.dfg.de/gepris/projekt/442077441}}), part of the
+German National Research Data Infrastructure (NFDI), funded by the
+Deutsche Forschungsgemeinschaft (DFG).
+
+Ana Trisovic is funded by the Alfred P. Sloan Foundation (grant number
+P-2020-13988).\footnote{\url{https://sloan.org/grant-detail/9555}} Harvard Data
+Commons is supported by an award from Harvard University Information
+Technology (HUIT).
+\end{acks}
+
+ \section*{Contributions}
+
+Author contributions to this article and the RO-Crate projet according
+to the Contributor Roles Taxonomy CASRAI
+CrEDiT\footnote{\url{https://casrai.org/credit/}} \cite{doi:10.1087/20150211}:
+\begin{description}
+\item[Stian Soiland-Reyes]  Conceptualization, Data curation, Formal Analysis, Funding
+acquisition, Investigation, Methodology, Project administration,
+Software, Visualization, Writing -- original draft, Writing -- review
+\& editing
+
+\item[Peter Sefton]  Conceptualization, Investigation, Methodology, Project
+administration, Resources, Software, Writing -- review \& editing
+
+\item[Merc\`{e} Crosas]  Writing -- review \& editing
+
+\item[Leyla Jael Castro]  Methodology, Writing -- review \& editing
+
+\item[Frederik Coppens]  Writing -- review \& editing
+
+\item[Jos\'{e} M. Fern\'{a}ndez]  Methodology, Software, Writing -- review \& editing
+
+\item[Daniel Garijo]  Methodology, Writing -- review \& editing
+
+\item[Bj\"{o}rn Gr\"{u}ning]  Writing -- review \& editing
+
+\item[Marco La Rosa]  Software, Methodology, Writing -- review \& editing
+
+\item[Simone Leo]  Software, Methodology, Writing -- review \& editing
+
+\item[Eoghan \'{O} Carrag\'{a}in]  Investigation, Methodology, Project administration, Writing -- review
+\& editing
+
+\item[Marc Portier]  Methodology, Writing -- review \& editing
+
+\item[Ana Trisovic]  Software, Writing -- review \& editing
+
+\item[RO-Crate Community]  Investigation, Software, Validation, Writing -- review \& editing
+
+\item[Paul Groth]  Methodology, Supervision, Writing -- original draft, Writing --
+review \& editing
+
+\item[Carole Goble]  Conceptualization, Funding acquisition, Methodology, Project
+administration, Supervision, Visualization, Writing -- review \& editing
+\end{description}
+
+We would also like to acknowledge contributions from:
+\begin{description}
+\item[Finn Bacall]
+ Software, Methodology
+
+\item[Herbert Van de Sompel]
+ Writing -- review \& editing
+
+\item[Ignacio Eguinoa]
+ Software, Methodology
+
+\item[Nick Juty]
+ Writing -- review \& editing
+
+\item[Oscar Corcho]
+ Writing -- review \& editing
+
+\item[Stuart Owen]
+ Writing -- review \& editing
+
+\item[Laura Rodr\'iguez-Navas]
+ Software, Visualization, Writing -- review \& editing
+
+\item[Alan R. Williams]
+ Writing -- review \& editing
+\end{description}
+\begin{appendix}
+
+\section{Formalizing RO-Crate in First Order Logic}
+\label{sec:formaldefinition}
+
+Below is a formalization of the concept of RO-Crate as a set of
+relations using First Order Logic.
+
+\subsection{Language}
+
+Definition of language $\mathcal{L}_{\mathit{rocrate}}$:
+%
+\begin{eqnarray*}
+\mathcal{L}_{\mathit{rocrate}} &=& \bigl\{\mathit{Property}(p), \mathit{Class}(c),
+\mathit{Value}(x), \mathbb{R}, \mathbb{S} \bigr\}
+\\
+\mathbb{D} &=& \mathbb{IRI}
+\\
+\mathbb{IRI} &\equiv& { \text{IRIs as defined in RFC3987} }
+\\
+\mathbb{R} &\equiv& { \text{real or integer numbers} }
+\\
+\mathbb{S} &\equiv& { \text{literal strings} }
+\end{eqnarray*}
+
+The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$
+identifiers \cite{doi:10.17487/rfc3987} (notation \texttt
+{<http://example.\noxml{\break}com/>}),\footnote{For simplicity, blank nodes are not included in this formalisation, as RO-Crate
+recommends the use of IRI identifiers: \url
+{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#describing-entities-in-json-ld}.} with additional descriptions using numbers $\mathbb{R}$ (notation
+$13.37$) and literal strings $\mathbb{S}$ (notation $\text{``Hello''}$).
+
+From this formalised language $\mathcal{L}_{\mathit{rocrate}}$ we can interpret
+an RO-Crate in any representation that can gather these descriptions,
+their properties, classes, and literal attributes.
+
+\subsection{Minimal RO-Crate}
+
+Below we use $\mathcal{L}_{\mathit{rocrate}}$ to define a minimal\footnote{The full
+list of types, relations and attribute properties from the
+RO-Crate specification are not included. Examples shown include
+$\textit{datePublished}$, $\textit{CreativeWork}$ and $\textit{name}$.
+} RO-Crate:
+%
+\begin{eqnarray*}
+\textit{ROCrate}(R) & \models& \textit{Root}(R) \land \textit{Mentions}(R, R)
+\land \textit{hasPart}(R, d) \land
+\\
+& & {}\textit{Mentions}(R, d) \land \textit{DataEntity}(d) \land
+\\
+& & {}\textit{Mentions}(R, c) \land \textit{ContextualEntity}(c)
+\\
+\forall r \ \textit{Root}(r) & \Rightarrow& \textit{Dataset}(r) \land
+\textit{name}(r, n) \land \textit{description}(r, d) \land
+\\
+& &{} \textit{datePublished}(r, date) \land \textit{license}(e, l)
+\\
+\forall e \forall n \ \textit{name}(e, n) & \Rightarrow& \textit{Value}(n)
+\\
+\forall e \forall s \ \textit{description}(e, s) & \Rightarrow&
+\textit{Value}(s)
+\\
+\forall e \forall d \ \textit{datePublished}(e, d) & \Rightarrow&
+\textit{Value}(d)
+\\
+\forall e \forall l \ \textit{license}(e, l) & \Rightarrow&
+\textit{ContextualEntity}(l)
+\\
+\textit{DataEntity}(e) & \equiv& \textit{File}(e) \oplus \textit{Dataset}(e)
+\\
+\textit{Entity}(e) & \equiv& \textit{DataEntity}(e) \lor \textit{ContextualEntity}(e)
+\\
+\forall e \ \textit{Entity}(e) & \Rightarrow& \textit{type}(e, c) \land
+\textit{Class}(c)
+\\
+\forall e \ \textit{ContextualEntity}(e) & \Rightarrow& \textit{name}(e, n)
+\\
+\textit{Mentions}(R, s) & \models& \textit{Relation}(s, p, e) \oplus
+\textit{Attribute}(s, p, l)
+\\
+\textit{Relation}(s, p, o) & \models& \textit{Entity}(s) \land
+\textit{Property}(p) \land \textit{Entity}(o)
+\\
+\textit{Attribute}(s, p, v) & \models& \textit{Entity}(s) \land
+\textit{Property}(p) \land \textit{Value}(v)
+\\
+\textit{Value}(v) & \equiv& v \in\mathbb{R} \oplus v \in\mathbb{S}
+\end{eqnarray*}
+
+An $\textit{ROCrate}(R)$ is defined as a self-described \emph{Root Data Entity},
+which describes and contains parts (\emph{data entities}), which are
+further described in \emph{contextual entities}. These terms align with
+their use in the RO-Crate 1.1 terminology.\footnote{\url{https://www.researchobject.org/ro-crate/1.1/terminology}}
+
+The $\textit{Root}(r)$ is a type of $\textit{Dataset}(r)$, and must have the metadata to
+literal attributes to provide a $\textit{name}$, $description$ and
+$\textit{datePublished}$, as well as a contextual entity identifying its
+license. These predicates correspond to the RO-Crate 1.1 requirements
+for the root data entity.\footnote{\url
+{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}}
+
+The concept of an $\textit{Entity}(e)$ is introduced as being either a
+$\textit{DataEntity}(e)$, a $\textit{ContextualEntity}(e)$, or both.\footnote{\url
+{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html\#contextual-vs-data-entities}} Any $\textit{Entity}(e)$ must be
+typed with at least one $\textit{Class}(c)$, and
+every $\textit{ContextualEntity}(e)$ must also have a $\textit{name}(e,n)$; this
+corresponds to expectations for any \emph{referenced contextual entity}
+(Section~\ref{sec:contextualentities}).
+
+For simplicity in this formalization (and to assist production rules
+below) $R$ is a constant representing a single RO-Crate, typically
+written to independent RO-Crate Metadata files. $R$ is used by
+$\textit{Mentions}(R, e)$ to indicate that $e$ is an Entity described by the
+RO-Crate and therefore its metadata (a set of $\textit{Relation}$ and
+$\textit{Attribute}$ predicates) form part of the RO-Crate serialization.
+$\textit{Relation}(s, p, o)$ and $\textit{Attribute}(s, p, x)$ are defined as a \emph
+{subject--predicate--object} triple pattern from an $\textit{Entity}(s)$ using
+a $\textit{Property}(p)$ to either another $\textit{Entity}(o)$ or a $\textit{Value}(x)$ value.
+
+\subsection{Example of formalized RO-Crate}
+
+The below is an example RO-Crate represented using the above
+formalisation, assuming a base URI of \texttt{<http://example.com/ro/123/>}:
+%
+{\fontsize{9.7}{11.7}\selectfont{\begin{gather*}
+\textit{ROCrate}(\texttt{<http://example.com/ro/123/>})
+\\
+\textit{name}\bigl(\texttt{<http://example.com/ro/123/>},
+\\
+\quad \text{``Data files associated with the manuscript:Effects of {
+\ldots}''}\bigr)
+\\
+\textit{description}\bigl(\texttt{<http://example.com/ro/123/>},
+\\
+\quad \text{``Palliative care planning for nursing home residents {
+\ldots}''}\bigr)
+\\
+\textit{license}(\texttt{<http://example.com/ro/123/>},
+\\
+\quad \texttt{<https://spdx.org/licenses/CC-BY-4.0>})
+\\
+\textit{datePublished}\bigl(\texttt{<http://example.com/ro/123/>}, \text
+{``2017-02-23''}\bigr)
+\\
+\textit{hasPart}(\texttt{<http://example.com/ro/123/>}, \texttt{<http://example.com/ro/123/file.txt>})
+\\
+\textit{hasPart}(\texttt{<http://example.com/ro/123/>}, \texttt{<http://example.com/ro/123/interviews/>})
+\\
+\\
+\textit{ContextualEntity}(\texttt{<https://spdx.org/licenses/CC-BY-4.0>})
+\\
+\textit{name}\bigl(\texttt{<https://spdx.org/licenses/CC-BY-4.0>},
+\\
+\quad \text{Creative Commons Attribution 4.0''}\bigr)
+\\
+\\
+\textit{ContextualEntity}(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>})
+\\
+\textit{name}\bigl(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>},
+\\
+\quad \text{Creative Commons Attribution Non Commercial 4.0''}
+\bigr)
+\\
+\\
+\textit{File}(\texttt{<http://example.com/ro/123/survey.csv>})
+\\
+\textit{name}\bigl(\texttt{<http://example.com/ro/123/survey.csv>},
+\text{``Survey of care providers''}\bigr)
+\\
+\\
+\textit{Dataset}(\texttt{<http://example.com/ro/123/interviews/>})
+\\
+\textit{name}\bigl(\texttt{<http://example.com/ro/123/interviews/>},
+\\
+\quad \text{``Audio recordings of care provider interviews''}
+\bigr)
+\\
+\textit{license}(\texttt{<http://example.com/ro/123/interviews/>},
+\\*
+\quad \texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>})
+\end{gather*}}}
+
+Notable from this triple-like formalization is that a RO-Crate $R$ is
+fully represented as a tree at depth 2 helped by the use of
+%PDAG%
+$\mathbh{lri}$
+%PDAG%
+nodes. For instance the aggregation from the root entity
+$\textit{hasPart}(\texttt{{\ldots}interviews/>})$ is at same level as the data
+entity's property \textit{license}(\texttt{{\ldots}CC-BY-\noxml{\break}NC-4.0>}) and that
+contextual entity's attribute $ \textit{name}(\text{{\ldots}Non Commercial
+4.0''})$. As shown in Section~\ref{sec:jsonld}, the RO-Crate Metadata
+File serialization is an equivalent shallow tree, although at depth 3
+to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.
+
+In reality many additional attributes and contextual types from
+\surl{schema.org} types like \url{http://schema.org/affiliation} and \url
+{http://schema.org/Organization} would be used to further describe the
+RO-Crate and its entities, but as these are optional (\textit{SHOULD}
+requirements) they do not form part of this formalization.
+
+\subsection{Mapping to RDF with schema.org}
+
+A formalized RO-Crate in $\mathcal{L}_{\mathit{rocrate}}$ can be mapped to
+different serializations.
+Assume a simplified\footnote{This simplification does not cover the extensive list of literal
+datatypes built-in to RDF 1.1, only strings and decimal real numbers.
+Likewise, language of literals are not included.} language $\mathcal{L}_{\textit{RDF}}$
+based on the RDF abstract syntax \cite{rdfworkinggroup_2014}:
+%
+{\fontsize{9.8}{11.8}\selectfont{\begin{eqnarray*}
+\mathcal{L}_{\textit{RDF}} & \equiv& \bigl\{ \textit{Triple}(s,p,o),
+\textit{IRI}(i), \textit{BlankNode}(b), \textit{Literal}(s), \mathbb{IRI},
+\mathbb{S}, \mathbb{R} \bigr\}
+\\
+\mathbb{D}_{\textit{RDF}} & \equiv& \mathbb{S}
+\\
+\forall i \ \textit{IRI}(i) & \Rightarrow& i \in\mathbb{IRI}
+\\
+\forall s \forall p \forall o \ \textit{Triple}(s,p,o) & \Rightarrow& \bigl(
+\textit{IRI}(s) \lor \textit{BlankNode}(s) \bigr) \land
+\\
+& & \textit{IRI}(p) \land
+\\
+& & \bigl(\textit{IRI}(o) \lor \textit{BlankNode}(o) \lor \textit{Literal}(o)
+\bigr)
+\\
+\textit{Literal}(v) & \models& \textit{Value}(v) \land \textit{Datatype}(v,t)
+\land \textit{IRI}(t)
+\\
+\forall v \ \textit{Value}(v) & \Rightarrow& v \in\mathbb{S}
+\\
+\textit{LanguageTag}(v, l) & \equiv& Datatype (v,
+\\
+&& \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#langString>} )
+\end{eqnarray*}}}
+
+Below follows a mapping from $\mathcal{L}_{\mathit{rocrate}}$ to $\mathcal
+{L}_{\textit{RDF}}$ using \surl{schema.org} as vocabulary:
+%
+{\fontsize{9.6}{11.6}\selectfont{\begin{eqnarray*}
+\textit{Property}(p) & \Rightarrow& \textit{type}(p, \texttt
+{<http://www.w3.org/2000/01/rdf-schema\#Property>})
+\\
+\textit{Class}(c) & \Rightarrow& \textit{type}(c, \texttt {<http://www.w3.org/2000/01/rdf-schema
+\#Class>})
+\\
+\textit{Dataset}(d) & \Rightarrow& \textit{type}(d, \texttt {<http://schema.org/Dataset>})
+\\
+\textit{File}(f) & \Rightarrow& \textit{type}(f, \texttt {<http://schema.org/MediaObject>})
+\\
+\textit{ContextualEntity}(e) & \Rightarrow& \textit{type}(f, \texttt
+{<http://schema.org/Thing>})
+\\
+\textit{CreativeWork}(e) & \Rightarrow& \textit{ContextualEntity}(e) \land
+\textit{type}(e, \texttt{<http://schema.org/CreativeWork>})
+\\
+\textit{hasPart}(e, t) & \Rightarrow& \textit{Relation}(e, \texttt
+{<http://schema.org/hasPart>}, t)
+\\
+\textit{name}(e, n) & \Rightarrow& \textit{Attribute}(e, \texttt
+{<http://schema.org/name>}, n)
+\\
+\textit{description}(e, s) & \Rightarrow& \textit{Attribute}(e, \texttt
+{<http://schema.org/description>}, s)
+\\
+\textit{datePublished}(e, d) & \Rightarrow& Attribute(e, \texttt
+{<http://schema.org/datePublished>}, d)
+\\
+\textit{license}(e, l) & \Rightarrow& \textit{Relation}(e, \texttt
+{<http://schema.org/license>}, l) \land \textit{CreativeWork}(l)
+\\
+\textit{type}(e, t) & \Rightarrow& \textit{Relation}(e, \texttt
+{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#type>}, t)
+\\
+& & {}\land \textit{Class}(t)
+\\
+\textit{String}(s) & \equiv& \textit{Value}(s) \land s \in\mathbb{S}
+\\
+\textit{String}(s) & \Rightarrow& \textit{Datatype}(s, \texttt
+{<http://www.w3.org/2001/XMLSchema\#string>})
+\\
+\textit{Decimal}(d) & \equiv& \textit{Value}(d) \land d \in\mathbb{R}
+\\
+\textit{Decimal}(d) & \Rightarrow& \textit{Datatype}(d, \texttt
+{<http://www.w3.org/2001/XMLSchema\#decimal>})
+\\
+\textit{Relation}(s,p,o) & \Rightarrow& \textit{Triple}(s,p,o) \land
+\textit{IRI}(s) \land \textit{IRI}(o)
+\\
+\textit{Attribute}(s,p,o) & \Rightarrow& \textit{Triple}(s,p,o) \land
+\textit{IRI}(s) \land \textit{Literal}(o)
+\end{eqnarray*}}}
+
+Note that in the JSON-LD serialization of RO-Crate, the expression of
+$\textit{Class}$ and $\textit{Property}$ is typically indirect: The JSON-LD \texttt
+{@context} maps to \surl{schema.org} IRIs, which, when resolved as Linked
+Data, embed their formal definition as RDFa. Extensions may however
+include such term definitions directly in the RO-Crate.
+
+\subsection{RO-Crate 1.1 Metadata File Descriptor}
+
+An important RO-Crate principle is that of being \textbf
+{self-described} Therefore the serialisation of the RO-Crate into a
+file should also describe itself in a Metadata File Descriptor,\footnote{\url
+{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#ro-crate-metadata-file-descriptor}} indicating it is $\textit{about}$ (describing) the RO-Crate root data entity,
+and that it $\textit{conformsTo}$ a particular version of the RO-Crate specification:
+%
+\begin{eqnarray*}
+\textit{about}(s,o) & \Rightarrow& \textit{Relation}(s, \texttt
+{<http://schema.org/about>}, o)
+\\
+\textit{conformsTo}(s,o) & \Rightarrow& \textit{Relation}(s, \texttt
+{<http://purl.org/dc/terms/conformsTo>}, o)
+\\
+\textit{MetadataFile}(m) & \Rightarrow& \textit{CreativeWork}(m) \land
+\textit{about}(m,R) {\wedge } \textit{ROCrate}(R) \land
+\\
+& &{} \textit{conformsTo}(m, \texttt{<https://w3id.org/ro/crate/1.1>})
+\end{eqnarray*}
+
+Note that although the metadata file necessarily is an \emph
+{information resource} written to disk or served over the network (as
+JSON-LD), it is not considered to be a contained \emph{part} of the
+RO-Crate in the form of a \emph{data entity}, rather it is described
+only as a \emph{contextual entity}.
+
+In the conceptual model the \emph{RO-Crate Metadata File} can be seen
+as the top-level node that describes the \emph{RO-Crate Root}, however
+in the formal model (and the JSON-LD format) the metadata file
+descriptor is an additional contextual entity that is not affecting the
+depth-limit of the RO-Crate.
+
+\subsection{Forward-chained production rules for JSON-LD}
+
+Combining the above predicates and \surl{schema.org} mapping with rudimentary
+JSON templates, these forward-chaining production rules can output
+JSON-LD according to the RO-Crate 1.1 specification:\footnote{\textbf{Limitations:}
+ Contextual entities not related from the
+RO-Crate (e.g. using inverse relations to a data entity) would not be
+covered by the single direction $Mentions(R, s)$ production rule; see
+\href{https://github.com/ResearchObject/ro-crate/issues/122}{GitHub
+issue ResearchObject/ro-crate\#122}. The $\textit{datePublished}(e, d)$ rule do
+not include syntax checks for the ISO 8601 datetime format. Compared
+with RO-Crate examples, this generated JSON-LD does not use a
+$\textit{@context}$ as the IRIs are produced unshortened, a post-step could do
+JSON-LD Flattening with a versioned RO-Crate context. The \texttt
+{@type} expansion is included for clarity, even though this is also
+implied by the $\textit{type}(e, t)$ expansion to $\textit{Relation}(e, \texttt{xsd:type})$.}
+%
+\begin{eqnarray*}
+\textit{Mentions}(R, s) \land \textit{Relation}(s,p,o) & \Rightarrow&
+\textit{Mentions}(R, o)
+\\
+\textit{IRI}(i) & \Rightarrow& \texttt{"} i \texttt{"}
+\\
+\textit{Decimal}(d) & \Rightarrow& d
+\\
+\textit{String}(s) & \Rightarrow& \texttt{"} s \texttt{"}
+\\
+\forall e \forall t \ \textit{type}(e, t) & \Rightarrow& \texttt{\{"@id":} e
+\texttt{,}
+\\
+&& \ \ \texttt{"@type":} t
+\\
+&& \texttt{\}}
+\\
+\forall s \forall p \forall o \ \textit{Relation}(s,p,o) & \Rightarrow& \texttt{
+\{"@id":} s \texttt{,}
+\\
+&& \ \ p \texttt{: \{ "@id":} o \texttt{\}}
+\\
+&& \texttt{\}}
+\\
+\forall s \forall p \forall v \ \textit{Attribute}(s,p,v) & \Rightarrow&
+\texttt{\{"@id":} s \texttt{,}
+\\
+&& \ \ p \texttt{:} v
+\\
+&& \texttt{\}}
+\\
+\forall r \forall c\ \textit{ROCrate}(r) & \Rightarrow& \texttt{\{
+"@graph": [}
+\\
+&& \ \ \ \ \textit{Mentions}(r, c)*
+\\
+&& \ \ \ \texttt{]}
+\\
+&& \texttt{\}}
+\\
+R & \equiv& \texttt{<./>}
+\\
+R & \Rightarrow& \textit{MetadataFile}(\texttt{<ro-crate-metadata.json>})
+\end{eqnarray*}
+
+This exposes the first order logic domain of discourse of IRIs, with
+rational numbers and strings as their corresponding JSON-LD
+representation. These production rules first grow the graph of $R$ by
+adding a transitive rule -- anything described in $R$ which is related
+to $o$, means that $o$ is also mentioned by the $\textit{ROCrate}(R)$. For
+simplicity this rule is one-way; in theory the graph can also contain
+free-standing contextual entities that have outgoing relations to data-
+and contextual entities, but these are proposed to be bound to the root
+data entity with \surl{schema.org} relation \url{http://schema.org/mentions}.
+
+\section{RO-Crate Community}
+
+\label{sec:communitylist}
+
+As of 2021-10-04, the \textit{RO-Crate} Community members are:
+\begin{itemize}
+\item[*] Peter Sefton \url{https://orcid.org/0000-0002-3545-944X} (co-chair)
+\item[*] Stian Soiland-Reyes \url{https://orcid.org/0000-0001-9842-9718} (co-chair)
+\item[*] Eoghan \'{O} Carrag\'{a}in \url
+{https://orcid.org/0000-0001-8131-2150} (emeritus chair)
+\item[*] Oscar Corcho \url{https://orcid.org/0000-0002-9260-0753}
+\item[*] Daniel Garijo \url{https://orcid.org/0000-0003-0454-7145}
+\item[*] Raul Palma \url{https://orcid.org/0000-0003-4289-4922}
+\item[*] Frederik Coppens \url{https://orcid.org/0000-0001-6565-5145}
+\item[*] Carole Goble \url{https://orcid.org/0000-0003-1219-2137}
+\item[*] Jos\'{e} Mar\'ia Fern\'{a}ndez \url{https://orcid.org/0000-0002-4806-5140}
+\item[*] Kyle Chard \url{https://orcid.org/0000-0002-7370-4805}
+\item[*] Jose Manuel Gomez-Perez \url{https://orcid.org/0000-0002-5491-6431}
+\item[*] Michael R Crusoe \url{https://orcid.org/0000-0002-2961-9670}
+\item[*] Ignacio Eguinoa \url{https://orcid.org/0000-0002-6190-122X}
+\item[*] Nick Juty \url{https://orcid.org/0000-0002-2036-8350}
+\item[*] Kristi Holmes \url{https://orcid.org/0000-0001-8420-5254}
+\item[*] Jason A. Clark \url{https://orcid.org/0000-0002-3588-6257}
+\item[*] Salvador Capella-Gutierrez \url{https://orcid.org/0000-0002-0309-604X}
+\item[*] Alasdair J. G. Gray \url{https://orcid.org/0000-0002-5711-4872}
+\item[*] Stuart Owen \url{https://orcid.org/0000-0003-2130-0865}
+\item[*] Alan R Williams \url{https://orcid.org/0000-0003-3156-2105}
+\item[*] Giacomo Tartari \url{https://orcid.org/0000-0003-1130-2154}
+\item[*] Finn Bacall \url{https://orcid.org/0000-0002-0048-3300}
+\item[*] Thomas Thelen \url{https://orcid.org/0000-0002-1756-2128}
+\item[*] Herv\'{e} M\'{e}nager \url{https://orcid.org/0000-0002-7552-1009}
+\item[*] Laura Rodr\'iguez-Navas \url{https://orcid.org/0000-0003-4929-1219}
+\item[*] Paul Walk \url{https://orcid.org/0000-0003-1541-5631}
+\item[*] brandon whitehead \url{https://orcid.org/0000-0002-0337-8610}
+\item[*] Mark Wilkinson \url{https://orcid.org/0000-0001-6960-357X}
+\item[*] Paul Groth \url{https://orcid.org/0000-0003-0183-6910}
+\item[*] Erich Bremer \url{https://orcid.org/0000-0003-0223-1059}
+\item[*] LJ Garcia Castro \url{https://orcid.org/0000-0003-3986-0510}
+\item[*] Karl Sebby \url{https://orcid.org/0000-0001-6022-9825}
+\item[*] Alexander Kanitz \url{https://orcid.org/0000-0002-3468-0652}
+\item[*] Ana Trisovic \url{https://orcid.org/0000-0003-1991-0533}
+\item[*] Gavin Kennedy \url{https://orcid.org/0000-0003-3910-0474}
+\item[*] Mark Graves \url{https://orcid.org/0000-0003-3486-8193}
+\item[*] Jasper Koehorst \url{https://orcid.org/0000-0001-8172-8981}
+\item[*] Simone Leo \url{https://orcid.org/0000-0001-8271-5429}
+\item[*] Marc Portier \url{https://orcid.org/0000-0002-9648-6484}
+\item[*] Paul Brack \url{https://orcid.org/0000-0002-5432-2748}
+\item[*] Milan Ojster\v{s}ek \url{https://orcid.org/0000-0003-1743-8300}
+\item[*] Bert Droesbeke \url{https://orcid.org/0000-0003-0522-5674}
+\item[*] Chenxu Niu \url{https://github.com/UstcChenxu}
+\item[*] Kosuke Tanabe \url{https://orcid.org/0000-0002-9986-7223}
+\item[*] Tomasz Miksa \url{https://orcid.org/0000-0002-4929-7875}
+\item[*] Marco La Rosa \url{https://orcid.org/0000-0001-5383-6993}
+\item[*] Cedric Decruw \url{https://orcid.org/0000-0001-6387-5988}
+\item[*] Andreas Czerniak \url{https://orcid.org/0000-0003-3883-4169}
+\item[*] Jeremy Jay \url{https://orcid.org/0000-0002-5761-7533}
+\item[*] Sergio Serra \url{https://orcid.org/0000-0002-0792-8157}
+\item[*] Ronald Siebes \url{https://orcid.org/0000-0001-8772-7904}
+\item[*] Shaun de Witt \url{https://orcid.org/0000-0003-4196-3658}
+\item[*] Shady El Damaty \url{https://orcid.org/0000-0002-2318-4477}
+\item[*] Douglas Lowe \url{https://orcid.org/0000-0002-1248-3594}
+\item[*] Sergio Serra \url{https://orcid.org/0000-0002-0792-8157}
+\item[*] Xuanqi Li \url{https://orcid.org/0000-0003-1498-6205}
+\end{itemize}
+\end{appendix}
+
+% structpyb loaded by imikolaityte, 2021-12-28 10:21:28
+%
+\begin{thebibliography}{}
+
+%%% bbsrt2.pl, ver. 2.5.8, 2017.08.09
+%b1 ###bbsrt2
+%b122 ###
+\bibitem{doi:10.1093/nar/gky379}
+%
+\begin{barticle}
+\bauthor{\binits{E.}~\bsnm{Afgan}},
+\bauthor{\binits{D.}~\bsnm{Baker}},
+\bauthor{\binits{B.}~\bsnm{Batut}},
+\bauthor{\binits{M.}~\bsnm{van den Beek}},
+\bauthor{\binits{D.}~\bsnm{Bouvier}},
+\bauthor{\binits{M.}~\bsnm{Cech}},
+\bauthor{\binits{J.}~\bsnm{Chilton}},
+\bauthor{\binits{D.}~\bsnm{Clements}},
+\bauthor{\binits{N.}~\bsnm{Coraor}},
+\bauthor{\binits{B.A.}~\bsnm{Gr\"{u}ning}},
+\bauthor{\binits{A.}~\bsnm{Guerler}},
+\bauthor{\binits{J.}~\bsnm{Hillman-Jackson}},
+\bauthor{\binits{S.}~\bsnm{Hiltemann}},
+\bauthor{\binits{V.}~\bsnm{Jalili}},
+\bauthor{\binits{H.}~\bsnm{Rasche}},
+\bauthor{\binits{N.}~\bsnm{Soranzo}},
+\bauthor{\binits{J.}~\bsnm{Goecks}},
+\bauthor{\binits{J.}~\bsnm{Taylor}},
+\bauthor{\binits{A.}~\bsnm{Nekrutenko}} and
+\bauthor{\binits{D.}~\bsnm{Blankenberg}},
+\batitle{The Galaxy platform for accessible, reproducible and collaborative
+biomedical analyses: 2018 update},
+\bjtitle{Nucleic Acids Research}
+\bvolume{46}(\bissue{W1})
+(\byear{2018}),
+\bfpage{W537}--\blpage{W544}.
+\bid{doi={10.1093/nar/gky379}}
+\end{barticle}
+%
+\OrigBibText
+E.~Afgan,
+D.~Baker,
+B.~Batut,
+M.~van~den Beek,
+D.~Bouvier,
+M.~Cech,
+J.~Chilton,
+D.~Clements,
+N.~Coraor,
+B.A.~Gr\"{u}ning,
+A.~Guerler,
+J.~Hillman-Jackson,
+S.~Hiltemann,
+V.~Jalili,
+H.~Rasche,
+N.~Soranzo,
+J.~Goecks,
+J.~Taylor,
+A.~Nekrutenko and
+D.~Blankenberg,
+\textit{The Galaxy platform for accessible, reproducible and collaborative
+biomedical analyses: 2018 update.},
+Nucleic Acids Research
+46(W1)
+(2018),
+W537--W544.
+doi:10.1093/nar/gky379.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b2 ###bbsrt2
+%b124 ###
+\bibitem{doi:10.5281/zenodo.4916734}
+%
+\begin{botherref}
+\oauthor{\binits{D.}~\bsnm{Agarwal}},
+\oauthor{\binits{C.}~\bsnm{Goble}},
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\oauthor{\binits{U.}~\bsnm{Sarkans}},
+\oauthor{\binits{D.}~\bsnm{Noesgaard}},
+\oauthor{\binits{U.}~\bsnm{Schindler}},
+\oauthor{\binits{M.}~\bsnm{Fenner}},
+\oauthor{\binits{P.}~\bsnm{Manghi}},
+\oauthor{\binits{S.}~\bsnm{Stall}},
+\oauthor{\binits{C.}~\bsnm{Coward}} and
+\oauthor{\binits{C.}~\bsnm{Erdmann}},
+Data Citation Community of Practice -- 8,
+Workshop, Zenodo/AGU, 2021.
+\url{https://data.agu.org/DataCitationCoP/2nd-workshop-data-citation}.
+\bid{doi={10.5281/zenodo.4916734}}
+\end{botherref}
+%
+\OrigBibText
+D.~Agarwal,
+C.~Goble,
+S.~Soiland-Reyes,
+U.~Sarkans,
+D.~Noesgaard,
+U.~Schindler,
+M.~Fenner,
+P.~Manghi,
+S.~Stall,
+C.~Coward and
+C.~Erdmann,
+Data Citation Community of Practice -- 8 June 2021 Workshop,
+Zenodo/AGU,
+2021.
+doi:10.5281/zenodo.4916734.
+\url{https://data.agu.org/DataCitationCoP/2nd-workshop-data-citation}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b3 ###bbsrt2
+%b87 ###
+\bibitem{dcat2}
+%
+\begin{botherref}
+\oauthor{\binits{R.}~\bsnm{Albertoni}},
+\oauthor{\binits{D.}~\bsnm{Browning}},
+\oauthor{\binits{S.}~\bsnm{Cox}},
+\oauthor{\binits{A.}~\bsnm{Gonzalez Beltran}},
+\oauthor{\binits{A.}~\bsnm{Perego}},
+\oauthor{\binits{P.}~\bsnm{Winstanley}} and
+\binstitute{Dataset Exchange Working Group},
+Data Catalog Vocabulary
+({DCAT}) -- Version 2,
+\textit{{W3C} Recommendation}
+(2020),
+\url{https://www.w3.org/TR/2020/REC-vocab-dcat-2-20200204/}.
+\end{botherref}
+%
+\OrigBibText
+R.~Albertoni,
+D.~Browning,
+S.~Cox,
+A.~Gonzalez~Beltran,
+A.~Perego,
+P.~Winstanley and
+Dataset Exchange Working Group:
+Data Catalog Vocabulary ({DCAT}) - Version 2,
+{W3C} Recommendation,
+2020.
+\url{https://www.w3.org/TR/2020/REC-vocab-dcat-2-20200204/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b4 ###bbsrt2
+%b111 ###
+\bibitem{doi:10.1038/s41586-019-0965-1}
+%
+\begin{barticle}
+\bauthor{\binits{A.}~\bsnm{Almeida}},
+\bauthor{\binits{A.L.}~\bsnm{Mitchell}},
+\bauthor{\binits{M.}~\bsnm{Boland}},
+\bauthor{\binits{S.C.}~\bsnm{Forster}},
+\bauthor{\binits{G.B.}~\bsnm{Gloor}},
+\bauthor{\binits{A.}~\bsnm{Tarkowska}},
+\bauthor{\binits{T.D.}~\bsnm{Lawley}} and
+\bauthor{\binits{R.D.}~\bsnm{Finn}},
+\batitle{A new genomic blueprint of the human gut microbiota},
+\bjtitle{Nature}
+\bvolume{568}(\bissue{7753})
+(\byear{2019}),
+\bfpage{499}--\blpage{504}.
+\bid{doi={10.1038/s41586-019-0965-1}}
+\end{barticle}
+%
+\OrigBibText
+A.~Almeida,
+A.L.~Mitchell,
+M.~Boland,
+S.C.~Forster,
+G.B.~Gloor,
+A.~Tarkowska,
+T.D.~Lawley and
+R.D.~Finn,
+\textit{A new genomic blueprint of the human gut microbiota.},
+Nature
+568(7753)
+(2019),
+499--504.
+doi:10.1038/s41586-019-0965-1.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b5 ###bbsrt2
+%b77 ###
+\bibitem{doi:10.1371/journal.pbio.3000099}
+%
+\begin{barticle}
+\bauthor{\binits{G.}~\bsnm{Alterovitz}},
+\bauthor{\binits{D.}~\bsnm{Dean}},
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\bauthor{\binits{A.}~\bsnm{Bell}},
+\bauthor{\binits{A.}~\bsnm{Hayes}},
+\bauthor{\binits{A.}~\bsnm{Suresh}},
+\bauthor{\binits{A.}~\bsnm{Purkayastha}},
+\bauthor{\binits{C.H.}~\bsnm{King}},
+\bauthor{\binits{D.}~\bsnm{Taylor}},
+\bauthor{\binits{E.}~\bsnm{Johanson}},
+\bauthor{\binits{E.E.}~\bsnm{Thompson}},
+\bauthor{\binits{E.}~\bsnm{Donaldson}},
+\bauthor{\binits{H.}~\bsnm{Morizono}},
+\bauthor{\binits{H.}~\bsnm{Tsang}},
+\bauthor{\binits{J.K.}~\bsnm{Vora}},
+\bauthor{\binits{J.}~\bsnm{Goecks}},
+\bauthor{\binits{J.}~\bsnm{Yao}},
+\bauthor{\binits{J.S.}~\bsnm{Almeida}},
+\bauthor{\binits{J.}~\bsnm{Keeney}},
+\bauthor{\binits{K.}~\bsnm{Addepalli}},
+\bauthor{\binits{K.}~\bsnm{Krampis}},
+\bauthor{\binits{K.M.}~\bsnm{Smith}},
+\bauthor{\binits{L.}~\bsnm{Guo}},
+\bauthor{\binits{M.}~\bsnm{Walderhaug}},
+\bauthor{\binits{M.}~\bsnm{Schito}},
+\bauthor{\binits{M.}~\bsnm{Ezewudo}},
+\bauthor{\binits{N.}~\bsnm{Guimera}},
+\bauthor{\binits{P.}~\bsnm{Walsh}},
+\bauthor{\binits{R.}~\bsnm{Kahsay}},
+\bauthor{\binits{S.}~\bsnm{Gottipati}},
+\bauthor{\binits{T.C.}~\bsnm{Rodwell}},
+\bauthor{\binits{T.}~\bsnm{Bloom}},
+\bauthor{\binits{Y.}~\bsnm{Lai}},
+\bauthor{\binits{V.}~\bsnm{Simonyan}} and
+\bauthor{\binits{R.}~\bsnm{Mazumder}},
+\batitle{Enabling precision medicine via standard communication of {HTS}
+provenance, analysis, and results},
+\bjtitle{PLoS Biology}
+\bvolume{16}(\bissue{12})
+(\byear{2018}),
+\bnumber{e3000099}.
+\bid{doi={10.1371/journal.pbio.3000099}}
+\end{barticle}
+%
+\OrigBibText
+G.~Alterovitz,
+D.~Dean,
+C.~Goble,
+M.R.~Crusoe,
+S.~Soiland-Reyes,
+A.~Bell,
+A.~Hayes,
+A.~Suresh,
+A.~Purkayastha,
+C.H.~King,
+D.~Taylor,
+E.~Johanson,
+E.E.~Thompson,
+E.~Donaldson,
+H.~Morizono,
+H.~Tsang,
+J.K.~Vora,
+J.~Goecks,
+J.~Yao,
+J.S.~Almeida,
+J.~Keeney,
+K.~Addepalli,
+K.~Krampis,
+K.M.~Smith,
+L.~Guo,
+M.~Walderhaug,
+M.~Schito,
+M.~Ezewudo,
+N.~Guimera,
+P.~Walsh,
+R.~Kahsay,
+S.~Gottipati,
+T.C.~Rodwell,
+T.~Bloom,
+Y.~Lai,
+V.~Simonyan and
+R.~Mazumder,
+\textit{Enabling precision medicine via standard communication of {HTS}
+provenance, analysis, and results.},
+{PLoS} Biology
+16(12)
+(2018),
+e3000099.
+doi:10.1371/journal.pbio.3000099.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b6 ###bbsrt2
+%b14 ###
+\bibitem{doi:10.1007/s10209-016-0475-y}
+%
+\begin{barticle}
+\bauthor{\binits{R.C.}~\bsnm{Amorim}},
+\bauthor{\binits{J.A.}~\bsnm{Castro}},
+\bauthor{\binits{J.}~\bsnm{Rocha da Silva}} and
+\bauthor{\binits{C.}~\bsnm{Ribeiro}},
+\batitle{A comparison of research data management platforms: Architecture,
+flexible metadata and interoperability},
+\bjtitle{Universal Access in the Information Society}
+(\byear{2016}),
+\bfpage{1}--\blpage{12}.
+\bid{doi={10.1007/s10209-016-0475-y}}
+\end{barticle}
+%
+\OrigBibText
+R.C.~Amorim,
+J.A.~Castro,
+J.~Rocha~da Silva and
+C.~Ribeiro,
+A comparison of research data management platforms: architecture, flexible
+metadata and interoperability,
+\textit{Universal Access in the Information Society}
+\textbf{00}
+(2016),
+1--12.
+doi:10.1007/s10209-016-0475-y.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b7 ###bbsrt2
+%b56 ###
+\bibitem{doi:10.5281/zenodo.3922136}
+%
+\begin{bbook}
+\bauthor{\binits{G.}~\bsnm{Arfaoui}} and
+\bauthor{\binits{M.}~\bsnm{Jaoua}},
+\bbtitle{{RO}-Crate {RDA} {maDMP} Mapper, Zenodo},
+\byear{2020}.
+\url{https://github.com/GhaithArf/ro-crate-rda-madmp-mapper}.
+\bid{doi={10.5281/zenodo.3922136}}
+\end{bbook}
+%
+\OrigBibText
+G.~Arfaoui and
+M.~Jaoua,
+\textit{{RO}-Crate {RDA} {maDMP} Mapper},
+Zenodo
+(2020).
+doi:10.5281/zenodo.3922136.
+\url{https://github.com/GhaithArf/ro-crate-rda-madmp-mapper}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b8 ###bbsrt2
+%b55 ###
+\bibitem{ro-composer}
+%
+\begin{botherref}
+\oauthor{\binits{F.}~\bsnm{Bacall}},
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}} and
+\oauthor{\binits{M.}~\bsnm{Soares e Silva}},
+eScienceLab: RO-Composer. \url
+{https://esciencelab.org.uk/projects/ro-composer/}.
+\end{botherref}
+%
+\OrigBibText
+F.~Bacall,
+S.~Soiland-Reyes and
+M.~Soares~e Silva,
+eScienceLab: RO-Composer.
+\url{https://esciencelab.org.uk/projects/ro-composer/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b9 ###bbsrt2
+%b46 ###
+\bibitem{ro-crate-ruby}
+%
+\begin{botherref}
+\oauthor{\binits{F.}~\bsnm{Bacall}} and
+\oauthor{\binits{M.}~\bsnm{Whitwell}},
+{GitHub} -- {ResearchObject}/ro-crate-ruby: A Ruby gem for creating,
+manipulating and reading {RO}-Crates.
+\url{https://github.com/ResearchObject/ro-crate-ruby}.
+\end{botherref}
+%
+\OrigBibText
+F.~Bacall and
+M.~Whitwell,
+{GitHub} - {ResearchObject}/ro-crate-ruby: A Ruby gem for creating,
+manipulating and reading {RO}-Crates.
+\url{https://github.com/ResearchObject/ro-crate-ruby}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b10 ###bbsrt2
+%b64 ###
+\bibitem{doi:10.1371/journal.ppat.1008643}
+%
+\begin{barticle}
+\bauthor{\binits{D.}~\bsnm{Baker}},
+\bauthor{\binits{M.}~\bsnm{van den Beek}},
+\bauthor{\binits{D.}~\bsnm{Blankenberg}},
+\bauthor{\binits{D.}~\bsnm{Bouvier}},
+\bauthor{\binits{J.}~\bsnm{Chilton}},
+\bauthor{\binits{N.}~\bsnm{Coraor}},
+\bauthor{\binits{F.}~\bsnm{Coppens}},
+\bauthor{\binits{I.}~\bsnm{Eguinoa}},
+\bauthor{\binits{S.}~\bsnm{Gladman}},
+\bauthor{\binits{B.}~\bsnm{Gr\"{u}ning}},
+\bauthor{\binits{N.}~\bsnm{Keener}},
+\bauthor{\binits{D.}~\bsnm{Larivi\`{e}re}},
+\bauthor{\binits{A.}~\bsnm{Lonie}},
+\bauthor{\binits{S.}~\bsnm{Kosakovsky Pond}},
+\bauthor{\binits{W.}~\bsnm{Maier}},
+\bauthor{\binits{A.}~\bsnm{Nekrutenko}},
+\bauthor{\binits{J.}~\bsnm{Taylor}} and
+\bauthor{\binits{S.}~\bsnm{Weaver}},
+\batitle{No more business as usual: Agile and effective responses to emerging
+pathogen threats require open data and open analytics},
+\bjtitle{PLoS Pathogens}
+\bvolume{16}(\bissue{8})
+(\byear{2020}),
+\bnumber{e1008643}.
+\bid{doi={10.1371/journal.ppat.1008643}}
+\end{barticle}
+%
+\OrigBibText
+D.~Baker,
+M.~van~den Beek,
+D.~Blankenberg,
+D.~Bouvier,
+J.~Chilton,
+N.~Coraor,
+F.~Coppens,
+I.~Eguinoa,
+S.~Gladman,
+B.~Gr\"{u}ning,
+N.~Keener,
+D.~Larivi\`{e}re,
+A.~Lonie,
+S.~Kosakovsky~Pond,
+W.~Maier,
+A.~Nekrutenko,
+J.~Taylor and
+S.~Weaver,
+\textit{No more business as usual: Agile and effective responses to emerging
+pathogen threats require open data and open analytics.},
+{PLoS} Pathogens
+16(8)
+(2020),
+e1008643.
+doi:10.1371/journal.ppat.1008643.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b11 ###bbsrt2
+%b89 ###
+\bibitem{doi:10.5334/dsj-2019-044}
+%
+\begin{barticle}
+\bauthor{\binits{M.}~\bsnm{Barker}},
+\bauthor{\binits{R.}~\bsnm{Wilkinson}} and
+\bauthor{\binits{A.}~\bsnm{Treloar}},
+\batitle{The Australian research data commons},
+\bjtitle{Data Science Journal}
+\bvolume{18}
+(\byear{2019}).
+\bid{doi={10.5334/dsj-2019-044}}
+\end{barticle}
+%
+\OrigBibText
+M.~Barker,
+R.~Wilkinson and
+A.~Treloar,
+The Australian Research Data Commons,
+\textit{Data Science Journal}
+\textbf{18}
+(2019), 00.
+doi:10.5334/dsj-2019-044.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b12 ###bbsrt2
+%b11 ###
+\bibitem{doi:10.1016/j.future.2011.08.004}
+%
+\begin{barticle}
+\bauthor{\binits{S.}~\bsnm{Bechhofer}},
+\bauthor{\binits{I.}~\bsnm{Buchan}},
+\bauthor{\binits{D.}~\bsnm{De Roure}},
+\bauthor{\binits{P.}~\bsnm{Missier}},
+\bauthor{\binits{J.}~\bsnm{Ainsworth}},
+\bauthor{\binits{J.}~\bsnm{Bhagat}},
+\bauthor{\binits{P.}~\bsnm{Couch}},
+\bauthor{\binits{D.}~\bsnm{Cruickshank}},
+\bauthor{\binits{M.}~\bsnm{Delderfield}},
+\bauthor{\binits{I.}~\bsnm{Dunlop}},
+\bauthor{\binits{M.}~\bsnm{Gamble}},
+\bauthor{\binits{D.}~\bsnm{Michaelides}},
+\bauthor{\binits{S.}~\bsnm{Owen}},
+\bauthor{\binits{D.}~\bsnm{Newman}},
+\bauthor{\binits{S.}~\bsnm{Sufi}} and
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\batitle{Why linked data is not enough for scientists},
+\bjtitle{Future Generation Computer Systems}
+\bvolume{29}(\bissue{2})
+(\byear{2013}),
+\bfpage{599}--\blpage{611}.
+\bid{doi={10.1016/j.future.2011.08.004}}
+\end{barticle}
+%
+\OrigBibText
+S.~Bechhofer,
+I.~Buchan,
+D.~De~Roure,
+P.~Missier,
+J.~Ainsworth,
+J.~Bhagat,
+P.~Couch,
+D.~Cruickshank,
+M.~Delderfield,
+I.~Dunlop,
+M.~Gamble,
+D.~Michaelides,
+S.~Owen,
+D.~Newman,
+S.~Sufi and
+C.~Goble,
+\textit{Why linked data is not enough for scientists},
+Future Generation Computer Systems
+29(2)
+(2013),
+599--611.
+doi:10.1016/j.future.2011.08.004.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b13 ###bbsrt2
+%b59 ###
+\bibitem{CheckMyCrate}
+%
+\begin{botherref}
+\oauthor{\binits{K.}~\bsnm{Belchev}},
+{KockataEPich}/{CheckMyCrate}: A command line application for
+validating a {RO}-Crate object against a {JSON} profile,
+\textit{GitHub}
+(2021).
+\url{https://github.com/KockataEPich/CheckMyCrate}.
+\end{botherref}
+%
+\OrigBibText
+K.~Belchev,
+{KockataEPich}/{CheckMyCrate}: A command line application for
+validating a
+{RO}-Crate object against a {JSON} profile.,
+GitHub,
+2021.
+\url{https://github.com/KockataEPich/CheckMyCrate}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b14 ###bbsrt2
+%b12 ###
+\bibitem{doi:10.1016/j.websem.2015.01.003}
+%
+\begin{barticle}
+\bauthor{\binits{K.}~\bsnm{Belhajjame}},
+\bauthor{\binits{J.}~\bsnm{Zhao}},
+\bauthor{\binits{D.}~\bsnm{Garijo}},
+\bauthor{\binits{M.}~\bsnm{Gamble}},
+\bauthor{\binits{K.}~\bsnm{Hettne}},
+\bauthor{\binits{R.}~\bsnm{Palma}},
+\bauthor{\binits{E.}~\bsnm{Mina}},
+\bauthor{\binits{O.}~\bsnm{Corcho}},
+\bauthor{\binits{J.M.}~\bsnm{G\'{o}mez-P\'{e}rez}},
+\bauthor{\binits{S.}~\bsnm{Bechhofer}},
+\bauthor{\binits{G.}~\bsnm{Klyne}} and
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\batitle{Using a suite of ontologies for preserving workflow-centric
+research objects},
+\bjtitle{Web Semantics: Science, Services and Agents on the World Wide Web}
+\bvolume{32}(\bissue{0})
+(\byear{2015}),
+\bfpage{16}--\blpage{42}.
+\bid{doi={10.1016/j.websem.2015.01.003}}
+\end{barticle}
+%
+\OrigBibText
+K.~Belhajjame,
+J.~Zhao,
+D.~Garijo,
+M.~Gamble,
+K.~Hettne,
+R.~Palma,
+E.~Mina,
+O.~Corcho,
+J.M.~G\'{o}mez-P\'{e}rez,
+S.~Bechhofer,
+G.~Klyne and
+C.~Goble,
+\textit{Using a suite of ontologies for preserving workflow-centric research
+objects},
+Web Semantics: Science, Services and Agents on the World Wide Web
+32(0)
+(2015),
+16--42.
+doi:10.1016/j.websem.2015.01.003.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b15 ###bbsrt2
+%b116 ###
+\bibitem{doi:10.3389/fninf.2017.00069}
+%
+\begin{barticle}
+\bauthor{\binits{F.C.Y.}~\bsnm{Benureau}} and
+\bauthor{\binits{N.P.}~\bsnm{Rougier}},
+\batitle{Re-run, repeat, reproduce, reuse, replicate: Transforming code
+into scientific contributions},
+\bjtitle{Frontiers in Neuroinformatics}
+\bvolume{11}
+(\byear{2017}),
+\bnumber{69}.
+\bid{doi={10.3389/fninf.2017.00069}}
+\end{barticle}
+%
+\OrigBibText
+F.C.Y.~Benureau and
+N.P.~Rougier,
+\textit{Re-run, Repeat, Reproduce, Reuse, Replicate: Transforming Code into
+Scientific Contributions.},
+Frontiers in Neuroinformatics
+11
+(2017),
+69.
+doi:10.3389/fninf.2017.00069.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b16 ###bbsrt2
+%b8 ###
+\bibitem{doi:10.1093/nar/gkl971}
+%
+\begin{barticle}
+\bauthor{\binits{H.}~\bsnm{Berman}},
+\bauthor{\binits{K.}~\bsnm{Henrick}},
+\bauthor{\binits{H.}~\bsnm{Nakamura}} and
+\bauthor{\binits{J.L.}~\bsnm{Markley}},
+\batitle{The worldwide Protein Data Bank ({wwPDB}): Ensuring a single,
+uniform archive of {PDB} data},
+\bjtitle{Nucleic Acids Research}
+\bvolume{35}\bcomment{(Database issue)}
+(\byear{2007}),
+\bfpage{D301}--\blpage{D303}.
+\bid{doi={10.1093/nar/gkl971}}
+\end{barticle}
+%
+\OrigBibText
+H.~Berman,
+K.~Henrick,
+H.~Nakamura and
+J.L.~Markley,
+\textit{The worldwide Protein Data Bank ({wwPDB}): ensuring a single, uniform
+archive of {PDB} data}.
+Nucleic Acids Research
+35 (Database issue)
+(2007),
+D301--3.
+doi:10.1093/nar/gkl971.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b17 ###bbsrt2
+%b68 ###
+\bibitem{doi:10.5281/zenodo.4705078}
+%
+\begin{barticle}
+\bauthor{\binits{F.}~\bsnm{Bietrix}},
+\bauthor{\binits{J.M.}~\bsnm{Carazo}},
+\bauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
+\bauthor{\binits{F.}~\bsnm{Coppens}},
+\bauthor{\binits{M.L.}~\bsnm{Chiusano}},
+\bauthor{\binits{R.}~\bsnm{David}},
+\bauthor{\binits{J.M.}~\bsnm{Fernandez}},
+\bauthor{\binits{M.}~\bsnm{Fratelli}},
+\bauthor{\binits{J.-K.}~\bsnm{Heriche}},
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{P.}~\bsnm{Gribbon}},
+\bauthor{\binits{P.}~\bsnm{Holub}},
+\bauthor{\binits{R.P.}~\bsnm{Joosten}},
+\bauthor{\binits{S.}~\bsnm{Leo}},
+\bauthor{\binits{S.}~\bsnm{Owen}},
+\bauthor{\binits{H.}~\bsnm{Parkinson}},
+\bauthor{\binits{R.}~\bsnm{Pieruschka}},
+\bauthor{\binits{L.}~\bsnm{Pireddu}},
+\bauthor{\binits{L.}~\bsnm{Porcu}},
+\bauthor{\binits{M.}~\bsnm{Raess}},
+\bauthor{\binits{L.}~\bsnm{Rodriguez- Navas}},
+\bauthor{\binits{A.}~\bsnm{Scherer}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}} and
+\bauthor{\binits{J.}~\bsnm{Tang}},
+\batitle{{EOSC}-life methodology framework to enhance reproducibility
+within {EOSC}-life},
+\bjtitle{Zenodo}
+(\byear{2021}).
+\bid{doi={10.5281/zenodo.4705078}}
+\end{barticle}
+%
+\OrigBibText
+F.~Bietrix,
+J.M.~Carazo,
+S.~Capella-Gutierrez,
+F.~Coppens,
+M.L.~Chiusano,
+R.~David,
+J.M.~Fernandez,
+M.~Fratelli,
+J.-K.~Heriche,
+C.~Goble,
+P.~Gribbon,
+P.~Holub,
+R.~P.~Joosten,
+S.~Leo,
+S.~Owen,
+H.~Parkinson,
+R.~Pieruschka,
+L.~Pireddu,
+L.~Porcu,
+M.~Raess,
+L.~Rodriguez-~Navas,
+A.~Scherer,
+S.~Soiland-Reyes and
+J.~Tang,
+\textit{{EOSC}-Life Methodology framework to enhance reproducibility within
+{EOSC}-Life},
+Zenodo
+(2021).
+doi:10.5281/zenodo.4705078.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b18 ###bbsrt2
+%b29 ###
+\bibitem{doi:10.4018/978-1-60960-593-3.ch008}
+%
+\begin{bchapter}
+\bauthor{\binits{C.}~\bsnm{Bizer}},
+\bauthor{\binits{T.}~\bsnm{Heath}} and
+\bauthor{\binits{T.}~\bsnm{Berners-Lee}},
+\bctitle{Linked data: The story so far},
+in: \bbtitle{Semantic Services, Interoperability and Web Applications:
+Emerging Concepts},
+\beditor{\binits{A.}~\bsnm{Sheth}}, ed.,
+\bsertitle{{IGI} Global},
+\byear{2011},
+pp.~\bfpage{205}--\blpage{227}.
+ISBN \bisbn{9781609605933}.
+\bid{doi={10.4018/978-1-60960-593-3.ch008}}
+\end{bchapter}
+%
+\OrigBibText
+C.~Bizer,
+T.~Heath and
+T.~Berners-Lee,
+\textit{Linked data: the story so far},
+in: Semantic services, interoperability and web applications: emerging
+concepts,
+A.~Sheth, ed.,
+{IGI} Global,
+2011,
+pp.~205--227.
+ISBN 9781609605933.
+doi:10.4018/978-1-60960-593-3.ch008.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b19 ###bbsrt2
+%b126 ###
+\bibitem{doi:10.1087/20150211}
+%
+\begin{barticle}
+\bauthor{\binits{A.}~\bsnm{Brand}},
+\bauthor{\binits{L.}~\bsnm{Allen}},
+\bauthor{\binits{M.}~\bsnm{Altman}},
+\bauthor{\binits{M.}~\bsnm{Hlava}} and
+\bauthor{\binits{J.}~\bsnm{Scott}},
+\batitle{Beyond authorship: Attribution, contribution, collaboration,
+and credit},
+\bjtitle{Learned Publishing}
+\bvolume{28}(\bissue{2})
+(\byear{2015}),
+\bfpage{151}--\blpage{155}.
+\bid{doi={10.1087/20150211}}
+\end{barticle}
+%
+\OrigBibText
+A.~Brand,
+L.~Allen,
+M.~Altman,
+M.~Hlava and
+J.~Scott,
+\textit{Beyond authorship: attribution, contribution, collaboration, and
+credit},
+Learned Publishing
+28(2)
+(2015),
+151--155.
+doi:10.1087/20150211.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b20 ###bbsrt2
+%b58 ###
+\bibitem{doi:10.5281/zenodo.3903463}
+%
+\begin{botherref}
+\oauthor{\binits{G.}~\bsnm{Brenner}},
+BrennerG/Ro-Crate\_2\_ma-DMP: v1.0.0,
+2020.
+\url{https://github.com/BrennerG/Ro-Crate\_2\_ma-DMP}.
+\bid{doi={10.5281/zenodo.3903463}}
+\end{botherref}
+%
+\OrigBibText
+G.~Brenner,
+BrennerG/Ro-Crate\_2\_ma-DMP: v1.0.0,
+2020.
+doi:10.5281/zenodo.3903463.
+\url{https://github.com/BrennerG/Ro-Crate_2_ma-DMP}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b21 ###bbsrt2
+%b86 ###
+\bibitem{doi:10.4126/frl01-006423289}
+%
+\begin{bbook}
+\bauthor{\binits{J.}~\bsnm{Cardoso}},
+\bauthor{\binits{L.J.}~\bsnm{Garcia Castro}},
+\bauthor{\binits{F.}~\bsnm{Ekaputra}},
+\bauthor{\binits{M.-C.}~\bsnm{Jacquemot-Perbal}},
+\bauthor{\binits{T.}~\bsnm{Miksa}} and
+\bauthor{\binits{J.}~\bsnm{Borbinha}},
+\bbtitle{Towards Semantic Representation of Machine-Actionable Data
+Management Plans, PUBLISSO},
+\byear{2020}.
+\url{https://repository.publisso.de/resource/frl:6423289}.
+\bid{doi={10.4126/frl01-006423289}}
+\end{bbook}
+%
+\OrigBibText
+J.~Cardoso,
+L.J.~Garcia~Castro,
+F.~Ekaputra,
+M.-C.~Jacquemot-Perbal,
+T.~Miksa and
+J.~Borbinha,
+\textit{Towards semantic representation of machine-actionable Data Management
+Plans},
+PUBLISSO
+(2020).
+doi:10.4126/frl01-006423289.
+\url{https://repository.publisso.de/resource/frl:6423289}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b22 ###bbsrt2
+%b84 ###
+\bibitem{doi:10.1007/978-3-030-45442-5_15}
+%
+\begin{bchapter}
+\bauthor{\binits{J.}~\bsnm{Cardoso}},
+\bauthor{\binits{D.}~\bsnm{Proen\c{c}a}} and
+\bauthor{\binits{J.}~\bsnm{Borbinha}},
+\bctitle{Machine-actionable data management plans: A knowledge retrieval
+approach to automate the assessment of funders' requirements},
+in: \bbtitle{Advances in Information Retrieval},
+\beditor{\binits{J.M.}~\bsnm{Jose}},
+\beditor{\binits{E.}~\bsnm{Yilmaz}},
+\beditor{\binits{J.}~\bsnm{Magalh\~{a}es}},
+\beditor{\binits{P.}~\bsnm{Castells}},
+\beditor{\binits{N.}~\bsnm{Ferro}},
+\beditor{\binits{M.J.}~\bsnm{Silva}} and
+\beditor{\binits{F.}~\bsnm{Martins}}, eds,
+\bpublisher{Springer International Publishing},
+\blocation{Cham},
+\byear{2020},
+pp.~\bfpage{118}--\blpage{125}.
+ISBN \bisbn{978-3-030-45442-5}.
+\bid{doi={10.1007/978-3-030-45442-5_15}}
+\end{bchapter}
+%
+\OrigBibText
+J.~Cardoso,
+D.~Proen{\c{c}}a and
+J.~Borbinha,
+\textit{Machine-Actionable Data Management Plans: A Knowledge Retrieval
+Approach to Automate the Assessment of Funders' Requirements},
+in: Advances in Information Retrieval,
+J.M.~Jose,
+E.~Yilmaz,
+J.~Magalh{\~a}es,
+P.~Castells,
+N.~Ferro,
+M.J.~Silva and
+F.~Martins, eds,
+Springer International Publishing,
+Cham,
+2020,
+pp.~118--125.
+ISBN 978-3-030-45442-5.
+doi:10.1007/978-3-030-45442-5\_15.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b23 ###bbsrt2
+%b13 ###
+\bibitem{doi:10.5281/zenodo.3250687}
+%
+\begin{barticle}
+\bauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{P.}~\bsnm{Sefton}} and
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\batitle{A lightweight approach to research object data packaging},
+\bjtitle{Zenodo}
+(\byear{2019}).
+\bid{doi={10.5281/zenodo.3250687}}
+\end{barticle}
+%
+\OrigBibText
+E.{\'{O}}.~Carrag\'{a}in,
+C.~Goble,
+P.~Sefton and
+S.~Soiland-Reyes,
+\textit{A lightweight approach to research object data packaging},
+Zenodo
+(2019).
+doi:10.5281/zenodo.3250687.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b24 ###bbsrt2
+%b17 ###
+\bibitem{chan_1995}
+%
+\begin{bbook}
+\bauthor{\binits{L.M.}~\bsnm{Chan}},
+\bbtitle{Library of Congress Subject Headings: Principles and Application},
+\bedition{3}rd edn,
+\bpublisher{Libraries Unlimited},
+\blocation{Englewood, Colo},
+\byear{1995},
+p.~\bfpage{556}.
+\url{https://eric.ed.gov/?id=ED387146}.
+ISBN \bisbn{9781563081910}.
+\end{bbook}
+%
+\OrigBibText
+L.M.~Chan,
+\textit{Library of Congress Subject Headings: Principles and Application},
+3rd edn,
+Libraries Unlimited,
+Englewood, Colo,
+1995,
+p.~556.
+ISBN 9781563081910.
+\url{https://eric.ed.gov/?id=ED387146}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b25 ###bbsrt2
+%b81 ###
+\bibitem{doi:10.1109/BigData.2016.7840618}
+%
+\begin{bchapter}
+\bauthor{\binits{K.}~\bsnm{Chard}},
+\bauthor{\binits{M.}~\bsnm{D'Arcy}},
+\bauthor{\binits{B.}~\bsnm{Heavner}},
+\bauthor{\binits{I.}~\bsnm{Foster}},
+\bauthor{\binits{C.}~\bsnm{Kesselman}},
+\bauthor{\binits{R.}~\bsnm{Madduri}},
+\bauthor{\binits{A.}~\bsnm{Rodriguez}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{K.}~\bsnm{Clark}},
+\bauthor{\binits{E.W.}~\bsnm{Deutsch}},
+\bauthor{\binits{I.}~\bsnm{Dinov}},
+\bauthor{\binits{N.}~\bsnm{Price}} and
+\bauthor{\binits{A.}~\bsnm{Toga}},
+\bctitle{I'll take that to go: Big data bags and minimal identifiers for
+exchange of large, complex datasets},
+in: \bbtitle{2016 {IEEE} International Conference on Big Data (Big Data)},
+\bpublisher{IEEE},
+\byear{2016},
+pp.~\bfpage{319}--\blpage{328}.
+\url{https://static.aminer.org/pdf/fa/bigdata2016/BigD418.pdf}.
+ISBN \bisbn{978-1-4673-9005-7}.
+\bid{doi={10.1109/BigData.2016.7840618}}
+\end{bchapter}
+%
+\OrigBibText
+K.~Chard,
+M.~D'Arcy,
+B.~Heavner,
+I.~Foster,
+C.~Kesselman,
+R.~Madduri,
+A.~Rodriguez,
+S.~Soiland-Reyes,
+C.~Goble,
+K.~Clark,
+E.W.~Deutsch,
+I.~Dinov,
+N.~Price and
+A.~Toga,
+\textit{I'll take that to go: Big data bags and minimal identifiers for
+exchange of large, complex datasets},
+in: 2016 {IEEE} International Conference on Big Data (Big Data),
+IEEE,
+2016,
+pp.~319--328.
+ISBN 978-1-4673-9005-7.
+doi:10.1109/BigData.2016.7840618.
+\url{https://static.aminer.org/pdf/fa/bigdata2016/BigD418.pdf}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b26 ###bbsrt2
+%b106 ###
+\bibitem{doi:10.1109/eScience.2019.00068}
+%
+\begin{bchapter}
+\bauthor{\binits{K.}~\bsnm{Chard}},
+\bauthor{\binits{N.}~\bsnm{Gaffney}},
+\bauthor{\binits{M.B.}~\bsnm{Jones}},
+\bauthor{\binits{K.}~\bsnm{Kowalik}},
+\bauthor{\binits{B.}~\bsnm{Ludascher}},
+\bauthor{\binits{T.}~\bsnm{McPhillips}},
+\bauthor{\binits{J.}~\bsnm{Nabrzyski}},
+\bauthor{\binits{V.}~\bsnm{Stodden}},
+\bauthor{\binits{I.}~\bsnm{Taylor}},
+\bauthor{\binits{T.}~\bsnm{Thelen}},
+\bauthor{\binits{M.J.}~\bsnm{Turk}} and
+\bauthor{\binits{C.}~\bsnm{Willis}},
+\bctitle{Application of {BagIt}-serialized research object bundles for
+packaging and re-execution of computational analyses},
+in: \bbtitle{15th International Conference on {eScience} ({eScience} 2019)},
+\bpublisher{IEEE},
+\byear{2019},
+pp.~\bfpage{514}--\blpage{521}.
+\url{https://zenodo.org/record/3381754}.
+ISBN \bisbn{978-1-7281-2451-3}.
+\bid{doi={10.1109/eScience.2019.00068}}
+\end{bchapter}
+%
+\OrigBibText
+K.~Chard,
+N.~Gaffney,
+M.B.~Jones,
+K.~Kowalik,
+B.~Ludascher,
+T.~{McPhillips},
+J.~Nabrzyski,
+V.~Stodden,
+I.~Taylor,
+T.~Thelen,
+M.J.~Turk and
+C.~Willis,
+\textit{Application of {BagIt}-Serialized Research Object Bundles for
+Packaging and Re-Execution of Computational Analyses},
+in: 15th International Conference on {eScience} ({eScience} 2019),
+IEEE,
+2019,
+pp.~514--521.
+ISBN 978-1-7281-2451-3.
+doi:10.1109/eScience.2019.00068.
+\url{https://zenodo.org/record/3381754}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b27 ###bbsrt2
+%b93 ###
+\bibitem{doi:10.1109/MCC.2014.52}
+%
+\begin{barticle}
+\bauthor{\binits{K.}~\bsnm{Chard}},
+\bauthor{\binits{S.}~\bsnm{Tuecke}} and
+\bauthor{\binits{I.}~\bsnm{Foster}},
+\batitle{Efficient and secure transfer, synchronization, and sharing of
+big data},
+\bjtitle{IEEE Cloud Computing}
+\bvolume{1}(\bissue{3})
+(\byear{2014}),
+\bfpage{46}--\blpage{55}.
+\bid{doi={10.1109/MCC.2014.52}}
+\end{barticle}
+%
+\OrigBibText
+K.~Chard,
+S.~Tuecke and
+I.~Foster,
+\textit{Efficient and Secure Transfer, Synchronization, and Sharing of Big
+Data},
+IEEE Cloud Computing
+1(3)
+(2014),
+46--55.
+doi:10.1109/MCC.2014.52.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b28 ###bbsrt2
+%b101 ###
+\bibitem{Ciccarese:17:WAD}
+%
+\begin{botherref}
+\oauthor{\binits{P.}~\bsnm{Ciccarese}},
+\oauthor{\binits{R.}~\bsnm{Sanderson}} and
+\oauthor{\binits{B.}~\bsnm{Young}},
+Web Annotation Data Model, {W3C} Recommendation,
+W3C,
+2017.
+\url{https://www.w3.org/TR/2017/REC-annotation-model-20170223/}.
+\end{botherref}
+%
+\OrigBibText
+P.~Ciccarese,
+R.~Sanderson and
+B.~Young,
+Web Annotation Data Model,
+{W3C} Recommendation,
+W3C,
+2017,
+https://www.w3.org/TR/2017/REC-annotation-model-20170223/.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b29 ###bbsrt2
+%b96 ###
+\bibitem{doi:10.1190/1.1822162}
+%
+\begin{bchapter}
+\bauthor{\binits{J.F.}~\bsnm{Claerbout}} and
+\bauthor{\binits{M.}~\bsnm{Karrenbach}},
+\bctitle{Electronic documents give reproducible research a new meaning},
+in: \bbtitle{{SEG} Technical Program Expanded Abstracts 1992, Society
+of Exploration Geophysicists},
+\byear{1992},
+pp.~\bfpage{601}--\blpage{604}.
+\bid{doi={10.1190/1.1822162}}
+\end{bchapter}
+%
+\OrigBibText
+J.F.~Claerbout and
+M.~Karrenbach,
+\textit{Electronic documents give reproducible research a new meaning},
+in: {SEG} Technical Program Expanded Abstracts 1992,
+Society of Exploration Geophysicists,
+1992,
+pp.~601--604.
+doi:10.1190/1.1822162.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b30 ###bbsrt2
+%b108 ###
+\bibitem{doi:10.1016/j.future.2017.01.012}
+%
+\begin{barticle}
+\bauthor{\binits{S.}~\bsnm{Cohen-Boulakia}},
+\bauthor{\binits{K.}~\bsnm{Belhajjame}},
+\bauthor{\binits{O.}~\bsnm{Collin}},
+\bauthor{\binits{J.}~\bsnm{Chopard}},
+\bauthor{\binits{C.}~\bsnm{Froidevaux}},
+\bauthor{\binits{A.}~\bsnm{Gaignard}},
+\bauthor{\binits{K.}~\bsnm{Hinsen}},
+\bauthor{\binits{P.}~\bsnm{Larmande}},
+\bauthor{\binits{Y.L.}~\bsnm{Bras}},
+\bauthor{\binits{F.}~\bsnm{Lemoine}},
+\bauthor{\binits{F.}~\bsnm{Mareuil}},
+\bauthor{\binits{H.}~\bsnm{M\'{e}nager}},
+\bauthor{\binits{C.}~\bsnm{Pradal}} and
+\bauthor{\binits{C.}~\bsnm{Blanchet}},
+\batitle{Scientific workflows for computational reproducibility in the
+life sciences: Status, challenges and opportunities},
+\bjtitle{Future Generation Computer Systems}
+\bvolume{75}
+(\byear{2017}),
+\bfpage{284}--\blpage{298}.
+\bid{doi={10.1016/j.future.2017.01.012}}
+\end{barticle}
+%
+\OrigBibText
+S.~Cohen-Boulakia,
+K.~Belhajjame,
+O.~Collin,
+J.~Chopard,
+C.~Froidevaux,
+A.~Gaignard,
+K.~Hinsen,
+P.~Larmande,
+Y.L.~Bras,
+F.~Lemoine,
+F.~Mareuil,
+H.~M\'{e}nager,
+C.~Pradal and
+C.~Blanchet,
+\textit{Scientific workflows for computational reproducibility in the life
+sciences: Status, challenges and opportunities},
+Future Generation Computer Systems
+75
+(2017),
+284--298.
+doi:10.1016/j.future.2017.01.012.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b31 ###bbsrt2
+%b36 ###
+\bibitem{pcdm}
+%
+\begin{botherref}
+\oauthor{\binits{S.}~\bsnm{Cossu}},
+\oauthor{\binits{E.}~\bsnm{Cowles}},
+\oauthor{\binits{K.}~\bsnm{Estlund}},
+\oauthor{\binits{C.}~\bsnm{Harlow}},
+\oauthor{\binits{T.}~\bsnm{Johnson}},
+\oauthor{\binits{M.}~\bsnm{Matienzo}},
+\oauthor{\binits{D.}~\bsnm{Lamb}},
+\oauthor{\binits{L.}~\bsnm{Rayle}},
+\oauthor{\binits{R.}~\bsnm{Sanderson}},
+\oauthor{\binits{J.}~\bsnm{Stroop}} and
+\oauthor{\binits{A.}~\bsnm{Woods}},
+Portland Common Data Model,
+2018.
+\url{https://github.com/duraspace/pcdm/wiki}.
+\end{botherref}
+%
+\OrigBibText
+S.~Cossu,
+E.~Cowles,
+K.~Estlund,
+C.~Harlow,
+T.~Johnson,
+M.~Matienzo,
+D.~Lamb,
+L.~Rayle,
+R.~Sanderson,
+J.~Stroop and
+A.~Woods,
+Portland Common Data Model,
+2018.
+\url{https://github.com/duraspace/pcdm/wiki}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b32 ###bbsrt2
+%b92 ###
+\bibitem{doi:10.1045/january2011-crosas}
+%
+\begin{barticle}
+\bauthor{\binits{M.}~\bsnm{Crosas}},
+\batitle{The dataverse network: An open-source application for sharing,
+discovering and preserving data},
+\bjtitle{D-Lib Magazine}
+\bvolume{17}(\bissue{1/2})
+(\byear{2011}).
+\bid{doi={10.1045/january2011-crosas}}
+\end{barticle}
+%
+\OrigBibText
+M.~Crosas,
+The Dataverse Network: An Open-Source Application for Sharing,
+Discovering and
+Preserving Data,
+\textit{D-Lib Magazine}
+\textbf{17}(1/2)
+(2011)
+00.
+doi:10.1045/january2011-crosas.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b33 ###bbsrt2
+%b91 ###
+\bibitem{doi:10.7557/5.5422}
+%
+\begin{botherref}
+\oauthor{\binits{M.}~\bsnm{Crosas}},
+Harvard Data Commons,
+European Dataverse Workshop 2020,
+Troms{\o}, Norway,
+2020. ISSN 2387-3086.
+\bid{doi={10.7557/5.5422}}
+\end{botherref}
+%
+\OrigBibText
+M.~Crosas,
+Harvard Data Commons,
+2020,
+European Dataverse Workshop 2020, Troms{\o}, Norway. 2020-01-23/--24.
+ISSN 2387-3086.
+doi:10.7557/5.5422.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b34 ###bbsrt2
+%b40 ###
+\bibitem{doi:10.1016/j.tibtech.2012.02.002}
+%
+\begin{barticle}
+\bauthor{\binits{L.C.}~\bsnm{Crosswell}} and
+\bauthor{\binits{J.M.}~\bsnm{Thornton}},
+\batitle{{ELIXIR}: A distributed infrastructure for European biological data},
+\bjtitle{Trends in Biotechnology}
+\bvolume{30}(\bissue{5})
+(\byear{2012}),
+\bfpage{241}--\blpage{242}.
+\bid{doi={10.1016/j.tibtech.2012.02.002}}
+\end{barticle}
+%
+\OrigBibText
+L.C.~Crosswell and
+J.M.~Thornton,
+\textit{{ELIXIR}: a distributed infrastructure for European biological data},
+Trends in Biotechnology
+30(5)
+(2012),
+241--242.
+doi:10.1016/j.tibtech.2012.02.002.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b35 ###bbsrt2
+%b49 ###
+\bibitem{about-lifemonitor}
+%
+\begin{botherref}
+{CRS4},
+{LifeMonitor}, a testing and monitoring service for scientific workflows.
+\url{https://about.lifemonitor.eu/}.
+\end{botherref}
+%
+\OrigBibText
+CRS4,
+{LifeMonitor}, a testing and monitoring service for scientific workflows.
+\url{https://about.lifemonitor.eu/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b36 ###bbsrt2
+%b69 ###
+\bibitem{doi:10.1145/3486897}
+%
+\begin{bbook}
+\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
+\bauthor{\binits{S.}~\bsnm{Abeln}},
+\bauthor{\binits{A.}~\bsnm{Iosup}},
+\bauthor{\binits{P.}~\bsnm{Amstutz}},
+\bauthor{\binits{J.}~\bsnm{Chilton}},
+\bauthor{\binits{N.}~\bsnm{Tijani\'{c}}},
+\bauthor{\binits{H.}~\bsnm{M\'{e}nager}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}} and
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bbtitle{Methods Included: Standardizing Computational Reuse and Portability
+with the Common Workflow Language},
+\bpublisher{Communications of the ACM},
+\byear{2021}.
+\bcomment{Accepted}.
+\bid{doi={10.1145/3486897}}
+\end{bbook}
+%
+\OrigBibText
+M.R.~Crusoe,
+S.~Abeln,
+A.~Iosup,
+P.~Amstutz,
+J.~Chilton,
+N.~Tijani\'{c},
+H.~M\'{e}nager,
+S.~Soiland-Reyes and
+C.~Goble,
+\textit{Methods Included: Standardizing Computational Reuse and Portability
+with the Common Workflow Language},
+Communications of the ACM
+(2021),
+Accepted.
+doi:10.1145/3486897.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b37 ###bbsrt2
+%b120 ###
+\bibitem{doi:10.1093/bioinformatics/btx192}
+%
+\begin{barticle}
+\bauthor{\binits{F.}~\bsnm{da~Veiga Leprevost}},
+\bauthor{\binits{B.A.}~\bsnm{Gr\"{u}ning}},
+\bauthor{\binits{S.}~\bsnm{Alves Aflitos}},
+\bauthor{\binits{H.L.}~\bsnm{R\"{o}st}},
+\bauthor{\binits{J.}~\bsnm{Uszkoreit}},
+\bauthor{\binits{H.}~\bsnm{Barsnes}},
+\bauthor{\binits{M.}~\bsnm{Vaudel}},
+\bauthor{\binits{P.}~\bsnm{Moreno}},
+\bauthor{\binits{L.}~\bsnm{Gatto}},
+\bauthor{\binits{J.}~\bsnm{Weber}},
+\bauthor{\binits{M.}~\bsnm{Bai}},
+\bauthor{\binits{R.C.}~\bsnm{Jimenez}},
+\bauthor{\binits{T.}~\bsnm{Sachsenberg}},
+\bauthor{\binits{J.}~\bsnm{Pfeuffer}},
+\bauthor{\binits{R.}~\bsnm{Vera Alvarez}},
+\bauthor{\binits{J.}~\bsnm{Griss}},
+\bauthor{\binits{A.I.}~\bsnm{Nesvizhskii}} and
+\bauthor{\binits{Y.}~\bsnm{Perez-Riverol}},
+\batitle{{BioContainers}: An open-source and community-driven framework
+for software standardization},
+\bjtitle{Bioinformatics}
+\bvolume{33}(\bissue{16})
+(\byear{2017}),
+\bfpage{2580}--\blpage{2582}.
+\bid{doi={10.1093/bioinformatics/btx192}}
+\end{barticle}
+%
+\OrigBibText
+F.~da~Veiga~Leprevost,
+B.A.~Gr\"{u}ning,
+S.~Alves~Aflitos,
+H.L.~R\"{o}st,
+J.~Uszkoreit,
+H.~Barsnes,
+M.~Vaudel,
+P.~Moreno,
+L.~Gatto,
+J.~Weber,
+M.~Bai,
+R.C.~Jimenez,
+T.~Sachsenberg,
+J.~Pfeuffer,
+R.~Vera~Alvarez,
+J.~Griss,
+A.I.~Nesvizhskii and
+Y.~Perez-Riverol,
+\textit{{BioContainers}: an open-source and community-driven framework for
+software standardization.},
+Bioinformatics
+33(16)
+(2017),
+2580--2582.
+doi:10.1093/bioinformatics/btx192.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b38 ###bbsrt2
+%b73 ###
+\bibitem{doi:10.3390/publications8020021}
+%
+\begin{barticle}
+\bauthor{\binits{K.}~\bsnm{De Smedt}},
+\bauthor{\binits{D.}~\bsnm{Koureas}} and
+\bauthor{\binits{P.}~\bsnm{Wittenburg}},
+\batitle{{FAIR} digital objects for science: From data pieces to
+actionable knowledge units},
+\bjtitle{Publications}
+\bvolume{8}(\bissue{2})
+(\byear{2020}),
+\bnumber{21}.
+\bid{doi={10.3390/publications8020021}}
+\end{barticle}
+%
+\OrigBibText
+K.~De~Smedt,
+D.~Koureas and
+P.~Wittenburg,
+\textit{{FAIR} digital objects for science: from data pieces to actionable
+knowledge units},
+Publications
+8(2)
+(2020),
+21.
+doi:10.3390/publications8020021.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b39 ###bbsrt2
+%b66 ###
+\bibitem{doi:10.1038/nbt.3820}
+%
+\begin{barticle}
+\bauthor{\binits{P.}~\bsnm{Di Tommaso}},
+\bauthor{\binits{M.}~\bsnm{Chatzou}},
+\bauthor{\binits{E.W.}~\bsnm{Floden}},
+\bauthor{\binits{P.P.}~\bsnm{Barja}},
+\bauthor{\binits{E.}~\bsnm{Palumbo}} and
+\bauthor{\binits{C.}~\bsnm{Notredame}},
+\batitle{Nextflow enables reproducible computational workflows},
+\bjtitle{Nature Biotechnology}
+\bvolume{35}(\bissue{4})
+(\byear{2017}),
+\bfpage{316}--\blpage{319}.
+\bid{doi={10.1038/nbt.3820}}
+\end{barticle}
+%
+\OrigBibText
+P.~Di~Tommaso,
+M.~Chatzou,
+E.W.~Floden,
+P.P.~Barja,
+E.~Palumbo and
+C.~Notredame,
+\textit{Nextflow enables reproducible computational workflows.},
+Nature Biotechnology
+35(4)
+(2017),
+316--319.
+doi:10.1038/nbt.3820.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b40 ###bbsrt2
+%b7 ###
+\bibitem{doi:10.3897/biss.3.37080}
+%
+\begin{barticle}
+\bauthor{\binits{M.}~\bsnm{Dillen}},
+\bauthor{\binits{Q.}~\bsnm{Groom}},
+\bauthor{\binits{D.}~\bsnm{Agosti}} and
+\bauthor{\binits{L.}~\bsnm{Nielsen}},
+\batitle{Zenodo, an archive and publishing repository: A tale of two herbarium
+specimen pilot projects},
+\bjtitle{Biodiversity Information Science and Standards}
+\bvolume{3}
+(\byear{2019}.
+\bid{doi={10.3897/biss.3.37080}}
+\end{barticle}
+%
+\OrigBibText
+M.~Dillen,
+Q.~Groom,
+D.~Agosti and
+L.~Nielsen,
+Zenodo, an Archive and Publishing Repository: A tale of two herbarium specimen
+pilot projects,
+\textit{Biodiversity Information Science and Standards}
+\textbf{3}
+(2019), 00-00.
+doi:10.3897/biss.3.37080.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b41 ###bbsrt2
+%b47 ###
+\bibitem{ro-crate-py}
+%
+\begin{botherref}
+\oauthor{\binits{B.}~\bsnm{Droesbeke}},
+\oauthor{\binits{I.}~\bsnm{Eguinoa}},
+\oauthor{\binits{A.}~\bsnm{Gaignard}},
+\oauthor{\binits{L.}~\bsnm{Simone}},
+\oauthor{\binits{L.}~\bsnm{Pireddu}},
+\oauthor{\binits{L.}~\bsnm{Rodr\'{\i}guez-Navas}} and
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+{GitHub} -- {ResearchObject}/ro-crate-py: Python library for {RO}-Crate.
+\url{https://github.com/researchobject/ro-crate-py}.
+\bid{doi={10.5281/zenodo.3956493}}
+\end{botherref}
+%
+\OrigBibText
+B.~Droesbeke,
+I.~Eguinoa,
+A.~Gaignard,
+L.~Simone,
+L.~Pireddu,
+L.~Rodr\'iguez-Navas and
+S.~Soiland-Reyes,
+{GitHub} - {ResearchObject}/ro-crate-py: Python library for {RO}-Crate.
+doi:10.5281/zenodo.3956493.
+\url{https://github.com/researchobject/ro-crate-py}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b42 ###bbsrt2
+%b30 ###
+\bibitem{doi:10.17487/rfc3987}
+%
+\begin{botherref}
+\oauthor{\binits{M.}~\bsnm{Duerst}} and
+\oauthor{\binits{M.}~\bsnm{Suignard}},
+Internationalized resource identifiers ({IRI}s),
+Technical Report,
+2005.
+\bid{doi={10.17487/rfc3987}}
+\end{botherref}
+%
+\OrigBibText
+M.~Duerst and
+M.~Suignard,
+Internationalized resource identifiers ({IRI}s),
+Technical Report,
+2005.
+doi:10.17487/rfc3987.
+\url{https://www.rfc-editor.org/info/rfc3987}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b43 ###bbsrt2
+%b112 ###
+\bibitem{ebi_ftp_umgs2019}
+%
+\begin{botherref}
+{EMBL-EBI Microbiome Informatics Team},
+{FTP} index of /pub/databases/metagenomics/umgs\_analyses/,
+2019.
+\url{http://ftp.ebi.ac.uk/pub/databases/metagenomics/umgs\_analyses/}.
+\end{botherref}
+%
+\OrigBibText
+{EMBL-EBI Microbiome Informatics Team},
+{FTP} index of /pub/databases/metagenomics/umgs\_analyses/,
+2019.
+\url{http://ftp.ebi.ac.uk/pub/databases/metagenomics/umgs\_analyses/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b44 ###bbsrt2
+% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
+%incorrectly.
+%b113 ###
+\bibitem{finn-lab-mgsgut}
+%
+\begin{botherref}
+EMBL-EBI Microbiome Informatics Team,
+{GitHub} -- Finn-Lab/{MGS}-gut: Analysing Metagenomic Species ({MGS}),
+\url{https://github.com/Finn-Lab/MGS-gut}.
+\end{botherref}
+%
+\OrigBibText
+{EMBL-EBI Microbiome Informatics Team},
+{GitHub} - Finn-Lab/{MGS}-gut: Analysing Metagenomic Species ({MGS}).
+\url{https://github.com/Finn-Lab/MGS-gut}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b45 ###bbsrt2
+%b65 ###
+\bibitem{doi:10.1038/s41587-020-0439-x}
+%
+\begin{barticle}
+\bauthor{\binits{P.A.}~\bsnm{Ewels}},
+\bauthor{\binits{A.}~\bsnm{Peltzer}},
+\bauthor{\binits{S.}~\bsnm{Fillinger}},
+\bauthor{\binits{H.}~\bsnm{Patel}},
+\bauthor{\binits{J.}~\bsnm{Alneberg}},
+\bauthor{\binits{A.}~\bsnm{Wilm}},
+\bauthor{\binits{M.U.}~\bsnm{Garcia}},
+\bauthor{\binits{P.}~\bsnm{Di Tommaso}} and
+\bauthor{\binits{S.}~\bsnm{Nahnsen}},
+\batitle{The nf-core framework for community-curated bioinformatics pipelines},
+\bjtitle{Nature Biotechnology}
+\bvolume{38}(\bissue{3})
+(\byear{2020}),
+\bfpage{276}--\blpage{278}.
+\bid{doi={10.1038/s41587-020-0439-x}}
+\end{barticle}
+%
+\OrigBibText
+P.A.~Ewels,
+A.~Peltzer,
+S.~Fillinger,
+H.~Patel,
+J.~Alneberg,
+A.~Wilm,
+M.U.~Garcia,
+P.~Di~Tommaso and
+S.~Nahnsen,
+\textit{The nf-core framework for community-curated bioinformatics
+pipelines.},
+Nature Biotechnology
+38(3)
+(2020),
+276--278.
+doi:10.1038/s41587-020-0439-x.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b46 ###bbsrt2
+%b15 ###
+\bibitem{farnel_2014}
+%
+\begin{bchapter}
+\bauthor{\binits{S.}~\bsnm{Farnel}} and
+\bauthor{\binits{A.}~\bsnm{Shiri}},
+\bctitle{Metadata for research data: Current practices and trends},
+in: \bbtitle{2014 Proceedings of the International Conference on Dublin
+Core and Metadata Applications},
+\beditor{\binits{W.}~\bsnm{Moen}} and
+\beditor{\binits{A.}~\bsnm{Rushing}}, eds,
+\bsertitle{Dublin Core Metadata Initiative},
+\byear{2014},
+\bcomment{ISSN 1939-1366}. \url
+{https://dcpapers.dublincore.org/pubs/article/view/3714}.
+\end{bchapter}
+%
+\OrigBibText
+S.~Farnel and
+A.~Shiri,
+\textit{Metadata for Research Data: Current Practices and Trends},
+in: 2014 Proceedings of the International Conference on Dublin Core
+and Metadata Applications,
+W.~Moen and
+A.~Rushing, eds,
+Dublin Core Metadata Initiative,
+2014.
+ISSN 1939-1366.
+\url{https://dcpapers.dublincore.org/pubs/article/view/3714}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b47 ###bbsrt2
+%b107 ###
+\bibitem{doip2.0}
+%
+\begin{botherref}
+\oauthor{\binits{D.}~\bsnm{Foundation}},
+Digital Object Interface Protocol Specification,
+version 2.0, Technical Report,
+(2018).
+\url{https://www.dona.net/sites/default/files/2018-11/DOIPv2Spec\_1.pdf}.
+\end{botherref}
+%
+\OrigBibText
+D.~Foundation,
+Digital Object Interface Protocol Specification, version 2.0,
+Technical Report,
+2018.
+\url{https://www.dona.net/sites/default/files/2018-11/DOIPv2Spec_1.pdf}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b48 ###bbsrt2
+%b104 ###
+\bibitem{doi:10.1016/j.future.2019.03.046}
+%
+\begin{barticle}
+\bauthor{\binits{A.}~\bsnm{Garcia-Silva}},
+\bauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
+\bauthor{\binits{R.}~\bsnm{Palma}},
+\bauthor{\binits{M.}~\bsnm{Krystek}},
+\bauthor{\binits{S.}~\bsnm{Mantovani}},
+\bauthor{\binits{F.}~\bsnm{Foglini}},
+\bauthor{\binits{V.}~\bsnm{Grande}},
+\bauthor{\binits{F.}~\bsnm{De Leo}},
+\bauthor{\binits{S.}~\bsnm{Salvi}},
+\bauthor{\binits{E.}~\bsnm{Trasatti}},
+\bauthor{\binits{V.}~\bsnm{Romaniello}},
+\bauthor{\binits{M.}~\bsnm{Albani}},
+\bauthor{\binits{C.}~\bsnm{Silvagni}},
+\bauthor{\binits{R.}~\bsnm{Leone}},
+\bauthor{\binits{F.}~\bsnm{Marelli}},
+\bauthor{\binits{S.}~\bsnm{Albani}},
+\bauthor{\binits{M.}~\bsnm{Lazzarini}},
+\bauthor{\binits{H.J.}~\bsnm{Napier}},
+\bauthor{\binits{H.M.}~\bsnm{Glaves}},
+\bauthor{\binits{T.}~\bsnm{Aldridge}},
+\bauthor{\binits{C.}~\bsnm{Meertens}},
+\bauthor{\binits{F.}~\bsnm{Boler}},
+\bauthor{\binits{H.W.}~\bsnm{Loescher}},
+\bauthor{\binits{C.}~\bsnm{Laney}},
+\bauthor{\binits{M.A.}~\bsnm{Genazzio}},
+\bauthor{\binits{D.}~\bsnm{Crawl}} and
+\bauthor{\binits{I.}~\bsnm{Altintas}},
+\batitle{Enabling {FAIR} research in Earth science through research objects},
+\bjtitle{Future Generation Computer Systems}
+\bvolume{98}
+(\byear{2019}),
+\bfpage{550}--\blpage{564}.
+\bid{doi={10.1016/j.future.2019.03.046}}
+\end{barticle}
+%
+\OrigBibText
+A.~Garcia-Silva,
+J.M.~Gomez-Perez,
+R.~Palma,
+M.~Krystek,
+S.~Mantovani,
+F.~Foglini,
+V.~Grande,
+F.~De~Leo,
+S.~Salvi,
+E.~Trasatti,
+V.~Romaniello,
+M.~Albani,
+C.~Silvagni,
+R.~Leone,
+F.~Marelli,
+S.~Albani,
+M.~Lazzarini,
+H.J.~Napier,
+H.M.~Glaves,
+T.~Aldridge,
+C.~Meertens,
+F.~Boler,
+H.W.~Loescher,
+C.~Laney,
+M.A.~Genazzio,
+D.~Crawl and
+I.~Altintas,
+\textit{Enabling {FAIR} research in Earth Science through research objects},
+Future Generation Computer Systems
+98
+(2019),
+550--564.
+doi:10.1016/j.future.2019.03.046.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b49 ###bbsrt2
+%b45 ###
+\bibitem{ro-crate-js}
+%
+\begin{botherref}
+{{GitHub}} -- {UTS}-{eResearch}/ro-crate-js: Research Object
+Crate ({RO}-Crate) utilities,
+\url{https://github.com/UTS-eResearch/ro-crate-js}.
+\end{botherref}
+%
+\OrigBibText
+{GitHub} - {UTS}-{eResearch}/ro-crate-js: Research Object Crate ({RO}-Crate)
+utilities.
+\url{https://github.com/UTS-eResearch/ro-crate-js}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b50 ###bbsrt2
+% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
+%incorrectly.
+%b51 ###
+\bibitem{galaxy2cwl}
+%
+\begin{botherref}
+{GitHub} -- workflowhub-eu/galaxy2cwl: Standalone version tool
+to get cwl descriptions (initially an abstract cwl interface)
+of galaxy workflows and Galaxy workflows executions.
+\url{https://github.com/workflowhub-eu/galaxy2cwl}.
+\end{botherref}
+%
+\OrigBibText
+{GitHub} - workflowhub-eu/galaxy2cwl: Standalone version tool to get cwl
+descriptions (initially an abstract cwl interface) of galaxy workflows and
+Galaxy workflows executions.
+\url{https://github.com/workflowhub-eu/galaxy2cwl}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b51 ###bbsrt2
+%b52 ###
+\bibitem{modpdsc}
+%
+\begin{botherref}
+{{GitHub} -- {CoEDL}/modpdsc},
+\url{https://github.com/CoEDL/modpdsc/}.
+\end{botherref}
+%
+\OrigBibText
+{GitHub} - {CoEDL}/modpdsc.
+\url{https://github.com/CoEDL/modpdsc/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b52 ###bbsrt2
+% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
+%incorrectly.
+%b54 ###
+\bibitem{ocfl-tools}
+%
+\begin{botherref}
+{GitHub} -- {CoEDL}/ocfl-tools: Tools to process and manipulate an
+{OCFL} tree.
+\url{https://github.com/CoEDL/ocfl-tools}.
+\end{botherref}
+%
+\OrigBibText
+{GitHub} - {CoEDL}/ocfl-tools: Tools to process and manipulate an
+{OCFL} tree.
+\url{https://github.com/CoEDL/ocfl-tools}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b53 ###bbsrt2
+% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
+%incorrectly.
+%b115 ###
+\bibitem{doi:10.1038/s41592-019-0350-x}
+%
+\begin{botherref}
+Giving software its due,
+\textit{Nature Methods}
+\textbf{16}(3)
+(2019),
+207--207.
+\bid{doi={10.1038/s41592-019-0350-x}}
+\end{botherref}
+%
+\OrigBibText
+Giving software its due,
+\textit{Nature Methods}
+\textbf{16}(3)
+(2019),
+207--207.
+doi:10.1038/s41592-019-0350-x.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b54 ###bbsrt2
+%b117 ###
+\bibitem{goble_presentation_2016}
+%
+\begin{bchapter}
+\bauthor{\binits{C.}~\bsnm{Goble}},
+in: \bbtitle{What Is Reproducibility? The R* Brouhaha},
+\bconflocation{Hannover, Germany},
+\byear{2016}.
+\url
+{http://repscience2016.research-infrastructures.eu/img/CaroleGoble-ReproScience2016v2.pdf}.
+\end{bchapter}
+%
+\OrigBibText
+C.~Goble,
+What is Reproducibility? The R* Brouhaha,
+Hannover, Germany, 2016.
+\url
+{http://repscience2016.research-infrastructures.eu/img/CaroleGoble-ReproScience2016v2.pdf}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b55 ###bbsrt2
+%b4 ###
+\bibitem{doi:10.1162/dint_a_00033}
+%
+\begin{barticle}
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{S.}~\bsnm{Cohen-Boulakia}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\bauthor{\binits{D.}~\bsnm{Garijo}},
+\bauthor{\binits{Y.}~\bsnm{Gil}},
+\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
+\bauthor{\binits{K.}~\bsnm{Peters}} and
+\bauthor{\binits{D.}~\bsnm{Schober}},
+\bjtitle{{FAIR} Computational Workflows, Data Intelligence}
+\bvolume{2}(\bissue{1--2})
+(\byear{2019}),
+\bfpage{108}--\blpage{121}.
+\bid{doi={10.1162/dint\_a\_00033}}
+\end{barticle}
+%
+\OrigBibText
+C.~Goble,
+S.~Cohen-Boulakia,
+S.~Soiland-Reyes,
+D.~Garijo,
+Y.~Gil,
+M.R.~Crusoe,
+K.~Peters and
+D.~Schober,
+\textit{{FAIR} Computational Workflows},
+Data Intelligence
+2(1--2)
+(2019),
+108--121.
+doi:10.1162/dint\_a\_00033.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b56 ###bbsrt2
+%b70 ###
+\bibitem{doi:10.5281/zenodo.4605654}
+%
+\begin{barticle}
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\bauthor{\binits{F.}~\bsnm{Bacall}},
+\bauthor{\binits{S.}~\bsnm{Owen}},
+\bauthor{\binits{A.}~\bsnm{Williams}},
+\bauthor{\binits{I.}~\bsnm{Eguinoa}},
+\bauthor{\binits{B.}~\bsnm{Droesbeke}},
+\bauthor{\binits{S.}~\bsnm{Leo}},
+\bauthor{\binits{L.}~\bsnm{Pireddu}},
+\bauthor{\binits{L.}~\bsnm{Rodr\'{\i}guez-Navas}},
+\bauthor{\binits{J.M.}~\bsnm{Fern\'{a}ndez}},
+\bauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
+\bauthor{\binits{H.}~\bsnm{M\'{e}nager}},
+\bauthor{\binits{B.}~\bsnm{Gr\"{u}ning}},
+\bauthor{\binits{B.}~\bsnm{Serrano-Solano}},
+\bauthor{\binits{P.}~\bsnm{Ewels}} and
+\bauthor{\binits{F.}~\bsnm{Coppens}},
+\batitle{Implementing {FAIR} digital objects in the {EOSC}-life
+workflow collaboratory},
+\bjtitle{Zenodo}
+(\byear{2021}).
+\bid{doi={10.5281/zenodo.4605654}}
+\end{barticle}
+%
+\OrigBibText
+C.~Goble,
+S.~Soiland-Reyes,
+F.~Bacall,
+S.~Owen,
+A.~Williams,
+I.~Eguinoa,
+B.~Droesbeke,
+S.~Leo,
+L.~Pireddu,
+L.~Rodr\'iguez-Navas,
+J.M.~Fern\'{a}ndez,
+S.~Capella-Gutierrez,
+H.~M\'{e}nager,
+B.~Gr\"{u}ning,
+B.~Serrano-Solano,
+P.~Ewels and
+F.~Coppens,
+\textit{Implementing {FAIR} Digital Objects in the {EOSC}-Life Workflow
+Collaboratory},
+Zenodo
+(2021).
+doi:10.5281/zenodo.4605654.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b57 ###bbsrt2
+%b98 ###
+\bibitem{doi:10.1093/nar/gkq429}
+%
+\begin{barticle}
+\bauthor{\binits{C.A.}~\bsnm{Goble}},
+\bauthor{\binits{J.}~\bsnm{Bhagat}},
+\bauthor{\binits{S.}~\bsnm{Aleksejevs}},
+\bauthor{\binits{D.}~\bsnm{Cruickshank}},
+\bauthor{\binits{D.}~\bsnm{Michaelides}},
+\bauthor{\binits{D.}~\bsnm{Newman}},
+\bauthor{\binits{M.}~\bsnm{Borkum}},
+\bauthor{\binits{S.}~\bsnm{Bechhofer}},
+\bauthor{\binits{M.}~\bsnm{Roos}},
+\bauthor{\binits{P.}~\bsnm{Li}} and
+\bauthor{\binits{D.}~\bsnm{De Roure}},
+\batitle{{myExperiment}: A repository and social network for the sharing
+of bioinformatics workflows},
+\bjtitle{Nucleic Acids Research}
+\bvolume{38}\bcomment{(Web Server issue)}
+(\byear{2010}),
+\bfpage{W677}--\blpage{W682}.
+\bid{doi={10.1093/nar/gkq429}}
+\end{barticle}
+%
+\OrigBibText
+C.A.~Goble,
+J.~Bhagat,
+S.~Aleksejevs,
+D.~Cruickshank,
+D.~Michaelides,
+D.~Newman,
+M.~Borkum,
+S.~Bechhofer,
+M.~Roos,
+P.~Li and
+D.~DeRoure,
+\textit{{myExperiment}: a repository and social network for the sharing of
+bioinformatics workflows}.
+Nucleic Acids Research
+38 (Web Server issue)
+(2010),
+W677--82.
+doi:10.1093/nar/gkq429.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b58 ###bbsrt2
+%b37 ###
+\bibitem{bioschemas_2017}
+%
+\begin{bchapter}
+\bauthor{\binits{A.}~\bsnm{Gray}},
+\bauthor{\binits{C.}~\bsnm{Goble}} and
+\bauthor{\binits{R.}~\bsnm{Jimenez}},
+in: \bbtitle{{Bioschemas Community}, Bioschemas: From Potato Salad to
+Protein Annotation},
+\bconflocation{Vienna, Austria},
+\byear{2017}.
+\url{https://iswc2017.semanticweb.org/paper-579/}.
+\end{bchapter}
+%
+\OrigBibText
+A.~Gray,
+C.~Goble,
+R.~Jimenez and
+{Bioschemas Community},
+Bioschemas: From Potato Salad to Protein Annotation,
+Vienna, Austria, 2017.
+\url{https://iswc2017.semanticweb.org/paper-579/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b59 ###bbsrt2
+%b88 ###
+\bibitem{doi:10.1109/MCSE.2016.92}
+%
+\begin{barticle}
+\bauthor{\binits{R.L.}~\bsnm{Grossman}},
+\bauthor{\binits{A.}~\bsnm{Heath}},
+\bauthor{\binits{M.}~\bsnm{Murphy}},
+\bauthor{\binits{M.}~\bsnm{Patterson}} and
+\bauthor{\binits{W.}~\bsnm{Wells}},
+\batitle{A case for data commons: Toward data science as a service},
+\bjtitle{Computing in science \& engineering}
+\bvolume{18}(\bissue{5})
+(\byear{2016}),
+\bfpage{10}--\blpage{20}.
+\bid{doi={10.1109/MCSE.2016.92}}
+\end{barticle}
+%
+\OrigBibText
+R.L.~Grossman,
+A.~Heath,
+M.~Murphy,
+M.~Patterson and
+W.~Wells,
+\textit{A case for data commons: toward data science as a service},
+Computing in science \& engineering
+18(5)
+(2016),
+10--20.
+doi:10.1109/MCSE.2016.92.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b60 ###bbsrt2
+%b110 ###
+\bibitem{doi:10.1016/j.cels.2018.03.014}
+%
+\begin{barticle}
+\bauthor{\binits{B.}~\bsnm{Gr\"{u}ning}},
+\bauthor{\binits{J.}~\bsnm{Chilton}},
+\bauthor{\binits{J.}~\bsnm{K\"{o}ster}},
+\bauthor{\binits{R.}~\bsnm{Dale}},
+\bauthor{\binits{N.}~\bsnm{Soranzo}},
+\bauthor{\binits{M.}~\bsnm{van den Beek}},
+\bauthor{\binits{J.}~\bsnm{Goecks}},
+\bauthor{\binits{R.}~\bsnm{Backofen}},
+\bauthor{\binits{A.}~\bsnm{Nekrutenko}} and
+\bauthor{\binits{J.}~\bsnm{Taylor}},
+\batitle{Practical computational reproducibility in the life sciences},
+\bjtitle{Cell Systems}
+\bvolume{6}(\bissue{6})
+(\byear{2018}),
+\bfpage{631}--\blpage{635}.
+\bid{doi={10.1016/j.cels.2018.03.014}}
+\end{barticle}
+%
+\OrigBibText
+B.~Gr\"{u}ning,
+J.~Chilton,
+J.~K\"{o}ster,
+R.~Dale,
+N.~Soranzo,
+M.~van~den Beek,
+J.~Goecks,
+R.~Backofen,
+A.~Nekrutenko and
+J.~Taylor,
+\textit{Practical computational reproducibility in the life sciences.},
+Cell Systems
+6(6)
+(2018),
+631--635.
+doi:10.1016/j.cels.2018.03.014.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b61 ###bbsrt2
+%b119 ###
+\bibitem{doi:10.1038/s41592-018-0046-7}
+%
+\begin{barticle}
+\bauthor{\binits{B.}~\bsnm{Gr\"{u}ning}},
+\bauthor{\binits{R.}~\bsnm{Dale}},
+\bauthor{\binits{A.}~\bsnm{Sj\"{o}din}},
+\bauthor{\binits{B.A.}~\bsnm{Chapman}},
+\bauthor{\binits{J.}~\bsnm{Rowe}},
+\bauthor{\binits{C.H.}~\bsnm{Tomkins-Tinch}},
+\bauthor{\binits{R.}~\bsnm{Valieris}},
+\bauthor{\binits{J.}~\bsnm{K\"{o}ster}} and
+\bauthor{\binits{B.}~\bsnm{Team}},
+\batitle{Bioconda: Sustainable and comprehensive software distribution
+for the life sciences},
+\bjtitle{Nature Methods}
+\bvolume{15}(\bissue{7})
+(\byear{2018}),
+\bfpage{475}--\blpage{476}.
+\bid{doi={10.1038/s41592-018-0046-7}}
+\end{barticle}
+%
+\OrigBibText
+B.~Gr\"{u}ning,
+R.~Dale,
+A.~Sj\"{o}din,
+B.A.~Chapman,
+J.~Rowe,
+C.H.~Tomkins-Tinch,
+R.~Valieris,
+J.~K\"{o}ster and
+B.~Team,
+\textit{Bioconda: sustainable and comprehensive software distribution
+for the
+life sciences.},
+Nature Methods
+15(7)
+(2018),
+475--476.
+doi:10.1038/s41592-018-0046-7.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b62 ###bbsrt2
+%b34 ###
+\bibitem{doi:10.1145/2857274.2857276}
+%
+\begin{barticle}
+\bauthor{\binits{R.V.}~\bsnm{Guha}},
+\bauthor{\binits{D.}~\bsnm{Brickley}} and
+\bauthor{\binits{S.}~\bsnm{Macbeth}},
+\batitle{Schema.org: Evolution of Structured Data on the Web: Big data
+makes common schemas even more necessary},
+\bjtitle{Queue}
+\bvolume{13}(\bissue{9})
+(\byear{2015}),
+\bfpage{10}--\blpage{37}.
+\bid{doi={10.1145/2857274.2857276}}
+\end{barticle}
+%
+\OrigBibText
+R.V.~Guha,
+D.~Brickley and
+S.~Macbeth,
+\textit{Schema.org: Evolution of Structured Data on the Web: Big data makes
+common schemas even more necessary},
+Queue
+13(9)
+(2015),
+10--37.
+doi:10.1145/2857274.2857276.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b63 ###bbsrt2
+%b24 ###
+\bibitem{doi:10.2200/S00334ED1V01Y201102WBE001}
+%
+\begin{bchapter}
+\bauthor{\binits{T.}~\bsnm{Heath}} and
+\bauthor{\binits{C.}~\bsnm{Bizer}},
+in: \bbtitle{Linked Data: Evolving the Web into a Global Data Space},
+Vol.~\bseriesno{1},
+\byear{2011},
+pp.~\bfpage{1}--\blpage{136},
+\bcomment{ISSN 2160-4711}.
+ISBN \bisbn{9781608454310}.
+\end{bchapter}
+%
+\OrigBibText
+T.~Heath and
+C.~Bizer,
+\textit{Linked Data: Evolving the Web into a Global Data Space},
+Vol.~1,
+2011,
+pp.~1--136.
+ISSN 2160-4711.
+ISBN 9781608454310.
+doi:10.2200/S00334ED1V01Y201102WBE001.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b64 ###bbsrt2
+% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
+%incorrectly.
+%b78 ###
+\bibitem{doi:10.1109/IEEESTD.2020.9094416}
+%
+\begin{botherref}
+IEEE Standard for Bioinformatics Analyses Generated by High-Throughput
+Sequencing (HTS) to Facilitate Communication,
+IEEE Std 2791-2020.
+ISBN 978-1-5044-6466-6.
+\bid{doi={10.1109/IEEESTD.2020.9094416}}
+\end{botherref}
+%
+\OrigBibText
+IEEE Standard for Bioinformatics Analyses Generated by High-Throughput
+Sequencing (HTS) to Facilitate Communication,
+IEEE Std 2791-2020.
+ISBN 978-1-5044-6466-6.
+doi:10.1109/IEEESTD.2020.9094416.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b65 ###bbsrt2
+%b90 ###
+\bibitem{doi:10.1182/blood-2017-03-735654}
+%
+\begin{barticle}
+\bauthor{\binits{M.A.}~\bsnm{Jensen}},
+\bauthor{\binits{V.}~\bsnm{Ferretti}},
+\bauthor{\binits{R.L.}~\bsnm{Grossman}} and
+\bauthor{\binits{L.M.}~\bsnm{Staudt}},
+\batitle{The {NCI} Genomic Data Commons as an engine for precision medicine},
+\bjtitle{Blood}
+\bvolume{130}(\bissue{4})
+(\byear{2017}),
+\bfpage{453}--\blpage{459}.
+\bid{doi={10.1182/blood-2017-03-735654}}
+\end{barticle}
+%
+\OrigBibText
+M.A.~Jensen,
+V.~Ferretti,
+R.L.~Grossman and
+L.M.~Staudt,
+\textit{The {NCI} Genomic Data Commons as an engine for precision medicine.},
+Blood
+130(4)
+(2017),
+453--459.
+doi:10.1182/blood-2017-03-735654.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b66 ###bbsrt2
+%b125 ###
+\bibitem{doi:10.5281/zenodo.4477164}
+%
+\begin{botherref}
+\oauthor{\binits{M.B.}~\bsnm{Jones}},
+\oauthor{\binits{S.}~\bsnm{Richard}},
+\oauthor{\binits{D.}~\bsnm{Vieglais}},
+\oauthor{\binits{A.}~\bsnm{Shepherd}},
+\oauthor{\binits{R.}~\bsnm{Duerr}},
+\oauthor{\binits{D.}~\bsnm{Fils}} and
+\oauthor{\binits{L.}~\bsnm{McGibbney}},
+Science-on-Schema.org v1.2.0, Technical Report,
+2021.
+\bid{doi={10.5281/zenodo.4477164}}
+\end{botherref}
+%
+\OrigBibText
+M.B.~Jones,
+S.~Richard,
+D.~Vieglais,
+A.~Shepherd,
+R.~Duerr,
+D.~Fils and
+L.~McGibbney,
+Science-on-Schema.org v1.2.0,
+Technical Report,
+2021.
+doi:10.5281/zenodo.4477164.
+\url{https://science-on-schema.org/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b67 ###bbsrt2
+%b103 ###
+\bibitem{doi:10.3233/978-1-61499-660-6-9}
+%
+\begin{bchapter}
+\bauthor{\binits{M.}~\bsnm{Katsumi}} and
+\bauthor{\binits{M.}~\bsnm{Gr\"{u}ninger}},
+\bctitle{What is ontology reuse?}
+in: \bbtitle{Formal Ontology in Information Systems},
+\beditor{\binits{R.}~\bsnm{Ferrario}} and
+\beditor{\binits{W.}~\bsnm{Kuhn}}, eds,
+\bsertitle{Frontiers in Artificial Intelligence and Applications},
+Vol.~\bseriesno{283},
+\bpublisher{{IOS} Press},
+\byear{2016},
+pp.~\bfpage{9}--\blpage{22}.
+ISBN \bisbn{978-1-61499-660-6}.
+\bid{doi={10.3233/978-1-61499-660-6-9}}
+\end{bchapter}
+%
+\OrigBibText
+M.~Katsumi and
+M.~Gr\"{u}ninger,
+\textit{What Is Ontology Reuse?},
+in: Formal Ontology in Information Systems,
+R.~Ferrario and
+W.~Kuhn, eds,
+Frontiers in Artificial Intelligence and Applications,
+Vol.~283,
+{IOS} Press,
+2016,
+pp.~9--22.
+ISBN 978-1-61499-660-6.
+doi:10.3233/978-1-61499-660-6-9.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b68 ###bbsrt2
+%b35 ###
+\bibitem{doi:10.1093/gigascience/giz095}
+%
+\begin{barticle}
+\bauthor{\binits{F.Z.}~\bsnm{Khan}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\bauthor{\binits{R.O.}~\bsnm{Sinnott}},
+\bauthor{\binits{A.}~\bsnm{Lonie}},
+\bauthor{\binits{C.}~\bsnm{Goble}} and
+\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
+\batitle{Sharing interoperable workflow provenance: A review of best practices
+and their practical application in {CWLProv}},
+\bjtitle{GigaScience}
+\bvolume{8}(\bissue{11})
+(\byear{2019}).
+\bid{doi={10.1093/gigascience/giz095}}
+\end{barticle}
+%
+\OrigBibText
+F.Z.~Khan,
+S.~Soiland-Reyes,
+R.O.~Sinnott,
+A.~Lonie,
+C.~Goble and
+M.R.~Crusoe,
+Sharing interoperable workflow provenance: A review of best practices
+and their
+practical application in {CWLProv},
+\textit{GigaScience}
+\textbf{8}(11)
+(2019)
+00.
+doi:10.1093/gigascience/giz095.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b69 ###bbsrt2
+%b109 ###
+\bibitem{doi:10.1002/cpe.1228}
+%
+\begin{barticle}
+\bauthor{\binits{J.}~\bsnm{Kim}},
+\bauthor{\binits{E.}~\bsnm{Deelman}},
+\bauthor{\binits{Y.}~\bsnm{Gil}},
+\bauthor{\binits{G.}~\bsnm{Mehta}} and
+\bauthor{\binits{V.}~\bsnm{Ratnakar}},
+\batitle{Provenance trails in the Wings/Pegasus system},
+\bjtitle{Concurrency and Computation: Practice and Experience}
+\bvolume{20}(\bissue{5})
+(\byear{2008}),
+\bfpage{587}--\blpage{597}.
+\bid{doi={10.1002/cpe.1228}}
+\end{barticle}
+%
+\OrigBibText
+J.~Kim,
+E.~Deelman,
+Y.~Gil,
+G.~Mehta and
+V.~Ratnakar,
+\textit{Provenance trails in the Wings/Pegasus system},
+Concurrency and Computation: Practice and Experience
+20(5)
+(2008),
+587--597.
+doi:10.1002/cpe.1228.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b70 ###bbsrt2
+%b23 ###
+\bibitem{doi:10.3233/978-1-61499-649-1-87}
+%
+\begin{bchapter}
+\bauthor{\binits{T.}~\bsnm{Kluyver}},
+\bauthor{\binits{B.}~\bsnm{Ragan-Kelley}},
+\bauthor{\binits{F.}~\bsnm{P\'{e}rez}},
+\bauthor{\binits{B.}~\bsnm{Granger}},
+\bauthor{\binits{M.}~\bsnm{Bussonnier}},
+\bauthor{\binits{J.}~\bsnm{Frederic}},
+\bauthor{\binits{K.}~\bsnm{Kelley}},
+\bauthor{\binits{J.}~\bsnm{Hamrick}},
+\bauthor{\binits{J.}~\bsnm{Grout}},
+\bauthor{\binits{S.}~\bsnm{Corlay}},
+\bauthor{\binits{P.}~\bsnm{Ivanov}},
+\bauthor{\binits{D.}~\bsnm{Avila}},
+\bauthor{\binits{S.}~\bsnm{Abdalla}},
+\bauthor{\binits{C.}~\bsnm{Willing}} and
+\binstitute{Jupyter Development Team},
+\bctitle{Jupyter Notebooks -- a publishing format for reproducible
+computational workflows},
+in: \bbtitle{Positioning and Power in Academic Publishing: Players,
+Agents and Agendas},
+\bsertitle{Proceedings of the 20th International Conference on
+Electronic Publishing},
+\bpublisher{IOS Press},
+\byear{2016},
+pp.~\bfpage{87}--\blpage{90},
+ISBN \bisbn{978-1-61499-649-1}.
+\bid{doi={10.3233/978-1-61499-649-1-87}}
+\end{bchapter}
+%
+\OrigBibText
+T.~Kluyver,
+B.~Ragan-Kelley,
+F.~P\'{e}rez,
+B.~Granger,
+M.~Bussonnier,
+J.~Frederic,
+K.~Kelley,
+J.~Hamrick,
+J.~Grout,
+S.~Corlay,
+P.~Ivanov,
+D.~Avila,
+S.~Abdalla,
+C.~Willing and
+{Jupyter Development Team},
+\textit{Jupyter Notebooks - a publishing format for reproducible computational
+workflows},
+in: Positioning and Power in Academic Publishing: Players, Agents and
+Agendas,
+IOS Press,
+2016,
+pp.~87--90,
+Proceedings of the 20th International Conference on Electronic
+Publishing.
+ISBN 978-1-61499-649-1.
+doi:10.3233/978-1-61499-649-1-87.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b71 ###bbsrt2
+%b10 ###
+\bibitem{doi:10.1016/j.ijhcs.2020.102562}
+%
+\begin{barticle}
+\bauthor{\binits{L.}~\bsnm{Koesten}},
+\bauthor{\binits{K.}~\bsnm{Gregory}},
+\bauthor{\binits{P.}~\bsnm{Groth}} and
+\bauthor{\binits{E.}~\bsnm{Simperl}},
+\batitle{Talking datasets -- understanding data sensemaking behaviours},
+\bjtitle{International journal of human-computer studies}
+\bvolume{146}
+(\byear{2021}),
+\bnumber{102562}.
+\bid{doi={10.1016/j.ijhcs.2020.102562}}
+\end{barticle}
+%
+\OrigBibText
+L.~Koesten,
+K.~Gregory,
+P.~Groth and
+E.~Simperl,
+\textit{Talking datasets -- Understanding data sensemaking behaviours},
+International journal of human-computer studies
+146
+(2021),
+102562.
+doi:10.1016/j.ijhcs.2020.102562.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b72 ###bbsrt2
+%b94 ###
+\bibitem{doi:10.1016/j.patter.2020.100136}
+%
+\begin{barticle}
+\bauthor{\binits{L.}~\bsnm{Koesten}},
+\bauthor{\binits{P.}~\bsnm{Vougiouklis}},
+\bauthor{\binits{E.}~\bsnm{Simperl}} and
+\bauthor{\binits{P.}~\bsnm{Groth}},
+\batitle{Dataset reuse: Toward translating principles to practice},
+\bjtitle{Patterns (New York, NY)}
+\bvolume{1}(\bissue{8})
+(\byear{2020}),
+\bnumber{100136}.
+\bid{doi={10.1016/j.patter.2020.100136}}
+\end{barticle}
+%
+\OrigBibText
+L.~Koesten,
+P.~Vougiouklis,
+E.~Simperl and
+P.~Groth,
+\textit{Dataset reuse: toward translating principles to practice},
+Patterns (New York, N.Y.)
+1(8)
+(2020),
+100136.
+doi:10.1016/j.patter.2020.100136.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b73 ###bbsrt2
+%b67 ###
+\bibitem{doi:10.1093/bioinformatics/bts480}
+%
+\begin{barticle}
+\bauthor{\binits{J.}~\bsnm{K\"{o}ster}} and
+\bauthor{\binits{S.}~\bsnm{Rahmann}},
+\batitle{Snakemake -- a scalable bioinformatics workflow engine},
+\bjtitle{Bioinformatics}
+\bvolume{28}(\bissue{19})
+(\byear{2012}),
+\bfpage{2520}--\blpage{2522}.
+\bid{doi={10.1093/bioinformatics/bts480}}
+\end{barticle}
+%
+\OrigBibText
+J.~K\"{o}ster and
+S.~Rahmann,
+\textit{Snakemake--a scalable bioinformatics workflow engine.},
+Bioinformatics
+28(19)
+(2012),
+2520--2522.
+doi:10.1093/bioinformatics/bts480.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b74 ###bbsrt2
+%b26 ###
+\bibitem{doi:10.17487/rfc8493}
+%
+\begin{botherref}
+\oauthor{\binits{J.}~\bsnm{Kunze}},
+\oauthor{\binits{J.}~\bsnm{Littman}},
+\oauthor{\binits{E.}~\bsnm{Madden}},
+\oauthor{\binits{J.}~\bsnm{Scancella}} and
+\oauthor{\binits{C.}~\bsnm{Adams}},
+The {BagIt} File Packaging Format,
+(V1.0), Technical Report,
+2018.
+\bid{doi={10.17487/RFC8493}}
+\end{botherref}
+%
+\OrigBibText
+J.~Kunze,
+J.~Littman,
+E.~Madden,
+J.~Scancella and
+C.~Adams,
+The {BagIt} File Packaging Format (V1.0),
+Technical Report,
+2018.
+doi:10.17487/RFC8493.
+\url{https://www.rfc-editor.org/info/rfc8493}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b75 ###bbsrt2
+%b16 ###
+\bibitem{doi:10.2777/620649}
+%
+\begin{botherref}
+\oauthor{\binits{K.}~\bsnm{Kurowski}},
+\oauthor{\binits{O.}~\bsnm{Corcho}},
+\oauthor{\binits{C.}~\bsnm{Choirat}},
+\oauthor{\binits{M.}~\bsnm{Eriksson}},
+\oauthor{\binits{F.}~\bsnm{Coppens}},
+\oauthor{\binits{M.}~\bsnm{van de Sanden}} and
+\oauthor{\binits{M.}~\bsnm{Ojster\v{s}ek}},
+{EOSC} Interoperability Framework, Technical Report,
+2021.
+\bid{doi={10.2777/620649}}
+\end{botherref}
+%
+\OrigBibText
+K.~Kurowski,
+O.~Corcho,
+C.~Choirat,
+M.~Eriksson,
+F.~Coppens,
+M.~van~de Sanden and
+M.~Ojster\v{s}ek,
+{EOSC} Interoperability Framework,
+Technical Report,
+2021.
+doi:10.2777/620649.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b76 ###bbsrt2
+%b105 ###
+\bibitem{doi:10.3233/APC200107}
+%
+\begin{barticle}
+\bauthor{\binits{C.}~\bsnm{Kyle}},
+\bauthor{\binits{G.}~\bsnm{Niall}},
+\bauthor{\binits{H.}~\bsnm{Mihael}},
+\bauthor{\binits{K.}~\bsnm{Kacper}},
+\bauthor{\binits{L.}~\bsnm{Bertram}},
+\bauthor{\binits{M.}~\bsnm{Timothy}},
+\bauthor{\binits{N.}~\bsnm{Jarek}},
+\bauthor{\binits{S.}~\bsnm{Victoria}},
+\bauthor{\binits{T.}~\bsnm{Ian}},
+\bauthor{\binits{T.}~\bsnm{Thomas}} \betal,
+\batitle{Toward enabling reproducibility for data-intensive research
+using the whole tale platform},
+\bjtitle{Advances in Parallel Computing 36 (Parallel Computing:
+Technology Trends)}
+(\byear{2020}),
+\bfpage{766}--\blpage{778}.
+\bid{doi={10.3233/APC200107}}
+\end{barticle}
+%
+\OrigBibText
+C.~Kyle,
+G.~Niall,
+H.~Mihael,
+K.~Kacper,
+L.~Bertram,
+M.~Timothy,
+N.~Jarek,
+S.~Victoria,
+T.~Ian,
+T.~Thomas and
+et~al.,
+\textit{Toward Enabling Reproducibility for Data-Intensive Research
+Using the
+Whole Tale Platform},
+Advances in Parallel Computing
+36 (Parallel Computing: Technology Trends), \textbf{00}
+(2020),
+766--778.
+doi:10.3233/APC200107.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b77 ###bbsrt2
+%b43 ###
+\bibitem{describo-online}
+%
+\begin{botherref}
+\oauthor{\binits{M.}~\bsnm{La Rosa}},
+Arkisto Platform: Describo Online.
+\url{https://arkisto-platform.github.io/describo-online/}.
+\end{botherref}
+%
+\OrigBibText
+M.~La~Rosa,
+Arkisto Platform: Describo Online.
+\url{https://arkisto-platform.github.io/describo-online/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b78 ###bbsrt2
+%b22 ###
+\bibitem{describo}
+%
+\begin{botherref}
+\oauthor{\binits{M.}~\bsnm{La Rosa}} and
+\oauthor{\binits{P.}~\bsnm{Sefton}},
+Arkisto Platform: Describo, \url{https://arkisto-platform.github.io/describo/}.
+\end{botherref}
+%
+\OrigBibText
+M.~La~Rosa and
+P.~Sefton,
+Arkisto Platform: Describo.
+\url{https://arkisto-platform.github.io/describo/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b79 ###bbsrt2
+%b28 ###
+\bibitem{doi:10.6087/kcse.192}
+%
+\begin{barticle}
+\bauthor{\binits{R.}~\bsnm{Lammey}},
+\batitle{Solutions for identification problems: A look at the research
+organization registry},
+\bjtitle{Science Editing}
+\bvolume{7}(\bissue{1})
+(\byear{2020}),
+\bfpage{65}--\blpage{69}.
+\bid{doi={10.6087/kcse.192}}
+\end{barticle}
+%
+\OrigBibText
+R.~Lammey,
+\textit{Solutions for identification problems: a look at the Research
+Organization Registry},
+Science Editing
+7(1)
+(2020),
+65--69.
+doi:10.6087/kcse.192.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b80 ###bbsrt2
+%b3 ###
+\bibitem{doi:10.3233/DS-190026}
+%
+\begin{barticle}
+\bauthor{\binits{A.-L.}~\bsnm{Lamprecht}},
+\bauthor{\binits{L.}~\bsnm{Garcia}},
+\bauthor{\binits{M.}~\bsnm{Kuzak}},
+\bauthor{\binits{C.}~\bsnm{Martinez}},
+\bauthor{\binits{R.}~\bsnm{Arcila}},
+\bauthor{\binits{E.}~\bsnm{Martin Del Pico}},
+\bauthor{\binits{V.}~\bsnm{Dominguez Del Angel}},
+\bauthor{\binits{S.}~\bsnm{van de Sandt}},
+\bauthor{\binits{J.}~\bsnm{Ison}},
+\bauthor{\binits{P.A.}~\bsnm{Martinez}},
+\bauthor{\binits{P.}~\bsnm{McQuilton}},
+\bauthor{\binits{A.}~\bsnm{Valencia}},
+\bauthor{\binits{J.}~\bsnm{Harrow}},
+\bauthor{\binits{F.}~\bsnm{Psomopoulos}},
+\bauthor{\binits{J.L.}~\bsnm{Gelpi}},
+\bauthor{\binits{N.}~\bsnm{Chue Hong}},
+\bauthor{\binits{C.}~\bsnm{Goble}} and
+\bauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
+\batitle{Towards {FAIR} principles for research software},
+\bjtitle{D\'{e}viance et soci\'{e}t\'{e}}
+(\byear{2019}),
+\bfpage{1}--\blpage{23}.
+\bid{doi={10.3233/DS-190026}}
+\end{barticle}
+%
+\OrigBibText
+A.-L.~Lamprecht,
+L.~Garcia,
+M.~Kuzak,
+C.~Martinez,
+R.~Arcila,
+E.~Martin Del~Pico,
+V.~Dominguez Del~Angel,
+S.~van~de Sandt,
+J.~Ison,
+P.A.~Martinez,
+P.~{McQuilton},
+A.~Valencia,
+J.~Harrow,
+F.~Psomopoulos,
+J.L.~Gelpi,
+N.~Chue~Hong,
+C.~Goble and
+S.~Capella-Gutierrez,
+Towards {FAIR} principles for~research~software,
+\textit{D\'{e}viance et soci\'{e}t\'{e}}, \textbf{00},
+(2019),
+1--23.
+doi:10.3233/DS-190026.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b81 ###bbsrt2
+%b102 ###
+\bibitem{PROVO}
+%
+\begin{botherref}
+\oauthor{\binits{T.}~\bsnm{Lebo}},
+\oauthor{\binits{S.}~\bsnm{Sahoo}},
+\oauthor{\binits{D.}~\bsnm{McGuinness}},
+\oauthor{\binits{K.}~\bsnm{Belhajjame}},
+\oauthor{\binits{J.}~\bsnm{Cheney}},
+\oauthor{\binits{D.}~\bsnm{Corsar}},
+\oauthor{\binits{D.}~\bsnm{Garijo}},
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\oauthor{\binits{S.}~\bsnm{Zednik}} and
+\oauthor{\binits{J.}~\bsnm{Zhao}},
+PROV-O: The PROV Ontology,
+Technical Report, W3C Recommendation Note 30,
+2013.
+\url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
+\end{botherref}
+%
+\OrigBibText
+T.~Lebo,
+S.~Sahoo,
+D.~McGuinness,
+K.~Belhajjame,
+J.~Cheney,
+D.~Corsar,
+D.~Garijo,
+S.~Soiland-Reyes,
+S.~Zednik and
+J.~Zhao,
+PROV-O: The PROV Ontology,
+Technical Report,
+2013,
+W3C Recommendation Note 30 April 2013.
+\url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b82 ###bbsrt2
+%b95 ###
+\bibitem{doi:10.1016/j.patter.2021.100322}
+%
+\begin{barticle}
+\bauthor{\binits{J.}~\bsnm{Leipzig}},
+\bauthor{\binits{D.}~\bsnm{N\"{u}st}},
+\bauthor{\binits{C.T.}~\bsnm{Hoyt}},
+\bauthor{\binits{K.}~\bsnm{Ram}} and
+\bauthor{\binits{J.}~\bsnm{Greenberg}},
+\batitle{The role of metadata in reproducible computational research},
+\bjtitle{Patterns}
+\bvolume{2}(\bissue{9})
+(\byear{2021}),
+\bnumber{100322}.
+\bid{doi={10.1016/j.patter.2021.100322}}
+\end{barticle}
+%
+\OrigBibText
+J.~Leipzig,
+D.~N\"{u}st,
+C.T.~Hoyt,
+K.~Ram and
+J.~Greenberg,
+\textit{The role of metadata in reproducible computational research},
+Patterns
+2(9)
+(2021),
+100322.
+doi:10.1016/j.patter.2021.100322.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b83 ###bbsrt2
+%b74 ###
+\bibitem{doi:10.48546/workflowhub.workflow.56.1}
+%
+\begin{botherref}
+\oauthor{\binits{D.}~\bsnm{Lowe}} and
+\oauthor{\binits{G.}~\bsnm{Bayarri}},
+Protein Ligand Complex {MD} Setup tutorial using {BioExcel} Building
+Blocks (biobb) (jupyter notebook),
+2021.
+\bid{doi={10.48546/workflowhub.workflow.56.1}}
+\end{botherref}
+%
+\OrigBibText
+D.~Lowe and
+G.~Bayarri,
+Protein Ligand Complex {MD} Setup tutorial using {BioExcel} Building Blocks
+(biobb) (jupyter notebook),
+2021.
+doi:10.48546/workflowhub.workflow.56.1.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b84 ###bbsrt2
+%b44 ###
+\bibitem{ro-crate-excel}
+%
+\begin{botherref}
+\oauthor{\binits{M.}~\bsnm{Lynch}} and
+\oauthor{\binits{P.}~\bsnm{Sefton}},
+npm: ro-crate-excel. \url{https://www.npmjs.com/package/ro-crate-excel}.
+\end{botherref}
+%
+\OrigBibText
+M.~Lynch and
+P.~Sefton,
+npm: ro-crate-excel.
+\url{https://www.npmjs.com/package/ro-crate-excel}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b85 ###bbsrt2
+%b80 ###
+\bibitem{github-lfs}
+%
+\begin{botherref}
+{Managing large files -- {GitHub} Docs}.
+\url
+{https://docs.github.com/en/repositories/working-with-files/managing-large-files}.
+\end{botherref}
+%
+\OrigBibText
+Managing large files - {GitHub} Docs.
+\url
+{https://docs.github.com/en/repositories/working-with-files/managing-large-files}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b86 ###bbsrt2
+%b25 ###
+\bibitem{doi:10.1371/journal.pbio.2001414}
+%
+\begin{barticle}
+\bauthor{\binits{J.A.}~\bsnm{McMurry}},
+\bauthor{\binits{N.}~\bsnm{Juty}},
+\bauthor{\binits{N.}~\bsnm{Blomberg}},
+\bauthor{\binits{T.}~\bsnm{Burdett}},
+\bauthor{\binits{T.}~\bsnm{Conlin}},
+\bauthor{\binits{N.}~\bsnm{Conte}},
+\bauthor{\binits{M.}~\bsnm{Courtot}},
+\bauthor{\binits{J.}~\bsnm{Deck}},
+\bauthor{\binits{M.}~\bsnm{Dumontier}},
+\bauthor{\binits{D.K.}~\bsnm{Fellows}},
+\bauthor{\binits{A.}~\bsnm{Gonzalez-Beltran}},
+\bauthor{\binits{P.}~\bsnm{Gormanns}},
+\bauthor{\binits{J.}~\bsnm{Grethe}},
+\bauthor{\binits{J.}~\bsnm{Hastings}},
+\bauthor{\binits{J.-K.}~\bsnm{H\'{e}rich\'{e}}},
+\bauthor{\binits{H.}~\bsnm{Hermjakob}},
+\bauthor{\binits{J.C.}~\bsnm{Ison}},
+\bauthor{\binits{R.C.}~\bsnm{Jimenez}},
+\bauthor{\binits{S.}~\bsnm{Jupp}},
+\bauthor{\binits{J.}~\bsnm{Kunze}},
+\bauthor{\binits{C.}~\bsnm{Laibe}},
+\bauthor{\binits{N.}~\bsnm{Le Nov\`{e}re}},
+\bauthor{\binits{J.}~\bsnm{Malone}},
+\bauthor{\binits{M.J.}~\bsnm{Martin}},
+\bauthor{\binits{J.R.}~\bsnm{McEntyre}},
+\bauthor{\binits{C.}~\bsnm{Morris}},
+\bauthor{\binits{J.}~\bsnm{Muilu}},
+\bauthor{\binits{W.}~\bsnm{M\"{u}ller}},
+\bauthor{\binits{P.}~\bsnm{Rocca-Serra}},
+\bauthor{\binits{S.-A.}~\bsnm{Sansone}},
+\bauthor{\binits{M.}~\bsnm{Sariyar}},
+\bauthor{\binits{J.L.}~\bsnm{Snoep}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\bauthor{\binits{N.J.}~\bsnm{Stanford}},
+\bauthor{\binits{N.}~\bsnm{Swainston}},
+\bauthor{\binits{N.}~\bsnm{Washington}},
+\bauthor{\binits{A.R.}~\bsnm{Williams}},
+\bauthor{\binits{S.M.}~\bsnm{Wimalaratne}},
+\bauthor{\binits{L.M.}~\bsnm{Winfree}},
+\bauthor{\binits{K.}~\bsnm{Wolstencroft}},
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{C.J.}~\bsnm{Mungall}},
+\bauthor{\binits{M.A.}~\bsnm{Haendel}} and
+\bauthor{\binits{H.}~\bsnm{Parkinson}},
+\batitle{Identifiers for the 21st century: How to design, provision, and
+reuse persistent identifiers to maximize utility and impact of life
+science data},
+\bjtitle{PLoS Biology}
+\bvolume{15}(\bissue{6})
+(\byear{2017}),
+\bnumber{e2001414}.
+\bid{doi={10.1371/journal.pbio.2001414}}
+\end{barticle}
+%
+\OrigBibText
+J.A.~{McMurry},
+N.~Juty,
+N.~Blomberg,
+T.~Burdett,
+T.~Conlin,
+N.~Conte,
+M.~Courtot,
+J.~Deck,
+M.~Dumontier,
+D.K.~Fellows,
+A.~Gonzalez-Beltran,
+P.~Gormanns,
+J.~Grethe,
+J.~Hastings,
+J.-K.~H\'{e}rich\'{e},
+H.~Hermjakob,
+J.C.~Ison,
+R.C.~Jimenez,
+S.~Jupp,
+J.~Kunze,
+C.~Laibe,
+N.~Le~Nov\`{e}re,
+J.~Malone,
+M.J.~Martin,
+J.R.~{McEntyre},
+C.~Morris,
+J.~Muilu,
+W.~M\"{u}ller,
+P.~Rocca-Serra,
+S.-A.~Sansone,
+M.~Sariyar,
+J.L.~Snoep,
+S.~Soiland-Reyes,
+N.J.~Stanford,
+N.~Swainston,
+N.~Washington,
+A.R.~Williams,
+S.M.~Wimalaratne,
+L.M.~Winfree,
+K.~Wolstencroft,
+C.~Goble,
+C.J.~Mungall,
+M.A.~Haendel and
+H.~Parkinson,
+\textit{Identifiers for the 21st century: How to design, provision, and reuse
+persistent identifiers to maximize utility and impact of life science data.},
+{PLoS} Biology
+15(6)
+(2017),
+e2001414.
+doi:10.1371/journal.pbio.2001414.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b87 ###bbsrt2
+%b57 ###
+\bibitem{doi:10.4126/frl01-006423291}
+%
+\begin{bchapter}
+\bauthor{\binits{T.}~\bsnm{Miksa}},
+\bauthor{\binits{M.}~\bsnm{Jaoua}} and
+\bauthor{\binits{G.}~\bsnm{Arfaoui}},
+\bctitle{Research object crates and machine-actionable data management plans},
+in: \bbtitle{1st Workshop on Research Data Management for Linked Open Science},
+\byear{2020}.
+\bid{doi={10.4126/frl01-006423291}}
+\end{bchapter}
+%
+\OrigBibText
+T.~Miksa,
+M.~Jaoua and
+G.~Arfaoui,
+\textit{Research Object Crates and Machine-actionable Data Management Plans},
+in: 1st Workshop on Research Data Management for Linked Open Science,
+2020.
+doi:10.4126/frl01-006423291.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b88 ###bbsrt2
+%b83 ###
+\bibitem{doi:10.1371/journal.pcbi.1006750}
+%
+\begin{barticle}
+\bauthor{\binits{T.}~\bsnm{Miksa}},
+\bauthor{\binits{S.}~\bsnm{Simms}},
+\bauthor{\binits{D.}~\bsnm{Mietchen}} and
+\bauthor{\binits{S.}~\bsnm{Jones}},
+\batitle{Ten principles for machine-actionable data management plans},
+\bjtitle{PLoS Computational Biology}
+\bvolume{15}(\bissue{3})
+(\byear{2019}),
+\bnumber{e1006750}.
+\bid{doi={10.1371/journal.pcbi.1006750}}
+\end{barticle}
+%
+\OrigBibText
+T.~Miksa,
+S.~Simms,
+D.~Mietchen and
+S.~Jones,
+\textit{Ten principles for machine-actionable data management plans},
+{PLoS} Computational Biology
+15(3)
+(2019),
+e1006750.
+doi:10.1371/journal.pcbi.1006750.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b89 ###bbsrt2
+%b121 ###
+\bibitem{doi:10.1186/1471-2105-11-S12-S5}
+%
+\begin{barticle}
+\bauthor{\binits{S.}~\bsnm{M\"{o}ller}},
+\bauthor{\binits{H.N.}~\bsnm{Krabbenh\"{o}ft}},
+\bauthor{\binits{A.}~\bsnm{Tille}},
+\bauthor{\binits{D.}~\bsnm{Paleino}},
+\bauthor{\binits{A.}~\bsnm{Williams}},
+\bauthor{\binits{K.}~\bsnm{Wolstencroft}},
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{R.}~\bsnm{Holland}},
+\bauthor{\binits{D.}~\bsnm{Belhachemi}} and
+\bauthor{\binits{C.}~\bsnm{Plessy}},
+\batitle{Community-driven computational biology with Debian Linux},
+\bjtitle{{BMC} Bioinformatics}
+\bvolume{11}(\bissue{Suppl 12})
+(\byear{2010}),
+\bnumber{S5}.
+\bid{doi={10.1186/1471-2105-11-S12-S5}}
+\end{barticle}
+%
+\OrigBibText
+S.~M\"{o}ller,
+H.N.~Krabbenh\"{o}ft,
+A.~Tille,
+D.~Paleino,
+A.~Williams,
+K.~Wolstencroft,
+C.~Goble,
+R.~Holland,
+D.~Belhachemi and
+C.~Plessy,
+\textit{Community-driven computational biology with Debian Linux.},
+{BMC} Bioinformatics
+11 Suppl 12
+(2010),
+S5.
+doi:10.1186/1471-2105-11-S12-S5.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b90 ###bbsrt2
+%b118 ###
+\bibitem{doi:10.1007/s41019-017-0050-4}
+%
+\begin{barticle}
+\bauthor{\binits{S.}~\bsnm{M\"{o}ller}},
+\bauthor{\binits{S.W.}~\bsnm{Prescott}},
+\bauthor{\binits{L.}~\bsnm{Wirzenius}},
+\bauthor{\binits{P.}~\bsnm{Reinholdtsen}},
+\bauthor{\binits{B.}~\bsnm{Chapman}},
+\bauthor{\binits{P.}~\bsnm{Prins}},
+\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\bauthor{\binits{F.}~\bsnm{Kl\"{o}tzl}},
+\bauthor{\binits{A.}~\bsnm{Bagnacani}},
+\bauthor{\binits{M.}~\bsnm{Kala\v{s}}},
+\bauthor{\binits{A.}~\bsnm{Tille}} and
+\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
+\batitle{Robust cross-platform workflows: How technical and scientific
+communities collaborate to develop, test and share best practices for
+data analysis},
+\bjtitle{Data Science and Engineering}
+\bvolume{2}(\bissue{3})
+(\byear{2017}),
+\bfpage{232}--\blpage{244}.
+\bid{doi={10.1007/s41019-017-0050-4}}
+\end{barticle}
+%
+\OrigBibText
+S.~M\"{o}ller,
+S.W.~Prescott,
+L.~Wirzenius,
+P.~Reinholdtsen,
+B.~Chapman,
+P.~Prins,
+S.~Soiland-Reyes,
+F.~Kl\"{o}tzl,
+A.~Bagnacani,
+M.~Kala\v{s},
+A.~Tille and
+M.R.~Crusoe,
+\textit{Robust Cross-Platform Workflows: How Technical and Scientific
+Communities Collaborate to Develop, Test and Share Best Practices for Data
+Analysis},
+Data Science and Engineering
+2(3)
+(2017),
+232--244.
+doi:10.1007/s41019-017-0050-4.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b91 ###bbsrt2
+%b6 ###
+\bibitem{isbn:9781315351148}
+%
+\begin{botherref}
+\oauthor{\binits{B.}~\bsnm{Mons}},
+Data Stewardship for Open Science,
+1st edn,
+Taylor \& Francis,
+p.~240.
+ISBN 9781315351148.
+\end{botherref}
+%
+\OrigBibText
+B.~Mons,
+\textit{Data Stewardship for Open Science},
+1st edn,
+Taylor \& Francis,
+p.~240.
+ISBN 9781315351148.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b92 ###bbsrt2
+%b100 ###
+\bibitem{myExperimentOntology2009}
+%
+\begin{botherref}
+{{myExperiment Ontology Modules}},
+2009.
+\url{http://web.archive.org/web/20091115080336/http\%3a\%2f\%2frdf.myexperiment.org/ontologies}.
+\end{botherref}
+%
+\OrigBibText
+{myExperiment} Ontology Modules,
+2009.
+\url{http://web.archive.org/web/20091115080336/}
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b93 ###bbsrt2
+%b99 ###
+\bibitem{newman2009}
+%
+\begin{bchapter}
+\bauthor{\binits{D.}~\bsnm{Newman}},
+\bauthor{\binits{S.}~\bsnm{Bechhofer}} and
+\bauthor{\binits{D.D.}~\bsnm{Roure}},
+\bctitle{{myExperiment}: An ontology for e-Research},
+in: \bbtitle{Proceedings of the Workshop on Semantic Web Applications
+in Scientific
+Discourse (SWASD 2009)},
+\beditor{\binits{T.}~\bsnm{Clark}},
+\beditor{\binits{J.S.}~\bsnm{Luciano}},
+\beditor{\binits{M.S.}~\bsnm{Marshall}},
+\beditor{\binits{E.}~\bsnm{Prud'Hommeaux}} and
+\beditor{\binits{S.}~\bsnm{Stephens}}, eds,
+\bsertitle{CEUR Workshop Proceedings},
+Vols~\bseriesno{523},
+\bpublisher{CEUR-WS},
+\byear{2009}.
+\bcomment{ISSN 1613-0073}. \url{http://ceur-ws.org/Vol-523/Newman.pdf}.
+\end{bchapter}
+%
+\OrigBibText
+D.~Newman,
+S.~Bechhofer and
+D.D.~Roure,
+\textit{{myExperiment}: An ontology for e-Research},
+in: Proceedings of the Workshop on Semantic Web Applications in
+Scientific Discourse (SWASD 2009),
+T.~Clark,
+J.S.~Luciano,
+M.S.~Marshall,
+E.~Prud'Hommeaux and
+S.~Stephens, eds,
+CEUR Workshop Proceedings,
+Vol.~523,
+CEUR-WS,
+2009,
+2009-10-25.
+ISSN 1613-0073.
+\url{http://ceur-ws.org/Vol-523/Newman.pdf}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b94 ###bbsrt2
+%b19 ###
+\bibitem{neylon_blog_post_2017}
+%
+\begin{botherref}
+\oauthor{\binits{C.}~\bsnm{Neylon}},
+As a researcher {\ldots} I'm a bit bloody fed up with Data Management,
+2017.
+\url
+{https://cameronneylon.net/blog/as-a-researcher-im-a-bit-bloody-fed-up-with-data-management/}.
+\end{botherref}
+%
+\OrigBibText
+C.~Neylon,
+As a researcher{\ldots}I'm a bit bloody fed up with Data Management,
+2017.
+\url
+{https://cameronneylon.net/blog/as-a-researcher-im-a-bit-bloody-fed-up-with-data-management/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b95 ###bbsrt2
+% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
+%incorrectly.
+%b38 ###
+\bibitem{ro-crate-html-js}
+%
+\begin{botherref}
+npm: ro-crate-html-js,
+\url{https://www.npmjs.com/package/ro-crate-html-js}.
+\end{botherref}
+%
+\OrigBibText
+npm: ro-crate-html-js.
+\url{https://www.npmjs.com/package/ro-crate-html-js}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b96 ###bbsrt2
+%b27 ###
+\bibitem{ocfl_2020}
+%
+\begin{botherref}
+{OCFL},
+Oxford Common File Layout Specification, Recommendation,
+2020.
+\url{https://ocfl.io/1.0/spec/}.
+\end{botherref}
+%
+\OrigBibText
+{OCFL},
+Oxford Common File Layout Specification,
+Recommendation,
+2020.
+\url{https://ocfl.io/1.0/spec/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b97 ###bbsrt2
+%b61 ###
+\bibitem{doi:10.1080/14490854.2020.1796500}
+%
+\begin{barticle}
+\bauthor{\binits{A.}~\bsnm{Piper}},
+\batitle{Digital crowdsourcing and public understandings of the past:
+Citizen historians meet criminal characters},
+\bjtitle{History Australia}
+\bvolume{17}(\bissue{3})
+(\byear{2020}),
+\bfpage{525}--\blpage{541}.
+\bid{doi={10.1080/14490854.2020.1796500}}
+\end{barticle}
+%
+\OrigBibText
+A.~Piper,
+\textit{Digital crowdsourcing and public understandings of the past: citizen
+historians meet Criminal Characters},
+History Australia
+17(3)
+(2020),
+525--541.
+doi:10.1080/14490854.2020.1796500.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b98 ###bbsrt2
+%b127 ###
+\bibitem{rdfworkinggroup_2014}
+%
+\begin{botherref}
+\binstitute{RDF Working Group},
+{RDF} 1.1 Concepts and Abstract Syntax, {W3C} Recommendation,
+2014.
+\url{https://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/}.
+\end{botherref}
+%
+\OrigBibText
+{RDF Working Group},
+{RDF} 1.1 Concepts and Abstract Syntax,
+{W3C} Recommendation,
+2014.
+\url{https://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b99 ###bbsrt2
+%b41 ###
+\bibitem{doi:10.1016/j.xgen.2021.100029}
+%
+\begin{barticle}
+\bauthor{\binits{H.L.}~\bsnm{Rehm}},
+\bauthor{\binits{A.J.H.}~\bsnm{Page}},
+\bauthor{\binits{L.}~\bsnm{Smith}},
+\bauthor{\binits{J.B.}~\bsnm{Adams}},
+\bauthor{\binits{G.}~\bsnm{Alterovitz}},
+\bauthor{\binits{L.J.}~\bsnm{Babb}},
+\bauthor{\binits{M.P.}~\bsnm{Barkley}},
+\bauthor{\binits{M.}~\bsnm{Baudis}},
+\bauthor{\binits{M.J.S.}~\bsnm{Beauvais}},
+\bauthor{\binits{T.}~\bsnm{Beck}},
+\bauthor{\binits{J.S.}~\bsnm{Beckmann}},
+\bauthor{\binits{S.}~\bsnm{Beltran}},
+\bauthor{\binits{D.}~\bsnm{Bernick}},
+\bauthor{\binits{A.}~\bsnm{Bernier}},
+\bauthor{\binits{J.K.}~\bsnm{Bonfield}},
+\bauthor{\binits{T.F.}~\bsnm{Boughtwood}},
+\bauthor{\binits{G.}~\bsnm{Bourque}},
+\bauthor{\binits{S.R.}~\bsnm{Bowers}},
+\bauthor{\binits{A.J.}~\bsnm{Brookes}},
+\bauthor{\binits{M.}~\bsnm{Brudno}},
+\bauthor{\binits{M.H.}~\bsnm{Brush}},
+\bauthor{\binits{D.}~\bsnm{Bujold}},
+\bauthor{\binits{T.}~\bsnm{Burdett}},
+\bauthor{\binits{O.J.}~\bsnm{Buske}},
+\bauthor{\binits{M.N.}~\bsnm{Cabili}},
+\bauthor{\binits{D.L.}~\bsnm{Cameron}},
+\bauthor{\binits{R.J.}~\bsnm{Carroll}},
+\bauthor{\binits{E.}~\bsnm{Casas-Silva}},
+\bauthor{\binits{D.}~\bsnm{Chakravarty}},
+\bauthor{\binits{B.P.}~\bsnm{Chaudhari}},
+\bauthor{\binits{S.H.}~\bsnm{Chen}},
+\bauthor{\binits{J.M.}~\bsnm{Cherry}},
+\bauthor{\binits{J.}~\bsnm{Chung}},
+\bauthor{\binits{M.}~\bsnm{Cline}},
+\bauthor{\binits{H.L.}~\bsnm{Clissold}},
+\bauthor{\binits{R.M.}~\bsnm{Cook-Deegan}},
+\bauthor{\binits{M.}~\bsnm{Courtot}},
+\bauthor{\binits{F.}~\bsnm{Cunningham}},
+\bauthor{\binits{M.}~\bsnm{Cupak}},
+\bauthor{\binits{R.M.}~\bsnm{Davies}},
+\bauthor{\binits{D.}~\bsnm{Denisko}},
+\bauthor{\binits{M.J.}~\bsnm{Doerr}},
+\bauthor{\binits{L.I.}~\bsnm{Dolman}},
+\bauthor{\binits{E.S.}~\bsnm{Dove}},
+\bauthor{\binits{L.J.}~\bsnm{Dursi}},
+\bauthor{\binits{S.O.M.}~\bsnm{Dyke}},
+\bauthor{\binits{J.A.}~\bsnm{Eddy}},
+\bauthor{\binits{K.}~\bsnm{Eilbeck}},
+\bauthor{\binits{K.P.}~\bsnm{Ellrott}},
+\bauthor{\binits{S.}~\bsnm{Fairley}},
+\bauthor{\binits{K.A.}~\bsnm{Fakhro}},
+\bauthor{\binits{H.V.}~\bsnm{Firth}},
+\bauthor{\binits{M.S.}~\bsnm{Fitzsimons}},
+\bauthor{\binits{M.}~\bsnm{Fiume}},
+\bauthor{\binits{P.}~\bsnm{Flicek}},
+\bauthor{\binits{I.M.}~\bsnm{Fore}},
+\bauthor{\binits{M.A.}~\bsnm{Freeberg}},
+\bauthor{\binits{R.R.}~\bsnm{Freimuth}},
+\bauthor{\binits{L.A.}~\bsnm{Fromont}},
+\bauthor{\binits{J.}~\bsnm{Fuerth}},
+\bauthor{\binits{C.L.}~\bsnm{Gaff}},
+\bauthor{\binits{W.}~\bsnm{Gan}},
+\bauthor{\binits{E.M.}~\bsnm{Ghanaim}},
+\bauthor{\binits{D.}~\bsnm{Glazer}},
+\bauthor{\binits{R.C.}~\bsnm{Green}},
+\bauthor{\binits{M.}~\bsnm{Griffith}},
+\bauthor{\binits{O.L.}~\bsnm{Griffith}},
+\bauthor{\binits{R.L.}~\bsnm{Grossman}},
+\bauthor{\binits{T.}~\bsnm{Groza}},
+\bauthor{\binits{J.M.}~\bsnm{Guidry Auvil}},
+\bauthor{\binits{R.}~\bsnm{Guig\'{o}}},
+\bauthor{\binits{D.}~\bsnm{Gupta}},
+\bauthor{\binits{M.A.}~\bsnm{Haendel}},
+\bauthor{\binits{A.}~\bsnm{Hamosh}},
+\bauthor{\binits{D.P.}~\bsnm{Hansen}},
+\bauthor{\binits{R.K.}~\bsnm{Hart}},
+\bauthor{\binits{D.M.}~\bsnm{Hartley}},
+\bauthor{\binits{D.}~\bsnm{Haussler}},
+\bauthor{\binits{R.M.}~\bsnm{Hendricks-Sturrup}},
+\bauthor{\binits{C.W.L.}~\bsnm{Ho}},
+\bauthor{\binits{A.E.}~\bsnm{Hobb}},
+\bauthor{\binits{M.M.}~\bsnm{Hoffman}},
+\bauthor{\binits{O.M.}~\bsnm{Hofmann}},
+\bauthor{\binits{P.}~\bsnm{Holub}},
+\bauthor{\binits{J.S.}~\bsnm{Hsu}},
+\bauthor{\binits{J.-P.}~\bsnm{Hubaux}},
+\bauthor{\binits{S.E.}~\bsnm{Hunt}},
+\bauthor{\binits{A.}~\bsnm{Husami}},
+\bauthor{\binits{J.O.}~\bsnm{Jacobsen}},
+\bauthor{\binits{S.S.}~\bsnm{Jamuar}},
+\bauthor{\binits{E.L.}~\bsnm{Janes}},
+\bauthor{\binits{F.}~\bsnm{Jeanson}},
+\bauthor{\binits{A.}~\bsnm{Jen\'{e}}},
+\bauthor{\binits{A.L.}~\bsnm{Johns}},
+\bauthor{\binits{Y.}~\bsnm{Joly}},
+\bauthor{\binits{S.J.M.}~\bsnm{Jones}},
+\bauthor{\binits{A.}~\bsnm{Kanitz}},
+\bauthor{\binits{K.}~\bsnm{Kato}},
+\bauthor{\binits{T.M.}~\bsnm{Keane}},
+\bauthor{\binits{K.}~\bsnm{Kekesi-Lafrance}},
+\bauthor{\binits{J.}~\bsnm{Kelleher}},
+\bauthor{\binits{G.}~\bsnm{Kerry}},
+\bauthor{\binits{S.-S.}~\bsnm{Khor}},
+\bauthor{\binits{B.M.}~\bsnm{Knoppers}},
+\bauthor{\binits{M.A.}~\bsnm{Konopko}},
+\bauthor{\binits{K.}~\bsnm{Kosaki}},
+\bauthor{\binits{M.}~\bsnm{Kuba}},
+\bauthor{\binits{J.}~\bsnm{Lawson}},
+\bauthor{\binits{R.}~\bsnm{Leinonen}},
+\bauthor{\binits{S.}~\bsnm{Li}},
+\bauthor{\binits{M.F.}~\bsnm{Lin}},
+\bauthor{\binits{M.}~\bsnm{Linden}},
+\bauthor{\binits{X.}~\bsnm{Liu}},
+\bauthor{\binits{I.U.}~\bsnm{Liyanage}},
+\bauthor{\binits{J.}~\bsnm{Lopez}},
+\bauthor{\binits{A.M.}~\bsnm{Lucassen}},
+\bauthor{\binits{M.}~\bsnm{Lukowski}},
+\bauthor{\binits{A.L.}~\bsnm{Mann}},
+\bauthor{\binits{J.}~\bsnm{Marshall}},
+\bauthor{\binits{M.}~\bsnm{Mattioni}},
+\bauthor{\binits{A.}~\bsnm{Metke-Jimenez}},
+\bauthor{\binits{A.}~\bsnm{Middleton}},
+\bauthor{\binits{R.J.}~\bsnm{Milne}},
+\bauthor{\binits{F.}~\bsnm{Moln\'{a}r-G\'{a}bor}},
+\bauthor{\binits{N.}~\bsnm{Mulder}},
+\bauthor{\binits{M.C.}~\bsnm{Munoz-Torres}},
+\bauthor{\binits{R.}~\bsnm{Nag}},
+\bauthor{\binits{H.}~\bsnm{Nakagawa}},
+\bauthor{\binits{J.}~\bsnm{Nasir}},
+\bauthor{\binits{A.}~\bsnm{Navarro}},
+\bauthor{\binits{T.H.}~\bsnm{Nelson}},
+\bauthor{\binits{A.}~\bsnm{Niewielska}},
+\bauthor{\binits{A.}~\bsnm{Nisselle}},
+\bauthor{\binits{J.}~\bsnm{Niu}},
+\bauthor{\binits{T.H.}~\bsnm{Nyr\"{o}nen}},
+\bauthor{\binits{B.D.}~\bsnm{O'Connor}},
+\bauthor{\binits{S.}~\bsnm{Oesterle}},
+\bauthor{\binits{S.}~\bsnm{Ogishima}},
+\bauthor{\binits{V.}~\bsnm{Ota Wang}},
+\bauthor{\binits{L.A.D.}~\bsnm{Paglione}},
+\bauthor{\binits{E.}~\bsnm{Palumbo}},
+\bauthor{\binits{H.E.}~\bsnm{Parkinson}},
+\bauthor{\binits{A.A.}~\bsnm{Philippakis}},
+\bauthor{\binits{A.D.}~\bsnm{Pizarro}},
+\bauthor{\binits{A.}~\bsnm{Prlic}},
+\bauthor{\binits{J.}~\bsnm{Rambla}},
+\bauthor{\binits{A.}~\bsnm{Rendon}},
+\bauthor{\binits{R.A.}~\bsnm{Rider}},
+\bauthor{\binits{P.N.}~\bsnm{Robinson}},
+\bauthor{\binits{K.W.}~\bsnm{Rodarmer}},
+\bauthor{\binits{L.L.}~\bsnm{Rodriguez}},
+\bauthor{\binits{A.F.}~\bsnm{Rubin}},
+\bauthor{\binits{M.}~\bsnm{Rueda}},
+\bauthor{\binits{G.A.}~\bsnm{Rushton}},
+\bauthor{\binits{R.S.}~\bsnm{Ryan}},
+\bauthor{\binits{G.I.}~\bsnm{Saunders}},
+\bauthor{\binits{H.}~\bsnm{Schuilenburg}},
+\bauthor{\binits{T.}~\bsnm{Schwede}},
+\bauthor{\binits{S.}~\bsnm{Scollen}},
+\bauthor{\binits{A.}~\bsnm{Senf}},
+\bauthor{\binits{N.C.}~\bsnm{Sheffield}},
+\bauthor{\binits{N.}~\bsnm{Skantharajah}},
+\bauthor{\binits{A.V.}~\bsnm{Smith}},
+\bauthor{\binits{H.J.}~\bsnm{Sofia}},
+\bauthor{\binits{D.}~\bsnm{Spalding}},
+\bauthor{\binits{A.B.}~\bsnm{Spurdle}},
+\bauthor{\binits{Z.}~\bsnm{Stark}},
+\bauthor{\binits{L.D.}~\bsnm{Stein}},
+\bauthor{\binits{M.}~\bsnm{Suematsu}},
+\bauthor{\binits{P.}~\bsnm{Tan}},
+\bauthor{\binits{J.A.}~\bsnm{Tedds}},
+\bauthor{\binits{A.A.}~\bsnm{Thomson}},
+\bauthor{\binits{A.}~\bsnm{Thorogood}},
+\bauthor{\binits{T.L.}~\bsnm{Tickle}},
+\bauthor{\binits{K.}~\bsnm{Tokunaga}},
+\bauthor{\binits{J.}~\bsnm{T\"{o}rnroos}},
+\bauthor{\binits{D.}~\bsnm{Torrents}},
+\bauthor{\binits{S.}~\bsnm{Upchurch}},
+\bauthor{\binits{A.}~\bsnm{Valencia}},
+\bauthor{\binits{R.V.}~\bsnm{Guimera}},
+\bauthor{\binits{J.}~\bsnm{Vamathevan}},
+\bauthor{\binits{S.}~\bsnm{Varma}},
+\bauthor{\binits{D.F.}~\bsnm{Vears}},
+\bauthor{\binits{C.}~\bsnm{Viner}},
+\bauthor{\binits{C.}~\bsnm{Voisin}},
+\bauthor{\binits{A.H.}~\bsnm{Wagner}},
+\bauthor{\binits{S.E.}~\bsnm{Wallace}},
+\bauthor{\binits{B.P.}~\bsnm{Walsh}},
+\bauthor{\binits{M.S.}~\bsnm{Williams}},
+\bauthor{\binits{E.C.}~\bsnm{Winkler}},
+\bauthor{\binits{B.J.}~\bsnm{Wold}},
+\bauthor{\binits{G.M.}~\bsnm{Wood}},
+\bauthor{\binits{J.P.}~\bsnm{Woolley}},
+\bauthor{\binits{C.}~\bsnm{Yamasaki}},
+\bauthor{\binits{A.D.}~\bsnm{Yates}},
+\bauthor{\binits{C.K.}~\bsnm{Yung}},
+\bauthor{\binits{L.J.}~\bsnm{Zass}},
+\bauthor{\binits{K.}~\bsnm{Zaytseva}},
+\bauthor{\binits{J.}~\bsnm{Zhang}},
+\bauthor{\binits{P.}~\bsnm{Goodhand}},
+\bauthor{\binits{K.}~\bsnm{North}} and
+\bauthor{\binits{E.}~\bsnm{Birney}},
+\batitle{{GA4GH}: International policies and standards for data sharing
+across genomic research and healthcare},
+\bjtitle{Cell Genomics}
+\bvolume{1}(\bissue{2})
+(\byear{2021}),
+\bnumber{100029}.
+\bid{doi={10.1016/j.xgen.2021.100029}}
+\end{barticle}
+%
+\OrigBibText
+H.L.~Rehm,
+A.J.H.~Page,
+L.~Smith,
+J.B.~Adams,
+G.~Alterovitz,
+L.J.~Babb,
+M.P.~Barkley,
+M.~Baudis,
+M.J.S.~Beauvais,
+T.~Beck,
+J.S.~Beckmann,
+S.~Beltran,
+D.~Bernick,
+A.~Bernier,
+J.K.~Bonfield,
+T.F.~Boughtwood,
+G.~Bourque,
+S.R.~Bowers,
+A.J.~Brookes,
+M.~Brudno,
+M.H.~Brush,
+D.~Bujold,
+T.~Burdett,
+O.J.~Buske,
+M.N.~Cabili,
+D.L.~Cameron,
+R.J.~Carroll,
+E.~Casas-Silva,
+D.~Chakravarty,
+B.P.~Chaudhari,
+S.H.~Chen,
+J.M.~Cherry,
+J.~Chung,
+M.~Cline,
+H.L.~Clissold,
+R.M.~Cook-Deegan,
+M.~Courtot,
+F.~Cunningham,
+M.~Cupak,
+R.M.~Davies,
+D.~Denisko,
+M.J.~Doerr,
+L.I.~Dolman,
+E.S.~Dove,
+L.J.~Dursi,
+S.O.M.~Dyke,
+J.A.~Eddy,
+K.~Eilbeck,
+K.P.~Ellrott,
+S.~Fairley,
+K.A.~Fakhro,
+H.V.~Firth,
+M.S.~Fitzsimons,
+M.~Fiume,
+P.~Flicek,
+I.M.~Fore,
+M.A.~Freeberg,
+R.R.~Freimuth,
+L.A.~Fromont,
+J.~Fuerth,
+C.L.~Gaff,
+W.~Gan,
+E.M.~Ghanaim,
+D.~Glazer,
+R.C.~Green,
+M.~Griffith,
+O.L.~Griffith,
+R.L.~Grossman,
+T.~Groza,
+J.M.~Guidry~Auvil,
+R.~Guig\'{o},
+D.~Gupta,
+M.A.~Haendel,
+A.~Hamosh,
+D.P.~Hansen,
+R.K.~Hart,
+D.M.~Hartley,
+D.~Haussler,
+R.M.~Hendricks-Sturrup,
+C.W.L.~Ho,
+A.E.~Hobb,
+M.M.~Hoffman,
+O.M.~Hofmann,
+P.~Holub,
+J.S.~Hsu,
+J.-P.~Hubaux,
+S.E.~Hunt,
+A.~Husami,
+J.O.~Jacobsen,
+S.S.~Jamuar,
+E.L.~Janes,
+F.~Jeanson,
+A.~Jen\'{e},
+A.L.~Johns,
+Y.~Joly,
+S.J.M.~Jones,
+A.~Kanitz,
+K.~Kato,
+T.M.~Keane,
+K.~Kekesi-Lafrance,
+J.~Kelleher,
+G.~Kerry,
+S.-S.~Khor,
+B.M.~Knoppers,
+M.A.~Konopko,
+K.~Kosaki,
+M.~Kuba,
+J.~Lawson,
+R.~Leinonen,
+S.~Li,
+M.F.~Lin,
+M.~Linden,
+X.~Liu,
+I.U.~Liyanage,
+J.~Lopez,
+A.M.~Lucassen,
+M.~Lukowski,
+A.L.~Mann,
+J.~Marshall,
+M.~Mattioni,
+A.~Metke-Jimenez,
+A.~Middleton,
+R.J.~Milne,
+F.~Moln\'{a}r-G\'{a}bor,
+N.~Mulder,
+M.C.~Munoz-Torres,
+R.~Nag,
+H.~Nakagawa,
+J.~Nasir,
+A.~Navarro,
+T.H.~Nelson,
+A.~Niewielska,
+A.~Nisselle,
+J.~Niu,
+T.H.~Nyr\"{o}nen,
+B.D.~{O\textquoteright Connor},
+S.~Oesterle,
+S.~Ogishima,
+V.~Ota~Wang,
+L.A.D.~Paglione,
+E.~Palumbo,
+H.E.~Parkinson,
+A.A.~Philippakis,
+A.D.~Pizarro,
+A.~Prlic,
+J.~Rambla,
+A.~Rendon,
+R.A.~Rider,
+P.N.~Robinson,
+K.W.~Rodarmer,
+L.L.~Rodriguez,
+A.F.~Rubin,
+M.~Rueda,
+G.A.~Rushton,
+R.S.~Ryan,
+G.I.~Saunders,
+H.~Schuilenburg,
+T.~Schwede,
+S.~Scollen,
+A.~Senf,
+N.C.~Sheffield,
+N.~Skantharajah,
+A.V.~Smith,
+H.J.~Sofia,
+D.~Spalding,
+A.B.~Spurdle,
+Z.~Stark,
+L.D.~Stein,
+M.~Suematsu,
+P.~Tan,
+J.A.~Tedds,
+A.A.~Thomson,
+A.~Thorogood,
+T.L.~Tickle,
+K.~Tokunaga,
+J.~T\"{o}rnroos,
+D.~Torrents,
+S.~Upchurch,
+A.~Valencia,
+R.V.~Guimera,
+J.~Vamathevan,
+S.~Varma,
+D.F.~Vears,
+C.~Viner,
+C.~Voisin,
+A.H.~Wagner,
+S.E.~Wallace,
+B.P.~Walsh,
+M.S.~Williams,
+E.C.~Winkler,
+B.J.~Wold,
+G.M.~Wood,
+J.P.~Woolley,
+C.~Yamasaki,
+A.D.~Yates,
+C.K.~Yung,
+L.J.~Zass,
+K.~Zaytseva,
+J.~Zhang,
+P.~Goodhand,
+K.~North and
+E.~Birney,
+\textit{{GA4GH}: International policies and standards for data sharing across
+genomic research and healthcare},
+Cell Genomics
+1(2)
+(2021),
+100029.
+doi:10.1016/j.xgen.2021.100029.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b100 ###bbsrt2
+%b42 ###
+\bibitem{rettberg_2015_openaire}
+%
+\begin{barticle}
+\bauthor{\binits{N.}~\bsnm{Rettberg}} and
+\bauthor{\binits{B.}~\bsnm{Schmidt}},
+\bjtitle{OpenAIRE, College \& Research Libraries News}
+\bvolume{76}(\bissue{6})
+(\byear{2015}),
+\bfpage{306}--\blpage{310}.
+\url{http://resolver.sub.uni-goettingen.de/purl?gs-1/11942}.
+\bid{doi={10.5860/crln.76.6.9326}}
+\end{barticle}
+%
+\OrigBibText
+N.~Rettberg and
+B.~Schmidt,
+\textit{OpenAIRE},
+College \& Research Libraries News
+76(6)
+(2015),
+306--310.
+\url{http://resolver.sub.uni-goettingen.de/purl?gs-1/11942}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b101 ###bbsrt2
+%b9 ###
+\bibitem{doi:10.1371/journal.pcbi.1003285}
+%
+\begin{barticle}
+\bauthor{\binits{G.K.}~\bsnm{Sandve}},
+\bauthor{\binits{A.}~\bsnm{Nekrutenko}},
+\bauthor{\binits{J.}~\bsnm{Taylor}} and
+\bauthor{\binits{E.}~\bsnm{Hovig}},
+\batitle{Ten simple rules for reproducible computational research},
+\bjtitle{PLoS Computational Biology}
+\bvolume{9}(\bissue{10})
+(\byear{2013}),
+\bnumber{e1003285}.
+\bid{doi={10.1371/journal.pcbi.1003285}}
+\end{barticle}
+%
+\OrigBibText
+G.K.~Sandve,
+A.~Nekrutenko,
+J.~Taylor and
+E.~Hovig,
+\textit{Ten simple rules for reproducible computational research.},
+{PLoS} Computational Biology
+9(10)
+(2013),
+e1003285.
+doi:10.1371/journal.pcbi.1003285.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b102 ###bbsrt2
+%b21 ###
+\bibitem{doi:10.1038/s41597-020-0524-5}
+%
+\begin{barticle}
+\bauthor{\binits{L.M.}~\bsnm{Schriml}},
+\bauthor{\binits{M.}~\bsnm{Chuvochina}},
+\bauthor{\binits{N.}~\bsnm{Davies}},
+\bauthor{\binits{E.A.}~\bsnm{Eloe-Fadrosh}},
+\bauthor{\binits{R.D.}~\bsnm{Finn}},
+\bauthor{\binits{P.}~\bsnm{Hugenholtz}},
+\bauthor{\binits{C.I.}~\bsnm{Hunter}},
+\bauthor{\binits{B.L.}~\bsnm{Hurwitz}},
+\bauthor{\binits{N.C.}~\bsnm{Kyrpides}},
+\bauthor{\binits{F.}~\bsnm{Meyer}},
+\bauthor{\binits{I.K.}~\bsnm{Mizrachi}},
+\bauthor{\binits{S.-A.}~\bsnm{Sansone}},
+\bauthor{\binits{G.}~\bsnm{Sutton}},
+\bauthor{\binits{S.}~\bsnm{Tighe}} and
+\bauthor{\binits{R.}~\bsnm{Walls}},
+\batitle{{COVID}-19 pandemic reveals the peril of ignoring metadata standards},
+\bjtitle{Scientific Data}
+\bvolume{7}(\bissue{1})
+(\byear{2020}),
+\bnumber{188}.
+\bid{doi={10.1038/s41597-020-0524-5}}
+\end{barticle}
+%
+\OrigBibText
+L.M.~Schriml,
+M.~Chuvochina,
+N.~Davies,
+E.A.~Eloe-Fadrosh,
+R.D.~Finn,
+P.~Hugenholtz,
+C.I.~Hunter,
+B.L.~Hurwitz,
+N.C.~Kyrpides,
+F.~Meyer,
+I.K.~Mizrachi,
+S.-A.~Sansone,
+G.~Sutton,
+S.~Tighe and
+R.~Walls,
+\textit{{COVID}-19 pandemic reveals the peril of ignoring metadata
+standards.},
+Scientific data
+7(1)
+(2020),
+188.
+doi:10.1038/s41597-020-0524-5.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b103 ###bbsrt2
+%b39 ###
+\bibitem{doi:10.5281/zenodo.1445817}
+%
+\begin{bchapter}
+\bauthor{\binits{P.}~\bsnm{Sefton}},
+in: \bbtitle{Datacrate Submisssion to the Workshop on Research Objects},
+\bconflocation{Zenodo},
+\byear{2018}.
+\bid{doi={10.5281/zenodo.1445817}}
+\end{bchapter}
+%
+\OrigBibText
+P.~Sefton,
+\textit{Datacrate Submisssion To The Workshop On Research Objects},
+Zenodo
+(2018).
+doi:10.5281/zenodo.1445817.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b104 ###bbsrt2
+%b1 ###
+\bibitem{sefton_blog_post_2021}
+%
+\begin{botherref}
+\oauthor{\binits{P.}~\bsnm{Sefton}},
+{FAIR} Data Management; It's a lifestyle not a lifecycle -- \surl{ptsefton.com},
+2021.
+\url{http://ptsefton.com/2021/04/07/rdmpic/}.
+\end{botherref}
+%
+\OrigBibText
+P.~Sefton,
+{FAIR} Data Management; It's a lifestyle not a lifecycle - ptsefton.com,
+2021.
+\url{http://ptsefton.com/2021/04/07/rdmpic/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b105 ###bbsrt2
+%b62 ###
+\bibitem{doi:10.5281/zenodo.3541888}
+%
+\begin{botherref}
+\oauthor{\binits{P.}~\bsnm{Sefton}},
+\oauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\oauthor{\binits{O.}~\bsnm{Corcho}},
+\oauthor{\binits{D.}~\bsnm{Garijo}},
+\oauthor{\binits{R.}~\bsnm{Palma}},
+\oauthor{\binits{F.}~\bsnm{Coppens}},
+\oauthor{\binits{C.}~\bsnm{Goble}},
+\oauthor{\binits{J.M.}~\bsnm{Fern\'{a}ndez}},
+\oauthor{\binits{K.}~\bsnm{Chard}},
+\oauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
+\oauthor{\binits{M.R.}~\bsnm{Crusoe}},
+\oauthor{\binits{I.}~\bsnm{Eguinoa}},
+\oauthor{\binits{N.}~\bsnm{Juty}},
+\oauthor{\binits{K.}~\bsnm{Holmes}},
+\oauthor{\binits{J.A.}~\bsnm{Clark}},
+\oauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
+\oauthor{\binits{A.J.G.}~\bsnm{Gray}},
+\oauthor{\binits{S.}~\bsnm{Owen}},
+\oauthor{\binits{A.R.}~\bsnm{Williams}},
+\oauthor{\binits{G.}~\bsnm{Tartari}},
+\oauthor{\binits{F.}~\bsnm{Bacall}} and
+\oauthor{\binits{T.}~\bsnm{Thelen}},
+{RO}-Crate Metadata Specification 1.0,
+Technical Report,
+2019.
+\bid{doi={10.5281/zenodo.3541888}}
+\end{botherref}
+%
+\OrigBibText
+P.~Sefton,
+E.{\'{O}}.~Carrag\'{a}in,
+S.~Soiland-Reyes,
+O.~Corcho,
+D.~Garijo,
+R.~Palma,
+F.~Coppens,
+C.~Goble,
+J.M.~Fern\'{a}ndez,
+K.~Chard,
+J.M.~Gomez-Perez,
+M.R.~Crusoe,
+I.~Eguinoa,
+N.~Juty,
+K.~Holmes,
+J.A.~Clark,
+S.~Capella-Gutierrez,
+A.J.G.~Gray,
+S.~Owen,
+A.R.~Williams,
+G.~Tartari,
+F.~Bacall and
+T.~Thelen,
+{RO}-Crate Metadata Specification 1.0,
+Technical Report,
+2019.
+doi:10.5281/zenodo.3541888.
+\url{https://w3id.org/ro/crate/1.0}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b106 ###bbsrt2
+%b33 ###
+\bibitem{doi:10.5281/zenodo.4541002}
+%
+\begin{botherref}
+\oauthor{\binits{P.}~\bsnm{Sefton}},
+\oauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\oauthor{\binits{O.}~\bsnm{Corcho}},
+\oauthor{\binits{D.}~\bsnm{Garijo}},
+\oauthor{\binits{R.}~\bsnm{Palma}},
+\oauthor{\binits{F.}~\bsnm{Coppens}},
+\oauthor{\binits{C.}~\bsnm{Goble}},
+\oauthor{\binits{J.M.}~\bsnm{Fern\'{a}ndez}},
+\oauthor{\binits{K.}~\bsnm{Chard}},
+\oauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
+\oauthor{\binits{M.R.}~\bsnm{Crusoe}},
+\oauthor{\binits{I.}~\bsnm{Eguinoa}},
+\oauthor{\binits{N.}~\bsnm{Juty}},
+\oauthor{\binits{K.}~\bsnm{Holmes}},
+\oauthor{\binits{J.A.}~\bsnm{Clark}},
+\oauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
+\oauthor{\binits{A.J.G.}~\bsnm{Gray}},
+\oauthor{\binits{S.}~\bsnm{Owen}},
+\oauthor{\binits{A.R.}~\bsnm{Williams}},
+\oauthor{\binits{G.}~\bsnm{Tartari}},
+\oauthor{\binits{F.}~\bsnm{Bacall}},
+\oauthor{\binits{T.}~\bsnm{Thelen}},
+\oauthor{\binits{H.}~\bsnm{M\'{e}nager}},
+\oauthor{\binits{L.R.-N.}~\bsnm{Navas}},
+\oauthor{\binits{P.}~\bsnm{Walk}},
+\oauthor{\binits{B.}~\bsnm{Whitehead}},
+\oauthor{\binits{M.}~\bsnm{Wilkinson}},
+\oauthor{\binits{P.}~\bsnm{Groth}},
+\oauthor{\binits{E.}~\bsnm{Bremer}},
+\oauthor{\binits{L.G.}~\bsnm{Castro}},
+\oauthor{\binits{K.}~\bsnm{Sebby}},
+\oauthor{\binits{A.}~\bsnm{Kanitz}},
+\oauthor{\binits{A.}~\bsnm{Trisovic}},
+\oauthor{\binits{G.}~\bsnm{Kennedy}},
+\oauthor{\binits{M.}~\bsnm{Graves}},
+\oauthor{\binits{J.}~\bsnm{Koehorst}},
+\oauthor{\binits{S.}~\bsnm{Leo}} and
+\oauthor{\binits{M.}~\bsnm{Portier}},
+{RO}-Crate Metadata Specification 1.1.1,
+Technical Report,
+2021.
+\bid{doi={10.5281/zenodo.4541002}}
+\end{botherref}
+%
+\OrigBibText
+P.~Sefton,
+E.{\'{O}}.~Carrag\'{a}in,
+S.~Soiland-Reyes,
+O.~Corcho,
+D.~Garijo,
+R.~Palma,
+F.~Coppens,
+C.~Goble,
+J.M.~Fern\'{a}ndez,
+K.~Chard,
+J.M.~Gomez-Perez,
+M.R.~Crusoe,
+I.~Eguinoa,
+N.~Juty,
+K.~Holmes,
+J.A.~Clark,
+S.~Capella-Gutierrez,
+A.J.G.~Gray,
+S.~Owen,
+A.R.~Williams,
+G.~Tartari,
+F.~Bacall,
+T.~Thelen,
+H.~M\'{e}nager,
+L.R.-N.~Navas,
+P.~Walk,
+B.~Whitehead,
+M.~Wilkinson,
+P.~Groth,
+E.~Bremer,
+L.G.~Castro,
+K.~Sebby,
+A.~Kanitz,
+A.~Trisovic,
+G.~Kennedy,
+M.~Graves,
+J.~Koehorst,
+S.~Leo and
+M.~Portier,
+{RO}-Crate Metadata Specification 1.1.1,
+Technical Report,
+2021.
+doi:10.5281/zenodo.4541002.
+\url{https://w3id.org/ro/crate/1.1}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b107 ###bbsrt2
+%b63 ###
+\bibitem{doi:10.5281/zenodo.4031327}
+%
+\begin{botherref}
+\oauthor{\binits{P.}~\bsnm{Sefton}},
+\oauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\oauthor{\binits{O.}~\bsnm{Corcho}},
+\oauthor{\binits{D.}~\bsnm{Garijo}},
+\oauthor{\binits{R.}~\bsnm{Palma}},
+\oauthor{\binits{F.}~\bsnm{Coppens}},
+\oauthor{\binits{C.}~\bsnm{Goble}},
+\oauthor{\binits{J.M.}~\bsnm{Fern\'{a}ndez}},
+\oauthor{\binits{K.}~\bsnm{Chard}},
+\oauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
+\oauthor{\binits{M.R.}~\bsnm{Crusoe}},
+\oauthor{\binits{I.}~\bsnm{Eguinoa}},
+\oauthor{\binits{N.}~\bsnm{Juty}},
+\oauthor{\binits{K.}~\bsnm{Holmes}},
+\oauthor{\binits{J.A.}~\bsnm{Clark}},
+\oauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
+\oauthor{\binits{A.J.G.}~\bsnm{Gray}},
+\oauthor{\binits{S.}~\bsnm{Owen}},
+\oauthor{\binits{A.R.}~\bsnm{Williams}},
+\oauthor{\binits{G.}~\bsnm{Tartari}},
+\oauthor{\binits{F.}~\bsnm{Bacall}},
+\oauthor{\binits{T.}~\bsnm{Thelen}},
+\oauthor{\binits{H.}~\bsnm{M\'{e}nager}},
+\oauthor{\binits{L.}~\bsnm{Rodr\'{\i}guez-Navas}},
+\oauthor{\binits{P.}~\bsnm{Walk}},
+\oauthor{\binits{B.}~\bsnm{Whitehead}},
+\oauthor{\binits{M.}~\bsnm{Wilkinson}},
+\oauthor{\binits{P.}~\bsnm{Groth}},
+\oauthor{\binits{E.}~\bsnm{Bremer}},
+\oauthor{\binits{L.G.}~\bsnm{Castro}},
+\oauthor{\binits{K.}~\bsnm{Sebby}},
+\oauthor{\binits{A.}~\bsnm{Kanitz}},
+\oauthor{\binits{A.}~\bsnm{Trisovic}},
+\oauthor{\binits{G.}~\bsnm{Kennedy}},
+\oauthor{\binits{M.}~\bsnm{Graves}},
+\oauthor{\binits{J.}~\bsnm{Koehorst}} and
+\oauthor{\binits{S.}~\bsnm{Leo}},
+{RO}-Crate Metadata Specification 1.1,
+Technical Report,
+2020.
+\bid{doi={10.5281/zenodo.4031327}}
+\end{botherref}
+%
+\OrigBibText
+P.~Sefton,
+E.{\'{O}}.~Carrag\'{a}in,
+S.~Soiland-Reyes,
+O.~Corcho,
+D.~Garijo,
+R.~Palma,
+F.~Coppens,
+C.~Goble,
+J.M.~Fern\'{a}ndez,
+K.~Chard,
+J.M.~Gomez-Perez,
+M.R.~Crusoe,
+I.~Eguinoa,
+N.~Juty,
+K.~Holmes,
+J.A.~Clark,
+S.~Capella-Gutierrez,
+A.J.G.~Gray,
+S.~Owen,
+A.R.~Williams,
+G.~Tartari,
+F.~Bacall,
+T.~Thelen,
+H.~M\'{e}nager,
+L.~Rodr\'iguez-Navas,
+P.~Walk,
+B.~Whitehead,
+M.~Wilkinson,
+P.~Groth,
+E.~Bremer,
+L.G.~Castro,
+K.~Sebby,
+A.~Kanitz,
+A.~Trisovic,
+G.~Kennedy,
+M.~Graves,
+J.~Koehorst and
+S.~Leo,
+{RO}-Crate Metadata Specification 1.1,
+Technical Report,
+2020.
+doi:10.5281/zenodo.4031327.
+\url{https://w3id.org/ro/crate/1.1}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b108 ###bbsrt2
+%b114 ###
+\bibitem{soilandreyes_tweet_2020}
+%
+\begin{botherref}
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+I am looking for which bioinformatics journals encourage authors
+to submit their code/pipeline/workflow supporting data analysis,
+2020.
+\url{https://twitter.com/soilandreyes/status/1250721245622079488}.
+\end{botherref}
+%
+\OrigBibText
+S.~Soiland-Reyes,
+I am looking for which bioinformatics journals encourage authors to submit
+their code/pipeline/workflow supporting data analysis,
+2020.
+\url{https://twitter.com/soilandreyes/status/1250721245622079488}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b109 ###bbsrt2
+%b79 ###
+\bibitem{doi:10.5281/zenodo.4633732}
+%
+\begin{botherref}
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+{Describing and packaging workflows using RO-Crate and BioCompute
+Objects}, Zenodo,
+2021,
+Webinar for U.S. Food and Drug Administration (FDA), 2021-05-12.
+\bid{doi={10.5281/zenodo.4633732}}
+\end{botherref}
+%
+\OrigBibText
+S.~Soiland-Reyes,
+{Describing and packaging workflows using RO-Crate and BioCompute Objects},
+Zenodo,
+2021,
+{Webinar for U.S. Food and Drug Administration (FDA), 2021-05-12}.
+doi:10.5281/zenodo.4633732.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b110 ###bbsrt2
+%b71 ###
+\bibitem{doi:10.5281/zenodo.51314}
+%
+\begin{botherref}
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\oauthor{\binits{P.}~\bsnm{Alper}} and
+\oauthor{\binits{C.}~\bsnm{Goble}},
+Tracking Workflow Execution With {TavernaPROV},
+{ProvenanceWeek} 2016, 2016, PROV: Three Years Later.
+\bid{doi={10.5281/zenodo.51314}}
+\end{botherref}
+%
+\OrigBibText
+S.~Soiland-Reyes,
+P.~Alper and
+C.~Goble,
+\textit{Tracking Workflow Execution With {TavernaPROV}},
+in: {ProvenanceWeek} 2016,
+2016,
+PROV: Three Years Later.
+doi:10.5281/zenodo.51314.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b111 ###bbsrt2
+%b76 ###
+\bibitem{doi:10.5281/zenodo.12586}
+%
+\begin{botherref}
+\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
+\oauthor{\binits{M.}~\bsnm{Gamble}} and
+\oauthor{\binits{R.}~\bsnm{Haines}},
+Research Object Bundle 1.0
+(2014).
+See \url{https://w3id.org/bundle/2014-11-05/}.
+doi: \doiurl{10.5281/zenodo.12586}.
+\end{botherref}
+%
+\OrigBibText
+S.~Soiland-Reyes,
+M.~Gamble and
+R.~Haines,
+\textit{Research Object Bundle 1.0},
+2014,
+See https://w3id.org/bundle/2014-11-05/.
+doi:10.5281/zenodo.12586.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b112 ###bbsrt2
+%b32 ###
+\bibitem{sporny_2014}
+%
+\begin{botherref}
+\oauthor{\binits{M.}~\bsnm{Sporny}},
+\oauthor{\binits{D.}~\bsnm{Longley}},
+\oauthor{\binits{G.}~\bsnm{Kellogg}},
+\oauthor{\binits{M.}~\bsnm{Lanthaler}} and
+\oauthor{\binits{N.}~\bsnm{Lindstr\"{o}m}},
+{JSON}-{LD} 1.0, {W3C} Recommendation, 2014. \url
+{https://www.w3.org/TR/2014/REC-json-ld-20140116/}.
+\end{botherref}
+%
+\OrigBibText
+M.~Sporny,
+D.~Longley,
+G.~Kellogg,
+M.~Lanthaler and
+N.~Lindstr\"{o}m,
+{JSON}-{LD} 1.0,
+{W3C} Recommendation,
+2014.
+\url{https://www.w3.org/TR/2014/REC-json-ld-20140116/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b113 ###bbsrt2
+%b2 ###
+\bibitem{doi:10.1126/science.aah6168}
+%
+\begin{barticle}
+\bauthor{\binits{V.}~\bsnm{Stodden}},
+\bauthor{\binits{M.}~\bsnm{McNutt}},
+\bauthor{\binits{D.H.}~\bsnm{Bailey}},
+\bauthor{\binits{E.}~\bsnm{Deelman}},
+\bauthor{\binits{Y.}~\bsnm{Gil}},
+\bauthor{\binits{B.}~\bsnm{Hanson}},
+\bauthor{\binits{M.A.}~\bsnm{Heroux}},
+\bauthor{\binits{J.P.A.}~\bsnm{Ioannidis}} and
+\bauthor{\binits{M.}~\bsnm{Taufer}},
+\batitle{Enhancing reproducibility for computational methods},
+\bjtitle{Science}
+\bvolume{354}(\bissue{6317})
+(\byear{2016}),
+\bfpage{1240}--\blpage{1241}.
+\bid{doi={10.1126/science.aah6168}}
+\end{barticle}
+%
+\OrigBibText
+V.~Stodden,
+M.~{McNutt},
+D.H.~Bailey,
+E.~Deelman,
+Y.~Gil,
+B.~Hanson,
+M.A.~Heroux,
+J.P.A.~Ioannidis and
+M.~Taufer,
+\textit{Enhancing reproducibility for computational methods},
+Science
+354(6317)
+(2016),
+1240--1241.
+doi:10.1126/science.aah6168.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b114 ###bbsrt2
+%b82 ###
+\bibitem{doi:10125/4567}
+%
+\begin{bchapter}
+\bauthor{\binits{N.}~\bsnm{Thieberger}} and
+\bauthor{\binits{L.}~\bsnm{Barwick}},
+\bctitle{Keeping records of language diversity in melanesia: The Pacific
+and regional archive for digital sources in endangered cultures ({PARADISEC})},
+in: \bbtitle{Melanesian Languages on the Edge of Asia: Challenges for
+the 21st Century},
+\beditor{\binits{N.}~\bsnm{Evans}} and
+\beditor{\binits{M.}~\bsnm{Klamer}}, eds,
+\bsertitle{Language Documentation \& Conservation Special Publication},
+Vol.~\bseriesno{SP05},
+\bpublisher{University of Hawai'i Press},
+\byear{2012},
+pp.~\bfpage{239}--\blpage{253}.
+ISBN \bisbn{978-0-9856211-2-4}.
+\bid{doi={10125/4567}}
+\end{bchapter}
+%
+\OrigBibText
+N.~Thieberger and
+L.~Barwick,
+\textit{Keeping records of language diversity in Melanesia: The Pacific and
+Regional Archive for Digital Sources in Endangered Cultures ({PARADISEC})},
+in: Melanesian Languages on the Edge of Asia: Challenges for the 21st
+Century,
+N.~Evans and
+M.~Klamer, eds,
+Language Documentation \& Conservation Special Publication,
+Vol.~SP05,
+University of Hawai'i Press,
+2012,
+pp.~239--253.
+ISBN 978-0-9856211-2-4.
+doi:10125/4567.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b115 ###bbsrt2
+%b53 ###
+\bibitem{arkisto-data-portal}
+%
+\begin{botherref}
+{Tools: Data Portal \& Discovery}.
+\url{https://arkisto-platform.github.io/tools/portal/}.
+\end{botherref}
+%
+\OrigBibText
+Tools: Data Portal \& Discovery.
+\url{https://arkisto-platform.github.io/tools/portal/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b116 ###bbsrt2
+%b123 ###
+\bibitem{doi:10.1007/s11042-009-0397-2}
+%
+\begin{barticle}
+\bauthor{\binits{R.}~\bsnm{Troncy}},
+\bauthor{\binits{W.}~\bsnm{Bailer}},
+\bauthor{\binits{M.}~\bsnm{H\"{o}ffernig}} and
+\bauthor{\binits{M.}~\bsnm{Hausenblas}},
+\batitle{{VAMP}: A service for validating {MPEG}-7 descriptions w.r.t.
+to formal profile definitions},
+\bjtitle{Multimedia tools and applications}
+\bvolume{46}(\bissue{2--3})
+(\byear{2010}),
+\bfpage{307}--\blpage{329}.
+\url{https://www.persistent-identifier.nl/urn:nbn:nl:ui:18-14511}.
+\bid{doi={10.1007/s11042-009-0397-2}}
+\end{barticle}
+%
+\OrigBibText
+R.~Troncy,
+W.~Bailer,
+M.~H\"{o}ffernig and
+M.~Hausenblas,
+\textit{{VAMP}: a service for validating {MPEG}-7 descriptions w.r.t. to
+formal profile definitions},
+Multimedia tools and applications
+46(2--3)
+(2010),
+307--329.
+doi:10.1007/s11042-009-0397-2.
+\url{https://www.persistent-identifier.nl/urn:nbn:nl:ui:18-14511}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b117 ###bbsrt2
+%b97 ###
+\bibitem{vandesompel_2007}
+%
+\begin{barticle}
+\bauthor{\binits{H.}~\bsnm{Van de Sompel}} and
+\bauthor{\binits{C.}~\bsnm{Lagoze}},
+\batitle{Interoperability for the discovery, use, and re-use of units
+of scholarly communication},
+\bjtitle{{CTWatch} Quarterly}
+\bvolume{3}(\bissue{3})
+(\byear{2007}).
+\url
+{http://icl.utk.edu/ctwatch/quarterly/articles/2007/08/interoperability-for-the-discovery-use-and-re-use-of-units-of-scholarly-communication/}.
+\end{barticle}
+%
+\OrigBibText
+H.~Van~de Sompel and
+C.~Lagoze,
+Interoperability for the Discovery, Use, and Re-Use of Units of Scholarly
+Communication,
+\textit{{CTWatch} Quarterly}
+\textbf{3}(3)
+(2007)
+00.
+\url
+{http://icl.utk.edu/ctwatch/quarterly/articles/2007/08/interoperability-for-the-discovery-use-and-re-use-of-units-of-scholarly-communication/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b118 ###bbsrt2
+%b50 ###
+\bibitem{vergoulis2021schema}
+%
+\begin{bbook}
+\bauthor{\binits{T.}~\bsnm{Vergoulis}},
+\bauthor{\binits{K.}~\bsnm{Zagganas}},
+\bauthor{\binits{L.}~\bsnm{Kavouras}},
+\bauthor{\binits{M.}~\bsnm{Reczko}},
+\bauthor{\binits{S.}~\bsnm{Sartzetakis}} and
+\bauthor{\binits{T.}~\bsnm{Dalamagas}},
+\bbtitle{{SCHeMa}: Scheduling Scientific Containers on a Cluster of
+Heterogeneous Machines},
+\byear{2021}.
+\url{https://arxiv.org/abs/2103.13138v1}.
+\end{bbook}
+%
+\OrigBibText
+T.~Vergoulis,
+K.~Zagganas,
+L.~Kavouras,
+M.~Reczko,
+S.~Sartzetakis and
+T.~Dalamagas,
+{SCHeMa}: Scheduling Scientific Containers on a Cluster of Heterogeneous
+Machines,
+\textit{arXiv}
+(2021),
+2103.13138--.
+\url{https://arxiv.org/abs/2103.13138v1}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b119 ###bbsrt2
+%b20 ###
+\bibitem{doi:10.1007/s00267-014-0258-2}
+%
+\begin{barticle}
+\bauthor{\binits{C.J.}~\bsnm{Volk}},
+\bauthor{\binits{Y.}~\bsnm{Lucero}} and
+\bauthor{\binits{K.}~\bsnm{Barnas}},
+\batitle{Why is data sharing in collaborative natural resource efforts
+so hard and what can we do to improve it?},
+\bjtitle{Environmental Management}
+\bvolume{53}(\bissue{5})
+(\byear{2014}),
+\bfpage{883}--\blpage{893}.
+\bid{doi={10.1007/s00267-014-0258-2}}
+\end{barticle}
+%
+\OrigBibText
+C.J.~Volk,
+Y.~Lucero and
+K.~Barnas,
+\textit{Why is data sharing in collaborative natural resource efforts
+so hard
+and what can we do to improve it?},
+Environmental Management
+53(5)
+(2014),
+883--893.
+doi:10.1007/s00267-014-0258-2.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b120 ###bbsrt2
+%b31 ###
+\bibitem{httprange14}
+%
+\begin{botherref}
+{W3C Technical Architecture Group}, Dereferencing {HTTP}
+{URIs}, Draft Tag Finding,
+2007.
+\url{https://www.w3.org/2001/tag/doc/httpRange-14/2007-08-31/HttpRange-14.html}.
+\end{botherref}
+%
+\OrigBibText
+{W3C Technical Architecture Group},
+Dereferencing {HTTP} {URIs},
+Draft Tag Finding,
+2007.
+\url{https://www.w3.org/2001/tag/doc/httpRange-14/2007-08-31/HttpRange-14.html}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b121 ###bbsrt2
+%b85 ###
+\bibitem{doi:10.15497/rda00039}
+%
+\begin{bbook}
+\bauthor{\binits{P.}~\bsnm{Walk}},
+\bauthor{\binits{T.}~\bsnm{Miksa}} and
+\bauthor{\binits{P.}~\bsnm{Neish}},
+\bbtitle{{RDA} {DMP} Common Standard for Machine-Actionable Data Management
+Plans, Research Data Alliance},
+\byear{2019}.
+\bid{doi={10.15497/rda00039}}
+\end{bbook}
+%
+\OrigBibText
+P.~Walk,
+T.~Miksa and
+P.~Neish,
+\textit{{RDA} {DMP} Common Standard for Machine-actionable Data Management
+Plans},
+Research Data Alliance
+(2019).
+doi:10.15497/rda00039.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b122 ###bbsrt2
+%b72 ###
+\bibitem{doi:10.3897/rio.6.e57602}
+%
+\begin{barticle}
+\bauthor{\binits{S.}~\bsnm{Walton}},
+\bauthor{\binits{L.}~\bsnm{Livermore}},
+\bauthor{\binits{O.}~\bsnm{B\'{a}nki}},
+\bauthor{\binits{R.}~\bsnm{Cubey}},
+\bauthor{\binits{R.}~\bsnm{Drinkwater}},
+\bauthor{\binits{M.}~\bsnm{Englund}},
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{Q.}~\bsnm{Groom}},
+\bauthor{\binits{C.}~\bsnm{Kermorvant}},
+\bauthor{\binits{I.}~\bsnm{Rey}},
+\bauthor{\binits{C.}~\bsnm{Santos}},
+\bauthor{\binits{B.}~\bsnm{Scott}},
+\bauthor{\binits{A.}~\bsnm{Williams}} and
+\bauthor{\binits{Z.}~\bsnm{Wu}},
+\batitle{Landscape analysis for the specimen data refinery},
+\bjtitle{Research Ideas and Outcomes}
+\bvolume{6}
+(\byear{2020}).
+\bid{doi={10.3897/rio.6.e57602}}
+\end{barticle}
+%
+\OrigBibText
+S.~Walton,
+L.~Livermore,
+O.~B\'{a}nki,
+R.~Cubey,
+R.~Drinkwater,
+M.~Englund,
+C.~Goble,
+Q.~Groom,
+C.~Kermorvant,
+I.~Rey,
+C.~Santos,
+B.~Scott,
+A.~Williams and
+Z.~Wu,
+Landscape analysis for the Specimen Data Refinery,
+\textit{Research Ideas and Outcomes}
+\textbf{6}
+(2020)
+00.
+doi:10.3897/rio.6.e57602.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b123 ###bbsrt2
+%b5 ###
+\bibitem{doi:10.1038/sdata.2016.18}
+%
+\begin{barticle}
+\bauthor{\binits{M.D.}~\bsnm{Wilkinson}},
+\bauthor{\binits{M.}~\bsnm{Dumontier}},
+\bauthor{\binits{I.J.J.}~\bsnm{Aalbersberg}},
+\bauthor{\binits{G.}~\bsnm{Appleton}},
+\bauthor{\binits{M.}~\bsnm{Axton}},
+\bauthor{\binits{A.}~\bsnm{Baak}},
+\bauthor{\binits{N.}~\bsnm{Blomberg}},
+\bauthor{\binits{J.-W.}~\bsnm{Boiten}},
+\bauthor{\binits{L.B.}~\bsnm{da~Silva Santos}},
+\bauthor{\binits{P.E.}~\bsnm{Bourne}},
+\bauthor{\binits{J.}~\bsnm{Bouwman}},
+\bauthor{\binits{A.J.}~\bsnm{Brookes}},
+\bauthor{\binits{T.}~\bsnm{Clark}},
+\bauthor{\binits{M.}~\bsnm{Crosas}},
+\bauthor{\binits{I.}~\bsnm{Dillo}},
+\bauthor{\binits{O.}~\bsnm{Dumon}},
+\bauthor{\binits{S.}~\bsnm{Edmunds}},
+\bauthor{\binits{C.T.}~\bsnm{Evelo}},
+\bauthor{\binits{R.}~\bsnm{Finkers}},
+\bauthor{\binits{A.}~\bsnm{Gonzalez-Beltran}},
+\bauthor{\binits{A.J.G.}~\bsnm{Gray}},
+\bauthor{\binits{P.}~\bsnm{Groth}},
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bauthor{\binits{J.S.}~\bsnm{Grethe}},
+\bauthor{\binits{J.}~\bsnm{Heringa}},
+\bauthor{\binits{P.A.C.}~\bsnm{'t~Hoen}},
+\bauthor{\binits{R.}~\bsnm{Hooft}},
+\bauthor{\binits{T.}~\bsnm{Kuhn}},
+\bauthor{\binits{R.}~\bsnm{Kok}},
+\bauthor{\binits{J.}~\bsnm{Kok}},
+\bauthor{\binits{S.J.}~\bsnm{Lusher}},
+\bauthor{\binits{M.E.}~\bsnm{Martone}},
+\bauthor{\binits{A.}~\bsnm{Mons}},
+\bauthor{\binits{A.L.}~\bsnm{Packer}},
+\bauthor{\binits{B.}~\bsnm{Persson}},
+\bauthor{\binits{P.}~\bsnm{Rocca-Serra}},
+\bauthor{\binits{M.}~\bsnm{Roos}},
+\bauthor{\binits{R.}~\bsnm{van Schaik}},
+\bauthor{\binits{S.-A.}~\bsnm{Sansone}},
+\bauthor{\binits{E.}~\bsnm{Schultes}},
+\bauthor{\binits{T.}~\bsnm{Sengstag}},
+\bauthor{\binits{T.}~\bsnm{Slater}},
+\bauthor{\binits{G.}~\bsnm{Strawn}},
+\bauthor{\binits{M.A.}~\bsnm{Swertz}},
+\bauthor{\binits{M.}~\bsnm{Thompson}},
+\bauthor{\binits{J.}~\bsnm{van der Lei}},
+\bauthor{\binits{E.}~\bsnm{van Mulligen}},
+\bauthor{\binits{J.}~\bsnm{Velterop}},
+\bauthor{\binits{A.}~\bsnm{Waagmeester}},
+\bauthor{\binits{P.}~\bsnm{Wittenburg}},
+\bauthor{\binits{K.}~\bsnm{Wolstencroft}},
+\bauthor{\binits{J.}~\bsnm{Zhao}} and
+\bauthor{\binits{B.}~\bsnm{Mons}},
+\batitle{The {FAIR} guiding principles for scientific data management
+and stewardship},
+\bjtitle{Scientific data}
+\bvolume{3}
+(\byear{2016}),
+\bnumber{160018}.
+\bid{doi={10.1038/sdata.2016.18}}
+\end{barticle}
+%
+\OrigBibText
+M.D.~Wilkinson,
+M.~Dumontier,
+I.J.J.~Aalbersberg,
+G.~Appleton,
+M.~Axton,
+A.~Baak,
+N.~Blomberg,
+J.-W.~Boiten,
+L.B.~da~Silva~Santos,
+P.E.~Bourne,
+J.~Bouwman,
+A.J.~Brookes,
+T.~Clark,
+M.~Crosas,
+I.~Dillo,
+O.~Dumon,
+S.~Edmunds,
+C.T.~Evelo,
+R.~Finkers,
+A.~Gonzalez-Beltran,
+A.J.G.~Gray,
+P.~Groth,
+C.~Goble,
+J.S.~Grethe,
+J.~Heringa,
+P.A.C.~'t~Hoen,
+R.~Hooft,
+T.~Kuhn,
+R.~Kok,
+J.~Kok,
+S.J.~Lusher,
+M.E.~Martone,
+A.~Mons,
+A.L.~Packer,
+B.~Persson,
+P.~Rocca-Serra,
+M.~Roos,
+R.~van Schaik,
+S.-A.~Sansone,
+E.~Schultes,
+T.~Sengstag,
+T.~Slater,
+G.~Strawn,
+M.A.~Swertz,
+M.~Thompson,
+J.~van~der Lei,
+E.~van Mulligen,
+J.~Velterop,
+A.~Waagmeester,
+P.~Wittenburg,
+K.~Wolstencroft,
+J.~Zhao and
+B.~Mons,
+\textit{The {FAIR} Guiding Principles for scientific data management and
+stewardship.},
+Scientific data
+3
+(2016),
+160018.
+doi:10.1038/sdata.2016.18.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b124 ###bbsrt2
+% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
+%incorrectly.
+%b48 ###
+\bibitem{about-workflowhub}
+%
+\begin{botherref}
+{WorkflowHub} project | Project pages for developing and
+running the {WorkflowHub}, a registry of scientific workflows.
+\url{https://w3id.org/workflowhub/}.
+\end{botherref}
+%
+\OrigBibText
+{WorkflowHub} project \textbar Project pages for developing and running the
+{WorkflowHub}, a registry of scientific workflows.
+\url{https://w3id.org/workflowhub/}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b125 ###bbsrt2
+%b75 ###
+\bibitem{doi:10.1109/eScience.2012.6404482}
+%
+\begin{bchapter}
+\bauthor{\binits{J.}~\bsnm{Zhao}},
+\bauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
+\bauthor{\binits{K.}~\bsnm{Belhajjame}},
+\bauthor{\binits{G.}~\bsnm{Klyne}},
+\bauthor{\binits{E.}~\bsnm{Garcia-Cuesta}},
+\bauthor{\binits{A.}~\bsnm{Garrido}},
+\bauthor{\binits{K.}~\bsnm{Hettne}},
+\bauthor{\binits{M.}~\bsnm{Roos}},
+\bauthor{\binits{D.}~\bsnm{De Roure}} and
+\bauthor{\binits{C.}~\bsnm{Goble}},
+\bctitle{Why workflows break -- understanding and combating decay in
+taverna workflows},
+in: \bbtitle{2012 {IEEE} 8th International Conference on e-Science},
+\bpublisher{IEEE},
+\byear{2012},
+pp.~\bfpage{1}--\blpage{9}.
+\url{https://www.research.manchester.ac.uk/portal/files/174861334/why\_
+decay.pdf}.
+ISBN \bisbn{978-1-4673-4466-1}.
+\bid{doi={10.1109/eScience.2012.6404482}}
+\end{bchapter}
+%
+\OrigBibText
+J.~Zhao,
+J.M.~Gomez-Perez,
+K.~Belhajjame,
+G.~Klyne,
+E.~Garcia-Cuesta,
+A.~Garrido,
+K.~Hettne,
+M.~Roos,
+D.~De~Roure and
+C.~Goble,
+\textit{Why workflows break -- Understanding and combating decay in Taverna
+workflows},
+in: 2012 {IEEE} 8th International Conference on E-Science,
+IEEE,
+2012,
+pp.~1--9.
+ISBN 978-1-4673-4466-1.
+doi:10.1109/eScience.2012.6404482.
+\url
+{https://www.research.manchester.ac.uk/portal/files/174861334/why_decay.pdf}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b126 ###bbsrt2
+%b60 ###
+\bibitem{doi:10.5281/zenodo.5068950}
+%
+\begin{botherref}
+\oauthor{\binits{F.}~\bsnm{Zoubek}} and
+\oauthor{\binits{M.}~\bsnm{Winkler}},
+RO Crates and Excel
+(2021).
+\url{https://github.com/e11938258/RO-Crates-and-Excel}.
+\bid{doi={10.5281/zenodo.5068950}}
+\end{botherref}
+%
+\OrigBibText
+F.~Zoubek and
+M.~Winkler,
+RO Crates and Excel,
+Zenodo,
+2021.
+doi:10.5281/zenodo.5068950.
+\url{https://github.com/e11938258/RO-Crates-and-Excel}.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+%b127 ###bbsrt2
+%b18 ###
+\bibitem{doi:10.1515/9783598441844}
+%
+\begin{bbook}
+\bauthor{\binits{M.}~\bsnm{\v{Z}umer}},
+\bbtitle{National Bibliographies in the Digital Age: Guidance and New
+Directions, {IFLA} Series on Bibliographic Control},
+\bsertitle{IFLA Working Group on Guidelines for National Bibliographies},
+\bpublisher{Walter de Gruyter -- K. G. Saur},
+\byear{2009},
+\bcomment{ISSN 1868-8438}.
+ISBN \bisbn{9783598441844}.
+\end{bbook}
+%
+\OrigBibText
+M.~\v{Z}umer,
+\textit{National bibliographies in the digital age: guidance and new
+directions},
+{IFLA} series on bibliographic control,
+Walter de Gruyter -- K. G. Saur,
+2009,
+{IFLA Working Group on Guidelines for National Bibliographies}.
+ISSN 1868-8438.
+ISBN 9783598441844.
+doi:10.1515/9783598441844.
+\endOrigBibText
+\bptok{structpyb}
+\endbibitem
+
+\end{thebibliography}
 \end{document}
+%

### 2022-01-01 latex/ro-crate.tex
avoid schema.org as hyperlinks

@@
 readable manner.
-RO-Crate is based on \surl{Schema.org} annotations in JSON-LD, aiming to
+RO-Crate is based on Schema.org annotations in JSON-LD, aiming to
 establish best
@@
 For our development methodology we followed the mantra of working
-closely with a small group to really get a deep understanding of
+closely with a small group to get a deep understanding of
 requirements and ensure rapid feedback loops. We created a pool of
@@
 interactive tools (Section~\ref{sec:tooling}) like
-Describo\footnote{\url{https://uts-eresearch.github.io/describo/}} \cite{describo} and
+Describo\footnote{\url{https://arkisto-platform.github.io/describo/}} \cite{describo} and
 Jupyter notebooks \cite{doi:10.3233/978-1-61499-649-1-87}, and by close
@@
 aggregated and non-aggregated data entities). Multiple types and
-relations from \surl{Schema.org} allow annotations to be more specific,
+relations from Schema.org allow annotations to be more specific,
 including figures, nested datasets, computational workflows, people,
@@
 opinionated and example-driven guide to writing
-\surl{Schema.org}\footnote{\url{https://schema.org/}} \cite{doi:10.1145/2857274.2857276}
+Schema.org\footnote{\url{https://schema.org/}} \cite{doi:10.1145/2857274.2857276}
 metadata as JSON-LD \cite{sporny_2014} (see Section~\ref{sec:implementation}), which leaves it open for implementers to include
-additional metadata using other \surl{Schema.org} types and properties, or
+additional metadata using other Schema.org types and properties, or
 even additional Linked Data vocabularies/ontologies or their own ad-hoc terms.
@@
 The RO-Crate conceptual model has been realised using JSON-LD and
-\surl{Schema.org} in a prescriptive form as discussed in Section~\ref{sec:conceptual}. These technical choices were made to cater for
+Schema.org in a prescriptive form as discussed in Section~\ref{sec:conceptual}. These technical choices were made to cater for
 simplicity from a developer perspective (as introduced in Section~\ref{sec:methodology}).
@@
 If processed as JSON-LD, this forms an RDF graph by matching the \texttt{@id}
-IRIs and applying the \texttt{@context} mapping to \surl{schema.org} terms.
+IRIs and applying the \texttt{@context} mapping to Schema.org terms.
 \normalsize
@@
 
-RO-Crate reuses vocabulary terms and IRIs from \surl{Schema.org}, but provides
+RO-Crate reuses vocabulary terms and IRIs from Schema.org, but provides
 its own versioned JSON-LD
@@
 otherwise unknown properties and types. Not reusing the official
-\surl{Schema.org} context means RO-Crate is also able to map in additional
+Schema.org context means RO-Crate is also able to map in additional
 vocabularies where needed, namely the \textit{Portland Common Data Model}
@@
 
-Similarly, while the \surl{Schema.org} context
+Similarly, while the Schema.org context
 currently\footnote{\url{https://schema.org/version/13.0/schemaorg-current-http.jsonld}}
@@
 Simplifying from the CWLProv approach, the planned Workflow Run
-RO-Crate profile will use a high level \surl{Schema.org}
+RO-Crate profile will use a high level Schema.org
 provenance\footnote{\url{https://www.researchobject.org/ro-crate/1.1/provenance.html\#software-used-to-create-files}}
@@
 the PARADISEC platform and can be archived, preserved and processed in
-its own right, using \surl{Schema.org} as base vocabulary and extended with
+its own right, using Schema.org as base vocabulary and extended with
 PARADISEC-specific terms.
@@
 software already has extensive support for metadata, including the Data
-Documentation Initiative (DDI), Dublin Core, DataCite, and \surl{Schema.org}.
+Documentation Initiative (DDI), Dublin Core, DataCite, and Schema.org.
 Incorporating RO-Crate, which has the flexibility to describe a wide
@@
 Metadata File.
-\item[4.] By being based on Linked Data and \surl{schema.org} vocabulary, meaning
+\item[4.] By being based on Linked Data and the Schema.org vocabulary, meaning
 that PIDs already exist for common types and properties.
@@
 However, just like the RO-Crate specification can be thought of as a
-\textit{core profile} of \surl{schema.org} in JSON-LD, we cannot stress the
+\textit{core profile} of Schema.org in JSON-LD, we cannot stress the
 importance of also establishing domain-specific RO-Crate profiles and
@@
 entity types requested by users, e.g. detailed workflow runs or
-container provenance, which do not have a good match in \surl{Schema.org}.
+container provenance, which do not have a good match in Schema.org.
 Such support could be added, for instance, by integrating other
@@
 this will help us to improve our specification and tools. By
-discovering commonalities in emerging usage (e.g. additional \surl{schema.org}
+discovering commonalities in emerging usage (e.g. additional Schema.org
 types), the community helps to reduce divergence that could otherwise
@@
 and software packages, tabular data structures; these are not always
-matched well with existing \surl{schema.org} types but may benefit from other
+matched well with existing Schema.org types but may benefit from other
 vocabularies or even separate metadata files, e.g. from Frictionless
@@
 In reality many additional attributes and contextual types from
-\surl{schema.org} types like \url{http://schema.org/affiliation} and \url
+Schema.org types like \url{http://schema.org/affiliation} and \url
 {http://schema.org/Organization} would be used to further describe the
@@
 Below follows a mapping from $\mathcal{L}_{\mathit{rocrate}}$ to $\mathcal
-{L}_{\textit{RDF}}$ using \surl{schema.org} as vocabulary:
+{L}_{\textit{RDF}}$ using Schema.org as vocabulary:
 %
@@
 $\textit{Class}$ and $\textit{Property}$ is typically indirect: The JSON-LD \texttt
-{@context} maps to \surl{schema.org} IRIs, which, when resolved as Linked
+{@context} maps to Schema.org IRIs, which, when resolved as Linked
 Data, embed their formal definition as RDFa. Extensions may however
@@
 
-Combining the above predicates and \surl{schema.org} mapping with rudimentary
+Combining the above predicates and Schema.org mapping with rudimentary
 JSON templates, these forward-chaining production rules can output
@@
 and contextual entities, but these are proposed to be bound to the root
-data entity with \surl{schema.org} relation \url{http://schema.org/mentions}.
+data entity with Schema.org relation \url{http://schema.org/mentions}.
 

### 2022-01-01 latex/ro-crate.tex
Avoid repeating "JSON-LD", add citation early

@@
 as a self-described \textbf{Root Data Entity} that describes and contains
-\textit{data entities}, which are further described by referencing {contextual
+\textit{data entities}, which are further described by referencing \textit{contextual
 entities}. A \textbf{data entity} is either a \textit{file} (i.e. a byte sequence
@@
 workflow language) and is stored solely by its metadata. The
-representation of a \textbf{data entity} as a byte sequence makes it
+representation of a \textit{data entity} as a byte sequence makes it
 possible to store a variety of research artefacts including not only
@@
 the presence of the \textbf{RO-Crate Metadata File} \texttt{ro-crate-metadata.json}
-(top of Fig.~\ref{fig:conceptual}). This is a JSON-LD file that
+(top of Fig.~\ref{fig:conceptual}). This file
 describes the RO-Crate, its content and related metadata using Linked
-Data. JSON-LD is a W3C standard RDF serialisation that has become
+Data in JSON-LD format \cite{sporny_2014}. This is a W3C standard RDF serialisation that has become
 popular as it is easy to read by humans while also offers some

### 2022-01-01 latex/ro-crate.tex
not directly

@@
 would themselves also have a human-readable description in the case
-where their PID does not resolve to human-readable documentation.
+where their PID does not directly resolve to human-readable documentation.
 

### 2022-01-01 latex/ro-crate.tex
associated with

@@
 of describing Research Objects in a structured language, while reducing
-the steep learning curve otherwise associated in Semantic Web
+the steep learning curve otherwise associated with Semantic Web
 adaptation, like development of ontologies, identifiers, namespaces,

### 2022-01-01 latex/ro-crate.tex
*ized -> ised (British spelling)

@@
 The conceptual notion of a \textit{Research Object}
-\cite{doi:10.1016/j.future.2011.08.004} is thus realized with the RO-Crate
-model and serialized using Linked Data constructs within the RO-Crate
+\cite{doi:10.1016/j.future.2011.08.004} is thus realised with the RO-Crate
+model and serialised using Linked Data constructs within the RO-Crate
 metadata file.
@@
 aspects that have already been specified by the RO-Crate community, but
-can extend and integrate RO-Crate in multiple standardized ways.
+can extend and integrate RO-Crate in multiple standardised ways.
 
@@
 scientific domains. However we have found that domain-specific use of
-RO-Crate will, implicitly or explicitly, form a specialized \textbf{profile}
+RO-Crate will, implicitly or explicitly, form a specialised \textbf{profile}
 of RO-Crate; i.e., {a set of conventions, types and properties that are
@@
 \texttt{ro-crate-metadata.json}\footnote{The avid reader may spot that the RO-Crate Metadata file use the
-extension \texttt{.json} instead of \texttt{.jsonld}, this is to emphasize the
+extension \texttt{.json} instead of \texttt{.jsonld}, this is to emphasise the
 developer expectations as a JSON format, while the file's JSON-LD
@@
 that provides mapping from JSON keys to Linked Data term IRIs, and can
-enable various JSON-LD directives to cater for customized JSON
+enable various JSON-LD directives to cater for customised JSON
 structures for translating to RDF.
@@
 endangered languages across more than 16,000 items, collected and
-digitized over many years by researchers interviewing and recording
+digitised over many years by researchers interviewing and recording
 native speakers across the region.
@@
 
-This initially workflow-centric approach was further formalized as the
+This initially workflow-centric approach was further formalised as the
 Wf4Ever Research Object Model \cite{doi:10.1016/j.websem.2015.01.003},
@@
 I2: \textit{Metadata use vocabularies that follow FAIR principles}), which in
-itself is recogized as a challenge \cite{doi:10.3233/978-1-61499-660-6-9}.
+itself is recognised as a challenge \cite{doi:10.3233/978-1-61499-660-6-9}.
 Adapters of the Wf4Ever RO model would have to navigate documentation
@@
 
-\subsection{Example of formalized RO-Crate}
+\subsection{Example of formalised RO-Crate}
 
@@
 
-A formalized RO-Crate in $\mathcal{L}_{\mathit{rocrate}}$ can be mapped to
+A formalised RO-Crate in $\mathcal{L}_{\mathit{rocrate}}$ can be mapped to
 different serializations.

### 2022-01-01 latex/ro-crate.tex
consistent Schema.org capitalisation

@@
 To further verify this idea of simplicity, we have formalised the
-RO-Crate definition (see \textit{Appendix~\ref{sec:formaldefinition})}). An
+RO-Crate definition (see \textit{Appendix~\ref{sec:formaldefinition}}). An
 important result of this exercise is that the underlying data structure
@@
 like CodeMeta\footnote{\url{https://codemeta.github.io/}} for software metadata,
-Science-on-schema.org\footnote{\url{https://science-on-schema.org/}}
+Science-on-Schema.org\footnote{\url{https://science-on-schema.org/}}
 \cite{doi:10.5281/zenodo.4477164} for datasets, FAIR Digital
@@
 
-\subsection{Mapping to RDF with schema.org}
+\subsection{Mapping to RDF with Schema.org}
 

### 2022-01-01 latex/ro-crate.tex
Remove confusing comments
attempt to use listing package

@@
 %%\newtheorem{}{}%[]
+
+% listing style for JSON/JavaScript - adapted from
+% https://github.com/ghammock/LaTeX_Listings_JavaScript_ES6
+%
+% MIT License
+% 
+% Copyright (c) 2018 Gary Hammock
+% 
+% Permission is hereby granted, free of charge, to any person obtaining a copy
+% of this software and associated documentation files (the "Software"), to deal
+% in the Software without restriction, including without limitation the rights
+% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+% copies of the Software, and to permit persons to whom the Software is
+% furnished to do so, subject to the following conditions:
+% 
+% The above copyright notice and this permission notice shall be included in all
+% copies or substantial portions of the Software.
+% 
+% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+% FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+% SOFTWARE.
+% 
+\lstdefinelanguage{JSON}{
+  morekeywords=[1]{break, continue, delete, else, for, function, if, in,
+    new, return, this, typeof, var, void, while, with},
+  % Literals, primitive types, and reference types.
+  morekeywords=[2]{false, null, true, boolean, number, undefined,
+    Array, Boolean, Date, Math, Number, String, Object},
+  % Built-ins.
+  morekeywords=[3]{eval, parseInt, parseFloat, escape, unescape},
+  sensitive,
+  morecomment=[s]{/*}{*/},
+  morecomment=[l]//,
+  morecomment=[s]{/**}{*/}, % JavaDoc style comments
+  morestring=[b]',
+  morestring=[b]"
+}[keywords, comments, strings]
+%% end listing style
+
 \endlocaldefs
@@
 %%%% Main text entry area:
-% Every *.md file in content/ in order, except 00, 90
 \section{Introduction}
@@
 \begin{listing}
-\includegraphics{210053i02}
+%\includegraphics{210053i02}
+\begin{lstlisting}[style=JSON]
+{ "@context": "https://w3id.org/ro/crate/1.1/context",
+  "@graph": [
+    { "@id": "ro-crate-metadata.json",      
+      "@type": "CreativeWork",
+      "conformsTo": {"@id": "https://w3id.org/ro/crate/1.1"},
+      "about": {"@id": "./"}
+    },
+    { "@id": "./",
+      "@type": "Dataset",
+      "name": "A simplified RO-Crate",
+      "author": {"@id": "#alice"},
+      "license": {"@id": "https://spdx.org/licenses/CC-BY-4.0"},
+      "datePublished": "2021-11-02T16:04:43Z",
+      "hasPart": [
+        {"@id": "survey-responses-2019.csv"},
+        {"@id": "https://example.com/pics/5707039334816454031_o.jpg"}
+      ]
+    },
+    { "@id": "survey-responses-2019.csv",
+      "@type": "File",
+      "about": {"@id": "https://example.com/pics/5707039334816454031_o.jpg"},
+      "author": {"@id": "#alice"}
+    },
+    { "@id": "https://example.com/pics/5707039334816454031_o.jpg",
+      "@type": ["File", "ImageObject"],
+      "contentLocation": {"@id": "http://sws.geonames.org/8152662/"},
+      "author": {"@id": "https://orcid.org/0000-0002-1825-0097"}
+    },
+    { "@id": "#alice",
+      "@type": "Person",
+      "name": "Alice"
+    },
+    { "@id": "https://orcid.org/0000-0002-1825-0097",
+      "@type": "Person",
+      "name": "Josiah Carberry"
+    },
+    { "@id": "http://sws.geonames.org/8152662/",
+      "@type": "Place",
+      "name": "Catalina Park"
+    },
+    { "@id": "https://spdx.org/licenses/CC-BY-4.0",
+      "@type": "CreativeWork",
+      "name": "Creative Commons Attribution 4.0"
+    }
+  ]
+}    
+\end{lstlisting}    
 \caption{%\qq{}\protect\footnotemark[31]
@@
 \texttt{contentLocation} and \texttt{author} that differs from that of the overall
-RO-Crate. In this example, \texttt{about} of the CSV data entity reference the
+RO-Crate. In this crate, \texttt{about} of the CSV data entity reference the
 \texttt{ImageObject}, which then take the roles of both a data entity and
@@
 including the EOSC Interoperability Framework \cite{doi:10.2777/620649}.
-
-% CrEDiT taxonomy:
-%
-% Conceptualization
-% Data curation
-% Formal Analysis
-% Funding acquisition
-% Investigation
-% Methodology
-% Project administration
-% Resources
-% Software
-% Supervision
-% Validation
-% Visualization
-% Writing - original draft
-% Writing - review \& editing
 

### 2022-01-01 latex/ro-crate.tex
ro-crate-excel is in beta
Use verbatim for listing
The listing was included as a picture instead of text, which mean it can't be copy-pasted and - I assume - would appear as a blurry bitmap on the Web. Replaced with basic use of verbatim in footnotesize.

--- 2022-01-01 latex/ro-crate.tex
@@
 \begin{listing}
-\includegraphics{210053i02}
+%\includegraphics{210053i02}
+\footnotesize
+\begin{verbatim}
+{ "@context": "https://w3id.org/ro/crate/1.1/context",
+  "@graph": [
+    { "@id": "ro-crate-metadata.json",      
+      "@type": "CreativeWork",
+      "conformsTo": {"@id": "https://w3id.org/ro/crate/1.1"},
+      "about": {"@id": "./"}
+    },
+    { "@id": "./",
+      "@type": "Dataset",
+      "name": "A simplified RO-Crate",
+      "author": {"@id": "#alice"},
+      "license": {"@id": "https://spdx.org/licenses/CC-BY-4.0"},
+      "datePublished": "2021-11-02T16:04:43Z",
+      "hasPart": [
+        {"@id": "survey-responses-2019.csv"},
+        {"@id": "https://example.com/pics/5707039334816454031_o.jpg"}
+      ]
+    },
+    { "@id": "survey-responses-2019.csv",
+      "@type": "File",
+      "about": {"@id": "https://example.com/pics/5707039334816454031_o.jpg"},
+      "author": {"@id": "#alice"}
+    },
+    { "@id": "https://example.com/pics/5707039334816454031_o.jpg",
+      "@type": ["File", "ImageObject"],
+      "contentLocation": {"@id": "http://sws.geonames.org/8152662/"},
+      "author": {"@id": "https://orcid.org/0000-0002-1825-0097"}
+    },
+    { "@id": "#alice",
+      "@type": "Person",
+      "name": "Alice"
+    },
+    { "@id": "https://orcid.org/0000-0002-1825-0097",
+      "@type": "Person",
+      "name": "Josiah Carberry"
+    },
+    { "@id": "http://sws.geonames.org/8152662/",
+      "@type": "Place",
+      "name": "Catalina Park"
+    },
+    { "@id": "https://spdx.org/licenses/CC-BY-4.0",
+      "@type": "CreativeWork",
+      "name": "Creative Commons Attribution 4.0"
+    }
+  ]
+}    
+\end{verbatim}    
 \caption{%\qq{}\protect\footnotemark[31]
@@
 \multicolumn{4}{@{}l@{}}{\textit{Web-based application to create RO-Crates using cloud storage}}                                                                                                               \\
-\textbf{ro-crate-excel} \cite{ro-crate-excel}                                                                                              & Data managers              & JavaScript       &       \\
+\textbf{ro-crate-excel} \cite{ro-crate-excel}                                                                                              & Data managers              & JavaScript       & Beta  \\
 \multicolumn{4}{@{}l@{}}{\textit{Command-line tool to create/edit RO-Crates with spreadsheets}}                                                                                                                \\

### 2022-01-01 latex/ro-crate.tex
escape {}

@@
 RO-Crate descriptions are on the root data entity \texttt{./} or any of the
-referenced data or contextual entities, as exemplified by the data
+referenced data or contextual entities. This is exemplified by the data
 entity \texttt{ImageObject} referencing contextual entities for
@@
 \url
-{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.} this flattened profile of JSON-LD, each \texttt{{entity}} are directly
+{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.} this flattened profile of JSON-LD, each \texttt{\{entity\}} is directly
 under \texttt{@graph} and represents the RDF triples with a common \textit{subject}
 (\texttt{@id}), mapped \textit{properties} like \texttt{hasPart}, and \textit{objects} -- as
-either literal \texttt{"string"} values, referenced \texttt{{objects}} (which
+either literal \texttt{"string"} values, referenced \texttt{\{objects\}} (which
 properties are listed in its own entity), or a JSON \texttt{[list]} of these.
@@
 JSON-LD distinguishes explicitly between references to other entities
-(\texttt{{"@id": "\#alice"}})
+(\texttt{\{"@id": "\#alice"}\})
 and string values (\texttt{"Alice"}) -- meaning

### 2022-01-01 latex/ro-crate.tex
more texttt/textit tweaks

@@
 currently\footnote{\url{https://schema.org/version/13.0/schemaorg-current-http.jsonld}}
-have \texttt{"@type": "@id"} annotations for object properties, RO-Crate
+have \texttt{"@type": "@id"} annotations for implicit object properties, RO-Crate
 JSON-LD distinguishes explicitly between references to other entities
-(\texttt{\{"@id": "\#alice"}\})
-and string values (\texttt{"Alice"}) -- meaning
+\texttt{\{"@id": "\#alice"}\}
+and string values \texttt{"Alice"} -- meaning
 RO-Crate applications can find references for corresponding entities
 and IRIs without parsing the \texttt{@context} to understand a particular
-property. Notably this is exploited by the \texttt{ro-crate-html-js}
+property. Notably this is exploited by the \textit{ro-crate-html-js}
 \cite{ro-crate-html-js} tool to provide reliable HTML rendering for

### 2022-01-01 latex/ro-crate.tex
Footnote placement

@@
 The RO-Crate community is strongly engaged with the European-wide
-biology/bioinformatics collaborative e-Infrastructure, ELIXIR
+biology/bioinformatics collaborative e-Infrastructure ELIXIR
 \cite{doi:10.1016/j.tibtech.2012.02.002}, along with European Open Science
-Cloud (EOSC\footnote{\url{https://eosc.eu/}}) projects including
+Cloud\footnote{\url{https://eosc.eu/}} (EOSC) projects including
 EOSC-Life,\footnote{\url{https://www.eosc-life.eu/}}
@@
 collaboration space for developing what is now the RO-Crate
-specification, maintained as
-Markdown\footnote{\url{https://github.com/researchobject/ro-crate/}} in GitHub Pages
+specification, maintained\footnote{\url{https://github.com/researchobject/ro-crate/}} as
+Markdown in GitHub Pages
 and published through Zenodo.
@@
 tools for creating and using RO-Crates. Table~\ref{tab1} shows the current set
-of implementations. Reviewing this list, one can see that tools support
+of implementations. Reviewing this list, one can see support for
 commonly used programming languages, including Python, JavaScript, and
-Ruby. Additionally, these tools can be integrated into commonly used
+Ruby. Additionally, the tools can be integrated into commonly used
 research environments, in particular, the command line tool
@@
 the Australian Criminal Characters
-project\footnote{\url{https://criminalcharacters.com/}} annotate historical prisoner
-records to gain greater insight into the history of Australia
+project\footnote{\url{https://criminalcharacters.com/}} to annotate historical prisoner
+records for greater insight into the history of Australia
 \cite{doi:10.1080/14490854.2020.1796500}.

### 2022-01-01 latex/ro-crate.tex
PROV-O citation, ieee2791 permalink

@@
 \cite{doi:10.5281/zenodo.51314}. In both cases, we used the PROV Ontology
-\cite{CheckMyCrate}, including details of every task execution with all the
+\cite{PROVO}, including details of every task execution with all the
 intermediate data, which required significant workflow engine
@@
 combination of JSON
-Schemas\footnote{\url{https://opensource.ieee.org/2791-object/ieee-2791-schema/}}
+Schemas\footnote{\url{https://w3id.org/ieee/ieee-2791-schema/}}
 that define the structure of JSON metadata files describing exemplar

### 2022-01-02 latex/ro-crate.tex
language fixes, URL for metadata in dataverse

@@
 \item[1.] Start a skeleton data management plan based on an existing RO-Crate
-dataset, e.g. from an RO-Crate from WorkflowHub.
+dataset, e.g. an RO-Crate from WorkflowHub.
 \item[2.] Instantiate an RO-Crate based on a data management plan.
@@
 institutional levels. For example, the Australian Research Data
-Commons (ARDC)\footnote{\url{https://ardc.edu.au}} \cite{doi:10.5334/dsj-2019-044} is a
+Commons (ARDC)\footnote{\url{https://ardc.edu.au/}} \cite{doi:10.5334/dsj-2019-044} is a
 national initiative that enables local researchers and industries to
@@
 necessary structure to make all research artefacts FAIR. The Dataverse
-software already has extensive support for metadata, including the Data
+software already has extensive support for metadata\footnote{\url{https://guides.dataverse.org/en/latest/user/appendix.html}}, including the Data
 Documentation Initiative (DDI), Dublin Core, DataCite, and Schema.org.
@@
 such as ROHub used by Earth Sciences
-\cite{doi:10.1016/j.future.2019.03.046}, which provides a interactive
+\cite{doi:10.1016/j.future.2019.03.046}, which provides a 
 user-interface for making Research Objects, along with Research Object
@@
 the life sciences. Workflows might be used primarily to improve
-computational scalability, as well as to also assist in making computed
+computational scalability, as well as to assist in making computed
 data results FAIR \cite{doi:10.1162/dint_a_00033}, for instance by

### 2022-01-02 latex/ro-crate.tex
Small typos

@@
 workflow provenance. Additional metadata can be added manually, e.g.
-for sensitive clinical data that cannot be publicly exposed\footnote{FAIR principle A2: \textit{M etadata are accessible, even when the
+for sensitive clinical data that cannot be publicly exposed\footnote{FAIR principle A2: \textit{Metadata are accessible, even when the
 data are no longer available} \cite{doi:10.1038/sdata.2016.18}.}, or
@@
 by an open and broad community. These guidelines show how to use ``just
-enough'' Linked Data standards in a consistent way. The use of
+enough'' standards in a consistent way. The use of
 structured metadata with a rich base vocabulary can cover
@@
 support, new and improved profiles and tools are being continuously
-added to the RO-Crate tooling landscape, making it easier for adopters
+added to the RO-Crate landscape, making it easier for adopters
 to find examples and support for their own use case.
@@
 RO-Crate are actively welcomed to participate and contribute feedback
-and requirements. In addition we are targeting a wider audience through
+and requirements. In addition, we are targeting a wider audience through
 extensive outreach
@@
 \cite{doi:10.5281/zenodo.4916734}, National Institute of Standards and
-Technology (NIST) on material science and
+Technology (NIST) on material science, and
 InvenioRDM\footnote{\url{https://inveniosoftware.org/products/rdm/}} used by the
@@
 and software packages, tabular data structures; these are not always
-matched well with existing Schema.org types but may benefit from other
+matched well with existing Schema.org types, but may benefit from other
 vocabularies or even separate metadata files, e.g. from Frictionless
-Data.\footnote{\url{https://frictionlessdata.io/}} We will be further aligning with
+Data.\footnote{\url{https://frictionlessdata.io/}} We will be further aligning
 and collaborating with related research artefact description efforts
@@
 
-Author contributions to this article and the RO-Crate projet according
+Author contributions to this article and the RO-Crate project according
 to the Contributor Roles Taxonomy CASRAI

### 2022-01-02 latex/ro-crate.tex
cite RFC3987 where mentioned

@@
 \\
-\mathbb{IRI} &\equiv& { \text{IRIs as defined in RFC3987} }
+\mathbb{IRI} &\equiv& { \text{IRIs as defined in RFC3987 \cite{doi:10.17487/rfc3987}} }
 \\
@@
 The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$
-identifiers \cite{doi:10.17487/rfc3987} (notation \texttt
+identifiers (notation \texttt
 {<http://example.\noxml{\break}com/>}),\footnote{For simplicity, blank nodes are not included in this formalisation, as RO-Crate

### 2022-01-02 latex/ro-crate.tex
URI->IRI.  license() is also a relation.

@@
 The possibilities of consuming\footnote{Some consideration is needed in processing of RO-Crates as
-knowledge graphs, e.g. establishing absolute URIs for files inside a
+knowledge graphs, e.g. establishing absolute IRIs for files inside a
 ZIP archive, detailed in the RO-Crate specification: \url
@@
 practices\footnote{\url{https://www.w3.org/TR/swbp-vocab-pub/}} is to establish a
-vocabulary with a new URI namespace, formalised using RDF
+vocabulary with a new IRI namespace, formalised using RDF
 Schema\footnote{\url{http://www.w3.org/TR/2014/REC-rdf-schema-20140225/}} or
@@
 context,\footnote{\url{https://w3id.org/ro/crate/1.1/context}} which has a flat list
-with the mapping from JSON-LD keys to their URI equivalents (e.g. key
+with the mapping from JSON-LD keys to their IRI equivalents (e.g. key
 \texttt{"author"} maps to the \url{http://schema.org/author} property).
@@
 An $\textit{ROCrate}(R)$ is defined as a self-described \emph{Root Data Entity},
-which describes and contains parts (\emph{data entities}), which are
+which contains parts (\emph{data entities}), which are
 further described in \emph{contextual entities}. These terms align with
@@
 
-The $\textit{Root}(r)$ is a type of $\textit{Dataset}(r)$, and must have the metadata to
-literal attributes to provide a $\textit{name}$, $description$ and
-$\textit{datePublished}$, as well as a contextual entity identifying its
-license. These predicates correspond to the RO-Crate 1.1 requirements
+The $\textit{Root}(r)$ is a type of $\textit{Dataset}(r)$, and must as metadata 
+have at least the attributes $\textit{name}$, $description$ and
+$\textit{datePublished}$ as literal values, as well as a contextual entity that identify its
+$\textit{license}$. These predicates correspond to the RO-Crate 1.1 minimal requirements
 for the root data entity.\footnote{\url
@@
 The below is an example RO-Crate represented using the above
-formalisation, assuming a base URI of \texttt{<http://example.com/ro/123/>}:
+formalisation, assuming a base IRI of \texttt{<http://example.com/ro/123/>}:
 %

### 2022-01-02 latex/ro-crate.tex
Some typos. LanguageTag is not used.

@@
 %PDAG%
-$\mathbh{lri}$
+$\mathbh{IRI}$
 %PDAG%
@@
 different serializations.
-Assume a simplified\footnote{This simplification does not cover the extensive list of literal
-datatypes built-in to RDF 1.1, only strings and decimal real numbers.
-Likewise, language of literals are not included.} language $\mathcal{L}_{\textit{RDF}}$
+Assume a simplified\footnote{This simplification and mapping does not cover the extensive list of literal
+datatypes built into RDF 1.1, only strings and decimal real numbers.
+Likewise, \textit{LanguageTag} is deliberately not utillised below.} language $\mathcal{L}_{\textit{RDF}}$
 based on the RDF abstract syntax \cite{rdfworkinggroup_2014}:
@@
 \\
-\textit{Class}(c) & \Rightarrow& \textit{type}(c, \texttt {<http://www.w3.org/2000/01/rdf-schema
-\#Class>})
+\textit{Class}(c) & \Rightarrow& \textit{type}(c, \texttt {<http://www.w3.org/2000/01/rdf-schema\#Class>})
 \\

### 2022-01-02 latex/ro-crate.tex
Formatting in appendix

@@
 An important RO-Crate principle is that of being \textbf
-{self-described} Therefore the serialisation of the RO-Crate into a
+{self-described}. Therefore the serialisation of the RO-Crate into a
 file should also describe itself in a Metadata File Descriptor,\footnote{\url
@@
 \textit{MetadataFile}(m) & \Rightarrow& \textit{CreativeWork}(m) \land
-\textit{about}(m,R) {\wedge } \textit{ROCrate}(R) \land
+\textit{about}(m,R) \land \textit{ROCrate}(R) \land
 \\
@@
 
-In the conceptual model the \emph{RO-Crate Metadata File} can be seen
+In the conceptual model, the \emph{RO-Crate Metadata File} can be seen
 as the top-level node that describes the \emph{RO-Crate Root}, however
@@
 \\
-\forall s \forall p \forall o \ \textit{Relation}(s,p,o) & \Rightarrow& \texttt{
-\{"@id":} s \texttt{,}
-\\
-&& \ \ p \texttt{: \{ "@id":} o \texttt{\}}
+\forall s \forall p \forall o \ \textit{Relation}(s,p,o) & \Rightarrow& 
+\texttt{\{"@id":} s \texttt{,}
+\\
+&& \ \ p \texttt{: \{"@id":} o \texttt{\}}
 \\
@@
 \\
-&& \ \ \ \ \textit{Mentions}(r, c)*
+&& \ \ \ \ \ \ \textit{Mentions}(r, c)*
 \\

### 2022-01-02 latex/ro-crate.tex
minor typos
bibliography fixes

@@
 Mapper \cite{doi:10.5281/zenodo.3922136}. A similar mapping has been
-implemented by \texttt{RO-Crate\_2\_ma-DMP} \cite{doi:10.5281/zenodo.3903463}. In
+implemented by \textit{RO-Crate\_2\_ma-DMP} \cite{doi:10.5281/zenodo.3903463}. In
 both cases, a maDMP can be converted to a RO-Crate, or vice versa. In
@@
 \\
-\quad \text{Creative Commons Attribution 4.0''}\bigr)
+\quad \text{``Creative Commons Attribution 4.0''}\bigr)
 \\
@@
 \\
-\quad \text{Creative Commons Attribution Non Commercial 4.0''}
+\quad \text{``Creative Commons Attribution Non Commercial 4.0''}
 \bigr)
@@
 \\
-\forall r \forall c\ \textit{ROCrate}(r) & \Rightarrow& \texttt{\{
-"@graph": [}
+\forall r \forall c\ \textit{ROCrate}(r) & \Rightarrow& 
+\texttt{\{"@graph": [}
 \\
@@
 provenance, analysis, and results},
-\bjtitle{PLoS Biology}
+\bjtitle{PLOS Biology}
 \bvolume{16}(\bissue{12})
@@
 provenance, analysis, and results.},
-{PLoS} Biology
+{PLOS} Biology
 16(12)
@@
 pathogen threats require open data and open analytics},
-\bjtitle{PLoS Pathogens}
+\bjtitle{PLOS Pathogens}
 \bvolume{16}(\bissue{8})
@@
 pathogen threats require open data and open analytics.},
-{PLoS} Pathogens
+{PLOS} Pathogens
 16(8)
@@
 \bauthor{\binits{A.}~\bsnm{Treloar}},
-\batitle{The Australian research data commons},
+\batitle{The Australian Research Data Commons},
 \bjtitle{Data Science Journal}
@@
 \bjtitle{Web Semantics: Science, Services and Agents on the World Wide Web}
-\bvolume{32}(\bissue{0})
+\bvolume{32}
 (\byear{2015}),
@@
 Web Semantics: Science, Services and Agents on the World Wide Web
-32(0)
+32
 (2015),
@@
 \bauthor{\binits{M.}~\bsnm{Crosas}},
-\batitle{The dataverse network: An open-source application for sharing,
+\batitle{The DataVerse Network: An open-source application for sharing,
 discovering and preserving data},
@@
 \bpublisher{Communications of the ACM},
-\byear{2021}.
+\byear{2022}.
 \bcomment{Accepted}.
@@
 Communications of the ACM
-(2021),
+(2022),
 Accepted.
@@
 \bvolume{3}
-(\byear{2019}.
+(\byear{2019}).
 \bid{doi={10.3897/biss.3.37080}}
@@
 Internationalized resource identifiers ({IRI}s),
-Technical Report,
-2005.
+RFC 3987,
+\bsertitle{Internet Requests for Comments},
+\bpublisher{RFC Editor},
+\byear(2005).
 \bid{doi={10.17487/rfc3987}}
@@
 Internationalized resource identifiers ({IRI}s),
-Technical Report,
+RFC 3987,
 2005.
@@
 \batitle{A case for data commons: Toward data science as a service},
-\bjtitle{Computing in science \& engineering}
+\bjtitle{Computing in Science \& Engineering}
 \bvolume{18}(\bissue{5})
@@
 \oauthor{\binits{L.}~\bsnm{McGibbney}},
-Science-on-Schema.org v1.2.0, Technical Report,
+Science-on-Schema.org v1.2.0, 
 2021.
@@
 Science-on-Schema.org v1.2.0,
-Technical Report,
 2021.
@@
 The {BagIt} File Packaging Format,
-(V1.0), Technical Report,
-2018.
+(V1.0), 
+RFC 8493,
+\bsertitle{Internet Requests for Comments},
+\bpublisher{RFC Editor},
+(\byear{2018}).
 \bid{doi={10.17487/RFC8493}}
@@
 \batitle{Toward enabling reproducibility for data-intensive research
-using the whole tale platform},
+using the Whole Tale platform},
 \bjtitle{Advances in Parallel Computing 36 (Parallel Computing:
@@
 \batitle{Towards {FAIR} principles for research software},
-\bjtitle{D\'{e}viance et soci\'{e}t\'{e}}
+\bjtitle{Data Science},
+\bvolume{3}(\bissue{1})
 (\byear{2019}),
@@
 Towards {FAIR} principles for~research~software,
-\textit{D\'{e}viance et soci\'{e}t\'{e}}, \textbf{00},
+\textit{Data Science}, \textbf{3},
 (2019),
@@
 PROV-O: The PROV Ontology,
-Technical Report, W3C Recommendation Note 30,
+W3C Recommendation 30,
 2013.
@@
 PROV-O: The PROV Ontology,
-Technical Report,
 2013,
-W3C Recommendation Note 30 April 2013.
+W3C Recommendation 30 April 2013.
 \url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
@@
 science data},
-\bjtitle{PLoS Biology}
+\bjtitle{PLOS Biology}
 \bvolume{15}(\bissue{6})
@@
 persistent identifiers to maximize utility and impact of life science data.},
-{PLoS} Biology
+{PLOS} Biology
 15(6)
@@
 \batitle{Ten principles for machine-actionable data management plans},
-\bjtitle{PLoS Computational Biology}
+\bjtitle{PLOS Computational Biology}
 \bvolume{15}(\bissue{3})
@@
 \textit{Ten principles for machine-actionable data management plans},
-{PLoS} Computational Biology
+{PLOS} Computational Biology
 15(3)
@@
 p.~240.
-ISBN 9781315351148.
+ISBN \bisbn{9781315351148}.
 \end{botherref}
@@
 \bauthor{\binits{S.}~\bsnm{Bechhofer}} and
-\bauthor{\binits{D.D.}~\bsnm{Roure}},
+\bauthor{\binits{D.}~\bsnm{De Roure}},
 \bctitle{{myExperiment}: An ontology for e-Research},
@@
 \batitle{Ten simple rules for reproducible computational research},
-\bjtitle{PLoS Computational Biology}
+\bjtitle{PLOS Computational Biology}
 \bvolume{9}(\bissue{10})
@@
 \textit{Ten simple rules for reproducible computational research.},
-{PLoS} Computational Biology
+{PLOS} Computational Biology
 9(10)
@@
 {RO}-Crate Metadata Specification 1.0,
-Technical Report,
 2019.
@@
 {RO}-Crate Metadata Specification 1.1.1,
-Technical Report,
 2021.
@@
 {RO}-Crate Metadata Specification 1.1,
-Technical Report,
 2020.
@@
 and stewardship},
-\bjtitle{Scientific data}
+\bjtitle{Scientific Data}
 \bvolume{3}

### 2022-01-02 latex/ro-crate.tex
break sentence

@@
 Data in JSON-LD format \cite{sporny_2014}. This is a W3C standard RDF serialisation that has become
-popular as it is easy to read by humans while also offers some
+popular; it is easy to read by humans while also offering some
 advantages for data exchange on the Internet. JSON-LD, a subset of the

### 2022-01-02 latex/ro-crate.tex
final bibliography fixes

@@
 RFC 3987,
-\bsertitle{Internet Requests for Comments},
-\bpublisher{RFC Editor},
+Internet Requests for Comments,
+RFC Editor,
 \byear(2005).
@@
 RFC 8493,
-\bsertitle{Internet Requests for Comments},
-\bpublisher{RFC Editor},
+Internet Requests for Comments,
+RFC Editor,
 (\byear{2018}).
@@
 PROV-O: The PROV Ontology,
-W3C Recommendation 30,
-2013.
+W3C Recommendation 30 April \byear{2013}.
 \url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
@@
 %
-\begin{botherref}
+\begin{bbook}
 \oauthor{\binits{B.}~\bsnm{Mons}},
-Data Stewardship for Open Science,
+\bbtitle{Data Stewardship for Open Science},
 1st edn,
-Taylor \& Francis,
+\bpublisher{Taylor \& Francis},
 p.~240.
 ISBN \bisbn{9781315351148}.
-\end{botherref}
+\end{bbook}
 %

### 2022-01-02 latex/ro-crate-proof.tex
ro-crate-proof separate

--- 2022-01-02 latex/ro-crate.tex
@@
 %
-\begin{barticle}
-\bauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
+\begin{bchapter}
+\bauthor{\binits{E.}~\bsnm{\'{O}~Carrag\'{a}in}},
 \bauthor{\binits{C.}~\bsnm{Goble}},
@@
 \bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\batitle{A lightweight approach to research object data packaging},
-\bjtitle{Zenodo}
+\bctitle{A lightweight approach to research object data packaging}.
+in: \bbtitle{Bioinformatics Open Source Conference (BOSC2019)}, 
+2019-07-24/2019-07-25, 
+\blocation{Basel, Switzerland}.
+\bpublisher{Zenodo}
 (\byear{2019}).
 \bid{doi={10.5281/zenodo.3250687}}
-\end{barticle}
+\end{bchapter}
 %
@@
 \bauthor{\binits{P.}~\bsnm{Sefton}},
-in: \bbtitle{Datacrate Submisssion to the Workshop on Research Objects},
-\bconflocation{Zenodo},
-\byear{2018}.
+\bauthor{\binits{G.}~\bsnm{Devine}},
+\bauthor{\binits{C.}~\bsnm{Evenhuis}},
+\bauthor{\binits{M.}~\bsnm{Lynch}},
+\bauthor{\binits{S.}~\bsnm{Wise}},
+\bauthor{\binits{M.}~\bsnm{Lake}} and
+\bauthor{\binits{D.}~\bsnm{Loxton}}: 
+\bctitle{DataCrate: a method of packaging, distributing, displaying and archiving Research Objects}.
+in: \bbtitle{Workshop on Research Objects (RO 2018)},
+29 Oct \byear{2018} at
+IEEE eScience 2018,
+\bconflocation{Amsterdam, Netherland}.
+\bpublisher{Zenodo}
 \bid{doi={10.5281/zenodo.1445817}}
@@
 \oauthor{\binits{P.}~\bsnm{Sefton}},
-\oauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
+\oauthor{\binits{E.}~\bsnm{\'{O}~Carrag\'{a}in}},
 \oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},

### 2022-01-02 latex/ro-crate.tex
swapped back ro-crate-from-markdown

--- 2022-01-02 latex/ro-crate.tex
@@
-%BeginFileInfo
-%%Publisher=IOS
-%%Project=DS
-%%Manuscript=DS210053
-%%MS position=
-%%Stage=201
-%%TID=imikolaityte
-%%Pages=0
-%%Format=2019
-%%Distribution=vtex
-%%Destination=DVI
-%%PDF type=
-%%DVI.Maker=luadvi
-%%PS.Maker=luaps
-%%History1=2021.12.28 09:11
-%EndFileInfo
+%% ro-crate paper
+%% (c) 2020-2021 The University of Manchester
+%% 
+%% SPDX-License-Identifier: CC-BY-4.0
+
+% add. options: [seceqn,secthm,crcready]
+\documentclass[ds,crcready]{iosart2x}
+
+%%%%%%%%%%% Put your definitions here
+
+% From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
+\usepackage{fontspec}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{hyperref}
+\usepackage{xcolor}
+\definecolor{darkerblue}{cmyk}{1,1,0,0.45}
+\hypersetup{
+    colorlinks=true,
+    linkcolor=black,
+    citecolor=black,
+    urlcolor=darkerblue}
+
+\usepackage{unicode-math}
+\usepackage{graphicx}
+
+% Hyperlinks on DOI permalinks – just 20 years after DOI invention!
 % 
-% Journal: Data Science (DS), IOS Press
-% Typeset in LaTeX2e by VTeX Ltd., Vilnius, Lithuania
-%
-%Spelling_date
-%\RequirePackage{float} % when algorithm.sty used in manuscript
-\documentclass[ds,v1.1.2,openaccess]{iosart2x}%% ,seceqn
-\usepackage{mathbh}
-%
-%\usepackage{}
-\usepackage{tmultirow}
-\usepackage{listings}
-%
-\PROOF
-%\PROOF
-%\CRC
-
-\volume{-1}
-\issue{-1}
+% We had to make our own copy of 
+% https://ctan.org/pkg/doi?lang=en 
+% where commands are renamed to \doix
+% because iosart2x.cls and ios1.bst have 
+% a different use of \doiurl \doi etc.
+% We also set the doixtext prefix empty
+% to avoid doi:doi:
+\usepackage{doix}
+\renewcommand{\doixtext}{}
+\renewcommand{\doiurl}[1]{\doix{#1}}
+
+\DeclareGraphicsExtensions{.pdf,.png,.svg}
+\graphicspath{../content/images/}
+\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
+\def\markdownOptionOutputDir{../output}
+\def\markdownOptionHeaderAttributes{true}
+
+% for table resizing
+\usepackage{booktabs}
+\usepackage{multirow}
+
+%%%%%%%%%%% End of definitions
+
 \pubyear{2021}
+\volume{0}
 \firstpage{1}
-\lastpage{42}
-\articletype{research-article} %%
-%http://www.vtex.vmt/docs/wiki/index.php/IOS_Press_instrukcijos_%28Kaunas%29#Straipsnio_tipas_.28.5Carticletype.7B.7D.29
-\aid{DS210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
-\doi{10.3233/DS-210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
-\vtexed{Ingrida}
-
-\startlocaldefs
-\setpkgattr{license}{type}{CC@BY} %% CC@BY@NC
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% indentation after theorem-like environments - auto
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\newtheorem*{}{} %% unnumbered
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\theoremstyle{plain}
-%%\newtheorem{}{}%[]
-%%\theoremstyle{definition}
-%%\newtheorem{}{}%[]
-\endlocaldefs
-%
+\lastpage{1}
+
 \begin{document}
-%
+
+% Welcome to the 2000s with UNICODE fonts!
+% Adapted from
+% https://www.overleaf.com/learn/latex/Articles%2FOpenType-based_math_typesetting:_An_introduction_to_the_STIX2_OpenType_fonts
+\setmathfont{STIX2Math}[
+    Extension={.otf},
+    Path=./STIX2fonts/,
+    Scale=1
+]
+\setmainfont{STIX2Text}[
+    Extension={.otf},
+    Path=./STIX2fonts/,
+    UprightFont={*-Regular},
+    BoldFont={*-Bold},
+    ItalicFont={*-Italic},
+    BoldItalicFont={*-BoldItalic}
+]
+
+% https://typeof.net/Iosevka/ mono font has 
+% all the logic symbols we need
+%\setmonofont{iosevka-slab}[
+%    Extension={.ttf},
+%    Path=./ttf-iosevka-slab-6.1.3/,
+%    Scale=MatchLowercase,
+%    UprightFont={*-light},
+%    ItalicFont={*-lightitalic}
+%]
+
 \begin{frontmatter}
-%
-\pretitle[XML]{Research Article}
+
 %\pretitle{}
 \title{Packaging research artefacts with RO-Crate}
-%\title{\thanks{}}
 \runtitle{Packaging research artefacts with RO-Crate}
 %\subtitle{}
-%
-%%% initials without spaces
-%
-% {\inits{}\fnms{} \snm{} \suffix{}\ead[label=e?]{}
-%\ead[label=p1,phone]{}\ead[label=f1,fax]{}\ead[label=o?,orcid]{}
-%\thanks{}}
-%
-\begin{aug}
-\author[A,B]{\inits{S.}\fnms{Stian} \snm{Soiland-Reyes}\ead
-[label=e1]{soiland-reyes@manchester.ac.uk}\ead
-[label=o1,orcid]{0000-0001-9842-9718}\thanks{Corresponding author.
-\printead{e1}.}}
-\author[C]{\inits{P.}\fnms{Peter} \snm{Sefton}\ead
-[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}}
-\author[D]{\inits{M.}\fnms{Merc\`{e}} \snm{Crosas}\ead
-[label=o3,orcid]{0000-0003-1304-1939}}
-\author[E]{\inits{L.J.}\fnms{Leyla Jael} \snm{Castro}\ead
-[label=o4,orcid]{0000-0003-3986-0510}}
-\author[F]{\inits{F.}\fnms{Frederik} \snm{Coppens}\ead
-[label=o5,orcid]{0000-0001-6565-5145}}
-\author[G]{\inits{J.M.}\fnms{Jos\'{e} M.} \snm{Fern\'{a}ndez}\ead
-[label=o6,orcid]{0000-0002-4806-5140}}
-\author[H]{\inits{D.}\fnms{Daniel} \snm{Garijo}\ead
-[label=o7,orcid]{0000-0003-0454-7145}}
-\author[I]{\inits{B.}\fnms{Bj\"{o}rn} \snm{Gr\"{u}ning}\ead
-[label=o8,orcid]{0000-0002-3079-6586}}
-\author[J]{\inits{M.}\fnms{Marco} \snm{La Rosa}\ead
-[label=o9,orcid]{0000-0001-5383-6993}}
-\author[K]{\inits{S.}\fnms{Simone} \snm{Leo}\ead
-[label=o10,orcid]{0000-0001-8271-5429}}
-\author[L]{\inits{E.}\fnms{Eoghan} \snm{\'{O} Carrag\'{a}in}\ead
-[label=o11,orcid]{0000-0001-8131-2150}}
-\author[M]{\inits{M.}\fnms{Marc} \snm{Portier}\ead
-[label=o12,orcid]{0000-0002-9648-6484}}
-\author[N]{\inits{A.}\fnms{Ana} \snm{Trisovic}\ead
-[label=o13,orcid]{0000-0003-1991-0533}}
-\author[O]{\snm{RO-Crate Community}\ead
-[label=u1,url]{https://www.researchobject.org/ro-crate/community}}
-\author[P]{\inits{P.}\fnms{Paul} \snm{Groth}\ead
-[label=e15]{p.t.groth@uva.nl}\ead[label=o15,orcid]{0000-0003-0183-6910}}
-\author[Q]{\inits{C.}\fnms{Carole} \snm{Goble}\ead
-[label=e16]{carole.goble@manchester.ac.uk}\ead
-[label=o16,orcid]{0000-0003-1219-2137}}
-%
-%%%%%% printead order: TEL FAX E-MAIL URL
-%\thanks{Corresponding author: name surname, address. \printead{}.}
-%\thanks{Corresponding author. \printead{}.}
-%\thanks{}%% Move support info to acks
-%
-\runauthor{S. Soiland-Reyes et al.} %% auto
-% {\orgname{} \cny{}\printead[presep={\\}]{}}
-\address[A]{Department of Computer Science, \orgname{The University of
-Manchester}, \cny{UK}\printead[presep={\\}]{e1,o1}}
-\address[B]{Informatics Institute, \orgname{University of Amsterdam},
+
+% For one author:
+%\author{\inits{N.}\fnms{Name1} \snm{Surname1}\ead[label=e1]{first@somewhere.com}\ead[label=o1,orcid]{0000-0000-0000-0001}}
+%\address{Department first, \orgname{University or Company name},
+%Abbreviate US states, \cny{Country}\printead[presep={\\}]{e1,o1}}
+
+% Two or more authors:
+\author[SSR1,SSR2]{\inits{S.}\fnms{Stian}
+    \snm{Soiland-Reyes}\ead[label=eSSR]{soiland-reyes@manchester.ac.uk}\ead[label=oSSR,orcid]{0000-0001-9842-9718}%
+    \thanks{Corresponding author. \printead{eSSR}.}}, 
+\author[PS]{\inits{P.}\fnms{Peter}
+    \snm{Sefton}\ead[label=ePS]{Peter.Sefton@uts.edu.au}\ead[label=oPS,orcid]{0000-0002-3545-944X}},
+\author[MC]{\inits{M.}\fnms{Mercè}
+    \snm{Crosas}\ead[label=eMC]{mcrosas@g.harvard.edu}\ead[label=oMC,orcid]{0000-0003-1304-1939}},
+\author[LJC]{\inits{LJ.}\fnms{Leyla Jael}
+    \snm{Castro}\ead[label=eLJC]{ljgarcia@zbmed.de}\ead[label=oLJC,orcid]{0000-0003-3986-0510}},
+\author[FC]{\inits{F.}\fnms{Frederik}
+    \snm{Coppens}\ead[label=eFC]{frcop@psb.vib-ugent.be}\ead[label=oFC,orcid]{0000-0001-6565-5145}},
+\author[JMF]{\inits{JM.}\fnms{José M.}
+    \snm{Fernández}\ead[label=eJMF]{jose.m.fernandez@bsc.es}\ead[label=oJMF,orcid]{0000-0002-4806-5140}},
+\author[DG]{\inits{D.}\fnms{Daniel}
+    \snm{Garijo}\ead[label=eDG]{dgarijo@fi.upm.es}\ead[label=oDG,orcid]{0000-0003-0454-7145}},
+\author[BG]{\inits{B.}\fnms{Björn}
+    \snm{Grüning}\ead[label=eBG]{gruening@informatik.uni-freiburg.de}\ead[label=oBG,orcid]{0000-0002-3079-6586}},
+\author[MLR]{\inits{M.}\fnms{Marco}
+    \snm{La Rosa}\ead[label=eMLR]{m@lr.id.au}\ead[label=oMLR,orcid]{0000-0001-5383-6993}},
+\author[SL]{\inits{S.}\fnms{Simone}
+    \snm{Leo}\ead[label=eSL]{simone.leo@crs4.it}\ead[label=oSL,orcid]{0000-0001-8271-5429}},
+\author[EOC]{\inits{E.}\fnms{Eoghan}
+    \snm{Ó Carragáin}\ead[label=eEOC]{eoghan.ocarragain@ucc.ie}\ead[label=oEOC,orcid]{0000-0001-8131-2150}},
+\author[MP]{\inits{M.}\fnms{Marc}
+    \snm{Portier}\ead[label=eMP]{marc.portier@vliz.be}\ead[label=oMP,orcid]{0000-0002-9648-6484}},
+\author[AT]{\inits{A.}\fnms{Ana}
+    \snm{Trisovic}\ead[label=eMP]{ana.trisovic@gmail.com}\ead[label=oAT,orcid]{0000-0003-1991-0533}},
+\author[RO]{RO-Crate Community},
+\author[PG]{\inits{P.}\fnms{Paul}
+    \snm{Groth}\ead[label=ePG]{p.t.groth@uva.nl}\ead[label=oPG,orcid]{0000-0003-0183-6910}},
+\author[CG]{\inits{C.}\fnms{Carole}
+    \snm{Goble}\ead[label=eCG]{carole.goble@manchester.ac.uk}\ead[label=oCG,orcid]{0000-0003-1219-2137}}
+
+\address[SSR1]{Department of Computer Science, 
+    \orgname{The University of Manchester}, \cny{UK}\printead[presep={\\}]{eSSR,oSSR}}
+\address[SSR2]{Informatics Institute, \orgname{University of Amsterdam},
     \cny{The Netherlands}}
-\address[C]{Faculty of Science, \orgname{University Technology Sydney},
-\cny{Australia}\printead[presep={\\}]{e2,o2}}
-\address[D]{Institute for Quantitative Social Science, \orgname{Harvard
-University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{o3}}
-\address[E]{\orgname{ZB MED Information Centre for Life Sciences},
-Cologne, \cny{Germany}\printead[presep={\\}]{o4}}
-\address[F]{\orgname{VIB-UGent Center for Plant Systems Biology}, Gent,
-\cny{Belgium}\printead[presep={\\}]{o5}}
-\address[G]{\orgname{Barcelona Supercomputing Center}, Barcelona, \cny
-{Spain}\printead[presep={\\}]{o6}}
-\address[H]{Ontology Engineering Group, \orgname{Universidad Polit\'
-{e}cnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{o7}}
-\address[I]{Bioinformatics Group, Department of Computer Science,
-\orgname{Albert-Ludwigs-University Freiburg}, Freiburg, \cny
-{Germany}\printead[presep={\\}]{o8}}
-\address[J]{\orgname{PARADISEC}, Melbourne, \cny{Australia}\printead
-[presep={\\}]{o9}}
-\address[K]{\orgname{Center for Advanced Studies, Research, and
-Development in Sardinia} (CRS4), Pula (CA), \cny{Italy}\printead
-[presep={\\}]{o10}}
-\address[L]{\orgname{University College Cork}, \cny{Ireland}\printead
-[presep={\\}]{o11}}
-\address[M]{\orgname{Vlaams Instituut voor de Zee}, Oostende, \cny
-{Belgium}\printead[presep={\\}]{o12}}
-\address[N]{Institute for Quantitative Social Science, \orgname{Harvard
-University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{o13}}
-\address[O]{\printead{u1}~(see Appendix B)}
-\address[P]{Informatics Institute, \orgname{University of Amsterdam},
-\cny{The Netherlands}\printead[presep={\\}]{e15,o15}}
-\address[Q]{Department of Computer Science, \orgname{The University of
-Manchester}, \cny{UK}\printead[presep={\\}]{e16,o16}}
-\end{aug}
-%
-%%%%%%
-%
-\begin{review}{editor}
-\reviewer{\fnms{Silvio} \snm{Peroni}\ead
-[label=oe1,orcid]{0000-0003-0530-4305}\printead[presep={\ }]{oe1}}
-%\reviewer{\fnms{} \snm{}\ead[label=oe2,orcid]{}\printead[presep={\
-%}]{oe2}}
-\end{review}
-
-\begin{review}{solicited}
-\reviewer{\fnms{Angelo A.} \snm{Salatino}\ead
-[label=os1,orcid]{0000-0002-4763-3943}\printead[presep={\ }]{os1}}
-\reviewer{\fnms{Alessia} \snm{Bardi}\ead
-[label=os2,orcid]{0000-0002-1112-1292}\printead[presep={\ }]{os2}}
-\reviewer{\snm{one anonymous reviewer}}
-\end{review}
-%
-%\dedicated{}
-%
-% HISTORY:
-\received{\sday{13} \smonth{8} \syear{2021}}
-\accepted{\sday{23} \smonth{9} \syear{2021}}
-%
+\address[PS]{Faculty of Science, \orgname{University Technology Sydney},
+    \cny{Australia}\printead[presep={\\}]{ePS,oPS}}
+\address[MC]{Institute for Quantitative Social Science, 
+    \orgname{Harvard University}, Cambridge, MA,  \cny{USA}\printead[presep={\\}]{oMC}}
+\address[LJC]{\orgname{ZB MED Information Centre for Life Sciences}, 
+    Cologne, \cny{Germany}\printead[presep={\\}]{oLJC}}
+\address[FC]{\orgname{VIB-UGent Center for Plant Systems Biology}, 
+    Gent, \cny{Belgium}\printead[presep={\\}]{oFC}}
+\address[JMF]{\orgname{Barcelona Supercomputing Center}, 
+    Barcelona, \cny{Spain}\printead[presep={\\}]{oJMF}}
+\address[DG]{Ontology Engineering Group, 
+    \orgname{Universidad Politécnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{oDG}}
+\address[BG]{Bioinformatics Group, Department of Computer Science,
+    \orgname{Albert-Ludwigs-University Freiburg}, Freiburg, \cny{Germany}\printead[presep={\\}]{oBG}}
+\address[MLR]{\orgname{PARADISEC}, Melbourne, \cny{Australia}\printead[presep={\\}]{oMLR}}
+\address[SL]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA),
+    \cny{Italy}\printead[presep={\\}]{oSL}}
+\address[EOC]{\orgname{University College Cork},
+    \cny{Ireland}\printead[presep={\\}]{oEOC}}
+\address[MP]{\orgname{Vlaams Instituut voor de Zee}, Oostende,
+    \cny{Belgium}\printead[presep={\\}]{oMP}}
+\address[AT]{Institute for Quantitative Social Science,
+    \orgname{Harvard University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{oAT}}
+\address[RO]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
+\address[PG]{Informatics Institute, \orgname{University of Amsterdam},
+    \cny{The Netherlands}\printead[presep={\\}]{ePG,oPG}}
+\address[CG]{Department of Computer Science, \orgname{The University of Manchester},
+    \cny{UK}\printead[presep={\\}]{eCG,oCG}}
+
+%\begin{review}{editor}
+%\reviewer{\fnms{First} \snm{Editor}\ead[label=oe1,orcid]{}\printead[presep={\ }]{oe1}}
+%\reviewer{\fnms{Second} \snm{Editor}\ead[label=oe2,orcid]{}\printead[presep={\ }]{oe2}}
+%\end{review}
+%\begin{review}{solicited}
+%\reviewer{\fnms{First} \snm{Solicited reviewer}\ead[label=os1,orcid]{}\printead[presep={\ }]{os1}}
+%\end{review}
+
 \begin{abstract}
-%\parttitle{}
-An increasing number of researchers support reproducibility by
-including pointers to and
-descriptions of datasets, software and methods in their publications.
-However, scientific
-articles may be ambiguous, incomplete and difficult to process by
-automated systems. In
-this paper we introduce RO-Crate, an open, community-driven, and
-lightweight approach
-to packaging research artefacts along with their metadata in a machine
-readable manner.
-RO-Crate is based on Schema.org annotations in JSON-LD, aiming to
-establish best
-practices to formally describe metadata in an accessible and practical
-way for their
-use in a wide variety of situations.
-
-An RO-Crate is a structured archive of all the items that contributed
-to a research
-outcome, including their identifiers, provenance, relations and
-annotations. As a general
-purpose packaging approach for data and their metadata, RO-Crate is
-used across multiple
-areas, including bioinformatics, digital humanities and regulatory
-sciences. By applying
-``just enough'' Linked Data standards, RO-Crate simplifies the process of
-making research outputs FAIR while also enhancing research reproducibility.
-
-An RO-Crate for this article\footnote{\url
-{https://w3id.org/ro/doi/10.5281/zenodo.5146227}}
-is archived at \url{https://doi.org/10.5281/zenodo.5146227}.
+\markdownInput{../content/01.abstract.md}
 \end{abstract}
-%
-\begin{keyword} %% 1 u.c.
+
+\begin{keyword}
 \kwd{Data publishing}
-\kwd{data packaging}
+\kwd{Data packaging}
 \kwd{FAIR}
 \kwd{Linked Data}
-\kwd{metadata}
-\kwd{reproducibility}
-\kwd{research object}
+\kwd{Metadata}
+\kwd{Reproducibility}
+\kwd{Research Object}
 \end{keyword}
-%
+
 \end{frontmatter}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%% Main text entry area:
-\section{Introduction}
-
-\label{sec:introduction}
-
-The move towards Open Science has increased the need and demand for the
-publication of artefacts of the research process
-\cite{sefton_blog_post_2021}. This is particularly apparent in domains that
-rely on computational experiments; for example, the publication of
-software, datasets and records of the dependencies that such
-experiments rely on \cite{doi:10.1126/science.aah6168}.
-
-It is often argued that the publication of these assets, and
-specifically software
-\cite{doi:10.3233/DS-190026}, workflows \cite{doi:10.1162/dint_a_00033} and
-data, should follow the FAIR principles \cite{doi:10.1038/sdata.2016.18};
-namely, that they are Findable, Accessible, Interoperable and Reusable.
-These principles are agnostic to the \textit{implementation} strategy needed
-to comply with them. Hence, there has been an increasing amount of work
-in the development of platforms and specifications that aim to fulfil
-these goals \cite{isbn:9781315351148}.
-
-Important examples include data publication with rich metadata (e.g.
-Zenodo \cite{doi:10.3897/biss.3.37080}), domain-specific data deposition
-(e.g. PDB \cite{doi:10.1093/nar/gkl971}) and following practices for
-reproducible research software \cite{doi:10.1371/journal.pcbi.1003285}
-(e.g. use of containers). While these platforms are useful, experience
-has shown that it is important to put greater emphasis on the
-interconnection of the multiple artefacts that make up the research
-process \cite{doi:10.1016/j.ijhcs.2020.102562}.
-
-The notion of \textbf{Research Objects} \cite{doi:10.1016/j.future.2011.08.004}
-(RO) was introduced to address this connectivity, providing
-semantically rich \textit{aggregations} of (potentially distributed) resources
-with a layer of structure over a research study; this is then to be
-delivered in a \textit{machine-readable format}.
-
-A Research Object combines the ability to bundle multiple types of
-artefacts together, such as spreadsheets, code, examples, and figures.
-The RO is augmented with annotations and relationships that describe
-the artefacts' \textit{context} (e.g. a CSV being used by a script, a figure
-being a result of a workflow).
-
-This notion of ROs provides a compelling vision as an approach for
-implementing FAIR data. However, existing Research Object
-implementations require a large technology stack
-\cite{doi:10.1016/j.websem.2015.01.003}, are typically tailored to a
-particular platform and are also not easily usable by end-users.
-
-To address this gap, a new community came together
-\cite{doi:10.5281/zenodo.3250687} to develop \textbf{RO-Crate} -- an \textit{approach
-to package and aggregate research artefacts with their metadata and
-relationships}. The aim of this paper is to introduce RO-Crate and
-assess it as a strategy for making multiple types of research artefacts
-FAIR. Specifically, the contributions of this paper are as follows:
-%
-\begin{enumerate}
-\item[1.] An introduction to RO-Crate, its purpose and context;
-\item[2.] A guide to the RO-Crate community and tooling;
-\item[3.] Examples of RO-Crate usage, demonstrating its value as connective
-tissue for different artefacts from different communities.
-\end{enumerate}
-
-The rest of this paper is organised as follows. We first describe
-RO-Crate through its development methodology that formed the RO-Crate
-concept, showing its foundations in Linked Data and emerging
-principles. We then define RO-Crate technically, before we introduce
-the community and tooling. We move to analyse RO-Crate with respect to
-usage in a diverse set of domains. Finally, we present related work and
-conclude with some remarks including RO-Crate highlights and future
-work. The appendix adds a formal definition of RO-Crate using
-First-Order logic.
-
-\section{RO-Crate} %{#rocrate}
-
-\label{sec:rocrate}
-
-RO-Crate aims to provide an approach to packaging research artefacts
-with their metadata that can be easily adopted. To illustrate this, let
-us imagine a research paper reporting on the sequence analysis of
-proteins obtained from an experiment on mice. The sequence output
-files, sequence analysis code, resulting data and reports summarising
-statistical measures are all important and inter-related research
-artefacts, and consequently would ideally all be co-located in a
-directory and accompanied with their corresponding metadata. In
-reality, some of the artefacts (e.g. data or software) will be recorded
-as external reference to repositories that are not necessarily
-following the FAIR principles. This conceptual directory, along with
-the relationships between its constituent digital artefacts, is what
-the RO-Crate model aims to represent, linking together all the elements
-of an experiment that are required for the experiment's reproducibility
-and reusability.
-
-The question then arises as to how the directory with all this material
-should be packaged in a manner that is accessible and usable by others.
-This means programmatically and automatically accessible by machines
-and human readable. A de facto approach to sharing collections of
-resources is through compressed archives (e.g. a zip file). This solves
-the problem of ``packaging'', but it does not guarantee downstream
-access to all artefacts in a programmatic fashion, nor describe the
-role of each file in that particular research. Both features, the
-ability to automatically access and reason about an object, are crucial
-and lead to the need for explicit metadata about the contents of the
-folder, describing each and linking them together.
-
-Examples of metadata descriptions across a wide range of
-domains\footnote{\url{https://rdamsc.bath.ac.uk/scheme-index}} abound within the
-literature, both in research data management
-\cite{doi:10.1007/s10209-016-0475-y,farnel_2014,doi:10.2777/620649}
-and within library and information
-systems\footnote{\url{https://www.loc.gov/librarians/standards}} \cite{chan_1995,doi:10.1515/9783598441844}. However, many of these approaches require
-knowledge of metadata schemas, particular annotation systems, or the
-use of complex software stacks. Indeed, particularly within research,
-these requirements have led to a lack of adoption and growing
-frustration with current tooling and specifications
-\cite{neylon_blog_post_2017,doi:10.1007/s00267-014-0258-2,doi:10.1038/s41597-020-0524-5}.
-
-RO-Crate seeks to address this complexity by:
-%
-\begin{enumerate}
-\item[1.] being conceptually simple and easy to understand for developers;
-\item[2.] providing strong, easy tooling for integration into community projects;
-\item[3.] providing a strong and opinionated guide regarding current best practices;
-\item[4.] adopting de-facto standards that are widely used on the Web.
-\end{enumerate}
-
-In the following sections we demonstrate how the RO-Crate specification
-and ecosystem achieve these goals.
-
+
+%%%%%%%%%%% The article body starts:
+
+% Every *.md file in content/ in order, except 00, 90
+\markdownInput{../content/10.background.md}
+
+\markdownInput{../content/20.rocrate.md}
 %% sub-subsections below to cater for figures in 
 %% markdown vs LaTex
-\subsection{Development methodology}
-
-\label{sec:methodology}
-
-It is a good question as to what base level we assume for `conceptually
-simple'. We take simplicity to apply at two levels: for the
-\textit{developers} who produce the platforms and for the \textit{data practitioners}
-and users of those platforms.
-
-For our development methodology we followed the mantra of working
-closely with a small group to get a deep understanding of
-requirements and ensure rapid feedback loops. We created a pool of
-early adopter projects from a range of disciplines and groups,
-primarily addressing developers of platforms. Thus the base level for
-simplicity was \textbf{developer friendliness}.
-
-We assumed a developer familiar with making Web applications with JSON
-data (who would then learn how to make \textit{RO-Crate JSON-LD}), which
-informed core design choices for our JSON-level documentation approach
-and RO-Crate serialization (Section~\ref{sec:implementation}). Our
-group of early adopters, growing as the community evolved, drove the
-RO-Crate requirements and provided feedback through our multiple
-communication channels including bi-monthly meetings, which we describe
-in Section~\ref{sec:community} along with the established norms.
-
-Addressing the simplicity of understanding and engaging with RO-Crate
-by data practitioners is through the platforms, for example with
-interactive tools (Section~\ref{sec:tooling}) like
-Describo\footnote{\url{https://arkisto-platform.github.io/describo/}} \cite{describo} and
-Jupyter notebooks \cite{doi:10.3233/978-1-61499-649-1-87}, and by close
-discussions with domain scientists on how to appropriately capture what
-they determine to be relevant metadata. This ultimately requires a new
-type of awareness and learning material separate from developer
-specifications, focusing on the simplicity of extensibility to serve
-the user needs, along with user-driven development of new RO-Crate
-Profiles specific for their needs (Section~\ref{sec:inuse}).
-
-\subsection{Conceptual definition}% {#conceptual}
-
-\label{sec:conceptual}
-
-A key premise of RO-Crate is the existence of a wide variety of
-resources on the Web that can help describe research. As such, RO-Crate
-relies on the Linked Data principles
-\cite{doi:10.2200/S00334ED1V01Y201102WBE001}. Figure~\ref{fig:conceptual}
-shows the main conceptual elements involved in an RO-Crate: The
-RO-Crate Metadata File (top) describes the Research Object using
-structured metadata including external references, coupled with the
-contained artefacts (bottom) bundled and described by the RO-Crate.
-
-The conceptual notion of a \textit{Research Object}
-\cite{doi:10.1016/j.future.2011.08.004} is thus realised with the RO-Crate
-model and serialised using Linked Data constructs within the RO-Crate
-metadata file.
-
-\begin{figure}%[t!]
-\includegraphics{210053f01}
-\caption{Conceptual overview of RO-Crate. A \emph{Persistent
-Identifier} (PID) \cite{doi:10.1371/journal.pbio.2001414} points to a
-\emph{Research Object} (RO), which may be archived using different
-packaging approaches like BagIt \cite{doi:10.17487/rfc8493}, OCFL \cite
-{ocfl_2020}, git or ZIP. The RO is described within a \emph{RO-Crate
-Metadata File}, providing identifiers for \emph{authors} using ORCID,
-\emph{organisations} using Research Organization Registry (ROR) \cite
-{doi:10.6087/kcse.192} and licences such as Creative Commons using SPDX
-identifiers. The \emph{RO-Crate content} is further described with
-additional metadata following a Linked Data approach. Data can be
-embedded files and directories, as well as links to external Web
-resources, PIDs and nested RO-Crates.}
-\label{fig:conceptual}
-\end{figure}
-
-\subsubsection{Linked Data as a foundation}% {#linkeddata}
-
-\label{sec:linkeddata}
-
-The \textbf{Linked Data} principles \cite{doi:10.4018/978-1-60960-593-3.ch008}
-(use of IRIs\footnote{\textbf{IRI}s \cite{doi:10.17487/rfc3987} are a generalisation
- of \textit{URI}s
-(which include well-known http/https URLs), permitting international
-Unicode characters without percent encoding, commonly used on the
-browser address bar and in HTML5.} to identify resources (i.e. artefacts), resolvable via
-HTTP, enriched with metadata and linked to each other) are core to
-RO-Crate; therefore IRIs are used to identify an RO-Crate, its
-constituent parts and metadata descriptions, and the properties and
-classes used in the metadata.
-
-RO-Crates are \textit{self-described} and follow the Linked Data principles to
-describe all of their resources in both human and machine readable
-manner. Hence, resources are identified using global identifiers
-(absolute IRIs) where possible; and relationships between two resources
-are defined with links.
-
-The foundation of Linked Data and shared vocabularies also means that
-multiple RO-Crates and other Linked Data resources can be indexed,
-combined, queried, validated or transformed using existing Semantic Web
-technologies such as SPARQL,\footnote{\url{https://www.w3.org/TR/sparql11-overview}}
-SHACL\footnote{\url{https://www.w3.org/TR/shacl/}} and well established \textit{knowledge
-graph} triple stores like Apache Jena\footnote{\url{https://jena.apache.org/}} and
-OntoText GraphDB.\footnote{\url{https://www.ontotext.com/products/graphdb/}}
-
-The possibilities of consuming\footnote{Some consideration is needed in processing of RO-Crates as
-knowledge graphs, e.g. establishing absolute IRIs for files inside a
-ZIP archive, detailed in the RO-Crate specification: \url
-{https://www.researchobject.org/ro-crate/1.1/appendix/relative-uris.html}.} RO-Crate metadata with such
-powerful tools gives another strong reason for using Linked Data as a
-foundation. This use of mature Web\footnote{Note that an RO-Crate is not required to be published on the
-Web, see Section~\ref{sec:selfdescribed}.} technologies also means its
-developers and consumers are not restricted to the Research Object
-aspects that have already been specified by the RO-Crate community, but
-can extend and integrate RO-Crate in multiple standardised ways.
-
-\subsubsection{RO-Crate is a self-described container}% {#selfdescribed}
-
-\label{sec:selfdescribed}
-
-An RO-Crate is defined\footnote{\url{https://www.researchobject.org/ro-crate/1.1/structure.html\#ro-crate-metadata-file-ro-crate-metadatajson}}
-as a self-described \textbf{Root Data Entity} that describes and contains
-\textit{data entities}, which are further described by referencing \textit{contextual
-entities}. A \textbf{data entity} is either a \textit{file} (i.e. a byte sequence
-stored on disk somewhere) or a \textit{directory} (i.e. set of named files and
-other directories). A~file does not need to be stored inside the
-RO-Crate root, it can be referenced via a PID/IRI. A \textbf{contextual
-entity} exists outside the information system (e.g. a Person, a
-workflow language) and is stored solely by its metadata. The
-representation of a \textit{data entity} as a byte sequence makes it
-possible to store a variety of research artefacts including not only
-data but also, for instance, software and text.
-
-The Root Data Entity is a directory, the \textit{RO-Crate Root}, identified by
-the presence of the \textbf{RO-Crate Metadata File} \texttt{ro-crate-metadata.json}
-(top of Fig.~\ref{fig:conceptual}). This file
-describes the RO-Crate, its content and related metadata using Linked
-Data in JSON-LD format \cite{sporny_2014}. This is a W3C standard RDF serialisation that has become
-popular; it is easy to read by humans while also offering some
-advantages for data exchange on the Internet. JSON-LD, a subset of the
-widely supported and well-known JSON format, has tooling available for
-many programming languages.\footnote{\url{https://json-ld.org/\#developers}}
-
-The minimal requirements for the root data entity
-metadata\footnote{\url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}}
-are \texttt{name}, \texttt{description} and \texttt{datePublished}, as well as a contextual
-entity identifying its \texttt{license} -- additional metadata are commonly
-added to entities depending on the purpose of the particular RO-Crate.
-
-RO-Crates can be stored, transferred or published in multiple ways,
-e.g. BagIt \cite{doi:10.17487/rfc8493}, Oxford Common File Layout
-\cite{ocfl_2020} (OCFL), downloadable ZIP archives in Zenodo or through
-dedicated online repositories, as well as published directly on the
-Web, e.g. using GitHub Pages.\footnote{\url{https://pages.github.com/}} Combined
-with Linked Data identifiers, this caters for a diverse set of storage
-and access requirements across different scientific domains, from
-metagenomics workflows producing hundreds of gigabytes of genome data
-to cultural heritage records with access restrictions for personally
-identifiable data. Specific \textit{RO-Crate profiles} (Section~\ref{sec:profiles}) may constrain serialization and publication
-expectations, and require additional contextual types and properties.
-
-\subsubsection{Data entities are described using contextual entities}% {#contextualentities}
-
-\label{sec:contextualentities}
-
-RO-Crate distinguishes between data and contextual
-entities\footnote{\url{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html\#contextual-vs-data-entities}}
-in a similar way to HTTP terminology's early attempt to separate
-\textit{information} (data) and \textit{non-information} (contextual) resources
-\cite{httprange14}. Data entities are usually files and directories located
-by relative IRI references within the RO-Crate Root, but they can also
-be Web resources or restricted data identified with absolute IRIs,
-including \textit{Persistent Identifiers} (PIDs) \cite{doi:10.1371/journal.pbio.2001414}.
-
-As both types of entities are identified by IRIs, their distinction is
-allowed to be blurry; data entities can be located anywhere and be
-complex, while contextual entities can have a Web presence beyond their
-description inside the RO-Crate. For instance
-\texttt{https://orcid.org/0000-0002-1825-0097} is primarily an identifier for
-a person, but secondarily it is also a Web page and a way to refer to
-their academic work.
-
-A particular IRI may appear as a contextual entity in one RO-Crate and
-as a data entity in another; the distinction lies in the fact that data
-entities can be considered to be \textit{contained} or captured by that
-RO-Crate (\textit{RO Content} in Fig.~\ref{fig:conceptual}), while
-contextual entities mainly \textit{explain} an RO-Crate or its content
-(although this distinction is not a formal requirement).
-
-In RO-Crate, a referenced contextual entity (e.g. a person identified
-by ORCID) should always be described within the RO-Crate Metadata File
-with at least a \textit{type} and \textit{name}, even where their PID might resolve
-to further Linked Data. This is so that clients are not required to
-follow every link for presentation purposes, for instance HTML
-rendering. Similarly any imported extension
-terms\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#extending-ro-crate}}
-would themselves also have a human-readable description in the case
-where their PID does not directly resolve to human-readable documentation.
-
-Figure~\ref{fig:uml} shows a simplified UML class diagram of RO-Crate,
-highlighting the different types of data entities and contextual
-entities that can be aggregated and related. While an RO-Crate would
-usually contain one or more data entities (\texttt{hasPart}), it may also be a
-pure aggregation of contextual entities (\texttt{mentions}).
-
-\begin{figure}%[t!]
-\includegraphics{210053f02}
-\caption{Simplified UML class diagram of RO-Crate. The \emph
-{RO-Crate Metadata File} conforms to a version of the specification;
-and contains a JSON-LD graph \cite{sporny_2014} that describes the
-entities that make up the RO-Crate. The \emph{RO-Crate Root Data
-Entity} represent the Research Object as a dataset. The RO-Crate
-aggregates \emph{data entities} (\texttt{hasPart}) which are further
-described using \emph{contextual entities} (which may include
-aggregated and non-aggregated data entities). Multiple types and
-relations from Schema.org allow annotations to be more specific,
-including figures, nested datasets, computational workflows, people,
-organisations, instruments and places. Contextual entities not
-otherwise cross-referenced from other entities' properties (\emph
-{describes}) can be grouped under the root entity (\texttt{mentions}).}
-\label{fig:uml}
-\end{figure}
-
-\subsubsection{Guide through recommended practices}% {#recommendedpractices}
-
-\label{sec:recommendedpractices}
-
-RO-Crate as a specification aims to build a set of recommended
-practices on how to practically apply existing standards in a common
-way to describe research outputs and their provenance, without having
-to learn each of the underlying technologies in detail.
-
-As such, the RO-Crate 1.1\footnote{\url{https://w3id.org/ro/crate/1.1}}
-specification \cite{doi:10.5281/zenodo.4541002} can be seen as an
-opinionated and example-driven guide to writing
-Schema.org\footnote{\url{https://schema.org/}} \cite{doi:10.1145/2857274.2857276}
-metadata as JSON-LD \cite{sporny_2014} (see Section~\ref{sec:implementation}), which leaves it open for implementers to include
-additional metadata using other Schema.org types and properties, or
-even additional Linked Data vocabularies/ontologies or their own ad-hoc terms.
-
-However the primary purpose of the RO-Crate specification is to assist
-developers in leveraging Linked Data principles for the focused purpose
-of describing Research Objects in a structured language, while reducing
-the steep learning curve otherwise associated with Semantic Web
-adaptation, like development of ontologies, identifiers, namespaces,
-and RDF serialization choices.
-\subsubsection{Ensuring simplicity}% {#simplicity}
-
-\label{sec:implicitly}
-
-One aim of RO-Crate is to be conceptually simple. This simplicity has
-been repeatedly checked and confirmed through an informal community
-review process. For instance, in the discussion on supporting ad-hoc
-vocabularies\footnote{\url{https://github.com/ResearchObject/ro-crate/issues/71}} in
-RO-Crate, the community explored potential Linked Data solutions. The
-conventional wisdom in RDF best
-practices\footnote{\url{https://www.w3.org/TR/swbp-vocab-pub/}} is to establish a
-vocabulary with a new IRI namespace, formalised using RDF
-Schema\footnote{\url{http://www.w3.org/TR/2014/REC-rdf-schema-20140225/}} or
-OWL\footnote{\url{http://www.w3.org/TR/2012/REC-owl2-overview-20121211/}}
-ontologies. However, this may seem an excessive learning curve for
-non-experts in semantic knowledge representation, and the RO-Crate
-community instead agreed on a dual lightweight approach: (i)
-Document\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#adding-new-or-ad-hoc-vocabulary-terms}}
-how projects with their own Web-presence can make a pure HTML-based
-vocabulary, and (ii) provide a community-wide PID namespace under
-\texttt{https://w3id.org/ro/terms} that redirect to simple CSV files
-maintained in GitHub.\footnote{\url{https://github.com/ResearchObject/ro-terms}}
-
-To further verify this idea of simplicity, we have formalised the
-RO-Crate definition (see \textit{Appendix~\ref{sec:formaldefinition}}). An
-important result of this exercise is that the underlying data structure
-of RO-Crate, although conceptually a graph, is represented as a
-depth-limited tree. This formalisation also emphasises the
-\textit{boundedness} of the structure; namely, the fact that elements are
-specifically identified as being either semantically \textit{contained} by the
-RO-Crate as \textit{Data Entities} (\texttt{hasPart}) or mainly referenced
-(\texttt{mentions}) and typed as \textit{external} to the Research Object as
-\textit{Contextual Entities}. It is worth pointing out that this semantic
-containment can extend beyond the physical containment of files
-residing within the RO-Crate Root directory on a given storage system,
-as the RO-Crate data entities may include any data resource globally
-identifiable using IRIs.
-\subsubsection{Extensibility and RO-Crate profiles}% {#profiles}
-
-\label{sec:profiles}
-
-The RO-Crate specification provides a core set of conventions to
-describe research outputs using types and properties applicable across
-scientific domains. However we have found that domain-specific use of
-RO-Crate will, implicitly or explicitly, form a specialised \textbf{profile}
-of RO-Crate; i.e., {a set of conventions, types and properties that are
-minimally required and one can expect to be present in that subset of
-RO-Crates}. For instance, RO-Crates used for exchange of workflows will
-have to contain a data entity of type \texttt{ComputationalWorkflow}, or
-cultural heritage records should have a \texttt{contentLocation}.
-
-Making such profiles explicit allow further reliable programmatic
-consumption and generation of RO-Crates beyond the core types defined
-in the RO-Crate specification. Following the RO-Crate mantra of
-\textit{guidance over strictness}, profiles are mainly \textit{duck-typing} rather
-than strict syntactic or semantic types, but may also have
-corresponding machine-readable schemas at multiple levels (file
-formats, JSON, RDF shapes, RDFS/OWL semantics).
-
-The next version of the RO-Crate specification 1.2 will define a
-formalization\footnote{\url{https://www.researchobject.org/ro-crate/1.2-DRAFT/profiles}}
-for publishing and declaring conformance to RO-Crate profiles. Such a
-profile is primarily a human-readable document of before-mentioned
-expectations and conventions, but may also define a machine-readable
-profile as a \textbf{Profile Crate}: Another RO-Crate that describe the
-profile and in addition can list schemas for validation, compatible
-software, applicable repositories, serialization/packaging formats,
-extension vocabularies, custom JSON-LD contexts and examples (see for
-example the Workflow RO-Crate
-profile\footnote{\url{https://w3id.org/workflowhub/workflow-ro-crate/}}).
-
-In addition, there are sometimes existing domain-specific metadata
-formats, but they are either not RDF-based (and thus time-consuming to
-construct terms for in JSON-LD) or are at a different granularity level
-that might become overwhelming if represented directly in the RO-Crate
-Metadata file (e.g. W3C PROV bundle detailing every step execution of a
-workflow run \cite{doi:10.1093/gigascience/giz095}). RO-Crate allows such
-\textit{alternative metadata files} to co-exist, and be described as data
-entities with references to the standards and vocabularies they conform
-to. This simplifies further programmatic consumption even where no
-filename or file extension conventions have emerged for those metadata formats.
-
-Section~\ref{sec:inuse} examines the observed specializations of
-RO-Crate use in several domains and their emerging profiles.
-
-\subsection{Technical implementation of the RO-Crate model}% {#implementation}
-
-\label{sec:implementation}
-
-The RO-Crate conceptual model has been realised using JSON-LD and
-Schema.org in a prescriptive form as discussed in Section~\ref{sec:conceptual}. These technical choices were made to cater for
-simplicity from a developer perspective (as introduced in Section~\ref{sec:methodology}).
-
-JSON-LD\footnote{\url{https://json-ld.org/}} \cite{sporny_2014} provides a way to
-express Linked Data as a JSON structure, where a \textit{context} provides
-mapping to RDF properties and classes. While JSON-LD cannot map
-arbitrary JSON structures to RDF, we found that it does lower the
-barrier compared to other RDF syntaxes, as the JSON syntax nowadays is
-a common and popular format for data exchange on the Web.
-
-However, JSON-LD alone has too many degrees of freedom and hidden
-complexities for software developers to reliably produce and consume
-without specialised expertise or large RDF software frameworks. A large
-part of the RO-Crate specification is therefore dedicated to describing
-the acceptable subset of JSON structures.
-
-\subsubsection{RO-Crate JSON-LD}% {#jsonld}
-
-\label{sec:jsonld}
-
-RO-Crate
-mandates\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html}}
-the use of flattened, compacted JSON-LD in the RO-Crate Metadata file
-\texttt{ro-crate-metadata.json}\footnote{The avid reader may spot that the RO-Crate Metadata file use the
-extension \texttt{.json} instead of \texttt{.jsonld}, this is to emphasise the
-developer expectations as a JSON format, while the file's JSON-LD
-nature is secondary. See \url
-{https://github.com/ResearchObject/ro-crate/issues/82}.}
-where a single \texttt{@graph} array contains all
-the data and contextual entities in a flat list. An example can be seen
-in the JSON-LD snippet in Listing \ref{lis1} below, describing a simple RO-Crate
-containing data entities described using contextual entities.
-
-\begin{listing}
-%\includegraphics{210053i02}
-\footnotesize
-\begin{verbatim}
-{ "@context": "https://w3id.org/ro/crate/1.1/context",
-  "@graph": [
-    { "@id": "ro-crate-metadata.json",      
-      "@type": "CreativeWork",
-      "conformsTo": {"@id": "https://w3id.org/ro/crate/1.1"},
-      "about": {"@id": "./"}
-    },
-    { "@id": "./",
-      "@type": "Dataset",
-      "name": "A simplified RO-Crate",
-      "author": {"@id": "#alice"},
-      "license": {"@id": "https://spdx.org/licenses/CC-BY-4.0"},
-      "datePublished": "2021-11-02T16:04:43Z",
-      "hasPart": [
-        {"@id": "survey-responses-2019.csv"},
-        {"@id": "https://example.com/pics/5707039334816454031_o.jpg"}
-      ]
-    },
-    { "@id": "survey-responses-2019.csv",
-      "@type": "File",
-      "about": {"@id": "https://example.com/pics/5707039334816454031_o.jpg"},
-      "author": {"@id": "#alice"}
-    },
-    { "@id": "https://example.com/pics/5707039334816454031_o.jpg",
-      "@type": ["File", "ImageObject"],
-      "contentLocation": {"@id": "http://sws.geonames.org/8152662/"},
-      "author": {"@id": "https://orcid.org/0000-0002-1825-0097"}
-    },
-    { "@id": "#alice",
-      "@type": "Person",
-      "name": "Alice"
-    },
-    { "@id": "https://orcid.org/0000-0002-1825-0097",
-      "@type": "Person",
-      "name": "Josiah Carberry"
-    },
-    { "@id": "http://sws.geonames.org/8152662/",
-      "@type": "Place",
-      "name": "Catalina Park"
-    },
-    { "@id": "https://spdx.org/licenses/CC-BY-4.0",
-      "@type": "CreativeWork",
-      "name": "Creative Commons Attribution 4.0"
-    }
-  ]
-}    
-\end{verbatim}    
-\caption{%\qq{}\protect\footnotemark[31]
-Simplified\protect\footnote{Recommended properties for types shown in
-Listing \ref{lis1} also include
-\texttt{affiliation}, \texttt{citation}, \texttt{contactPoint},
-\texttt{description},
-\texttt{encodingFormat}, \texttt{funder}, \texttt{geo}, \texttt{identifier},
-\texttt{keywords},
-\texttt{publisher}; these properties and corresponding contextual
-entities are
-excluded here for brevity. See complete example
-\url
-{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.}
-RO-Crate metadata file showing the
-flattened compacted JSON-LD \texttt{@graph} array containing the data entities
-and contextual entities, cross-referenced using \texttt{@id}. The
-\texttt{ro-crate-metadata.json} entity self-declares conformance with the
-RO-Crate specification using a versioned persistent identifier, further
-RO-Crate descriptions are on the root data entity \texttt{./} or any of the
-referenced data or contextual entities. This is exemplified by the data
-entity \texttt{ImageObject} referencing contextual entities for
-\texttt{contentLocation} and \texttt{author} that differs from that of the overall
-RO-Crate. In this crate, \texttt{about} of the CSV data entity reference the
-\texttt{ImageObject}, which then take the roles of both a data entity and
-contextual entity. While \texttt{Person} entities ideally are identified with
-ORCID PIDs as for Josiah, \texttt{\#alice} is here in contrast an RO-Crate
-local identifier, highlighting the pragmatic ``just enough'' Linked
-Data approach.}\label{lis1}
-\end{listing}
-
-In\footnotetext{Recommended properties for types shown in
-Listing \ref{lis1} also include
-\texttt{affiliation}, \texttt{citation}, \texttt{contactPoint},
-\texttt{description},
-\texttt{encodingFormat}, \texttt{funder}, \texttt{geo}, \texttt{identifier},
-\texttt{keywords},
-\texttt{publisher}; these properties and corresponding contextual
-entities are
-excluded here for brevity. See complete example
-\url
-{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.} this flattened profile of JSON-LD, each \texttt{\{entity\}} is directly
-under \texttt{@graph} and represents the RDF triples with a common \textit{subject}
-(\texttt{@id}), mapped \textit{properties} like \texttt{hasPart}, and \textit{objects} -- as
-either literal \texttt{"string"} values, referenced \texttt{\{objects\}} (which
-properties are listed in its own entity), or a JSON \texttt{[list]} of these.
-If processed as JSON-LD, this forms an RDF graph by matching the \texttt{@id}
-IRIs and applying the \texttt{@context} mapping to Schema.org terms.
-\normalsize
-
- \subsubsection{Flattened JSON-LD}
-
-When JSON-LD 1.0 \cite{sporny_2014} was proposed, one of the motivations
-was to seamlessly apply an RDF nature on top of regular JSON as
-frequently used by Web APIs. JSON objects in APIs are frequently nested
-with objects at multiple levels, and the perhaps most common form of
-JSON-LD is the compacted
-form\footnote{\url{https://json-ld.org/spec/REC/json-ld/20140116/\#compacted-document-form}}
-which follows this expectation (JSON-LD
-1.1\footnote{\url{https://www.w3.org/TR/2020/REC-json-ld11-20200716/}} further
-expands these capabilities, e.g. allowing nested \texttt{@context} definitions).
-
-While this feature of JSON-LD can be seen as a way to ``hide'' its RDF
-nature, we found that the use of nested trees (e.g. a \texttt{Person} entity
-appearing as \texttt{author} of a \texttt{File} which nests under a \texttt{Dataset} with
-\texttt{hasPart}) counter-intuitively forces consumers to consider the JSON-LD
-as an RDF Graph, since an identified \texttt{Person} entity can appear at
-multiple and repeated points of the tree (e.g. author of multiple
-files), necessitating node merging or duplication, which can become
-complicated as this approach also invites the use of \textit{blank nodes}
-(entities missing \texttt{@id}).
-
-By comparison, a single flat \texttt{@graph} array approach, as required by
-RO-Crate, means that applications can choose to process and edit each
-entity as pure JSON by a simple lookup based on \texttt{@id}. At the same
-time, lifting all entities to the same level reflects the Research
-Object principles \cite{doi:10.1016/j.future.2011.08.004} in that
-describing the context and provenance is just as important as
-describing the data, and the requirement of \texttt{@id} of every entity
-forces RO-Crate generators to consciously consider existing IRIs and
-identifiers.\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#describing-entities-in-json-ld}}
-
-\subsubsection{JSON-LD context}
-
-In JSON-LD, the \texttt{@context} is a reference to another JSON-LD document
-that provides mapping from JSON keys to Linked Data term IRIs, and can
-enable various JSON-LD directives to cater for customised JSON
-structures for translating to RDF.
-
-RO-Crate reuses vocabulary terms and IRIs from Schema.org, but provides
-its own versioned JSON-LD
-context,\footnote{\url{https://w3id.org/ro/crate/1.1/context}} which has a flat list
-with the mapping from JSON-LD keys to their IRI equivalents (e.g. key
-\texttt{"author"} maps to the \url{http://schema.org/author} property).
-
-The rationale behind this decision is to support JSON-based RO-Crate
-applications that are largely unaware of JSON-LD, that still may want
-to process the \texttt{@context} to find or add Linked Data definitions of
-otherwise unknown properties and types. Not reusing the official
-Schema.org context means RO-Crate is also able to map in additional
-vocabularies where needed, namely the \textit{Portland Common Data Model}
-(PCDM) \cite{pcdm} for repositories and Bioschemas \cite{bioschemas_2017} for
-describing computational workflows. RO-Crate profiles may
-extend\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#extending-ro-crate}}
-the \texttt{@context} to re-use additional domain-specific ontologies.
-
-Similarly, while the Schema.org context
-currently\footnote{\url{https://schema.org/version/13.0/schemaorg-current-http.jsonld}}
-have \texttt{"@type": "@id"} annotations for implicit object properties, RO-Crate
-JSON-LD distinguishes explicitly between references to other entities
-\texttt{\{"@id": "\#alice"}\}
-and string values \texttt{"Alice"} -- meaning
-RO-Crate applications can find references for corresponding entities
-and IRIs without parsing the \texttt{@context} to understand a particular
-property. Notably this is exploited by the \textit{ro-crate-html-js}
-\cite{ro-crate-html-js} tool to provide reliable HTML rendering for
-otherwise unknown properties and types.
-
-\subsection{RO-Crate community}% {#community}
-
-\label{sec:community}
-
-The RO-Crate conceptual model, implementation and best practices are
-developed by a growing community of researchers, developers and
-publishers. RO-Crate's community is a key aspect of its effectiveness
-in making research artefacts FAIR. Fundamentally, the community
-provides the overall context of the implementation and model and
-ensures its interoperability.
-
-The RO-Crate community consists of:
-\begin{enumerate}
-\item[1.] a diverse set of people representing a variety of stakeholders;
-\item[2.] a set of collective norms;
-\item[3.] an open platform that facilitates communication (GitHub, Google
-Docs, monthly teleconferences).
-\end{enumerate}
-\subsubsection{People}
-
-The initial concept of RO-Crate was formed at the first Workshop on
-Research Objects (RO2018\footnote{\url{https://www.researchobject.org/ro2018/}}),
-held as part of the IEEE conference on eScience. This workshop followed
-up on considerations made at a Research Data Alliance (RDA) meeting on
-Research Data
-Packaging\footnote{\url{https://rd-alliance.org/approaches-research-data-packaging-rda-11th-plenary-bof-meeting}}
-that found similar goals across multiple data packaging efforts
-\cite{doi:10.5281/zenodo.3250687}: simplicity, structured metadata and the
-use of JSON-LD.
-
-An important outcome of discussions that took place at RO2018 was the
-conclusion that the original Wf4Ever Research Object ontologies
-\cite{doi:10.1016/j.websem.2015.01.003}, in principle sufficient for
-packaging research artefacts with rich descriptions, were, in practice,
-considered inaccessible for regular programmers (e.g., Web developers)
-and in danger of being incomprehensible for domain scientists due to
-their reliance on Semantic Web technologies and other ontologies.
-
-DataCrate \cite{doi:10.5281/zenodo.1445817} was presented at RO2018 as a
-promising lightweight alternative approach, and an agreement was made
-by a group of volunteers to attempt building what was initially called
-\textit{``RO Lite''} as a combination of DataCrate's implementation and
-Research Object's principles.
-
-This group, originally made up of library and Semantic Web experts, has
-subsequently grown to include domain scientists, developers, publishers
-and more. This perspective of multiple views led to the specification
-being used in a variety of domains, from bioinformatics and regulatory
-submissions to humanities and cultural heritage preservation.
-
-The RO-Crate community is strongly engaged with the European-wide
-biology/bioinformatics collaborative e-Infrastructure ELIXIR
-\cite{doi:10.1016/j.tibtech.2012.02.002}, along with European Open Science
-Cloud\footnote{\url{https://eosc.eu/}} (EOSC) projects including
-EOSC-Life,\footnote{\url{https://www.eosc-life.eu/}}
-FAIRplus,\footnote{\url{https://fairplus-project.eu/}}
-CS3MESH4EOSC\footnote{\url{https://cs3mesh4eosc.eu/}} and
-BY-COVID.\footnote{\url{https://by-covid.eu/}} RO-Crate has also established
-collaborations with Bioschemas \cite{bioschemas_2017}, GA4GH
-\cite{doi:10.1016/j.xgen.2021.100029}, OpenAIRE \cite{rettberg_2015_openaire}
-and multiple H2020 projects.
-
-A key set of stakeholders are developers: the RO-Crate community has
-made a point of attracting developers who can implement the
-specifications but, importantly, keeps ``developer user experience'' in
-mind. This means that the specifications are straightforward to
-implement and thus do not require expertise in technologies that are
-not widely deployed.
-
-This notion of catering to ``developer user experience'' is an example
-of the set of norms that have developed and now define the community.
-
-\subsubsection{Norms}
-
-The RO-Crate community is driven by informal conventions and notions
-that are prevalent but not neccessarily written down. Here, we distil
-what we as authors believe are the critical set of norms that have
-facilitated the development of RO-Crate and contributed to the ability
-for RO-Crate research packages to be FAIR. This is not to say that
-there are no other norms within the community nor that everyone in the
-community holds these uniformly. Instead, what we emphasise is that
-these norms are helpful and also shaped by community practices.
-\begin{enumerate}
-\item[1.] Simplicity
-\item[2.] Developer friendliness
-\item[3.] Focus on examples and best practices rather than rigorous specification
-\item[4.] Reuse ``just enough'' Web standards
-\end{enumerate}
-
-A core norm of RO-Crate is that of \textbf{simplicity}, which sets the scene
-for how we guide developers to structure metadata with RO-Crate. We
-focus mainly on documenting simple approaches to the most common use
-cases, such as authors having an affiliation. This norm also influences
-our take on \textbf{developer friendliness}; for instance, we are using the
-Web-native JSON format, allowing only a few of JSON-LD's flexible
-Linked Data features. Moreover, the RO-Crate documentation is largely
-built up by \textbf{examples} showcasing \textbf{best practices}, rather than
-rigorous specifications. We build on existing \textbf{Web standards} that
-themselves are defined rigorously, which we utilise \textit{``\textbf{just
-enough}''} in order to benefit from the advantages of Linked Data
-(e.g., extensions by namespaced vocabularies), without imposing too
-many developer choices or uncertainties (e.g., having to choose between
-the many RDF syntaxes).
-
-While the above norms alone could easily lead to the creation of ``yet
-another'' JSON format, we keep the goal of \textbf{FAIR interoperability} of
-the captured metadata, and therefore follow closely FAIR best practices
-and current developments such as data citations, PIDs, open
-repositories and recommendations for sharing research outputs and software.
-
-\subsubsection{Open platforms}
-
-The critical infrastructure that enables the community around RO-Crate
-is the use of open development platforms. This underpins the importance
-of open community access to supporting FAIR. Specifically, it is
-difficult to build and consume FAIR research artefacts without being
-able to access the specifications, understand how they are developed,
-know about any potential implementation issues, and discuss usage to
-evolve best practices.
-
-The development of RO-Crate was driven by capturing documentation of
-real-life examples and best practices rather than creating a rigorous
-specification. At the same time, we agreed to be opinionated on the
-syntactic form to reduce the jungle of implementation choices; we
-wanted to keep the important aspects of Linked Data to adhere to the
-FAIR principles while retaining the option of combining and extending
-the structured metadata using the existing Semantic Web stack, not just
-build a standalone JSON format.
-
-Further work during 2019 started adapting the DataCrate documentation
-through a more collaborative and exploratory \textit{RO Lite} phase, initially
-using Google Docs for review and discussion, then moving to GitHub as a
-collaboration space for developing what is now the RO-Crate
-specification, maintained\footnote{\url{https://github.com/researchobject/ro-crate/}} as
-Markdown in GitHub Pages
-and published through Zenodo.
-
-In addition to the typical Open Source-style development with GitHub
-issues and pull requests, the RO-Crate Community have, at time of
-writing, two regular monthly calls, a Slack channel and a mailing list
-for coordinating the project; also many of its participants collaborate
-on RO-Crate at multiple conferences and coding events such as the
-ELIXIR BioHackathon.\footnote{\url{https://biohackathon-europe.org/}} The community
-is jointly developing the RO-Crate specification and Open Source tools,
-as well as providing support and considering new use cases. The
-RO-Crate Community\footnote{\url{https://www.researchobject.org/ro-crate/community}}
-is open for anyone to join, to equally participate under a code of
-conduct, and as of October 2021 has more than 50 members (see Appendix~\ref{sec:communitylist}).
-
-% Please update in 31.tooling-table.md first!
-%
-\begin{table}[t!]%[htbp]
-\tabcolsep=0pt
-\caption{Applications and libraries implementing RO-Crate, targeting
-different types of users across multiple programming languages. Status
-is indicative as assessed by this work (Alpha $<$ Beta $<$ Release
-Candidate (RC) $<$ Release)}
-\label{tab1}
-%
-\begin{tabular*}{\textwidth}{@{\extracolsep{4in minus 4in}}llcc@{}}
-\hline
-\textbf{Tool name} & \multicolumn{1}{c}{\tmultirow{2}{*}{Targets}}
-& \tmultirow{2}{*}{Language/Platform} & \tmultirow{2}{*}{Status} \\
-\textit{Brief Description}&&& \\
-\hline
-\textbf{Describo} \cite{describo}                                                                                                          & Research Data Managers     & NodeJS (Desktop) & RC    \\
-\multicolumn{4}{@{}l@{}}{\textit{Interactive desktop application to create, update and export RO-Crates for different profiles}}                                                                               \\
-\textbf{Describo Online} \cite{describo-online}                                                                                            & Platform developers        & NodeJS (Web)     & Alpha \\
-\multicolumn{4}{@{}l@{}}{\textit{Web-based application to create RO-Crates using cloud storage}}                                                                                                               \\
-\textbf{ro-crate-excel} \cite{ro-crate-excel}                                                                                              & Data managers              & JavaScript       & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Command-line tool to create/edit RO-Crates with spreadsheets}}                                                                                                                \\
-\textbf{ro-crate-html-js} \cite{ro-crate-html-js}                                                                                          & Developers                 & JavaScript       & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{HTML rendering of RO-Crate}}                                                                                                                                                  \\
-\textbf{ro-crate-js} \cite{ro-crate-js}                                                                                                    & Research Data Managers     & JavaScript       & Alpha \\
-\multicolumn{4}{@{}l@{}}{\textit{Library for creating/manipulating crates; basic validation code}}                                                                                                             \\
-\textbf{ro-crate-ruby} \cite{ro-crate-ruby}                                                                                                & Developers                 & Ruby             & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Ruby library for reading/writing RO-Crate, with workflow support}}                                                                                                            \\
-\textbf{ro-crate-py} \cite{ro-crate-py}                                                                                                    & Developers                 & Python           & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Object-oriented Python library for reading/writing RO-Crate and use by Jupyter Notebook}}                                                                                     \\
-\textbf{WorkflowHub} \cite{about-workflowhub}                                                                                              & Workflow users             & Ruby             & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Workflow repository; imports and exports Workflow RO-Crate}}                                                                                                                  \\
-\textbf{Life Monitor} \cite{about-lifemonitor}                                                                                             & Workflow developers        & Python           & Alpha \\
-\multicolumn{4}{@{}l@{}}{\textit{Workflow testing and monitoring service; Workflow Testing profile of RO-Crate}}                                                                                               \\
-\textbf{SCHeMa} \cite{vergoulis2021schema}                                                                                                 & Workflow users             & PHP              & Alpha \\
-\multicolumn{4}{@{}l@{}}{\textit{Workflow execution using RO-Crate as exchange mechanism}}                                                                                                                     \\
-\textbf{galaxy2cwl} \cite{galaxy2cwl}                                                                                                      & Workflow developers        & Python           & Alpha \\
-\multicolumn{4}{@{}l@{}}{\textit{Wraps Galaxy workflow as Workflow RO-Crate}}                                                                                                                                  \\
-\textbf{Modern PARADISEC} \cite{modpdsc}                                                                                                   & Repository managers        & Platform         & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Cultural Heritage portal based on OCFL and RO-Crate}}                                                                                                                        \\
-\textbf{ONI express} \cite{arkisto-data-portal}                                                                                            & Repository managers        & Platform         & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Platform for publishing data and documents stored in an OCFL repository via a Web interface}}                                                                                 \\
-\textbf{ocfl-tools} \cite{ocfl-tools}                                                                                                      & Developers                 & JavaScript (CLI) & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Tools for managing RO-Crates in an OCFL repository}}                                                                                                                          \\
-\textbf{RO Composer} \cite{ro-composer}                                                                                                    & Repository developers      & Java             & Alpha \\
-\multicolumn{4}{@{}l@{}}{\textit{REST API for gradually building ROs for given profile}}                                                                                                                       \\
-\textbf{RDA maDMP Mapper} \cite{doi:10.5281/zenodo.3922136}                                                                                & Data Management Plan users & Python           & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Mapping between machine-actionable data management plans (maDMP) and RO-Crate \cite{doi:10.4126/frl01-006423291}}}                                                         \\
-\textbf{Ro-Crate\_2\_ma-DMP} \cite{doi:10.5281/zenodo.3903463}                                                                             & Data Management Plan users & Python           & Beta  \\
-\multicolumn{4}{@{}l@{}}{\textit{Convert between machine-actionable data management plans (maDMP) and RO-Crate}}                                                                                              \\
-\textbf{CheckMyCrate} \cite{CheckMyCrate}                                                                                                  & Developers                 & Python (CLI)     & Alpha \\
-\multicolumn{4}{@{}l@{}}{\textit{Validation according to Workflow RO-Crate profile}}                                                                                                                           \\
-\textbf{RO-Crates-and-Excel} \cite{doi:10.5281/zenodo.5068950}                                                                             & Data Managers              & Java (CLI)       & Alpha \\
-\multicolumn{4}{@{}l@{}}{\textit{Describe column/data details of spreadsheets as RO-Crate using DataCube vocabulary}}                                                                                          \\
-\hline
-\end{tabular*}\vspace*{10pt}
-\end{table}
-
-\section{RO-Crate tooling}% {#tooling}
-
-\label{sec:tooling}
-
-The work of the community has led to the development of a number of
-tools for creating and using RO-Crates. Table~\ref{tab1} shows the current set
-of implementations. Reviewing this list, one can see support for
-commonly used programming languages, including Python, JavaScript, and
-Ruby. Additionally, the tools can be integrated into commonly used
-research environments, in particular, the command line tool
-\textit{ro-crate-html-js} \cite{ro-crate-html-js} for creating a human-readable
-preview of an RO-Crate as a sidecar HTML file. Furthermore, there are
-tools that cater to end-users (\textit{Describo} \cite{describo}, \textit{WorkflowHub}
-\cite{about-workflowhub}), in order to simplify creating and managing
-RO-Crate. For example, Describo was developed to help researchers of
-the Australian Criminal Characters
-project\footnote{\url{https://criminalcharacters.com/}} to annotate historical prisoner
-records for greater insight into the history of Australia
-\cite{doi:10.1080/14490854.2020.1796500}.
-
-While the development of these tools is promising, our analysis of
-their maturity status shows that the majority of them are in the Beta
-stage. This is partly due to the fact that the RO-Crate specification
-itself only recently reached 1.0 status, in November 2019
-\cite{doi:10.5281/zenodo.3541888}. Now that there is a fixed point of
-reference: With version 1.1 (October 2020)
-\cite{doi:10.5281/zenodo.4031327} RO-Crate has stabilised based on feedback
-from application development, and now we are seeing a further increase
-in the maturity of these tools, along with the creation of new ones.
-
-Given the stage of the specification, these tools have been primarily
-targeting developers, essentially providing them with the core
-libraries for working with RO-Crate. Another target has been that of
-research data managers who need to manage and curate large amounts of data.
-
-\section{Profiles of RO-Crate in use}% {#inuse}
-
-\label{sec:inuse}
-
-RO-Crate fundamentally forms part of an infrastructure to help build
-FAIR research artefacts. In other words, the key question is whether
-RO-Crate can be used to share and (re)use research artefacts. Here we
-look at three research domains where RO-Crate is being applied:
-Bioinformatics, Regulatory Science and Cultural Heritage. In addition,
-we note how RO-Crate may have an important role as part of
-machine-actionable data management plans and institutional repositories.
-
-From these varied uses of RO-Crate we observe natural differences in
-their detail level and the type of entities described by the RO-Crate.
-For instance, on submission of an RO-Crate to a workflow repository, it
-is reasonable to expect the RO-Crate to contain at least one workflow,
-ideally with a declared licence and workflow language. Specific
-additional recommendations such as on identifiers is also needed to
-meet the emerging requirements of FAIR Digital
-Objects.\footnote{\url{https://fairdo.org/}} Work has now
-begun\footnote{\url{https://github.com/ResearchObject/ro-crate/issues/153}} to
-formalise these different \textit{profiles} of RO-Crates, which may impose
-additional constraints based on the needs of a specific domain or use case.
-
-\subsection{Bioinformatics workflows}% {#workflows}
-
-\label{sec:workflows}
-
-WorkflowHub.eu\footnote{\url{https://workflowhub.eu/}} is a European cross-domain
-registry of computational workflows, supported by European Open Science
-Cloud projects, e.g. EOSC-Life,\footnote{\url{https://www.eosc-life.eu/}} and
-research infrastructures including the pan-European bioinformatics
-network ELIXIR\footnote{\url{https://elixir-europe.org/}}
-\cite{doi:10.1016/j.tibtech.2012.02.002}. As part of promoting workflows as
-reusable tools, WorkflowHub includes documentation and high-level
-rendering of the workflow structure independent of its native workflow
-definition format. The rationale is that a domain scientist can browse
-all relevant workflows for their domain, before narrowing down their
-workflow engine requirements. As such, the WorkflowHub is intended
-largely as a registry of workflows already deposited in repositories
-specific to particular workflow languages and domains, such as
-UseGalaxy.eu \cite{doi:10.1371/journal.ppat.1008643} and Nextflow nf-core
-\cite{doi:10.1038/s41587-020-0439-x}.
-
-We here describe three different RO-Crate profiles developed for use
-with WorkflowHub.
-
-\subsubsection{Profile for describing workflows}
-
-Being cross-domain, WorkflowHub has to cater for many different
-workflow systems. Many of these, for instance Nextflow
-\cite{doi:10.1038/nbt.3820} and Snakemake
-\cite{doi:10.1093/bioinformatics/bts480}, by virtue of their script-like
-nature, reference multiple neighbouring files typically maintained in a
-GitHub repository. This calls for a data exchange method that allows
-keeping related files together. WorkflowHub has tackled this problem by
-adopting RO-Crate as the packaging mechanism
-\cite{doi:10.5281/zenodo.4705078}, typing and annotating the constituent
-files of a workflow and -- crucially -- marking up the workflow
-language, as many workflow engines use common file extensions like
-\texttt{*.xml} and \texttt{*.json}. Workflows are further described with authors,
-license, diagram previews and a listing of their inputs and outputs.
-RO-Crates can thus be used for interoperable deposition of workflows to
-WorkflowHub, but are also used as an archive for downloading workflows,
-embedding metadata registered with the WorkflowHub entry and translated
-workflow files such as abstract Common Workflow Language (CWL)
-\cite{doi:10.1145/3486897} definitions and diagrams \cite{doi:10.5281/zenodo.4605654}.
-
-RO-Crate acts therefore as an interoperability layer between
-registries, repositories and users in WorkflowHub. The iterative
-development between WorkflowHub developers and the RO-Crate community
-heavily informed the creation of the Bioschemas \cite{bioschemas_2017}
-profile for Computational
-Workflows,\footnote{\url{https://bioschemas.org/profiles/ComputationalWorkflow/1.0-RELEASE/}}
-which again informed the RO-Crate 1.1 specification on
-workflows\footnote{\url{https://www.researchobject.org/ro-crate/1.1/workflows.html}}
-and led to the RO-Crate Python library \cite{ro-crate-py} and WorkflowHub's
-\textbf{Workflow RO-Crate
-profile},\footnote{\url{https://w3id.org/workflowhub/workflow-ro-crate/1.0}} which,
-in a similar fashion to RO-Crate itself, recommends which workflow
-resources and descriptions are required. This co-development across
-project boundaries exemplifies the drive for simplicity and for
-establishing best practices.
-
-\subsubsection{Profile for recording workflow runs}
-
-RO-Crates in WorkflowHub have so far been focused on workflows that are
-ready to be run, and development of WorkflowHub is now creating a
-\textbf{Workflow Run RO-Crate profile} for the purposes of benchmarking,
-testing and executing workflows. As such, RO-Crate serves as a
-container of both a \textit{workflow definition} that may be executed and of a
-particular \textit{workflow execution with test results}.
-
-This workflow run profile is a continuation of our previous work with
-capturing workflow provenance in a Research Object in CWLProv
-\cite{doi:10.1093/gigascience/giz095} and TavernaPROV
-\cite{doi:10.5281/zenodo.51314}. In both cases, we used the PROV Ontology
-\cite{PROVO}, including details of every task execution with all the
-intermediate data, which required significant workflow engine
-integration.\footnote{CWLProv and TavernaProv predate RO-Crate, but use
-RO-Bundle\cite{doi:10.5281/zenodo.12586}, a similar Research Object
-packaging method with JSON-LD metadata.}
-
-Simplifying from the CWLProv approach, the planned Workflow Run
-RO-Crate profile will use a high level Schema.org
-provenance\footnote{\url{https://www.researchobject.org/ro-crate/1.1/provenance.html\#software-used-to-create-files}}
-for the input/output boundary of the overall workflow execution. This
-\textit{Level 1 workflow provenance} \cite{doi:10.1093/gigascience/giz095} can be
-expressed generally across workflow languages with minimal workflow
-engine changes, with the option of more detailed provenance traces as
-separate PROV artefacts in the RO-Crate as data entities. In the
-current development of Specimen Data
-Refinery\footnote{\url{https://github.com/DiSSCo/SDR}} \cite{doi:10.3897/rio.6.e57602}
-these RO-Crates will document the text recognition workflow runs of
-digitised biological specimens, exposed as FAIR Digital Objects
-\cite{doi:10.3390/publications8020021}.
-
-WorkflowHub has recently enabled minting of Digital Object Identifiers
-(DOIs), a PID commonly used for scholarly artefacts, for registered
-workflows, e.g. \texttt{10.48546/workflowhub.workflow.56.1}
-\cite{doi:10.48546/workflowhub.workflow.56.1}, lowering the barrier for
-citing workflows as computational methods along with their FAIR
-metadata -- captured within an RO-Crate. While it is not an aim for
-WorkflowHub to be a repository of workflow runs and their data,
-RO-Crates of \textit{exemplar workflow runs} serve as useful workflow
-documentation, as well as being an exchange mechanism that preserves
-FAIR metadata in a diverse workflow execution environment.
-
-\subsubsection{Profile for testing workflows}
-
-The value of computational workflows, however, is potentially
-undermined by the ``collapse'' over time of the software and services
-they depend upon: for instance, software dependencies can change in a
-non-backwards-compatible manner, or active maintenance may cease; an
-external resource, such as a reference index or a database query
-service, could shift to a different URL or modify its access protocol;
-or the workflow itself may develop hard-to-find bugs as it is updated.
-This \textit{workflow decay} can take a big toll on the workflow's reusability
-and on the reproducibility of any processes it evokes
-\cite{doi:10.1109/eScience.2012.6404482}.
-
-For this reason, WorkflowHub is complemented by a monitoring and
-testing service called LifeMonitor \cite{about-lifemonitor}, also supported
-by EOSC-Life. LifeMonitor's main goal is to assist in the creation,
-periodic execution and monitoring of workflow tests, enabling the early
-detection of software collapse in order to minimise its detrimental
-effects. The communication of metadata related to workflow testing is
-achieved through the adoption of a \textbf{Workflow Testing RO-Crate
-profile}\footnote{\url{https://lifemonitor.eu/workflow_testing_ro_crate}} stacked on
-top of the \textit{Workflow RO-Crate} profile. This further specialisation of
-Workflow RO-Crate allows to specify additional testing-related entities
-(test suites, instances, services, etc.), leveraging RO-Crate's
-extension
-mechanism\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#extending-ro-crate}}
-through the addition of terms from custom namespaces.
-
-In addition to showcasing RO-Crate's extensibility, the testing profile
-is an example of the format's flexibility and adaptability to the
-different needs of the research community. Though ultimately related to
-a computational workflow, in fact, most of the testing-specific
-entities are more about describing a protocol for interacting with a
-monitoring service than a set of research outputs and its associated
-metadata. Indeed, one of LifeMonitor's main functionalities is
-monitoring and reporting on test suites running on existing Continuous
-Integration (CI) services, which is described in terms of service URLs
-and job identifiers in the testing profile. In principle, in this
-context, data could disappear altogether, leading to an RO-Crate
-consisting entirely of contextual entities. Such an RO-Crate acts more
-as an exchange format for communication between services (WorkflowHub
-and LifeMonitor) than as an aggregator for research data and metadata,
-providing a good example of the format's high versatility.
-
-\subsection{Regulatory sciences}% {#regulatorysciences}
-
-\label{sec:regulatorysciences}
-
-BioCompute Objects\footnote{\url{https://biocomputeobject.org/}} (BCO)
-\cite{doi:10.1371/journal.pbio.3000099} is a community-led effort to
-standardise submissions of computational workflows to biomedical
-regulators. For instance, a genomics sequencing pipeline, as part of a
-personalised cancer treatment study, can be submitted to the US Food
-and Drugs Administration (FDA) for approval. BCOs are formalised in the
-standard IEEE 2791-2020 \cite{doi:10.1109/IEEESTD.2020.9094416} as a
-combination of JSON
-Schemas\footnote{\url{https://w3id.org/ieee/ieee-2791-schema/}}
-that define the structure of JSON metadata files describing exemplar
-workflow runs in detail, covering aspects such as the usability and
-error domain of the workflow, its runtime requirements, the reference
-datasets used and representative output data produced.
-
-BCOs provide a structured view over a particular workflow, informing
-regulators about its workings independently of the underlying workflow
-definition language. However, BCOs have only limited support for
-additional metadata.\footnote{IEEE 2791-2020 do permit user extensions in the \textit{extension
-domain} by referencing additional JSON Schemas.} For instance, while the BCO itself can
-indicate authors and contributors, and in particular regulators and
-their review decisions, it cannot describe the provenance of individual
-data files or workflow definitions.
-
-As a custom JSON format, BCOs cannot be extended with Linked Data
-concepts, except by adding an additional top-level JSON object
-formalised in another JSON Schema. A BCO and workflow submitted by
-upload to a regulator will also frequently consist of multiple
-cross-related files. Crucially, there is no way to tell whether a given
-\texttt{*.json} file is a BCO file, except by reading its content and check
-for its \texttt{spec\_version}.
-
-We can then consider how a BCO and its referenced artefacts can be
-packaged and transferred following FAIR principles. \textbf{BCO
-RO-Crate}\footnote{\url{https://biocompute-objects.github.io/bco-ro-crate/}} \cite{doi:10.5281/zenodo.4633732},
-part of the BioCompute Object user guides, defines a set of best
-practices for wrapping a BCO with a workflow, together with its
-exemplar outputs in an RO-Crate, which then provides typing and
-additional provenance metadata of the individual files, workflow
-definition, referenced data and the BCO metadata itself.
-
-Here the BCO is responsible for describing the \textit{purpose} of a workflow
-and its run at an abstraction level suitable for a domain scientist,
-while the more open-ended RO-Crate describes the surroundings of the
-workflow, classifying and relating its resources and providing
-provenance of their existence beyond the BCO. This emerging \textit{separation
-of concerns} is shown in Fig.~\ref{fig:sep_concerns}, and highlights
-how RO-Crate is used side-by-side of existing standards and tooling,
-even where there are apparent partial overlaps.
-
-A similar separation of concerns can be found if considering the
-RO-Crate as a set of files, where the \textit{transport-level} metadata, such
-as checksum of files, are delegated to separate
-BagIt\footnote{\url{https://www.researchobject.org/ro-crate/1.1/appendix/implementation-notes.html\#adding-ro-crate-to-bagit}}
-manifests, a standard focusing on the preservation challenges of
-digital libraries \cite{doi:10.17487/rfc8493}. As such, RO-Crate metadata
-files are not required to iterate all the files in their folder
-hierarchy, only those that benefit from being described.
-
-Specifically, a BCO description alone is insufficient for reliable
-re-execution of a workflow, which would need a compatible workflow
-engine depending on the original workflow definition language, so IEEE
-2791 recommends using Common Workflow Language (CWL)
-\cite{doi:10.1145/3486897} for interoperable pipeline execution. CWL itself
-relies on tool packaging in software containers using
-Docker\footnote{\url{https://www.docker.com/}} or Conda.\footnote{\url{https://docs.conda.io/}}
-Thus, we can consider BCO RO-Crate as a stack: transport-level
-manifests of files (BagIt), provenance, typing and context of those
-files (RO-Crate), workflow overview and purpose (BCO), interoperable
-workflow definition (CWL) and tool distribution (Docker).
-
-%
-\begin{figure}%[t]
-\includegraphics{210053f03}
-\caption{Separation of Concerns in BCO RO-Crate. BioCompute
-Object (IEEE2791) is a JSON file that structurally explains the purpose
-and implementation of a computational workflow, for instance
-implemented in Common Workflow Language (CWL), that installs the
-workflow's software dependencies as Docker containers or BioConda
-packages. An example execution of the workflow shows the different
-kinds of result outputs, which may be external, using GitHub LFS \cite
-{github-lfs} to support larger data. RO-Crate gathers all these local
-and external resources, relating them and giving individual
-descriptions, for instance permanent DOI identifiers for reused
-datasets accessed from Zenodo, but also adding external identifiers to
-attribute authors using ORCID or to identify which licences apply to
-individual resources. The RO-Crate and its local files are captured in
-a BagIt whose checksum ensures completeness, combined with Big Data Bag
-\cite{doi:10.1109/BigData.2016.7840618} features to ``complete'' the
-bag with large external files such as the workflow outputs.}
-\label{fig:sep_concerns}
-\end{figure}
-%
-\subsection{Digital humanities: Cultural heritage}% {#culturalheritage}
-
-\label{sec:culturalheritage}
-
-The Pacific And Regional Archive for Digital Sources in Endangered
-Cultures (PARADISEC\footnote{\url{https://www.paradisec.org.au/}}) \cite{doi:10125/4567}
-maintains a repository of more than 500,000 files documenting
-endangered languages across more than 16,000 items, collected and
-digitised over many years by researchers interviewing and recording
-native speakers across the region.
-
-The Modern PARADISEC demonstrator\footnote{\url{https://mod.paradisec.org.au/}} has
-been
-proposed\footnote{\url{https://arkisto-platform.github.io/case-studies/paradisec/}}
-as an update to the 18 year old infrastructure, to also help long-term
-preservation of these artefacts in their digital form. The demonstrator
-uses RO-Crate to describe the overall structure and to capture the
-metadata of each item. The existing PARADISEC data collection has been
-ported and captured as RO-Crates. A Web portal then exposes the
-repository and its entries by indexing the RO-Crate metadata files,
-presenting a domain-specific view of the items -- the RO-Crate is
-``hidden'' and does not change the user interface.
-
-The PARADISEC use case takes advantage of several RO-Crate features and
-principles. Firstly, the transcribed metadata are now independent of
-the PARADISEC platform and can be archived, preserved and processed in
-its own right, using Schema.org as base vocabulary and extended with
-PARADISEC-specific terms.
-
-In this approach, RO-Crate is the holder of itemised metadata, stored
-in regular files that are organised using Oxford Common File
-Layout\footnote{\url{https://ocfl.io/1.0/spec/}} (OCFL) \cite{ocfl_2020}, which ensures
-file integrity and versioning on a regular shared file system. This
-lightweight infrastructure also gives flexibility for future
-developments and maintenance. For example a consumer can use Linked
-Data software such as a graph database and query the whole corpora
-using SPARQL triple patterns across multiple RO-Crates. For long term
-digital preservation, beyond the lifetime of PARADISEC portals, a
-``last resort'' fallback is storing the generic RO-Crate HTML preview
-\cite{ro-crate-html-js}. Such human-readable rendering of RO-Crates can be
-hosted as static files by any Web server, in line with the approach
-taken by the Endings Project.\footnote{The Endings Project \url{https://endings.uvic.ca/} is a five-year
-project funded by the Social Sciences and Humanities Research Council
-(SSHRC) that is creating tools, principles, policies and
-recommendations for digital scholarship practitioners to create
-accessible, stable, long-lasting resources in the humanities.}
-
-\subsection{Machine-actionable Data Management Plans}% {#dmp}
-
-\label{sec:dmp}
-
-Machine-actionable Data Management Plans (maDMPs) have been proposed as
-an improvement to automate FAIR data management tasks in research
-\cite{doi:10.1371/journal.pcbi.1006750}; maDMPs use PIDs and controlled
-vocabularies to describe what happens to data over the research life
-cycle \cite{doi:10.1007/978-3-030-45442-5_15}. The Research Data Alliance's
-\textit{DMP Common Standard} for maDMPs \cite{doi:10.15497/rda00039} is one such
-formalisation for expressing maDMPs, which can be expressed as Linked
-Data using the DMP Common Standard Ontology
-\cite{doi:10.4126/frl01-006423289}, a specialisation of the W3C Data
-Catalog Vocabulary (DCAT) \cite{dcat2}. RDA maDMPs are usually expressed
-using regular JSON, conforming to the DMP JSON Schema.
-
-A mapping has been produced between Research Object Crates and
-Machine-actionable Data Management Plans
-\cite{doi:10.4126/frl01-006423291}, implemented by the RO-Crate RDA maDMP
-Mapper \cite{doi:10.5281/zenodo.3922136}. A similar mapping has been
-implemented by \textit{RO-Crate\_2\_ma-DMP} \cite{doi:10.5281/zenodo.3903463}. In
-both cases, a maDMP can be converted to a RO-Crate, or vice versa. In
-\cite{doi:10.4126/frl01-006423291} this functionality caters for two use cases:
-\begin{enumerate}
-\item[1.] Start a skeleton data management plan based on an existing RO-Crate
-dataset, e.g. an RO-Crate from WorkflowHub.
-\item[2.] Instantiate an RO-Crate based on a data management plan.
-\end{enumerate}
-
-An important nuance here is that data management plans are (ideally)
-written in \textit{advance} of data production, while RO-Crates are typically
-created to describe data \textit{after} it has been generated. What is
-significant to note in this approach is the importance of
-\textbf{templating} in order to make both tasks automatable and achievable,
-and how RO-Crate can fit into earlier stages of the research life cycle.
-\subsection{Institutional data repositories -- Harvard Data Commons}% {#institutionalrepos}
-
-\label{sec:institutionalrepos}
-
-The concept of a \textbf{Data Commons} for research collaboration was
-originally defined as \textit{``cyber-infrastructure that co-locates data,
-storage, and computing infrastructure with commonly used tools for
-analysing and sharing data to create an interoperable resource for the
-research community''} \cite{doi:10.1109/MCSE.2016.92}. More recently, Data
-Commons has been established to mean integration of active
-data-intensive research with data management and archival best
-practices, along with a supporting computational infrastructure.
-Furthermore, the Commons features tools and services, such as
-computation clusters and storage for scalability, data repositories for
-disseminating and preserving regular, but also large or sensitive
-datasets, and other research assets. Multiple initiatives were
-undertaken to create Data Commons on national, research, and
-institutional levels. For example, the Australian Research Data
-Commons (ARDC)\footnote{\url{https://ardc.edu.au/}} \cite{doi:10.5334/dsj-2019-044} is a
-national initiative that enables local researchers and industries to
-access computing infrastructure, training, and curated datasets for
-data-intensive research. NCI's Genomic Data
-Commons\footnote{\url{https://gdc.cancer.gov/}} (GDC)
-\cite{doi:10.1182/blood-2017-03-735654} provides the cancer research
-community with access to a vast volume of genomic and clinical data.
-Initiatives such as Research Data Alliance (RDA) Global Open Research
-Commons\footnote{\url{https://www.rd-alliance.org/groups/global-open-research-commons-ig}}
-propose standards for the implementation of Data Commons to prevent
-them becoming ``data silos'' and thus, enable interoperability from one
-Data Commons to another.
-
-\textbf{Harvard Data Commons} \cite{doi:10.7557/5.5422} aims to address the
-challenges of data access and cross-disciplinary research within a
-research institution. It brings together multiple institutional
-schools, libraries, computing centres and the Harvard
-Dataverse\footnote{\url{https://dataverse.harvard.edu/}} data repository.
-Dataverse\footnote{\url{https://dataverse.org/}} \cite{doi:10.1045/january2011-crosas}
-is a free and open-source software platform to archive, share and cite
-research data. The Harvard Dataverse repository is the largest of 70
-Dataverse installations worldwide, containing over 120K datasets with
-about 1.3M data files (as of 2021-11-16). Working toward the goal of
-facilitating collaboration and data discoverability and management
-within the university, Harvard Data Commons has the following primary
-objectives:\looseness=1
-\begin{enumerate}
-\item[1.] the integration of Harvard Research Computing with Harvard Dataverse
-by leveraging Globus endpoints \cite{doi:10.1109/MCC.2014.52}; this will
-allow an automatic transfer of large datasets to the repository. In
-some cases, only the metadata will be transferred while the data stays
-stored in remote storage;
-\item[2.] support for advanced research workflows and providing packaging
-options for assets such as code and workflows in the Harvard Dataverse
-repository to enable reproducibility and reuse, and
-\item[3.] interation of repositories supported by Harvard, which include
-DASH,\footnote{\url{https://dash.harvard.edu/}} the open access institutional
-repository, the Digital Repository Services (DRS) for preserving
-digital asset collections, and the Harvard Dataverse.
-\end{enumerate}
-
-Particularly relevant to this article is the second objective of the
-Harvard Data Commons, which aims to support the deposit of research
-artefacts to Harvard Dataverse with sufficient information in the
-metadata to allow their future reuse (Fig.~\ref{fig:hdc}). To support
-the incorporation of data, code, and other artefacts from various
-institutional infrastructures, Harvard Data Commons is currently
-working on RO-Crate adaptation. The RO-Crate metadata provides the
-necessary structure to make all research artefacts FAIR. The Dataverse
-software already has extensive support for metadata\footnote{\url{https://guides.dataverse.org/en/latest/user/appendix.html}}, including the Data
-Documentation Initiative (DDI), Dublin Core, DataCite, and Schema.org.
-Incorporating RO-Crate, which has the flexibility to describe a wide
-range of research resources, will facilitate their seamless transition
-from one infrastructure to the other within the Harvard Data Commons.
-
-Even though the Harvard Data Commons is specific to Harvard University,
-the overall vision and the three objectives can be abstracted and
-applied to other universities or research organisations. The Commons
-will be designed and implemented using standards and commonly-used
-approaches to make it interoperable and reusable by others.
-
-\begin{figure}%[t!]
-\includegraphics{210053f04}
-\caption{One aspect of Harvard Data Commons. Automatic
-encapsulation and deposit of artefacts from data management tools used
-during active research at the Harvard Dataverse repository.}
-\label{fig:hdc}
-\end{figure}
-
- \section{Related work}% {#relatedwork}
-
-\label{sec:relatedwork}
-
-With the increasing digitisation of research processes, there has been
-a significant call for the wider adoption of interoperable sharing of
-data and its associated metadata. We refer to
-\cite{doi:10.1016/j.patter.2020.100136} for a comprehensive overview and
-recommendations, in particular for data; notably that review highlights
-the wide variety of metadata and documentation that the literature
-prescribes for enabling data reuse. Likewise, we suggest
-\cite{doi:10.1016/j.patter.2021.100322} that covers the importance of
-metadata standards in reproducible computational research.
-
-Here we focus on approaches for bundling research artefacts along with
-their metadata. This notion of publishing compound objects for
-scholarly communication has a long history behind it
-\cite{doi:10.1190/1.1822162,vandesompel_2007}, but recent approaches
-have followed three main strands: (1) publishing to centralised
-repositories; (2) packaging approaches similar to RO-Crate; and (3)
-bundling the computational workflow around a scientific experiment.
-
- \subsection{Bundling and packaging digital research artefacts}
-
-Early work making the case for publishing compound scholarly
-communication units \cite{vandesompel_2007} led to the development of the
-Object Re-Use and Exchange
-model\footnote{\url{http://www.openarchives.org/ore/1.0/primer}} (OAI-ORE), providing
-a structured \textbf{resource map} of the digital artefacts that together
-support a scholarly output.
-
-The challenge of describing computational workflows was one of the main
-motivations for the early proposal of \textit{Research Objects} (RO)
-\cite{doi:10.1016/j.future.2011.08.004} as first-class citizens for sharing
-and publishing. The RO approach involves bundling datasets, workflows,
-scripts and results along with traditional dissemination materials like
-journal articles and presentations, forming a single package.
-Crucially, these resources are not just gathered, but also individually
-typed, described and related to each other using semantic vocabularies.
-As pointed out in \cite{doi:10.1016/j.future.2011.08.004} an open-ended
-\textit{Linked Data} approach is not sufficient for scholarly communication: a
-common data model is also needed in addition to common and best
-practices for managing and annotating lifecycle, ownership, versioning
-and attributions.
-
-Considering the FAIR principles \cite{doi:10.1038/sdata.2016.18}, we can
-say with hindsight that the initial RO approaches strongly targeted
-\textit{Interoperability}, with a particular focus on the reproducibility of
-\textit{in-silico experiments} involving computational workflows and the reuse
-of existing RDF vocabularies.
-
-The first implementation of Research Objects for sharing workflows in
-myExperiment \cite{doi:10.1093/nar/gkq429} was based on RDF ontologies
-\cite{newman2009}, building on Dublin Core, FOAF, SIOC, Creative Commons
-and OAI-ORE to form myExperiment ontologies for describing social
-networking, attribution and credit, annotations, aggregation packs,
-experiments, view statistics, contributions, and workflow components
-\cite{myExperimentOntology2009}.
-
-This initially workflow-centric approach was further formalised as the
-Wf4Ever Research Object Model \cite{doi:10.1016/j.websem.2015.01.003},
-which is a general-purpose research artefact description framework.
-This model is based on existing ontologies (FOAF, Dublin Core Terms,
-OAI-ORE and AO/OAC precursors to the W3C Web Annotation Model
-\cite{Ciccarese:17:WAD}) and adds specializations for workflow models and
-executions using W3C PROV-O \cite{PROVO}. The Research Object statements
-are saved in a \textit{manifest} (the OAI-ORE \textit{resource map}), with additional
-annotation resources containing user-provided details such as title and
-description.
-
-We now claim that one barrier for wider adoption of the Wf4Eer Research
-Object model for general packaging digital research artefacts was
-exactly this re-use of multiple existing vocabularies (FAIR principle
-I2: \textit{Metadata use vocabularies that follow FAIR principles}), which in
-itself is recognised as a challenge \cite{doi:10.3233/978-1-61499-660-6-9}.
-Adapters of the Wf4Ever RO model would have to navigate documentation
-of multiple overlapping ontologies, in addition to facing the usual
-Semantic Web development choices for RDF serialization formats,
-identifier minting and publishing resources on the Web.
-
-Several developments for Research Objects improved on this situation,
-such as ROHub used by Earth Sciences
-\cite{doi:10.1016/j.future.2019.03.046}, which provides a 
-user-interface for making Research Objects, along with Research Object
-Bundle \cite{doi:10.5281/zenodo.12586} (RO Bundle), which is a ZIP-archive
-embedding data files and a JSON-LD serialization of the manifest with
-mappings for a limited set of terms. RO Bundle was also used for
-storing detailed workflow run provenance (TavernaPROV
-\cite{doi:10.5281/zenodo.51314}).
-
-RO-Bundle evolved to Research Object BagIt
-archives,\footnote{\url{https://w3id.org/ro/bagit}} a variant of RO Bundle as a BagIt
-archive \cite{doi:10.17487/rfc8493}, used by Big Data Bags
-\cite{doi:10.1109/BigData.2016.7840618}, CWLProv
-\cite{doi:10.1093/gigascience/giz095} and WholeTale
-\cite{doi:10.3233/APC200107,doi:10.1109/eScience.2019.00068}.
-
-\subsection{FAIR Digital Objects}
-
-FAIR Digital Objects (FDO) \cite{doi:10.3390/publications8020021} have been
-proposed as a conceptual framework for making digital resources
-available in a Digital Objects (DO) architecture which encourages
-active use of the objects and their metadata. In particular, an FDO has
-five parts: (i) The FDO \textit{content}, bit sequences stored in an
-accessible repository; (ii) a \textit{Persistent Identifier} (PID) such as a
-DOI that identifies the FDO and can resolve these same parts; (iii)
-Associated rich \textit{metadata}, as separate FDOs; (iv) Type definitions,
-also separate FDOs; (v) Associated \textit{operations} for the given types. A
-Digital Object typed as a Collection aggregates other DOs by reference.
-
-The Digital Object Interface Protocol \cite{doip2.0} can be considered an
-``abstract protocol'' of requirements, DOs could be implemented in
-multiple ways. One suggested implementation is the FAIR Digital Object
-Framework,\footnote{\url{https://fairdigitalobjectframework.org/}} based on HTTP and
-the Linked Data Principles. While there is agreement on using PIDs
-based on DOIs, consensus on how to represent common metadata, core
-types and collections as FDOs has not yet been reached. We argue that
-RO-Crate can play an important role for FDOs:
-\begin{enumerate}
-\item[1.] By providing a predictable and extensible serialisation of
-structured metadata.
-\item[2.] By formalising how to aggregate digital objects as collections (and
-adding their context).
-\item[3.] By providing a natural Metadata FDO in the form of the RO-Crate
-Metadata File.
-\item[4.] By being based on Linked Data and the Schema.org vocabulary, meaning
-that PIDs already exist for common types and properties.
-\end{enumerate}
-
-At the same time, it is clear that the goal of FDO is broader than that
-of RO-Crate; namely, FDOs are active objects with distributed
-operations, and add further constraints such as PIDs for every element.
-These features improve FAIR features of digital objects and are also
-useful for RO-Crate, but they also severely restrict the infrastructure
-that needs to be implemented and maintained in order for FDOs to remain
-accessible. RO-Crate, on the other hand, is more flexible: it can
-minimally be used within any file system structure, or ideally exposed
-through a range of Web-based scenarios. A \textit{FAIR profile of RO-Crate}
-(e.g. enforcing PID usage) will fit well within a FAIR Digital Object ecosystem.
-
- \subsection{Packaging workflows}
-
-The use of computational workflows, typically combining a chain of
-tools in an analytical pipeline, has gained prominence in particular in
-the life sciences. Workflows might be used primarily to improve
-computational scalability, as well as to assist in making computed
-data results FAIR \cite{doi:10.1162/dint_a_00033}, for instance by
-improving reproducibility \cite{doi:10.1016/j.future.2017.01.012}, but also
-because programmatic data usage help propagate their metadata and
-provenance \cite{doi:10.1002/cpe.1228}. At the same time, workflows raise
-additional FAIR challenges, since they can be considered important
-research artefacts themselves. This viewpoint poses the problem of
-capturing and explaining the computational methods of a pipeline in
-sufficient machine-readable detail \cite{doi:10.3233/DS-190026}.
-
-Even when researchers follow current best practices for workflow
-reproducibility \cite{doi:10.1016/j.cels.2018.03.014,doi:10.1016/j.future.2017.01.012}, the communication of computational
-outcomes through traditional academic publishing routes effectively
-adds barriers as authors are forced to rely on a textual manuscript
-representations. This hinder reproducibility and FAIR use of the
-knowledge previously captured in the workflow.
-
-As a real-life example, let us look at a metagenomics article
-\cite{doi:10.1038/s41586-019-0965-1} that describes a computational
-pipeline. Here the authors have gone to extraordinary efforts to
-document the individual tools that have been reused, including their
-citations, versions, settings, parameters and combinations. The
-\textit{Methods} section is two pages in tight double-columns with twenty four
-additional references, supported by the availability of data on an FTP
-server (60 GB) \cite{ebi_ftp_umgs2019} and of open source code in GitHub
-Finn-Lab/MGS-gut\footnote{\url{https://github.com/Finn-Lab/MGS-gut}}
-\cite{finn-lab-mgsgut}, including the pipeline as shell scripts and
-associated analysis scripts in R and Python.
-
-This attention to reporting detail for computational workflows is
-unfortunately not yet the norm, and although bioinformatics journals
-have strong \textit{data availability} requirements, they frequently do not
-require authors to include or cite \textit{software, scripts and pipelines}
-used for analysing and producing results \cite{soilandreyes_tweet_2020}.
-Indeed, in the absence of a specific requirement and an editorial
-policy to back it up -- such as eliminating the reference limit --
-authors are effectively discouraged from properly and comprehensively
-citing software \cite{doi:10.1038/s41592-019-0350-x}.
-
-However detailed this additional information might be, another
-researcher who wants to reuse a particular computational method may
-first want to assess if the described tool or workflow is Re-runnable
-(executable at all), Repeatable (same results for original inputs on
-same platform), Reproducible (same results for original inputs with
-different platform or newer tools) and ultimately Reusable (similar
-results for different input data), Repurposable (reusing parts of the
-method for making a new method) or Replicable (rewriting the workflow
-following the method description)
-\cite{doi:10.3389/fninf.2017.00069,goble_presentation_2016}.
-
-Following the textual description alone, researchers would be forced to
-jump straight to evaluate ``Replicable'' by rewriting the pipeline from
-scratch. This can be expensive and error-prone. They would firstly need
-to install all the software dependencies and download reference
-datasets. This can be a daunting task, which may have to be repeated
-multiple times as workflows typically are developed at small scale on
-desktop computers, scaled up to local clusters, and potentially put
-into production using cloud instances, each of which will have
-different requirements for software installations.
-
-In recent years the situation has been greatly improved by software
-packaging and container technologies like Docker and Conda, these
-technologies have been increasingly adopted in life sciences
-\cite{doi:10.1007/s41019-017-0050-4} thanks to collaborative efforts such
-as BioConda \cite{doi:10.1038/s41592-018-0046-7} and BioContainers
-\cite{doi:10.1093/bioinformatics/btx192}, and support by Linux
-distributions (e.g. Debian Med \cite{doi:10.1186/1471-2105-11-S12-S5}). As
-of November 2021, more than 9,000 software packages are available in
-BioConda alone,\footnote{\url{https://anaconda.org/bioconda/}} and 10,000 containers
-in BioContainers.\footnote{\url{https://biocontainers.pro/\#/registry}}
-
-Docker and Conda have been integrated into workflow systems such as
-Snakemake \cite{doi:10.1093/bioinformatics/bts480}, Galaxy
-\cite{doi:10.1093/nar/gky379} and Nextflow \cite{doi:10.1038/nbt.3820}, meaning
-a downloaded workflow definition can now be executed on a ``blank''
-machine (except for the workflow engine) with the underlying analytical
-tools installed on demand. Even with using containers there is a
-reproducibility challenge, for instance Docker Hub's retention policy
-will expire container images after six
-months,\footnote{\url{https://www.docker.com/blog/docker-hub-image-retention-policy-delayed-and-subscription-updates/}}
-or a lack of recording versions of transitive dependencies of Conda
-packages could cause incompatibilities if the packages are subsequently updated.
-
-These container and package systems only capture small amounts of
-metadata.\footnote{Docker and Conda can use \textit{build recipes}, a set of commands
-that construct the container image through downloading and installing
-its requirements. However these recipes are effectively another piece
-of software code, which may itself decay and become difficult to rerun.} In particular, they do not capture any of the semantic
-relationships between their content. Understanding these relationships
-is made harder by the opaque wrapping of arbitrary tools with unclear
-functionality, licenses and attributions.
-
-From this we see that computational workflows are themselves complex
-digital objects that need to be recorded not just as files, but in the
-context of their execution environment, dependencies and analytical
-purpose in research -- as well as other metadata (e.g. version,
-license, attribution and identifiers).
-
-It is important to note that having all these computational details in
-order to represent them in an RO-Crate is an ideal scenario -- in
-practice there will always be gaps of knowledge, and exposing all
-provenance details automatically would require improvements to the data
-sources, workflow, workflow engine and its dependencies. RO-Crate can
-be seen as a flexible annotation mechanism for augmenting automatic
-workflow provenance. Additional metadata can be added manually, e.g.
-for sensitive clinical data that cannot be publicly exposed\footnote{FAIR principle A2: \textit{Metadata are accessible, even when the
-data are no longer available} \cite{doi:10.1038/sdata.2016.18}.}, or
-to cite software that lack persistent identifiers. This inline
-\textit{FAIRifying} allows researchers to achieve ``just enough FAIR'' to
-explain their computational experiments.
-
-\section{Conclusion}
-
-\label{sec:conclusion}
-
-RO-Crate has been established as an approach to packaging digital
-research artefacts with structured metadata. This approach assists
-developers and researchers to produce and consume FAIR archives of
-their research.
-
-RO-Crate is formed by a set of best practice recommendations, developed
-by an open and broad community. These guidelines show how to use ``just
-enough'' standards in a consistent way. The use of
-structured metadata with a rich base vocabulary can cover
-general-purpose contextual relations, with a Linked Data foundation
-that ensures extensibility to domain- and application-specific uses. We
-can therefore consider an RO-Crate not just as a structured data
-archive, but as a multimodal scholarly knowledge graph that can help
-``FAIRify'' and combine metadata of existing resources.
-
-The adoption of simple Web technologies in the RO-Crate specification
-has helped a rapid development of a wide variety of supporting open
-source tools and libraries. RO-Crate fits into the larger landscape of
-open scholarly communication and FAIR Digital Object infrastructure,
-and can be integrated into data repository platforms. RO-Crate can be
-applied as a data/metadata exchange mechanism, assist in long-term
-archival preservation of metadata and data, or simply used at a small
-scale by individual researchers. Thanks to its strong community
-support, new and improved profiles and tools are being continuously
-added to the RO-Crate landscape, making it easier for adopters
-to find examples and support for their own use case.
-
-\subsection{Strictness vs flexibility}
-
-There is always a tradeoff between flexibility and strictness
-\cite{doi:10.1007/s11042-009-0397-2} when deciding on semantics of metadata
-models. Strict requirements make it easier for users and code to
-consume and populate a model, by reducing choices and having mandated
-``slots'' to fill in. But such rigidity can also restrict richness and
-applicability of the model, as it in turn enforce the initial
-assumptions about what can be described.
-
-RO-Crate attempts to strike a balance between these tensions, and
-provides a common metadata framework that encourages extensions.
-However, just like the RO-Crate specification can be thought of as a
-\textit{core profile} of Schema.org in JSON-LD, we cannot stress the
-importance of also establishing domain-specific RO-Crate profiles and
-conventions, as explored in Sections~\ref{sec:profiles} and \ref
-{sec:inuse}. Specialization comes hand-in-hand with the principle of
-\textit{graceful degradation}; RO-Crate applications and users are free to
-choose the semantic detail level they participate at, as long as they
-follow the common syntactic requirements.
-
- \section{Future work}% {#futurework}
-
-\label{sec:futurework}
-
-The direction of future RO-Crate work is determined by the community
-around it as a collaborative effort. We currently plan on further
-outreach, building training material (including a comprehensive
-entry-level tutorial) and maturing the reference implementation
-libraries. We will also collect and build examples of RO-Crate
-\textit{consumption}, e.g. Jupyter Notebooks that query multiple crates using
-knowledge graphs. In addition, we are exploring ways to support some
-entity types requested by users, e.g. detailed workflow runs or
-container provenance, which do not have a good match in Schema.org.
-Such support could be added, for instance, by integrating other
-vocabularies or by having separated (but linked) metadata files.
-
-Furthermore, we want to better understand how the community uses
-RO-Crate in practice and how it contrasts with other related efforts;
-this will help us to improve our specification and tools. By
-discovering commonalities in emerging usage (e.g. additional Schema.org
-types), the community helps to reduce divergence that could otherwise
-occur with proliferation of further RO-Crate profiles. We plan to
-gather feedback via user studies, with the Linked Open Data community
-or as part of EOSC Bring-your-own-Data training events.
-
-We operate in an open community where future and potential users of
-RO-Crate are actively welcomed to participate and contribute feedback
-and requirements. In addition, we are targeting a wider audience through
-extensive outreach
-activities\footnote{\url{https://www.researchobject.org/ro-crate/outreach.html}} and
-by initiating new connections. Recent contacts include American
-Geophysical Union (AGU) on Data Citation Reliquary
-\cite{doi:10.5281/zenodo.4916734}, National Institute of Standards and
-Technology (NIST) on material science, and
-InvenioRDM\footnote{\url{https://inveniosoftware.org/products/rdm/}} used by the
-Zenodo data repository. New Horizon Europe projects adapting RO-Crate
-include BY-COVID,\footnote{\url{https://by-covid.org/}} which aims to improve FAIR
-access to data on COVID-19 and other infectious diseases.
-
-The main addition in the upcoming 1.2 release of the RO-Crate
-specifications will be the formalization of
-profiles\footnote{\url{https://www.researchobject.org/ro-crate/1.2-DRAFT/profiles}}
-for different categories of crates. Additional entity types have been
-requested by users, e.g. workflow runs, business workflows, containers
-and software packages, tabular data structures; these are not always
-matched well with existing Schema.org types, but may benefit from other
-vocabularies or even separate metadata files, e.g. from Frictionless
-Data.\footnote{\url{https://frictionlessdata.io/}} We will be further aligning
-and collaborating with related research artefact description efforts
-like CodeMeta\footnote{\url{https://codemeta.github.io/}} for software metadata,
-Science-on-Schema.org\footnote{\url{https://science-on-schema.org/}}
-\cite{doi:10.5281/zenodo.4477164} for datasets, FAIR Digital
-Objects\footnote{\url{https://fairdo.org/}} \cite{doi:10.3390/publications8020021} and
-activities in EOSC task forces\footnote{\url{https://www.eosc.eu/task-force-faq}}
-including the EOSC Interoperability Framework \cite{doi:10.2777/620649}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%% Supplementary data
-%\begin{esm}%[title=]
-%\esmtitle{Supplementary data}
-%\esmdescription{} is available at: \url{http://dx.doi.org/???}.
-%\esmfiletype{pdf} %file type list:
-%%http://www.freeformatter.com/mime-types-list.html#mime-types-list
-%\esmfilename{MANUSCRIPTsupp.pdf}
-%\end{esm}
-
-%% Put support info here:
-%\begin{ack}
-%\end{ack}
-\begin{acks}
-This work has received funding from the European Commission's Horizon
-2020 research and innovation programme for projects
-BioExcel-2\footnote{\url{https://cordis.europa.eu/project/id/823830}}
-(H2020-INFRAEDI-2018-1 823830), IBISBA
-1.0\footnote{\url{https://cordis.europa.eu/project/id/730976}}
-(H2020-INFRAIA-2017-1-two-stage 730976),
-PREP-IBISBA\footnote{\url{https://cordis.europa.eu/project/id/871118}}
-(H2020-INFRADEV-2019-2 871118),
-EOSC-Life\footnote{\url{https://cordis.europa.eu/project/id/824087}}
-(H2020-INFRAEOSC-2018-2 824087),
-SyntheSys+\footnote{\url{https://cordis.europa.eu/project/id/823827}}
-(H2020-INFRAIA-2018-1 823827). From the Horizon Europe Framework
-Programme this work has received funding for
-BY-COVID\footnote{\url{https://cordis.europa.eu/project/id/101046203}}
-(HORIZON-INFRA-2021-EMERGENCY-01 101046203).
-
-Bj\"{o}rn Gr\"{u}ning is supported by DataPLANT (NFDI 7/1 --
-42077441\footnote{\url{https://gepris.dfg.de/gepris/projekt/442077441}}), part of the
-German National Research Data Infrastructure (NFDI), funded by the
-Deutsche Forschungsgemeinschaft (DFG).
-
-Ana Trisovic is funded by the Alfred P. Sloan Foundation (grant number
-P-2020-13988).\footnote{\url{https://sloan.org/grant-detail/9555}} Harvard Data
-Commons is supported by an award from Harvard University Information
-Technology (HUIT).
-\end{acks}
-
- \section*{Contributions}
-
-Author contributions to this article and the RO-Crate project according
-to the Contributor Roles Taxonomy CASRAI
-CrEDiT\footnote{\url{https://casrai.org/credit/}} \cite{doi:10.1087/20150211}:
-\begin{description}
-\item[Stian Soiland-Reyes]  Conceptualization, Data curation, Formal Analysis, Funding
-acquisition, Investigation, Methodology, Project administration,
-Software, Visualization, Writing -- original draft, Writing -- review
-\& editing
-
-\item[Peter Sefton]  Conceptualization, Investigation, Methodology, Project
-administration, Resources, Software, Writing -- review \& editing
-
-\item[Merc\`{e} Crosas]  Writing -- review \& editing
-
-\item[Leyla Jael Castro]  Methodology, Writing -- review \& editing
-
-\item[Frederik Coppens]  Writing -- review \& editing
-
-\item[Jos\'{e} M. Fern\'{a}ndez]  Methodology, Software, Writing -- review \& editing
-
-\item[Daniel Garijo]  Methodology, Writing -- review \& editing
-
-\item[Bj\"{o}rn Gr\"{u}ning]  Writing -- review \& editing
-
-\item[Marco La Rosa]  Software, Methodology, Writing -- review \& editing
-
-\item[Simone Leo]  Software, Methodology, Writing -- review \& editing
-
-\item[Eoghan \'{O} Carrag\'{a}in]  Investigation, Methodology, Project administration, Writing -- review
-\& editing
-
-\item[Marc Portier]  Methodology, Writing -- review \& editing
-
-\item[Ana Trisovic]  Software, Writing -- review \& editing
-
-\item[RO-Crate Community]  Investigation, Software, Validation, Writing -- review \& editing
-
-\item[Paul Groth]  Methodology, Supervision, Writing -- original draft, Writing --
-review \& editing
-
-\item[Carole Goble]  Conceptualization, Funding acquisition, Methodology, Project
-administration, Supervision, Visualization, Writing -- review \& editing
-\end{description}
-
-We would also like to acknowledge contributions from:
-\begin{description}
-\item[Finn Bacall]
- Software, Methodology
-
-\item[Herbert Van de Sompel]
- Writing -- review \& editing
-
-\item[Ignacio Eguinoa]
- Software, Methodology
-
-\item[Nick Juty]
- Writing -- review \& editing
-
-\item[Oscar Corcho]
- Writing -- review \& editing
-
-\item[Stuart Owen]
- Writing -- review \& editing
-
-\item[Laura Rodr\'iguez-Navas]
- Software, Visualization, Writing -- review \& editing
-
-\item[Alan R. Williams]
- Writing -- review \& editing
-\end{description}
-\begin{appendix}
-
-\section{Formalizing RO-Crate in First Order Logic}
-\label{sec:formaldefinition}
-
-Below is a formalization of the concept of RO-Crate as a set of
-relations using First Order Logic.
-
-\subsection{Language}
-
-Definition of language $\mathcal{L}_{\mathit{rocrate}}$:
-%
-\begin{eqnarray*}
-\mathcal{L}_{\mathit{rocrate}} &=& \bigl\{\mathit{Property}(p), \mathit{Class}(c),
-\mathit{Value}(x), \mathbb{R}, \mathbb{S} \bigr\}
-\\
-\mathbb{D} &=& \mathbb{IRI}
-\\
-\mathbb{IRI} &\equiv& { \text{IRIs as defined in RFC3987 \cite{doi:10.17487/rfc3987}} }
-\\
-\mathbb{R} &\equiv& { \text{real or integer numbers} }
-\\
-\mathbb{S} &\equiv& { \text{literal strings} }
-\end{eqnarray*}
-
-The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$
-identifiers (notation \texttt
-{<http://example.\noxml{\break}com/>}),\footnote{For simplicity, blank nodes are not included in this formalisation, as RO-Crate
-recommends the use of IRI identifiers: \url
-{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#describing-entities-in-json-ld}.} with additional descriptions using numbers $\mathbb{R}$ (notation
-$13.37$) and literal strings $\mathbb{S}$ (notation $\text{``Hello''}$).
-
-From this formalised language $\mathcal{L}_{\mathit{rocrate}}$ we can interpret
-an RO-Crate in any representation that can gather these descriptions,
-their properties, classes, and literal attributes.
-
-\subsection{Minimal RO-Crate}
-
-Below we use $\mathcal{L}_{\mathit{rocrate}}$ to define a minimal\footnote{The full
-list of types, relations and attribute properties from the
-RO-Crate specification are not included. Examples shown include
-$\textit{datePublished}$, $\textit{CreativeWork}$ and $\textit{name}$.
-} RO-Crate:
-%
-\begin{eqnarray*}
-\textit{ROCrate}(R) & \models& \textit{Root}(R) \land \textit{Mentions}(R, R)
-\land \textit{hasPart}(R, d) \land
-\\
-& & {}\textit{Mentions}(R, d) \land \textit{DataEntity}(d) \land
-\\
-& & {}\textit{Mentions}(R, c) \land \textit{ContextualEntity}(c)
-\\
-\forall r \ \textit{Root}(r) & \Rightarrow& \textit{Dataset}(r) \land
-\textit{name}(r, n) \land \textit{description}(r, d) \land
-\\
-& &{} \textit{datePublished}(r, date) \land \textit{license}(e, l)
-\\
-\forall e \forall n \ \textit{name}(e, n) & \Rightarrow& \textit{Value}(n)
-\\
-\forall e \forall s \ \textit{description}(e, s) & \Rightarrow&
-\textit{Value}(s)
-\\
-\forall e \forall d \ \textit{datePublished}(e, d) & \Rightarrow&
-\textit{Value}(d)
-\\
-\forall e \forall l \ \textit{license}(e, l) & \Rightarrow&
-\textit{ContextualEntity}(l)
-\\
-\textit{DataEntity}(e) & \equiv& \textit{File}(e) \oplus \textit{Dataset}(e)
-\\
-\textit{Entity}(e) & \equiv& \textit{DataEntity}(e) \lor \textit{ContextualEntity}(e)
-\\
-\forall e \ \textit{Entity}(e) & \Rightarrow& \textit{type}(e, c) \land
-\textit{Class}(c)
-\\
-\forall e \ \textit{ContextualEntity}(e) & \Rightarrow& \textit{name}(e, n)
-\\
-\textit{Mentions}(R, s) & \models& \textit{Relation}(s, p, e) \oplus
-\textit{Attribute}(s, p, l)
-\\
-\textit{Relation}(s, p, o) & \models& \textit{Entity}(s) \land
-\textit{Property}(p) \land \textit{Entity}(o)
-\\
-\textit{Attribute}(s, p, v) & \models& \textit{Entity}(s) \land
-\textit{Property}(p) \land \textit{Value}(v)
-\\
-\textit{Value}(v) & \equiv& v \in\mathbb{R} \oplus v \in\mathbb{S}
-\end{eqnarray*}
-
-An $\textit{ROCrate}(R)$ is defined as a self-described \emph{Root Data Entity},
-which contains parts (\emph{data entities}), which are
-further described in \emph{contextual entities}. These terms align with
-their use in the RO-Crate 1.1 terminology.\footnote{\url{https://www.researchobject.org/ro-crate/1.1/terminology}}
-
-The $\textit{Root}(r)$ is a type of $\textit{Dataset}(r)$, and must as metadata 
-have at least the attributes $\textit{name}$, $description$ and
-$\textit{datePublished}$ as literal values, as well as a contextual entity that identify its
-$\textit{license}$. These predicates correspond to the RO-Crate 1.1 minimal requirements
-for the root data entity.\footnote{\url
-{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}}
-
-The concept of an $\textit{Entity}(e)$ is introduced as being either a
-$\textit{DataEntity}(e)$, a $\textit{ContextualEntity}(e)$, or both.\footnote{\url
-{https://www.researchobject.org/ro-crate/1.1/contextual-entities.html\#contextual-vs-data-entities}} Any $\textit{Entity}(e)$ must be
-typed with at least one $\textit{Class}(c)$, and
-every $\textit{ContextualEntity}(e)$ must also have a $\textit{name}(e,n)$; this
-corresponds to expectations for any \emph{referenced contextual entity}
-(Section~\ref{sec:contextualentities}).
-
-For simplicity in this formalization (and to assist production rules
-below) $R$ is a constant representing a single RO-Crate, typically
-written to independent RO-Crate Metadata files. $R$ is used by
-$\textit{Mentions}(R, e)$ to indicate that $e$ is an Entity described by the
-RO-Crate and therefore its metadata (a set of $\textit{Relation}$ and
-$\textit{Attribute}$ predicates) form part of the RO-Crate serialization.
-$\textit{Relation}(s, p, o)$ and $\textit{Attribute}(s, p, x)$ are defined as a \emph
-{subject--predicate--object} triple pattern from an $\textit{Entity}(s)$ using
-a $\textit{Property}(p)$ to either another $\textit{Entity}(o)$ or a $\textit{Value}(x)$ value.
-
-\subsection{Example of formalised RO-Crate}
-
-The below is an example RO-Crate represented using the above
-formalisation, assuming a base IRI of \texttt{<http://example.com/ro/123/>}:
-%
-{\fontsize{9.7}{11.7}\selectfont{\begin{gather*}
-\textit{ROCrate}(\texttt{<http://example.com/ro/123/>})
-\\
-\textit{name}\bigl(\texttt{<http://example.com/ro/123/>},
-\\
-\quad \text{``Data files associated with the manuscript:Effects of {
-\ldots}''}\bigr)
-\\
-\textit{description}\bigl(\texttt{<http://example.com/ro/123/>},
-\\
-\quad \text{``Palliative care planning for nursing home residents {
-\ldots}''}\bigr)
-\\
-\textit{license}(\texttt{<http://example.com/ro/123/>},
-\\
-\quad \texttt{<https://spdx.org/licenses/CC-BY-4.0>})
-\\
-\textit{datePublished}\bigl(\texttt{<http://example.com/ro/123/>}, \text
-{``2017-02-23''}\bigr)
-\\
-\textit{hasPart}(\texttt{<http://example.com/ro/123/>}, \texttt{<http://example.com/ro/123/file.txt>})
-\\
-\textit{hasPart}(\texttt{<http://example.com/ro/123/>}, \texttt{<http://example.com/ro/123/interviews/>})
-\\
-\\
-\textit{ContextualEntity}(\texttt{<https://spdx.org/licenses/CC-BY-4.0>})
-\\
-\textit{name}\bigl(\texttt{<https://spdx.org/licenses/CC-BY-4.0>},
-\\
-\quad \text{``Creative Commons Attribution 4.0''}\bigr)
-\\
-\\
-\textit{ContextualEntity}(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>})
-\\
-\textit{name}\bigl(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>},
-\\
-\quad \text{``Creative Commons Attribution Non Commercial 4.0''}
-\bigr)
-\\
-\\
-\textit{File}(\texttt{<http://example.com/ro/123/survey.csv>})
-\\
-\textit{name}\bigl(\texttt{<http://example.com/ro/123/survey.csv>},
-\text{``Survey of care providers''}\bigr)
-\\
-\\
-\textit{Dataset}(\texttt{<http://example.com/ro/123/interviews/>})
-\\
-\textit{name}\bigl(\texttt{<http://example.com/ro/123/interviews/>},
-\\
-\quad \text{``Audio recordings of care provider interviews''}
-\bigr)
-\\
-\textit{license}(\texttt{<http://example.com/ro/123/interviews/>},
-\\*
-\quad \texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>})
-\end{gather*}}}
-
-Notable from this triple-like formalization is that a RO-Crate $R$ is
-fully represented as a tree at depth 2 helped by the use of
-%PDAG%
-$\mathbh{IRI}$
-%PDAG%
-nodes. For instance the aggregation from the root entity
-$\textit{hasPart}(\texttt{{\ldots}interviews/>})$ is at same level as the data
-entity's property \textit{license}(\texttt{{\ldots}CC-BY-\noxml{\break}NC-4.0>}) and that
-contextual entity's attribute $ \textit{name}(\text{{\ldots}Non Commercial
-4.0''})$. As shown in Section~\ref{sec:jsonld}, the RO-Crate Metadata
-File serialization is an equivalent shallow tree, although at depth 3
-to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.
-
-In reality many additional attributes and contextual types from
-Schema.org types like \url{http://schema.org/affiliation} and \url
-{http://schema.org/Organization} would be used to further describe the
-RO-Crate and its entities, but as these are optional (\textit{SHOULD}
-requirements) they do not form part of this formalization.
-
-\subsection{Mapping to RDF with Schema.org}
-
-A formalised RO-Crate in $\mathcal{L}_{\mathit{rocrate}}$ can be mapped to
-different serializations.
-Assume a simplified\footnote{This simplification and mapping does not cover the extensive list of literal
-datatypes built into RDF 1.1, only strings and decimal real numbers.
-Likewise, \textit{LanguageTag} is deliberately not utillised below.} language $\mathcal{L}_{\textit{RDF}}$
-based on the RDF abstract syntax \cite{rdfworkinggroup_2014}:
-%
-{\fontsize{9.8}{11.8}\selectfont{\begin{eqnarray*}
-\mathcal{L}_{\textit{RDF}} & \equiv& \bigl\{ \textit{Triple}(s,p,o),
-\textit{IRI}(i), \textit{BlankNode}(b), \textit{Literal}(s), \mathbb{IRI},
-\mathbb{S}, \mathbb{R} \bigr\}
-\\
-\mathbb{D}_{\textit{RDF}} & \equiv& \mathbb{S}
-\\
-\forall i \ \textit{IRI}(i) & \Rightarrow& i \in\mathbb{IRI}
-\\
-\forall s \forall p \forall o \ \textit{Triple}(s,p,o) & \Rightarrow& \bigl(
-\textit{IRI}(s) \lor \textit{BlankNode}(s) \bigr) \land
-\\
-& & \textit{IRI}(p) \land
-\\
-& & \bigl(\textit{IRI}(o) \lor \textit{BlankNode}(o) \lor \textit{Literal}(o)
-\bigr)
-\\
-\textit{Literal}(v) & \models& \textit{Value}(v) \land \textit{Datatype}(v,t)
-\land \textit{IRI}(t)
-\\
-\forall v \ \textit{Value}(v) & \Rightarrow& v \in\mathbb{S}
-\\
-\textit{LanguageTag}(v, l) & \equiv& Datatype (v,
-\\
-&& \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#langString>} )
-\end{eqnarray*}}}
-
-Below follows a mapping from $\mathcal{L}_{\mathit{rocrate}}$ to $\mathcal
-{L}_{\textit{RDF}}$ using Schema.org as vocabulary:
-%
-{\fontsize{9.6}{11.6}\selectfont{\begin{eqnarray*}
-\textit{Property}(p) & \Rightarrow& \textit{type}(p, \texttt
-{<http://www.w3.org/2000/01/rdf-schema\#Property>})
-\\
-\textit{Class}(c) & \Rightarrow& \textit{type}(c, \texttt {<http://www.w3.org/2000/01/rdf-schema\#Class>})
-\\
-\textit{Dataset}(d) & \Rightarrow& \textit{type}(d, \texttt {<http://schema.org/Dataset>})
-\\
-\textit{File}(f) & \Rightarrow& \textit{type}(f, \texttt {<http://schema.org/MediaObject>})
-\\
-\textit{ContextualEntity}(e) & \Rightarrow& \textit{type}(f, \texttt
-{<http://schema.org/Thing>})
-\\
-\textit{CreativeWork}(e) & \Rightarrow& \textit{ContextualEntity}(e) \land
-\textit{type}(e, \texttt{<http://schema.org/CreativeWork>})
-\\
-\textit{hasPart}(e, t) & \Rightarrow& \textit{Relation}(e, \texttt
-{<http://schema.org/hasPart>}, t)
-\\
-\textit{name}(e, n) & \Rightarrow& \textit{Attribute}(e, \texttt
-{<http://schema.org/name>}, n)
-\\
-\textit{description}(e, s) & \Rightarrow& \textit{Attribute}(e, \texttt
-{<http://schema.org/description>}, s)
-\\
-\textit{datePublished}(e, d) & \Rightarrow& Attribute(e, \texttt
-{<http://schema.org/datePublished>}, d)
-\\
-\textit{license}(e, l) & \Rightarrow& \textit{Relation}(e, \texttt
-{<http://schema.org/license>}, l) \land \textit{CreativeWork}(l)
-\\
-\textit{type}(e, t) & \Rightarrow& \textit{Relation}(e, \texttt
-{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#type>}, t)
-\\
-& & {}\land \textit{Class}(t)
-\\
-\textit{String}(s) & \equiv& \textit{Value}(s) \land s \in\mathbb{S}
-\\
-\textit{String}(s) & \Rightarrow& \textit{Datatype}(s, \texttt
-{<http://www.w3.org/2001/XMLSchema\#string>})
-\\
-\textit{Decimal}(d) & \equiv& \textit{Value}(d) \land d \in\mathbb{R}
-\\
-\textit{Decimal}(d) & \Rightarrow& \textit{Datatype}(d, \texttt
-{<http://www.w3.org/2001/XMLSchema\#decimal>})
-\\
-\textit{Relation}(s,p,o) & \Rightarrow& \textit{Triple}(s,p,o) \land
-\textit{IRI}(s) \land \textit{IRI}(o)
-\\
-\textit{Attribute}(s,p,o) & \Rightarrow& \textit{Triple}(s,p,o) \land
-\textit{IRI}(s) \land \textit{Literal}(o)
-\end{eqnarray*}}}
-
-Note that in the JSON-LD serialization of RO-Crate, the expression of
-$\textit{Class}$ and $\textit{Property}$ is typically indirect: The JSON-LD \texttt
-{@context} maps to Schema.org IRIs, which, when resolved as Linked
-Data, embed their formal definition as RDFa. Extensions may however
-include such term definitions directly in the RO-Crate.
-
-\subsection{RO-Crate 1.1 Metadata File Descriptor}
-
-An important RO-Crate principle is that of being \textbf
-{self-described}. Therefore the serialisation of the RO-Crate into a
-file should also describe itself in a Metadata File Descriptor,\footnote{\url
-{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#ro-crate-metadata-file-descriptor}} indicating it is $\textit{about}$ (describing) the RO-Crate root data entity,
-and that it $\textit{conformsTo}$ a particular version of the RO-Crate specification:
-%
-\begin{eqnarray*}
-\textit{about}(s,o) & \Rightarrow& \textit{Relation}(s, \texttt
-{<http://schema.org/about>}, o)
-\\
-\textit{conformsTo}(s,o) & \Rightarrow& \textit{Relation}(s, \texttt
-{<http://purl.org/dc/terms/conformsTo>}, o)
-\\
-\textit{MetadataFile}(m) & \Rightarrow& \textit{CreativeWork}(m) \land
-\textit{about}(m,R) \land \textit{ROCrate}(R) \land
-\\
-& &{} \textit{conformsTo}(m, \texttt{<https://w3id.org/ro/crate/1.1>})
-\end{eqnarray*}
-
-Note that although the metadata file necessarily is an \emph
-{information resource} written to disk or served over the network (as
-JSON-LD), it is not considered to be a contained \emph{part} of the
-RO-Crate in the form of a \emph{data entity}, rather it is described
-only as a \emph{contextual entity}.
-
-In the conceptual model, the \emph{RO-Crate Metadata File} can be seen
-as the top-level node that describes the \emph{RO-Crate Root}, however
-in the formal model (and the JSON-LD format) the metadata file
-descriptor is an additional contextual entity that is not affecting the
-depth-limit of the RO-Crate.
-
-\subsection{Forward-chained production rules for JSON-LD}
-
-Combining the above predicates and Schema.org mapping with rudimentary
-JSON templates, these forward-chaining production rules can output
-JSON-LD according to the RO-Crate 1.1 specification:\footnote{\textbf{Limitations:}
- Contextual entities not related from the
-RO-Crate (e.g. using inverse relations to a data entity) would not be
-covered by the single direction $Mentions(R, s)$ production rule; see
-\href{https://github.com/ResearchObject/ro-crate/issues/122}{GitHub
-issue ResearchObject/ro-crate\#122}. The $\textit{datePublished}(e, d)$ rule do
-not include syntax checks for the ISO 8601 datetime format. Compared
-with RO-Crate examples, this generated JSON-LD does not use a
-$\textit{@context}$ as the IRIs are produced unshortened, a post-step could do
-JSON-LD Flattening with a versioned RO-Crate context. The \texttt
-{@type} expansion is included for clarity, even though this is also
-implied by the $\textit{type}(e, t)$ expansion to $\textit{Relation}(e, \texttt{xsd:type})$.}
-%
-\begin{eqnarray*}
-\textit{Mentions}(R, s) \land \textit{Relation}(s,p,o) & \Rightarrow&
-\textit{Mentions}(R, o)
-\\
-\textit{IRI}(i) & \Rightarrow& \texttt{"} i \texttt{"}
-\\
-\textit{Decimal}(d) & \Rightarrow& d
-\\
-\textit{String}(s) & \Rightarrow& \texttt{"} s \texttt{"}
-\\
-\forall e \forall t \ \textit{type}(e, t) & \Rightarrow& \texttt{\{"@id":} e
-\texttt{,}
-\\
-&& \ \ \texttt{"@type":} t
-\\
-&& \texttt{\}}
-\\
-\forall s \forall p \forall o \ \textit{Relation}(s,p,o) & \Rightarrow& 
-\texttt{\{"@id":} s \texttt{,}
-\\
-&& \ \ p \texttt{: \{"@id":} o \texttt{\}}
-\\
-&& \texttt{\}}
-\\
-\forall s \forall p \forall v \ \textit{Attribute}(s,p,v) & \Rightarrow&
-\texttt{\{"@id":} s \texttt{,}
-\\
-&& \ \ p \texttt{:} v
-\\
-&& \texttt{\}}
-\\
-\forall r \forall c\ \textit{ROCrate}(r) & \Rightarrow& 
-\texttt{\{"@graph": [}
-\\
-&& \ \ \ \ \ \ \textit{Mentions}(r, c)*
-\\
-&& \ \ \ \texttt{]}
-\\
-&& \texttt{\}}
-\\
-R & \equiv& \texttt{<./>}
-\\
-R & \Rightarrow& \textit{MetadataFile}(\texttt{<ro-crate-metadata.json>})
-\end{eqnarray*}
-
-This exposes the first order logic domain of discourse of IRIs, with
-rational numbers and strings as their corresponding JSON-LD
-representation. These production rules first grow the graph of $R$ by
-adding a transitive rule -- anything described in $R$ which is related
-to $o$, means that $o$ is also mentioned by the $\textit{ROCrate}(R)$. For
-simplicity this rule is one-way; in theory the graph can also contain
-free-standing contextual entities that have outgoing relations to data-
-and contextual entities, but these are proposed to be bound to the root
-data entity with Schema.org relation \url{http://schema.org/mentions}.
-
-\section{RO-Crate Community}
-
-\label{sec:communitylist}
-
-As of 2021-10-04, the \textit{RO-Crate} Community members are:
-\begin{itemize}
-\item[*] Peter Sefton \url{https://orcid.org/0000-0002-3545-944X} (co-chair)
-\item[*] Stian Soiland-Reyes \url{https://orcid.org/0000-0001-9842-9718} (co-chair)
-\item[*] Eoghan \'{O} Carrag\'{a}in \url
-{https://orcid.org/0000-0001-8131-2150} (emeritus chair)
-\item[*] Oscar Corcho \url{https://orcid.org/0000-0002-9260-0753}
-\item[*] Daniel Garijo \url{https://orcid.org/0000-0003-0454-7145}
-\item[*] Raul Palma \url{https://orcid.org/0000-0003-4289-4922}
-\item[*] Frederik Coppens \url{https://orcid.org/0000-0001-6565-5145}
-\item[*] Carole Goble \url{https://orcid.org/0000-0003-1219-2137}
-\item[*] Jos\'{e} Mar\'ia Fern\'{a}ndez \url{https://orcid.org/0000-0002-4806-5140}
-\item[*] Kyle Chard \url{https://orcid.org/0000-0002-7370-4805}
-\item[*] Jose Manuel Gomez-Perez \url{https://orcid.org/0000-0002-5491-6431}
-\item[*] Michael R Crusoe \url{https://orcid.org/0000-0002-2961-9670}
-\item[*] Ignacio Eguinoa \url{https://orcid.org/0000-0002-6190-122X}
-\item[*] Nick Juty \url{https://orcid.org/0000-0002-2036-8350}
-\item[*] Kristi Holmes \url{https://orcid.org/0000-0001-8420-5254}
-\item[*] Jason A. Clark \url{https://orcid.org/0000-0002-3588-6257}
-\item[*] Salvador Capella-Gutierrez \url{https://orcid.org/0000-0002-0309-604X}
-\item[*] Alasdair J. G. Gray \url{https://orcid.org/0000-0002-5711-4872}
-\item[*] Stuart Owen \url{https://orcid.org/0000-0003-2130-0865}
-\item[*] Alan R Williams \url{https://orcid.org/0000-0003-3156-2105}
-\item[*] Giacomo Tartari \url{https://orcid.org/0000-0003-1130-2154}
-\item[*] Finn Bacall \url{https://orcid.org/0000-0002-0048-3300}
-\item[*] Thomas Thelen \url{https://orcid.org/0000-0002-1756-2128}
-\item[*] Herv\'{e} M\'{e}nager \url{https://orcid.org/0000-0002-7552-1009}
-\item[*] Laura Rodr\'iguez-Navas \url{https://orcid.org/0000-0003-4929-1219}
-\item[*] Paul Walk \url{https://orcid.org/0000-0003-1541-5631}
-\item[*] brandon whitehead \url{https://orcid.org/0000-0002-0337-8610}
-\item[*] Mark Wilkinson \url{https://orcid.org/0000-0001-6960-357X}
-\item[*] Paul Groth \url{https://orcid.org/0000-0003-0183-6910}
-\item[*] Erich Bremer \url{https://orcid.org/0000-0003-0223-1059}
-\item[*] LJ Garcia Castro \url{https://orcid.org/0000-0003-3986-0510}
-\item[*] Karl Sebby \url{https://orcid.org/0000-0001-6022-9825}
-\item[*] Alexander Kanitz \url{https://orcid.org/0000-0002-3468-0652}
-\item[*] Ana Trisovic \url{https://orcid.org/0000-0003-1991-0533}
-\item[*] Gavin Kennedy \url{https://orcid.org/0000-0003-3910-0474}
-\item[*] Mark Graves \url{https://orcid.org/0000-0003-3486-8193}
-\item[*] Jasper Koehorst \url{https://orcid.org/0000-0001-8172-8981}
-\item[*] Simone Leo \url{https://orcid.org/0000-0001-8271-5429}
-\item[*] Marc Portier \url{https://orcid.org/0000-0002-9648-6484}
-\item[*] Paul Brack \url{https://orcid.org/0000-0002-5432-2748}
-\item[*] Milan Ojster\v{s}ek \url{https://orcid.org/0000-0003-1743-8300}
-\item[*] Bert Droesbeke \url{https://orcid.org/0000-0003-0522-5674}
-\item[*] Chenxu Niu \url{https://github.com/UstcChenxu}
-\item[*] Kosuke Tanabe \url{https://orcid.org/0000-0002-9986-7223}
-\item[*] Tomasz Miksa \url{https://orcid.org/0000-0002-4929-7875}
-\item[*] Marco La Rosa \url{https://orcid.org/0000-0001-5383-6993}
-\item[*] Cedric Decruw \url{https://orcid.org/0000-0001-6387-5988}
-\item[*] Andreas Czerniak \url{https://orcid.org/0000-0003-3883-4169}
-\item[*] Jeremy Jay \url{https://orcid.org/0000-0002-5761-7533}
-\item[*] Sergio Serra \url{https://orcid.org/0000-0002-0792-8157}
-\item[*] Ronald Siebes \url{https://orcid.org/0000-0001-8772-7904}
-\item[*] Shaun de Witt \url{https://orcid.org/0000-0003-4196-3658}
-\item[*] Shady El Damaty \url{https://orcid.org/0000-0002-2318-4477}
-\item[*] Douglas Lowe \url{https://orcid.org/0000-0002-1248-3594}
-\item[*] Sergio Serra \url{https://orcid.org/0000-0002-0792-8157}
-\item[*] Xuanqi Li \url{https://orcid.org/0000-0003-1498-6205}
-\end{itemize}
-\end{appendix}
-
-% structpyb loaded by imikolaityte, 2021-12-28 10:21:28
-%
-\begin{thebibliography}{}
-
-%%% bbsrt2.pl, ver. 2.5.8, 2017.08.09
-%b1 ###bbsrt2
-%b122 ###
-\bibitem{doi:10.1093/nar/gky379}
-%
-\begin{barticle}
-\bauthor{\binits{E.}~\bsnm{Afgan}},
-\bauthor{\binits{D.}~\bsnm{Baker}},
-\bauthor{\binits{B.}~\bsnm{Batut}},
-\bauthor{\binits{M.}~\bsnm{van den Beek}},
-\bauthor{\binits{D.}~\bsnm{Bouvier}},
-\bauthor{\binits{M.}~\bsnm{Cech}},
-\bauthor{\binits{J.}~\bsnm{Chilton}},
-\bauthor{\binits{D.}~\bsnm{Clements}},
-\bauthor{\binits{N.}~\bsnm{Coraor}},
-\bauthor{\binits{B.A.}~\bsnm{Gr\"{u}ning}},
-\bauthor{\binits{A.}~\bsnm{Guerler}},
-\bauthor{\binits{J.}~\bsnm{Hillman-Jackson}},
-\bauthor{\binits{S.}~\bsnm{Hiltemann}},
-\bauthor{\binits{V.}~\bsnm{Jalili}},
-\bauthor{\binits{H.}~\bsnm{Rasche}},
-\bauthor{\binits{N.}~\bsnm{Soranzo}},
-\bauthor{\binits{J.}~\bsnm{Goecks}},
-\bauthor{\binits{J.}~\bsnm{Taylor}},
-\bauthor{\binits{A.}~\bsnm{Nekrutenko}} and
-\bauthor{\binits{D.}~\bsnm{Blankenberg}},
-\batitle{The Galaxy platform for accessible, reproducible and collaborative
-biomedical analyses: 2018 update},
-\bjtitle{Nucleic Acids Research}
-\bvolume{46}(\bissue{W1})
-(\byear{2018}),
-\bfpage{W537}--\blpage{W544}.
-\bid{doi={10.1093/nar/gky379}}
-\end{barticle}
-%
-\OrigBibText
-E.~Afgan,
-D.~Baker,
-B.~Batut,
-M.~van~den Beek,
-D.~Bouvier,
-M.~Cech,
-J.~Chilton,
-D.~Clements,
-N.~Coraor,
-B.A.~Gr\"{u}ning,
-A.~Guerler,
-J.~Hillman-Jackson,
-S.~Hiltemann,
-V.~Jalili,
-H.~Rasche,
-N.~Soranzo,
-J.~Goecks,
-J.~Taylor,
-A.~Nekrutenko and
-D.~Blankenberg,
-\textit{The Galaxy platform for accessible, reproducible and collaborative
-biomedical analyses: 2018 update.},
-Nucleic Acids Research
-46(W1)
-(2018),
-W537--W544.
-doi:10.1093/nar/gky379.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b2 ###bbsrt2
-%b124 ###
-\bibitem{doi:10.5281/zenodo.4916734}
-%
-\begin{botherref}
-\oauthor{\binits{D.}~\bsnm{Agarwal}},
-\oauthor{\binits{C.}~\bsnm{Goble}},
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\oauthor{\binits{U.}~\bsnm{Sarkans}},
-\oauthor{\binits{D.}~\bsnm{Noesgaard}},
-\oauthor{\binits{U.}~\bsnm{Schindler}},
-\oauthor{\binits{M.}~\bsnm{Fenner}},
-\oauthor{\binits{P.}~\bsnm{Manghi}},
-\oauthor{\binits{S.}~\bsnm{Stall}},
-\oauthor{\binits{C.}~\bsnm{Coward}} and
-\oauthor{\binits{C.}~\bsnm{Erdmann}},
-Data Citation Community of Practice -- 8,
-Workshop, Zenodo/AGU, 2021.
-\url{https://data.agu.org/DataCitationCoP/2nd-workshop-data-citation}.
-\bid{doi={10.5281/zenodo.4916734}}
-\end{botherref}
-%
-\OrigBibText
-D.~Agarwal,
-C.~Goble,
-S.~Soiland-Reyes,
-U.~Sarkans,
-D.~Noesgaard,
-U.~Schindler,
-M.~Fenner,
-P.~Manghi,
-S.~Stall,
-C.~Coward and
-C.~Erdmann,
-Data Citation Community of Practice -- 8 June 2021 Workshop,
-Zenodo/AGU,
-2021.
-doi:10.5281/zenodo.4916734.
-\url{https://data.agu.org/DataCitationCoP/2nd-workshop-data-citation}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b3 ###bbsrt2
-%b87 ###
-\bibitem{dcat2}
-%
-\begin{botherref}
-\oauthor{\binits{R.}~\bsnm{Albertoni}},
-\oauthor{\binits{D.}~\bsnm{Browning}},
-\oauthor{\binits{S.}~\bsnm{Cox}},
-\oauthor{\binits{A.}~\bsnm{Gonzalez Beltran}},
-\oauthor{\binits{A.}~\bsnm{Perego}},
-\oauthor{\binits{P.}~\bsnm{Winstanley}} and
-\binstitute{Dataset Exchange Working Group},
-Data Catalog Vocabulary
-({DCAT}) -- Version 2,
-\textit{{W3C} Recommendation}
-(2020),
-\url{https://www.w3.org/TR/2020/REC-vocab-dcat-2-20200204/}.
-\end{botherref}
-%
-\OrigBibText
-R.~Albertoni,
-D.~Browning,
-S.~Cox,
-A.~Gonzalez~Beltran,
-A.~Perego,
-P.~Winstanley and
-Dataset Exchange Working Group:
-Data Catalog Vocabulary ({DCAT}) - Version 2,
-{W3C} Recommendation,
-2020.
-\url{https://www.w3.org/TR/2020/REC-vocab-dcat-2-20200204/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b4 ###bbsrt2
-%b111 ###
-\bibitem{doi:10.1038/s41586-019-0965-1}
-%
-\begin{barticle}
-\bauthor{\binits{A.}~\bsnm{Almeida}},
-\bauthor{\binits{A.L.}~\bsnm{Mitchell}},
-\bauthor{\binits{M.}~\bsnm{Boland}},
-\bauthor{\binits{S.C.}~\bsnm{Forster}},
-\bauthor{\binits{G.B.}~\bsnm{Gloor}},
-\bauthor{\binits{A.}~\bsnm{Tarkowska}},
-\bauthor{\binits{T.D.}~\bsnm{Lawley}} and
-\bauthor{\binits{R.D.}~\bsnm{Finn}},
-\batitle{A new genomic blueprint of the human gut microbiota},
-\bjtitle{Nature}
-\bvolume{568}(\bissue{7753})
-(\byear{2019}),
-\bfpage{499}--\blpage{504}.
-\bid{doi={10.1038/s41586-019-0965-1}}
-\end{barticle}
-%
-\OrigBibText
-A.~Almeida,
-A.L.~Mitchell,
-M.~Boland,
-S.C.~Forster,
-G.B.~Gloor,
-A.~Tarkowska,
-T.D.~Lawley and
-R.D.~Finn,
-\textit{A new genomic blueprint of the human gut microbiota.},
-Nature
-568(7753)
-(2019),
-499--504.
-doi:10.1038/s41586-019-0965-1.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b5 ###bbsrt2
-%b77 ###
-\bibitem{doi:10.1371/journal.pbio.3000099}
-%
-\begin{barticle}
-\bauthor{\binits{G.}~\bsnm{Alterovitz}},
-\bauthor{\binits{D.}~\bsnm{Dean}},
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\bauthor{\binits{A.}~\bsnm{Bell}},
-\bauthor{\binits{A.}~\bsnm{Hayes}},
-\bauthor{\binits{A.}~\bsnm{Suresh}},
-\bauthor{\binits{A.}~\bsnm{Purkayastha}},
-\bauthor{\binits{C.H.}~\bsnm{King}},
-\bauthor{\binits{D.}~\bsnm{Taylor}},
-\bauthor{\binits{E.}~\bsnm{Johanson}},
-\bauthor{\binits{E.E.}~\bsnm{Thompson}},
-\bauthor{\binits{E.}~\bsnm{Donaldson}},
-\bauthor{\binits{H.}~\bsnm{Morizono}},
-\bauthor{\binits{H.}~\bsnm{Tsang}},
-\bauthor{\binits{J.K.}~\bsnm{Vora}},
-\bauthor{\binits{J.}~\bsnm{Goecks}},
-\bauthor{\binits{J.}~\bsnm{Yao}},
-\bauthor{\binits{J.S.}~\bsnm{Almeida}},
-\bauthor{\binits{J.}~\bsnm{Keeney}},
-\bauthor{\binits{K.}~\bsnm{Addepalli}},
-\bauthor{\binits{K.}~\bsnm{Krampis}},
-\bauthor{\binits{K.M.}~\bsnm{Smith}},
-\bauthor{\binits{L.}~\bsnm{Guo}},
-\bauthor{\binits{M.}~\bsnm{Walderhaug}},
-\bauthor{\binits{M.}~\bsnm{Schito}},
-\bauthor{\binits{M.}~\bsnm{Ezewudo}},
-\bauthor{\binits{N.}~\bsnm{Guimera}},
-\bauthor{\binits{P.}~\bsnm{Walsh}},
-\bauthor{\binits{R.}~\bsnm{Kahsay}},
-\bauthor{\binits{S.}~\bsnm{Gottipati}},
-\bauthor{\binits{T.C.}~\bsnm{Rodwell}},
-\bauthor{\binits{T.}~\bsnm{Bloom}},
-\bauthor{\binits{Y.}~\bsnm{Lai}},
-\bauthor{\binits{V.}~\bsnm{Simonyan}} and
-\bauthor{\binits{R.}~\bsnm{Mazumder}},
-\batitle{Enabling precision medicine via standard communication of {HTS}
-provenance, analysis, and results},
-\bjtitle{PLOS Biology}
-\bvolume{16}(\bissue{12})
-(\byear{2018}),
-\bnumber{e3000099}.
-\bid{doi={10.1371/journal.pbio.3000099}}
-\end{barticle}
-%
-\OrigBibText
-G.~Alterovitz,
-D.~Dean,
-C.~Goble,
-M.R.~Crusoe,
-S.~Soiland-Reyes,
-A.~Bell,
-A.~Hayes,
-A.~Suresh,
-A.~Purkayastha,
-C.H.~King,
-D.~Taylor,
-E.~Johanson,
-E.E.~Thompson,
-E.~Donaldson,
-H.~Morizono,
-H.~Tsang,
-J.K.~Vora,
-J.~Goecks,
-J.~Yao,
-J.S.~Almeida,
-J.~Keeney,
-K.~Addepalli,
-K.~Krampis,
-K.M.~Smith,
-L.~Guo,
-M.~Walderhaug,
-M.~Schito,
-M.~Ezewudo,
-N.~Guimera,
-P.~Walsh,
-R.~Kahsay,
-S.~Gottipati,
-T.C.~Rodwell,
-T.~Bloom,
-Y.~Lai,
-V.~Simonyan and
-R.~Mazumder,
-\textit{Enabling precision medicine via standard communication of {HTS}
-provenance, analysis, and results.},
-{PLOS} Biology
-16(12)
-(2018),
-e3000099.
-doi:10.1371/journal.pbio.3000099.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b6 ###bbsrt2
-%b14 ###
-\bibitem{doi:10.1007/s10209-016-0475-y}
-%
-\begin{barticle}
-\bauthor{\binits{R.C.}~\bsnm{Amorim}},
-\bauthor{\binits{J.A.}~\bsnm{Castro}},
-\bauthor{\binits{J.}~\bsnm{Rocha da Silva}} and
-\bauthor{\binits{C.}~\bsnm{Ribeiro}},
-\batitle{A comparison of research data management platforms: Architecture,
-flexible metadata and interoperability},
-\bjtitle{Universal Access in the Information Society}
-(\byear{2016}),
-\bfpage{1}--\blpage{12}.
-\bid{doi={10.1007/s10209-016-0475-y}}
-\end{barticle}
-%
-\OrigBibText
-R.C.~Amorim,
-J.A.~Castro,
-J.~Rocha~da Silva and
-C.~Ribeiro,
-A comparison of research data management platforms: architecture, flexible
-metadata and interoperability,
-\textit{Universal Access in the Information Society}
-\textbf{00}
-(2016),
-1--12.
-doi:10.1007/s10209-016-0475-y.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b7 ###bbsrt2
-%b56 ###
-\bibitem{doi:10.5281/zenodo.3922136}
-%
-\begin{bbook}
-\bauthor{\binits{G.}~\bsnm{Arfaoui}} and
-\bauthor{\binits{M.}~\bsnm{Jaoua}},
-\bbtitle{{RO}-Crate {RDA} {maDMP} Mapper, Zenodo},
-\byear{2020}.
-\url{https://github.com/GhaithArf/ro-crate-rda-madmp-mapper}.
-\bid{doi={10.5281/zenodo.3922136}}
-\end{bbook}
-%
-\OrigBibText
-G.~Arfaoui and
-M.~Jaoua,
-\textit{{RO}-Crate {RDA} {maDMP} Mapper},
-Zenodo
-(2020).
-doi:10.5281/zenodo.3922136.
-\url{https://github.com/GhaithArf/ro-crate-rda-madmp-mapper}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b8 ###bbsrt2
-%b55 ###
-\bibitem{ro-composer}
-%
-\begin{botherref}
-\oauthor{\binits{F.}~\bsnm{Bacall}},
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}} and
-\oauthor{\binits{M.}~\bsnm{Soares e Silva}},
-eScienceLab: RO-Composer. \url
-{https://esciencelab.org.uk/projects/ro-composer/}.
-\end{botherref}
-%
-\OrigBibText
-F.~Bacall,
-S.~Soiland-Reyes and
-M.~Soares~e Silva,
-eScienceLab: RO-Composer.
-\url{https://esciencelab.org.uk/projects/ro-composer/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b9 ###bbsrt2
-%b46 ###
-\bibitem{ro-crate-ruby}
-%
-\begin{botherref}
-\oauthor{\binits{F.}~\bsnm{Bacall}} and
-\oauthor{\binits{M.}~\bsnm{Whitwell}},
-{GitHub} -- {ResearchObject}/ro-crate-ruby: A Ruby gem for creating,
-manipulating and reading {RO}-Crates.
-\url{https://github.com/ResearchObject/ro-crate-ruby}.
-\end{botherref}
-%
-\OrigBibText
-F.~Bacall and
-M.~Whitwell,
-{GitHub} - {ResearchObject}/ro-crate-ruby: A Ruby gem for creating,
-manipulating and reading {RO}-Crates.
-\url{https://github.com/ResearchObject/ro-crate-ruby}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b10 ###bbsrt2
-%b64 ###
-\bibitem{doi:10.1371/journal.ppat.1008643}
-%
-\begin{barticle}
-\bauthor{\binits{D.}~\bsnm{Baker}},
-\bauthor{\binits{M.}~\bsnm{van den Beek}},
-\bauthor{\binits{D.}~\bsnm{Blankenberg}},
-\bauthor{\binits{D.}~\bsnm{Bouvier}},
-\bauthor{\binits{J.}~\bsnm{Chilton}},
-\bauthor{\binits{N.}~\bsnm{Coraor}},
-\bauthor{\binits{F.}~\bsnm{Coppens}},
-\bauthor{\binits{I.}~\bsnm{Eguinoa}},
-\bauthor{\binits{S.}~\bsnm{Gladman}},
-\bauthor{\binits{B.}~\bsnm{Gr\"{u}ning}},
-\bauthor{\binits{N.}~\bsnm{Keener}},
-\bauthor{\binits{D.}~\bsnm{Larivi\`{e}re}},
-\bauthor{\binits{A.}~\bsnm{Lonie}},
-\bauthor{\binits{S.}~\bsnm{Kosakovsky Pond}},
-\bauthor{\binits{W.}~\bsnm{Maier}},
-\bauthor{\binits{A.}~\bsnm{Nekrutenko}},
-\bauthor{\binits{J.}~\bsnm{Taylor}} and
-\bauthor{\binits{S.}~\bsnm{Weaver}},
-\batitle{No more business as usual: Agile and effective responses to emerging
-pathogen threats require open data and open analytics},
-\bjtitle{PLOS Pathogens}
-\bvolume{16}(\bissue{8})
-(\byear{2020}),
-\bnumber{e1008643}.
-\bid{doi={10.1371/journal.ppat.1008643}}
-\end{barticle}
-%
-\OrigBibText
-D.~Baker,
-M.~van~den Beek,
-D.~Blankenberg,
-D.~Bouvier,
-J.~Chilton,
-N.~Coraor,
-F.~Coppens,
-I.~Eguinoa,
-S.~Gladman,
-B.~Gr\"{u}ning,
-N.~Keener,
-D.~Larivi\`{e}re,
-A.~Lonie,
-S.~Kosakovsky~Pond,
-W.~Maier,
-A.~Nekrutenko,
-J.~Taylor and
-S.~Weaver,
-\textit{No more business as usual: Agile and effective responses to emerging
-pathogen threats require open data and open analytics.},
-{PLOS} Pathogens
-16(8)
-(2020),
-e1008643.
-doi:10.1371/journal.ppat.1008643.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b11 ###bbsrt2
-%b89 ###
-\bibitem{doi:10.5334/dsj-2019-044}
-%
-\begin{barticle}
-\bauthor{\binits{M.}~\bsnm{Barker}},
-\bauthor{\binits{R.}~\bsnm{Wilkinson}} and
-\bauthor{\binits{A.}~\bsnm{Treloar}},
-\batitle{The Australian Research Data Commons},
-\bjtitle{Data Science Journal}
-\bvolume{18}
-(\byear{2019}).
-\bid{doi={10.5334/dsj-2019-044}}
-\end{barticle}
-%
-\OrigBibText
-M.~Barker,
-R.~Wilkinson and
-A.~Treloar,
-The Australian Research Data Commons,
-\textit{Data Science Journal}
-\textbf{18}
-(2019), 00.
-doi:10.5334/dsj-2019-044.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b12 ###bbsrt2
-%b11 ###
-\bibitem{doi:10.1016/j.future.2011.08.004}
-%
-\begin{barticle}
-\bauthor{\binits{S.}~\bsnm{Bechhofer}},
-\bauthor{\binits{I.}~\bsnm{Buchan}},
-\bauthor{\binits{D.}~\bsnm{De Roure}},
-\bauthor{\binits{P.}~\bsnm{Missier}},
-\bauthor{\binits{J.}~\bsnm{Ainsworth}},
-\bauthor{\binits{J.}~\bsnm{Bhagat}},
-\bauthor{\binits{P.}~\bsnm{Couch}},
-\bauthor{\binits{D.}~\bsnm{Cruickshank}},
-\bauthor{\binits{M.}~\bsnm{Delderfield}},
-\bauthor{\binits{I.}~\bsnm{Dunlop}},
-\bauthor{\binits{M.}~\bsnm{Gamble}},
-\bauthor{\binits{D.}~\bsnm{Michaelides}},
-\bauthor{\binits{S.}~\bsnm{Owen}},
-\bauthor{\binits{D.}~\bsnm{Newman}},
-\bauthor{\binits{S.}~\bsnm{Sufi}} and
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\batitle{Why linked data is not enough for scientists},
-\bjtitle{Future Generation Computer Systems}
-\bvolume{29}(\bissue{2})
-(\byear{2013}),
-\bfpage{599}--\blpage{611}.
-\bid{doi={10.1016/j.future.2011.08.004}}
-\end{barticle}
-%
-\OrigBibText
-S.~Bechhofer,
-I.~Buchan,
-D.~De~Roure,
-P.~Missier,
-J.~Ainsworth,
-J.~Bhagat,
-P.~Couch,
-D.~Cruickshank,
-M.~Delderfield,
-I.~Dunlop,
-M.~Gamble,
-D.~Michaelides,
-S.~Owen,
-D.~Newman,
-S.~Sufi and
-C.~Goble,
-\textit{Why linked data is not enough for scientists},
-Future Generation Computer Systems
-29(2)
-(2013),
-599--611.
-doi:10.1016/j.future.2011.08.004.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b13 ###bbsrt2
-%b59 ###
-\bibitem{CheckMyCrate}
-%
-\begin{botherref}
-\oauthor{\binits{K.}~\bsnm{Belchev}},
-{KockataEPich}/{CheckMyCrate}: A command line application for
-validating a {RO}-Crate object against a {JSON} profile,
-\textit{GitHub}
-(2021).
-\url{https://github.com/KockataEPich/CheckMyCrate}.
-\end{botherref}
-%
-\OrigBibText
-K.~Belchev,
-{KockataEPich}/{CheckMyCrate}: A command line application for
-validating a
-{RO}-Crate object against a {JSON} profile.,
-GitHub,
-2021.
-\url{https://github.com/KockataEPich/CheckMyCrate}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b14 ###bbsrt2
-%b12 ###
-\bibitem{doi:10.1016/j.websem.2015.01.003}
-%
-\begin{barticle}
-\bauthor{\binits{K.}~\bsnm{Belhajjame}},
-\bauthor{\binits{J.}~\bsnm{Zhao}},
-\bauthor{\binits{D.}~\bsnm{Garijo}},
-\bauthor{\binits{M.}~\bsnm{Gamble}},
-\bauthor{\binits{K.}~\bsnm{Hettne}},
-\bauthor{\binits{R.}~\bsnm{Palma}},
-\bauthor{\binits{E.}~\bsnm{Mina}},
-\bauthor{\binits{O.}~\bsnm{Corcho}},
-\bauthor{\binits{J.M.}~\bsnm{G\'{o}mez-P\'{e}rez}},
-\bauthor{\binits{S.}~\bsnm{Bechhofer}},
-\bauthor{\binits{G.}~\bsnm{Klyne}} and
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\batitle{Using a suite of ontologies for preserving workflow-centric
-research objects},
-\bjtitle{Web Semantics: Science, Services and Agents on the World Wide Web}
-\bvolume{32}
-(\byear{2015}),
-\bfpage{16}--\blpage{42}.
-\bid{doi={10.1016/j.websem.2015.01.003}}
-\end{barticle}
-%
-\OrigBibText
-K.~Belhajjame,
-J.~Zhao,
-D.~Garijo,
-M.~Gamble,
-K.~Hettne,
-R.~Palma,
-E.~Mina,
-O.~Corcho,
-J.M.~G\'{o}mez-P\'{e}rez,
-S.~Bechhofer,
-G.~Klyne and
-C.~Goble,
-\textit{Using a suite of ontologies for preserving workflow-centric research
-objects},
-Web Semantics: Science, Services and Agents on the World Wide Web
-32
-(2015),
-16--42.
-doi:10.1016/j.websem.2015.01.003.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b15 ###bbsrt2
-%b116 ###
-\bibitem{doi:10.3389/fninf.2017.00069}
-%
-\begin{barticle}
-\bauthor{\binits{F.C.Y.}~\bsnm{Benureau}} and
-\bauthor{\binits{N.P.}~\bsnm{Rougier}},
-\batitle{Re-run, repeat, reproduce, reuse, replicate: Transforming code
-into scientific contributions},
-\bjtitle{Frontiers in Neuroinformatics}
-\bvolume{11}
-(\byear{2017}),
-\bnumber{69}.
-\bid{doi={10.3389/fninf.2017.00069}}
-\end{barticle}
-%
-\OrigBibText
-F.C.Y.~Benureau and
-N.P.~Rougier,
-\textit{Re-run, Repeat, Reproduce, Reuse, Replicate: Transforming Code into
-Scientific Contributions.},
-Frontiers in Neuroinformatics
-11
-(2017),
-69.
-doi:10.3389/fninf.2017.00069.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b16 ###bbsrt2
-%b8 ###
-\bibitem{doi:10.1093/nar/gkl971}
-%
-\begin{barticle}
-\bauthor{\binits{H.}~\bsnm{Berman}},
-\bauthor{\binits{K.}~\bsnm{Henrick}},
-\bauthor{\binits{H.}~\bsnm{Nakamura}} and
-\bauthor{\binits{J.L.}~\bsnm{Markley}},
-\batitle{The worldwide Protein Data Bank ({wwPDB}): Ensuring a single,
-uniform archive of {PDB} data},
-\bjtitle{Nucleic Acids Research}
-\bvolume{35}\bcomment{(Database issue)}
-(\byear{2007}),
-\bfpage{D301}--\blpage{D303}.
-\bid{doi={10.1093/nar/gkl971}}
-\end{barticle}
-%
-\OrigBibText
-H.~Berman,
-K.~Henrick,
-H.~Nakamura and
-J.L.~Markley,
-\textit{The worldwide Protein Data Bank ({wwPDB}): ensuring a single, uniform
-archive of {PDB} data}.
-Nucleic Acids Research
-35 (Database issue)
-(2007),
-D301--3.
-doi:10.1093/nar/gkl971.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b17 ###bbsrt2
-%b68 ###
-\bibitem{doi:10.5281/zenodo.4705078}
-%
-\begin{barticle}
-\bauthor{\binits{F.}~\bsnm{Bietrix}},
-\bauthor{\binits{J.M.}~\bsnm{Carazo}},
-\bauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
-\bauthor{\binits{F.}~\bsnm{Coppens}},
-\bauthor{\binits{M.L.}~\bsnm{Chiusano}},
-\bauthor{\binits{R.}~\bsnm{David}},
-\bauthor{\binits{J.M.}~\bsnm{Fernandez}},
-\bauthor{\binits{M.}~\bsnm{Fratelli}},
-\bauthor{\binits{J.-K.}~\bsnm{Heriche}},
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{P.}~\bsnm{Gribbon}},
-\bauthor{\binits{P.}~\bsnm{Holub}},
-\bauthor{\binits{R.P.}~\bsnm{Joosten}},
-\bauthor{\binits{S.}~\bsnm{Leo}},
-\bauthor{\binits{S.}~\bsnm{Owen}},
-\bauthor{\binits{H.}~\bsnm{Parkinson}},
-\bauthor{\binits{R.}~\bsnm{Pieruschka}},
-\bauthor{\binits{L.}~\bsnm{Pireddu}},
-\bauthor{\binits{L.}~\bsnm{Porcu}},
-\bauthor{\binits{M.}~\bsnm{Raess}},
-\bauthor{\binits{L.}~\bsnm{Rodriguez- Navas}},
-\bauthor{\binits{A.}~\bsnm{Scherer}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}} and
-\bauthor{\binits{J.}~\bsnm{Tang}},
-\batitle{{EOSC}-life methodology framework to enhance reproducibility
-within {EOSC}-life},
-\bjtitle{Zenodo}
-(\byear{2021}).
-\bid{doi={10.5281/zenodo.4705078}}
-\end{barticle}
-%
-\OrigBibText
-F.~Bietrix,
-J.M.~Carazo,
-S.~Capella-Gutierrez,
-F.~Coppens,
-M.L.~Chiusano,
-R.~David,
-J.M.~Fernandez,
-M.~Fratelli,
-J.-K.~Heriche,
-C.~Goble,
-P.~Gribbon,
-P.~Holub,
-R.~P.~Joosten,
-S.~Leo,
-S.~Owen,
-H.~Parkinson,
-R.~Pieruschka,
-L.~Pireddu,
-L.~Porcu,
-M.~Raess,
-L.~Rodriguez-~Navas,
-A.~Scherer,
-S.~Soiland-Reyes and
-J.~Tang,
-\textit{{EOSC}-Life Methodology framework to enhance reproducibility within
-{EOSC}-Life},
-Zenodo
-(2021).
-doi:10.5281/zenodo.4705078.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b18 ###bbsrt2
-%b29 ###
-\bibitem{doi:10.4018/978-1-60960-593-3.ch008}
-%
-\begin{bchapter}
-\bauthor{\binits{C.}~\bsnm{Bizer}},
-\bauthor{\binits{T.}~\bsnm{Heath}} and
-\bauthor{\binits{T.}~\bsnm{Berners-Lee}},
-\bctitle{Linked data: The story so far},
-in: \bbtitle{Semantic Services, Interoperability and Web Applications:
-Emerging Concepts},
-\beditor{\binits{A.}~\bsnm{Sheth}}, ed.,
-\bsertitle{{IGI} Global},
-\byear{2011},
-pp.~\bfpage{205}--\blpage{227}.
-ISBN \bisbn{9781609605933}.
-\bid{doi={10.4018/978-1-60960-593-3.ch008}}
-\end{bchapter}
-%
-\OrigBibText
-C.~Bizer,
-T.~Heath and
-T.~Berners-Lee,
-\textit{Linked data: the story so far},
-in: Semantic services, interoperability and web applications: emerging
-concepts,
-A.~Sheth, ed.,
-{IGI} Global,
-2011,
-pp.~205--227.
-ISBN 9781609605933.
-doi:10.4018/978-1-60960-593-3.ch008.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b19 ###bbsrt2
-%b126 ###
-\bibitem{doi:10.1087/20150211}
-%
-\begin{barticle}
-\bauthor{\binits{A.}~\bsnm{Brand}},
-\bauthor{\binits{L.}~\bsnm{Allen}},
-\bauthor{\binits{M.}~\bsnm{Altman}},
-\bauthor{\binits{M.}~\bsnm{Hlava}} and
-\bauthor{\binits{J.}~\bsnm{Scott}},
-\batitle{Beyond authorship: Attribution, contribution, collaboration,
-and credit},
-\bjtitle{Learned Publishing}
-\bvolume{28}(\bissue{2})
-(\byear{2015}),
-\bfpage{151}--\blpage{155}.
-\bid{doi={10.1087/20150211}}
-\end{barticle}
-%
-\OrigBibText
-A.~Brand,
-L.~Allen,
-M.~Altman,
-M.~Hlava and
-J.~Scott,
-\textit{Beyond authorship: attribution, contribution, collaboration, and
-credit},
-Learned Publishing
-28(2)
-(2015),
-151--155.
-doi:10.1087/20150211.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b20 ###bbsrt2
-%b58 ###
-\bibitem{doi:10.5281/zenodo.3903463}
-%
-\begin{botherref}
-\oauthor{\binits{G.}~\bsnm{Brenner}},
-BrennerG/Ro-Crate\_2\_ma-DMP: v1.0.0,
-2020.
-\url{https://github.com/BrennerG/Ro-Crate\_2\_ma-DMP}.
-\bid{doi={10.5281/zenodo.3903463}}
-\end{botherref}
-%
-\OrigBibText
-G.~Brenner,
-BrennerG/Ro-Crate\_2\_ma-DMP: v1.0.0,
-2020.
-doi:10.5281/zenodo.3903463.
-\url{https://github.com/BrennerG/Ro-Crate_2_ma-DMP}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b21 ###bbsrt2
-%b86 ###
-\bibitem{doi:10.4126/frl01-006423289}
-%
-\begin{bbook}
-\bauthor{\binits{J.}~\bsnm{Cardoso}},
-\bauthor{\binits{L.J.}~\bsnm{Garcia Castro}},
-\bauthor{\binits{F.}~\bsnm{Ekaputra}},
-\bauthor{\binits{M.-C.}~\bsnm{Jacquemot-Perbal}},
-\bauthor{\binits{T.}~\bsnm{Miksa}} and
-\bauthor{\binits{J.}~\bsnm{Borbinha}},
-\bbtitle{Towards Semantic Representation of Machine-Actionable Data
-Management Plans, PUBLISSO},
-\byear{2020}.
-\url{https://repository.publisso.de/resource/frl:6423289}.
-\bid{doi={10.4126/frl01-006423289}}
-\end{bbook}
-%
-\OrigBibText
-J.~Cardoso,
-L.J.~Garcia~Castro,
-F.~Ekaputra,
-M.-C.~Jacquemot-Perbal,
-T.~Miksa and
-J.~Borbinha,
-\textit{Towards semantic representation of machine-actionable Data Management
-Plans},
-PUBLISSO
-(2020).
-doi:10.4126/frl01-006423289.
-\url{https://repository.publisso.de/resource/frl:6423289}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b22 ###bbsrt2
-%b84 ###
-\bibitem{doi:10.1007/978-3-030-45442-5_15}
-%
-\begin{bchapter}
-\bauthor{\binits{J.}~\bsnm{Cardoso}},
-\bauthor{\binits{D.}~\bsnm{Proen\c{c}a}} and
-\bauthor{\binits{J.}~\bsnm{Borbinha}},
-\bctitle{Machine-actionable data management plans: A knowledge retrieval
-approach to automate the assessment of funders' requirements},
-in: \bbtitle{Advances in Information Retrieval},
-\beditor{\binits{J.M.}~\bsnm{Jose}},
-\beditor{\binits{E.}~\bsnm{Yilmaz}},
-\beditor{\binits{J.}~\bsnm{Magalh\~{a}es}},
-\beditor{\binits{P.}~\bsnm{Castells}},
-\beditor{\binits{N.}~\bsnm{Ferro}},
-\beditor{\binits{M.J.}~\bsnm{Silva}} and
-\beditor{\binits{F.}~\bsnm{Martins}}, eds,
-\bpublisher{Springer International Publishing},
-\blocation{Cham},
-\byear{2020},
-pp.~\bfpage{118}--\blpage{125}.
-ISBN \bisbn{978-3-030-45442-5}.
-\bid{doi={10.1007/978-3-030-45442-5_15}}
-\end{bchapter}
-%
-\OrigBibText
-J.~Cardoso,
-D.~Proen{\c{c}}a and
-J.~Borbinha,
-\textit{Machine-Actionable Data Management Plans: A Knowledge Retrieval
-Approach to Automate the Assessment of Funders' Requirements},
-in: Advances in Information Retrieval,
-J.M.~Jose,
-E.~Yilmaz,
-J.~Magalh{\~a}es,
-P.~Castells,
-N.~Ferro,
-M.J.~Silva and
-F.~Martins, eds,
-Springer International Publishing,
-Cham,
-2020,
-pp.~118--125.
-ISBN 978-3-030-45442-5.
-doi:10.1007/978-3-030-45442-5\_15.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b23 ###bbsrt2
-%b13 ###
-\bibitem{doi:10.5281/zenodo.3250687}
-%
-\begin{barticle}
-\bauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{P.}~\bsnm{Sefton}} and
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\batitle{A lightweight approach to research object data packaging},
-\bjtitle{Zenodo}
-(\byear{2019}).
-\bid{doi={10.5281/zenodo.3250687}}
-\end{barticle}
-%
-\OrigBibText
-E.{\'{O}}.~Carrag\'{a}in,
-C.~Goble,
-P.~Sefton and
-S.~Soiland-Reyes,
-\textit{A lightweight approach to research object data packaging},
-Zenodo
-(2019).
-doi:10.5281/zenodo.3250687.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b24 ###bbsrt2
-%b17 ###
-\bibitem{chan_1995}
-%
-\begin{bbook}
-\bauthor{\binits{L.M.}~\bsnm{Chan}},
-\bbtitle{Library of Congress Subject Headings: Principles and Application},
-\bedition{3}rd edn,
-\bpublisher{Libraries Unlimited},
-\blocation{Englewood, Colo},
-\byear{1995},
-p.~\bfpage{556}.
-\url{https://eric.ed.gov/?id=ED387146}.
-ISBN \bisbn{9781563081910}.
-\end{bbook}
-%
-\OrigBibText
-L.M.~Chan,
-\textit{Library of Congress Subject Headings: Principles and Application},
-3rd edn,
-Libraries Unlimited,
-Englewood, Colo,
-1995,
-p.~556.
-ISBN 9781563081910.
-\url{https://eric.ed.gov/?id=ED387146}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b25 ###bbsrt2
-%b81 ###
-\bibitem{doi:10.1109/BigData.2016.7840618}
-%
-\begin{bchapter}
-\bauthor{\binits{K.}~\bsnm{Chard}},
-\bauthor{\binits{M.}~\bsnm{D'Arcy}},
-\bauthor{\binits{B.}~\bsnm{Heavner}},
-\bauthor{\binits{I.}~\bsnm{Foster}},
-\bauthor{\binits{C.}~\bsnm{Kesselman}},
-\bauthor{\binits{R.}~\bsnm{Madduri}},
-\bauthor{\binits{A.}~\bsnm{Rodriguez}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{K.}~\bsnm{Clark}},
-\bauthor{\binits{E.W.}~\bsnm{Deutsch}},
-\bauthor{\binits{I.}~\bsnm{Dinov}},
-\bauthor{\binits{N.}~\bsnm{Price}} and
-\bauthor{\binits{A.}~\bsnm{Toga}},
-\bctitle{I'll take that to go: Big data bags and minimal identifiers for
-exchange of large, complex datasets},
-in: \bbtitle{2016 {IEEE} International Conference on Big Data (Big Data)},
-\bpublisher{IEEE},
-\byear{2016},
-pp.~\bfpage{319}--\blpage{328}.
-\url{https://static.aminer.org/pdf/fa/bigdata2016/BigD418.pdf}.
-ISBN \bisbn{978-1-4673-9005-7}.
-\bid{doi={10.1109/BigData.2016.7840618}}
-\end{bchapter}
-%
-\OrigBibText
-K.~Chard,
-M.~D'Arcy,
-B.~Heavner,
-I.~Foster,
-C.~Kesselman,
-R.~Madduri,
-A.~Rodriguez,
-S.~Soiland-Reyes,
-C.~Goble,
-K.~Clark,
-E.W.~Deutsch,
-I.~Dinov,
-N.~Price and
-A.~Toga,
-\textit{I'll take that to go: Big data bags and minimal identifiers for
-exchange of large, complex datasets},
-in: 2016 {IEEE} International Conference on Big Data (Big Data),
-IEEE,
-2016,
-pp.~319--328.
-ISBN 978-1-4673-9005-7.
-doi:10.1109/BigData.2016.7840618.
-\url{https://static.aminer.org/pdf/fa/bigdata2016/BigD418.pdf}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b26 ###bbsrt2
-%b106 ###
-\bibitem{doi:10.1109/eScience.2019.00068}
-%
-\begin{bchapter}
-\bauthor{\binits{K.}~\bsnm{Chard}},
-\bauthor{\binits{N.}~\bsnm{Gaffney}},
-\bauthor{\binits{M.B.}~\bsnm{Jones}},
-\bauthor{\binits{K.}~\bsnm{Kowalik}},
-\bauthor{\binits{B.}~\bsnm{Ludascher}},
-\bauthor{\binits{T.}~\bsnm{McPhillips}},
-\bauthor{\binits{J.}~\bsnm{Nabrzyski}},
-\bauthor{\binits{V.}~\bsnm{Stodden}},
-\bauthor{\binits{I.}~\bsnm{Taylor}},
-\bauthor{\binits{T.}~\bsnm{Thelen}},
-\bauthor{\binits{M.J.}~\bsnm{Turk}} and
-\bauthor{\binits{C.}~\bsnm{Willis}},
-\bctitle{Application of {BagIt}-serialized research object bundles for
-packaging and re-execution of computational analyses},
-in: \bbtitle{15th International Conference on {eScience} ({eScience} 2019)},
-\bpublisher{IEEE},
-\byear{2019},
-pp.~\bfpage{514}--\blpage{521}.
-\url{https://zenodo.org/record/3381754}.
-ISBN \bisbn{978-1-7281-2451-3}.
-\bid{doi={10.1109/eScience.2019.00068}}
-\end{bchapter}
-%
-\OrigBibText
-K.~Chard,
-N.~Gaffney,
-M.B.~Jones,
-K.~Kowalik,
-B.~Ludascher,
-T.~{McPhillips},
-J.~Nabrzyski,
-V.~Stodden,
-I.~Taylor,
-T.~Thelen,
-M.J.~Turk and
-C.~Willis,
-\textit{Application of {BagIt}-Serialized Research Object Bundles for
-Packaging and Re-Execution of Computational Analyses},
-in: 15th International Conference on {eScience} ({eScience} 2019),
-IEEE,
-2019,
-pp.~514--521.
-ISBN 978-1-7281-2451-3.
-doi:10.1109/eScience.2019.00068.
-\url{https://zenodo.org/record/3381754}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b27 ###bbsrt2
-%b93 ###
-\bibitem{doi:10.1109/MCC.2014.52}
-%
-\begin{barticle}
-\bauthor{\binits{K.}~\bsnm{Chard}},
-\bauthor{\binits{S.}~\bsnm{Tuecke}} and
-\bauthor{\binits{I.}~\bsnm{Foster}},
-\batitle{Efficient and secure transfer, synchronization, and sharing of
-big data},
-\bjtitle{IEEE Cloud Computing}
-\bvolume{1}(\bissue{3})
-(\byear{2014}),
-\bfpage{46}--\blpage{55}.
-\bid{doi={10.1109/MCC.2014.52}}
-\end{barticle}
-%
-\OrigBibText
-K.~Chard,
-S.~Tuecke and
-I.~Foster,
-\textit{Efficient and Secure Transfer, Synchronization, and Sharing of Big
-Data},
-IEEE Cloud Computing
-1(3)
-(2014),
-46--55.
-doi:10.1109/MCC.2014.52.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b28 ###bbsrt2
-%b101 ###
-\bibitem{Ciccarese:17:WAD}
-%
-\begin{botherref}
-\oauthor{\binits{P.}~\bsnm{Ciccarese}},
-\oauthor{\binits{R.}~\bsnm{Sanderson}} and
-\oauthor{\binits{B.}~\bsnm{Young}},
-Web Annotation Data Model, {W3C} Recommendation,
-W3C,
-2017.
-\url{https://www.w3.org/TR/2017/REC-annotation-model-20170223/}.
-\end{botherref}
-%
-\OrigBibText
-P.~Ciccarese,
-R.~Sanderson and
-B.~Young,
-Web Annotation Data Model,
-{W3C} Recommendation,
-W3C,
-2017,
-https://www.w3.org/TR/2017/REC-annotation-model-20170223/.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b29 ###bbsrt2
-%b96 ###
-\bibitem{doi:10.1190/1.1822162}
-%
-\begin{bchapter}
-\bauthor{\binits{J.F.}~\bsnm{Claerbout}} and
-\bauthor{\binits{M.}~\bsnm{Karrenbach}},
-\bctitle{Electronic documents give reproducible research a new meaning},
-in: \bbtitle{{SEG} Technical Program Expanded Abstracts 1992, Society
-of Exploration Geophysicists},
-\byear{1992},
-pp.~\bfpage{601}--\blpage{604}.
-\bid{doi={10.1190/1.1822162}}
-\end{bchapter}
-%
-\OrigBibText
-J.F.~Claerbout and
-M.~Karrenbach,
-\textit{Electronic documents give reproducible research a new meaning},
-in: {SEG} Technical Program Expanded Abstracts 1992,
-Society of Exploration Geophysicists,
-1992,
-pp.~601--604.
-doi:10.1190/1.1822162.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b30 ###bbsrt2
-%b108 ###
-\bibitem{doi:10.1016/j.future.2017.01.012}
-%
-\begin{barticle}
-\bauthor{\binits{S.}~\bsnm{Cohen-Boulakia}},
-\bauthor{\binits{K.}~\bsnm{Belhajjame}},
-\bauthor{\binits{O.}~\bsnm{Collin}},
-\bauthor{\binits{J.}~\bsnm{Chopard}},
-\bauthor{\binits{C.}~\bsnm{Froidevaux}},
-\bauthor{\binits{A.}~\bsnm{Gaignard}},
-\bauthor{\binits{K.}~\bsnm{Hinsen}},
-\bauthor{\binits{P.}~\bsnm{Larmande}},
-\bauthor{\binits{Y.L.}~\bsnm{Bras}},
-\bauthor{\binits{F.}~\bsnm{Lemoine}},
-\bauthor{\binits{F.}~\bsnm{Mareuil}},
-\bauthor{\binits{H.}~\bsnm{M\'{e}nager}},
-\bauthor{\binits{C.}~\bsnm{Pradal}} and
-\bauthor{\binits{C.}~\bsnm{Blanchet}},
-\batitle{Scientific workflows for computational reproducibility in the
-life sciences: Status, challenges and opportunities},
-\bjtitle{Future Generation Computer Systems}
-\bvolume{75}
-(\byear{2017}),
-\bfpage{284}--\blpage{298}.
-\bid{doi={10.1016/j.future.2017.01.012}}
-\end{barticle}
-%
-\OrigBibText
-S.~Cohen-Boulakia,
-K.~Belhajjame,
-O.~Collin,
-J.~Chopard,
-C.~Froidevaux,
-A.~Gaignard,
-K.~Hinsen,
-P.~Larmande,
-Y.L.~Bras,
-F.~Lemoine,
-F.~Mareuil,
-H.~M\'{e}nager,
-C.~Pradal and
-C.~Blanchet,
-\textit{Scientific workflows for computational reproducibility in the life
-sciences: Status, challenges and opportunities},
-Future Generation Computer Systems
-75
-(2017),
-284--298.
-doi:10.1016/j.future.2017.01.012.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b31 ###bbsrt2
-%b36 ###
-\bibitem{pcdm}
-%
-\begin{botherref}
-\oauthor{\binits{S.}~\bsnm{Cossu}},
-\oauthor{\binits{E.}~\bsnm{Cowles}},
-\oauthor{\binits{K.}~\bsnm{Estlund}},
-\oauthor{\binits{C.}~\bsnm{Harlow}},
-\oauthor{\binits{T.}~\bsnm{Johnson}},
-\oauthor{\binits{M.}~\bsnm{Matienzo}},
-\oauthor{\binits{D.}~\bsnm{Lamb}},
-\oauthor{\binits{L.}~\bsnm{Rayle}},
-\oauthor{\binits{R.}~\bsnm{Sanderson}},
-\oauthor{\binits{J.}~\bsnm{Stroop}} and
-\oauthor{\binits{A.}~\bsnm{Woods}},
-Portland Common Data Model,
-2018.
-\url{https://github.com/duraspace/pcdm/wiki}.
-\end{botherref}
-%
-\OrigBibText
-S.~Cossu,
-E.~Cowles,
-K.~Estlund,
-C.~Harlow,
-T.~Johnson,
-M.~Matienzo,
-D.~Lamb,
-L.~Rayle,
-R.~Sanderson,
-J.~Stroop and
-A.~Woods,
-Portland Common Data Model,
-2018.
-\url{https://github.com/duraspace/pcdm/wiki}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b32 ###bbsrt2
-%b92 ###
-\bibitem{doi:10.1045/january2011-crosas}
-%
-\begin{barticle}
-\bauthor{\binits{M.}~\bsnm{Crosas}},
-\batitle{The DataVerse Network: An open-source application for sharing,
-discovering and preserving data},
-\bjtitle{D-Lib Magazine}
-\bvolume{17}(\bissue{1/2})
-(\byear{2011}).
-\bid{doi={10.1045/january2011-crosas}}
-\end{barticle}
-%
-\OrigBibText
-M.~Crosas,
-The Dataverse Network: An Open-Source Application for Sharing,
-Discovering and
-Preserving Data,
-\textit{D-Lib Magazine}
-\textbf{17}(1/2)
-(2011)
-00.
-doi:10.1045/january2011-crosas.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b33 ###bbsrt2
-%b91 ###
-\bibitem{doi:10.7557/5.5422}
-%
-\begin{botherref}
-\oauthor{\binits{M.}~\bsnm{Crosas}},
-Harvard Data Commons,
-European Dataverse Workshop 2020,
-Troms{\o}, Norway,
-2020. ISSN 2387-3086.
-\bid{doi={10.7557/5.5422}}
-\end{botherref}
-%
-\OrigBibText
-M.~Crosas,
-Harvard Data Commons,
-2020,
-European Dataverse Workshop 2020, Troms{\o}, Norway. 2020-01-23/--24.
-ISSN 2387-3086.
-doi:10.7557/5.5422.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b34 ###bbsrt2
-%b40 ###
-\bibitem{doi:10.1016/j.tibtech.2012.02.002}
-%
-\begin{barticle}
-\bauthor{\binits{L.C.}~\bsnm{Crosswell}} and
-\bauthor{\binits{J.M.}~\bsnm{Thornton}},
-\batitle{{ELIXIR}: A distributed infrastructure for European biological data},
-\bjtitle{Trends in Biotechnology}
-\bvolume{30}(\bissue{5})
-(\byear{2012}),
-\bfpage{241}--\blpage{242}.
-\bid{doi={10.1016/j.tibtech.2012.02.002}}
-\end{barticle}
-%
-\OrigBibText
-L.C.~Crosswell and
-J.M.~Thornton,
-\textit{{ELIXIR}: a distributed infrastructure for European biological data},
-Trends in Biotechnology
-30(5)
-(2012),
-241--242.
-doi:10.1016/j.tibtech.2012.02.002.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b35 ###bbsrt2
-%b49 ###
-\bibitem{about-lifemonitor}
-%
-\begin{botherref}
-{CRS4},
-{LifeMonitor}, a testing and monitoring service for scientific workflows.
-\url{https://about.lifemonitor.eu/}.
-\end{botherref}
-%
-\OrigBibText
-CRS4,
-{LifeMonitor}, a testing and monitoring service for scientific workflows.
-\url{https://about.lifemonitor.eu/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b36 ###bbsrt2
-%b69 ###
-\bibitem{doi:10.1145/3486897}
-%
-\begin{bbook}
-\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
-\bauthor{\binits{S.}~\bsnm{Abeln}},
-\bauthor{\binits{A.}~\bsnm{Iosup}},
-\bauthor{\binits{P.}~\bsnm{Amstutz}},
-\bauthor{\binits{J.}~\bsnm{Chilton}},
-\bauthor{\binits{N.}~\bsnm{Tijani\'{c}}},
-\bauthor{\binits{H.}~\bsnm{M\'{e}nager}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}} and
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bbtitle{Methods Included: Standardizing Computational Reuse and Portability
-with the Common Workflow Language},
-\bpublisher{Communications of the ACM},
-\byear{2022}.
-\bcomment{Accepted}.
-\bid{doi={10.1145/3486897}}
-\end{bbook}
-%
-\OrigBibText
-M.R.~Crusoe,
-S.~Abeln,
-A.~Iosup,
-P.~Amstutz,
-J.~Chilton,
-N.~Tijani\'{c},
-H.~M\'{e}nager,
-S.~Soiland-Reyes and
-C.~Goble,
-\textit{Methods Included: Standardizing Computational Reuse and Portability
-with the Common Workflow Language},
-Communications of the ACM
-(2022),
-Accepted.
-doi:10.1145/3486897.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b37 ###bbsrt2
-%b120 ###
-\bibitem{doi:10.1093/bioinformatics/btx192}
-%
-\begin{barticle}
-\bauthor{\binits{F.}~\bsnm{da~Veiga Leprevost}},
-\bauthor{\binits{B.A.}~\bsnm{Gr\"{u}ning}},
-\bauthor{\binits{S.}~\bsnm{Alves Aflitos}},
-\bauthor{\binits{H.L.}~\bsnm{R\"{o}st}},
-\bauthor{\binits{J.}~\bsnm{Uszkoreit}},
-\bauthor{\binits{H.}~\bsnm{Barsnes}},
-\bauthor{\binits{M.}~\bsnm{Vaudel}},
-\bauthor{\binits{P.}~\bsnm{Moreno}},
-\bauthor{\binits{L.}~\bsnm{Gatto}},
-\bauthor{\binits{J.}~\bsnm{Weber}},
-\bauthor{\binits{M.}~\bsnm{Bai}},
-\bauthor{\binits{R.C.}~\bsnm{Jimenez}},
-\bauthor{\binits{T.}~\bsnm{Sachsenberg}},
-\bauthor{\binits{J.}~\bsnm{Pfeuffer}},
-\bauthor{\binits{R.}~\bsnm{Vera Alvarez}},
-\bauthor{\binits{J.}~\bsnm{Griss}},
-\bauthor{\binits{A.I.}~\bsnm{Nesvizhskii}} and
-\bauthor{\binits{Y.}~\bsnm{Perez-Riverol}},
-\batitle{{BioContainers}: An open-source and community-driven framework
-for software standardization},
-\bjtitle{Bioinformatics}
-\bvolume{33}(\bissue{16})
-(\byear{2017}),
-\bfpage{2580}--\blpage{2582}.
-\bid{doi={10.1093/bioinformatics/btx192}}
-\end{barticle}
-%
-\OrigBibText
-F.~da~Veiga~Leprevost,
-B.A.~Gr\"{u}ning,
-S.~Alves~Aflitos,
-H.L.~R\"{o}st,
-J.~Uszkoreit,
-H.~Barsnes,
-M.~Vaudel,
-P.~Moreno,
-L.~Gatto,
-J.~Weber,
-M.~Bai,
-R.C.~Jimenez,
-T.~Sachsenberg,
-J.~Pfeuffer,
-R.~Vera~Alvarez,
-J.~Griss,
-A.I.~Nesvizhskii and
-Y.~Perez-Riverol,
-\textit{{BioContainers}: an open-source and community-driven framework for
-software standardization.},
-Bioinformatics
-33(16)
-(2017),
-2580--2582.
-doi:10.1093/bioinformatics/btx192.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b38 ###bbsrt2
-%b73 ###
-\bibitem{doi:10.3390/publications8020021}
-%
-\begin{barticle}
-\bauthor{\binits{K.}~\bsnm{De Smedt}},
-\bauthor{\binits{D.}~\bsnm{Koureas}} and
-\bauthor{\binits{P.}~\bsnm{Wittenburg}},
-\batitle{{FAIR} digital objects for science: From data pieces to
-actionable knowledge units},
-\bjtitle{Publications}
-\bvolume{8}(\bissue{2})
-(\byear{2020}),
-\bnumber{21}.
-\bid{doi={10.3390/publications8020021}}
-\end{barticle}
-%
-\OrigBibText
-K.~De~Smedt,
-D.~Koureas and
-P.~Wittenburg,
-\textit{{FAIR} digital objects for science: from data pieces to actionable
-knowledge units},
-Publications
-8(2)
-(2020),
-21.
-doi:10.3390/publications8020021.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b39 ###bbsrt2
-%b66 ###
-\bibitem{doi:10.1038/nbt.3820}
-%
-\begin{barticle}
-\bauthor{\binits{P.}~\bsnm{Di Tommaso}},
-\bauthor{\binits{M.}~\bsnm{Chatzou}},
-\bauthor{\binits{E.W.}~\bsnm{Floden}},
-\bauthor{\binits{P.P.}~\bsnm{Barja}},
-\bauthor{\binits{E.}~\bsnm{Palumbo}} and
-\bauthor{\binits{C.}~\bsnm{Notredame}},
-\batitle{Nextflow enables reproducible computational workflows},
-\bjtitle{Nature Biotechnology}
-\bvolume{35}(\bissue{4})
-(\byear{2017}),
-\bfpage{316}--\blpage{319}.
-\bid{doi={10.1038/nbt.3820}}
-\end{barticle}
-%
-\OrigBibText
-P.~Di~Tommaso,
-M.~Chatzou,
-E.W.~Floden,
-P.P.~Barja,
-E.~Palumbo and
-C.~Notredame,
-\textit{Nextflow enables reproducible computational workflows.},
-Nature Biotechnology
-35(4)
-(2017),
-316--319.
-doi:10.1038/nbt.3820.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b40 ###bbsrt2
-%b7 ###
-\bibitem{doi:10.3897/biss.3.37080}
-%
-\begin{barticle}
-\bauthor{\binits{M.}~\bsnm{Dillen}},
-\bauthor{\binits{Q.}~\bsnm{Groom}},
-\bauthor{\binits{D.}~\bsnm{Agosti}} and
-\bauthor{\binits{L.}~\bsnm{Nielsen}},
-\batitle{Zenodo, an archive and publishing repository: A tale of two herbarium
-specimen pilot projects},
-\bjtitle{Biodiversity Information Science and Standards}
-\bvolume{3}
-(\byear{2019}).
-\bid{doi={10.3897/biss.3.37080}}
-\end{barticle}
-%
-\OrigBibText
-M.~Dillen,
-Q.~Groom,
-D.~Agosti and
-L.~Nielsen,
-Zenodo, an Archive and Publishing Repository: A tale of two herbarium specimen
-pilot projects,
-\textit{Biodiversity Information Science and Standards}
-\textbf{3}
-(2019), 00-00.
-doi:10.3897/biss.3.37080.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b41 ###bbsrt2
-%b47 ###
-\bibitem{ro-crate-py}
-%
-\begin{botherref}
-\oauthor{\binits{B.}~\bsnm{Droesbeke}},
-\oauthor{\binits{I.}~\bsnm{Eguinoa}},
-\oauthor{\binits{A.}~\bsnm{Gaignard}},
-\oauthor{\binits{L.}~\bsnm{Simone}},
-\oauthor{\binits{L.}~\bsnm{Pireddu}},
-\oauthor{\binits{L.}~\bsnm{Rodr\'{\i}guez-Navas}} and
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-{GitHub} -- {ResearchObject}/ro-crate-py: Python library for {RO}-Crate.
-\url{https://github.com/researchobject/ro-crate-py}.
-\bid{doi={10.5281/zenodo.3956493}}
-\end{botherref}
-%
-\OrigBibText
-B.~Droesbeke,
-I.~Eguinoa,
-A.~Gaignard,
-L.~Simone,
-L.~Pireddu,
-L.~Rodr\'iguez-Navas and
-S.~Soiland-Reyes,
-{GitHub} - {ResearchObject}/ro-crate-py: Python library for {RO}-Crate.
-doi:10.5281/zenodo.3956493.
-\url{https://github.com/researchobject/ro-crate-py}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b42 ###bbsrt2
-%b30 ###
-\bibitem{doi:10.17487/rfc3987}
-%
-\begin{botherref}
-\oauthor{\binits{M.}~\bsnm{Duerst}} and
-\oauthor{\binits{M.}~\bsnm{Suignard}},
-Internationalized resource identifiers ({IRI}s),
-RFC 3987,
-\bsertitle{Internet Requests for Comments},
-\bpublisher{RFC Editor},
-\byear(2005).
-\bid{doi={10.17487/rfc3987}}
-\end{botherref}
-%
-\OrigBibText
-M.~Duerst and
-M.~Suignard,
-Internationalized resource identifiers ({IRI}s),
-RFC 3987,
-2005.
-doi:10.17487/rfc3987.
-\url{https://www.rfc-editor.org/info/rfc3987}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b43 ###bbsrt2
-%b112 ###
-\bibitem{ebi_ftp_umgs2019}
-%
-\begin{botherref}
-{EMBL-EBI Microbiome Informatics Team},
-{FTP} index of /pub/databases/metagenomics/umgs\_analyses/,
-2019.
-\url{http://ftp.ebi.ac.uk/pub/databases/metagenomics/umgs\_analyses/}.
-\end{botherref}
-%
-\OrigBibText
-{EMBL-EBI Microbiome Informatics Team},
-{FTP} index of /pub/databases/metagenomics/umgs\_analyses/,
-2019.
-\url{http://ftp.ebi.ac.uk/pub/databases/metagenomics/umgs\_analyses/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b44 ###bbsrt2
-% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
-%incorrectly.
-%b113 ###
-\bibitem{finn-lab-mgsgut}
-%
-\begin{botherref}
-EMBL-EBI Microbiome Informatics Team,
-{GitHub} -- Finn-Lab/{MGS}-gut: Analysing Metagenomic Species ({MGS}),
-\url{https://github.com/Finn-Lab/MGS-gut}.
-\end{botherref}
-%
-\OrigBibText
-{EMBL-EBI Microbiome Informatics Team},
-{GitHub} - Finn-Lab/{MGS}-gut: Analysing Metagenomic Species ({MGS}).
-\url{https://github.com/Finn-Lab/MGS-gut}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b45 ###bbsrt2
-%b65 ###
-\bibitem{doi:10.1038/s41587-020-0439-x}
-%
-\begin{barticle}
-\bauthor{\binits{P.A.}~\bsnm{Ewels}},
-\bauthor{\binits{A.}~\bsnm{Peltzer}},
-\bauthor{\binits{S.}~\bsnm{Fillinger}},
-\bauthor{\binits{H.}~\bsnm{Patel}},
-\bauthor{\binits{J.}~\bsnm{Alneberg}},
-\bauthor{\binits{A.}~\bsnm{Wilm}},
-\bauthor{\binits{M.U.}~\bsnm{Garcia}},
-\bauthor{\binits{P.}~\bsnm{Di Tommaso}} and
-\bauthor{\binits{S.}~\bsnm{Nahnsen}},
-\batitle{The nf-core framework for community-curated bioinformatics pipelines},
-\bjtitle{Nature Biotechnology}
-\bvolume{38}(\bissue{3})
-(\byear{2020}),
-\bfpage{276}--\blpage{278}.
-\bid{doi={10.1038/s41587-020-0439-x}}
-\end{barticle}
-%
-\OrigBibText
-P.A.~Ewels,
-A.~Peltzer,
-S.~Fillinger,
-H.~Patel,
-J.~Alneberg,
-A.~Wilm,
-M.U.~Garcia,
-P.~Di~Tommaso and
-S.~Nahnsen,
-\textit{The nf-core framework for community-curated bioinformatics
-pipelines.},
-Nature Biotechnology
-38(3)
-(2020),
-276--278.
-doi:10.1038/s41587-020-0439-x.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b46 ###bbsrt2
-%b15 ###
-\bibitem{farnel_2014}
-%
-\begin{bchapter}
-\bauthor{\binits{S.}~\bsnm{Farnel}} and
-\bauthor{\binits{A.}~\bsnm{Shiri}},
-\bctitle{Metadata for research data: Current practices and trends},
-in: \bbtitle{2014 Proceedings of the International Conference on Dublin
-Core and Metadata Applications},
-\beditor{\binits{W.}~\bsnm{Moen}} and
-\beditor{\binits{A.}~\bsnm{Rushing}}, eds,
-\bsertitle{Dublin Core Metadata Initiative},
-\byear{2014},
-\bcomment{ISSN 1939-1366}. \url
-{https://dcpapers.dublincore.org/pubs/article/view/3714}.
-\end{bchapter}
-%
-\OrigBibText
-S.~Farnel and
-A.~Shiri,
-\textit{Metadata for Research Data: Current Practices and Trends},
-in: 2014 Proceedings of the International Conference on Dublin Core
-and Metadata Applications,
-W.~Moen and
-A.~Rushing, eds,
-Dublin Core Metadata Initiative,
-2014.
-ISSN 1939-1366.
-\url{https://dcpapers.dublincore.org/pubs/article/view/3714}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b47 ###bbsrt2
-%b107 ###
-\bibitem{doip2.0}
-%
-\begin{botherref}
-\oauthor{\binits{D.}~\bsnm{Foundation}},
-Digital Object Interface Protocol Specification,
-version 2.0, Technical Report,
-(2018).
-\url{https://www.dona.net/sites/default/files/2018-11/DOIPv2Spec\_1.pdf}.
-\end{botherref}
-%
-\OrigBibText
-D.~Foundation,
-Digital Object Interface Protocol Specification, version 2.0,
-Technical Report,
-2018.
-\url{https://www.dona.net/sites/default/files/2018-11/DOIPv2Spec_1.pdf}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b48 ###bbsrt2
-%b104 ###
-\bibitem{doi:10.1016/j.future.2019.03.046}
-%
-\begin{barticle}
-\bauthor{\binits{A.}~\bsnm{Garcia-Silva}},
-\bauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
-\bauthor{\binits{R.}~\bsnm{Palma}},
-\bauthor{\binits{M.}~\bsnm{Krystek}},
-\bauthor{\binits{S.}~\bsnm{Mantovani}},
-\bauthor{\binits{F.}~\bsnm{Foglini}},
-\bauthor{\binits{V.}~\bsnm{Grande}},
-\bauthor{\binits{F.}~\bsnm{De Leo}},
-\bauthor{\binits{S.}~\bsnm{Salvi}},
-\bauthor{\binits{E.}~\bsnm{Trasatti}},
-\bauthor{\binits{V.}~\bsnm{Romaniello}},
-\bauthor{\binits{M.}~\bsnm{Albani}},
-\bauthor{\binits{C.}~\bsnm{Silvagni}},
-\bauthor{\binits{R.}~\bsnm{Leone}},
-\bauthor{\binits{F.}~\bsnm{Marelli}},
-\bauthor{\binits{S.}~\bsnm{Albani}},
-\bauthor{\binits{M.}~\bsnm{Lazzarini}},
-\bauthor{\binits{H.J.}~\bsnm{Napier}},
-\bauthor{\binits{H.M.}~\bsnm{Glaves}},
-\bauthor{\binits{T.}~\bsnm{Aldridge}},
-\bauthor{\binits{C.}~\bsnm{Meertens}},
-\bauthor{\binits{F.}~\bsnm{Boler}},
-\bauthor{\binits{H.W.}~\bsnm{Loescher}},
-\bauthor{\binits{C.}~\bsnm{Laney}},
-\bauthor{\binits{M.A.}~\bsnm{Genazzio}},
-\bauthor{\binits{D.}~\bsnm{Crawl}} and
-\bauthor{\binits{I.}~\bsnm{Altintas}},
-\batitle{Enabling {FAIR} research in Earth science through research objects},
-\bjtitle{Future Generation Computer Systems}
-\bvolume{98}
-(\byear{2019}),
-\bfpage{550}--\blpage{564}.
-\bid{doi={10.1016/j.future.2019.03.046}}
-\end{barticle}
-%
-\OrigBibText
-A.~Garcia-Silva,
-J.M.~Gomez-Perez,
-R.~Palma,
-M.~Krystek,
-S.~Mantovani,
-F.~Foglini,
-V.~Grande,
-F.~De~Leo,
-S.~Salvi,
-E.~Trasatti,
-V.~Romaniello,
-M.~Albani,
-C.~Silvagni,
-R.~Leone,
-F.~Marelli,
-S.~Albani,
-M.~Lazzarini,
-H.J.~Napier,
-H.M.~Glaves,
-T.~Aldridge,
-C.~Meertens,
-F.~Boler,
-H.W.~Loescher,
-C.~Laney,
-M.A.~Genazzio,
-D.~Crawl and
-I.~Altintas,
-\textit{Enabling {FAIR} research in Earth Science through research objects},
-Future Generation Computer Systems
-98
-(2019),
-550--564.
-doi:10.1016/j.future.2019.03.046.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b49 ###bbsrt2
-%b45 ###
-\bibitem{ro-crate-js}
-%
-\begin{botherref}
-{{GitHub}} -- {UTS}-{eResearch}/ro-crate-js: Research Object
-Crate ({RO}-Crate) utilities,
-\url{https://github.com/UTS-eResearch/ro-crate-js}.
-\end{botherref}
-%
-\OrigBibText
-{GitHub} - {UTS}-{eResearch}/ro-crate-js: Research Object Crate ({RO}-Crate)
-utilities.
-\url{https://github.com/UTS-eResearch/ro-crate-js}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b50 ###bbsrt2
-% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
-%incorrectly.
-%b51 ###
-\bibitem{galaxy2cwl}
-%
-\begin{botherref}
-{GitHub} -- workflowhub-eu/galaxy2cwl: Standalone version tool
-to get cwl descriptions (initially an abstract cwl interface)
-of galaxy workflows and Galaxy workflows executions.
-\url{https://github.com/workflowhub-eu/galaxy2cwl}.
-\end{botherref}
-%
-\OrigBibText
-{GitHub} - workflowhub-eu/galaxy2cwl: Standalone version tool to get cwl
-descriptions (initially an abstract cwl interface) of galaxy workflows and
-Galaxy workflows executions.
-\url{https://github.com/workflowhub-eu/galaxy2cwl}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b51 ###bbsrt2
-%b52 ###
-\bibitem{modpdsc}
-%
-\begin{botherref}
-{{GitHub} -- {CoEDL}/modpdsc},
-\url{https://github.com/CoEDL/modpdsc/}.
-\end{botherref}
-%
-\OrigBibText
-{GitHub} - {CoEDL}/modpdsc.
-\url{https://github.com/CoEDL/modpdsc/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b52 ###bbsrt2
-% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
-%incorrectly.
-%b54 ###
-\bibitem{ocfl-tools}
-%
-\begin{botherref}
-{GitHub} -- {CoEDL}/ocfl-tools: Tools to process and manipulate an
-{OCFL} tree.
-\url{https://github.com/CoEDL/ocfl-tools}.
-\end{botherref}
-%
-\OrigBibText
-{GitHub} - {CoEDL}/ocfl-tools: Tools to process and manipulate an
-{OCFL} tree.
-\url{https://github.com/CoEDL/ocfl-tools}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b53 ###bbsrt2
-% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
-%incorrectly.
-%b115 ###
-\bibitem{doi:10.1038/s41592-019-0350-x}
-%
-\begin{botherref}
-Giving software its due,
-\textit{Nature Methods}
-\textbf{16}(3)
-(2019),
-207--207.
-\bid{doi={10.1038/s41592-019-0350-x}}
-\end{botherref}
-%
-\OrigBibText
-Giving software its due,
-\textit{Nature Methods}
-\textbf{16}(3)
-(2019),
-207--207.
-doi:10.1038/s41592-019-0350-x.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b54 ###bbsrt2
-%b117 ###
-\bibitem{goble_presentation_2016}
-%
-\begin{bchapter}
-\bauthor{\binits{C.}~\bsnm{Goble}},
-in: \bbtitle{What Is Reproducibility? The R* Brouhaha},
-\bconflocation{Hannover, Germany},
-\byear{2016}.
-\url
-{http://repscience2016.research-infrastructures.eu/img/CaroleGoble-ReproScience2016v2.pdf}.
-\end{bchapter}
-%
-\OrigBibText
-C.~Goble,
-What is Reproducibility? The R* Brouhaha,
-Hannover, Germany, 2016.
-\url
-{http://repscience2016.research-infrastructures.eu/img/CaroleGoble-ReproScience2016v2.pdf}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b55 ###bbsrt2
-%b4 ###
-\bibitem{doi:10.1162/dint_a_00033}
-%
-\begin{barticle}
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{S.}~\bsnm{Cohen-Boulakia}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\bauthor{\binits{D.}~\bsnm{Garijo}},
-\bauthor{\binits{Y.}~\bsnm{Gil}},
-\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
-\bauthor{\binits{K.}~\bsnm{Peters}} and
-\bauthor{\binits{D.}~\bsnm{Schober}},
-\bjtitle{{FAIR} Computational Workflows, Data Intelligence}
-\bvolume{2}(\bissue{1--2})
-(\byear{2019}),
-\bfpage{108}--\blpage{121}.
-\bid{doi={10.1162/dint\_a\_00033}}
-\end{barticle}
-%
-\OrigBibText
-C.~Goble,
-S.~Cohen-Boulakia,
-S.~Soiland-Reyes,
-D.~Garijo,
-Y.~Gil,
-M.R.~Crusoe,
-K.~Peters and
-D.~Schober,
-\textit{{FAIR} Computational Workflows},
-Data Intelligence
-2(1--2)
-(2019),
-108--121.
-doi:10.1162/dint\_a\_00033.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b56 ###bbsrt2
-%b70 ###
-\bibitem{doi:10.5281/zenodo.4605654}
-%
-\begin{barticle}
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\bauthor{\binits{F.}~\bsnm{Bacall}},
-\bauthor{\binits{S.}~\bsnm{Owen}},
-\bauthor{\binits{A.}~\bsnm{Williams}},
-\bauthor{\binits{I.}~\bsnm{Eguinoa}},
-\bauthor{\binits{B.}~\bsnm{Droesbeke}},
-\bauthor{\binits{S.}~\bsnm{Leo}},
-\bauthor{\binits{L.}~\bsnm{Pireddu}},
-\bauthor{\binits{L.}~\bsnm{Rodr\'{\i}guez-Navas}},
-\bauthor{\binits{J.M.}~\bsnm{Fern\'{a}ndez}},
-\bauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
-\bauthor{\binits{H.}~\bsnm{M\'{e}nager}},
-\bauthor{\binits{B.}~\bsnm{Gr\"{u}ning}},
-\bauthor{\binits{B.}~\bsnm{Serrano-Solano}},
-\bauthor{\binits{P.}~\bsnm{Ewels}} and
-\bauthor{\binits{F.}~\bsnm{Coppens}},
-\batitle{Implementing {FAIR} digital objects in the {EOSC}-life
-workflow collaboratory},
-\bjtitle{Zenodo}
-(\byear{2021}).
-\bid{doi={10.5281/zenodo.4605654}}
-\end{barticle}
-%
-\OrigBibText
-C.~Goble,
-S.~Soiland-Reyes,
-F.~Bacall,
-S.~Owen,
-A.~Williams,
-I.~Eguinoa,
-B.~Droesbeke,
-S.~Leo,
-L.~Pireddu,
-L.~Rodr\'iguez-Navas,
-J.M.~Fern\'{a}ndez,
-S.~Capella-Gutierrez,
-H.~M\'{e}nager,
-B.~Gr\"{u}ning,
-B.~Serrano-Solano,
-P.~Ewels and
-F.~Coppens,
-\textit{Implementing {FAIR} Digital Objects in the {EOSC}-Life Workflow
-Collaboratory},
-Zenodo
-(2021).
-doi:10.5281/zenodo.4605654.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b57 ###bbsrt2
-%b98 ###
-\bibitem{doi:10.1093/nar/gkq429}
-%
-\begin{barticle}
-\bauthor{\binits{C.A.}~\bsnm{Goble}},
-\bauthor{\binits{J.}~\bsnm{Bhagat}},
-\bauthor{\binits{S.}~\bsnm{Aleksejevs}},
-\bauthor{\binits{D.}~\bsnm{Cruickshank}},
-\bauthor{\binits{D.}~\bsnm{Michaelides}},
-\bauthor{\binits{D.}~\bsnm{Newman}},
-\bauthor{\binits{M.}~\bsnm{Borkum}},
-\bauthor{\binits{S.}~\bsnm{Bechhofer}},
-\bauthor{\binits{M.}~\bsnm{Roos}},
-\bauthor{\binits{P.}~\bsnm{Li}} and
-\bauthor{\binits{D.}~\bsnm{De Roure}},
-\batitle{{myExperiment}: A repository and social network for the sharing
-of bioinformatics workflows},
-\bjtitle{Nucleic Acids Research}
-\bvolume{38}\bcomment{(Web Server issue)}
-(\byear{2010}),
-\bfpage{W677}--\blpage{W682}.
-\bid{doi={10.1093/nar/gkq429}}
-\end{barticle}
-%
-\OrigBibText
-C.A.~Goble,
-J.~Bhagat,
-S.~Aleksejevs,
-D.~Cruickshank,
-D.~Michaelides,
-D.~Newman,
-M.~Borkum,
-S.~Bechhofer,
-M.~Roos,
-P.~Li and
-D.~DeRoure,
-\textit{{myExperiment}: a repository and social network for the sharing of
-bioinformatics workflows}.
-Nucleic Acids Research
-38 (Web Server issue)
-(2010),
-W677--82.
-doi:10.1093/nar/gkq429.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b58 ###bbsrt2
-%b37 ###
-\bibitem{bioschemas_2017}
-%
-\begin{bchapter}
-\bauthor{\binits{A.}~\bsnm{Gray}},
-\bauthor{\binits{C.}~\bsnm{Goble}} and
-\bauthor{\binits{R.}~\bsnm{Jimenez}},
-in: \bbtitle{{Bioschemas Community}, Bioschemas: From Potato Salad to
-Protein Annotation},
-\bconflocation{Vienna, Austria},
-\byear{2017}.
-\url{https://iswc2017.semanticweb.org/paper-579/}.
-\end{bchapter}
-%
-\OrigBibText
-A.~Gray,
-C.~Goble,
-R.~Jimenez and
-{Bioschemas Community},
-Bioschemas: From Potato Salad to Protein Annotation,
-Vienna, Austria, 2017.
-\url{https://iswc2017.semanticweb.org/paper-579/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b59 ###bbsrt2
-%b88 ###
-\bibitem{doi:10.1109/MCSE.2016.92}
-%
-\begin{barticle}
-\bauthor{\binits{R.L.}~\bsnm{Grossman}},
-\bauthor{\binits{A.}~\bsnm{Heath}},
-\bauthor{\binits{M.}~\bsnm{Murphy}},
-\bauthor{\binits{M.}~\bsnm{Patterson}} and
-\bauthor{\binits{W.}~\bsnm{Wells}},
-\batitle{A case for data commons: Toward data science as a service},
-\bjtitle{Computing in Science \& Engineering}
-\bvolume{18}(\bissue{5})
-(\byear{2016}),
-\bfpage{10}--\blpage{20}.
-\bid{doi={10.1109/MCSE.2016.92}}
-\end{barticle}
-%
-\OrigBibText
-R.L.~Grossman,
-A.~Heath,
-M.~Murphy,
-M.~Patterson and
-W.~Wells,
-\textit{A case for data commons: toward data science as a service},
-Computing in science \& engineering
-18(5)
-(2016),
-10--20.
-doi:10.1109/MCSE.2016.92.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b60 ###bbsrt2
-%b110 ###
-\bibitem{doi:10.1016/j.cels.2018.03.014}
-%
-\begin{barticle}
-\bauthor{\binits{B.}~\bsnm{Gr\"{u}ning}},
-\bauthor{\binits{J.}~\bsnm{Chilton}},
-\bauthor{\binits{J.}~\bsnm{K\"{o}ster}},
-\bauthor{\binits{R.}~\bsnm{Dale}},
-\bauthor{\binits{N.}~\bsnm{Soranzo}},
-\bauthor{\binits{M.}~\bsnm{van den Beek}},
-\bauthor{\binits{J.}~\bsnm{Goecks}},
-\bauthor{\binits{R.}~\bsnm{Backofen}},
-\bauthor{\binits{A.}~\bsnm{Nekrutenko}} and
-\bauthor{\binits{J.}~\bsnm{Taylor}},
-\batitle{Practical computational reproducibility in the life sciences},
-\bjtitle{Cell Systems}
-\bvolume{6}(\bissue{6})
-(\byear{2018}),
-\bfpage{631}--\blpage{635}.
-\bid{doi={10.1016/j.cels.2018.03.014}}
-\end{barticle}
-%
-\OrigBibText
-B.~Gr\"{u}ning,
-J.~Chilton,
-J.~K\"{o}ster,
-R.~Dale,
-N.~Soranzo,
-M.~van~den Beek,
-J.~Goecks,
-R.~Backofen,
-A.~Nekrutenko and
-J.~Taylor,
-\textit{Practical computational reproducibility in the life sciences.},
-Cell Systems
-6(6)
-(2018),
-631--635.
-doi:10.1016/j.cels.2018.03.014.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b61 ###bbsrt2
-%b119 ###
-\bibitem{doi:10.1038/s41592-018-0046-7}
-%
-\begin{barticle}
-\bauthor{\binits{B.}~\bsnm{Gr\"{u}ning}},
-\bauthor{\binits{R.}~\bsnm{Dale}},
-\bauthor{\binits{A.}~\bsnm{Sj\"{o}din}},
-\bauthor{\binits{B.A.}~\bsnm{Chapman}},
-\bauthor{\binits{J.}~\bsnm{Rowe}},
-\bauthor{\binits{C.H.}~\bsnm{Tomkins-Tinch}},
-\bauthor{\binits{R.}~\bsnm{Valieris}},
-\bauthor{\binits{J.}~\bsnm{K\"{o}ster}} and
-\bauthor{\binits{B.}~\bsnm{Team}},
-\batitle{Bioconda: Sustainable and comprehensive software distribution
-for the life sciences},
-\bjtitle{Nature Methods}
-\bvolume{15}(\bissue{7})
-(\byear{2018}),
-\bfpage{475}--\blpage{476}.
-\bid{doi={10.1038/s41592-018-0046-7}}
-\end{barticle}
-%
-\OrigBibText
-B.~Gr\"{u}ning,
-R.~Dale,
-A.~Sj\"{o}din,
-B.A.~Chapman,
-J.~Rowe,
-C.H.~Tomkins-Tinch,
-R.~Valieris,
-J.~K\"{o}ster and
-B.~Team,
-\textit{Bioconda: sustainable and comprehensive software distribution
-for the
-life sciences.},
-Nature Methods
-15(7)
-(2018),
-475--476.
-doi:10.1038/s41592-018-0046-7.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b62 ###bbsrt2
-%b34 ###
-\bibitem{doi:10.1145/2857274.2857276}
-%
-\begin{barticle}
-\bauthor{\binits{R.V.}~\bsnm{Guha}},
-\bauthor{\binits{D.}~\bsnm{Brickley}} and
-\bauthor{\binits{S.}~\bsnm{Macbeth}},
-\batitle{Schema.org: Evolution of Structured Data on the Web: Big data
-makes common schemas even more necessary},
-\bjtitle{Queue}
-\bvolume{13}(\bissue{9})
-(\byear{2015}),
-\bfpage{10}--\blpage{37}.
-\bid{doi={10.1145/2857274.2857276}}
-\end{barticle}
-%
-\OrigBibText
-R.V.~Guha,
-D.~Brickley and
-S.~Macbeth,
-\textit{Schema.org: Evolution of Structured Data on the Web: Big data makes
-common schemas even more necessary},
-Queue
-13(9)
-(2015),
-10--37.
-doi:10.1145/2857274.2857276.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b63 ###bbsrt2
-%b24 ###
-\bibitem{doi:10.2200/S00334ED1V01Y201102WBE001}
-%
-\begin{bchapter}
-\bauthor{\binits{T.}~\bsnm{Heath}} and
-\bauthor{\binits{C.}~\bsnm{Bizer}},
-in: \bbtitle{Linked Data: Evolving the Web into a Global Data Space},
-Vol.~\bseriesno{1},
-\byear{2011},
-pp.~\bfpage{1}--\blpage{136},
-\bcomment{ISSN 2160-4711}.
-ISBN \bisbn{9781608454310}.
-\end{bchapter}
-%
-\OrigBibText
-T.~Heath and
-C.~Bizer,
-\textit{Linked Data: Evolving the Web into a Global Data Space},
-Vol.~1,
-2011,
-pp.~1--136.
-ISSN 2160-4711.
-ISBN 9781608454310.
-doi:10.2200/S00334ED1V01Y201102WBE001.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b64 ###bbsrt2
-% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
-%incorrectly.
-%b78 ###
-\bibitem{doi:10.1109/IEEESTD.2020.9094416}
-%
-\begin{botherref}
-IEEE Standard for Bioinformatics Analyses Generated by High-Throughput
-Sequencing (HTS) to Facilitate Communication,
-IEEE Std 2791-2020.
-ISBN 978-1-5044-6466-6.
-\bid{doi={10.1109/IEEESTD.2020.9094416}}
-\end{botherref}
-%
-\OrigBibText
-IEEE Standard for Bioinformatics Analyses Generated by High-Throughput
-Sequencing (HTS) to Facilitate Communication,
-IEEE Std 2791-2020.
-ISBN 978-1-5044-6466-6.
-doi:10.1109/IEEESTD.2020.9094416.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b65 ###bbsrt2
-%b90 ###
-\bibitem{doi:10.1182/blood-2017-03-735654}
-%
-\begin{barticle}
-\bauthor{\binits{M.A.}~\bsnm{Jensen}},
-\bauthor{\binits{V.}~\bsnm{Ferretti}},
-\bauthor{\binits{R.L.}~\bsnm{Grossman}} and
-\bauthor{\binits{L.M.}~\bsnm{Staudt}},
-\batitle{The {NCI} Genomic Data Commons as an engine for precision medicine},
-\bjtitle{Blood}
-\bvolume{130}(\bissue{4})
-(\byear{2017}),
-\bfpage{453}--\blpage{459}.
-\bid{doi={10.1182/blood-2017-03-735654}}
-\end{barticle}
-%
-\OrigBibText
-M.A.~Jensen,
-V.~Ferretti,
-R.L.~Grossman and
-L.M.~Staudt,
-\textit{The {NCI} Genomic Data Commons as an engine for precision medicine.},
-Blood
-130(4)
-(2017),
-453--459.
-doi:10.1182/blood-2017-03-735654.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b66 ###bbsrt2
-%b125 ###
-\bibitem{doi:10.5281/zenodo.4477164}
-%
-\begin{botherref}
-\oauthor{\binits{M.B.}~\bsnm{Jones}},
-\oauthor{\binits{S.}~\bsnm{Richard}},
-\oauthor{\binits{D.}~\bsnm{Vieglais}},
-\oauthor{\binits{A.}~\bsnm{Shepherd}},
-\oauthor{\binits{R.}~\bsnm{Duerr}},
-\oauthor{\binits{D.}~\bsnm{Fils}} and
-\oauthor{\binits{L.}~\bsnm{McGibbney}},
-Science-on-Schema.org v1.2.0, 
-2021.
-\bid{doi={10.5281/zenodo.4477164}}
-\end{botherref}
-%
-\OrigBibText
-M.B.~Jones,
-S.~Richard,
-D.~Vieglais,
-A.~Shepherd,
-R.~Duerr,
-D.~Fils and
-L.~McGibbney,
-Science-on-Schema.org v1.2.0,
-2021.
-doi:10.5281/zenodo.4477164.
-\url{https://science-on-schema.org/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b67 ###bbsrt2
-%b103 ###
-\bibitem{doi:10.3233/978-1-61499-660-6-9}
-%
-\begin{bchapter}
-\bauthor{\binits{M.}~\bsnm{Katsumi}} and
-\bauthor{\binits{M.}~\bsnm{Gr\"{u}ninger}},
-\bctitle{What is ontology reuse?}
-in: \bbtitle{Formal Ontology in Information Systems},
-\beditor{\binits{R.}~\bsnm{Ferrario}} and
-\beditor{\binits{W.}~\bsnm{Kuhn}}, eds,
-\bsertitle{Frontiers in Artificial Intelligence and Applications},
-Vol.~\bseriesno{283},
-\bpublisher{{IOS} Press},
-\byear{2016},
-pp.~\bfpage{9}--\blpage{22}.
-ISBN \bisbn{978-1-61499-660-6}.
-\bid{doi={10.3233/978-1-61499-660-6-9}}
-\end{bchapter}
-%
-\OrigBibText
-M.~Katsumi and
-M.~Gr\"{u}ninger,
-\textit{What Is Ontology Reuse?},
-in: Formal Ontology in Information Systems,
-R.~Ferrario and
-W.~Kuhn, eds,
-Frontiers in Artificial Intelligence and Applications,
-Vol.~283,
-{IOS} Press,
-2016,
-pp.~9--22.
-ISBN 978-1-61499-660-6.
-doi:10.3233/978-1-61499-660-6-9.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b68 ###bbsrt2
-%b35 ###
-\bibitem{doi:10.1093/gigascience/giz095}
-%
-\begin{barticle}
-\bauthor{\binits{F.Z.}~\bsnm{Khan}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\bauthor{\binits{R.O.}~\bsnm{Sinnott}},
-\bauthor{\binits{A.}~\bsnm{Lonie}},
-\bauthor{\binits{C.}~\bsnm{Goble}} and
-\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
-\batitle{Sharing interoperable workflow provenance: A review of best practices
-and their practical application in {CWLProv}},
-\bjtitle{GigaScience}
-\bvolume{8}(\bissue{11})
-(\byear{2019}).
-\bid{doi={10.1093/gigascience/giz095}}
-\end{barticle}
-%
-\OrigBibText
-F.Z.~Khan,
-S.~Soiland-Reyes,
-R.O.~Sinnott,
-A.~Lonie,
-C.~Goble and
-M.R.~Crusoe,
-Sharing interoperable workflow provenance: A review of best practices
-and their
-practical application in {CWLProv},
-\textit{GigaScience}
-\textbf{8}(11)
-(2019)
-00.
-doi:10.1093/gigascience/giz095.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b69 ###bbsrt2
-%b109 ###
-\bibitem{doi:10.1002/cpe.1228}
-%
-\begin{barticle}
-\bauthor{\binits{J.}~\bsnm{Kim}},
-\bauthor{\binits{E.}~\bsnm{Deelman}},
-\bauthor{\binits{Y.}~\bsnm{Gil}},
-\bauthor{\binits{G.}~\bsnm{Mehta}} and
-\bauthor{\binits{V.}~\bsnm{Ratnakar}},
-\batitle{Provenance trails in the Wings/Pegasus system},
-\bjtitle{Concurrency and Computation: Practice and Experience}
-\bvolume{20}(\bissue{5})
-(\byear{2008}),
-\bfpage{587}--\blpage{597}.
-\bid{doi={10.1002/cpe.1228}}
-\end{barticle}
-%
-\OrigBibText
-J.~Kim,
-E.~Deelman,
-Y.~Gil,
-G.~Mehta and
-V.~Ratnakar,
-\textit{Provenance trails in the Wings/Pegasus system},
-Concurrency and Computation: Practice and Experience
-20(5)
-(2008),
-587--597.
-doi:10.1002/cpe.1228.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b70 ###bbsrt2
-%b23 ###
-\bibitem{doi:10.3233/978-1-61499-649-1-87}
-%
-\begin{bchapter}
-\bauthor{\binits{T.}~\bsnm{Kluyver}},
-\bauthor{\binits{B.}~\bsnm{Ragan-Kelley}},
-\bauthor{\binits{F.}~\bsnm{P\'{e}rez}},
-\bauthor{\binits{B.}~\bsnm{Granger}},
-\bauthor{\binits{M.}~\bsnm{Bussonnier}},
-\bauthor{\binits{J.}~\bsnm{Frederic}},
-\bauthor{\binits{K.}~\bsnm{Kelley}},
-\bauthor{\binits{J.}~\bsnm{Hamrick}},
-\bauthor{\binits{J.}~\bsnm{Grout}},
-\bauthor{\binits{S.}~\bsnm{Corlay}},
-\bauthor{\binits{P.}~\bsnm{Ivanov}},
-\bauthor{\binits{D.}~\bsnm{Avila}},
-\bauthor{\binits{S.}~\bsnm{Abdalla}},
-\bauthor{\binits{C.}~\bsnm{Willing}} and
-\binstitute{Jupyter Development Team},
-\bctitle{Jupyter Notebooks -- a publishing format for reproducible
-computational workflows},
-in: \bbtitle{Positioning and Power in Academic Publishing: Players,
-Agents and Agendas},
-\bsertitle{Proceedings of the 20th International Conference on
-Electronic Publishing},
-\bpublisher{IOS Press},
-\byear{2016},
-pp.~\bfpage{87}--\blpage{90},
-ISBN \bisbn{978-1-61499-649-1}.
-\bid{doi={10.3233/978-1-61499-649-1-87}}
-\end{bchapter}
-%
-\OrigBibText
-T.~Kluyver,
-B.~Ragan-Kelley,
-F.~P\'{e}rez,
-B.~Granger,
-M.~Bussonnier,
-J.~Frederic,
-K.~Kelley,
-J.~Hamrick,
-J.~Grout,
-S.~Corlay,
-P.~Ivanov,
-D.~Avila,
-S.~Abdalla,
-C.~Willing and
-{Jupyter Development Team},
-\textit{Jupyter Notebooks - a publishing format for reproducible computational
-workflows},
-in: Positioning and Power in Academic Publishing: Players, Agents and
-Agendas,
-IOS Press,
-2016,
-pp.~87--90,
-Proceedings of the 20th International Conference on Electronic
-Publishing.
-ISBN 978-1-61499-649-1.
-doi:10.3233/978-1-61499-649-1-87.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b71 ###bbsrt2
-%b10 ###
-\bibitem{doi:10.1016/j.ijhcs.2020.102562}
-%
-\begin{barticle}
-\bauthor{\binits{L.}~\bsnm{Koesten}},
-\bauthor{\binits{K.}~\bsnm{Gregory}},
-\bauthor{\binits{P.}~\bsnm{Groth}} and
-\bauthor{\binits{E.}~\bsnm{Simperl}},
-\batitle{Talking datasets -- understanding data sensemaking behaviours},
-\bjtitle{International journal of human-computer studies}
-\bvolume{146}
-(\byear{2021}),
-\bnumber{102562}.
-\bid{doi={10.1016/j.ijhcs.2020.102562}}
-\end{barticle}
-%
-\OrigBibText
-L.~Koesten,
-K.~Gregory,
-P.~Groth and
-E.~Simperl,
-\textit{Talking datasets -- Understanding data sensemaking behaviours},
-International journal of human-computer studies
-146
-(2021),
-102562.
-doi:10.1016/j.ijhcs.2020.102562.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b72 ###bbsrt2
-%b94 ###
-\bibitem{doi:10.1016/j.patter.2020.100136}
-%
-\begin{barticle}
-\bauthor{\binits{L.}~\bsnm{Koesten}},
-\bauthor{\binits{P.}~\bsnm{Vougiouklis}},
-\bauthor{\binits{E.}~\bsnm{Simperl}} and
-\bauthor{\binits{P.}~\bsnm{Groth}},
-\batitle{Dataset reuse: Toward translating principles to practice},
-\bjtitle{Patterns (New York, NY)}
-\bvolume{1}(\bissue{8})
-(\byear{2020}),
-\bnumber{100136}.
-\bid{doi={10.1016/j.patter.2020.100136}}
-\end{barticle}
-%
-\OrigBibText
-L.~Koesten,
-P.~Vougiouklis,
-E.~Simperl and
-P.~Groth,
-\textit{Dataset reuse: toward translating principles to practice},
-Patterns (New York, N.Y.)
-1(8)
-(2020),
-100136.
-doi:10.1016/j.patter.2020.100136.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b73 ###bbsrt2
-%b67 ###
-\bibitem{doi:10.1093/bioinformatics/bts480}
-%
-\begin{barticle}
-\bauthor{\binits{J.}~\bsnm{K\"{o}ster}} and
-\bauthor{\binits{S.}~\bsnm{Rahmann}},
-\batitle{Snakemake -- a scalable bioinformatics workflow engine},
-\bjtitle{Bioinformatics}
-\bvolume{28}(\bissue{19})
-(\byear{2012}),
-\bfpage{2520}--\blpage{2522}.
-\bid{doi={10.1093/bioinformatics/bts480}}
-\end{barticle}
-%
-\OrigBibText
-J.~K\"{o}ster and
-S.~Rahmann,
-\textit{Snakemake--a scalable bioinformatics workflow engine.},
-Bioinformatics
-28(19)
-(2012),
-2520--2522.
-doi:10.1093/bioinformatics/bts480.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b74 ###bbsrt2
-%b26 ###
-\bibitem{doi:10.17487/rfc8493}
-%
-\begin{botherref}
-\oauthor{\binits{J.}~\bsnm{Kunze}},
-\oauthor{\binits{J.}~\bsnm{Littman}},
-\oauthor{\binits{E.}~\bsnm{Madden}},
-\oauthor{\binits{J.}~\bsnm{Scancella}} and
-\oauthor{\binits{C.}~\bsnm{Adams}},
-The {BagIt} File Packaging Format,
-(V1.0), 
-RFC 8493,
-\bsertitle{Internet Requests for Comments},
-\bpublisher{RFC Editor},
-(\byear{2018}).
-\bid{doi={10.17487/RFC8493}}
-\end{botherref}
-%
-\OrigBibText
-J.~Kunze,
-J.~Littman,
-E.~Madden,
-J.~Scancella and
-C.~Adams,
-The {BagIt} File Packaging Format (V1.0),
-Technical Report,
-2018.
-doi:10.17487/RFC8493.
-\url{https://www.rfc-editor.org/info/rfc8493}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b75 ###bbsrt2
-%b16 ###
-\bibitem{doi:10.2777/620649}
-%
-\begin{botherref}
-\oauthor{\binits{K.}~\bsnm{Kurowski}},
-\oauthor{\binits{O.}~\bsnm{Corcho}},
-\oauthor{\binits{C.}~\bsnm{Choirat}},
-\oauthor{\binits{M.}~\bsnm{Eriksson}},
-\oauthor{\binits{F.}~\bsnm{Coppens}},
-\oauthor{\binits{M.}~\bsnm{van de Sanden}} and
-\oauthor{\binits{M.}~\bsnm{Ojster\v{s}ek}},
-{EOSC} Interoperability Framework, Technical Report,
-2021.
-\bid{doi={10.2777/620649}}
-\end{botherref}
-%
-\OrigBibText
-K.~Kurowski,
-O.~Corcho,
-C.~Choirat,
-M.~Eriksson,
-F.~Coppens,
-M.~van~de Sanden and
-M.~Ojster\v{s}ek,
-{EOSC} Interoperability Framework,
-Technical Report,
-2021.
-doi:10.2777/620649.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b76 ###bbsrt2
-%b105 ###
-\bibitem{doi:10.3233/APC200107}
-%
-\begin{barticle}
-\bauthor{\binits{C.}~\bsnm{Kyle}},
-\bauthor{\binits{G.}~\bsnm{Niall}},
-\bauthor{\binits{H.}~\bsnm{Mihael}},
-\bauthor{\binits{K.}~\bsnm{Kacper}},
-\bauthor{\binits{L.}~\bsnm{Bertram}},
-\bauthor{\binits{M.}~\bsnm{Timothy}},
-\bauthor{\binits{N.}~\bsnm{Jarek}},
-\bauthor{\binits{S.}~\bsnm{Victoria}},
-\bauthor{\binits{T.}~\bsnm{Ian}},
-\bauthor{\binits{T.}~\bsnm{Thomas}} \betal,
-\batitle{Toward enabling reproducibility for data-intensive research
-using the Whole Tale platform},
-\bjtitle{Advances in Parallel Computing 36 (Parallel Computing:
-Technology Trends)}
-(\byear{2020}),
-\bfpage{766}--\blpage{778}.
-\bid{doi={10.3233/APC200107}}
-\end{barticle}
-%
-\OrigBibText
-C.~Kyle,
-G.~Niall,
-H.~Mihael,
-K.~Kacper,
-L.~Bertram,
-M.~Timothy,
-N.~Jarek,
-S.~Victoria,
-T.~Ian,
-T.~Thomas and
-et~al.,
-\textit{Toward Enabling Reproducibility for Data-Intensive Research
-Using the
-Whole Tale Platform},
-Advances in Parallel Computing
-36 (Parallel Computing: Technology Trends), \textbf{00}
-(2020),
-766--778.
-doi:10.3233/APC200107.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b77 ###bbsrt2
-%b43 ###
-\bibitem{describo-online}
-%
-\begin{botherref}
-\oauthor{\binits{M.}~\bsnm{La Rosa}},
-Arkisto Platform: Describo Online.
-\url{https://arkisto-platform.github.io/describo-online/}.
-\end{botherref}
-%
-\OrigBibText
-M.~La~Rosa,
-Arkisto Platform: Describo Online.
-\url{https://arkisto-platform.github.io/describo-online/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b78 ###bbsrt2
-%b22 ###
-\bibitem{describo}
-%
-\begin{botherref}
-\oauthor{\binits{M.}~\bsnm{La Rosa}} and
-\oauthor{\binits{P.}~\bsnm{Sefton}},
-Arkisto Platform: Describo, \url{https://arkisto-platform.github.io/describo/}.
-\end{botherref}
-%
-\OrigBibText
-M.~La~Rosa and
-P.~Sefton,
-Arkisto Platform: Describo.
-\url{https://arkisto-platform.github.io/describo/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b79 ###bbsrt2
-%b28 ###
-\bibitem{doi:10.6087/kcse.192}
-%
-\begin{barticle}
-\bauthor{\binits{R.}~\bsnm{Lammey}},
-\batitle{Solutions for identification problems: A look at the research
-organization registry},
-\bjtitle{Science Editing}
-\bvolume{7}(\bissue{1})
-(\byear{2020}),
-\bfpage{65}--\blpage{69}.
-\bid{doi={10.6087/kcse.192}}
-\end{barticle}
-%
-\OrigBibText
-R.~Lammey,
-\textit{Solutions for identification problems: a look at the Research
-Organization Registry},
-Science Editing
-7(1)
-(2020),
-65--69.
-doi:10.6087/kcse.192.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b80 ###bbsrt2
-%b3 ###
-\bibitem{doi:10.3233/DS-190026}
-%
-\begin{barticle}
-\bauthor{\binits{A.-L.}~\bsnm{Lamprecht}},
-\bauthor{\binits{L.}~\bsnm{Garcia}},
-\bauthor{\binits{M.}~\bsnm{Kuzak}},
-\bauthor{\binits{C.}~\bsnm{Martinez}},
-\bauthor{\binits{R.}~\bsnm{Arcila}},
-\bauthor{\binits{E.}~\bsnm{Martin Del Pico}},
-\bauthor{\binits{V.}~\bsnm{Dominguez Del Angel}},
-\bauthor{\binits{S.}~\bsnm{van de Sandt}},
-\bauthor{\binits{J.}~\bsnm{Ison}},
-\bauthor{\binits{P.A.}~\bsnm{Martinez}},
-\bauthor{\binits{P.}~\bsnm{McQuilton}},
-\bauthor{\binits{A.}~\bsnm{Valencia}},
-\bauthor{\binits{J.}~\bsnm{Harrow}},
-\bauthor{\binits{F.}~\bsnm{Psomopoulos}},
-\bauthor{\binits{J.L.}~\bsnm{Gelpi}},
-\bauthor{\binits{N.}~\bsnm{Chue Hong}},
-\bauthor{\binits{C.}~\bsnm{Goble}} and
-\bauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
-\batitle{Towards {FAIR} principles for research software},
-\bjtitle{Data Science},
-\bvolume{3}(\bissue{1})
-(\byear{2019}),
-\bfpage{1}--\blpage{23}.
-\bid{doi={10.3233/DS-190026}}
-\end{barticle}
-%
-\OrigBibText
-A.-L.~Lamprecht,
-L.~Garcia,
-M.~Kuzak,
-C.~Martinez,
-R.~Arcila,
-E.~Martin Del~Pico,
-V.~Dominguez Del~Angel,
-S.~van~de Sandt,
-J.~Ison,
-P.A.~Martinez,
-P.~{McQuilton},
-A.~Valencia,
-J.~Harrow,
-F.~Psomopoulos,
-J.L.~Gelpi,
-N.~Chue~Hong,
-C.~Goble and
-S.~Capella-Gutierrez,
-Towards {FAIR} principles for~research~software,
-\textit{Data Science}, \textbf{3},
-(2019),
-1--23.
-doi:10.3233/DS-190026.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b81 ###bbsrt2
-%b102 ###
-\bibitem{PROVO}
-%
-\begin{botherref}
-\oauthor{\binits{T.}~\bsnm{Lebo}},
-\oauthor{\binits{S.}~\bsnm{Sahoo}},
-\oauthor{\binits{D.}~\bsnm{McGuinness}},
-\oauthor{\binits{K.}~\bsnm{Belhajjame}},
-\oauthor{\binits{J.}~\bsnm{Cheney}},
-\oauthor{\binits{D.}~\bsnm{Corsar}},
-\oauthor{\binits{D.}~\bsnm{Garijo}},
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\oauthor{\binits{S.}~\bsnm{Zednik}} and
-\oauthor{\binits{J.}~\bsnm{Zhao}},
-PROV-O: The PROV Ontology,
-W3C Recommendation 30,
-2013.
-\url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
-\end{botherref}
-%
-\OrigBibText
-T.~Lebo,
-S.~Sahoo,
-D.~McGuinness,
-K.~Belhajjame,
-J.~Cheney,
-D.~Corsar,
-D.~Garijo,
-S.~Soiland-Reyes,
-S.~Zednik and
-J.~Zhao,
-PROV-O: The PROV Ontology,
-2013,
-W3C Recommendation 30 April 2013.
-\url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b82 ###bbsrt2
-%b95 ###
-\bibitem{doi:10.1016/j.patter.2021.100322}
-%
-\begin{barticle}
-\bauthor{\binits{J.}~\bsnm{Leipzig}},
-\bauthor{\binits{D.}~\bsnm{N\"{u}st}},
-\bauthor{\binits{C.T.}~\bsnm{Hoyt}},
-\bauthor{\binits{K.}~\bsnm{Ram}} and
-\bauthor{\binits{J.}~\bsnm{Greenberg}},
-\batitle{The role of metadata in reproducible computational research},
-\bjtitle{Patterns}
-\bvolume{2}(\bissue{9})
-(\byear{2021}),
-\bnumber{100322}.
-\bid{doi={10.1016/j.patter.2021.100322}}
-\end{barticle}
-%
-\OrigBibText
-J.~Leipzig,
-D.~N\"{u}st,
-C.T.~Hoyt,
-K.~Ram and
-J.~Greenberg,
-\textit{The role of metadata in reproducible computational research},
-Patterns
-2(9)
-(2021),
-100322.
-doi:10.1016/j.patter.2021.100322.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b83 ###bbsrt2
-%b74 ###
-\bibitem{doi:10.48546/workflowhub.workflow.56.1}
-%
-\begin{botherref}
-\oauthor{\binits{D.}~\bsnm{Lowe}} and
-\oauthor{\binits{G.}~\bsnm{Bayarri}},
-Protein Ligand Complex {MD} Setup tutorial using {BioExcel} Building
-Blocks (biobb) (jupyter notebook),
-2021.
-\bid{doi={10.48546/workflowhub.workflow.56.1}}
-\end{botherref}
-%
-\OrigBibText
-D.~Lowe and
-G.~Bayarri,
-Protein Ligand Complex {MD} Setup tutorial using {BioExcel} Building Blocks
-(biobb) (jupyter notebook),
-2021.
-doi:10.48546/workflowhub.workflow.56.1.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b84 ###bbsrt2
-%b44 ###
-\bibitem{ro-crate-excel}
-%
-\begin{botherref}
-\oauthor{\binits{M.}~\bsnm{Lynch}} and
-\oauthor{\binits{P.}~\bsnm{Sefton}},
-npm: ro-crate-excel. \url{https://www.npmjs.com/package/ro-crate-excel}.
-\end{botherref}
-%
-\OrigBibText
-M.~Lynch and
-P.~Sefton,
-npm: ro-crate-excel.
-\url{https://www.npmjs.com/package/ro-crate-excel}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b85 ###bbsrt2
-%b80 ###
-\bibitem{github-lfs}
-%
-\begin{botherref}
-{Managing large files -- {GitHub} Docs}.
-\url
-{https://docs.github.com/en/repositories/working-with-files/managing-large-files}.
-\end{botherref}
-%
-\OrigBibText
-Managing large files - {GitHub} Docs.
-\url
-{https://docs.github.com/en/repositories/working-with-files/managing-large-files}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b86 ###bbsrt2
-%b25 ###
-\bibitem{doi:10.1371/journal.pbio.2001414}
-%
-\begin{barticle}
-\bauthor{\binits{J.A.}~\bsnm{McMurry}},
-\bauthor{\binits{N.}~\bsnm{Juty}},
-\bauthor{\binits{N.}~\bsnm{Blomberg}},
-\bauthor{\binits{T.}~\bsnm{Burdett}},
-\bauthor{\binits{T.}~\bsnm{Conlin}},
-\bauthor{\binits{N.}~\bsnm{Conte}},
-\bauthor{\binits{M.}~\bsnm{Courtot}},
-\bauthor{\binits{J.}~\bsnm{Deck}},
-\bauthor{\binits{M.}~\bsnm{Dumontier}},
-\bauthor{\binits{D.K.}~\bsnm{Fellows}},
-\bauthor{\binits{A.}~\bsnm{Gonzalez-Beltran}},
-\bauthor{\binits{P.}~\bsnm{Gormanns}},
-\bauthor{\binits{J.}~\bsnm{Grethe}},
-\bauthor{\binits{J.}~\bsnm{Hastings}},
-\bauthor{\binits{J.-K.}~\bsnm{H\'{e}rich\'{e}}},
-\bauthor{\binits{H.}~\bsnm{Hermjakob}},
-\bauthor{\binits{J.C.}~\bsnm{Ison}},
-\bauthor{\binits{R.C.}~\bsnm{Jimenez}},
-\bauthor{\binits{S.}~\bsnm{Jupp}},
-\bauthor{\binits{J.}~\bsnm{Kunze}},
-\bauthor{\binits{C.}~\bsnm{Laibe}},
-\bauthor{\binits{N.}~\bsnm{Le Nov\`{e}re}},
-\bauthor{\binits{J.}~\bsnm{Malone}},
-\bauthor{\binits{M.J.}~\bsnm{Martin}},
-\bauthor{\binits{J.R.}~\bsnm{McEntyre}},
-\bauthor{\binits{C.}~\bsnm{Morris}},
-\bauthor{\binits{J.}~\bsnm{Muilu}},
-\bauthor{\binits{W.}~\bsnm{M\"{u}ller}},
-\bauthor{\binits{P.}~\bsnm{Rocca-Serra}},
-\bauthor{\binits{S.-A.}~\bsnm{Sansone}},
-\bauthor{\binits{M.}~\bsnm{Sariyar}},
-\bauthor{\binits{J.L.}~\bsnm{Snoep}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\bauthor{\binits{N.J.}~\bsnm{Stanford}},
-\bauthor{\binits{N.}~\bsnm{Swainston}},
-\bauthor{\binits{N.}~\bsnm{Washington}},
-\bauthor{\binits{A.R.}~\bsnm{Williams}},
-\bauthor{\binits{S.M.}~\bsnm{Wimalaratne}},
-\bauthor{\binits{L.M.}~\bsnm{Winfree}},
-\bauthor{\binits{K.}~\bsnm{Wolstencroft}},
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{C.J.}~\bsnm{Mungall}},
-\bauthor{\binits{M.A.}~\bsnm{Haendel}} and
-\bauthor{\binits{H.}~\bsnm{Parkinson}},
-\batitle{Identifiers for the 21st century: How to design, provision, and
-reuse persistent identifiers to maximize utility and impact of life
-science data},
-\bjtitle{PLOS Biology}
-\bvolume{15}(\bissue{6})
-(\byear{2017}),
-\bnumber{e2001414}.
-\bid{doi={10.1371/journal.pbio.2001414}}
-\end{barticle}
-%
-\OrigBibText
-J.A.~{McMurry},
-N.~Juty,
-N.~Blomberg,
-T.~Burdett,
-T.~Conlin,
-N.~Conte,
-M.~Courtot,
-J.~Deck,
-M.~Dumontier,
-D.K.~Fellows,
-A.~Gonzalez-Beltran,
-P.~Gormanns,
-J.~Grethe,
-J.~Hastings,
-J.-K.~H\'{e}rich\'{e},
-H.~Hermjakob,
-J.C.~Ison,
-R.C.~Jimenez,
-S.~Jupp,
-J.~Kunze,
-C.~Laibe,
-N.~Le~Nov\`{e}re,
-J.~Malone,
-M.J.~Martin,
-J.R.~{McEntyre},
-C.~Morris,
-J.~Muilu,
-W.~M\"{u}ller,
-P.~Rocca-Serra,
-S.-A.~Sansone,
-M.~Sariyar,
-J.L.~Snoep,
-S.~Soiland-Reyes,
-N.J.~Stanford,
-N.~Swainston,
-N.~Washington,
-A.R.~Williams,
-S.M.~Wimalaratne,
-L.M.~Winfree,
-K.~Wolstencroft,
-C.~Goble,
-C.J.~Mungall,
-M.A.~Haendel and
-H.~Parkinson,
-\textit{Identifiers for the 21st century: How to design, provision, and reuse
-persistent identifiers to maximize utility and impact of life science data.},
-{PLOS} Biology
-15(6)
-(2017),
-e2001414.
-doi:10.1371/journal.pbio.2001414.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b87 ###bbsrt2
-%b57 ###
-\bibitem{doi:10.4126/frl01-006423291}
-%
-\begin{bchapter}
-\bauthor{\binits{T.}~\bsnm{Miksa}},
-\bauthor{\binits{M.}~\bsnm{Jaoua}} and
-\bauthor{\binits{G.}~\bsnm{Arfaoui}},
-\bctitle{Research object crates and machine-actionable data management plans},
-in: \bbtitle{1st Workshop on Research Data Management for Linked Open Science},
-\byear{2020}.
-\bid{doi={10.4126/frl01-006423291}}
-\end{bchapter}
-%
-\OrigBibText
-T.~Miksa,
-M.~Jaoua and
-G.~Arfaoui,
-\textit{Research Object Crates and Machine-actionable Data Management Plans},
-in: 1st Workshop on Research Data Management for Linked Open Science,
-2020.
-doi:10.4126/frl01-006423291.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b88 ###bbsrt2
-%b83 ###
-\bibitem{doi:10.1371/journal.pcbi.1006750}
-%
-\begin{barticle}
-\bauthor{\binits{T.}~\bsnm{Miksa}},
-\bauthor{\binits{S.}~\bsnm{Simms}},
-\bauthor{\binits{D.}~\bsnm{Mietchen}} and
-\bauthor{\binits{S.}~\bsnm{Jones}},
-\batitle{Ten principles for machine-actionable data management plans},
-\bjtitle{PLOS Computational Biology}
-\bvolume{15}(\bissue{3})
-(\byear{2019}),
-\bnumber{e1006750}.
-\bid{doi={10.1371/journal.pcbi.1006750}}
-\end{barticle}
-%
-\OrigBibText
-T.~Miksa,
-S.~Simms,
-D.~Mietchen and
-S.~Jones,
-\textit{Ten principles for machine-actionable data management plans},
-{PLOS} Computational Biology
-15(3)
-(2019),
-e1006750.
-doi:10.1371/journal.pcbi.1006750.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b89 ###bbsrt2
-%b121 ###
-\bibitem{doi:10.1186/1471-2105-11-S12-S5}
-%
-\begin{barticle}
-\bauthor{\binits{S.}~\bsnm{M\"{o}ller}},
-\bauthor{\binits{H.N.}~\bsnm{Krabbenh\"{o}ft}},
-\bauthor{\binits{A.}~\bsnm{Tille}},
-\bauthor{\binits{D.}~\bsnm{Paleino}},
-\bauthor{\binits{A.}~\bsnm{Williams}},
-\bauthor{\binits{K.}~\bsnm{Wolstencroft}},
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{R.}~\bsnm{Holland}},
-\bauthor{\binits{D.}~\bsnm{Belhachemi}} and
-\bauthor{\binits{C.}~\bsnm{Plessy}},
-\batitle{Community-driven computational biology with Debian Linux},
-\bjtitle{{BMC} Bioinformatics}
-\bvolume{11}(\bissue{Suppl 12})
-(\byear{2010}),
-\bnumber{S5}.
-\bid{doi={10.1186/1471-2105-11-S12-S5}}
-\end{barticle}
-%
-\OrigBibText
-S.~M\"{o}ller,
-H.N.~Krabbenh\"{o}ft,
-A.~Tille,
-D.~Paleino,
-A.~Williams,
-K.~Wolstencroft,
-C.~Goble,
-R.~Holland,
-D.~Belhachemi and
-C.~Plessy,
-\textit{Community-driven computational biology with Debian Linux.},
-{BMC} Bioinformatics
-11 Suppl 12
-(2010),
-S5.
-doi:10.1186/1471-2105-11-S12-S5.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b90 ###bbsrt2
-%b118 ###
-\bibitem{doi:10.1007/s41019-017-0050-4}
-%
-\begin{barticle}
-\bauthor{\binits{S.}~\bsnm{M\"{o}ller}},
-\bauthor{\binits{S.W.}~\bsnm{Prescott}},
-\bauthor{\binits{L.}~\bsnm{Wirzenius}},
-\bauthor{\binits{P.}~\bsnm{Reinholdtsen}},
-\bauthor{\binits{B.}~\bsnm{Chapman}},
-\bauthor{\binits{P.}~\bsnm{Prins}},
-\bauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\bauthor{\binits{F.}~\bsnm{Kl\"{o}tzl}},
-\bauthor{\binits{A.}~\bsnm{Bagnacani}},
-\bauthor{\binits{M.}~\bsnm{Kala\v{s}}},
-\bauthor{\binits{A.}~\bsnm{Tille}} and
-\bauthor{\binits{M.R.}~\bsnm{Crusoe}},
-\batitle{Robust cross-platform workflows: How technical and scientific
-communities collaborate to develop, test and share best practices for
-data analysis},
-\bjtitle{Data Science and Engineering}
-\bvolume{2}(\bissue{3})
-(\byear{2017}),
-\bfpage{232}--\blpage{244}.
-\bid{doi={10.1007/s41019-017-0050-4}}
-\end{barticle}
-%
-\OrigBibText
-S.~M\"{o}ller,
-S.W.~Prescott,
-L.~Wirzenius,
-P.~Reinholdtsen,
-B.~Chapman,
-P.~Prins,
-S.~Soiland-Reyes,
-F.~Kl\"{o}tzl,
-A.~Bagnacani,
-M.~Kala\v{s},
-A.~Tille and
-M.R.~Crusoe,
-\textit{Robust Cross-Platform Workflows: How Technical and Scientific
-Communities Collaborate to Develop, Test and Share Best Practices for Data
-Analysis},
-Data Science and Engineering
-2(3)
-(2017),
-232--244.
-doi:10.1007/s41019-017-0050-4.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b91 ###bbsrt2
-%b6 ###
-\bibitem{isbn:9781315351148}
-%
-\begin{botherref}
-\oauthor{\binits{B.}~\bsnm{Mons}},
-Data Stewardship for Open Science,
-1st edn,
-Taylor \& Francis,
-p.~240.
-ISBN \bisbn{9781315351148}.
-\end{botherref}
-%
-\OrigBibText
-B.~Mons,
-\textit{Data Stewardship for Open Science},
-1st edn,
-Taylor \& Francis,
-p.~240.
-ISBN 9781315351148.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b92 ###bbsrt2
-%b100 ###
-\bibitem{myExperimentOntology2009}
-%
-\begin{botherref}
-{{myExperiment Ontology Modules}},
-2009.
-\url{http://web.archive.org/web/20091115080336/http\%3a\%2f\%2frdf.myexperiment.org/ontologies}.
-\end{botherref}
-%
-\OrigBibText
-{myExperiment} Ontology Modules,
-2009.
-\url{http://web.archive.org/web/20091115080336/}
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b93 ###bbsrt2
-%b99 ###
-\bibitem{newman2009}
-%
-\begin{bchapter}
-\bauthor{\binits{D.}~\bsnm{Newman}},
-\bauthor{\binits{S.}~\bsnm{Bechhofer}} and
-\bauthor{\binits{D.}~\bsnm{De Roure}},
-\bctitle{{myExperiment}: An ontology for e-Research},
-in: \bbtitle{Proceedings of the Workshop on Semantic Web Applications
-in Scientific
-Discourse (SWASD 2009)},
-\beditor{\binits{T.}~\bsnm{Clark}},
-\beditor{\binits{J.S.}~\bsnm{Luciano}},
-\beditor{\binits{M.S.}~\bsnm{Marshall}},
-\beditor{\binits{E.}~\bsnm{Prud'Hommeaux}} and
-\beditor{\binits{S.}~\bsnm{Stephens}}, eds,
-\bsertitle{CEUR Workshop Proceedings},
-Vols~\bseriesno{523},
-\bpublisher{CEUR-WS},
-\byear{2009}.
-\bcomment{ISSN 1613-0073}. \url{http://ceur-ws.org/Vol-523/Newman.pdf}.
-\end{bchapter}
-%
-\OrigBibText
-D.~Newman,
-S.~Bechhofer and
-D.D.~Roure,
-\textit{{myExperiment}: An ontology for e-Research},
-in: Proceedings of the Workshop on Semantic Web Applications in
-Scientific Discourse (SWASD 2009),
-T.~Clark,
-J.S.~Luciano,
-M.S.~Marshall,
-E.~Prud'Hommeaux and
-S.~Stephens, eds,
-CEUR Workshop Proceedings,
-Vol.~523,
-CEUR-WS,
-2009,
-2009-10-25.
-ISSN 1613-0073.
-\url{http://ceur-ws.org/Vol-523/Newman.pdf}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b94 ###bbsrt2
-%b19 ###
-\bibitem{neylon_blog_post_2017}
-%
-\begin{botherref}
-\oauthor{\binits{C.}~\bsnm{Neylon}},
-As a researcher {\ldots} I'm a bit bloody fed up with Data Management,
-2017.
-\url
-{https://cameronneylon.net/blog/as-a-researcher-im-a-bit-bloody-fed-up-with-data-management/}.
-\end{botherref}
-%
-\OrigBibText
-C.~Neylon,
-As a researcher{\ldots}I'm a bit bloody fed up with Data Management,
-2017.
-\url
-{https://cameronneylon.net/blog/as-a-researcher-im-a-bit-bloody-fed-up-with-data-management/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b95 ###bbsrt2
-% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
-%incorrectly.
-%b38 ###
-\bibitem{ro-crate-html-js}
-%
-\begin{botherref}
-npm: ro-crate-html-js,
-\url{https://www.npmjs.com/package/ro-crate-html-js}.
-\end{botherref}
-%
-\OrigBibText
-npm: ro-crate-html-js.
-\url{https://www.npmjs.com/package/ro-crate-html-js}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b96 ###bbsrt2
-%b27 ###
-\bibitem{ocfl_2020}
-%
-\begin{botherref}
-{OCFL},
-Oxford Common File Layout Specification, Recommendation,
-2020.
-\url{https://ocfl.io/1.0/spec/}.
-\end{botherref}
-%
-\OrigBibText
-{OCFL},
-Oxford Common File Layout Specification,
-Recommendation,
-2020.
-\url{https://ocfl.io/1.0/spec/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b97 ###bbsrt2
-%b61 ###
-\bibitem{doi:10.1080/14490854.2020.1796500}
-%
-\begin{barticle}
-\bauthor{\binits{A.}~\bsnm{Piper}},
-\batitle{Digital crowdsourcing and public understandings of the past:
-Citizen historians meet criminal characters},
-\bjtitle{History Australia}
-\bvolume{17}(\bissue{3})
-(\byear{2020}),
-\bfpage{525}--\blpage{541}.
-\bid{doi={10.1080/14490854.2020.1796500}}
-\end{barticle}
-%
-\OrigBibText
-A.~Piper,
-\textit{Digital crowdsourcing and public understandings of the past: citizen
-historians meet Criminal Characters},
-History Australia
-17(3)
-(2020),
-525--541.
-doi:10.1080/14490854.2020.1796500.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b98 ###bbsrt2
-%b127 ###
-\bibitem{rdfworkinggroup_2014}
-%
-\begin{botherref}
-\binstitute{RDF Working Group},
-{RDF} 1.1 Concepts and Abstract Syntax, {W3C} Recommendation,
-2014.
-\url{https://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/}.
-\end{botherref}
-%
-\OrigBibText
-{RDF Working Group},
-{RDF} 1.1 Concepts and Abstract Syntax,
-{W3C} Recommendation,
-2014.
-\url{https://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b99 ###bbsrt2
-%b41 ###
-\bibitem{doi:10.1016/j.xgen.2021.100029}
-%
-\begin{barticle}
-\bauthor{\binits{H.L.}~\bsnm{Rehm}},
-\bauthor{\binits{A.J.H.}~\bsnm{Page}},
-\bauthor{\binits{L.}~\bsnm{Smith}},
-\bauthor{\binits{J.B.}~\bsnm{Adams}},
-\bauthor{\binits{G.}~\bsnm{Alterovitz}},
-\bauthor{\binits{L.J.}~\bsnm{Babb}},
-\bauthor{\binits{M.P.}~\bsnm{Barkley}},
-\bauthor{\binits{M.}~\bsnm{Baudis}},
-\bauthor{\binits{M.J.S.}~\bsnm{Beauvais}},
-\bauthor{\binits{T.}~\bsnm{Beck}},
-\bauthor{\binits{J.S.}~\bsnm{Beckmann}},
-\bauthor{\binits{S.}~\bsnm{Beltran}},
-\bauthor{\binits{D.}~\bsnm{Bernick}},
-\bauthor{\binits{A.}~\bsnm{Bernier}},
-\bauthor{\binits{J.K.}~\bsnm{Bonfield}},
-\bauthor{\binits{T.F.}~\bsnm{Boughtwood}},
-\bauthor{\binits{G.}~\bsnm{Bourque}},
-\bauthor{\binits{S.R.}~\bsnm{Bowers}},
-\bauthor{\binits{A.J.}~\bsnm{Brookes}},
-\bauthor{\binits{M.}~\bsnm{Brudno}},
-\bauthor{\binits{M.H.}~\bsnm{Brush}},
-\bauthor{\binits{D.}~\bsnm{Bujold}},
-\bauthor{\binits{T.}~\bsnm{Burdett}},
-\bauthor{\binits{O.J.}~\bsnm{Buske}},
-\bauthor{\binits{M.N.}~\bsnm{Cabili}},
-\bauthor{\binits{D.L.}~\bsnm{Cameron}},
-\bauthor{\binits{R.J.}~\bsnm{Carroll}},
-\bauthor{\binits{E.}~\bsnm{Casas-Silva}},
-\bauthor{\binits{D.}~\bsnm{Chakravarty}},
-\bauthor{\binits{B.P.}~\bsnm{Chaudhari}},
-\bauthor{\binits{S.H.}~\bsnm{Chen}},
-\bauthor{\binits{J.M.}~\bsnm{Cherry}},
-\bauthor{\binits{J.}~\bsnm{Chung}},
-\bauthor{\binits{M.}~\bsnm{Cline}},
-\bauthor{\binits{H.L.}~\bsnm{Clissold}},
-\bauthor{\binits{R.M.}~\bsnm{Cook-Deegan}},
-\bauthor{\binits{M.}~\bsnm{Courtot}},
-\bauthor{\binits{F.}~\bsnm{Cunningham}},
-\bauthor{\binits{M.}~\bsnm{Cupak}},
-\bauthor{\binits{R.M.}~\bsnm{Davies}},
-\bauthor{\binits{D.}~\bsnm{Denisko}},
-\bauthor{\binits{M.J.}~\bsnm{Doerr}},
-\bauthor{\binits{L.I.}~\bsnm{Dolman}},
-\bauthor{\binits{E.S.}~\bsnm{Dove}},
-\bauthor{\binits{L.J.}~\bsnm{Dursi}},
-\bauthor{\binits{S.O.M.}~\bsnm{Dyke}},
-\bauthor{\binits{J.A.}~\bsnm{Eddy}},
-\bauthor{\binits{K.}~\bsnm{Eilbeck}},
-\bauthor{\binits{K.P.}~\bsnm{Ellrott}},
-\bauthor{\binits{S.}~\bsnm{Fairley}},
-\bauthor{\binits{K.A.}~\bsnm{Fakhro}},
-\bauthor{\binits{H.V.}~\bsnm{Firth}},
-\bauthor{\binits{M.S.}~\bsnm{Fitzsimons}},
-\bauthor{\binits{M.}~\bsnm{Fiume}},
-\bauthor{\binits{P.}~\bsnm{Flicek}},
-\bauthor{\binits{I.M.}~\bsnm{Fore}},
-\bauthor{\binits{M.A.}~\bsnm{Freeberg}},
-\bauthor{\binits{R.R.}~\bsnm{Freimuth}},
-\bauthor{\binits{L.A.}~\bsnm{Fromont}},
-\bauthor{\binits{J.}~\bsnm{Fuerth}},
-\bauthor{\binits{C.L.}~\bsnm{Gaff}},
-\bauthor{\binits{W.}~\bsnm{Gan}},
-\bauthor{\binits{E.M.}~\bsnm{Ghanaim}},
-\bauthor{\binits{D.}~\bsnm{Glazer}},
-\bauthor{\binits{R.C.}~\bsnm{Green}},
-\bauthor{\binits{M.}~\bsnm{Griffith}},
-\bauthor{\binits{O.L.}~\bsnm{Griffith}},
-\bauthor{\binits{R.L.}~\bsnm{Grossman}},
-\bauthor{\binits{T.}~\bsnm{Groza}},
-\bauthor{\binits{J.M.}~\bsnm{Guidry Auvil}},
-\bauthor{\binits{R.}~\bsnm{Guig\'{o}}},
-\bauthor{\binits{D.}~\bsnm{Gupta}},
-\bauthor{\binits{M.A.}~\bsnm{Haendel}},
-\bauthor{\binits{A.}~\bsnm{Hamosh}},
-\bauthor{\binits{D.P.}~\bsnm{Hansen}},
-\bauthor{\binits{R.K.}~\bsnm{Hart}},
-\bauthor{\binits{D.M.}~\bsnm{Hartley}},
-\bauthor{\binits{D.}~\bsnm{Haussler}},
-\bauthor{\binits{R.M.}~\bsnm{Hendricks-Sturrup}},
-\bauthor{\binits{C.W.L.}~\bsnm{Ho}},
-\bauthor{\binits{A.E.}~\bsnm{Hobb}},
-\bauthor{\binits{M.M.}~\bsnm{Hoffman}},
-\bauthor{\binits{O.M.}~\bsnm{Hofmann}},
-\bauthor{\binits{P.}~\bsnm{Holub}},
-\bauthor{\binits{J.S.}~\bsnm{Hsu}},
-\bauthor{\binits{J.-P.}~\bsnm{Hubaux}},
-\bauthor{\binits{S.E.}~\bsnm{Hunt}},
-\bauthor{\binits{A.}~\bsnm{Husami}},
-\bauthor{\binits{J.O.}~\bsnm{Jacobsen}},
-\bauthor{\binits{S.S.}~\bsnm{Jamuar}},
-\bauthor{\binits{E.L.}~\bsnm{Janes}},
-\bauthor{\binits{F.}~\bsnm{Jeanson}},
-\bauthor{\binits{A.}~\bsnm{Jen\'{e}}},
-\bauthor{\binits{A.L.}~\bsnm{Johns}},
-\bauthor{\binits{Y.}~\bsnm{Joly}},
-\bauthor{\binits{S.J.M.}~\bsnm{Jones}},
-\bauthor{\binits{A.}~\bsnm{Kanitz}},
-\bauthor{\binits{K.}~\bsnm{Kato}},
-\bauthor{\binits{T.M.}~\bsnm{Keane}},
-\bauthor{\binits{K.}~\bsnm{Kekesi-Lafrance}},
-\bauthor{\binits{J.}~\bsnm{Kelleher}},
-\bauthor{\binits{G.}~\bsnm{Kerry}},
-\bauthor{\binits{S.-S.}~\bsnm{Khor}},
-\bauthor{\binits{B.M.}~\bsnm{Knoppers}},
-\bauthor{\binits{M.A.}~\bsnm{Konopko}},
-\bauthor{\binits{K.}~\bsnm{Kosaki}},
-\bauthor{\binits{M.}~\bsnm{Kuba}},
-\bauthor{\binits{J.}~\bsnm{Lawson}},
-\bauthor{\binits{R.}~\bsnm{Leinonen}},
-\bauthor{\binits{S.}~\bsnm{Li}},
-\bauthor{\binits{M.F.}~\bsnm{Lin}},
-\bauthor{\binits{M.}~\bsnm{Linden}},
-\bauthor{\binits{X.}~\bsnm{Liu}},
-\bauthor{\binits{I.U.}~\bsnm{Liyanage}},
-\bauthor{\binits{J.}~\bsnm{Lopez}},
-\bauthor{\binits{A.M.}~\bsnm{Lucassen}},
-\bauthor{\binits{M.}~\bsnm{Lukowski}},
-\bauthor{\binits{A.L.}~\bsnm{Mann}},
-\bauthor{\binits{J.}~\bsnm{Marshall}},
-\bauthor{\binits{M.}~\bsnm{Mattioni}},
-\bauthor{\binits{A.}~\bsnm{Metke-Jimenez}},
-\bauthor{\binits{A.}~\bsnm{Middleton}},
-\bauthor{\binits{R.J.}~\bsnm{Milne}},
-\bauthor{\binits{F.}~\bsnm{Moln\'{a}r-G\'{a}bor}},
-\bauthor{\binits{N.}~\bsnm{Mulder}},
-\bauthor{\binits{M.C.}~\bsnm{Munoz-Torres}},
-\bauthor{\binits{R.}~\bsnm{Nag}},
-\bauthor{\binits{H.}~\bsnm{Nakagawa}},
-\bauthor{\binits{J.}~\bsnm{Nasir}},
-\bauthor{\binits{A.}~\bsnm{Navarro}},
-\bauthor{\binits{T.H.}~\bsnm{Nelson}},
-\bauthor{\binits{A.}~\bsnm{Niewielska}},
-\bauthor{\binits{A.}~\bsnm{Nisselle}},
-\bauthor{\binits{J.}~\bsnm{Niu}},
-\bauthor{\binits{T.H.}~\bsnm{Nyr\"{o}nen}},
-\bauthor{\binits{B.D.}~\bsnm{O'Connor}},
-\bauthor{\binits{S.}~\bsnm{Oesterle}},
-\bauthor{\binits{S.}~\bsnm{Ogishima}},
-\bauthor{\binits{V.}~\bsnm{Ota Wang}},
-\bauthor{\binits{L.A.D.}~\bsnm{Paglione}},
-\bauthor{\binits{E.}~\bsnm{Palumbo}},
-\bauthor{\binits{H.E.}~\bsnm{Parkinson}},
-\bauthor{\binits{A.A.}~\bsnm{Philippakis}},
-\bauthor{\binits{A.D.}~\bsnm{Pizarro}},
-\bauthor{\binits{A.}~\bsnm{Prlic}},
-\bauthor{\binits{J.}~\bsnm{Rambla}},
-\bauthor{\binits{A.}~\bsnm{Rendon}},
-\bauthor{\binits{R.A.}~\bsnm{Rider}},
-\bauthor{\binits{P.N.}~\bsnm{Robinson}},
-\bauthor{\binits{K.W.}~\bsnm{Rodarmer}},
-\bauthor{\binits{L.L.}~\bsnm{Rodriguez}},
-\bauthor{\binits{A.F.}~\bsnm{Rubin}},
-\bauthor{\binits{M.}~\bsnm{Rueda}},
-\bauthor{\binits{G.A.}~\bsnm{Rushton}},
-\bauthor{\binits{R.S.}~\bsnm{Ryan}},
-\bauthor{\binits{G.I.}~\bsnm{Saunders}},
-\bauthor{\binits{H.}~\bsnm{Schuilenburg}},
-\bauthor{\binits{T.}~\bsnm{Schwede}},
-\bauthor{\binits{S.}~\bsnm{Scollen}},
-\bauthor{\binits{A.}~\bsnm{Senf}},
-\bauthor{\binits{N.C.}~\bsnm{Sheffield}},
-\bauthor{\binits{N.}~\bsnm{Skantharajah}},
-\bauthor{\binits{A.V.}~\bsnm{Smith}},
-\bauthor{\binits{H.J.}~\bsnm{Sofia}},
-\bauthor{\binits{D.}~\bsnm{Spalding}},
-\bauthor{\binits{A.B.}~\bsnm{Spurdle}},
-\bauthor{\binits{Z.}~\bsnm{Stark}},
-\bauthor{\binits{L.D.}~\bsnm{Stein}},
-\bauthor{\binits{M.}~\bsnm{Suematsu}},
-\bauthor{\binits{P.}~\bsnm{Tan}},
-\bauthor{\binits{J.A.}~\bsnm{Tedds}},
-\bauthor{\binits{A.A.}~\bsnm{Thomson}},
-\bauthor{\binits{A.}~\bsnm{Thorogood}},
-\bauthor{\binits{T.L.}~\bsnm{Tickle}},
-\bauthor{\binits{K.}~\bsnm{Tokunaga}},
-\bauthor{\binits{J.}~\bsnm{T\"{o}rnroos}},
-\bauthor{\binits{D.}~\bsnm{Torrents}},
-\bauthor{\binits{S.}~\bsnm{Upchurch}},
-\bauthor{\binits{A.}~\bsnm{Valencia}},
-\bauthor{\binits{R.V.}~\bsnm{Guimera}},
-\bauthor{\binits{J.}~\bsnm{Vamathevan}},
-\bauthor{\binits{S.}~\bsnm{Varma}},
-\bauthor{\binits{D.F.}~\bsnm{Vears}},
-\bauthor{\binits{C.}~\bsnm{Viner}},
-\bauthor{\binits{C.}~\bsnm{Voisin}},
-\bauthor{\binits{A.H.}~\bsnm{Wagner}},
-\bauthor{\binits{S.E.}~\bsnm{Wallace}},
-\bauthor{\binits{B.P.}~\bsnm{Walsh}},
-\bauthor{\binits{M.S.}~\bsnm{Williams}},
-\bauthor{\binits{E.C.}~\bsnm{Winkler}},
-\bauthor{\binits{B.J.}~\bsnm{Wold}},
-\bauthor{\binits{G.M.}~\bsnm{Wood}},
-\bauthor{\binits{J.P.}~\bsnm{Woolley}},
-\bauthor{\binits{C.}~\bsnm{Yamasaki}},
-\bauthor{\binits{A.D.}~\bsnm{Yates}},
-\bauthor{\binits{C.K.}~\bsnm{Yung}},
-\bauthor{\binits{L.J.}~\bsnm{Zass}},
-\bauthor{\binits{K.}~\bsnm{Zaytseva}},
-\bauthor{\binits{J.}~\bsnm{Zhang}},
-\bauthor{\binits{P.}~\bsnm{Goodhand}},
-\bauthor{\binits{K.}~\bsnm{North}} and
-\bauthor{\binits{E.}~\bsnm{Birney}},
-\batitle{{GA4GH}: International policies and standards for data sharing
-across genomic research and healthcare},
-\bjtitle{Cell Genomics}
-\bvolume{1}(\bissue{2})
-(\byear{2021}),
-\bnumber{100029}.
-\bid{doi={10.1016/j.xgen.2021.100029}}
-\end{barticle}
-%
-\OrigBibText
-H.L.~Rehm,
-A.J.H.~Page,
-L.~Smith,
-J.B.~Adams,
-G.~Alterovitz,
-L.J.~Babb,
-M.P.~Barkley,
-M.~Baudis,
-M.J.S.~Beauvais,
-T.~Beck,
-J.S.~Beckmann,
-S.~Beltran,
-D.~Bernick,
-A.~Bernier,
-J.K.~Bonfield,
-T.F.~Boughtwood,
-G.~Bourque,
-S.R.~Bowers,
-A.J.~Brookes,
-M.~Brudno,
-M.H.~Brush,
-D.~Bujold,
-T.~Burdett,
-O.J.~Buske,
-M.N.~Cabili,
-D.L.~Cameron,
-R.J.~Carroll,
-E.~Casas-Silva,
-D.~Chakravarty,
-B.P.~Chaudhari,
-S.H.~Chen,
-J.M.~Cherry,
-J.~Chung,
-M.~Cline,
-H.L.~Clissold,
-R.M.~Cook-Deegan,
-M.~Courtot,
-F.~Cunningham,
-M.~Cupak,
-R.M.~Davies,
-D.~Denisko,
-M.J.~Doerr,
-L.I.~Dolman,
-E.S.~Dove,
-L.J.~Dursi,
-S.O.M.~Dyke,
-J.A.~Eddy,
-K.~Eilbeck,
-K.P.~Ellrott,
-S.~Fairley,
-K.A.~Fakhro,
-H.V.~Firth,
-M.S.~Fitzsimons,
-M.~Fiume,
-P.~Flicek,
-I.M.~Fore,
-M.A.~Freeberg,
-R.R.~Freimuth,
-L.A.~Fromont,
-J.~Fuerth,
-C.L.~Gaff,
-W.~Gan,
-E.M.~Ghanaim,
-D.~Glazer,
-R.C.~Green,
-M.~Griffith,
-O.L.~Griffith,
-R.L.~Grossman,
-T.~Groza,
-J.M.~Guidry~Auvil,
-R.~Guig\'{o},
-D.~Gupta,
-M.A.~Haendel,
-A.~Hamosh,
-D.P.~Hansen,
-R.K.~Hart,
-D.M.~Hartley,
-D.~Haussler,
-R.M.~Hendricks-Sturrup,
-C.W.L.~Ho,
-A.E.~Hobb,
-M.M.~Hoffman,
-O.M.~Hofmann,
-P.~Holub,
-J.S.~Hsu,
-J.-P.~Hubaux,
-S.E.~Hunt,
-A.~Husami,
-J.O.~Jacobsen,
-S.S.~Jamuar,
-E.L.~Janes,
-F.~Jeanson,
-A.~Jen\'{e},
-A.L.~Johns,
-Y.~Joly,
-S.J.M.~Jones,
-A.~Kanitz,
-K.~Kato,
-T.M.~Keane,
-K.~Kekesi-Lafrance,
-J.~Kelleher,
-G.~Kerry,
-S.-S.~Khor,
-B.M.~Knoppers,
-M.A.~Konopko,
-K.~Kosaki,
-M.~Kuba,
-J.~Lawson,
-R.~Leinonen,
-S.~Li,
-M.F.~Lin,
-M.~Linden,
-X.~Liu,
-I.U.~Liyanage,
-J.~Lopez,
-A.M.~Lucassen,
-M.~Lukowski,
-A.L.~Mann,
-J.~Marshall,
-M.~Mattioni,
-A.~Metke-Jimenez,
-A.~Middleton,
-R.J.~Milne,
-F.~Moln\'{a}r-G\'{a}bor,
-N.~Mulder,
-M.C.~Munoz-Torres,
-R.~Nag,
-H.~Nakagawa,
-J.~Nasir,
-A.~Navarro,
-T.H.~Nelson,
-A.~Niewielska,
-A.~Nisselle,
-J.~Niu,
-T.H.~Nyr\"{o}nen,
-B.D.~{O\textquoteright Connor},
-S.~Oesterle,
-S.~Ogishima,
-V.~Ota~Wang,
-L.A.D.~Paglione,
-E.~Palumbo,
-H.E.~Parkinson,
-A.A.~Philippakis,
-A.D.~Pizarro,
-A.~Prlic,
-J.~Rambla,
-A.~Rendon,
-R.A.~Rider,
-P.N.~Robinson,
-K.W.~Rodarmer,
-L.L.~Rodriguez,
-A.F.~Rubin,
-M.~Rueda,
-G.A.~Rushton,
-R.S.~Ryan,
-G.I.~Saunders,
-H.~Schuilenburg,
-T.~Schwede,
-S.~Scollen,
-A.~Senf,
-N.C.~Sheffield,
-N.~Skantharajah,
-A.V.~Smith,
-H.J.~Sofia,
-D.~Spalding,
-A.B.~Spurdle,
-Z.~Stark,
-L.D.~Stein,
-M.~Suematsu,
-P.~Tan,
-J.A.~Tedds,
-A.A.~Thomson,
-A.~Thorogood,
-T.L.~Tickle,
-K.~Tokunaga,
-J.~T\"{o}rnroos,
-D.~Torrents,
-S.~Upchurch,
-A.~Valencia,
-R.V.~Guimera,
-J.~Vamathevan,
-S.~Varma,
-D.F.~Vears,
-C.~Viner,
-C.~Voisin,
-A.H.~Wagner,
-S.E.~Wallace,
-B.P.~Walsh,
-M.S.~Williams,
-E.C.~Winkler,
-B.J.~Wold,
-G.M.~Wood,
-J.P.~Woolley,
-C.~Yamasaki,
-A.D.~Yates,
-C.K.~Yung,
-L.J.~Zass,
-K.~Zaytseva,
-J.~Zhang,
-P.~Goodhand,
-K.~North and
-E.~Birney,
-\textit{{GA4GH}: International policies and standards for data sharing across
-genomic research and healthcare},
-Cell Genomics
-1(2)
-(2021),
-100029.
-doi:10.1016/j.xgen.2021.100029.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b100 ###bbsrt2
-%b42 ###
-\bibitem{rettberg_2015_openaire}
-%
-\begin{barticle}
-\bauthor{\binits{N.}~\bsnm{Rettberg}} and
-\bauthor{\binits{B.}~\bsnm{Schmidt}},
-\bjtitle{OpenAIRE, College \& Research Libraries News}
-\bvolume{76}(\bissue{6})
-(\byear{2015}),
-\bfpage{306}--\blpage{310}.
-\url{http://resolver.sub.uni-goettingen.de/purl?gs-1/11942}.
-\bid{doi={10.5860/crln.76.6.9326}}
-\end{barticle}
-%
-\OrigBibText
-N.~Rettberg and
-B.~Schmidt,
-\textit{OpenAIRE},
-College \& Research Libraries News
-76(6)
-(2015),
-306--310.
-\url{http://resolver.sub.uni-goettingen.de/purl?gs-1/11942}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b101 ###bbsrt2
-%b9 ###
-\bibitem{doi:10.1371/journal.pcbi.1003285}
-%
-\begin{barticle}
-\bauthor{\binits{G.K.}~\bsnm{Sandve}},
-\bauthor{\binits{A.}~\bsnm{Nekrutenko}},
-\bauthor{\binits{J.}~\bsnm{Taylor}} and
-\bauthor{\binits{E.}~\bsnm{Hovig}},
-\batitle{Ten simple rules for reproducible computational research},
-\bjtitle{PLOS Computational Biology}
-\bvolume{9}(\bissue{10})
-(\byear{2013}),
-\bnumber{e1003285}.
-\bid{doi={10.1371/journal.pcbi.1003285}}
-\end{barticle}
-%
-\OrigBibText
-G.K.~Sandve,
-A.~Nekrutenko,
-J.~Taylor and
-E.~Hovig,
-\textit{Ten simple rules for reproducible computational research.},
-{PLOS} Computational Biology
-9(10)
-(2013),
-e1003285.
-doi:10.1371/journal.pcbi.1003285.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b102 ###bbsrt2
-%b21 ###
-\bibitem{doi:10.1038/s41597-020-0524-5}
-%
-\begin{barticle}
-\bauthor{\binits{L.M.}~\bsnm{Schriml}},
-\bauthor{\binits{M.}~\bsnm{Chuvochina}},
-\bauthor{\binits{N.}~\bsnm{Davies}},
-\bauthor{\binits{E.A.}~\bsnm{Eloe-Fadrosh}},
-\bauthor{\binits{R.D.}~\bsnm{Finn}},
-\bauthor{\binits{P.}~\bsnm{Hugenholtz}},
-\bauthor{\binits{C.I.}~\bsnm{Hunter}},
-\bauthor{\binits{B.L.}~\bsnm{Hurwitz}},
-\bauthor{\binits{N.C.}~\bsnm{Kyrpides}},
-\bauthor{\binits{F.}~\bsnm{Meyer}},
-\bauthor{\binits{I.K.}~\bsnm{Mizrachi}},
-\bauthor{\binits{S.-A.}~\bsnm{Sansone}},
-\bauthor{\binits{G.}~\bsnm{Sutton}},
-\bauthor{\binits{S.}~\bsnm{Tighe}} and
-\bauthor{\binits{R.}~\bsnm{Walls}},
-\batitle{{COVID}-19 pandemic reveals the peril of ignoring metadata standards},
-\bjtitle{Scientific Data}
-\bvolume{7}(\bissue{1})
-(\byear{2020}),
-\bnumber{188}.
-\bid{doi={10.1038/s41597-020-0524-5}}
-\end{barticle}
-%
-\OrigBibText
-L.M.~Schriml,
-M.~Chuvochina,
-N.~Davies,
-E.A.~Eloe-Fadrosh,
-R.D.~Finn,
-P.~Hugenholtz,
-C.I.~Hunter,
-B.L.~Hurwitz,
-N.C.~Kyrpides,
-F.~Meyer,
-I.K.~Mizrachi,
-S.-A.~Sansone,
-G.~Sutton,
-S.~Tighe and
-R.~Walls,
-\textit{{COVID}-19 pandemic reveals the peril of ignoring metadata
-standards.},
-Scientific data
-7(1)
-(2020),
-188.
-doi:10.1038/s41597-020-0524-5.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b103 ###bbsrt2
-%b39 ###
-\bibitem{doi:10.5281/zenodo.1445817}
-%
-\begin{bchapter}
-\bauthor{\binits{P.}~\bsnm{Sefton}},
-in: \bbtitle{Datacrate Submisssion to the Workshop on Research Objects},
-\bconflocation{Zenodo},
-\byear{2018}.
-\bid{doi={10.5281/zenodo.1445817}}
-\end{bchapter}
-%
-\OrigBibText
-P.~Sefton,
-\textit{Datacrate Submisssion To The Workshop On Research Objects},
-Zenodo
-(2018).
-doi:10.5281/zenodo.1445817.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b104 ###bbsrt2
-%b1 ###
-\bibitem{sefton_blog_post_2021}
-%
-\begin{botherref}
-\oauthor{\binits{P.}~\bsnm{Sefton}},
-{FAIR} Data Management; It's a lifestyle not a lifecycle -- \surl{ptsefton.com},
-2021.
-\url{http://ptsefton.com/2021/04/07/rdmpic/}.
-\end{botherref}
-%
-\OrigBibText
-P.~Sefton,
-{FAIR} Data Management; It's a lifestyle not a lifecycle - ptsefton.com,
-2021.
-\url{http://ptsefton.com/2021/04/07/rdmpic/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b105 ###bbsrt2
-%b62 ###
-\bibitem{doi:10.5281/zenodo.3541888}
-%
-\begin{botherref}
-\oauthor{\binits{P.}~\bsnm{Sefton}},
-\oauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\oauthor{\binits{O.}~\bsnm{Corcho}},
-\oauthor{\binits{D.}~\bsnm{Garijo}},
-\oauthor{\binits{R.}~\bsnm{Palma}},
-\oauthor{\binits{F.}~\bsnm{Coppens}},
-\oauthor{\binits{C.}~\bsnm{Goble}},
-\oauthor{\binits{J.M.}~\bsnm{Fern\'{a}ndez}},
-\oauthor{\binits{K.}~\bsnm{Chard}},
-\oauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
-\oauthor{\binits{M.R.}~\bsnm{Crusoe}},
-\oauthor{\binits{I.}~\bsnm{Eguinoa}},
-\oauthor{\binits{N.}~\bsnm{Juty}},
-\oauthor{\binits{K.}~\bsnm{Holmes}},
-\oauthor{\binits{J.A.}~\bsnm{Clark}},
-\oauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
-\oauthor{\binits{A.J.G.}~\bsnm{Gray}},
-\oauthor{\binits{S.}~\bsnm{Owen}},
-\oauthor{\binits{A.R.}~\bsnm{Williams}},
-\oauthor{\binits{G.}~\bsnm{Tartari}},
-\oauthor{\binits{F.}~\bsnm{Bacall}} and
-\oauthor{\binits{T.}~\bsnm{Thelen}},
-{RO}-Crate Metadata Specification 1.0,
-2019.
-\bid{doi={10.5281/zenodo.3541888}}
-\end{botherref}
-%
-\OrigBibText
-P.~Sefton,
-E.{\'{O}}.~Carrag\'{a}in,
-S.~Soiland-Reyes,
-O.~Corcho,
-D.~Garijo,
-R.~Palma,
-F.~Coppens,
-C.~Goble,
-J.M.~Fern\'{a}ndez,
-K.~Chard,
-J.M.~Gomez-Perez,
-M.R.~Crusoe,
-I.~Eguinoa,
-N.~Juty,
-K.~Holmes,
-J.A.~Clark,
-S.~Capella-Gutierrez,
-A.J.G.~Gray,
-S.~Owen,
-A.R.~Williams,
-G.~Tartari,
-F.~Bacall and
-T.~Thelen,
-{RO}-Crate Metadata Specification 1.0,
-2019.
-doi:10.5281/zenodo.3541888.
-\url{https://w3id.org/ro/crate/1.0}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b106 ###bbsrt2
-%b33 ###
-\bibitem{doi:10.5281/zenodo.4541002}
-%
-\begin{botherref}
-\oauthor{\binits{P.}~\bsnm{Sefton}},
-\oauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\oauthor{\binits{O.}~\bsnm{Corcho}},
-\oauthor{\binits{D.}~\bsnm{Garijo}},
-\oauthor{\binits{R.}~\bsnm{Palma}},
-\oauthor{\binits{F.}~\bsnm{Coppens}},
-\oauthor{\binits{C.}~\bsnm{Goble}},
-\oauthor{\binits{J.M.}~\bsnm{Fern\'{a}ndez}},
-\oauthor{\binits{K.}~\bsnm{Chard}},
-\oauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
-\oauthor{\binits{M.R.}~\bsnm{Crusoe}},
-\oauthor{\binits{I.}~\bsnm{Eguinoa}},
-\oauthor{\binits{N.}~\bsnm{Juty}},
-\oauthor{\binits{K.}~\bsnm{Holmes}},
-\oauthor{\binits{J.A.}~\bsnm{Clark}},
-\oauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
-\oauthor{\binits{A.J.G.}~\bsnm{Gray}},
-\oauthor{\binits{S.}~\bsnm{Owen}},
-\oauthor{\binits{A.R.}~\bsnm{Williams}},
-\oauthor{\binits{G.}~\bsnm{Tartari}},
-\oauthor{\binits{F.}~\bsnm{Bacall}},
-\oauthor{\binits{T.}~\bsnm{Thelen}},
-\oauthor{\binits{H.}~\bsnm{M\'{e}nager}},
-\oauthor{\binits{L.R.-N.}~\bsnm{Navas}},
-\oauthor{\binits{P.}~\bsnm{Walk}},
-\oauthor{\binits{B.}~\bsnm{Whitehead}},
-\oauthor{\binits{M.}~\bsnm{Wilkinson}},
-\oauthor{\binits{P.}~\bsnm{Groth}},
-\oauthor{\binits{E.}~\bsnm{Bremer}},
-\oauthor{\binits{L.G.}~\bsnm{Castro}},
-\oauthor{\binits{K.}~\bsnm{Sebby}},
-\oauthor{\binits{A.}~\bsnm{Kanitz}},
-\oauthor{\binits{A.}~\bsnm{Trisovic}},
-\oauthor{\binits{G.}~\bsnm{Kennedy}},
-\oauthor{\binits{M.}~\bsnm{Graves}},
-\oauthor{\binits{J.}~\bsnm{Koehorst}},
-\oauthor{\binits{S.}~\bsnm{Leo}} and
-\oauthor{\binits{M.}~\bsnm{Portier}},
-{RO}-Crate Metadata Specification 1.1.1,
-2021.
-\bid{doi={10.5281/zenodo.4541002}}
-\end{botherref}
-%
-\OrigBibText
-P.~Sefton,
-E.{\'{O}}.~Carrag\'{a}in,
-S.~Soiland-Reyes,
-O.~Corcho,
-D.~Garijo,
-R.~Palma,
-F.~Coppens,
-C.~Goble,
-J.M.~Fern\'{a}ndez,
-K.~Chard,
-J.M.~Gomez-Perez,
-M.R.~Crusoe,
-I.~Eguinoa,
-N.~Juty,
-K.~Holmes,
-J.A.~Clark,
-S.~Capella-Gutierrez,
-A.J.G.~Gray,
-S.~Owen,
-A.R.~Williams,
-G.~Tartari,
-F.~Bacall,
-T.~Thelen,
-H.~M\'{e}nager,
-L.R.-N.~Navas,
-P.~Walk,
-B.~Whitehead,
-M.~Wilkinson,
-P.~Groth,
-E.~Bremer,
-L.G.~Castro,
-K.~Sebby,
-A.~Kanitz,
-A.~Trisovic,
-G.~Kennedy,
-M.~Graves,
-J.~Koehorst,
-S.~Leo and
-M.~Portier,
-{RO}-Crate Metadata Specification 1.1.1,
-2021.
-doi:10.5281/zenodo.4541002.
-\url{https://w3id.org/ro/crate/1.1}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b107 ###bbsrt2
-%b63 ###
-\bibitem{doi:10.5281/zenodo.4031327}
-%
-\begin{botherref}
-\oauthor{\binits{P.}~\bsnm{Sefton}},
-\oauthor{\binits{E.\'{O}.}~\bsnm{Carrag\'{a}in}},
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\oauthor{\binits{O.}~\bsnm{Corcho}},
-\oauthor{\binits{D.}~\bsnm{Garijo}},
-\oauthor{\binits{R.}~\bsnm{Palma}},
-\oauthor{\binits{F.}~\bsnm{Coppens}},
-\oauthor{\binits{C.}~\bsnm{Goble}},
-\oauthor{\binits{J.M.}~\bsnm{Fern\'{a}ndez}},
-\oauthor{\binits{K.}~\bsnm{Chard}},
-\oauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
-\oauthor{\binits{M.R.}~\bsnm{Crusoe}},
-\oauthor{\binits{I.}~\bsnm{Eguinoa}},
-\oauthor{\binits{N.}~\bsnm{Juty}},
-\oauthor{\binits{K.}~\bsnm{Holmes}},
-\oauthor{\binits{J.A.}~\bsnm{Clark}},
-\oauthor{\binits{S.}~\bsnm{Capella-Gutierrez}},
-\oauthor{\binits{A.J.G.}~\bsnm{Gray}},
-\oauthor{\binits{S.}~\bsnm{Owen}},
-\oauthor{\binits{A.R.}~\bsnm{Williams}},
-\oauthor{\binits{G.}~\bsnm{Tartari}},
-\oauthor{\binits{F.}~\bsnm{Bacall}},
-\oauthor{\binits{T.}~\bsnm{Thelen}},
-\oauthor{\binits{H.}~\bsnm{M\'{e}nager}},
-\oauthor{\binits{L.}~\bsnm{Rodr\'{\i}guez-Navas}},
-\oauthor{\binits{P.}~\bsnm{Walk}},
-\oauthor{\binits{B.}~\bsnm{Whitehead}},
-\oauthor{\binits{M.}~\bsnm{Wilkinson}},
-\oauthor{\binits{P.}~\bsnm{Groth}},
-\oauthor{\binits{E.}~\bsnm{Bremer}},
-\oauthor{\binits{L.G.}~\bsnm{Castro}},
-\oauthor{\binits{K.}~\bsnm{Sebby}},
-\oauthor{\binits{A.}~\bsnm{Kanitz}},
-\oauthor{\binits{A.}~\bsnm{Trisovic}},
-\oauthor{\binits{G.}~\bsnm{Kennedy}},
-\oauthor{\binits{M.}~\bsnm{Graves}},
-\oauthor{\binits{J.}~\bsnm{Koehorst}} and
-\oauthor{\binits{S.}~\bsnm{Leo}},
-{RO}-Crate Metadata Specification 1.1,
-2020.
-\bid{doi={10.5281/zenodo.4031327}}
-\end{botherref}
-%
-\OrigBibText
-P.~Sefton,
-E.{\'{O}}.~Carrag\'{a}in,
-S.~Soiland-Reyes,
-O.~Corcho,
-D.~Garijo,
-R.~Palma,
-F.~Coppens,
-C.~Goble,
-J.M.~Fern\'{a}ndez,
-K.~Chard,
-J.M.~Gomez-Perez,
-M.R.~Crusoe,
-I.~Eguinoa,
-N.~Juty,
-K.~Holmes,
-J.A.~Clark,
-S.~Capella-Gutierrez,
-A.J.G.~Gray,
-S.~Owen,
-A.R.~Williams,
-G.~Tartari,
-F.~Bacall,
-T.~Thelen,
-H.~M\'{e}nager,
-L.~Rodr\'iguez-Navas,
-P.~Walk,
-B.~Whitehead,
-M.~Wilkinson,
-P.~Groth,
-E.~Bremer,
-L.G.~Castro,
-K.~Sebby,
-A.~Kanitz,
-A.~Trisovic,
-G.~Kennedy,
-M.~Graves,
-J.~Koehorst and
-S.~Leo,
-{RO}-Crate Metadata Specification 1.1,
-2020.
-doi:10.5281/zenodo.4031327.
-\url{https://w3id.org/ro/crate/1.1}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b108 ###bbsrt2
-%b114 ###
-\bibitem{soilandreyes_tweet_2020}
-%
-\begin{botherref}
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-I am looking for which bioinformatics journals encourage authors
-to submit their code/pipeline/workflow supporting data analysis,
-2020.
-\url{https://twitter.com/soilandreyes/status/1250721245622079488}.
-\end{botherref}
-%
-\OrigBibText
-S.~Soiland-Reyes,
-I am looking for which bioinformatics journals encourage authors to submit
-their code/pipeline/workflow supporting data analysis,
-2020.
-\url{https://twitter.com/soilandreyes/status/1250721245622079488}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b109 ###bbsrt2
-%b79 ###
-\bibitem{doi:10.5281/zenodo.4633732}
-%
-\begin{botherref}
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-{Describing and packaging workflows using RO-Crate and BioCompute
-Objects}, Zenodo,
-2021,
-Webinar for U.S. Food and Drug Administration (FDA), 2021-05-12.
-\bid{doi={10.5281/zenodo.4633732}}
-\end{botherref}
-%
-\OrigBibText
-S.~Soiland-Reyes,
-{Describing and packaging workflows using RO-Crate and BioCompute Objects},
-Zenodo,
-2021,
-{Webinar for U.S. Food and Drug Administration (FDA), 2021-05-12}.
-doi:10.5281/zenodo.4633732.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b110 ###bbsrt2
-%b71 ###
-\bibitem{doi:10.5281/zenodo.51314}
-%
-\begin{botherref}
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\oauthor{\binits{P.}~\bsnm{Alper}} and
-\oauthor{\binits{C.}~\bsnm{Goble}},
-Tracking Workflow Execution With {TavernaPROV},
-{ProvenanceWeek} 2016, 2016, PROV: Three Years Later.
-\bid{doi={10.5281/zenodo.51314}}
-\end{botherref}
-%
-\OrigBibText
-S.~Soiland-Reyes,
-P.~Alper and
-C.~Goble,
-\textit{Tracking Workflow Execution With {TavernaPROV}},
-in: {ProvenanceWeek} 2016,
-2016,
-PROV: Three Years Later.
-doi:10.5281/zenodo.51314.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b111 ###bbsrt2
-%b76 ###
-\bibitem{doi:10.5281/zenodo.12586}
-%
-\begin{botherref}
-\oauthor{\binits{S.}~\bsnm{Soiland-Reyes}},
-\oauthor{\binits{M.}~\bsnm{Gamble}} and
-\oauthor{\binits{R.}~\bsnm{Haines}},
-Research Object Bundle 1.0
-(2014).
-See \url{https://w3id.org/bundle/2014-11-05/}.
-doi: \doiurl{10.5281/zenodo.12586}.
-\end{botherref}
-%
-\OrigBibText
-S.~Soiland-Reyes,
-M.~Gamble and
-R.~Haines,
-\textit{Research Object Bundle 1.0},
-2014,
-See https://w3id.org/bundle/2014-11-05/.
-doi:10.5281/zenodo.12586.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b112 ###bbsrt2
-%b32 ###
-\bibitem{sporny_2014}
-%
-\begin{botherref}
-\oauthor{\binits{M.}~\bsnm{Sporny}},
-\oauthor{\binits{D.}~\bsnm{Longley}},
-\oauthor{\binits{G.}~\bsnm{Kellogg}},
-\oauthor{\binits{M.}~\bsnm{Lanthaler}} and
-\oauthor{\binits{N.}~\bsnm{Lindstr\"{o}m}},
-{JSON}-{LD} 1.0, {W3C} Recommendation, 2014. \url
-{https://www.w3.org/TR/2014/REC-json-ld-20140116/}.
-\end{botherref}
-%
-\OrigBibText
-M.~Sporny,
-D.~Longley,
-G.~Kellogg,
-M.~Lanthaler and
-N.~Lindstr\"{o}m,
-{JSON}-{LD} 1.0,
-{W3C} Recommendation,
-2014.
-\url{https://www.w3.org/TR/2014/REC-json-ld-20140116/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b113 ###bbsrt2
-%b2 ###
-\bibitem{doi:10.1126/science.aah6168}
-%
-\begin{barticle}
-\bauthor{\binits{V.}~\bsnm{Stodden}},
-\bauthor{\binits{M.}~\bsnm{McNutt}},
-\bauthor{\binits{D.H.}~\bsnm{Bailey}},
-\bauthor{\binits{E.}~\bsnm{Deelman}},
-\bauthor{\binits{Y.}~\bsnm{Gil}},
-\bauthor{\binits{B.}~\bsnm{Hanson}},
-\bauthor{\binits{M.A.}~\bsnm{Heroux}},
-\bauthor{\binits{J.P.A.}~\bsnm{Ioannidis}} and
-\bauthor{\binits{M.}~\bsnm{Taufer}},
-\batitle{Enhancing reproducibility for computational methods},
-\bjtitle{Science}
-\bvolume{354}(\bissue{6317})
-(\byear{2016}),
-\bfpage{1240}--\blpage{1241}.
-\bid{doi={10.1126/science.aah6168}}
-\end{barticle}
-%
-\OrigBibText
-V.~Stodden,
-M.~{McNutt},
-D.H.~Bailey,
-E.~Deelman,
-Y.~Gil,
-B.~Hanson,
-M.A.~Heroux,
-J.P.A.~Ioannidis and
-M.~Taufer,
-\textit{Enhancing reproducibility for computational methods},
-Science
-354(6317)
-(2016),
-1240--1241.
-doi:10.1126/science.aah6168.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b114 ###bbsrt2
-%b82 ###
-\bibitem{doi:10125/4567}
-%
-\begin{bchapter}
-\bauthor{\binits{N.}~\bsnm{Thieberger}} and
-\bauthor{\binits{L.}~\bsnm{Barwick}},
-\bctitle{Keeping records of language diversity in melanesia: The Pacific
-and regional archive for digital sources in endangered cultures ({PARADISEC})},
-in: \bbtitle{Melanesian Languages on the Edge of Asia: Challenges for
-the 21st Century},
-\beditor{\binits{N.}~\bsnm{Evans}} and
-\beditor{\binits{M.}~\bsnm{Klamer}}, eds,
-\bsertitle{Language Documentation \& Conservation Special Publication},
-Vol.~\bseriesno{SP05},
-\bpublisher{University of Hawai'i Press},
-\byear{2012},
-pp.~\bfpage{239}--\blpage{253}.
-ISBN \bisbn{978-0-9856211-2-4}.
-\bid{doi={10125/4567}}
-\end{bchapter}
-%
-\OrigBibText
-N.~Thieberger and
-L.~Barwick,
-\textit{Keeping records of language diversity in Melanesia: The Pacific and
-Regional Archive for Digital Sources in Endangered Cultures ({PARADISEC})},
-in: Melanesian Languages on the Edge of Asia: Challenges for the 21st
-Century,
-N.~Evans and
-M.~Klamer, eds,
-Language Documentation \& Conservation Special Publication,
-Vol.~SP05,
-University of Hawai'i Press,
-2012,
-pp.~239--253.
-ISBN 978-0-9856211-2-4.
-doi:10125/4567.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b115 ###bbsrt2
-%b53 ###
-\bibitem{arkisto-data-portal}
-%
-\begin{botherref}
-{Tools: Data Portal \& Discovery}.
-\url{https://arkisto-platform.github.io/tools/portal/}.
-\end{botherref}
-%
-\OrigBibText
-Tools: Data Portal \& Discovery.
-\url{https://arkisto-platform.github.io/tools/portal/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b116 ###bbsrt2
-%b123 ###
-\bibitem{doi:10.1007/s11042-009-0397-2}
-%
-\begin{barticle}
-\bauthor{\binits{R.}~\bsnm{Troncy}},
-\bauthor{\binits{W.}~\bsnm{Bailer}},
-\bauthor{\binits{M.}~\bsnm{H\"{o}ffernig}} and
-\bauthor{\binits{M.}~\bsnm{Hausenblas}},
-\batitle{{VAMP}: A service for validating {MPEG}-7 descriptions w.r.t.
-to formal profile definitions},
-\bjtitle{Multimedia tools and applications}
-\bvolume{46}(\bissue{2--3})
-(\byear{2010}),
-\bfpage{307}--\blpage{329}.
-\url{https://www.persistent-identifier.nl/urn:nbn:nl:ui:18-14511}.
-\bid{doi={10.1007/s11042-009-0397-2}}
-\end{barticle}
-%
-\OrigBibText
-R.~Troncy,
-W.~Bailer,
-M.~H\"{o}ffernig and
-M.~Hausenblas,
-\textit{{VAMP}: a service for validating {MPEG}-7 descriptions w.r.t. to
-formal profile definitions},
-Multimedia tools and applications
-46(2--3)
-(2010),
-307--329.
-doi:10.1007/s11042-009-0397-2.
-\url{https://www.persistent-identifier.nl/urn:nbn:nl:ui:18-14511}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b117 ###bbsrt2
-%b97 ###
-\bibitem{vandesompel_2007}
-%
-\begin{barticle}
-\bauthor{\binits{H.}~\bsnm{Van de Sompel}} and
-\bauthor{\binits{C.}~\bsnm{Lagoze}},
-\batitle{Interoperability for the discovery, use, and re-use of units
-of scholarly communication},
-\bjtitle{{CTWatch} Quarterly}
-\bvolume{3}(\bissue{3})
-(\byear{2007}).
-\url
-{http://icl.utk.edu/ctwatch/quarterly/articles/2007/08/interoperability-for-the-discovery-use-and-re-use-of-units-of-scholarly-communication/}.
-\end{barticle}
-%
-\OrigBibText
-H.~Van~de Sompel and
-C.~Lagoze,
-Interoperability for the Discovery, Use, and Re-Use of Units of Scholarly
-Communication,
-\textit{{CTWatch} Quarterly}
-\textbf{3}(3)
-(2007)
-00.
-\url
-{http://icl.utk.edu/ctwatch/quarterly/articles/2007/08/interoperability-for-the-discovery-use-and-re-use-of-units-of-scholarly-communication/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b118 ###bbsrt2
-%b50 ###
-\bibitem{vergoulis2021schema}
-%
-\begin{bbook}
-\bauthor{\binits{T.}~\bsnm{Vergoulis}},
-\bauthor{\binits{K.}~\bsnm{Zagganas}},
-\bauthor{\binits{L.}~\bsnm{Kavouras}},
-\bauthor{\binits{M.}~\bsnm{Reczko}},
-\bauthor{\binits{S.}~\bsnm{Sartzetakis}} and
-\bauthor{\binits{T.}~\bsnm{Dalamagas}},
-\bbtitle{{SCHeMa}: Scheduling Scientific Containers on a Cluster of
-Heterogeneous Machines},
-\byear{2021}.
-\url{https://arxiv.org/abs/2103.13138v1}.
-\end{bbook}
-%
-\OrigBibText
-T.~Vergoulis,
-K.~Zagganas,
-L.~Kavouras,
-M.~Reczko,
-S.~Sartzetakis and
-T.~Dalamagas,
-{SCHeMa}: Scheduling Scientific Containers on a Cluster of Heterogeneous
-Machines,
-\textit{arXiv}
-(2021),
-2103.13138--.
-\url{https://arxiv.org/abs/2103.13138v1}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b119 ###bbsrt2
-%b20 ###
-\bibitem{doi:10.1007/s00267-014-0258-2}
-%
-\begin{barticle}
-\bauthor{\binits{C.J.}~\bsnm{Volk}},
-\bauthor{\binits{Y.}~\bsnm{Lucero}} and
-\bauthor{\binits{K.}~\bsnm{Barnas}},
-\batitle{Why is data sharing in collaborative natural resource efforts
-so hard and what can we do to improve it?},
-\bjtitle{Environmental Management}
-\bvolume{53}(\bissue{5})
-(\byear{2014}),
-\bfpage{883}--\blpage{893}.
-\bid{doi={10.1007/s00267-014-0258-2}}
-\end{barticle}
-%
-\OrigBibText
-C.J.~Volk,
-Y.~Lucero and
-K.~Barnas,
-\textit{Why is data sharing in collaborative natural resource efforts
-so hard
-and what can we do to improve it?},
-Environmental Management
-53(5)
-(2014),
-883--893.
-doi:10.1007/s00267-014-0258-2.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b120 ###bbsrt2
-%b31 ###
-\bibitem{httprange14}
-%
-\begin{botherref}
-{W3C Technical Architecture Group}, Dereferencing {HTTP}
-{URIs}, Draft Tag Finding,
-2007.
-\url{https://www.w3.org/2001/tag/doc/httpRange-14/2007-08-31/HttpRange-14.html}.
-\end{botherref}
-%
-\OrigBibText
-{W3C Technical Architecture Group},
-Dereferencing {HTTP} {URIs},
-Draft Tag Finding,
-2007.
-\url{https://www.w3.org/2001/tag/doc/httpRange-14/2007-08-31/HttpRange-14.html}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b121 ###bbsrt2
-%b85 ###
-\bibitem{doi:10.15497/rda00039}
-%
-\begin{bbook}
-\bauthor{\binits{P.}~\bsnm{Walk}},
-\bauthor{\binits{T.}~\bsnm{Miksa}} and
-\bauthor{\binits{P.}~\bsnm{Neish}},
-\bbtitle{{RDA} {DMP} Common Standard for Machine-Actionable Data Management
-Plans, Research Data Alliance},
-\byear{2019}.
-\bid{doi={10.15497/rda00039}}
-\end{bbook}
-%
-\OrigBibText
-P.~Walk,
-T.~Miksa and
-P.~Neish,
-\textit{{RDA} {DMP} Common Standard for Machine-actionable Data Management
-Plans},
-Research Data Alliance
-(2019).
-doi:10.15497/rda00039.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b122 ###bbsrt2
-%b72 ###
-\bibitem{doi:10.3897/rio.6.e57602}
-%
-\begin{barticle}
-\bauthor{\binits{S.}~\bsnm{Walton}},
-\bauthor{\binits{L.}~\bsnm{Livermore}},
-\bauthor{\binits{O.}~\bsnm{B\'{a}nki}},
-\bauthor{\binits{R.}~\bsnm{Cubey}},
-\bauthor{\binits{R.}~\bsnm{Drinkwater}},
-\bauthor{\binits{M.}~\bsnm{Englund}},
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{Q.}~\bsnm{Groom}},
-\bauthor{\binits{C.}~\bsnm{Kermorvant}},
-\bauthor{\binits{I.}~\bsnm{Rey}},
-\bauthor{\binits{C.}~\bsnm{Santos}},
-\bauthor{\binits{B.}~\bsnm{Scott}},
-\bauthor{\binits{A.}~\bsnm{Williams}} and
-\bauthor{\binits{Z.}~\bsnm{Wu}},
-\batitle{Landscape analysis for the specimen data refinery},
-\bjtitle{Research Ideas and Outcomes}
-\bvolume{6}
-(\byear{2020}).
-\bid{doi={10.3897/rio.6.e57602}}
-\end{barticle}
-%
-\OrigBibText
-S.~Walton,
-L.~Livermore,
-O.~B\'{a}nki,
-R.~Cubey,
-R.~Drinkwater,
-M.~Englund,
-C.~Goble,
-Q.~Groom,
-C.~Kermorvant,
-I.~Rey,
-C.~Santos,
-B.~Scott,
-A.~Williams and
-Z.~Wu,
-Landscape analysis for the Specimen Data Refinery,
-\textit{Research Ideas and Outcomes}
-\textbf{6}
-(2020)
-00.
-doi:10.3897/rio.6.e57602.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b123 ###bbsrt2
-%b5 ###
-\bibitem{doi:10.1038/sdata.2016.18}
-%
-\begin{barticle}
-\bauthor{\binits{M.D.}~\bsnm{Wilkinson}},
-\bauthor{\binits{M.}~\bsnm{Dumontier}},
-\bauthor{\binits{I.J.J.}~\bsnm{Aalbersberg}},
-\bauthor{\binits{G.}~\bsnm{Appleton}},
-\bauthor{\binits{M.}~\bsnm{Axton}},
-\bauthor{\binits{A.}~\bsnm{Baak}},
-\bauthor{\binits{N.}~\bsnm{Blomberg}},
-\bauthor{\binits{J.-W.}~\bsnm{Boiten}},
-\bauthor{\binits{L.B.}~\bsnm{da~Silva Santos}},
-\bauthor{\binits{P.E.}~\bsnm{Bourne}},
-\bauthor{\binits{J.}~\bsnm{Bouwman}},
-\bauthor{\binits{A.J.}~\bsnm{Brookes}},
-\bauthor{\binits{T.}~\bsnm{Clark}},
-\bauthor{\binits{M.}~\bsnm{Crosas}},
-\bauthor{\binits{I.}~\bsnm{Dillo}},
-\bauthor{\binits{O.}~\bsnm{Dumon}},
-\bauthor{\binits{S.}~\bsnm{Edmunds}},
-\bauthor{\binits{C.T.}~\bsnm{Evelo}},
-\bauthor{\binits{R.}~\bsnm{Finkers}},
-\bauthor{\binits{A.}~\bsnm{Gonzalez-Beltran}},
-\bauthor{\binits{A.J.G.}~\bsnm{Gray}},
-\bauthor{\binits{P.}~\bsnm{Groth}},
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bauthor{\binits{J.S.}~\bsnm{Grethe}},
-\bauthor{\binits{J.}~\bsnm{Heringa}},
-\bauthor{\binits{P.A.C.}~\bsnm{'t~Hoen}},
-\bauthor{\binits{R.}~\bsnm{Hooft}},
-\bauthor{\binits{T.}~\bsnm{Kuhn}},
-\bauthor{\binits{R.}~\bsnm{Kok}},
-\bauthor{\binits{J.}~\bsnm{Kok}},
-\bauthor{\binits{S.J.}~\bsnm{Lusher}},
-\bauthor{\binits{M.E.}~\bsnm{Martone}},
-\bauthor{\binits{A.}~\bsnm{Mons}},
-\bauthor{\binits{A.L.}~\bsnm{Packer}},
-\bauthor{\binits{B.}~\bsnm{Persson}},
-\bauthor{\binits{P.}~\bsnm{Rocca-Serra}},
-\bauthor{\binits{M.}~\bsnm{Roos}},
-\bauthor{\binits{R.}~\bsnm{van Schaik}},
-\bauthor{\binits{S.-A.}~\bsnm{Sansone}},
-\bauthor{\binits{E.}~\bsnm{Schultes}},
-\bauthor{\binits{T.}~\bsnm{Sengstag}},
-\bauthor{\binits{T.}~\bsnm{Slater}},
-\bauthor{\binits{G.}~\bsnm{Strawn}},
-\bauthor{\binits{M.A.}~\bsnm{Swertz}},
-\bauthor{\binits{M.}~\bsnm{Thompson}},
-\bauthor{\binits{J.}~\bsnm{van der Lei}},
-\bauthor{\binits{E.}~\bsnm{van Mulligen}},
-\bauthor{\binits{J.}~\bsnm{Velterop}},
-\bauthor{\binits{A.}~\bsnm{Waagmeester}},
-\bauthor{\binits{P.}~\bsnm{Wittenburg}},
-\bauthor{\binits{K.}~\bsnm{Wolstencroft}},
-\bauthor{\binits{J.}~\bsnm{Zhao}} and
-\bauthor{\binits{B.}~\bsnm{Mons}},
-\batitle{The {FAIR} guiding principles for scientific data management
-and stewardship},
-\bjtitle{Scientific Data}
-\bvolume{3}
-(\byear{2016}),
-\bnumber{160018}.
-\bid{doi={10.1038/sdata.2016.18}}
-\end{barticle}
-%
-\OrigBibText
-M.D.~Wilkinson,
-M.~Dumontier,
-I.J.J.~Aalbersberg,
-G.~Appleton,
-M.~Axton,
-A.~Baak,
-N.~Blomberg,
-J.-W.~Boiten,
-L.B.~da~Silva~Santos,
-P.E.~Bourne,
-J.~Bouwman,
-A.J.~Brookes,
-T.~Clark,
-M.~Crosas,
-I.~Dillo,
-O.~Dumon,
-S.~Edmunds,
-C.T.~Evelo,
-R.~Finkers,
-A.~Gonzalez-Beltran,
-A.J.G.~Gray,
-P.~Groth,
-C.~Goble,
-J.S.~Grethe,
-J.~Heringa,
-P.A.C.~'t~Hoen,
-R.~Hooft,
-T.~Kuhn,
-R.~Kok,
-J.~Kok,
-S.J.~Lusher,
-M.E.~Martone,
-A.~Mons,
-A.L.~Packer,
-B.~Persson,
-P.~Rocca-Serra,
-M.~Roos,
-R.~van Schaik,
-S.-A.~Sansone,
-E.~Schultes,
-T.~Sengstag,
-T.~Slater,
-G.~Strawn,
-M.A.~Swertz,
-M.~Thompson,
-J.~van~der Lei,
-E.~van Mulligen,
-J.~Velterop,
-A.~Waagmeester,
-P.~Wittenburg,
-K.~Wolstencroft,
-J.~Zhao and
-B.~Mons,
-\textit{The {FAIR} Guiding Principles for scientific data management and
-stewardship.},
-Scientific data
-3
-(2016),
-160018.
-doi:10.1038/sdata.2016.18.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b124 ###bbsrt2
-% NO SURNAME/INITIALS TAGS! Surnames and initials may be identified
-%incorrectly.
-%b48 ###
-\bibitem{about-workflowhub}
-%
-\begin{botherref}
-{WorkflowHub} project | Project pages for developing and
-running the {WorkflowHub}, a registry of scientific workflows.
-\url{https://w3id.org/workflowhub/}.
-\end{botherref}
-%
-\OrigBibText
-{WorkflowHub} project \textbar Project pages for developing and running the
-{WorkflowHub}, a registry of scientific workflows.
-\url{https://w3id.org/workflowhub/}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b125 ###bbsrt2
-%b75 ###
-\bibitem{doi:10.1109/eScience.2012.6404482}
-%
-\begin{bchapter}
-\bauthor{\binits{J.}~\bsnm{Zhao}},
-\bauthor{\binits{J.M.}~\bsnm{Gomez-Perez}},
-\bauthor{\binits{K.}~\bsnm{Belhajjame}},
-\bauthor{\binits{G.}~\bsnm{Klyne}},
-\bauthor{\binits{E.}~\bsnm{Garcia-Cuesta}},
-\bauthor{\binits{A.}~\bsnm{Garrido}},
-\bauthor{\binits{K.}~\bsnm{Hettne}},
-\bauthor{\binits{M.}~\bsnm{Roos}},
-\bauthor{\binits{D.}~\bsnm{De Roure}} and
-\bauthor{\binits{C.}~\bsnm{Goble}},
-\bctitle{Why workflows break -- understanding and combating decay in
-taverna workflows},
-in: \bbtitle{2012 {IEEE} 8th International Conference on e-Science},
-\bpublisher{IEEE},
-\byear{2012},
-pp.~\bfpage{1}--\blpage{9}.
-\url{https://www.research.manchester.ac.uk/portal/files/174861334/why\_
-decay.pdf}.
-ISBN \bisbn{978-1-4673-4466-1}.
-\bid{doi={10.1109/eScience.2012.6404482}}
-\end{bchapter}
-%
-\OrigBibText
-J.~Zhao,
-J.M.~Gomez-Perez,
-K.~Belhajjame,
-G.~Klyne,
-E.~Garcia-Cuesta,
-A.~Garrido,
-K.~Hettne,
-M.~Roos,
-D.~De~Roure and
-C.~Goble,
-\textit{Why workflows break -- Understanding and combating decay in Taverna
-workflows},
-in: 2012 {IEEE} 8th International Conference on E-Science,
-IEEE,
-2012,
-pp.~1--9.
-ISBN 978-1-4673-4466-1.
-doi:10.1109/eScience.2012.6404482.
-\url
-{https://www.research.manchester.ac.uk/portal/files/174861334/why_decay.pdf}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b126 ###bbsrt2
-%b60 ###
-\bibitem{doi:10.5281/zenodo.5068950}
-%
-\begin{botherref}
-\oauthor{\binits{F.}~\bsnm{Zoubek}} and
-\oauthor{\binits{M.}~\bsnm{Winkler}},
-RO Crates and Excel
-(2021).
-\url{https://github.com/e11938258/RO-Crates-and-Excel}.
-\bid{doi={10.5281/zenodo.5068950}}
-\end{botherref}
-%
-\OrigBibText
-F.~Zoubek and
-M.~Winkler,
-RO Crates and Excel,
-Zenodo,
-2021.
-doi:10.5281/zenodo.5068950.
-\url{https://github.com/e11938258/RO-Crates-and-Excel}.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-%b127 ###bbsrt2
-%b18 ###
-\bibitem{doi:10.1515/9783598441844}
-%
-\begin{bbook}
-\bauthor{\binits{M.}~\bsnm{\v{Z}umer}},
-\bbtitle{National Bibliographies in the Digital Age: Guidance and New
-Directions, {IFLA} Series on Bibliographic Control},
-\bsertitle{IFLA Working Group on Guidelines for National Bibliographies},
-\bpublisher{Walter de Gruyter -- K. G. Saur},
-\byear{2009},
-\bcomment{ISSN 1868-8438}.
-ISBN \bisbn{9783598441844}.
-\end{bbook}
-%
-\OrigBibText
-M.~\v{Z}umer,
-\textit{National bibliographies in the digital age: guidance and new
-directions},
-{IFLA} series on bibliographic control,
-Walter de Gruyter -- K. G. Saur,
-2009,
-{IFLA Working Group on Guidelines for National Bibliographies}.
-ISSN 1868-8438.
-ISBN 9783598441844.
-doi:10.1515/9783598441844.
-\endOrigBibText
-\bptok{structpyb}
-\endbibitem
-
-\end{thebibliography}
+\markdownInput{../content/20a.methodology.md}
+\markdownInput{../content/21.conceptual.md}
+\input{../content/21a.conceptual-img.tex}
+\markdownInput{../content/21b.foundation.md}
+\markdownInput{../content/21c.container.md}
+\markdownInput{../content/21d.contextual-entities.md}
+\input{../content/21e.contextual-entities-img.tex}
+\markdownInput{../content/21f.guiding-practices.md}
+\markdownInput{../content/21g.simplicity.md}
+\markdownInput{../content/21h.extensibility.md}
+
+\markdownInput{../content/22.implementation.md}
+\markdownInput{../content/23.community.md}
+
+\input{../content/31.tooling-table.tex}
+\markdownInput{../content/30.tooling.md}
+
+\markdownInput{../content/40.inuse.md}
+\markdownInput{../content/41.workflows.md}
+\markdownInput{../content/42.regulatory.md}
+\input{../content/42a.regulatory-img.tex}
+\markdownInput{../content/43.culturalheritage.md}
+\markdownInput{../content/44.dmp.md}
+\markdownInput{../content/45.institutional-repos.md}
+\input{../content/46.institutional-repos-img.tex}
+
+\markdownInput{../content/50.relatedwork.md}
+
+\markdownInput{../content/60.conclusion.md}
+\markdownInput{../content/61.futurework.md}
+
+\markdownInput{../content/70.acknowledgements.md}
+
+%%%%%%%%%%% The bibliography starts:
+\nocite{*} 
+\bibliographystyle{ios1}
+\bibliography{ro-crate}
+
+%%%% Appendix
+
+\newpage
+\appendix
+%\markdownInput{../content/80.formal-definition.md}
+\input{../content/80.formal-definition}
+
+\newpage
+\markdownInput{../content/89.community.md}
+
 \end{document}
-%

### 2022-01-06 response-to-review.md
Updates from Overleaf
response to review as submitted

@@
 
-Response to  open peer review at <https://datasciencehub.net/paper/packaging-research-artefacts-ro-crate>
+Response to  open peer review at <https://datasciencehub.net/paper/packaging-research-artefacts-ro-crate> of “Packaging research artefacts with RO-Crate” submitted to Data Science._
 
-## Reviewer 1
+We thank the reviewers for their attentive and detailed comments, and we very much appreciate the time and care they have given that has helped us improve the article.
 
-_By Angelo A. <https://orcid.org/0000-0002-4763-3943>_
+## Response to Reviewer 1
 
->  I will leave the latter to its future users in defining how it can be improved, based on their needs.
+_Open Review was by Angelo A. <https://orcid.org/0000-0002-4763-3943>_
+
+We thank Angelo A for the comprehensive comments that we respond to below.
+
+### Summary of changes
+
+The reviewer raised important points on ensuring RO-Crate targets its aim of simplicity.
+This prompted us to include an explicit description of our user requirement and feedback methodology, with references to specific sections.
+
+We have reflected this discussion in the following places:
+
+* A new [section 2.1 Development Methodology](https://stain.github.io/ro-crate-paper/#methodology) that explicitly describe this requirement/feedback methodology in the article
+* In section [3 RO-Crate Tooling](https://stain.github.io/ro-crate-paper/#tooling), by clarifying that the role of the tools aim to make creating and managing RO-Crates simple.
+* A new section [7 Future Work](https://stain.github.io/ro-crate-paper/#futurework) that outlines some new directions and plans for RO-Crate 1.2 and a future user study linked to EOSC.
+
+In addition, the reviewer has helped us improve the article with:
+
+1.
+Language/typography fixes in [section 2](https://stain.github.io/ro-crate-paper/#rocrate), in [Figure 1](https://stain.github.io/ro-crate-paper/#fig:conceptual), [section 4.4](https://stain.github.io/ro-crate-paper/#dmp)
+2. Missing [figure 4](https://stain.github.io/ro-crate-paper/#fig:hdc) restored
+3. Citations added in [section 3](https://stain.github.io/ro-crate-paper/#tooling), [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows)
+4. Clarifications added [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities),
+5. Consistent section numbers HTML/PDF, added more cross references
+
+### Detailed response per comment
 
@@
 
-> Typos:
+We agree - we have added section numbers to both PDF and HTML.
+
+> **Typos**:
 > - In section RO-Crate. “abuot” -> “about”
+
+Fixed in [section 2](https://stain.github.io/ro-crate-paper/#rocrate).
+
 > - Figure 1 caption. “The RO is described within a ,”. The object of the sentence is missing.
-> - In Data Entities are described using Contextual Entities section, you wrote “person, but secondarily also a web page and their academic work.” I think the sentence is incomplete.
+
+Fixed in [Figure 1](https://stain.github.io/ro-crate-paper/#fig:conceptual) - this was a Markdown error.
+
+> - In Data Entities are described using Contextual Entities section, you wrote _“person, but secondarily also a web page and their academic work.”_ I think the sentence is incomplete.
+
+The sentence was indeed incomplete. We have changed it in [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities) to:
+
+_For instance https://orcid.org/0000-0002-1825-0097 is primarily an identifier for a person, but secondarily it is also a Web page and a way to refer to their academic work._
+
 > - In Table 1, the tool SCHeMa has the squared parenthesis signalling a citation, but there are question marks.
+
+Fixed in [section 3](https://stain.github.io/ro-crate-paper/#tooling).
+
 > - In Machine-actionable Data Management Plans section, there is a left curly bracket ({)
+
+Fixed in [section 4.4.](https://stain.github.io/ro-crate-paper/#dmp)
+
 > - In Institutional data repositories – Harvard Data Commons, the reference to figure 3 is missing
+
+Fixed [section 4.5](https://stain.github.io/ro-crate-paper/#institutionalrepos) with reference to (now) [figure 4](https://stain.github.io/ro-crate-paper/#fig:hdc).
+
 > - In Institutional data repositories – Harvard Data Commons (PDF version) figure 3 is missing. In the HTML version, the figure 3 is available.
+
+We have fixed this inconsistency (caused by graphics file extensions), Harvard Data Commons is now consistently [figure 4](https://stain.github.io/ro-crate-paper/#fig:hdc).
+
 > - In Packaging Workflows, there is `[cite?]` signalling a missing citation.
 
+We have reworded the sentence in [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows) to further clarify, adding the missing citation:
+
+_Indeed, in the absence of a specific requirement and an editorial policy to back it up – such as eliminating the reference limit – authors are effectively discouraged from properly and comprehensively citing software [[114](https://doi.org/10.1038/s41592-019-0350-x)]._
+
 > While reading the paper, I was eager to get to the evaluation section.
-At some point, in the Ensuring simplicity section, you state “One aim of RO-Crate is to be conceptually simple.
-> This simplicity has been repeatedly checked and confirmed through a community review process.”
-> I was indeed curious to see if you had run a survey on what users thought about RO-Crate.
-Perhaps, a survey gathering feedback on the user experience.
-Also, the concept of simplicity is bound with the skills of the practitioner.
-To what extent, they found it simple?
+At some point, in the Ensuring simplicity section, you state _“One aim of RO-Crate is to be conceptually simple.
+This simplicity has been repeatedly checked and confirmed through a community review process.”_
+>
+> I was indeed curious to see if you had **run a survey** on what users thought about RO-Crate. Perhaps, a survey gathering feedback on the user experience.
+
+For our early development methodology we created a pool of early adopter projects from a range of disciplines and groups that we refer to in the article in new section [2.1 Development methodology](https://stain.github.io/ro-crate-paper/#methodology).
+
+We are now ready for the next stage, which is to widen user adoption coupled with user evaluation to confirm our claims of conceptual simplicity and developer friendliness.
+We have added [section 7 Future Work](https://stain.github.io/ro-crate-paper/#futurework) where we outline our plans for developing novice-level tutorials and training events with feedback through user studies.
+
+> Also, the concept of simplicity is bound with the skills of the practitioner. To what extent, they found it simple?
+
+We have expanded on this in new section [2.1 Development methodology](https://stain.github.io/ro-crate-paper/#methodology).
 
@@
 
-## Reviewer 2
+We have outlined the next directions and recent new contacts in new section [7 Future Work](https://stain.github.io/ro-crate-paper/#futurework).
+
+## Response to Reviewer 2
 
@@
 
+We thank Alessia Bardi for her deep dive into the article, and for thoroughly checking the RO-Crate and the formal logic statements.
+
+### Summary of changes
+
+The reviewer found some bugs in the associated RO-Crate which we have fixed. We appreciate all the language improvements, which prompted us to revise the article readability.
+
+1.
+Language and typographical fixes in [section 2.1](https://stain.github.io/ro-crate-paper/#conceptual), [Figure 1](https://stain.github.io/ro-crate-paper/#fig:conceptual), [section 2.2.1](https://stain.github.io/ro-crate-paper/#linkeddata), [section 2.4.1](https://stain.github.io/ro-crate-paper/#people), [section 4](https://stain.github.io/ro-crate-paper/#inuse), [section 4.4](https://stain.github.io/ro-crate-paper/#dmp), [section 4.5](https://stain.github.io/ro-crate-paper/#institutionalrepos), [section 9.2](https://stain.github.io/ro-crate-paper/#minimal-ro-crate)
+2. Language/grammar revised throughout to improve readability (thanks Alan R Williams <https://orcid.org/0000-0003-3156-2105>)
+3.
+Clarifications added in [section 2.2](https://stain.github.io/ro-crate-paper/#conceptual), [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities), [Figure 2](https://stain.github.io/ro-crate-paper/#fig:uml), [section 5.2](https://stain.github.io/ro-crate-paper/#fair-digital-objects), [section 7](https://stain.github.io/ro-crate-paper/#futurework)
+4. [Listing 1](https://stain.github.io/ro-crate-paper/#jsonld) modified to show data entities refencing other data entities
+5.
+References added [section 2.4.1](https://stain.github.io/ro-crate-paper/#people), [section 3](https://stain.github.io/ro-crate-paper/#tooling), [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows)
+6. Fixed `ContextualEntity` mapping in formal logic statements of [appendix A section 9.2](https://stain.github.io/ro-crate-paper/#formaldefinition)
+
+For the benefit of readers of _this response_ we have also below included an example of using the production rules from appendix A.
+
+### Detailed response per comment
+
 > The most important remark is that I had not a good experience with the RO-Crate of this publication.
-> I clicked on supplementary files 1128.html and went to https://datasciencehub.net/system/files/ds-supplementary-706-1128.html#r...
-> - conformsTo is missing
+
+> I clicked on supplementary files 1128.html and went to <https://datasciencehub.net/system/files/ds-supplementary-706-1128.html#ro-crate-metadata.json>
+
+> - `conformsTo` is missing
+
+We apologise for this bad experience and recognise that we should have done the validation that she did on our behalf. Mea culpa - we must “eat our own dog food”!
+
+This turned out to be a [bug in ro-crate-html-js](https://github.com/UTS-eResearch/ro-crate-html-js/issues/10) rendering, it expected an contextual entity to display the linked RO-Crate standard.
+We have added such an entity in order to address this issue, but would point out this is not a formal requirement on a JSON level - see <https://www.researchobject.org/ro-crate/1.1/root-data-entity.html>
+
 > - the link to download returns 404
+
+This relative link was unfortunately broken by the datasciencehub.net submission website, which, out of our control, renamed the uploaded files and broke the relative links.
+We have checked with the IOS Press publisher, which unfortunately replied they are not able to host JSON files as supplementary material.
+In the accepted version we have published the corresponding RO-Crate separately using [GitHub pages](https://w3id.org/ro/doi/10.5281/zenodo.5146227) (with w3id.org PID) and Zenodo (DOI).
+
 > - the RO-Crate check says it is not a valid RO-Crate
+
+This was an unfortunate oversight in that the ro-crate-html-js rendering had a [more stringent view](https://github.com/UTS-eResearch/ro-crate-html-js/issues/11) than the specification on the requirements of license entities.
+We have fixed the updated RO-Crate.
+
 > The same happens for the version deposited on Zenodo (https://doi.org/10.5281/zenodo.5146227) linked in the abstract.
 
-Section "Conceptual definition":
-- Change ';' with ':' in "Figure 1 shows the main conceptual elements involved in an RO-Crate;"
-- Figure 1: check the caption, there are missing words at "The RO is described within a , "
-- Conceptual definition does not clarify what's the difference between a generic Research Object (as introduced in the previous section) and a RO-Crate
+We have updated the Zenodo deposit with the corrected RO-Crate.
 
-Section "Linked data as a foundation"
-- Move the footnote about IRIs the first time you use the acronym, if you think the footnote is needed
+_> Section "Conceptual definition":_
 
-Section "Data Entities are described using Contextual Entities"
-The example of orcid is not very clear to me. Why would I need to include a contextual entity that describes the person? For example, as shown in listing 1, we have
-"author": {"@id": "https://orcid.org/0000-0002-1825-0097"}
+> - Change ';' with ':' in _"Figure 1 shows the main conceptual elements involved in an RO-Crate;"_
 
-I can understand the suggestion to add
+Fixed in [section 2.1](https://stain.github.io/ro-crate-paper/#conceptual).
 
-```
-{ "@id": "https://orcid.org/0000-0002-1825-0097",
-"@type": "Person",
-"name": "Josiah Carberry"
-},
-```
+> - Figure 1: check the caption, there are missing words at _"The RO is described within a , "_
 
-for discoverability (and convenience) purposes, but following the linked data practices I could get the jsonld of the person with
-`curl --location --request GET 'https://orcid.org/0000-0002-1825-0097' --header 'Accept: application/ld+json'`
+Fixed [Figure 1](https://stain.github.io/ro-crate-paper/#fig:conceptual), this was an unfortunate Markdown/LaTeX rendering error.
 
-- Figure 2 does not clearly show how the data entities can be linked to each other. Neither does listing 1.
+> - Conceptual definition does not clarify what's the difference between a generic Research Object (as introduced in the previous section) and a RO-Crate
 
-Section "People"
-- add links to all mentioned projects.
-- Change ';' with ':' in "A key set of stakeholders are developers;"
-- Table 1: check missing link for SCHeMa
+We have added to [section 2.2](https://stain.github.io/ro-crate-paper/#conceptual):
 
-Section "Profiles of RO-Crate in use"
-- Cultural Heritages: remove the final 's'
+_The conceptual notion of a Research Object [[11](https://stain.github.io/ro-crate-paper/#ref-sRYUCzCq)] is thus realized with the RO-Crate model and serialized using Linked Data constructs within the RO-Crate metadata file._
 
-Section "Machine-actionable Data Management Plans"
-Pending brace at " RO-Crate {RDA maDMP Mapper"
+> Section "Linked data as a foundation"
+>
+> - Move the footnote about IRIs the first time you use the acronym, if you think the footnote is needed
 
-Section 'Institutional data repositories – Harvard Data Commons"
-- missing figure number at "in the metadata to allow their future reuse (Figure~)"
+Agreed, IRI footnote moved to first mention in [section 2.2.1](https://stain.github.io/ro-crate-paper/#linkeddata).
 
-Section "FAIR Digital Objects"
-- the link behind "As an “abstract protocol”" returns 404
+> Section "Data Entities are described using Contextual Entities"
+>
+> The example of orcid is not very clear to me. Why would I need to include a contextual entity that describes the person? For example, as shown in listing 1, we have:
+> 
+>    	"author": {"@id": "https://orcid.org/0000-0002-1825-0097"}
+>
+> I can understand the suggestion to add
+> 
+>    	{ "@id": "https://orcid.org/0000-0002-1825-0097",
+>         "@type": "Person",
+>         "name": "Josiah Carberry"
+>     },
+>
+> for discoverability (and convenience) purposes, but following the linked data practices I could get the jsonld of the person with
+>
+>     curl --location --request GET 'https://orcid.org/0000-0002-1825-0097' --header 'Accept: application/ld+json'
 
-Section "packaging workflows"
-- Missing citation at "authors might be penalised for doing so [cite?] "
+We are aware of this capability, and we believe other readers experienced with LOD will have similar thoughts.
+However, this showcases an important philosophical difference between RO-Crate and traditional Linked Data practices.
+We’ve added to [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities):
 
-Section "Minimal RO-Crate"
-- Align terms in the definition of minimal RO-crate (published vs datePublished)
-- Check footnote 8
+_In RO-Crate, a referenced contextual entity (e.g.
+a person identified by ORCID) should always be described within the RO-Crate Metadata File with at least a _type_ and _name_, even where their PID might resolve to further Linked Data.
+This is so that clients are not required to follow every link for presentation purposes, for instance HTML rendering.
+Similarly any imported [extension terms](https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#extending-ro-crate) would themselves also have a human-readable description in the case where their PID does not resolve to human-readable documentation._
 
-Section "Mapping to RDF with schema.org"
-- The mapping for ContextualEntity(e) is missing
+> Figure 2 does not clearly show how the data entities can be linked to each other. Neither does listing 1.
 
-A couple of final questions, that can be addressed in authors' future work, is about the profiles.
+The UML is a simplification of a Linked Data model with an Open World Assumption.
+We have modified the [Figure 2](https://stain.github.io/ro-crate-paper/#fig:uml) title to _“Simplified UML class diagram of RO-Crate**”_ **and expanded _“_contextual entities_ (which may include aggregated and non-aggregated data entities).
+”_
+
+We have modified [Listing 1](https://stain.github.io/ro-crate-paper/#jsonld) to showcase this possibility of data entities linking each others, and added _“In this example `about `of the CSV data entity reference the `ImageObject`, which then take the roles of both a data entity and contextual entity”_
+
+> Section "People"
+>
+> - add links to all mentioned projects.
+
+In [section 2.4.1](https://stain.github.io/ro-crate-paper/#people) we have added links and citations for the projects.
+
+> - Change ';' with ':' in "A key set of stakeholders are developers;"
+
+Fixed in [section 2.4.1](https://stain.github.io/ro-crate-paper/#people).
+
+> - Table 1: check missing link for SCHeMa
+
+Fixed in [section 3](https://stain.github.io/ro-crate-paper/#tooling).
+
+> Section "Profiles of RO-Crate in use"
+>
+> - Cultural Heritages: remove the final 's'
+
+Removed in [section 4](https://stain.github.io/ro-crate-paper/#inuse).
+
+> Section "Machine-actionable Data Management Plans"
+>
+> Pending brace at _" RO-Crate {RDA maDMP Mapper"_
+
+Fixed in [section 4.4](https://stain.github.io/ro-crate-paper/#dmp).
+
+> Section 'Institutional data repositories – Harvard Data Commons"
+> 
+> - missing figure number at "in the metadata to allow their future reuse (Figure~)"
+
+Fixed in [section 4.5](https://stain.github.io/ro-crate-paper/#institutionalrepos)
+
+> Section "FAIR Digital Objects"
+
+> - the link behind "As an “abstract protocol”" returns 404
+
+The link was accidentally broken by figure handling code in the manuscript HTML build process. Rephrased text in [section 5.2](https://stain.github.io/ro-crate-paper/#fair-digital-objects) to:
+
+_The Digital Object Interface Protocol [[@doip2.0](https://www.dona.net/sites/default/files/2018-11/DOIPv2Spec_1.pdf)] can be considered an "abstract protocol" of requirements, DOs could be implemented in multiple ways._
+
+> Section "packaging workflows"
+>
+> - Missing citation at _"authors might be penalised for doing so [cite?] "_
+
+We rephrased this in [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows) to clarify, and added the missing citation:
+
+_Indeed, in the absence of a specific requirement and an editorial policy to back it up – such as eliminating the reference limit – authors are effectively discouraged from properly and comprehensively citing software [[116](https://doi.org/10.1038/s41592-019-0350-x)]._
+
+> Section "Minimal RO-Crate"
+> 
+> - Align terms in the definition of minimal RO-crate (`published` vs `datePublished`)
+> 
+> - Check footnote 8
+
+Fixed to `datePublished()` in [section 9.2](https://stain.github.io/ro-crate-paper/#minimal-ro-crate) to be consistent [with footnote](https://stain.github.io/ro-crate-paper/#fn11).
+
+> Section "Mapping to RDF with schema.org"
+>
+> - The mapping for ContextualEntity(e) is missing
+
+Thanks for spotting this. For any interested readers, below we explore the mapping following the production rules.
+
+The old mapping in [appendix A](https://stain.github.io/ro-crate-paper/#formaldefinition) assumed there was an explicit type(e, c) for the contextual entity, e.g. if we have:
+
+    E ≡ <http://example.com/>
+    L ≡ <https://spdx.org/licenses/CC0-1.0>
+    license(E, L)
+
+Then following the 𝕃𝖗𝖉𝖋 production rules:
+
+    license(E, L)
+    ⇒ Relation(E, &lt;http://schema.org/license>, L) ∧ CreativeWork(L)
+
+    CreativeWork(L)
+    ⇒ ContextualEntity(L) ∧
+
+    type(L, <http://schema.org/CreativeWork>)
+    ⇒ Relation(L, <http://schema.org/CreativeWork>)
+    ∧ Class(<http://schema.org/CreativeWork>)
+
+    Relation(E, <http://schema.org/license>, L)
+    ⇒ Triple(E, <http://schema.org/license>, L)
+    ∧IRI(E) ∧ IRI(L)
+
+    Relation(L,<http://www.w3.org/1999/02/22-rdf-syntax-ns#type>, <http://schema.org/CreativeWork>)
+    ⇒ Triple(L,<http://www.w3.org/1999/02/22-rdf-syntax-ns#type>,<http://schema.org/CreativeWork>)
+    ∧IRI(L) ∧ IRI(&lt;http://schema.org/CreativeWork>)
+
+In this example, `ContextualEntity(L)` did not need to get expanded to 𝕃𝖗𝖉𝖋, as we already got an implied type `CreativeWork(L)` from `license(E, L)`.
+
+Expanding back E and L we get the corresponding triples:
+
+    Triple(<http://example.com/>, <http://schema.org/license>, <https://spdx.org/licenses/CC0-1.0>)
+
+    Triple(<https://spdx.org/licenses/CC0-1.0>, <http://www.w3.org/1999/02/22-rdf-syntax-ns#type>, <http://schema.org/CreativeWork>)
+
+However this expectation of an explicit `type(e, l)` was indeed not enforced in our original definition of 𝕃𝖗𝖔𝖈𝖗𝖆𝖙𝖊.
+
+Considering the recommendation in updated [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities): _“a referenced contextual entity should always be described within the RO-Crate Metadata File with at least a type and name, even where their PID might resolve to further Linked Data_”, then such a requirement should be added also to the first order logic rules.
+
+Investigating this, you also helped uncover a variable name typo in the [section 9.2](https://stain.github.io/ro-crate-paper/#minimal-ro-crate) Class declaration:
+
+    ∀e Entity(e) ⇒  Class(e)
+
+In the updated manuscript, the above has been replaced with:
+
+    ∀e Entity(e) ⇒  type(e, c) ∧ Class(c)
+   	∀e ContextualEntity(e) ⇒  name(e, n)
+
+The updated mapping to 𝕃𝖗𝖉𝖋 adds this relatively harmless fallback, to ensure a triple is produced for untyped URI references:
+
+    ContextualEntity(e) ⇒ type(e, &lt;http://schema.org/Thing>)
+
+The mapping for any arbitrary `ContextualEntity(E)` will then proceed as in the license example.
+
+An alternative rule, which we dismissed as it is adding RDFS inferencing, would be to represent the property <http://schema.org/name> which is defined with domain <http://schema.org/Thing>.
+This rule would then be triggered by the previous `ContextualEntity` existential requirement of `n` in `name(e, n)`:
+
+    ∀e∀n name(e, n) ⇒ type(e, <http://schema.org/Thing>)
+
+> A couple of final questions, that can be addressed in authors' future work, is about the **profiles**.
 Could they limit the re-usability of RO-Crate and the data entities in it in different domain?
@@
 
-## Reviewer 3
+We appreciate these questions, and refer to the new [subsection 6.1](https://stain.github.io/ro-crate-paper/#strictness-vs-flexibility) that includes:
 
-_By anonymous:_
+_Specialization comes hand-in-hand with the principle of _graceful degradation_; RO-Crate applications and users are free to choose the semantic detail level they participate at, as long as they follow the common syntactic requirements._
+
+Even though there is common ground across profiles thanks to graceful degradation, the reviewer is right to point out that a proliferation would move the interoperability goal posts and risk different RO-Crate extensions diverging with different vocabularies and conventions.
+We have thus expanded new [section 7](https://stain.github.io/ro-crate-paper/#futurework) to include:
+
+_By discovering commonalities in emerging usage (e.g.
+additional schema.org types), the community helps to reduce divergence that could otherwise occur with proliferation of further RO-Crate profiles._
+
+## Response to Reviewer 3
+
+_By anonymous_
+
+The reviewer provided helpful perspectives on Linked Open Data potential of RO-Crate and reminded us that we should not underplay this.
+We agree on the need for a user study and have added this to the planned future work, as well as highlighting the development methodology.
+
+The suggestions on expanding the related work in a couple of directions are understandable.
+However, after careful consideration, we have concluded that to give justice in such a review we would need to significantly increase the scope and length of this article.
+We therefore rather propose to address this in a future work with a wider review, and have cited relevant existing review articles in [Section 5](https://stain.github.io/ro-crate-paper/#relatedwork).
+
+For the benefit of future readers we have expanded on our reasoning below.
+
+### Overview of changes
+
+1. Reference added to [Section 5](https://stain.github.io/ro-crate-paper/#relatedwork)
+2. Clarified user engagement strategy in new [2.1 Development methodology](https://stain.github.io/ro-crate-paper/#methodology).
+3. Added [Section 7 Future Work](https://stain.github.io/ro-crate-paper/#futurework), including plans for a user study
+4.
+Highlighted potential for consuming RO-Crate as RDF/knowledge graph in [section 2.2.1](https://stain.github.io/ro-crate-paper/#linkeddata), [section 4.3](https://stain.github.io/ro-crate-paper/#culturalheritage), [section 7](https://stain.github.io/ro-crate-paper/#futurework)
+5. Expanded [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows) on practical limitation of workflow provenance
+6. Clarifications in to [section 2.2.6](https://stain.github.io/ro-crate-paper/#profiles), [section 6](https://stain.github.io/ro-crate-paper/#conclusion)
+
+### Detailed response per comment
 
@@
 
+In [Section 5](https://stain.github.io/ro-crate-paper/#relatedwork) we have added a reference to newly published <https://doi.org/10.1016/j.patter.2021.100322> which we have contributed to, and which provides an extensive list of metadata standards with importance for reproducible computational research.
+A similar review on data packaging literature would be welcome, but we feel that could get too extensive for the scope of this manuscript as a Resource Article.
+
 > Although the idea behind the RO-Crate seems interesting, in its current shape that is designed for a limited community with limited purposes.
+
 > The impact of RO-Crate for large uptake and long term could be formulated better.
+
+We realised from this comment and other reviewers that we should be much more explicit in our user engagement strategy, which was to start with a small group and build firm foundations and making sure that the approach worked before we set out for wider adoption.
+We have added this in a new section [2.1 Development methodology](https://stain.github.io/ro-crate-paper/#methodology).
+
+It is right to point out that RO-Crate has a relatively small user base today, however it is also a fairly broad base in terms of scientific domains involved, which is part of our argument for RO-Crate being applicable for general purpose description of research artefacts.
+
+We have added [Section 7 Future Work](https://stain.github.io/ro-crate-paper/#futurework) on how we are working on increasing uptake, current outreach activities and new connections.
 
@@
 
+We have focused this article on RO-Crate as a platform-neutral mechanism to capture data, metadata and their context.
+The sphere of Open Science-supporting platforms/services is very exciting and important for progress on the practice of FAIR and reproducible data, but we are concerned that comparing a metadata format with these varied platforms could get confusing to the reader.
+
 > Though a resource paper, there is no **evaluation or statistics** for comparison of RO-Crate to any other similar service. Even a **user study** from the LOD community would be helpful.
+
+To clarify this we have added to new [section 2.1 on our user methodology](https://stain.github.io/ro-crate-paper/#methodology) that we have stabilized the RO-Crate specification, and are now at the stage of building wider adoption.
+
+We agree on the suggestion of a user study, particularly after development of tutorials.
+We are planning a user study with the Linked Open Data community or as part of EOSC Bring-your-own-Data training events, and have included this in the new [section 7 Future Work](https://stain.github.io/ro-crate-paper/#futurework).
 
@@
 Providing a statistic from repositories will be helpful.
+
+We recognize this limitation, and have expanded [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows) to add:
+
+_It is important to note that having all these computational details in order to represent them in an RO-Crate is an ideal scenario -- in practice there will always be gaps of knowledge, and exposing all provenance details automatically would require improvements to the data sources, workflow, workflow engine and its dependencies.
+RO-Crate can be seen as a flexible annotation mechanism for augmenting automatic workflow provenance.
+Additional metadata can be added manually, e.g.
+for sensitive clinical data that cannot be publicly exposed<sup>1</sup>, or to identify prototype software without a web presence.
+This inline _FAIRifying _allows researchers to achieve “just enough FAIR” to explain their computational experiments._
+<sup>1</sup>: FAIR principle A2: _Metadata are accessible, even when the data are no longer available_. [[10.1038/sdata.2016.18](https://doi.org/10.1038/sdata.2016.18)]
+
+We recognise the value of gaining understanding of how RO-Crate users deal with missing data and descriptions.
+This can be discussed as part of the user study mentioned above, which will be a good opportunity to get direct feedback from users that might not be fully familiar with RO-Crate.
+
+Regarding statistics from existing repositories, we have left out of the scope of this article the analysis of the current practices in Open Science Registries and repositories.
+This is, however, an intriguing line of work that may motivate even further the need for RO-Crates.
+
 > - Although the primary purpose of the RO-Crate specification is to assist developers in leveraging LOD, how can individual researchers benefit from RO-Crate or is it just a community-wide service for LOD developers?
+
+Individual researchers benefit from RO-Crates by having access to bundles of information with connected resources, in context (and through the applications aimed to help their consumption).
+While this is available in a structured human-readable HTML format, RO-Crate JSON is intended to be machine-readable, meaning software is needed to consume it, primarily with existing RO-Crate tools ([section 3](https://stain.github.io/ro-crate-paper/#tooling)), general Linked Data tools ([section 2.2.1](https://stain.github.io/ro-crate-paper/#linkeddata)) or alternatively with new code at JSON level.
+
+We have thus so far, in the development of specifications for writing RO-Crate as JSON, primarily targeted developers who may not be familiar with Linked Open Data principles, but which are familiar with Web technologies.
+We describe this further in new [section 2.1](https://stain.github.io/ro-crate-paper/#methodology).
+
+As part of this we have also prototyped user-level applications for individual researchers (e.g.
+Describo) as well as integration into data platforms.
+The next step, as shown in new [section 7](https://stain.github.io/ro-crate-paper/#futurework), is to further develop both of these aspects along with training material to onboard wider communities of users which are not interested in aspects of JSON serialization.
+
 > - It is mentioned that ‘RO-Crate is used across multiple areas’, but how can other less-technical domains benefit from RO-Crate?
+
+Firstly, we would like to object to the term “_less technical domains_”, as for instance humanities, palaeontology, art, history, psychology and language studies are all examples of non-STEM domains which are adapting digital data management practices in their every-day research, as well as using high tech instruments for a variety of purposes.
+
+A large part of RO-Crate’s work by the Australian partners is focused on this _eScience_ cross-over with humanities and social sciences, both by building user-level tooling for describing individual data files (Describo, ro-crate-html-js, ro-crate-excel) as well as integration into existing data platforms like PARADISEC ([section 4.3](https://stain.github.io/ro-crate-paper/#culturalheritage)).
+
+One challenge for a new domain is to use RO-Crate beyond the bibliographic level, to form what is effectively a new RO-Crate profile with additional contextual types and properties that are meaningful for their desired metadata.
+
+We have found that, to a large degree, existing terms from schema.org can be used, however their match is not always obvious to the domain scientist and may still need help from a LOD expert.
+Tools like Describo help by providing a user interface for a given type, and we have experimented with supporting the full schema.org vocabulary without otherwise dealing with low level Linked Data details.
+
+However we did find that, without the filtering from a profile, it can be overwhelming for users to navigate or understand how a type could be used.
+For instance, [https://schema.org/HowTo](https://schema.org/HowTo) is seemingly explained as a generalization of food recipes which sounds unrelated to science, but on closer inspection this type can work equally well to structure the steps of an industrial biochemical process or how to prepare an insect specimen in a museum collection.
+
+When we identify such common solutions within the RO-Crate Community we “lift” a practice from specific profiles to the general RO-Crate specification, adding the explanation “glue” on how it would apply to different domains, e.g.
+for [equipment used to create files](https://www.researchobject.org/ro-crate/1.1/provenance.html#equipment-used-to-create-files)<span style="text-decoration:underline;">.</span>
+
+While we have added some of these details to [section 2.2.6](https://stain.github.io/ro-crate-paper/#profiles) and [section 7](https://stain.github.io/ro-crate-paper/#futurework), we plan to expand further on this in a separate future work detailing RO-Crate profile development and formalization.
+
 > - How RO-Crate relates and differs from ORKG (https://www.orkg.org/orkg/) and other similar platforms?
+
+ORKG, as stated in [https://www.orkg.org/orkg/about](https://www.orkg.org/orkg/about) and [https://www.orkg.org/orkg/about/2/Features](https://www.orkg.org/orkg/about/2/Features), focuses on scholarly literature, the description of granular elements reported in scientific articles and the comparison across articles.
+
+RO-Crate aims at packing all sorts of research outcomes as a coherent unit that can be FAIRly shared, and does not require a corresponding scholarly article.
+
+In terms of granularity, an RO-Crate is generally a structured “map” to other digital resources, which themselves convey knowledge in other formats, rather than aiming to be a container of all the captured scientific knowledge.
+
+A knowledge graph can be generated/derived from multiple RO-Crates and queried in detail, which can be particularly powerful if the crates follow the same profile.
+We have perhaps not sufficiently emphasized this opportunity in the [current specification](https://www.researchobject.org/ro-crate/1.1/appendix/relative-uris.html#parsing-as-rdf-with-a-different-ro-crate-root), although the manuscript briefly suggests use of SPARQL queries in [section 2.2.1](https://stain.github.io/ro-crate-paper/#linkeddata) and [section 4.3](https://stain.github.io/ro-crate-paper/#culturalheritage).
+
+We will explore the consumption using knowledge graphs further in developing Jupyter Notebook tutorials, and have highlighted this in new [section 7 Future work](https://stain.github.io/ro-crate-paper/#futurework).
+
+With relation to ORKG and RO-Crate, we do however not see how we can do a fair comparison, as we think they are pursuing different aims at different granularities.
+Other related work in this field, delving into formalizing arguments from literature and representing individual scientific facts, include [nanopublications](https://doi.org/10.7717/peerj-cs.387) and [FAIRSCAPE](https://doi.org/10.1007/s12021-021-09529-4).
+Expanding the related work would naturally then also need to cover these, which we deem out of scope for this work as a resource article.
+
 > - RO-Crate is presented as ‘a structured archive’ and not called a knowledge graph (KG) as it seems to be more than a KG, but looking into the details of the service, it looks like a multimodal scholarly KG - what are the authors’ viewpoints here?
+
+We like the categorization of _multimodal scholarly knowledge graph_, and have taken the liberty of adding this point to [section 6](https://stain.github.io/ro-crate-paper/#conclusion).
+
+We think of RO-Crate mainly as a way to structure data and its metadata, but clearly the power of knowledge graph is one reason why we have adapted Linked Data technologies.
+We have emphasized this point stronger in updated [section 2.2.1](https://stain.github.io/ro-crate-paper/#linkeddata).
+
+In new [section 7 Future Work](https://stain.github.io/ro-crate-paper/#futurework) we have also highlighted the plans for demonstrating more of consumption using knowledge graph technologies:
+
+We will also collect and build examples of_ RO-Crate consumption_, e.g. Jupyter Notebooks that query multiple crates using knowledge graphs.
 
@@
 
-I have to confess it has been a pleasure reading this paper since it presents, with a lot of details, an important resource for the scholarly community at large, in particular in the context of current practices (such as those concerning FAIR and EOSC).
-All the reviewers agreed on this aspect, even if they provided some insights and questions that should be appropriately addressed in the camera-ready and that I will carefully check before publication.
+> I have to confess it has been a pleasure reading this paper since it presents, with a lot of details, an important resource for the scholarly community at large, in particular in the context of current practices (such as those concerning FAIR and EOSC).
+
+> All the reviewers agreed on this aspect, even if they provided some insights and questions that should be appropriately addressed in the camera-ready and that I will carefully check before publication.
+
+We thank the editors and the reviewers for the insightful comments and suggestions, which we feel have significantly improved the article.
+We hope you will find our edits to address the issues raised to your satisfaction.

### 2022-01-06 content/80.formal-definition.tex
typo again!
schema.org -> Schema.org (in text)

@@
 Entity(e)                                   & \equiv &      DataEntity(e) \lor ContextualEntity(e) \\
-\forall e \ Entity(e)                       & \Rightarrow & type(e, c) \land Class(e) \\
+\forall e \ Entity(e)                       & \Rightarrow & type(e, c) \land Class(c) \\
 \forall e \ ContextualEntity(e)             & \Rightarrow & name(e, n)  \\
@@
 
-In reality many additional attributes and contextual types from schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (\textit{SHOULD} requirements) they do not form part of this formalization.
+In reality many additional attributes and contextual types from Schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (\textit{SHOULD} requirements) they do not form part of this formalization.
 
-\subsection{Mapping to RDF with schema.org}
+\subsection{Mapping to RDF with Schema.org}
 
@@
 
-Below follows a mapping from $\mathcal{L}_{rocrate}$ to $\mathcal{L}_{RDF}$ using schema.org as vocabulary:
+Below follows a mapping from $\mathcal{L}_{rocrate}$ to $\mathcal{L}_{RDF}$ using Schema.org as vocabulary:
 
@@
 
-Note that in the JSON-LD serialization of RO-Crate, the expression of $Class$ and $Property$ is typically indirect: The JSON-LD \texttt{@context} maps to schema.org IRIs, which, when resolved as Linked Data, embed their formal definition as RDFa.
+Note that in the JSON-LD serialization of RO-Crate, the expression of $Class$ and $Property$ is typically indirect: The JSON-LD \texttt{@context} maps to Schema.org IRIs, which, when resolved as Linked Data, embed their formal definition as RDFa.
 Extensions may however include such term definitions directly in the RO-Crate.
@@
 
-Combining the above predicates and schema.org mapping with rudimentary JSON templates, these forward-chaining production rules can output JSON-LD according to the RO-Crate 1.1 specification\footnote{
+Combining the above predicates and Schema.org mapping with rudimentary JSON templates, these forward-chaining production rules can output JSON-LD according to the RO-Crate 1.1 specification\footnote{
     \textbf{Limitations:} Contextual entities not related from the RO-Crate (e.g.
@@
 These production rules first grow the graph of $R$ by adding a transitive rule – anything described in $R$ which is related to $o$, means that $o$ is also mentioned by the $ROCrate(R)$.
-For simplicity this rule is one-way; in theory the graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities, but these are proposed to be bound to the root data entity with schema.org relation \href{http://schema.org/mentions}.
+For simplicity this rule is one-way; in theory the graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities, but these are proposed to be bound to the root data entity with Schema.org relation \href{http://schema.org/mentions}.

### 2022-01-06 content/31.tooling-table.tex
Alan's typo fixes and copy-editing
Signed-off-by: Stian Soiland-Reyes <stain@apache.org>
Typos fixed as in proof to IOS Press

@@
 
-		\textbf{ro-crate-excel} \citep{ro-crate-excel} & Data managers & JavaScript \\
+		\textbf{ro-crate-excel} \citep{ro-crate-excel} & Data managers & JavaScript & Beta \\
 		\multicolumn{4}{l}{\it Command-line tool to create/edit RO-Crates with spreadsheets} \\
@@
 		\textbf{ONI express} \citep{arkisto-data-portal} & Repository managers & Platform & Beta \\
-		\multicolumn{4}{l}{\it Platform for publishing data and documents stored in an OCFL repository via a web interface} \\
+		\multicolumn{4}{l}{\it Platform for publishing data and documents stored in an OCFL repository via a Web interface} \\
 

### 2022-01-07 content/80.formal-definition.tex
URIs -> IRIs
typo fixes as in proof

@@
     \mathbb{D}              & = & \mathbb{IRI} \\
-    \mathbb{IRI}            & \equiv & { \text{IRIs as defined in RFC3987} } \\
+    \mathbb{IRI}            & \equiv & { \text{IRIs as defined in RFC3987 \cite{doi:10.17487/rfc3987}}} } \\
     \mathbb{R}              & \equiv & { \text{real or integer numbers} } \\
@@
 
-The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers \cite{doi:10.17487/rfc3987} (notation $\texttt{<http://example.com/>}$)\footnote{
+The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers (notation $\texttt{<http://example.com/>}$)\footnote{
     For simplicity, blank nodes are not included in this formalisation, as RO-Crate
@@
 
-An $ROCrate(R)$ is defined as a self-described \emph{Root Data Entity}, which describes and contains parts (\emph{data entities}), which are further described in \emph{contextual entities}.
+An $ROCrate(R)$ is defined as a self-described \emph{Root Data Entity}, which contains parts (\emph{data entities}), which are further described in \emph{contextual entities}.
 These terms align with their use in the RO-Crate 1.1 terminology\footnote{
@@
 
-The $Root(r)$ is a type of $Dataset(r)$, and must have the metadata to literal attributes to provide a $name$, $description$ and $datePublished$, as well as a contextual entity identifying its license.
-These predicates correspond to the RO-Crate 1.1 requirements for the root data entity\footnote{
+The $Root(r)$ is a type of $Dataset(r)$, and must as metadata have at least the attributes $name$, $description$ and $datePublished$, as well as a contextual entity that identify its $license$.
+These predicates correspond to the RO-Crate 1.1 minimal requirements for the root data entity\footnote{
     \url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}
@@
 
-The below is an example RO-Crate represented using the above formalisation, assuming a base URI of \texttt{<http://example.com/ro/123/>}:
+The below is an example RO-Crate represented using the above formalisation, assuming a base IRI of \texttt{<http://example.com/ro/123/>}:
 
@@
 && name(\texttt{<https://spdx.org/licenses/CC-BY-4.0>},  \\
-&& \ \ \ \ \  \text{Creative Commons Attribution 4.0”}) \\
+&& \ \ \ \ \  \text{“Creative Commons Attribution 4.0”}) \\
 \\
@@
 && name(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>},  \\
-&& \ \ \ \ \  \text{Creative Commons Attribution Non Commercial 4.0”}) \\
+&& \ \ \ \ \  \text{“Creative Commons Attribution Non Commercial 4.0”}) \\
 \\
@@
 Assume a simplified\footnote{
- This simplification does not cover the extensive list of literal datatypes built-in to RDF 1.1, only strings and decimal real numbers. Likewise, language of literals are not included.
+ This simplification and mapping does not cover the extensive list of literal datatypes built into RDF 1.1, only strings and decimal real numbers.
+Likewise, $LanguageTag$ is deliberately not utillised below.
 } language $\mathcal{L}_{RDF}$
@@
 
-An important RO-Crate principle is that of being \textbf{self-described} Therefore the serialisation of the RO-Crate into a file should also describe itself in a Metadata File Descriptor\footnote{
+An important RO-Crate principle is that of being \textbf{self-described}. Therefore the serialisation of the RO-Crate into a file should also describe itself in a Metadata File Descriptor\footnote{
     \url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#ro-crate-metadata-file-descriptor}
@@
 
-In the conceptual model the \emph{RO-Crate Metadata File} can be seen as the top-level node that describes the \emph{RO-Crate Root}, however in the formal model (and the JSON-LD format) the metadata file descriptor is an additional contextual entity that is not affecting the depth-limit of the RO-Crate.
+In the conceptual model, the \emph{RO-Crate Metadata File} can be seen as the top-level node that describes the \emph{RO-Crate Root}, however in the formal model (and the JSON-LD format) the metadata file descriptor is an additional contextual entity that is not affecting the depth-limit of the RO-Crate.
 

### 2022-01-07 response-to-review.md

@@
 
-Response to  open peer review at <https://datasciencehub.net/paper/packaging-research-artefacts-ro-crate> of “Packaging research artefacts with RO-Crate” submitted to Data Science._
+_Response to  open peer review at <https://datasciencehub.net/paper/packaging-research-artefacts-ro-crate> of “Packaging research artefacts with RO-Crate” submitted to Data Science._
 

### 2022-01-08 latex/ro-crate.tex
DOI as published in Data Science
Alphabetical bibliography as in proof
latex content+bibliography from ro-crate-proof.tex
but with our images

--- 2022-01-02 latex/ro-crate.tex
@@
-%BeginFileInfo
-%%Publisher=IOS
-%%Project=DS
-%%Manuscript=DS210053
-%%MS position=
-%%Stage=201
-%%TID=imikolaityte
-%%Pages=0
-%%Format=2019
-%%Distribution=vtex
-%%Destination=DVI
-%%PDF type=
-%%DVI.Maker=luadvi
-%%PS.Maker=luaps
-%%History1=2021.12.28 09:11
-%EndFileInfo
-%
-% Journal: Data Science (DS), IOS Press
-% Typeset in LaTeX2e by VTeX Ltd., Vilnius, Lithuania
-%
-%Spelling_date
-%\RequirePackage{float} % when algorithm.sty used in manuscript
-\documentclass[ds,v1.1.2,openaccess]{iosart2x}%% ,seceqn
-\usepackage{mathbh}
-%
-%\usepackage{}
-\usepackage{tmultirow}
-\usepackage{listings}
-%
-\PROOF
-%\PROOF
-%\CRC
-
-\volume{-1}
-\issue{-1}
-\pubyear{2021}
+%% ro-crate paper
+%% (c) 2020-2021 The University of Manchester
+%% 
+%% SPDX-License-Identifier: CC-BY-4.0
+
+% add. options: [seceqn,secthm,crcready]
+\documentclass[ds,crcready]{iosart2x}
+
+%%%%%%%%%%% Put your definitions here
+
+% From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
+\usepackage{fontspec}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{hyperref}
+\usepackage{xcolor}
+\definecolor{darkerblue}{cmyk}{1,1,0,0.45}
+\hypersetup{
+    colorlinks=true,
+    linkcolor=black,
+    citecolor=black,
+    urlcolor=darkerblue}
+
+\usepackage{unicode-math}
+\usepackage{graphicx}
+
+% Hyperlinks on DOI permalinks – just 20 years after DOI invention!
+% 
+% We had to make our own copy of 
+% https://ctan.org/pkg/doi?lang=en 
+% where commands are renamed to \doix
+% because iosart2x.cls and ios1.bst have 
+% a different use of \doiurl \doi etc.
+% We also set the doixtext prefix empty
+% to avoid doi:doi:
+\usepackage{doix}
+\renewcommand{\doixtext}{}
+\renewcommand{\doiurl}[1]{\doix{#1}}
+
+\DeclareGraphicsExtensions{.pdf,.png,.svg}
+\graphicspath{../content/images/}
+\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
+\def\markdownOptionOutputDir{../output}
+\def\markdownOptionHeaderAttributes{true}
+
+% for table resizing
+\usepackage{booktabs}
+\usepackage{multirow}
+
+%%%%%%%%%%% End of definitions
+
+\pubyear{2022}
+\volume{0}
 \firstpage{1}
-\lastpage{42}
-\articletype{research-article} %%
-%http://www.vtex.vmt/docs/wiki/index.php/IOS_Press_instrukcijos_%28Kaunas%29#Straipsnio_tipas_.28.5Carticletype.7B.7D.29
-\aid{DS210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
-\doi{10.3233/DS-210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
-\vtexed{Ingrida}
-
-\startlocaldefs
-\setpkgattr{license}{type}{CC@BY} %% CC@BY@NC
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% indentation after theorem-like environments - auto
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\newtheorem*{}{} %% unnumbered
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\theoremstyle{plain}
-%%\newtheorem{}{}%[]
-%%\theoremstyle{definition}
-%%\newtheorem{}{}%[]
-\endlocaldefs
-%
+\lastpage{1}
+\def\doi{10.3233/DS-210053}
+
 \begin{document}
-%
+
+% Welcome to the 2000s with UNICODE fonts!
+% Adapted from
+% https://www.overleaf.com/learn/latex/Articles%2FOpenType-based_math_typesetting:_An_introduction_to_the_STIX2_OpenType_fonts
+\setmathfont{STIX2Math}[
+    Extension={.otf},
+    Path=./STIX2fonts/,
+    Scale=1
+]
+\setmainfont{STIX2Text}[
+    Extension={.otf},
+    Path=./STIX2fonts/,
+    UprightFont={*-Regular},
+    BoldFont={*-Bold},
+    ItalicFont={*-Italic},
+    BoldItalicFont={*-BoldItalic}
+]
+
+% https://typeof.net/Iosevka/ mono font has 
+% all the logic symbols we need
+%\setmonofont{iosevka-slab}[
+%    Extension={.ttf},
+%    Path=./ttf-iosevka-slab-6.1.3/,
+%    Scale=MatchLowercase,
+%    UprightFont={*-light},
+%    ItalicFont={*-lightitalic}
+%]
+
 \begin{frontmatter}
-%
-\pretitle[XML]{Research Article}
+
 %\pretitle{}
 \title{Packaging research artefacts with RO-Crate}
-%\title{\thanks{}}
 \runtitle{Packaging research artefacts with RO-Crate}
 %\subtitle{}
-%
-%%% initials without spaces
-%
-% {\inits{}\fnms{} \snm{} \suffix{}\ead[label=e?]{}
-%\ead[label=p1,phone]{}\ead[label=f1,fax]{}\ead[label=o?,orcid]{}
-%\thanks{}}
-%
-\begin{aug}
-\author[A,B]{\inits{S.}\fnms{Stian} \snm{Soiland-Reyes}\ead
-[label=e1]{soiland-reyes@manchester.ac.uk}\ead
-[label=o1,orcid]{0000-0001-9842-9718}\thanks{Corresponding author.
-\printead{e1}.}}
-\author[C]{\inits{P.}\fnms{Peter} \snm{Sefton}\ead
-[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}}
-\author[D]{\inits{M.}\fnms{Merc\`{e}} \snm{Crosas}\ead
-[label=o3,orcid]{0000-0003-1304-1939}}
-\author[E]{\inits{L.J.}\fnms{Leyla Jael} \snm{Castro}\ead
-[label=o4,orcid]{0000-0003-3986-0510}}
-\author[F]{\inits{F.}\fnms{Frederik} \snm{Coppens}\ead
-[label=o5,orcid]{0000-0001-6565-5145}}
-\author[G]{\inits{J.M.}\fnms{Jos\'{e} M.} \snm{Fern\'{a}ndez}\ead
-[label=o6,orcid]{0000-0002-4806-5140}}
-\author[H]{\inits{D.}\fnms{Daniel} \snm{Garijo}\ead
-[label=o7,orcid]{0000-0003-0454-7145}}
-\author[I]{\inits{B.}\fnms{Bj\"{o}rn} \snm{Gr\"{u}ning}\ead
-[label=o8,orcid]{0000-0002-3079-6586}}
-\author[J]{\inits{M.}\fnms{Marco} \snm{La Rosa}\ead
-[label=o9,orcid]{0000-0001-5383-6993}}
-\author[K]{\inits{S.}\fnms{Simone} \snm{Leo}\ead
-[label=o10,orcid]{0000-0001-8271-5429}}
-\author[L]{\inits{E.}\fnms{Eoghan} \snm{\'{O} Carrag\'{a}in}\ead
-[label=o11,orcid]{0000-0001-8131-2150}}
-\author[M]{\inits{M.}\fnms{Marc} \snm{Portier}\ead
-[label=o12,orcid]{0000-0002-9648-6484}}
-\author[N]{\inits{A.}\fnms{Ana} \snm{Trisovic}\ead
-[label=o13,orcid]{0000-0003-1991-0533}}
-\author[O]{\snm{RO-Crate Community}\ead
-[label=u1,url]{https://www.researchobject.org/ro-crate/community}}
-\author[P]{\inits{P.}\fnms{Paul} \snm{Groth}\ead
-[label=e15]{p.t.groth@uva.nl}\ead[label=o15,orcid]{0000-0003-0183-6910}}
-\author[Q]{\inits{C.}\fnms{Carole} \snm{Goble}\ead
-[label=e16]{carole.goble@manchester.ac.uk}\ead
-[label=o16,orcid]{0000-0003-1219-2137}}
-%
-%%%%%% printead order: TEL FAX E-MAIL URL
-%\thanks{Corresponding author: name surname, address. \printead{}.}
-%\thanks{Corresponding author. \printead{}.}
-%\thanks{}%% Move support info to acks
-%
-\runauthor{S. Soiland-Reyes et al.} %% auto
-% {\orgname{} \cny{}\printead[presep={\\}]{}}
-\address[A]{Department of Computer Science, \orgname{The University of
-Manchester}, \cny{UK}\printead[presep={\\}]{e1,o1}}
-\address[B]{Informatics Institute, \orgname{University of Amsterdam},
+
+% For one author:
+%\author{\inits{N.}\fnms{Name1} \snm{Surname1}\ead[label=e1]{first@somewhere.com}\ead[label=o1,orcid]{0000-0000-0000-0001}}
+%\address{Department first, \orgname{University or Company name},
+%Abbreviate US states, \cny{Country}\printead[presep={\\}]{e1,o1}}
+
+% Two or more authors:
+\author[SSR1,SSR2]{\inits{S.}\fnms{Stian}
+    \snm{Soiland-Reyes}\ead[label=eSSR]{soiland-reyes@manchester.ac.uk}\ead[label=oSSR,orcid]{0000-0001-9842-9718}%
+    \thanks{Corresponding author. \printead{eSSR}.}}, 
+\author[PS]{\inits{P.}\fnms{Peter}
+    \snm{Sefton}\ead[label=ePS]{Peter.Sefton@uts.edu.au}\ead[label=oPS,orcid]{0000-0002-3545-944X}},
+\author[MC]{\inits{M.}\fnms{Mercè}
+    \snm{Crosas}\ead[label=eMC]{mcrosas@g.harvard.edu}\ead[label=oMC,orcid]{0000-0003-1304-1939}},
+\author[LJC]{\inits{LJ.}\fnms{Leyla Jael}
+    \snm{Castro}\ead[label=eLJC]{ljgarcia@zbmed.de}\ead[label=oLJC,orcid]{0000-0003-3986-0510}},
+\author[FC]{\inits{F.}\fnms{Frederik}
+    \snm{Coppens}\ead[label=eFC]{frcop@psb.vib-ugent.be}\ead[label=oFC,orcid]{0000-0001-6565-5145}},
+\author[JMF]{\inits{JM.}\fnms{José M.}
+    \snm{Fernández}\ead[label=eJMF]{jose.m.fernandez@bsc.es}\ead[label=oJMF,orcid]{0000-0002-4806-5140}},
+\author[DG]{\inits{D.}\fnms{Daniel}
+    \snm{Garijo}\ead[label=eDG]{dgarijo@fi.upm.es}\ead[label=oDG,orcid]{0000-0003-0454-7145}},
+\author[BG]{\inits{B.}\fnms{Björn}
+    \snm{Grüning}\ead[label=eBG]{gruening@informatik.uni-freiburg.de}\ead[label=oBG,orcid]{0000-0002-3079-6586}},
+\author[MLR]{\inits{M.}\fnms{Marco}
+    \snm{La Rosa}\ead[label=eMLR]{m@lr.id.au}\ead[label=oMLR,orcid]{0000-0001-5383-6993}},
+\author[SL]{\inits{S.}\fnms{Simone}
+    \snm{Leo}\ead[label=eSL]{simone.leo@crs4.it}\ead[label=oSL,orcid]{0000-0001-8271-5429}},
+\author[EOC]{\inits{E.}\fnms{Eoghan}
+    \snm{Ó Carragáin}\ead[label=eEOC]{eoghan.ocarragain@ucc.ie}\ead[label=oEOC,orcid]{0000-0001-8131-2150}},
+\author[MP]{\inits{M.}\fnms{Marc}
+    \snm{Portier}\ead[label=eMP]{marc.portier@vliz.be}\ead[label=oMP,orcid]{0000-0002-9648-6484}},
+\author[AT]{\inits{A.}\fnms{Ana}
+    \snm{Trisovic}\ead[label=eMP]{ana.trisovic@gmail.com}\ead[label=oAT,orcid]{0000-0003-1991-0533}},
+\author[RO]{RO-Crate Community},
+\author[PG]{\inits{P.}\fnms{Paul}
+    \snm{Groth}\ead[label=ePG]{p.t.groth@uva.nl}\ead[label=oPG,orcid]{0000-0003-0183-6910}},
+\author[CG]{\inits{C.}\fnms{Carole}
+    \snm{Goble}\ead[label=eCG]{carole.goble@manchester.ac.uk}\ead[label=oCG,orcid]{0000-0003-1219-2137}}
+
+\address[SSR1]{Department of Computer Science, 
+    \orgname{The University of Manchester}, \cny{UK}\printead[presep={\\}]{eSSR,oSSR}}
+\address[SSR2]{Informatics Institute, \orgname{University of Amsterdam},
     \cny{The Netherlands}}
-\address[C]{Faculty of Science, \orgname{University Technology Sydney},
-\cny{Australia}\printead[presep={\\}]{e2,o2}}
-\address[D]{Institute for Quantitative Social Science, \orgname{Harvard
-University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{o3}}
-\address[E]{\orgname{ZB MED Information Centre for Life Sciences},
-Cologne, \cny{Germany}\printead[presep={\\}]{o4}}
-\address[F]{\orgname{VIB-UGent Center for Plant Systems Biology}, Gent,
-\cny{Belgium}\printead[presep={\\}]{o5}}
-\address[G]{\orgname{Barcelona Supercomputing Center}, Barcelona, \cny
-{Spain}\printead[presep={\\}]{o6}}
-\address[H]{Ontology Engineering Group, \orgname{Universidad Polit\'
-{e}cnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{o7}}
-\address[I]{Bioinformatics Group, Department of Computer Science,
-\orgname{Albert-Ludwigs-University Freiburg}, Freiburg, \cny
-{Germany}\printead[presep={\\}]{o8}}
-\address[J]{\orgname{PARADISEC}, Melbourne, \cny{Australia}\printead
-[presep={\\}]{o9}}
-\address[K]{\orgname{Center for Advanced Studies, Research, and
-Development in Sardinia} (CRS4), Pula (CA), \cny{Italy}\printead
-[presep={\\}]{o10}}
-\address[L]{\orgname{University College Cork}, \cny{Ireland}\printead
-[presep={\\}]{o11}}
-\address[M]{\orgname{Vlaams Instituut voor de Zee}, Oostende, \cny
-{Belgium}\printead[presep={\\}]{o12}}
-\address[N]{Institute for Quantitative Social Science, \orgname{Harvard
-University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{o13}}
-\address[O]{\printead{u1}~(see Appendix B)}
-\address[P]{Informatics Institute, \orgname{University of Amsterdam},
-\cny{The Netherlands}\printead[presep={\\}]{e15,o15}}
-\address[Q]{Department of Computer Science, \orgname{The University of
-Manchester}, \cny{UK}\printead[presep={\\}]{e16,o16}}
-\end{aug}
-%
-%%%%%%
-%
-\begin{review}{editor}
-\reviewer{\fnms{Silvio} \snm{Peroni}\ead
-[label=oe1,orcid]{0000-0003-0530-4305}\printead[presep={\ }]{oe1}}
-%\reviewer{\fnms{} \snm{}\ead[label=oe2,orcid]{}\printead[presep={\
-%}]{oe2}}
-\end{review}
-
-\begin{review}{solicited}
-\reviewer{\fnms{Angelo A.} \snm{Salatino}\ead
-[label=os1,orcid]{0000-0002-4763-3943}\printead[presep={\ }]{os1}}
-\reviewer{\fnms{Alessia} \snm{Bardi}\ead
-[label=os2,orcid]{0000-0002-1112-1292}\printead[presep={\ }]{os2}}
-\reviewer{\snm{one anonymous reviewer}}
-\end{review}
-%
-%\dedicated{}
-%
-% HISTORY:
-\received{\sday{13} \smonth{8} \syear{2021}}
-\accepted{\sday{23} \smonth{9} \syear{2021}}
-%
+\address[PS]{Faculty of Science, \orgname{University Technology Sydney},
+    \cny{Australia}\printead[presep={\\}]{ePS,oPS}}
+\address[MC]{Institute for Quantitative Social Science, 
+    \orgname{Harvard University}, Cambridge, MA,  \cny{USA}\printead[presep={\\}]{oMC}}
+\address[LJC]{\orgname{ZB MED Information Centre for Life Sciences}, 
+    Cologne, \cny{Germany}\printead[presep={\\}]{oLJC}}
+\address[FC]{\orgname{VIB-UGent Center for Plant Systems Biology}, 
+    Gent, \cny{Belgium}\printead[presep={\\}]{oFC}}
+\address[JMF]{\orgname{Barcelona Supercomputing Center}, 
+    Barcelona, \cny{Spain}\printead[presep={\\}]{oJMF}}
+\address[DG]{Ontology Engineering Group, 
+    \orgname{Universidad Politécnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{oDG}}
+\address[BG]{Bioinformatics Group, Department of Computer Science,
+    \orgname{Albert-Ludwigs-University Freiburg}, Freiburg, \cny{Germany}\printead[presep={\\}]{oBG}}
+\address[MLR]{\orgname{PARADISEC}, Melbourne, \cny{Australia}\printead[presep={\\}]{oMLR}}
+\address[SL]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA),
+    \cny{Italy}\printead[presep={\\}]{oSL}}
+\address[EOC]{\orgname{University College Cork},
+    \cny{Ireland}\printead[presep={\\}]{oEOC}}
+\address[MP]{\orgname{Vlaams Instituut voor de Zee}, Oostende,
+    \cny{Belgium}\printead[presep={\\}]{oMP}}
+\address[AT]{Institute for Quantitative Social Science,
+    \orgname{Harvard University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{oAT}}
+\address[RO]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
+\address[PG]{Informatics Institute, \orgname{University of Amsterdam},
+    \cny{The Netherlands}\printead[presep={\\}]{ePG,oPG}}
+\address[CG]{Department of Computer Science, \orgname{The University of Manchester},
+    \cny{UK}\printead[presep={\\}]{eCG,oCG}}
+
+%\begin{review}{editor}
+%\reviewer{\fnms{First} \snm{Editor}\ead[label=oe1,orcid]{}\printead[presep={\ }]{oe1}}
+%\reviewer{\fnms{Second} \snm{Editor}\ead[label=oe2,orcid]{}\printead[presep={\ }]{oe2}}
+%\end{review}
+%\begin{review}{solicited}
+%\reviewer{\fnms{First} \snm{Solicited reviewer}\ead[label=os1,orcid]{}\printead[presep={\ }]{os1}}
+%\end{review}
+
 \begin{abstract}
-%\parttitle{}
-An increasing number of researchers support reproducibility by
-including pointers to and
-descriptions of datasets, software and methods in their publications.
-However, scientific
-articles may be ambiguous, incomplete and difficult to process by
-automated systems. In
-this paper we introduce RO-Crate, an open, community-driven, and
-lightweight approach
-to packaging research artefacts along with their metadata in a machine
-readable manner.
-RO-Crate is based on Schema.org annotations in JSON-LD, aiming to
-establish best
-practices to formally describe metadata in an accessible and practical
-way for their
-use in a wide variety of situations.
-
-An RO-Crate is a structured archive of all the items that contributed
-to a research
-outcome, including their identifiers, provenance, relations and
-annotations. As a general
-purpose packaging approach for data and their metadata, RO-Crate is
-used across multiple
-areas, including bioinformatics, digital humanities and regulatory
-sciences. By applying
-``just enough'' Linked Data standards, RO-Crate simplifies the process of
-making research outputs FAIR while also enhancing research reproducibility.
-
-An RO-Crate for this article\footnote{\url
-{https://w3id.org/ro/doi/10.5281/zenodo.5146227}}
-is archived at \url{https://doi.org/10.5281/zenodo.5146227}.
+\markdownInput{../content/01.abstract.md}
 \end{abstract}
-%
-\begin{keyword} %% 1 u.c.
+
+\begin{keyword}
 \kwd{Data publishing}
-\kwd{data packaging}
+\kwd{Data packaging}
 \kwd{FAIR}
 \kwd{Linked Data}
-\kwd{metadata}
-\kwd{reproducibility}
-\kwd{research object}
+\kwd{Metadata}
+\kwd{Reproducibility}
+\kwd{Research Object}
 \end{keyword}
-%
+
 \end{frontmatter}
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}%[t!]
-\includegraphics{210053f01}
+\includegraphics{content/images/ro-crate-overview.pdf}
 \caption{Conceptual overview of RO-Crate. A \emph{Persistent
@@
 \begin{figure}%[t!]
-\includegraphics{210053f02}
+\includegraphics[width=0.9\textwidth]{content/images/ro-crate-uml.pdf}
 \caption{Simplified UML class diagram of RO-Crate. The \emph
@@
 \begin{figure}%[t]
-\includegraphics{210053f03}
+\includegraphics{content/images/ro-crate-bco-sep-of-concerns.pdf}
 \caption{Separation of Concerns in BCO RO-Crate. BioCompute
@@
 \begin{figure}%[t!]
-\includegraphics{210053f04}
+\includegraphics{content/images/data-commons-ro-crate-figure-5.pdf}
 \caption{One aspect of Harvard Data Commons. Automatic
@@
 
+%\markdownInput{../content/70.acknowledgements.md}
+
+%%%% Appendix
+
+%\newpage
+%\appendix
+%%\markdownInput{../content/80.formal-definition.md}
+%\input{../content/80.formal-definition}
+
+%\newpage
+%\markdownInput{../content/89.community.md}
+
+%%%%%%%%%%% The bibliography starts:
+%\nocite{label}
+%\bibliographystyle{ios1}
+%\bibliography{ro-crate}
+
+% Explicit bibliography from proof
+%
 % structpyb loaded by imikolaityte, 2021-12-28 10:21:28
@@
 \end{thebibliography}
+
 \end{document}
-%

### 2022-01-08 latex/ro-crate.tex
Sufficient (but not all) packages from ro-crate-proof

@@
 % add. options: [seceqn,secthm,crcready]
-\documentclass[ds,crcready]{iosart2x}
+\documentclass[ds,v1.1.2,openaccess]{iosart2x}
 
@@
 % From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
-\usepackage{fontspec}
-\usepackage{amsmath}
-\usepackage{amssymb}
+%\usepackage{fontspec}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
 \usepackage{hyperref}
@@
 
-\usepackage{unicode-math}
-\usepackage{graphicx}
+%\usepackage{unicode-math}
+%\usepackage{graphicx}
 
@@
 \graphicspath{../content/images/}
-\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
-\def\markdownOptionOutputDir{../output}
-\def\markdownOptionHeaderAttributes{true}
+%\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
+%\def\markdownOptionOutputDir{../output}
+%\def\markdownOptionHeaderAttributes{true}
 
@@
 \usepackage{multirow}
+%\usepackage{tmultirow}
 
@@
 
+\volume{-1}
+\issue{-1}
 \pubyear{2022}
-\volume{0}
 \firstpage{1}
-\lastpage{1}
-\def\doi{10.3233/DS-210053}
+\lastpage{42}
+\aid{DS210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
+\doi{10.3233/DS-210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
+
+\startlocaldefs
+\setpkgattr{license}{type}{CC@BY} %% CC@BY@NC
+\endlocaldefs
 
 \begin{document}
-
-% Welcome to the 2000s with UNICODE fonts!
-% Adapted from
-% https://www.overleaf.com/learn/latex/Articles%2FOpenType-based_math_typesetting:_An_introduction_to_the_STIX2_OpenType_fonts
-\setmathfont{STIX2Math}[
-    Extension={.otf},
-    Path=./STIX2fonts/,
-    Scale=1
-]
-\setmainfont{STIX2Text}[
-    Extension={.otf},
-    Path=./STIX2fonts/,
-    UprightFont={*-Regular},
-    BoldFont={*-Bold},
-    ItalicFont={*-Italic},
-    BoldItalicFont={*-BoldItalic}
-]
-
-% https://typeof.net/Iosevka/ mono font has 
-% all the logic symbols we need
-%\setmonofont{iosevka-slab}[
-%    Extension={.ttf},
-%    Path=./ttf-iosevka-slab-6.1.3/,
-%    Scale=MatchLowercase,
-%    UprightFont={*-light},
-%    ItalicFont={*-lightitalic}
-%]
 
@@
 
-%\begin{review}{editor}
-%\reviewer{\fnms{First} \snm{Editor}\ead[label=oe1,orcid]{}\printead[presep={\ }]{oe1}}
-%\reviewer{\fnms{Second} \snm{Editor}\ead[label=oe2,orcid]{}\printead[presep={\ }]{oe2}}
-%\end{review}
-%\begin{review}{solicited}
-%\reviewer{\fnms{First} \snm{Solicited reviewer}\ead[label=os1,orcid]{}\printead[presep={\ }]{os1}}
-%\end{review}
-
+\begin{review}{editor}
+\reviewer{\fnms{Silvio} \snm{Peroni}\ead
+[label=oe1,orcid]{0000-0003-0530-4305}\printead[presep={\ }]{oe1}}
+%\reviewer{\fnms{} \snm{}\ead[label=oe2,orcid]{}\printead[presep={\
+%}]{oe2}}
+\end{review}
+
+\begin{review}{solicited}
+\reviewer{\fnms{Angelo A.} \snm{Salatino}\ead
+[label=os1,orcid]{0000-0002-4763-3943}\printead[presep={\ }]{os1}}
+\reviewer{\fnms{Alessia} \snm{Bardi}\ead
+[label=os2,orcid]{0000-0002-1112-1292}\printead[presep={\ }]{os2}}
+\reviewer{\snm{one anonymous reviewer}}
+\end{review}
+%
+%\dedicated{}
+%
+% HISTORY:
+\received{\sday{13} \smonth{8} \syear{2021}}
+\accepted{\sday{23} \smonth{9} \syear{2021}}
+%
 \begin{abstract}
-\markdownInput{../content/01.abstract.md}
+%\parttitle{}
+An increasing number of researchers support reproducibility by
+including pointers to and
+descriptions of datasets, software and methods in their publications.
+However, scientific
+articles may be ambiguous, incomplete and difficult to process by
+automated systems. In
+this paper we introduce RO-Crate, an open, community-driven, and
+lightweight approach
+to packaging research artefacts along with their metadata in a machine
+readable manner.
+RO-Crate is based on Schema.org annotations in JSON-LD, aiming to
+establish best
+practices to formally describe metadata in an accessible and practical
+way for their
+use in a wide variety of situations.
+
+An RO-Crate is a structured archive of all the items that contributed
+to a research
+outcome, including their identifiers, provenance, relations and
+annotations. As a general
+purpose packaging approach for data and their metadata, RO-Crate is
+used across multiple
+areas, including bioinformatics, digital humanities and regulatory
+sciences. By applying
+``just enough'' Linked Data standards, RO-Crate simplifies the process of
+making research outputs FAIR while also enhancing research reproducibility.
+
+An RO-Crate for this article\footnote{\url
+{https://w3id.org/ro/doi/10.5281/zenodo.5146227}}
+is archived at \url{https://doi.org/10.5281/zenodo.5146227}.
 \end{abstract}
-
-\begin{keyword}
+%
+\begin{keyword} %% 1 u.c.
 \kwd{Data publishing}
-\kwd{Data packaging}
+\kwd{data packaging}
 \kwd{FAIR}
 \kwd{Linked Data}
-\kwd{Metadata}
-\kwd{Reproducibility}
-\kwd{Research Object}
+\kwd{metadata}
+\kwd{reproducibility}
+\kwd{research object}
 \end{keyword}
-
+%
 \end{frontmatter}
@@
 \hline
-\textbf{Tool name} & \multicolumn{1}{c}{\tmultirow{2}{*}{Targets}}
-& \tmultirow{2}{*}{Language/Platform} & \tmultirow{2}{*}{Status} \\
+\textbf{Tool name} & \multicolumn{1}{c}{\multirow{2}{*}{Targets}}
+& \multirow{2}{*}{Language/Platform} & \multirow{2}{*}{Status} \\
 \textit{Brief Description}&&& \\
@@
 fully represented as a tree at depth 2 helped by the use of
-%PDAG%
-$\mathbh{IRI}$
-%PDAG%
+$\mathbb{IRI}$
 nodes. For instance the aggregation from the root entity
@@
 ISBN \bisbn{978-3-030-45442-5}.
-\bid{doi={10.1007/978-3-030-45442-5_15}}
+\bid{doi={10.1007/978-3-030-45442-5\_15}}
 \end{bchapter}
@@
 \oauthor{\binits{P.}~\bsnm{Sefton}},
-{FAIR} Data Management; It's a lifestyle not a lifecycle -- \surl{ptsefton.com},
+{FAIR} Data Management; It's a lifestyle not a lifecycle -- ptsefton.com,
 2021.
